### 2018-09-26 Cytokine_Analysis.Rmd
First Commit

---
title: "Cytokine_Analysis"
author: "Taryn McLaughlin"
date: "9/21/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 10.5, fig.width = 7.5)

library(dplyr)
library(reshape2)
Cytokine<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/Background Subtracted Cytokine.csv")
Cytokine <- mutate(Cytokine, G_plus = G_4_13_x + G_4_x_x + G_x_13_x + G_x_x_x) %>%
    mutate(T_plus = x_4_13_T + x_4_x_T + x_x_13_T + x_x_x_T) %>%
    mutate(G_T_plus = G_4_x_T + G_x_13_T + G_x_x_T)%>%
    mutate(IL4_plus = G_4_x_T + G_4_x_x + x_4_x_T + x_4_x_x) %>%
    mutate(IL13_plus = G_x_13_T + G_x_13_x + x_x_13_T + x_x_13_x) %>%
    mutate(IL4_IL13_plus = G_4_13_x + x_4_13_T + x_4_13_x)
Cytokine$TB<-factor(Cytokine$TB, levels=c("N","HC","LTBI","TB"))
Cytokine$SM<-factor(Cytokine$SM, levels=c("N","X","SM"))
```

#Statistical Analysis
I wrote a function called pvals which evaluates the relevant two sample two-tailed  p-values using the non-parametric Mann Whitney test (the command is wilcox.test in R). The function then corrects for multiple comparisons in two ways. The first uses  a Bonferonni correction to control the Family Wise Error Rate. The second controls the False Discovery Rate, the expected proportion of false discoveries amongst the rejected hypotheses. The false discovery rate is a less stringent condition than the family-wise error rate, so these methods are more powerful than the others. The function is below if you would like to see the details.

```{r p-value function, echo=TRUE}
pvals<-function(data, column){

    data<-filter(data, SM!="N")
    data$SM<-factor(data$SM, levels=c("X","SM"))
    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))

    x<-split(data,data$TB) 
    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"]))

    y<-split(data,data$SM) 
    B<-lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "LTBI")))
    C<- lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("LTBI", "TB")))
    D<- lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "TB")))

    labels<-c("HC SM+ to HC SM-", "LTBI SM+ to LTBI SM-", "TB SM+ to TB SM-",
              "HC SM+ to LTBI SM+", "LTBI SM+ to TB SM+", "HC SM+ to TB SM+",
              "HC SM- to LTBI SM-", "LTBI SM- to TB SM-", "HC SM- to TB SM-"
    )
    pvals<-c(A$HC$p.value, A$LTBI$p.value, A$TB$p.value,
             B$SM$p.value, C$SM$p.value,  D$SM$p.value,
             B$X$p.value,  C$X$p.value,   D$X$p.value)

    bonf_adj_pvals<-round(p.adjust(pvals, method="bonferroni"), 4)

    fdr_adj_pvals<-round(p.adjust(pvals, method="fdr"), 4)

    table<-(cbind(labels, round(pvals, 4), bonf_adj_pvals, fdr_adj_pvals))
    colnames(table)<-c("Comparison", paste(column,"p-values"," "), 
                       "Bonferonni Corrected p-values", "FDR Corrected p-vales")
    table
}
```

#Boolean Graphs
The first graph is just meant to give a snapshot of all the data at once to see if there are weird anomalous data points that we will have to examine. I did not include all the individual data points on any of the graphs because it made it incredibly cluttered looking, however I did allow outliers to display.

```{r Boolean Graphs, cache=TRUE, echo=FALSE, warning=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
Boolean<-select(Cytokine, Donor, TB, SM, Stim, G_4_13_T, G_4_x_T, G_x_13_T, G_4_13_x, x_4_13_T, G_4_x_x, G_x_13_x,
                x_4_x_T, x_x_13_T, G_x_x_T , G_x_x_x, x_x_x_T,  x_4_13_x, x_4_x_x, x_x_13_x)

melt<-melt(Boolean,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", 
                                           "G_x_13_x","x_4_x_T","x_x_13_T", "G_x_x_T" , "G_x_x_x","x_x_x_T",  
                                           "x_4_13_x","x_4_x_x", "x_x_13_x"))
melt<-filter(melt, !is.na(label))

#all the boolean plots
g<-ggplot(melt, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(Stim~TB, scale="free")+
    labs(title="Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

#subset by stim
data<-filter(melt, Stim=="PMA")
g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(TB~., scale="free")+
    labs(title="PMA Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)");

data<-filter(melt, Stim=="WCL")
g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(TB~., scale="free")+
    labs(title="WCL Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)");data<-filter(melt, Stim=="PEP")

data<-filter(melt, Stim=="PEP")
g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(TB~., scale="free")+
    labs(title="PEP Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)");

data<-filter(melt, Stim=="SEA")
g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(TB~., scale="free")+
    labs(title="SEA Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)");

data<-filter(melt, Stim=="SWAP")
g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(TB~., scale="free")+
    labs(title="SWAP Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)");
```

#T cell Lineage Specific Grouped Analysis
These are my standard lineage specific groups.

```{r TH, echo=FALSE, fig.cap="Breakdown of Cytokine Definitions"}
knitr::include_graphics("/Applications/Old Computer/Day Lab/Data_Flow/TH_Group_Explanation.png")
```

```{r Grouped TH Graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
TH<-select(Cytokine, Donor, TB, SM, Stim, Cyt.Freq, TH1.Freq, TH1.2.Freq, TH2.Freq)

melt<-melt(TH,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("Cyt.Freq", "TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
melt<-filter(melt, !is.na(label))

##### PMA
data<-filter(melt, Stim=="PMA")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(TH, Stim=="PMA")
kable(pvals(data, "TH1.Freq"), caption = "Non Specific CD4+ T cell Responses")
kable(pvals(data, "TH2.Freq"))
kable(pvals(data, "TH1.2.Freq"))

##### WCL
data<-filter(melt, Stim=="WCL")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(TH, Stim=="WCL")
kable(pvals(data, "TH1.Freq"), caption = "Mtb Whole Cell Lysate Specific CD4 T cell Responses")
kable(pvals(data, "TH2.Freq"))
kable(pvals(data, "TH1.2.Freq"))

##### Peptide
data<-filter(melt, Stim=="PEP")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(TH, Stim=="PEP")
kable(pvals(data, "TH1.Freq"), caption = "Mtb Peptide Specific CD4 T cell Responses")
kable(pvals(data, "TH2.Freq"))
kable(pvals(data, "TH1.2.Freq"))

```
\newpage

#Individual Cytokines
These are the frequencies of cells producing individual cytokines without considering whether or not they are single positive, double positive etc. So any cell producing IL-4 is included in the IL-4 analysis.

```{r Individual Cytokines but not single producers}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
cyt<-select(Cytokine, Donor, TB, SM, Stim, IFNg, TNFa, IL4, IL13)

melt<-melt(cyt,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("IFNg", "TNFa", "IL4", "IL13"))
melt<-filter(melt, !is.na(label))

#### PMA
data<-filter(melt, Stim=="PMA")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(cyt, Stim=="PMA")
kable(pvals(data, "IFNg"), caption = "Non Specific CD4 T cell Responses")
kable(pvals(data, "TNFa"))
kable(pvals(data, "IL4"))
kable(pvals(data, "IL13"))

#### WCL
data<-filter(melt, Stim=="WCL")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(cyt, Stim=="WCL")
kable(pvals(data, "IFNg"), caption = "Mtb Whole Cell Lysate Specific CD4 T cell Responses")
kable(pvals(data, "TNFa"))
kable(pvals(data, "IL4"))
kable(pvals(data, "IL13"))

#### Peptide
data<-filter(melt, Stim=="PEP")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(cyt, Stim=="PEP")
kable(pvals(data, "IFNg"), caption = "Mtb Peptide Specific CD4 T cell Responses")
kable(pvals(data, "TNFa"))
kable(pvals(data, "IL4"))
kable(pvals(data, "IL13"))

```

#True Single and Double Cytokine Positive CD4+ T Cells
These are the frequencies of cells truly ONLY producing:  
1. IFNg or TNFa or the combination of both AND NOT IL4 or IL13  
2. IL-4 or IL-13 or the combination of both AND NOT IFNg or TNFa  
These are the subgroups within the grouped TH1 vs TH2 lineage analysis. They are essentially the last 6 groups of the boolean data.

```{r Single and Double Producers Lin Spec graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
SP_lin<-rename(Cytokine, G_single =  G_x_x_x, T_single = x_x_x_T, IL4_single = x_4_x_x, IL13_single = x_x_13_x, 
               G_T = G_x_x_T, IL4_IL13 = x_4_13_x) %>%
    select(Donor, TB, SM, Stim, G_single, T_single, IL4_single, IL13_single, G_T, IL4_IL13)

melt<-melt(SP_lin,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("G_single", "T_single", "G_T", "IL4_single", "IL13_single", "IL4_IL13"))
melt<-filter(melt, !is.na(label))

#### PMA
# Th1 plots
data<-filter(melt, Stim=="PMA" & label %in% c("G_single", "T_single", "G_T"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non-Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="PMA" & label %in% c("IL4_single", "IL13_single", "IL4_IL13"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non-Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_lin, Stim=="PMA")
kable(pvals(data, "G_single"), caption = "Non-Specific CD4+ T cell Responses")
kable(pvals(data, "T_single")) 
kable(pvals(data,"G_T"))
kable(pvals(data, "IL4_single"))
kable(pvals(data, "IL13_single")) 
kable(pvals(data, "IL4_IL13"))

#### WCL
# Th1 plots
data<-filter(melt, Stim=="WCL" & label %in% c("G_single", "T_single", "G_T"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="WCL" & label %in% c("IL4_single", "IL13_single", "IL4_IL13"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_lin, Stim=="WCL")
kable(pvals(data, "G_single"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
kable(pvals(data, "T_single")) 
kable(pvals(data,"G_T"))
kable(pvals(data, "IL4_single"))
kable(pvals(data, "IL13_single")) 
kable(pvals(data, "IL4_IL13"))

#### PEP
# Th1 plots
data<-filter(melt, Stim=="PEP" & label %in% c("G_single", "T_single", "G_T"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="PEP" & label %in% c("IL4_single", "IL13_single", "IL4_IL13"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_lin, Stim=="PEP")
kable(pvals(data, "G_single"), caption = "Mtb Peptide Specific CD4 T cell Responses")
kable(pvals(data, "T_single")) 
kable(pvals(data,"G_T"))
kable(pvals(data, "IL4_single"))
kable(pvals(data, "IL13_single")) 
kable(pvals(data, "IL4_IL13"))

```

\newpage

#Lineage "Agnostic" Single and Double Cytokine Positive CD4+ T Cells
I don't really believe that this is an appropriate way to look at the data because the cells are technically shared between some of these groupings, but I'm almost certain someone would ask the question "What if you didn't consider Il-4 and IL-13 in defining Gamma and TNF single and double positive cells" so I did it for them.

```{r TH1, echo=FALSE, fig.cap="Breakdown of TH1 Cytokine Definitions"}
knitr::include_graphics("/Applications/Old Computer/Day Lab/Data_Flow/TH1 plus explanations.png")
```

```{r TH2, echo=FALSE, fig.cap="Breakdown of TH2 Cytokine Definitions"}
knitr::include_graphics("/Applications/Old Computer/Day Lab/Data_Flow/TH2 plus explanations.png")
```

```{r Single and Double Producers Lin Ag graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)

SP_Ag<-select(Cytokine, Donor, TB, SM, Stim, G_plus, T_plus, IL4_plus, IL13_plus, G_T_plus, IL4_IL13_plus)
melt<-melt(SP_Ag,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("G_plus", "T_plus", "G_T_plus", "IL4_plus", "IL13_plus", "IL4_IL13_plus"))
melt<-filter(melt, !is.na(label))

### PMA
#TH1 plots
data<-filter(melt, Stim=="PMA" & label %in% c("G_plus", "T_plus", "G_T_plus"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="PMA" & label %in% c("IL4_plus", "IL13_plus", "IL4_IL13_plus"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_Ag, Stim=="PMA")
kable(pvals(data, "G_plus"), caption = "Non Specific CD4+ T cell Responses")
kable(pvals(data, "T_plus"))
kable(pvals(data, "G_T_plus"))
kable(pvals(data, "IL4_plus"))
kable(pvals(data, "IL13_plus"))
kable(pvals(data, "IL4_IL13_plus"))

### WCL
#TH1 plots
data<-filter(melt, Stim=="WCL" & label %in% c("G_plus", "T_plus", "G_T_plus"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="WCL" & label %in% c("IL4_plus", "IL13_plus", "IL4_IL13_plus"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_Ag, Stim=="WCL")
kable(pvals(data, "G_plus"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
kable(pvals(data, "T_plus"))
kable(pvals(data, "G_T_plus"))
kable(pvals(data, "IL4_plus"))
kable(pvals(data, "IL13_plus"))
kable(pvals(data, "IL4_IL13_plus"))

### Peptide
#TH1 plots
data<-filter(melt, Stim=="PEP" & label %in% c("G_plus", "T_plus", "G_T_plus"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ T cell Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="PEP" & label %in% c("IL4_plus", "IL13_plus", "IL4_IL13_plus"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ T cell Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_Ag, Stim=="PEP")
kable(pvals(data, "G_plus"), caption = "Mtb Peptide Specific CD4+ T cell Responses")
kable(pvals(data, "T_plus"))
kable(pvals(data, "G_T_plus"))
kable(pvals(data, "IL4_plus"))
kable(pvals(data, "IL13_plus"))
kable(pvals(data, "IL4_IL13_plus"))
```

### 2018-10-21 CD4 Cleaning.Rmd
Add files via upload

---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---
Load and rough clean all data sets
```{r}
setwd("/Applications/Old Computer/Day Lab/Data_Flow")
CD4<-read.csv("CD4.csv")
CD4<-filter(CD4,!(Sample=="Mean"))%>%
filter(!Sample=="StdDev")%>%
filter(!Sample=="Sample")
CD4[CD4=="¥"]<-NA

Cyt<-read.csv("CD4_Cytokine.csv")
Cyt<-filter(Cyt,!(Sample=="Mean"))%>%
filter(!Sample=="StdDev")%>%
filter(!Sample=="Sample")
Cyt[Cyt=="¥"]<-NA

AG<-read.table("/Applications/Old Computer/Day Lab/Data_Flow/CD4_Positive.csv", header= TRUE, sep = ",")
AG<-filter(AG,!(Sample=="Mean"))%>%
filter(!Sample=="StdDev")%>%
filter(!Sample=="Sample")
AG[AG=="¥"]<-NA
AG<-filter(AG,!(Sample=="Mean")) %>%
  filter(!Sample=="StdDev") %>%
  filter(!Sample=="Sample") %>%
  filter(!Sample=="NK2126_HC_X_SEA_005.fcs") %>%
  filter(!Sample=="NK2450_HC_SM_SEA_047.fcs") %>%
  filter(!Sample=="NK2402_LTBI_SM_SWAP_024.fcs") %>%
  filter(!Sample=="NK2342_TB_X_SWAP_006.fcs") %>%
  filter(!Sample=="NK2136_TB_X_WCL_003.fcs") %>%
  filter(!Sample=="NK2168_HC_SM_WCL_040.fcs") %>%
  filter(!Sample=="NK2421_LTBI_SM_PEP_010.fcs") %>%
  filter(!Sample=="NK2063_TB_X_SEA_005.fcs") %>%
  filter(!Sample=="NK2115_LTBI_X_PEP_010.fcs") %>%
  filter(!Sample=="NK2115_LTBI_X_WCL_009.fcs") %>%
  filter(!Sample=="NK2171_HC_X_WCL_009.fcs") %>%
  filter(!Sample=="NK2162_HC_X_SEA_023.fcs")

CYT<-AG[,c(1:18)]
TH1<-AG[,c(1,19:35)]
TH2<-AG[,c(1,53:69)]
TH1.2<-AG[,c(1,36:52)]

write.csv(CD4,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/CD4.csv")
write.csv(Cyt,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/Cytokine.csv")
write.csv(CYT,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/Cyt.csv")
write.csv(TH1,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH1.csv")
write.csv(TH2,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH2.csv")
write.csv(TH1.2,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH1-2.csv")

remove(AG, CD4, Cyt, CYT, TH1, TH2, TH1.2)
```

PBMC
```{r}
PBMC<-read.csv("/Applications/Old computer/Day Lab/Data_Flow/PBMC.csv")
PBMC<-filter(PBMC,!(Helminth.Status=="Ascaris+"))%>%
        filter(!Helminth.Status=="Ascarus+ & Trichuris+")%>%
        filter(!(Helminth.Status=="Hookworm+"))%>%
        filter(!(Helminth.Status=="Trichuris+"))
PBMC$TB.Status = factor(PBMC$TB.Status,levels=c("Healthy Control", "LTBI", "Active TB"))
```

cleaning total CD4 data
```{r}
## Names Labels Etc. for X and SM
#remember to check for _ issues
setwd("/Applications/Old Computer/Day Lab/Data_Flow/Clean")
CD4<-read.csv("CD4.csv")

CD4<-filter(CD4,!(Sample=="Mean")) %>%
  filter(!Sample=="StdDev") %>%
  filter(!Sample=="Sample") %>%
  filter(!Sample=="NK2126_HC_X_SEA_005.fcs") %>%
  filter(!Sample=="NK2450_HC_SM_SEA_047.fcs") %>%
  filter(!Sample=="NK2402_LTBI_SM_SWAP_024.fcs") %>%
  filter(!Sample=="NK2342_TB_X_SWAP_006.fcs") %>%
  filter(!Sample=="NK2136_TB_X_WCL_003.fcs") %>%
  filter(!Sample=="NK2168_HC_SM_WCL_040.fcs") %>%
  filter(!Sample=="NK2421_LTBI_SM_PEP_010.fcs") %>%
  filter(!Sample=="NK2063_TB_X_SEA_005.fcs") %>%
  filter(!Sample=="NK2115_LTBI_X_PEP_010.fcs") %>%
  filter(!Sample=="NK2115_LTBI_X_WCL_009.fcs") %>%
  filter(!Sample=="NK2171_HC_X_WCL_009.fcs") %>%
  filter(!Sample=="NK2162_HC_X_SEA_023.fcs")

Split<-strsplit(as.character(CD4$Sample),"_");
CD4$Donor = as.factor(sapply(Split,"[",1));
CD4$TB = as.factor(sapply(Split,"[",2));
CD4$SM = as.factor(sapply(Split,"[",3));
CD4$Antigen = as.factor(sapply(Split,"[",4));

CD4<-as.data.table(CD4)
CD4<-CD4[CD4$SM!="HEL"]

CD4$SM<-factor(CD4$SM, levels=c("N","X","SM"))
CD4$TB<-factor(CD4$TB, levels=c("N","HC", "LTBI", "TB"))
CD4$Stim = factor(CD4$Antigen,levels=c("UN", "WCL","PEP","SEA", "SWAP", "PMA"))
CD4<-CD4[!(is.na(CD4$Sample))]

write.csv(CD4,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/CD4.csv")
```
cleaning cytokine frequency data
```{r}
Cyt<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/Cytokine.csv")

Cyt<-filter(Cyt,!(Sample=="Mean")) %>%
  filter(!Sample=="StdDev") %>%
  filter(!Sample=="Sample") %>%
  filter(!Sample=="NK2126_HC_X_SEA_005.fcs") %>%
  filter(!Sample=="NK2450_HC_SM_SEA_047.fcs") %>%
  filter(!Sample=="NK2402_LTBI_SM_SWAP_024.fcs") %>%
  filter(!Sample=="NK2342_TB_X_SWAP_006.fcs") %>%
  filter(!Sample=="NK2136_TB_X_WCL_003.fcs") %>%
  filter(!Sample=="NK2168_HC_SM_WCL_040.fcs") %>%
  filter(!Sample=="NK2421_LTBI_SM_PEP_010.fcs") %>%
  filter(!Sample=="NK2063_TB_X_SEA_005.fcs") %>%
  filter(!Sample=="NK2115_LTBI_X_PEP_010.fcs") %>%
  filter(!Sample=="NK2115_LTBI_X_WCL_009.fcs") %>%
  filter(!Sample=="NK2171_HC_X_WCL_009.fcs") %>%
  filter(!Sample=="NK2162_HC_X_SEA_023.fcs")

Split<-strsplit(as.character(Cyt$Sample),"_");
Cyt$Donor = as.factor(sapply(Split,"[",1));
Cyt$TB = as.factor(sapply(Split,"[",2));
Cyt$SM = as.factor(sapply(Split,"[",3))
Cyt$Antigen = as.factor(sapply(Split,"[",4));

Cyt<-as.data.table(Cyt)
Cyt<-Cyt[Cyt$SM!="HEL"]

Cyt$SM<-factor(Cyt$SM, levels=c("N","X","SM"))
Cyt$TB<-factor(Cyt$TB, levels=c("N","HC", "LTBI", "TB"))
Cyt$Stim = factor(Cyt$Antigen,levels=c("UN", "WCL","PEP","SEA", "SWAP", "PMA"))

Cyt<-Cyt[!(is.na(Cyt$Sample))]
remove(Split)
write.csv(Cyt,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/Cytokine.csv")

```

Determining antigen specific threshold vlaues
```{r}
UN<-filter(Cyt,Stim=="UN")
ag<-filter(Cyt,Stim!="UN")
ag[ag<0.03]<-NA
Dat<-merge(ag,UN,by="Donor")
Dat<-as.data.table(Dat)

Dat$Sample<-Dat$Sample.x
Dat$Stim<-Dat$Stim.x
Dat$TB<-Dat$TB.x
Dat$SM<-Dat$SM.x
Dat$G_4_13_T<-(Dat$G_4_13_T.x-3*Dat$G_4_13_T.y)
Dat$G_4_13_x<-(Dat$G_4_13_x.x-3*Dat$G_4_13_x.y)
Dat$G_4_x_T<-(Dat$G_4_x_T.x-3*Dat$G_4_x_T.y)
Dat$G_4_x_x<-(Dat$G_4_x_x.x-3*Dat$G_4_x_x.y)
Dat$G_x_13_T<-(Dat$G_x_13_T.x-3*Dat$G_x_13_T.y)
Dat$G_x_13_x<-(Dat$G_x_13_x.x-3*Dat$G_x_13_x.y)
Dat$G_x_x_T<-(Dat$G_x_x_T.x-3*Dat$G_x_x_T.y)
Dat$G_x_x_x<-(Dat$G_x_x_x.x-3*Dat$G_x_x_x.y)
Dat$x_4_13_T<-(Dat$x_4_13_T.x-3*Dat$x_4_13_T.y)
Dat$x_4_13_x<-(Dat$x_4_13_x.x-3*Dat$x_4_13_x.y)
Dat$x_4_x_T<-(Dat$x_4_x_T.x-3*Dat$x_4_x_T.y)
Dat$x_4_x_x<-(Dat$x_4_x_x.x-3*Dat$x_4_x_x.y)
Dat$x_x_13_T<-(Dat$x_x_13_T.x-3*Dat$x_x_13_T.y)
Dat$x_x_13_x<-(Dat$x_x_13_x.x-3*Dat$x_x_13_x.y)
Dat$x_x_x_T<-(Dat$x_x_x_T.x-3*Dat$x_x_x_T.y)
Dat$x_x_x_x<-Dat$x_x_x_x.x
Dat$IFNg<-(Dat$IFNg.x-3*Dat$IFNg.y)
Dat$IL13<-(Dat$IL13.x-3*Dat$IL13.y)
Dat$IL4<-(Dat$IL4.x-3*Dat$IL4.y)
Dat$TNFa<-(Dat$TNFa.x-3*Dat$TNFa.y)
Dat$Cyt.Count<-(Dat$Cyt.Count.x-3*Dat$Cyt.Count.y)
Dat$Cyt.Freq<-(Dat$Cyt.Freq.x-3*Dat$Cyt.Freq.y)
Dat$TH1.Count<-(Dat$TH1.Count.x-3*Dat$TH1.Count.y)
Dat$TH1.Freq<-(Dat$Th1.Freq.x-3*Dat$Th1.Freq.y)
Dat$TH1.2.Count<-(Dat$TH1.2.Count.x-3*Dat$TH1.2.Count.y)
Dat$TH1.2.Freq<-(Dat$TH1.2.Freq.x-3*Dat$TH1.2.Freq.y)
Dat$TH2.Count<-(Dat$TH2.Count.x-3*Dat$TH2.Count.y)
Dat$TH2.Freq<-(Dat$TH2.Freq.x-3*Dat$TH2.Freq.y)

AG<-Dat[,c(1,70:88,90:93,95,97,99,101)]

AG[AG <0 ]<-as.numeric(0)
AG[AG >0 ]<-as.numeric(1)

AG$Sum<-rowSums(AG[,6:28], na.rm=TRUE)
AG$Cyt.Sum<-rowSums(AG[,21:24], na.rm = TRUE)
AG$TH1.Sum<-rowSums(AG[,c(12,13,20,21,24,26)], na.rm = TRUE)
AG$TH2.Sum<-rowSums(AG[,c(15,17,19,22,23,28)], na.rm = TRUE)
AG$TH1.2.Sum<-rowSums(AG[,c(6:11,14,16,18,27)], na.rm = TRUE)

x <- split(AG, AG$Stim)

c<-lapply(x, function(g) colSums(g[,6:28], na.rm = TRUE))
c2<-as.data.table(c)
c3<-cbind(names(AG))
cyt.tally<-cbind(c3[6:28],c2)

r<-lapply(x, function(g) rowSums(g[,6:28], na.rm = TRUE))
r2<-as.data.table(r)
r3<-unique(cbind(as.character(AG$Donor)))
donor.tally<-cbind(r3,r2)

remove(UN, ag, Dat, c, c2, c3, r, r2, r3, x)
write.csv(cyt.tally,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/cyt-tally.csv")
write.csv(donor.tally,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/donor-tally.csv")
write.csv(AG,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/threshold-data.csv")

```

Background subtracting cytokine frequencies
```{r}
UN<-filter(Cyt,Stim=="UN")
ag<-filter(Cyt,Stim!="UN")
Dat<-merge(ag,UN,by="Donor")
Dat<-as.data.table(Dat)

Dat$Sample<-Dat$Sample.x
Dat$Stim<-Dat$Stim.x
Dat$TB<-Dat$TB.x
Dat$SM<-Dat$SM.x
Dat$G_4_13_T<-(Dat$G_4_13_T.x-Dat$G_4_13_T.y)
Dat$G_4_13_x<-(Dat$G_4_13_x.x-Dat$G_4_13_x.y)
Dat$G_4_x_T<-(Dat$G_4_x_T.x-Dat$G_4_x_T.y)
Dat$G_4_x_x<-(Dat$G_4_x_x.x-Dat$G_4_x_x.y)
Dat$G_x_13_T<-(Dat$G_x_13_T.x-Dat$G_x_13_T.y)
Dat$G_x_13_x<-(Dat$G_x_13_x.x-Dat$G_x_13_x.y)
Dat$G_x_x_T<-(Dat$G_x_x_T.x-Dat$G_x_x_T.y)
Dat$G_x_x_x<-(Dat$G_x_x_x.x-Dat$G_x_x_x.y)
Dat$x_4_13_T<-(Dat$x_4_13_T.x-Dat$x_4_13_T.y)
Dat$x_4_13_x<-(Dat$x_4_13_x.x-Dat$x_4_13_x.y)
Dat$x_4_x_T<-(Dat$x_4_x_T.x-Dat$x_4_x_T.y)
Dat$x_4_x_x<-(Dat$x_4_x_x.x-Dat$x_4_x_x.y)
Dat$x_x_13_T<-(Dat$x_x_13_T.x-Dat$x_x_13_T.y)
Dat$x_x_13_x<-(Dat$x_x_13_x.x-Dat$x_x_13_x.y)
Dat$x_x_x_T<-(Dat$x_x_x_T.x-Dat$x_x_x_T.y)
Dat$x_x_x_x<-Dat$x_x_x_x.x
Dat$IFNg<-(Dat$IFNg.x-Dat$IFNg.y)
Dat$IL13<-(Dat$IL13.x-Dat$IL13.y)
Dat$IL4<-(Dat$IL4.x-Dat$IL4.y)
Dat$TNFa<-(Dat$TNFa.x-Dat$TNFa.y)
Dat$Cyt.Count<-(Dat$Cyt.Count.x-Dat$Cyt.Count.y)
Dat$Cyt.Freq<-(Dat$Cyt.Freq.x-Dat$Cyt.Freq.y)
Dat$TH1.Count<-(Dat$TH1.Count.x-Dat$TH1.Count.y)
Dat$TH1.Freq<-(Dat$Th1.Freq.x-Dat$Th1.Freq.y)
Dat$TH1.2.Count<-(Dat$TH1.2.Count.x-Dat$TH1.2.Count.y)
Dat$TH1.2.Freq<-(Dat$TH1.2.Freq.x-Dat$TH1.2.Freq.y)
Dat$TH2.Count<-(Dat$TH2.Count.x-Dat$TH2.Count.y)
Dat$TH2.Freq<-(Dat$TH2.Freq.x-Dat$TH2.Freq.y)

CYT<-Dat[,c(1,70:101)]
CYT[CYT<0]<-0
remove(UN, ag, Dat, Cyt)

write.csv(CYT,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/Background Subtracted Cytokine.csv")
```

#Antigen specific responses using threshold values
```{r}
Cyt<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/Cyt.csv")
TH1<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH1.csv")
TH2<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH2.csv")
TH1.2<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH1-2.csv")

TH1<-merge(AG,TH1,by="Sample", all=TRUE)
TH2<-merge(AG,TH2,by="Sample", all=TRUE)
TH1.2<-merge(AG,TH1.2,by="Sample", all=TRUE)
Cyt<-merge(AG,Cyt,by="Sample", all=TRUE)

Cyt<-Cyt[,c(1,3:5,21:25,29,30,35:51)]
TH1<-TH1[,c(1:5,21,24,26,31,35:51)]
TH2<-TH2[,c(1:5,22:23,28,32,35:51)]
TH1.2<-TH1.2[,c(1:5,21:24,27,33,35:51)]

write.csv(Cyt,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/Cyt_pos.csv")
write.csv(TH1,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH1_pos.csv")
write.csv(TH2,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH2_pos.csv")
write.csv(TH1.2,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH1-2_pos.csv")
```

Load and rough clean all counts sets
```{r}
setwd("/Applications/Old Computer/Day Lab/Data_Flow")
CD4_count<-read.csv("CD4_counts.csv")
CD4_count<-filter(CD4_count,!(Sample=="Mean"))%>%
    filter(!Sample=="StdDev")%>%
    filter(!Sample=="Sample")%>%
    filter(!Sample=="NK2126_HC_X_SEA_005.fcs") %>%
    filter(!Sample=="NK2450_HC_SM_SEA_047.fcs") %>%
  filter(!Sample=="NK2402_LTBI_SM_SWAP_024.fcs") %>%
  filter(!Sample=="NK2342_TB_X_SWAP_006.fcs") %>%
  filter(!Sample=="NK2136_TB_X_WCL_003.fcs") %>%
  filter(!Sample=="NK2168_HC_SM_WCL_040.fcs") %>%
  filter(!Sample=="NK2421_LTBI_SM_PEP_010.fcs") %>%
  filter(!Sample=="NK2063_TB_X_SEA_005.fcs") %>%
  filter(!Sample=="NK2115_LTBI_X_PEP_010.fcs") %>%
  filter(!Sample=="NK2115_LTBI_X_WCL_009.fcs") %>%
  filter(!Sample=="NK2171_HC_X_WCL_009.fcs") %>%
  filter(!Sample=="NK2162_HC_X_SEA_023.fcs")
CD4_count[CD4_count=="¥"]<-NA

Split<-strsplit(as.character(CD4_count$Sample),"_");
CD4_count$ID = as.factor(sapply(Split,"[",1));
CD4_count$TB = as.factor(sapply(Split,"[",2));
CD4_count$SM = as.factor(sapply(Split,"[",3));
CD4_count$Antigen = as.factor(sapply(Split,"[",4));
CD4_count$cell<-"CD4"

CD4_count<-as.data.table(CD4_count)
CD4_count<-CD4_count[CD4_count$SM!="HEL"]

CD4_count$SM<-factor(CD4_count$SM, levels=c("N","X","SM"))
CD4_count$TB<-factor(CD4_count$TB, levels=c("N","HC", "LTBI", "TB"))
CD4_count$Stim = factor(CD4_count$Antigen,levels=c("UN", "WCL","PEP","SEA", "SWAP", "PMA"))
CD4_count<-CD4_count[!(is.na(CD4_count$Sample))]

write.csv(CD4_count,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/CD4_count.csv")

CD8<-read.csv("CD8_counts.csv")
CD8<-filter(CD8,!(Sample=="Mean"))%>%
filter(!Sample=="StdDev")%>%
filter(!Sample=="Sample")%>%
    filter(!Sample=="NK2126_HC_X_SEA_005.fcs") %>%
  filter(!Sample=="NK2450_HC_SM_SEA_047.fcs") %>%
  filter(!Sample=="NK2402_LTBI_SM_SWAP_024.fcs") %>%
  filter(!Sample=="NK2342_TB_X_SWAP_006.fcs") %>%
  filter(!Sample=="NK2136_TB_X_WCL_003.fcs") %>%
  filter(!Sample=="NK2168_HC_SM_WCL_040.fcs") %>%
  filter(!Sample=="NK2421_LTBI_SM_PEP_010.fcs") %>%
  filter(!Sample=="NK2063_TB_X_SEA_005.fcs") %>%
  filter(!Sample=="NK2115_LTBI_X_PEP_010.fcs") %>%
  filter(!Sample=="NK2115_LTBI_X_WCL_009.fcs") %>%
  filter(!Sample=="NK2171_HC_X_WCL_009.fcs") %>%
  filter(!Sample=="NK2162_HC_X_SEA_023.fcs")
CD8[CD8=="¥"]<-NA

Split<-strsplit(as.character(CD8$Sample),"_");
CD8$ID = as.factor(sapply(Split,"[",1));
CD8$TB = as.factor(sapply(Split,"[",2));
CD8$SM = as.factor(sapply(Split,"[",3));
CD8$Antigen = as.factor(sapply(Split,"[",4));
CD8$cell<-"CD8"

CD8<-as.data.table(CD8)
CD8<-CD8[CD8$SM!="HEL"]

CD8$SM<-factor(CD8$SM, levels=c("N","X","SM"))
CD8$TB<-factor(CD8$TB, levels=c("N","HC", "LTBI", "TB"))
CD8$Stim = factor(CD8$Antigen,levels=c("UN", "WCL","PEP","SEA", "SWAP", "PMA"))
CD8<-CD8[!(is.na(CD8$Sample))]

write.csv(CD8,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/CD8_count.csv")

GD<-read.csv("GD_counts.csv")
GD<-filter(GD,!(Sample=="Mean"))%>%
filter(!Sample=="StdDev")%>%
filter(!Sample=="Sample")%>%
    filter(!Sample=="NK2126_HC_X_SEA_005.fcs") %>%
  filter(!Sample=="NK2450_HC_SM_SEA_047.fcs") %>%
  filter(!Sample=="NK2402_LTBI_SM_SWAP_024.fcs") %>%
  filter(!Sample=="NK2342_TB_X_SWAP_006.fcs") %>%
  filter(!Sample=="NK2136_TB_X_WCL_003.fcs") %>%
  filter(!Sample=="NK2168_HC_SM_WCL_040.fcs") %>%
  filter(!Sample=="NK2421_LTBI_SM_PEP_010.fcs") %>%
  filter(!Sample=="NK2063_TB_X_SEA_005.fcs") %>%
  filter(!Sample=="NK2115_LTBI_X_PEP_010.fcs") %>%
  filter(!Sample=="NK2115_LTBI_X_WCL_009.fcs") %>%
  filter(!Sample=="NK2171_HC_X_WCL_009.fcs") %>%
  filter(!Sample=="NK2162_HC_X_SEA_023.fcs")
GD[GD=="¥"]<-NA

Split<-strsplit(as.character(GD$Sample),"_");
GD$ID = as.factor(sapply(Split,"[",1));
GD$TB = as.factor(sapply(Split,"[",2));
GD$SM = as.factor(sapply(Split,"[",3));
GD$Antigen = as.factor(sapply(Split,"[",4));
GD$cell<-"GD"

GD<-as.data.table(GD)
GD<-GD[GD$SM!="HEL"]

GD$SM<-factor(GD$SM, levels=c("N","X","SM"))
GD$TB<-factor(GD$TB, levels=c("N","HC", "LTBI", "TB"))
GD$Stim = factor(GD$Antigen,levels=c("UN", "WCL","PEP","SEA", "SWAP", "PMA"))
GD<-GD[!(is.na(GD$Sample))]

write.csv(GD,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/GD_count.csv")

```

### 2018-10-21 Committee Meeting 2017.Rmd
Add files via upload

---
title: "Committee Meeting 2017"
output:
  pdf_document: default
  html_notebook: default
---
setwd("/Users/oldmac/Desktop/Day Lab/Data")

```{r}
FP<- read.csv("Fix Perm.csv")
splStudyID = strsplit(as.character(FP$Sample),":");
FP$Sample = as.factor(sapply(splStudyID,"[",2));
splStudyID = strsplit(as.character(FP$Sample),"_");
FP$Donor = as.factor(sapply(splStudyID,"[",1));
FP$Antigen = as.factor(sapply(splStudyID,"[",2));
FP$FP = as.factor(sapply(splStudyID,"[",3));
FP$Stim = factor(FP$Antigen,levels=c("UN", "WCL","Pep","PMA"))
FP$Condition <- as.character("Frozen")
```

```{r}
Fresh<-read.csv("Fresh.csv")
splStudyID = strsplit(as.character(Fresh$Sample),":");
Fresh$Sample = as.factor(sapply(splStudyID,"[",2));
splStudyID = strsplit(as.character(Fresh$Sample),"_");
Fresh$Donor = as.factor(sapply(splStudyID,"[",1));
Fresh$Antigen = as.factor(sapply(splStudyID,"[",2));
Fresh$Stim = factor(Fresh$Antigen,levels=c("UN", "WCL","Pep","PMA"))
Fresh$Condition <- as.character("Fresh")
```

```{r}
Frozen<-read.csv("Frozen.csv")
splStudyID = strsplit(as.character(Frozen$Sample),":");
Frozen$Sample = as.factor(sapply(splStudyID,"[",2));
splStudyID = strsplit(as.character(Frozen$Sample),"_");
Frozen$Donor = as.factor(sapply(splStudyID,"[",1));
Frozen$Antigen = as.factor(sapply(splStudyID,"[",2));
Frozen$Stim = factor(Frozen$Antigen,levels=c("UN", "WCL","Pep","PMA"))
Frozen$FP = as.factor(sapply(splStudyID,"[",3));
Frozen$Condition <- as.character("Frozen")
```

```{r}
TB<-rbind.fill(Fresh,Frozen)
TB<-data.table(TB)
TB<-TB[!(Stim=="PMA")]
```

```{r}
pd <- position_dodge(width=0.5)

ggplot(TB, aes(x=Stim, y=CD4.TH1, color=FP)) + 
  geom_point(position=pd, size=10) +
  xlab("Stimulation") +
  ylab("% TH1 Cytokine+ CD4+ T cells") +
  ggtitle("TH1") +
  theme_bw()+
  theme(text = element_text(size = 30))+
  theme(legend.position="bottom")
```

```{r}
ggplot(TB, aes(x=Stim, y=CD4.TH1.2, color=FP)) + 
  geom_point(position=pd, size=10) +
  xlab("Stimulation") +
  ylab("% TH1/2 Cytokine+ CD4+ T cells") +
  ggtitle("TH1/TH2") +
  theme_bw()+
  theme(text = element_text(size = 30))+
  theme(legend.position="bottom")
```

```{r}
ggplot(TB, aes(x=Stim, y=CD4.TH2, color=FP)) + 
  geom_point(position=pd, size=10) +
  xlab("Stimulation") +
  ylab("% TH2 Cytokine+ CD4+ T cells") +
  ggtitle("TH2") +
  theme_bw()+
  theme(text = element_text(size = 30))+
  theme(legend.position="bottom")
```

```{r}
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  library(plyr)

  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }

  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )

  # Rename the "mean" column    
  datac <- rename(datac, c("mean" = measurevar))

  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult

  return(datac)
}
```

```{r}
Dat1<-summarySE(FP, measurevar="CD4.TH1", groupvars=c("Stim","FP"))
ggplot(Dat1, aes(x=Stim, y=CD4.TH1, color=FP)) + 
  geom_errorbar(aes(ymin=CD4.TH1-se, ymax=CD4.TH1+se), colour="black", width=.1, position=pd) +
  geom_line(position=pd) +
  geom_point(position=pd, size=10) +
  xlab("Stimulation") +
  ylab("% TH1 Cytokine+ CD4+ T cells") +
  ggtitle("TH1") +
  theme_bw()+
  theme(text = element_text(size = 30))+
  theme(legend.position="bottom")
```

```{r}
Dat2<-summarySE(FP, measurevar="CD4.TH1.2", groupvars=c("Stim","FP"))
ggplot(Dat2, aes(x=Stim, y=CD4.TH1.2, color=FP)) + 
  geom_errorbar(aes(ymin=CD4.TH1.2-se, ymax=CD4.TH1.2+se), colour="black", width=.1, position=pd) +
  geom_line(position=pd) +
  geom_point(position=pd, size=10) +
  xlab("Stimulation") +
  ylab("% TH1/2 Cytokine+ CD4+ T cells") +
  ggtitle("TH1/TH2") +
  theme_bw()+
  theme(text = element_text(size = 30))+
  theme(legend.position="bottom")
```

```{r}
Dat3<-summarySE(FP, measurevar="CD4.TH2", groupvars=c("Stim","FP"))
ggplot(Dat3, aes(x=Stim, y=CD4.TH2, color=FP)) + 
  geom_errorbar(aes(ymin=CD4.TH2-se, ymax=CD4.TH2+se), colour="black", width=.1, position=pd) +
  geom_line(position=pd) +
  geom_point(position=pd, size=10) +
  xlab("Stimulation") +
  ylab("% TH2 Cytokine+ CD4+ T cells") +
  ggtitle("TH2") +
  theme_bw()+
  theme(text = element_text(size = 30))+
  theme(legend.position="bottom")
```

### 2018-10-21 Compass_analysis.Rmd
Add files via upload

---
title: "COMPASS_analysis"
author: "Taryn McLaughlin"
date: "9/23/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 4, fig.width = 3)
```

```{r p-value function, echo=TRUE}
library(dplyr)
pvals<-function(data, column){

    data<-filter(data, SM!="N")
    data$SM<-factor(data$SM, levels=c("X","SM"))
    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))

    x<-split(data,data$TB) 
    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"]))

    y<-split(data,data$SM) 
    B<-lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "LTBI")))
    C<- lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("LTBI", "TB")))
    D<- lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "TB")))

    labels<-c("HC SM+ to HC SM-", "LTBI SM+ to LTBI SM-", "TB SM+ to TB SM-",
              "HC SM+ to LTBI SM+", "LTBI SM+ to TB SM+", "HC SM+ to TB SM+",
              "HC SM- to LTBI SM-", "LTBI SM- to TB SM-", "HC SM- to TB SM-"
    )
    pvals<-c(A$HC$p.value, A$LTBI$p.value, A$TB$p.value,
             B$SM$p.value, C$SM$p.value,  D$SM$p.value,
             B$X$p.value,  C$X$p.value,   D$X$p.value)

    bonf_adj_pvals<-round(p.adjust(pvals, method="bonferroni"), 4)

    fdr_adj_pvals<-round(p.adjust(pvals, method="fdr"), 4)

    table<-(cbind(labels, round(pvals, 4), bonf_adj_pvals, fdr_adj_pvals))
    colnames(table)<-c("Comparison", paste(column,"p-values"," "), 
                       "Bonferonni Corrected p-values", "FDR Corrected p-vales")
    table
}
```

```{r COMPASS fits with Yerkes Donors, cache=TRUE, fig.height=10, fig.width=7}
library(dplyr)
library(COMPASS)
DF<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/CD4_count.csv")
compass_data<-DF
compass_data<-rename(compass_data, 
    "IFNg&IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count",
    "IFNg&IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.1",
    "IFNg&IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.2",
    "IFNg&IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.3",
    "IFNg&!IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.4",
    "IFNg&!IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.5",
    "IFNg&!IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.6",
    "IFNg&!IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.7",
    "!IFNg&IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.8",
    "!IFNg&IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.9",
    "!IFNg&IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.10",
    "!IFNg&IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.11",
    "!IFNg&!IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.12",
    "!IFNg&!IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.13",
    "!IFNg&!IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.14",
    "!IFNg&!IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.15")

compass_data$TB<-factor(compass_data$TB, levels= c("N","HC","LTBI","TB"))
compass_data$SM<-factor(compass_data$SM, levels= c("N", "X", "SM"))
compass_data$Stim<-factor(compass_data$Stim, levels=c("UN", "PMA", "WCL", "PEP", "SEA", "SWAP"))
metadata<-dplyr::select(compass_data, c("ID","TB","SM","Stim"))
metadata$group<-paste(metadata$TB,metadata$SM)
counts<- as.matrix(select(compass_data, "IFNg&IL4&IL13&TNFa", 
         "IFNg&IL4&IL13&!TNFa", 
         "IFNg&IL4&!IL13&TNFa",
         "IFNg&IL4&!IL13&!TNFa",
         "IFNg&!IL4&IL13&TNFa",
         "IFNg&!IL4&IL13&!TNFa",
         "IFNg&!IL4&!IL13&TNFa",
         "IFNg&!IL4&!IL13&!TNFa",
         "!IFNg&IL4&IL13&TNFa",
         "!IFNg&IL4&IL13&!TNFa",
         "!IFNg&IL4&!IL13&TNFa",
         "!IFNg&IL4&!IL13&!TNFa",
         "!IFNg&!IL4&IL13&TNFa",
         "!IFNg&!IL4&IL13&!TNFa",
         "!IFNg&!IL4&!IL13&TNFa",
         "!IFNg&!IL4&!IL13&!TNFa"))

metadata$unique_id <- metadata$ID

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")

#WCL
n_wcl <- subset(counts, metadata$Stim =="WCL");
rownames(n_wcl) <- subset(metadata$unique_id,metadata$Stim=="WCL")
metadata_wcl<-subset(metadata,metadata$Stim=="WCL")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_wcl))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_wcl<-COMPASS::SimpleCOMPASS (n_s=n_wcl, n_u=n_u, meta=metadata_wcl, individual_id = "unique_id", iterations = 10000, replications = 8, verbose = TRUE)

#PEP
n_pep <- subset(counts, metadata$Stim =="PEP")
rownames(n_pep) <- subset(metadata$unique_id,metadata$Stim=="PEP")
metadata_pep<-subset(metadata,metadata$Stim=="PEP")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_pep))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_pep<-COMPASS::SimpleCOMPASS (n_s=n_pep, n_u=n_u, meta=metadata_pep, individual_id = "unique_id", iterations = 10000, 
                                 replications = 8, verbose = TRUE)

#PMA
n_pma <- subset(counts, metadata$Stim =="PMA")
rownames(n_pma) <- subset(metadata$unique_id,metadata$Stim=="PMA")
metadata_pma<-subset(metadata,metadata$Stim=="PMA")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_pma))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_pma<-COMPASS::SimpleCOMPASS (n_s=n_pma, n_u=n_u, meta=metadata_pma, individual_id = "unique_id", iterations = 10000, 
                                 replications = 8, verbose = TRUE)

#SEA
n_sea <- subset(counts, metadata$Stim =="SEA")
rownames(n_sea) <- subset(metadata$unique_id,metadata$Stim=="SEA")
metadata_sea<-subset(metadata,metadata$Stim=="SEA")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_sea))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_sea<-COMPASS::SimpleCOMPASS (n_s=n_sea, n_u=n_u, meta=metadata_sea, individual_id = "unique_id", iterations = 10000, 
                                 replications = 8, verbose = TRUE)

#SWAP
n_swap<- subset(counts, metadata$Stim =="SWAP")
rownames(n_swap) <- subset(metadata$unique_id,metadata$Stim=="SWAP")
metadata_swap<-subset(metadata,metadata$Stim=="SWAP")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_swap))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_swap<-COMPASS::SimpleCOMPASS (n_s=n_swap, n_u=n_u, meta=metadata_swap, individual_id = "unique_id", iterations = 10000, 
                                  replications = 8, verbose = TRUE)
```

```{r COMPASS heatmaps yerkes, error =TRUE}
library(COMPASS)
plot(fit_pma, order_by_max_functionality = FALSE, markers=c("IL13","IL4","IFNg","TNFa"), 
     row_annotation=c("TB","SM"), border_color=NA, fontsize=14, threshold=0)
plot(fit_wcl, order_by_max_functionality = FALSE, markers=c("IL13","IL4","IFNg","TNFa"), 
     row_annotation=c("TB","SM"), border_color=NA, fontsize=14, threshold=0)
plot(fit_pep, order_by_max_functionality = FALSE, markers=c("IL13","IL4","IFNg","TNFa"), 
     row_annotation=c("TB","SM"), border_color=NA, fontsize=14, threshold=0)
```

```{r COMPASS scores yerkes, error=TRUE}
library(COMPASS)
library(ggplot2)
library(knitr)
library(dplyr)

####### PMA
scores_pma<-COMPASS::scores(fit_pma)
g<-ggplot(scores_pma, aes(x=SM, y=FS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Responses",
         subtitle="Functionality Score",
         caption="",
         x="TB and S. mansoni Status",
         y="Functionality Score")

g<-ggplot(scores_pma, aes(x=SM, y=PFS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Responses",
         subtitle="Polyfunctionality Scores",
         caption="",
         x="TB and S. mansoni Status",
         y="Polyunctionality Score")

kable(pvals(scores_pma, "FS") , caption = "Non Specific CD4+ T cell Responses")
kable(pvals(scores_pma, "PFS"))

#### WHOLE CELL LYSATE
scores_wcl<-COMPASS::scores(fit_wcl)
g<-ggplot(scores_wcl, aes(x=SM, y=FS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="Functionality Scores",
         caption="",
         x="TB and S. mansoni Status",
         y="Functionality Score")
g<-ggplot(scores_wcl, aes(x=SM, y=PFS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="Polyfunctionality Scores",
         caption="",
         x="TB and S. mansoni Status",
         y="Polyunctionality Score")
kable(pvals(scores_wcl, "FS"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
kable(pvals(scores_wcl, "PFS"))

######## PEPTIDE
scores_pep<-COMPASS::scores(fit_pep)
g<-ggplot(scores_pep, aes(x=SM, y=FS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Peptide Specific CD4 Responses", 
         subtitle="Functionality Scores",
         caption="",
         x="Disease Status",
         y="Functionality Score")
g<-ggplot(scores_pep, aes(x=SM, y=PFS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Peptide Specific CD4 Responses", 
         subtitle="Polyfunctionality Score",
         caption="",
         x="Disease Status",
         y="Polyunctionality Score")

kable(pvals(scores_pep, "FS"), caption = "Mtb Peptide Specific CD4 T cell Responses")
kable(pvals(scores_pep, "PFS"))
```

```{r COMPASS fits without Yerkes Donors, cache=TRUE, fig.height=10, fig.width=7}
library(dplyr)
library(COMPASS)
DF<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/CD4_count.csv")
compass_data<-filter(DF, TB!="N")
compass_data$SM<-factor(compass_data$SM, levels=c("X","SM"))
compass_data$TB<-factor(compass_data$TB, levels=c("HC","LTBI","TB"))
compass_data<-rename(compass_data, 
    "IFNg&IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count",
    "IFNg&IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.1",
    "IFNg&IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.2",
    "IFNg&IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.3",
    "IFNg&!IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.4",
    "IFNg&!IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.5",
    "IFNg&!IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.6",
    "IFNg&!IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.7",
    "!IFNg&IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.8",
    "!IFNg&IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.9",
    "!IFNg&IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.10",
    "!IFNg&IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.11",
    "!IFNg&!IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.12",
    "!IFNg&!IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.13",
    "!IFNg&!IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.14",
    "!IFNg&!IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.15")

compass_data$Stim<-factor(compass_data$Stim, levels=c("UN", "PMA", "WCL", "PEP", "SEA", "SWAP"))
metadata<-dplyr::select(compass_data, c("ID","TB","SM","Stim"))
metadata$group<-paste(metadata$TB,metadata$SM)
counts<- as.matrix(select(compass_data, "IFNg&IL4&IL13&TNFa", 
         "IFNg&IL4&IL13&!TNFa", 
         "IFNg&IL4&!IL13&TNFa",
         "IFNg&IL4&!IL13&!TNFa",
         "IFNg&!IL4&IL13&TNFa",
         "IFNg&!IL4&IL13&!TNFa",
         "IFNg&!IL4&!IL13&TNFa",
         "IFNg&!IL4&!IL13&!TNFa",
         "!IFNg&IL4&IL13&TNFa",
         "!IFNg&IL4&IL13&!TNFa",
         "!IFNg&IL4&!IL13&TNFa",
         "!IFNg&IL4&!IL13&!TNFa",
         "!IFNg&!IL4&IL13&TNFa",
         "!IFNg&!IL4&IL13&!TNFa",
         "!IFNg&!IL4&!IL13&TNFa",
         "!IFNg&!IL4&!IL13&!TNFa"))

metadata$unique_id <- metadata$ID

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")

#WCL
n_wcl <- subset(counts, metadata$Stim =="WCL");
rownames(n_wcl) <- subset(metadata$unique_id,metadata$Stim=="WCL")
metadata_wcl<-subset(metadata,metadata$Stim=="WCL")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_wcl))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_wcl_noN<-COMPASS::SimpleCOMPASS (n_s=n_wcl, n_u=n_u, meta=metadata_wcl, individual_id = "unique_id", iterations = 10000, replications = 8, verbose = TRUE)

#PEP
n_pep <- subset(counts, metadata$Stim =="PEP")
rownames(n_pep) <- subset(metadata$unique_id,metadata$Stim=="PEP")
metadata_pep<-subset(metadata,metadata$Stim=="PEP")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_pep))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_pep_noN<-COMPASS::SimpleCOMPASS (n_s=n_pep, n_u=n_u, meta=metadata_pep, individual_id = "unique_id", iterations = 10000, 
                                 replications = 8, verbose = TRUE)

#PMA
n_pma <- subset(counts, metadata$Stim =="PMA")
rownames(n_pma) <- subset(metadata$unique_id,metadata$Stim=="PMA")
metadata_pma<-subset(metadata,metadata$Stim=="PMA")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_pma))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_pma_noN<-COMPASS::SimpleCOMPASS (n_s=n_pma, n_u=n_u, meta=metadata_pma, individual_id = "unique_id", iterations = 10000, 
                                 replications = 8, verbose = TRUE)

#SEA
n_sea <- subset(counts, metadata$Stim =="SEA")
rownames(n_sea) <- subset(metadata$unique_id,metadata$Stim=="SEA")
metadata_sea<-subset(metadata,metadata$Stim=="SEA")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_sea))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_sea_noN<-COMPASS::SimpleCOMPASS (n_s=n_sea, n_u=n_u, meta=metadata_sea, individual_id = "unique_id", iterations = 10000, 
                                 replications = 8, verbose = TRUE)

#SWAP
n_swap<- subset(counts, metadata$Stim =="SWAP")
rownames(n_swap) <- subset(metadata$unique_id,metadata$Stim=="SWAP")
metadata_swap<-subset(metadata,metadata$Stim=="SWAP")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_swap))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_swap_noN<-COMPASS::SimpleCOMPASS (n_s=n_swap, n_u=n_u, meta=metadata_swap, individual_id = "unique_id", iterations = 10000, 
                                  replications = 8, verbose = TRUE)
```

```{r COMPASS heatmaps, error =TRUE}
library(COMPASS)
plot(fit_pma_noN, order_by_max_functionality = FALSE, markers=c("IL13","IL4","IFNg","TNFa"),
     row_annotation=c("TB","SM"), border_color=NA, fontsize=14, threshold=0)
plot(fit_wcl_noN, order_by_max_functionality = FALSE, markers=c("IL13","IL4","IFNg","TNFa"),
     row_annotation=c("TB","SM"), border_color=NA, fontsize=14, threshold=0)
plot(fit_pep_noN, order_by_max_functionality = FALSE, markers=c("IL13","IL4","IFNg","TNFa"),
     row_annotation=c("TB","SM"), border_color=NA, fontsize=14, threshold=0)
```

```{r COMPASS scores, error =TRUE}
library(COMPASS)
library(ggplot2)
library(knitr)

####### PMA
scores_pma_noN<-COMPASS::scores(fit_pma_noN)
g<-ggplot(scores_pma_noN, aes(x=SM, y=FS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Responses",
         subtitle="Functionality Score",
         caption="",
         x="TB and S. mansoni Status",
         y="Functionality Score")
g<-ggplot(scores_pma_noN, aes(x=SM, y=PFS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Responses",
         subtitle="Polyfunctionality Scores",
         caption="",
         x="TB and S. mansoni Status",
         y="Polyunctionality Score")

kable(pvals(scores_pma_noN, "FS"), caption = "Non-Specific CD4+ T cell Responses")
kable(pvals(scores_pma_noN, "PFS"))

#### WHOLE CELL LYSATE
scores_wcl_noN<-COMPASS::scores(fit_wcl_noN)
g<-ggplot(scores_wcl_noN, aes(x=SM, y=FS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="Functionality Scores",
         caption="",
         x="Disease Status",
         y="Functionality Score")
g<-ggplot(scores_wcl_noN, aes(x=SM, y=PFS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="Polyfunctionality Scores",
         caption="",
         x="TB and S. mansoni Status",
         y="Polyunctionality Score")

kable(pvals(scores_wcl_noN, "FS"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
kable(pvals(scores_wcl_noN, "PFS"))

######## PEPTIDE
scores_pep_noN<-COMPASS::scores(fit_pep_noN)
g<-ggplot(scores_pep_noN, aes(x=SM, y=FS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Peptide Specific CD4+ T cell Responses", 
         subtitle="Functionality Scores",
         caption="",
         x="TB and S. mansoni Status",
         y="Functionality Score")
g<-ggplot(scores_pep_noN, aes(x=SM, y=PFS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Peptide Specific CD4+ T cell Responses", 
         subtitle="Polyfunctionality Score",
         caption="",
         x="TB and S. mansoni Status",
         y="Polyunctionality Score")

kable(pvals(scores_pep_noN, "FS"), caption = "Mtb Peptide Specific CD4+ T cell Responses")
kable(pvals(scores_pep_noN, "PFS"))

```

### 2018-10-21 Graphics.Rmd
Add files via upload

---
title: "Graphs"
output: html_document
---
http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
library(RColorBrewer)
display.brewer.all()
http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html

#67001f dark red
#b2182b red
#d1e5f0 light blue
#2166ac dark blue
#1a9850 green
#91cf60 light green

```{r}
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  library(plyr)

  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }

  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )

  # Rename the "mean" column    
  datac <- rename(datac, c("mean" = measurevar))

  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult

  return(datac)
}
```

```{r}
dodge <- position_dodge(width = 1)
Infection=c("#9e9e9e","#900C3F", "#1E88E5")
Helper=c("#C86CC2","#3787CE","#000A61")
Palette=c( "#1E88E5", "#900C3F")
Q<-c("#009999","#663399","#CC0066")
```

boolean
```{r setup, include=FALSE}
#HC
HC<-subset(melt, TB %in% "HC")
g<-ggplot(HC, aes(x=label, y=value, color=SM))

g +
  xlab("") +
  ylab("Frequency Cytokine+ CD4+ T-cells (%)") +
  ggtitle("HC") +
  geom_boxplot(outlier.shape = NA, size=1.8, position=dodge)+
  scale_color_manual((values=Infection), name="Helminth Status")+
  theme_bw()+
  theme(axis.text=element_text(size=20,face="bold"),axis.title=element_text(size=24,face="bold"),plot.title=element_text(size=30,face="bold"))+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.key.size = unit(2.5, "cm"))+
  theme(legend.text = element_text(size = 20))

```

TB groups +/- SM for whatever
```{r}
data<-CYT[Stim=="WCL"]
data<-subset(CYT,CYT$Stim=="PMA")
g <- ggplot(data, aes(x=SM, y=Cyt.Freq, color=SM))

g + geom_boxplot(outlier.shape = NA, size=1) + 
        geom_jitter(width=.1,height=0, shape=1,size=2) +
        scale_color_manual(labels = c("SM(-)", "SM(+)"), values = Palette) +
        scale_x_discrete(breaks=c("X","SM"),
                           labels=c("SM(-)", "SM(+)"))+
        theme_bw()+
        theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
        facet_wrap(~TB,nrow = 1, scale="free_x")+
        labs(title="Cytokine Frequencies", 
             subtitle="",
             caption="",
             x="TB and Helminth Status",
             y="Frequency of Cytokine+ CD4+ T cells")

#
geom_boxplot(outlier.shape = NA, size=2, position=dodge)+
coord_cartesian(ylim=c(0,100))+

theme(legend.key.size = unit(2.5, "cm"))+
theme(legend.text = element_text(size = 20))+
theme_bw()+
theme(plot.title = element_text(hjust = 0.5))+
theme(axis.text=element_text(size=16),axis.title=element_text(size=20,face="bold"),plot.title=element_text(size=24,face="bold"))+
theme(
  axis.text.x = element_blank(),
  axis.text.y = element_blank(),
  axis.ticks = element_blank())+

```

Ways to save images
```{r}
jpeg("jpeg", width = 15, height = 9, units = 'in', res = 300)
###plot
dev.off()

tiff('tiff', units="in", width=15, height=9, res=300, compression = 'lzw')
###plot
dev.off()

png("png", width = 15, height = 9, units = 'in', res = 300)
###plot
dev.off()
```

pie charts
```{r}
pie(tf$mean[tf$TB=="TB" & tf$SM=="SM"],labels=tf$variable, col=Palette)
pie(tf$mean[tf$TB=="LTBI" & tf$SM=="SM"],labels=tf$variable, col=Palette)
pie(tf$mean[tf$TB=="HC" & tf$SM=="SM"],labels=tf$variable, col=Palette)
pie(tf$mean[tf$TB=="TB" & tf$SM=="X"],labels=tf$variable, col=Palette)
pie(tf$mean[tf$TB=="LTBI" & tf$SM=="X"],labels=tf$variable, col=Palette)
pie(tf$mean[tf$TB=="HC" & tf$SM=="X"],labels=tf$variable, col=Palette)

pie(cc$mean[cc$TB=="TB" & cc$SM=="SM"],labels=cc$variable, col=Palette)
pie(cc$mean[cc$TB=="LTBI" & cc$SM=="SM"],labels=cc$variable, col=Palette)
pie(cc$mean[cc$TB=="HC" & cc$SM=="SM"],labels=cc$variable, col=Palette)
pie(cc$mean[cc$TB=="TB" & cc$SM=="X"],labels=cc$variable, col=Palette)
pie(cc$mean[cc$TB=="LTBI" & cc$SM=="X"],labels=cc$variable, col=Palette)
pie(cc$mean[cc$TB=="HC" & cc$SM=="X"],labels=cc$variable, col=Palette)
```

prep for a stacked bar
```{r}
df<-PMA[,c(4,5,27:34)]
df$TH1.Portion<-(df$TH1.Count/df$Cyt.Count)
df$TH2.Portion<-(df$TH2.Count/df$Cyt.Count)
df$TH1.2.Portion<-(df$TH1.2.Count/df$Cyt.Count)
df$TH1.Port<-(df$TH1.Freq/df$Cyt.Freq)
df$TH2.Port<-(df$TH2.Freq/df$Cyt.Freq)
df$TH1.2.Port<-(df$TH1.2.Freq/df$Cyt.Freq)

#unscaled frequencies
df1<-df[,c(1,2,6,8,10)]
df2<-melt(df1,id.vars=c("TB","SM"))
df3<-as.data.frame(
  ddply(df2, c("TB", "SM", "variable"), summarise,
        median = median(value), sd = sd(value),
        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
df3$upper<-(df3$median+df3$error)
df3$lower<-(df3$median-df3$error)

df3$variable<-factor(df3$variable, levels=c('TH2.Freq','TH1.2.Freq','TH1.Freq',ordered=TRUE))
df3$Label <- paste(df3$TB, df3$SM, sep='_')
```

unscaled stacked bar
```{r}
g<-ggplot(df3, aes(SM, median))

g+ geom_bar(aes(fill = factor(variable)), stat="identity", width = 0.7) +
  scale_colour_brewer(palette = "Helper")+
  facet_wrap(~TB,nrow = 1)+
  geom_errorbar(aes(ymax=upper , ymin= lower ),width=.5) +
  theme(panel.background = element_rect(fill='white', colour='white'), 
        panel.grid = element_line(color = NA),
        panel.grid.minor = element_line(color = NA),
        panel.border = element_rect(fill = NA, color = "black"),
        axis.text.x  = element_text(size=10, colour="black", face = "bold"), 
        axis.title.x = element_text(vjust=0.1, face = "bold"),
        axis.text.y = element_text(size=12, colour="black"),
        axis.title.y = element_text(vjust=0.2, size = 12, face = "bold"))

```

### 2018-10-21 ICS_CD4.Rmd
Add files via upload

---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---
library(dplyr)
library(reshape2)
Boolean<-select(Cytokine, Donor, TB, SM, Stim, G_4_13_T, G_4_x_T, G_x_13_T, G_4_13_x, x_4_13_T, G_4_x_x, G_x_13_x,
                x_4_x_T, x_x_13_T, G_x_x_T , G_x_x_x, x_x_x_T,  x_4_13_x, x_4_x_x, x_x_13_x)

melt<-melt(Boolean,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T", "G_x_x_T" , "G_x_x_x","x_x_x_T",  "x_4_13_x","x_4_x_x", "x_x_13_x"))
melt$TB<-factor(melt$TB, levels=c("N","HC","LTBI","TB"))
melt$SM<-factor(melt$SM, levels=c("N","X","SM"))
melt<-filter(melt, !is.na(label))

data<-filter(melt, TB %in% c("HC", "N") & Stim=="PMA")

g<-ggplot(melt, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#67001f"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(Stim~TB, scale="free")

    scale_color_manual(labels = c("SM(-)", "SM(+)"), values = Palette) +
    scale_x_discrete(breaks=c("X","SM"),
                     labels=c("SM(-)", "SM(+)"))+
    theme_bw()+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_wrap(~TB,nrow = 1, scale="free_x")+
    labs(title="Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and Helminth Status",
         y="Frequency of Cytokine+ CD4+ T cells")

```{r}
#Antigen specific responses using total cytokine
Cyt<-read.csv("Cyt.csv")
TH1<-read.csv("TH1.csv")
TH2<-read.csv("TH2.csv")
TH1.2<-read.csv("TH1-2.csv")

Split<- strsplit(as.character(Cyt$Sample)," ")
Cyt$Sample = as.factor(sapply(Split,"[",2))
TH1$Sample = as.factor(sapply(Split,"[",2))
TH2$Sample = as.factor(sapply(Split,"[",2))
TH1.2$Sample = as.factor(sapply(Split,"[",2))

AG<-AG[!(AG$Sample=="NK2037_TB_HEL_T_PMA_020.fcs")]
AG<-AG[!(AG$Sample=="NK2126_HC_X_SEA_005.fcs")]
AG<-AG[!(AG$Sample=="NK2450_HC_SM_SEA_047.fcs")]
AG<-AG[!(AG$Sample=="NK2402_LTBI_SM_SWAP_024.fcs")]
AG<-AG[!(AG$Sample=="NK2342_TB_X_SWAP_006.fcs")]
AG<-AG[!(AG$Sample=="NK2136_TB_X_WCL_003.fcs")]
AG<-AG[!(AG$Sample=="NK2168_HC_SM_WCL_040.fcs")]
AG<-AG[!(AG$Sample=="NK2088_LTBI_HEL_H_PEP_010.fcs")]

AG<-as.data.table(AG)

TH1<-merge(AG,TH1,by="Sample")
TH2<-merge(AG,TH2,by="Sample")
TH1.2<-merge(AG,TH1.2,by="Sample")
Cyt<-merge(AG,Cyt,by="Sample")

Cyt<-Cyt[,c(1:5,21:25,29,30,35:51)]
TH1<-TH1[,c(1:5,21,24,26,31,35:51)]
TH2<-TH2[,c(1:5,22:23,28,32,35:51)]
TH1.2<-TH1.2[,c(1:5,21:24,27,33,35:51)]

```

```{r}

```

Prep for Stacked Bars PMA as example
```{r}
df<-WCL[,c(4,5,27:34)]
df$TH1.Portion<-(df$TH1.Count/df$Cyt.Count)
df$TH2.Portion<-(df$TH2.Count/df$Cyt.Count)
df$TH1.2.Portion<-(df$TH1.2.Count/df$Cyt.Count)
df$TH1.Port<-(df$TH1.Freq/df$Cyt.Freq)
df$TH2.Port<-(df$TH2.Freq/df$Cyt.Freq)
df$TH1.2.Port<-(df$TH1.2.Freq/df$Cyt.Freq)

#unscaled frequencies
df1<-df[,c(1,2,6,8,10)]
df2<-melt(df1,id.vars=c("TB","SM"))
df3<-as.data.frame(
  ddply(df2, c("TB", "SM", "variable"), summarise,
        median = median(value), sd = sd(value),
        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
df3$upper<-(df3$median+df3$error)
df3$lower<-(df3$median-df3$error)

#you have to do these to get the bars to be right
df3$variable<-factor(df3$variable, levels=c('TH2.Freq','TH1.2.Freq','TH1.Freq',ordered=TRUE))
df3$Label <- paste(df3$TB, df3$SM, sep='_')

#adjusting error HC_X
df3$upper[df3$variable=="TH1.2.Freq" & df3$Label=="HC_X"]<-
  df3$median[df3$variable=="TH1.2.Freq" & df3$Label=="HC_X"]+
  df3$median[df3$variable=="TH1.Freq" & df3$Label=="HC_X"]+
  df3$error[df3$variable=="TH1.2.Freq" & df3$Label=="HC_X"]

```

Boolean Cytokines with PMA as an example
```{r}
# to make the cytokine frequency graphs example with PMA
library(dplyr)
library(reshape2)
Boolean<-select(data, Donor, TB, SM, G_4_13_T, G_4_x_T, G_x_13_T, G_4_13_x, x_4_13_T,G_4_x_x, G_x_13_x,x_4_x_T,x_x_13_T, G_x_x_T , G_x_x_x,x_x_x_T,  x_4_13_x,x_4_x_x, x_x_13_x)

melt<-melt(Boolean,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T", "G_x_x_T" , "G_x_x_x","x_x_x_T",  "x_4_13_x","x_4_x_x", "x_x_13_x"))

HC<-subset(melt, TB %in% "HC")
LTBI<-subset(melt, TB %in% "LTBI")
ATB<-subset(melt, TB %in% "TB")

```

Surface Markers on CD4
```{r}
df1<-CD4[,c(7:9,19,20)]
df2<-melt(df1,id.vars=c("TB","SM"))
df3<-as.data.frame(
  ddply(df2, c("TB", "SM", "variable"), summarise,
        median = median(value), sd = sd(value),
        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
df3$upper<-(df3$median+df3$error)
df3$lower<-(df3$median-df3$error)

#you have to do these to get the bars to be right
df3$Label <- paste(df3$TB, df3$SM, sep='_')

df3$variable<-factor(df3$variable, levels=c('CCR4.P_CXCR3.P','CCR4.N_CXCR3.P','CCR4.P_CXCR3.N', ordered=TRUE))
```

Transcription Factors on CD4
```{r}
df1<-data[,c(22:24,4,5)]
df2<-melt(df1,id.vars=c("TB","SM"))
df3<-as.data.frame(
  ddply(df2, c("TB", "SM", "variable"), summarise,
        median = median(value), sd = sd(value),
        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
df3$upper<-(df3$median+df3$error)
df3$lower<-(df3$median-df3$error)

#you have to do these to get the bars to be right
# (grey, neither, top), (cornflower, mix), (orchid, type 2), (navy, type 1, bottom)
df3$Label <- paste(df3$TB, df3$SM, sep='_')
df3$variable<-factor(df3$variable, levels=c("TH1_GATA3.P_Tbet.P",'TH1_GATA3.P_Tbet.N','TH1_GATA3.N_Tbet.P', ordered=TRUE))
``` 
Surface Markers on AG specific
```{r}
df<-TH1[TH1$Stim=="WCL"]
df1<-df[,c(9:11,20,21)]
df2<-melt(df1,id.vars=c("TB","SM"))
df3<-as.data.frame(
  ddply(df2, c("TB", "SM", "variable"), summarise,
        median = median(value), sd = sd(value),
        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
df3$upper<-(df3$median+df3$error)
df3$lower<-(df3$median-df3$error)

#you have to do these to get the bars to be right
# (grey, neither, top), (cornflower, mix), (orchid, type 2), (navy, type 1, bottom)
df3$Label <- paste(df3$TB, df3$SM, sep='_')
df3$variable<-factor(df3$variable, levels=c('CCR4.P_CXCR3.P','CCR4.N_CXCR3.P','CCR4.P_CXCR3.N', ordered=TRUE))
```

Transcription Factors on AG specific
```{r}
df<-TH1[TH1$Stim=="WCL"]
df1<-df[,c(15:17,20,21)]
df2<-melt(df1,id.vars=c("TB","SM"))
df3<-as.data.frame(
  ddply(df2, c("TB", "SM", "variable"), summarise,
        median = median(value), sd = sd(value),
        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
df3$upper<-(df3$median+df3$error)
df3$lower<-(df3$median-df3$error)

#you have to do these to get the bars to be right
# (grey, neither, top), (cornflower, mix), (orchid, type 2), (navy, type 1, bottom)
df3$Label <- paste(df3$TB, df3$SM, sep='_')
df3$variable<-factor(df3$variable, levels=c("GATA3.P_Tbet.P",'GATA3.P_Tbet.N','GATA3.N_Tbet.P', ordered=TRUE))
```

Marker Co-expression
```{r}
CD4_melt<-melt(data,id.vars=c("TB","SM"))
CD4_melt$value<-as.numeric(CD4_melt$value)
CD4_melt<-CD4_melt[!(is.na(CD4_melt$value))]
CD4_stats<-as.data.frame(
  ddply(CD4_melt, c("TB", "SM", "variable"), summarise,
        mean = mean(value), sd = sd(value),
        sem = sd(value)/sqrt(length(value))))

TF<-subset(CD4_melt, variable %in% c("GATA3.N_Tbet.P", "GATA3.P_Tbet.N", "GATA3.P_Tbet.P"))
tf<-subset(CD4_stats, variable %in% c("GATA3.N_Tbet.P", "GATA3.P_Tbet.N", "GATA3.P_Tbet.P"))
CC<-subset(CD4_melt, variable %in% c("CCR4.N_CXCR3.P", "CCR4.P_CXCR3.N", "CCR4.P_CXCR3.P"))
cc<-subset(CD4_stats, variable %in% c("CCR4.N_CXCR3.P", "CCR4.P_CXCR3.N", "CCR4.P_CXCR3.P","CCR4.N_CXCR3.N"))
```

stats? sort of read this http://rcompanion.org/rcompanion/d_08.html
```{r}
data<-subset(CYT,CYT$Stim=="WCL")
data<-filter(TH1,(Stim=="WCL" & TH1.Sum>0))

x <- split(data, data$TB)
lapply(x, function(g) wilcox.test(g$Cyt.Freq~g$SM))
lapply(x, function(g) wilcox.test(g$TH1.Freq~g$SM))
lapply(x, function(g) wilcox.test(g$TH1.2.Freq~g$SM))
lapply(x, function(g) wilcox.test(g$TH2.Freq~g$SM))

y <- split(data, data$SM)
lapply(y, function(g) kruskal.test(g$Cyt.Freq~g$TB))
lapply(y, function(g) kruskal.test(g$TH1.Freq~g$TB))
lapply(y, function(g) kruskal.test(g$TH1.2.Freq~g$TB))
lapply(y, function(g) kruskal.test(g$TH2.Freq~g$TB))

data<-filter(TH1,(Stim=="WCL" & TH1.Sum>0))
x <- split(data, data$TB)
lapply(x, function(g) wilcox.test(g$TH1_Tbet~g$SM))
lapply(x, function(g) wilcox.test(g$TH1_GATA3~g$SM))

y <- split(data, data$SM)
lapply(y, function(g) kruskal.test(g$TH1_Tbet~g$TB))
lapply(y, function(g) kruskal.test(g$TH1_GATA3~g$TB))
```

### 2018-10-21 Stock Code.Rmd
Add files via upload

---
title: "Stock Code"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
#concatenating tables- "fill" includes columns that are different between the two
OG<-rbind.fill(OG.5,OG.8)

#renaming columns
setnames(OG, 
         old=c("Sample","Singlets.Lymphocytes.LD..CD3..Q1..CD4_...CD8..Count") ,
         new=c("StudyID","CD8.Pos"))

#splitting values of a variable by some parameter (only works for strings: factors & characters)
split = strsplit(as.character(OG$StudyID),":")

#create a new column with the a values from the split
OG$Donor = as.factor(sapply(splStudyID,"[",2))

# REORDER A FACTOR VARIABLE
OG$TEST = factor(OG$Antigen,levels=c("UN","PPD","WCL","PEP","SEB"))

#delete entire rows based on the value in a cell- must be a data.table
OG<- OG[!(Donor=="NA")]

#Melt your data table- id.vars are the variables which you want to "melt" by
#can melt a data.table or a data.frame 
melted<-melt(OG, id.vars=c("Antigen", "RE"))
#for some reason it stores all the values as character so coerce them to numeric to run stats
melted$value<-as.numeric(as.character(melted$value))

#summary statistics spit out as a data fram
stats<-as.data.frame(
  ddply(DATA, c("Antigen", "RE", "variable"), summarise,
        mean = mean(value), sd = sd(value),
        sem = sd(value)/sqrt(length(value))))

# trying to mass turn data to numeric from factor
jfhuigls <- data.frame(lapply(OG.RE, function(x) as.numeric(as.character(x))))
melted$value<-as.numeric(as.character(melted$value))

#do a T-test on two subsets of the data
t.test(subset(melted$value,melted$Antigen1 == "UN" & melted$variable=="IFN"),
       subset(melted$value,melted$Antigen1 == "SEB" & melted$variable=="IFN"))

#summary statistics for graphing purposes--> makes a function which can be applied to any data you want
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  library(plyr)

  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }

  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )

  # Rename the "mean" column    
  datac <- rename(datac, c("mean" = measurevar))

  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult

  return(datac)
}

# how to apply the summary statistic function
Dat<-summarySE(OG, measurevar="CD4_OG.L", groupvars=c("Antigen","Time"))

#and then graph it
pd <- position_dodge(0.1) # move error bars .05 to the left and right to avoid overlap
ggplot(Dat, aes(x=Antigen, y=CD4_OG.L, colour=Time)) + 
  geom_errorbar(aes(ymin=CD4_OG.L-se, ymax=CD4_OG.L+se), colour="black", width=.1, position=pd) +
  geom_line(position=pd) +
  geom_point(position=pd, size=3) +
  xlab("Stimulation") +
  ylab("Frequency of Oregon Green Low CD4+ T cells") +
  ggtitle("Proliferation Restimulation Assay") +
  theme_bw()

#simplify graph and use 95% confidence interval instead of SEM
ggplot(tgc, aes(x=dose, y=len, colour=supp)) + 
  geom_errorbar(aes(ymin=len-ci, ymax=len+ci), width=.1, position=pd) +
  geom_line(position=pd) +
  geom_point(position=pd)

#change individual cells based on their content
levels(TB$TBAntigen)[levels(TB$TBAntigen)==">10"] <- "10"
ELISA$Result[ELISA$Result>500]<-500
HIV$HIVResult[HIV$HIVResult==1]<-as.character("POS")

#remove columns
TB<-TB[,c(2,4,5,9,11:15):=NULL]

#subset data to make it easier to work with
NKS<-subset(Final,Final$Study=="NKS")
TBRU<-subset(Final,Final$Study=="TRU")

## create new column with NA as value
TBRU[,"TB.Status"]<-NA

## for each row in the data table
## update new column with value based on other column
for (i in 1:nrow(TBRU)){
  if(TBRU$QualResult[i]=="1") {TBRU$TB.Status[i]<-as.character("LTBI")}
  else {TBRU$TB.Status[i]<-as.character("Healthy Control")}
}

for (i in 1:length(Final$HIV.Status)){
  Final$HIV.Status[i] <- ifelse(is.na(Final$HIVResult[i]), as.character(Final$HIV.STATUS[i]),
                                as.character(Final$HIVResult[i]))
}

#Aggregate data from the same donor in a data set and perform some function on it
##this makes a new data frame with the StudyID and the aggregated data
A = aggregate(AscarisPositive ~ StudyID, data = HEL, function(x) {

  if (1 %in% x) {
    return(1);
  } else {
    return(0);
  }
});

#merge two data sets. match rows by study IDs 
##but keep all rows regardless of whether the StudyID is in both sets
I<-merge(Disease,TB,by="StudyID",all=TRUE)

#Check for duplicated IDs
dups<-Final$StudyID[duplicated(Final$StudyID)]
#NK2081 NK2194 NK2269 NK2464 NK2667 NK2712 80008  80019

#find common StudyIDs and remove the non common StudyIDs from a data set
Inter<- intersect(HEL$StudyID,TB$StudyID)
CleanTB<-TB[TB$StudyID %in% Inter,];

#no clue figure these out
s<-split(Final, Final$HelmithPositive)
i<-interaction(Final$HelmithPositive, Final$QualResult)
S<-str(split(Final,list(Final$HelmithPositive, Final$QualResult), drop=TRUE))

#graph
qplot(test$SchistosomaIntensity,test$QFT, col=test$SchistosomaPositive.y)

#correlation test
# the [] provides a restriction on the data used in the correlation
cor.test((test$SchistosomaIntensity)[test$SchistosomaPositive==1],test$QFT[test$SchistosomaPositive==1])

# linear model
lm((test$SchistosomaIntensity)[test$SchistosomaPositive==1]~test$QFT[test$SchistosomaPositive==1])

#ANOVA
aov((test$SchistosomaIntensity)[test$SchistosomaPositive==1]~test$QFT[test$SchistosomaPositive==1])

###power for proportions
trans.arcsine <- function(x){
  asin(sign(x) * sqrt(abs(x)))
}
h=2*trans.arcsine(b/d)-2*trans.arcsine(a/c)
pwr.2p.test(h=h,n=NULL,sig.level=0.05,power=.8)

for (i in 1:length(Cyt$IFNg)) {
  Cyt$IFNg[i]<-ifelse(Cyt$Stim[i]=="UN",Cyt$IFNg[i],(Cyt$IFNg[i]-Cyt$IFNg[Cyt$Stim=="UN"]))
}

#PCA Google this https://www.r-bloggers.com/computing-and-visualizing-pca-in-r/
#http://www.statmethods.net/stats/frequencies.html

#graphing data http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/#Helper functions
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2018-10-21 Broken Axis.R

@@ -1,0 +1,25 @@
+data<-Cytokine[Cytokine$Stim=="PMA"]
+data$mask = 0
+data$mask[data$TH2.Freq > 1.5] = 1
+data$mask = factor(data$mask, levels = c('1', '0'))
+
+library(ggplot2)
+base.plot <- function(DF, group1, group2, value) {
+        p <- ggplot(DF, aes(x=group1, y=value, col=group1))
+        p <- p + theme_bw()
+        p <- p + theme(legend.position="bottom")
+        p <- p + geom_jitter(width=.1,height=0, shape=1,size=2)
+        p <- p + scale_color_brewer(palette="Set1", guide=guide_legend(ncol=6, title=NULL))
+        p <- p + xlab("") + ylab("")
+        return(p)
+}
+
+p <- base.plot(DF=data, SM, TB, Value)
+
+p <- p + facet_grid(mask~TB, scales="free_y")
+p <- p + theme(strip.text.y = element_blank())
+#The faceting in the Y is backwards and I don't know why or how to change it
+#I have tried releveling the mask variable and that didn't work
+#And why is the bottom chunk huge compared to the top?
+#I tried using the shrink and the space options in facet_grid 
+#to make the bottom box smaller but it didn't work

### 2018-10-21 Committee Meeting 1:30:2017 copy 2.R

@@ -1,0 +1,220 @@
+setwd("/Users/oldmac/Desktop/Day Lab/Data")
+
+######FIX PERM
+FP<- read.csv("Fix Perm.csv")
+splStudyID = strsplit(as.character(FP$Sample),":");
+FP$Sample = as.factor(sapply(splStudyID,"[",2));
+splStudyID = strsplit(as.character(FP$Sample),"_");
+FP$Donor = as.factor(sapply(splStudyID,"[",1));
+FP$Antigen = as.factor(sapply(splStudyID,"[",2));
+FP$FP = as.factor(sapply(splStudyID,"[",3));
+FP$Stim = factor(FP$Antigen,levels=c("UN", "WCL","Pep","PMA"))
+FP$Condition <- as.character("Frozen")
+
+Fresh<-read.csv("Fresh.csv")
+splStudyID = strsplit(as.character(Fresh$Sample),":");
+Fresh$Sample = as.factor(sapply(splStudyID,"[",2));
+splStudyID = strsplit(as.character(Fresh$Sample),"_");
+Fresh$Donor = as.factor(sapply(splStudyID,"[",1));
+Fresh$Antigen = as.factor(sapply(splStudyID,"[",2));
+Fresh$Stim = factor(Fresh$Antigen,levels=c("UN", "WCL","Pep","PMA"))
+Fresh$Condition <- as.character("Fresh")
+
+#plot TB
+TB<-rbind.fill(Fresh,FP)
+TB<-data.table(TB)
+TB<-TB[!(Donor=="HD136")]
+TB<-TB[!(Donor=="HD398")]
+TB<-TB[!(Donor=="HD118")]
+TB<-TB[!(Stim=="PMA")]
+ggplot(TB, aes(x=Stim, y=CD4.TH1, color=FP)) + 
+  geom_point(position=pd, size=10) +
+  xlab("Stimulation") +
+  ylab("Frequency of TH1 Cytokine+ CD4+ T cells") +
+  ggtitle("TH1") +
+  theme_bw()+
+  theme(text = element_text(size = 30))+
+  theme(legend.position="bottom")
+ggplot(TB, aes(x=Stim, y=CD4.TH1.2, color=FP)) + 
+  geom_point(position=pd, size=10) +
+  xlab("Stimulation") +
+  ylab("Frequency of TH1/2 Cytokine+ CD4+ T cells") +
+  ggtitle("TH1/TH2") +
+  theme_bw()+
[…]

### 2018-10-21 Committee Meeting 9:9:2017.R

@@ -1,0 +1,49 @@
+PBMC<-read.csv("PBMC.csv")
+PBMC$Viability<-as.numeric(PBMC$Viability)
+PBMC$SM<-factor(PBMC$SM, levels=c("Negative","Schisto+","HEL"))
+
+ggplot(PBMC, aes(x=TB.Status, y=Viability, color=SM))+
+  xlab("") +
+  ylab("Frequency of Viable Cells") +
+  ggtitle("Viability") +
+  geom_boxplot(outlier.shape = NA, size=1.8, position=dodge)+
+  scale_color_manual(values=Infection, name="Helminth Status")+
+  theme_bw()+
+  theme(axis.text=element_text(size=20,face="bold"),axis.title=element_text(size=24,face="bold"),plot.title=element_text(size=30,face="bold"))+
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(legend.key.size = unit(2.5, "cm"))+
+  theme(legend.text = element_text(size = 20))
+
+ggplot(PBMC, aes(x=TB.Status, y=Cells.Vial, color=SM))+
+  xlab("") +
+  ylab("Number Cells Per Vial Thawed") +
+  ggtitle("Cell Counts") +
+  geom_boxplot(outlier.shape = NA, size=1.8, position=dodge)+
+  scale_color_manual(values=Infection, name="Helminth Status")+
+  theme_bw()+
+  theme(axis.text=element_text(size=20,face="bold"),axis.title=element_text(size=24,face="bold"),plot.title=element_text(size=30,face="bold"))+
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(legend.key.size = unit(2.5, "cm"))+
+  theme(legend.text = element_text(size = 20))+
+  coord_cartesian(ylim = c(0,2E7))
+
+#Boolean plots
+ggplot(DF, aes(x=label, y=value, color=SM))+
+  ylab("Frequency Cytokine+ CD4+ T-cells (%)") +
+  ggtitle(DF$TB) +
+  geom_boxplot(outlier.shape = NA, size=1.8, position=dodge)+
+  scale_color_manual("Helminth Status", values=Infection)+
+  theme_bw()+
+  theme(axis.text=element_text(size=20,face="bold"),axis.title=element_text(size=24,face="bold"),plot.title=element_text(size=30,face="bold"))+
+  theme(plot.title = element_text(hjust = 0.5))+
[…]

### 2018-10-21 Counts.R

@@ -1,0 +1,35 @@
+DF<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Counts.csv", stringsAsFactors = TRUE)
+library(dplyr)
+DF<-filter(DF,!(Sample=="Mean")) %>%
+  filter(!Sample=="StdDev") %>%
+  filter(!Sample=="Sample") %>%
+  filter(!Sample=="NK2126_HC_X_SEA_005.fcs") %>%
+  filter(!Sample=="NK2450_HC_SM_SEA_047.fcs") %>%
+  filter(!Sample=="NK2402_LTBI_SM_SWAP_024.fcs") %>%
+  filter(!Sample=="NK2342_TB_X_SWAP_006.fcs") %>%
+  filter(!Sample=="NK2136_TB_X_WCL_003.fcs") %>%
+  filter(!Sample=="NK2168_HC_SM_WCL_040.fcs") %>%
+  filter(!Sample=="NK2421_LTBI_SM_PEP_010.fcs") %>%
+  filter(!Sample=="NK2063_TB_X_SEA_005.fcs") %>%
+  filter(!Sample=="NK2115_LTBI_X_PEP_010.fcs") %>%
+  filter(!Sample=="NK2115_LTBI_X_WCL_009.fcs") %>%
+  filter(!Sample=="NK2171_HC_X_WCL_009.fcs") %>%
+  filter(!Sample=="NK2162_HC_X_SEA_023.fcs")
+
+source("/Applications/Old Computer/DataScience/Flow Shit/tryna function.R")
+library(data.table)
+data<-Split(DF, column="Sample", splitter="_", one="ID", two="TB", three="SM", four="Stim")
+
+DF2<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/Cytokine.csv", stringsAsFactors = TRUE)
+DF2<-DF2[,c(3,24:27)]
+df1<-merge(data, DF2, by="Sample")
+
+df2<-df1[,2:17]
+colnames(df2)<-compass_nms
+data2<-cbind(df1[,1], df1[,18:21], df2, df1[,22:25])
+
+write.csv(data2,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/All Counts.csv")
+
+
+
+

### 2018-10-21 EVC copy.R

@@ -1,0 +1,154 @@
+setwd("/Users/oldmac/Desktop/Day Lab/Data")
+NK_ICS<-read.csv("NK_ICS.csv")
+OG<-read.csv("EVC_OG.csv")
+IFN<-read.csv("ELISA_IFN.csv")
+IL13<-read.csv("ELISA_IL13.csv")
+
+##ELISA###
+IFN$Cytokine<-as.character("IFNg")
+IL13$Cytokine<-as.character("13")
+setnames(IFN,old="Dilution.Factor",new="Dilution")
+ELISA<-rbind(IFN,IL13)
+splStudyID = strsplit(as.character(ELISA$Sample)," ");
+ELISA$Donor = as.factor(sapply(splStudyID,"[",1));
+ELISA$Antigen <- as.factor(sapply(splStudyID, "[", 2));
+ELISA$Result[ELISA$Result>500]<-500
+ELISA$Adjusted<-(ELISA$Result*ELISA$Dilution)
+
+## combine and melt
+melted<-melt(ELISA, id.vars=c("Donor", "Antigen", "Cytokine"))
+#for some reason it stores all the values as character so coerce them to numeric to run stats
+melted$value<-as.numeric(as.character(melted$value))
+stats<-as.data.frame(
+  ddply(melted, c("Antigen", "Donor", "Cytokine", "variable"), summarise,
+        mean = mean(value), sd = sd(value),
+        sem = sd(value)/sqrt(length(value))))
+
+# re-separate data sets
+ELISA.Mean<-subset(stats,stats$variable=="Adjusted")
+ELISA.Mean<-ELISA.Mean[!("Antigen"=="Pep")]
+
+
+ELISA.Mean[,"Group"]<-NA
+ELISA.Mean$Group[ELISA.Mean$Donor=="HD235"]<-as.character("TH2")
+ELISA.Mean$Group[ELISA.Mean$Donor=="HD355"]<-as.character("HD")
+ELISA.Mean$Group[ELISA.Mean$Donor=="HD203"]<-as.character("HD")
+ELISA.Mean$Group[ELISA.Mean$Donor=="PS16-1001"]<-as.character("SA")
+ELISA.Mean$Group[ELISA.Mean$Donor=="PS16-1002"]<-as.character("SA")
+ELISA.Mean$Group[is.na(ELISA.Mean$Group)]<-as.character("SM")
[…]

### 2018-10-21 Exploratory Pairs Scatterplots.R

@@ -1,0 +1,32 @@
+# another option
+makePairs <- function(data) 
+{
+        grid <- expand.grid(x = 1:ncol(data), y = 1:ncol(data))
+        grid <- subset(grid, x != y)
+        all <- do.call("rbind", lapply(1:nrow(grid), function(i) {
+                xcol <- grid[i, "x"]
+                ycol <- grid[i, "y"]
+                data.frame(xvar = names(data)[ycol], yvar = names(data)[xcol], 
+                           x = data[, xcol], y = data[, ycol], data)
+        }))
+        all$xvar <- factor(all$xvar, levels = names(data))
+        all$yvar <- factor(all$yvar, levels = names(data))
+        densities <- do.call("rbind", lapply(1:ncol(data), function(i) {
+                data.frame(xvar = names(data)[i], yvar = names(data)[i], x = data[, i])
+        }))
+        list(all=all, densities=densities)
+}
+
+# expand iris data frame for pairs plot
+gg1 = makePairs(CD4[,3:18])
+
+# new data frame mega iris
+mega_CD4 = data.frame(gg1$all, Species=rep(CD4$TB, length=nrow(gg1$all)))
+
+# pairs plot
+ggplot(mega_CD4, aes_string(x = "x", y = "y")) + 
+        facet_grid(xvar ~ yvar, scales = "free") + 
+        geom_point(aes(colour=Species), na.rm = TRUE, alpha=0.8) + 
+        stat_density(aes(x = x, y = ..scaled.. * diff(range(x)) + min(x)), 
+                     data = gg1$densities, position = "identity", 
+                     colour = "grey20", geom = "line")

### 2018-10-21 Merging 3 Columns in R For Loop.R

@@ -1,0 +1,43 @@
+TB.1<-c("LTBI", "LTBI", "LTBI", "LTBI", "LTBI", 
+        "ACTIVE", "ACTIVE", "ACTIVE", "HEALTHY CONTROL", "HEALTHY CONTROL",
+        NA,NA,NA,NA,NA)
+TB.2<-c(NA, NA, "LTBI", "LTBI", "HEALTHY CONTROL", 
+        "ACTIVE", "ACTIVE", NA, "HEALTHY CONTROL", "HEALTHY CONTROL",
+        "HEALTHY CONTROL","ACTIVE","LTBI",NA,NA)
+TB.3<-c(NA, "ACTIVE", NA, NA, "HEALTHY CONTROL", 
+        NA, NA, NA,"HEALTHY CONTROL", NA,
+        "HEALTHY CONTROL","ACTIVE","LTBI","LTBI", "HEALTHY CONTROL")
+library(data.table)
+TB<-as.data.table(cbind(TB.1,TB.2,TB.3))
+TB[,"TB.Final"]<-NA
+
+###I'm trying to write a for-loop that combines the information in columns 1-3
+#such that if TB.1 is not NA then TB.Final takes that value
+#if TB.1 is NA then TB.Final takes the value in TB.2
+#if TB.2 is NA then TB.Final takes the value in TB.3
+
+#This one doesn't incorporate the information in TB.3 without an error message
+for (i in 1:length(TB$TB.Final)){
+        if(!is.na(TB$TB.1[i])){
+                TB$TB.Final[i] <- TB$TB.1[i]}
+        
+        else if(!is.na(TB$TB.2[i])){
+                TB$TB.Final[i] <- TB$TB.2[i]} 
+        
+        else{
+                TB$TB.Final[i] <- TB$TB.3[i]} 
+}
+
+#This one also doesn't incorporate the information in TB.3 without an error message
+for (i in 1:length(TB$TB.Final)){
+        if(!is.na(TB$TB.1[i])){
+                TB$TB.Final[i] <- TB$TB.1[i]}
+        
+        else if(is.na(TB$TB.1[i])){
+                TB$TB.Final[i] <- TB$TB.2[i]} 
+        
+        else if (is.na(TB$TB.1[i] & is.na(TB$TB.2[i]))){
+                TB$TB.Final[i] <- TB$TB.3[i]} 
+}
+
+

### 2018-10-21 Trying to make NA values in specific cells.R

@@ -1,0 +1,42 @@
+names<-c("ID1","ID2","ID3", "ID4", "ID5", "ID6", "ID7", "ID8", "ID9", "ID10")
+data<-matrix(data=rep(1:8, 5), nrow=10)
+data<-as.data.table(cbind(names,data))
+names(data)<-c("ID","CXCR3_MFI","blank","blank2","CXCR3_Freq")
+
+#I'm super lost on how to combine various tools to achieve the following
+#I want to make certain values of columns that contain the string "CXCR3" NA
+#based on the ID in the ID column
+
+#I know I can select columns containing "CXCR3" using contains in dplyr
+library(dplyr)
+df1<-data %>% select(contains("CXCR3"))
+#but I can't combine that with "ID" these both give errors
+df1<-data %>% select(contains("CXCR3|ID"))
+df1<-data %>% select(contains("CXCR3") | contains("ID"))
+
+#So then I tried a million variations of grep commands
+#This one gives me a 3 rows all columns
+df2 <- data [grep("CXCR3|ID", names(data))]
+#This one has an error
+df3 <- data [grepl("CXCR3|ID", names(data))]
+#This one give me a logical vector of all FALSE
+df4 <- grepl("ID | CXCR3", names(data))
+
+#this gives me the correct column names but I can never seem to subset the data based on this
+colnames<-names(data)[grep("ID|CXCR3", names(data))]
+df<-select(data, (names(data) %in% colnames))
+df<-data[(names(data) %in% colnames)]
+df<-subset(data, names(data) %in% colnames)
+
+#So then I gave up but this is what I want to do 
+#but without having to type out every column name <-NA
+important_cols = grep('CXCR3', names(data))
+for (i in 1:length(data)){
+        if(data$ID[i] %in% c("ID1", "ID4", "ID5")){
+                for(j in important_cols){
+                        data[i,j] <- NA;
+                }        
+        }
+}
+
+

### 2018-10-21 multiple correlation analysis.R

@@ -1,0 +1,17 @@
+multiple correlation analysis
+
+mosthighlycorrelated <- function(mydataframe,numtoreport)
+{
+  # find the correlations
+  cormatrix <- cor(mydataframe)
+  # set the correlations on the diagonal or lower triangle to zero,
+  # so they will not be reported as the highest ones:
+  diag(cormatrix) <- 0
+  cormatrix[lower.tri(cormatrix)] <- 0
+  # flatten the matrix into a dataframe for easy sorting
+  fm <- as.data.frame(as.table(cormatrix))
+  # assign human-friendly names
+  names(fm) <- c("First.Variable", "Second.Variable","Correlation")
+  # sort and print the top n correlations
+  head(fm[order(abs(fm$Correlation),decreasing=T),],n=numtoreport)
+}

### 2018-10-22 MIMOSA.R
adding original ICS  mimosa scripts

@@ -1,0 +1,115 @@
+biocLite("devtools")
+library(devtools) #load it
+install_github("MIMOSA","RGLab",branch="master") #Need R 3.0.0, and a bunch of dependencies. The install will fail with various error messages until you install those dependencies.
+biocLite("MIMOSA")
+library(MIMOSA)
+
+DF<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/All Counts.csv")
+DF<-as.data.table(DF) 
+DF$Neg<-DF[,21]
+DF<-DF[DF$Stim!="PMA"]
+
+#All cytokine
+data<-DF[,c(2,5,26,22)]
+E<-ConstructMIMOSAExpressionSet(data,
+                                reference=Stim%in%'UN',
+                                measure.columns=c('Neg','Cyt.Count'),
+                                other.annotations=c('Stim','ID'),
+                                default.cast.formula=component~ID+Stim,
+                                .variables=.(ID))
+test <- MIMOSA(Neg+Cyt.Count~ID|Stim,
+               data=E)
+probabilities<-MIMOSA::getZ(test)
+counts<-MIMOSA::countsTable(test)
+fdr<-MIMOSA::fdr(test)
+MIMOSA.Results<-cbind(counts,probabilities,fdr)
+MIMOSA.summary<-MIMOSA::getW(test)
+write.csv(MIMOSA.Results,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/MIMOSA-Results.csv")
+
+#TH1 
+data1<-DF[,c(2,5,26,23)]
+E1<-ConstructMIMOSAExpressionSet(data1,
+                                reference=Stim%in%'UN',
+                                measure.columns=c('Neg','TH1.Count'),
+                                other.annotations=c('Stim','ID'),
+                                default.cast.formula=component~ID+Stim,
+                                .variables=.(ID))
+
+test1 <- MIMOSA(Neg+TH1.Count~ID|Stim,
+                data=E1)
+probabilities1<-MIMOSA::getZ(test1)
+counts1<-MIMOSA::countsTable(test1)
+fdr1<-MIMOSA::fdr(test1)
+MIMOSA.Results1<-cbind(counts1,probabilities1,fdr1)
+MIMOSA.summary1<-MIMOSA::getW(test1)
[…]

### 2018-11-19 ICS_CD4.Rmd
cleaning functions - subtract unstim

@@
 ---
+
+```{r}
 library(dplyr)
@@
          y="Frequency of Cytokine+ CD4+ T cells")
+```
 

### 2019-03-27 Cytokine P-values.Rmd
adding before I prune files

---
title: "Untitled"
author: "Taryn McLaughlin"
date: "11/30/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#P-value Functions

```{r}
# a function to evaluate differences in the expression of a marker (designated by the "column" argument) between disease groups
pvals_helminth<-function(data, column){

    x<-split(data,data$TB) #creates individual dataframes for each level of TB
    #does a wilcox test for the column of interest comparing SM- to SM+ within each level of TB
    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"], exact=FALSE))

    y<-split(data,data$SM) #creates two dataframes for SM- and SM+
    #each lapply does a wilcox test on subsets these dataframes to compare 2 levels of TB
    #stores the output in a variable. So B will contain the comparison of HC to LTBI for both SM- and SM+
    B<-lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "LTBI"), exact=FALSE))
    C<- lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("LTBI", "TB"), exact=FALSE))
    D<- lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "TB"), exact=FALSE))

    #To keep track of all the p-values... these should match the order below
    labels<-c("HC SM+ to HC SM-", "LTBI SM+ to LTBI SM-", "TB SM+ to TB SM-",
              "HC SM+ to LTBI SM+", "LTBI SM+ to TB SM+", "HC SM+ to TB SM+",
              "HC SM- to LTBI SM-", "LTBI SM- to TB SM-", "HC SM- to TB SM-"
    )

    #extract the p-values from all the stored wilcox tests
    pvals<-c(A$HC$p.value, A$LTBI$p.value, A$TB$p.value,
             B$SM$p.value, C$SM$p.value,  D$SM$p.value,
             B$X$p.value,  C$X$p.value,   D$X$p.value)

    #adjusts the p-values using the false discovery rate strategy
    fdr_adj_pvals<-round(p.adjust(pvals, method="fdr"), 4)

    #create a table with the output
    table<-(cbind(labels, round(pvals, 4), fdr_adj_pvals))
    colnames(table)<-c("Comparison", paste(column,"p-values"," "), 
                       "FDR Corrected")
    table #return the table
}
```

#Read in Data
```{r}
library(dplyr)
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
Cyt<-read.csv("CD4_Cytokine_clean.csv")
    data<-filter(Cyt, SM!="N")
    data$SM<-factor(data$SM, levels=c("X","SM"))
    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
Cyt<-data

MFI<-read.csv("CD4_MFI_clean.csv")
    data<-filter(MFI, SM!="N")
    data$SM<-factor(data$SM, levels=c("X","SM"))
    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
MFI<-data

Compass<-read.csv("ICS_Compass_scores_complete.csv")
    data<-filter(Compass, SM!="N")
    data$SM<-factor(data$SM, levels=c("X","SM"))
    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
Compass<-data
```

#Standard Analysis
This will literally p-value the frequencies of every cytokine subset as well as combination gate from my elaborate gating strategy.

##PMA
```{r}
library(dplyr)
library(knitr)
DF<-Cyt
colnames<-names(dplyr::select(DF, -Sample, -Donor, -Stim, -TB, -SM)[1:16])
colnames<-names(dplyr::select(DF, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){
    x<-subset(DF, DF$Stim=="PMA")
    y<-pvals_helminth(x, col)
    print(y)
}
```

##PEP
```{r}
library(dplyr)
library(knitr)
DF<-Cyt
colnames<-names(dplyr::select(DF, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){
    x<-subset(DF, DF$Stim=="PEP")
    y<-pvals_helminth(x, col)
    print(y)
}
```

##WCL
```{r}
library(dplyr)
library(knitr)
DF<-Cyt
colnames<-names(dplyr::select(DF, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){
    x<-subset(DF, DF$Stim=="WCL")
    y<-pvals_helminth(x, col)
    print(y)
}
```

##SEA
```{r}
library(dplyr)
library(knitr)
DF<-Cyt
colnames<-names(dplyr::select(DF, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){
    x<-subset(DF, DF$Stim=="SEA")
    y<-pvals_helminth(x, col)
    print(y)
}
```

##SWAP
```{r}
library(dplyr)
library(knitr)
DF<-Cyt
colnames<-names(dplyr::select(DF, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){
    x<-subset(DF, DF$Stim=="SWAP")
    y<-pvals_helminth(x, col)
    print(y)
}
```

#Compass Scores
The actually data processessing is done elsewhere because it's computationally intensive and is stored locally. This simply analyzes the output data.

```{r}
library(dplyr)
library(knitr)
#Since there are only two parameters of interest this is pretty straightforward, 
#we don't need elaborate loops
#we split the data by celltype (since GD and CD8 data is included) and stim
x<-split(Compass, list(Compass$celltype, Compass$Stim))
#and apply over the subsequent list of dataframes the pvals function to see if FS is different between disease groups
y<-lapply(x, function(g) pvals_helminth(g, "FS"))
y
#or if the PFS is different between disease groups
z<-lapply(x, function(g) pvals_helminth(g, "PFS"))
z
```

#Cytokine MFI
During the cleaning phase of this project (see ICS_Cleaning.Rmd), cytokine MFI values were set to NA if there were fewer than 20 cells contributing to that MFI value. As such, we can only make statistical comparisons between groups when the groups contain enough observed values. I have annotated this code for PMA only but the mechanism is the same for other stims.

##PMA
```{r}
library(dplyr)
library(knitr)
kable(colSums(!is.na(MFI)))
#since the extent of missingness varies by stim we have to do this each time
DF<-subset(MFI, MFI$Stim=="PMA")
#split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
x<-split(DF, list(DF$TB, DF$SM))
#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0. It's a weird double negative. But basically makes a list of the columns that have sufficient data.
y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
#each entry in the list has to be extracted and named
a<-unlist(y[1])
b<-unlist(y[2])
c<-unlist(y[3])
d<-unlist(y[5])
e<-unlist(y[6])
#then find the intersection of all the lists. Aka which column names have sufficient data in all 6 TB-SM groups
colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
colnames<-colnames[7:length(colnames)]
for(col in colnames){
    y<-pvals_helminth(DF, col)
    print(y)
}
```

##PEP
```{r}
library(dplyr)
library(knitr)
DF<-subset(MFI, MFI$Stim=="PEP")
x<-split(DF, list(DF$TB, DF$SM))
y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
a<-unlist(y[1])
b<-unlist(y[2])
c<-unlist(y[3])
d<-unlist(y[5])
e<-unlist(y[6])
colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
colnames<-colnames[7:length(colnames)]
for(col in colnames){
    y<-pvals_helminth(DF, col)
    print(y)
}
```

##WCL
```{r}
library(dplyr)
library(knitr)
DF<-subset(MFI, MFI$Stim=="WCL")
x<-split(DF, list(DF$TB, DF$SM))
y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
a<-unlist(y[1])
b<-unlist(y[2])
c<-unlist(y[3])
d<-unlist(y[5])
e<-unlist(y[6])
colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
colnames<-colnames[7:length(colnames)]
for(col in colnames){
    y<-pvals_helminth(DF, col)
    print(y)
}
```

##SEA
```{r}
library(dplyr)
library(knitr)
DF<-subset(MFI, MFI$Stim=="SEA")
x<-split(DF, list(DF$TB, DF$SM))
y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
a<-unlist(y[1])
b<-unlist(y[2])
c<-unlist(y[3])
d<-unlist(y[5])
e<-unlist(y[6])
colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
colnames<-colnames[7:length(colnames)]
for(col in colnames){
    y<-pvals_helminth(DF, col)
    print(y)
}
```

##SWAP
```{r}
library(dplyr)
library(knitr)
DF<-subset(MFI, MFI$Stim=="SWAP")
x<-split(DF, list(DF$TB, DF$SM))
y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
a<-unlist(y[1])
b<-unlist(y[2])
c<-unlist(y[3])
d<-unlist(y[5])
e<-unlist(y[6])
colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
colnames<-colnames[7:length(colnames)]
for(col in colnames){
    y<-pvals_helminth(DF, col)
    print(y)
}
```

### 2019-03-27 SUBUN.R

@@ -1,0 +1,60 @@
+# All of this generates a fake data table that contains a "key" variable (i.e. donor)
+# a condition variable (i.e. "stim") and a set of columns of output data
+# the goal is for every donor, in each column of output data subtract out the value in that column
+# that corresponds to a specific value in the condition column
+donor<-c(rep("john", 6), rep("mary", 6), rep("jack", 6), rep("suzy", 6))
+stim<-rep(c("un","w","p","s","sw","sb"),4)
+ifng<-rnorm(24, 2)
+tnfa<-rnorm(24,10)
+IL4<-rnorm(24,5)
+IL5<-rnorm(24,7)
+IL10<-rnorm(24,13)
+IL13<-rnorm(24,19)
+IL17<-rnorm(24,17)
+IL21<-rnorm(24,11)
+IL22<-rnorm(24,3)
+
+fake<-data.frame(cbind(donor,stim,ifng,tnfa,IL4, IL5, IL10, IL13, IL17, IL21, IL22))
+remove(donor,stim,ifng,tnfa,IL4, IL5, IL10, IL13, IL17, IL21, IL22)
+#can only use data frames to use bracketed string column indexing so make sure this is a data frame not a data table
+
+#when binding data together the output data becomes facgtors and we need them to be numeric
+class_change<-function(datatable){
+    y<-dplyr::select(datatable, donor, stim)
+    x<-dplyr::select(datatable, -donor, -stim)
+    for(i in 1:length(x)){x[[i]]<-as.numeric(as.character(x[[i]]))}
+    df<-cbind(y, x)
+}
+fake<-class_change(fake)
+#we want to copy the original data table to reference back to and check our math
+fakecopy<-fake
+
+#This is going to generate the list of columns names that we want to iterate over
+numeric.only <- function(X,...){
+    returnCols <- names(X)
+    a<-sapply(X, is.numeric)
+    print(returnCols[a == "TRUE"])
+}
+colnms<-numeric.only(fake)
+
+dt<-split(fake, fake[,"donor"]) #generate a list of donor specific mini data frames 
+
+#split --> lists in alphabetical order not by original key order
+#To solve this I arranged the original data frame in alphabetical order
+
+fake<-arrange(fake, donor)
+#can we instead arrange the split to match the original order
+
+#for each column name in the list of column names
+for(col in colnms){
+    #this is a check to make sure it goes through all the columns
+    print(col) 
+    #dt.r will make a list of length of dt where each item in the list is the new column for a particular donor
[…]

### 2019-03-27 remove-samples-from-key-table.R

@@ -1,0 +1,77 @@
+# All of this generates a fake data table that contains a "key" variable (i.e. donor)
+# a condition variable (i.e. "stim") and a set of columns of output data
+# the goal is for every donor, in each column of output data subtract out the value in that column
+# that corresponds to a specific value in the condition column
+donor<-c(rep("john", 6), rep("mary", 6), rep("jack", 6), rep("suzy", 6))
+stim<-rep(c("un","w","p","s","sw","sb"),4)
+Cyt<-sample(20:200, 24)
+TH1<-sample(20:200, 24)
+TH2<-sample(20:200, 24)
+TH1.2<-sample(20:200, 24)
+
+Cyt_ifng<-sample(100:1000, 24)
+Cyt_tnfa<-sample(100:1000, 24)
+Cyt_IL4<-sample(100:1000, 24)
+Cyt_IL13<-sample(100:1000, 24)
+TH1_ifng<-sample(100:1000, 24)
+TH1_tnfa<-sample(100:1000, 24)
+TH1_IL4<-sample(100:1000, 24)
+TH1_IL13<-sample(100:1000, 24)
+TH2_ifng<-sample(100:1000, 24)
+TH2_tnfa<-sample(100:1000, 24)
+TH2_IL4<-sample(100:1000, 24)
+TH2_IL13<-sample(100:1000, 24)
+TH1.2_ifng<-sample(100:1000, 24)
+TH1.2_tnfa<-sample(100:1000, 24)
+TH1.2_IL4<-sample(100:1000, 24)
+TH1.2_IL13<-sample(100:1000, 24)
+
+fake_count<-data.frame(cbind(donor,stim, Cyt, TH1, TH2, TH1.2))
+fake_mfi<-data.frame(cbind(donor,stim,Cyt_ifng,Cyt_tnfa,Cyt_IL4,Cyt_IL13,TH1_ifng,TH1_tnfa,
+                           TH1_IL4,TH1_IL13,TH2_ifng,TH2_tnfa,TH2_IL4,TH2_IL13,TH1.2_ifng,TH1.2_tnfa,
+                           TH1.2_IL4,TH1.2_IL13))
+remove(donor, stim, Cyt, TH1, TH2, TH1.2,Cyt_ifng,Cyt_tnfa,Cyt_IL4,Cyt_IL13,TH1_ifng,TH1_tnfa,
[…]

### 2019-03-27 subtract-background.R

@@ -1,0 +1,42 @@
+#used to generate a list of character strings that are the names of the columns containing numeric data
+numeric.only <- function(X,...){
+    returnCols <- names(X)
+    a<-sapply(X, is.numeric)
+    print(returnCols[a == "TRUE"])
+}
+
+#takes a dataframe (v. important its a dataframe), splitvar is a character string of a column name
+#colnms is a list of strings indicating which columns to subtract background from
+#condition is a character string with the key to indicate what is being subtracted
+background.subtract<-function(df, splitvar, keyvar, condition){
+    dt<-split(df, df[, splitvar]) #generate a list of mini data frames for each level of splitvar
+    #split --> lists in alphabetical order not by original key order
+    #To solve this I arranged the original data frame in alphabetical order
+    #can we instead arrange the split to match the original order
+    
+    colnms<-numeric.only(df)
+    
+    #for each column name in the list of column names
+    for(col in colnms){
+        #this is a check to make sure it goes through all the columns
+        print(col) 
+        #dt.r will make a list of length of dt where each item in the list is the new column for a particular donor
+        dt.r<-lapply(dt, function(g)
+            #take the values in each column and subtract out the value indicated by an "un" the stim column 
+            (g[,col] - subset(g, g[,keyvar]==condition)[,col]))
+        #unlist the list so you get a single column and insert it back into the original data frame
+        df[,col]<-unlist(dt.r)
+    }
+    df
+}
+
+filenames = c("CD4_Cytokine.csv", "CD8_Cytokine.csv", "GD_Cytokine.csv")
+l<-list(CD4_Cytokine.csv, CD8_Cytokine.csv, GD_Cytokine.csv)
+l<-lapply(l, function(x) arrange(x, Donor))
+l<-lapply(l, function(x) background.subtract(x, "Donor", "Stim", "UN"))
+for (i in 1:length(filenames)) {
+    assign(filenames[i], 
+           data.frame(l[i]))}
+CD4_Cytokine.csv[CD4_Cytokine.csv<0]<-0
+CD8_Cytokine.csv[CD8_Cytokine.csv<0]<-0
+GD_Cytokine.csv[GD_Cytokine.csv<0]<-0

### 2019-03-27 tryna function.R

@@ -1,0 +1,110 @@
+splitter<-function(data,column,splitter,one=NULL,two=NULL,three=NULL,four=NULL){
+        library(data.table)
+        
+        labels<-tstrsplit(data[,column], splitter)
+        data[,one]<-labels[[1]]
+        data[,two]<-labels[[2]]
+        data[,three]<-labels[[3]]
+        data[,four]<-labels[[4]]
+        return(data)
+}
+
+base.plot <- function(DF, group1, group2, value) {
+        p <- ggplot(DF, aes(x=group1, y=value, col=group1))
+        p <- p + theme_bw()
+        p <- p + theme(legend.position="bottom")
+        p <- p + geom_jitter(width=.1,height=0, shape=1,size=2)
+        p <- p + scale_color_brewer(palette="Set1", guide=guide_legend(ncol=6, title=NULL))
+        p <- p + xlab("") + ylab("")
+        return(p)
+}
+
+replace.btwn.tables2<-function(datatable1, datatable2, id.var, ref.var, new.var, new.vals){
+    library(dplyr)
+    inter <- intersect(datatable1[,id.var], datatable2[,ref.var])
+    tbl1a <- subset(datatable1, datatable1[,id.var] %in% inter)
+    tbl1b <- subset(datatable1, !(datatable1[,id.var] %in% inter))
+    tbl2  <- subset(datatable2, datatable2[,ref.var] %in% inter)
+    tbl1a[,new.var] <- replace(x = tbl1a[,new.var],
+                           list = (tbl1a[,id.var] %in% tbl2[,ref.var]),
+                           values = as.character(tbl2[,new.vals]))
+    new.table<-rbind(tbl1a, tbl1b)
+}
+
+pvals<-function(data, column){
+    
+    data<-filter(data, SM!="N")
+    data$SM<-factor(data$SM, levels=c("X","SM"))
+    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
+    
+    x<-split(data,data$TB) 
+    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"]))
+    
+    y<-split(data,data$SM) 
+    B<-lapply(y, function(g) wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "LTBI")))
[…]

### 2019-03-27 CD4 Cleaning.Rmd
most recent version

@@
 ---
-Load and rough clean all data sets
-```{r}
-setwd("/Applications/Old Computer/Day Lab/Data_Flow")
-CD4<-read.csv("CD4.csv")
-CD4<-filter(CD4,!(Sample=="Mean"))%>%
-filter(!Sample=="StdDev")%>%
-filter(!Sample=="Sample")
-CD4[CD4=="¥"]<-NA
-
-Cyt<-read.csv("CD4_Cytokine.csv")
-Cyt<-filter(Cyt,!(Sample=="Mean"))%>%
-filter(!Sample=="StdDev")%>%
-filter(!Sample=="Sample")
-Cyt[Cyt=="¥"]<-NA
-
-AG<-read.table("/Applications/Old Computer/Day Lab/Data_Flow/CD4_Positive.csv", header= TRUE, sep = ",")
-AG<-filter(AG,!(Sample=="Mean"))%>%
-filter(!Sample=="StdDev")%>%
-filter(!Sample=="Sample")
-AG[AG=="¥"]<-NA
-AG<-filter(AG,!(Sample=="Mean")) %>%
-  filter(!Sample=="StdDev") %>%
-  filter(!Sample=="Sample") %>%
-  filter(!Sample=="NK2126_HC_X_SEA_005.fcs") %>%
-  filter(!Sample=="NK2450_HC_SM_SEA_047.fcs") %>%
-  filter(!Sample=="NK2402_LTBI_SM_SWAP_024.fcs") %>%
-  filter(!Sample=="NK2342_TB_X_SWAP_006.fcs") %>%
-  filter(!Sample=="NK2136_TB_X_WCL_003.fcs") %>%
-  filter(!Sample=="NK2168_HC_SM_WCL_040.fcs") %>%
-  filter(!Sample=="NK2421_LTBI_SM_PEP_010.fcs") %>%
-  filter(!Sample=="NK2063_TB_X_SEA_005.fcs") %>%
-  filter(!Sample=="NK2115_LTBI_X_PEP_010.fcs") %>%
-  filter(!Sample=="NK2115_LTBI_X_WCL_009.fcs") %>%
-  filter(!Sample=="NK2171_HC_X_WCL_009.fcs") %>%
-  filter(!Sample=="NK2162_HC_X_SEA_023.fcs")
-
-CYT<-AG[,c(1:18)]
-TH1<-AG[,c(1,19:35)]
-TH2<-AG[,c(1,53:69)]
-TH1.2<-AG[,c(1,36:52)]
-
-write.csv(CD4,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/CD4.csv")
-write.csv(Cyt,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/Cytokine.csv")
-write.csv(CYT,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/Cyt.csv")
-write.csv(TH1,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH1.csv")
-write.csv(TH2,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH2.csv")
-write.csv(TH1.2,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH1-2.csv")
-
-remove(AG, CD4, Cyt, CYT, TH1, TH2, TH1.2)
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE, message = FALSE )
 ```
 
-PBMC
-```{r}
-PBMC<-read.csv("/Applications/Old computer/Day Lab/Data_Flow/PBMC.csv")
-PBMC<-filter(PBMC,!(Helminth.Status=="Ascaris+"))%>%
-        filter(!Helminth.Status=="Ascarus+ & Trichuris+")%>%
-        filter(!(Helminth.Status=="Hookworm+"))%>%
-        filter(!(Helminth.Status=="Trichuris+"))
-PBMC$TB.Status = factor(PBMC$TB.Status,levels=c("Healthy Control", "LTBI", "Active TB"))
+```{r functions}
+clean_flowjo_output<-function(datatable){
+    library(dplyr)
+    datatable$Sample<-gsub("1: |2: |3: |4: |5: |6: ", "", datatable$Sample)
+    datatable$Sample<-gsub("XUN", "X_UN", datatable$Sample)
+    datatable$Sample<-gsub("XPMA", "X_PMA", datatable$Sample)
+    datatable$Sample<-gsub("XWCL", "X_WCL", datatable$Sample)
+    datatable$Sample<-gsub("XPEP", "X_PEP", datatable$Sample)
+    datatable$Sample<-gsub("XSEA", "X_SEA", datatable$Sample)
+    datatable$Sample<-gsub("XSWAP", "X_SWAP", datatable$Sample)
+    datatable$Sample<-gsub("_ ", "_", datatable$Sample)
+    datatable$Sample<-gsub(" _", "_", datatable$Sample)
+    datatable<-filter(datatable,!(Sample=="Mean"))%>%
+        filter(!Sample=="StdDev")%>%
+        filter(!Sample=="Sample")%>%
+        filter(!Sample=="Sample")
+    datatable[datatable=="¥"]<-NA
+    datatable[datatable=="•"]<-NA
+    datatable
+
+}
+
+splitter<-function(data,column,splitter,one=NULL,two=NULL,three=NULL,four=NULL){
+    library(data.table)
+
+    labels<-tstrsplit(data[,column], splitter)
+    data[,one]<-labels[[1]]
+    data[,two]<-labels[[2]]
+    data[,three]<-labels[[3]]
+    data[,four]<-labels[[4]]
+    return(data)
+}
+#demo: HEL<-splitter(HEL, "StudyID", "-", one="StudyID", two="Visit", three="Iteration")
+
+remove_samples<-function(datatable){
+    library(dplyr)
+    excluded_samples<-c("NK2126_HC_X_SEA_005.fcs", 
+                        "NK2450_HC_SM_SEA_047.fcs" , 
+                        "NK2402_LTBI_SM_SWAP_024.fcs",
+                        "NK2342_TB_X_SWAP_006.fcs",
+                        "NK2136_TB_X_WCL_003.fcs",
+                        "NK2168_HC_SM_WCL_040.fcs",
+                        "NK2421_LTBI_SM_PEP_010.fcs",
+                        "NK2063_TB_X_SEA_005.fcs",
+                        "NK2115_LTBI_X_PEP_010.fcs",
+                        "NK2115_LTBI_X_WCL_009.fcs",
+                        "NK2171_HC_X_WCL_009.fcs",
+                        "NK2162_HC_X_SEA_023.fcs")
+    datatable<-filter(datatable, !(Sample %in% excluded_samples))
+    datatable
+}
+
+class_change<-function(datatable){
+    y<-dplyr::select(datatable, Sample, Donor, TB, SM, Stim)
+    x<-dplyr::select(datatable, -Sample, -Donor, -TB, -SM, -Stim)
+    for(i in 1:length(x)){x[[i]]<-as.numeric(as.character(x[[i]]))}
+    df<-cbind(y, x)
+}
+
+#used to generate a list of character strings that are the names of the columns containing numeric data
+numeric.only <- function(X,...){
+      returnCols <- names(X)
+      a<-sapply(X, is.numeric)
+      print(returnCols[a == "TRUE"])
+}
+
+#takes a dataframe (v. important its a dataframe), splitvar is a character string of a column name
+#colnms is a list of strings indicating which columns to subtract background from
+#condition is a character string with the key to indicate what is being subtracted
+background.subtract<-function(df, splitvar, keyvar, condition){
+    dt<-split(df, df[, splitvar]) #generate a list of mini data frames for each level of splitvar
+    #split --> lists in alphabetical order not by original key order
+    #To solve this I arranged the original data frame in alphabetical order
+    #can we instead arrange the split to match the original order
+
+    colnms<-numeric.only(df)
+
+    #for each column name in the list of column names
+    for(col in colnms){
+        #this is a check to make sure it goes through all the columns
+        print(col) 
+        #dt.r will make a list of length of dt where each item in the list is the new column for a particular donor
+        dt.r<-lapply(dt, function(g)
+        #take the values in each column and subtract out the value indicated by an "un" the stim column 
+        (g[,col] - subset(g, g[,keyvar]==condition)[,col]))
+        #unlist the list so you get a single column and insert it back into the original data frame
+        df[,col]<-unlist(dt.r)
+    }
+    df
+}
 ```
 
-cleaning total CD4 data
-```{r}
-## Names Labels Etc. for X and SM
-#remember to check for _ issues
-setwd("/Applications/Old Computer/Day Lab/Data_Flow/Clean")
-CD4<-read.csv("CD4.csv")
+```{r Donor Info}
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/")
+PBMC<-read.csv("ICS_PBMC.csv")
+PBMC<-dplyr::filter(PBMC, Helminth.Status=="Schisto+" | Helminth.Status== "Negative" | Helminth.Status== "Not US" | Helminth.Status=="US" | Helminth.Status== "Unknown")
+PBMC$TB.Status = factor(PBMC$TB.Status,levels=c("Naïve","BCG","Healthy Control", "LTBI", "Active TB", "Unknown"))
+PBMC$Viability <- gsub("%/%" , "", PBMC$Viability)
+PBMC$Exp.Date <- gsub("TM", "", PBMC$Exp.Date)
+PBMC<- filter(PBMC, !(Excluded=="EX"))%>%
+    dplyr::rename(TB = TB.Status, SM = Helminth.Status, StudyID = Study.ID)%>%
+    select(StudyID, TB, SM, Viability, Cells.Vial, Excluded)
 
-CD4<-filter(CD4,!(Sample=="Mean")) %>%
-  filter(!Sample=="StdDev") %>%
-  filter(!Sample=="Sample") %>%
-  filter(!Sample=="NK2126_HC_X_SEA_005.fcs") %>%
-  filter(!Sample=="NK2450_HC_SM_SEA_047.fcs") %>%
-  filter(!Sample=="NK2402_LTBI_SM_SWAP_024.fcs") %>%
-  filter(!Sample=="NK2342_TB_X_SWAP_006.fcs") %>%
-  filter(!Sample=="NK2136_TB_X_WCL_003.fcs") %>%
-  filter(!Sample=="NK2168_HC_SM_WCL_040.fcs") %>%
-  filter(!Sample=="NK2421_LTBI_SM_PEP_010.fcs") %>%
-  filter(!Sample=="NK2063_TB_X_SEA_005.fcs") %>%
-  filter(!Sample=="NK2115_LTBI_X_PEP_010.fcs") %>%
-  filter(!Sample=="NK2115_LTBI_X_WCL_009.fcs") %>%
-  filter(!Sample=="NK2171_HC_X_WCL_009.fcs") %>%
-  filter(!Sample=="NK2162_HC_X_SEA_023.fcs")
+QFT<- read.csv("/Applications/Old Computer/Epi Project/Data_clean/complete-kenya.csv")
+QFT<- filter(QFT, StudyID %in% PBMC$StudyID)%>%
+    select(StudyID, QFT, SchistosomaIntensity)
 
-Split<-strsplit(as.character(CD4$Sample),"_");
-CD4$Donor = as.factor(sapply(Split,"[",1));
-CD4$TB = as.factor(sapply(Split,"[",2));
-CD4$SM = as.factor(sapply(Split,"[",3));
-CD4$Antigen = as.factor(sapply(Split,"[",4));
-
-CD4<-as.data.table(CD4)
-CD4<-CD4[CD4$SM!="HEL"]
-
-CD4$SM<-factor(CD4$SM, levels=c("N","X","SM"))
-CD4$TB<-factor(CD4$TB, levels=c("N","HC", "LTBI", "TB"))
-CD4$Stim = factor(CD4$Antigen,levels=c("UN", "WCL","PEP","SEA", "SWAP", "PMA"))
-CD4<-CD4[!(is.na(CD4$Sample))]
-
-write.csv(CD4,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/CD4.csv")
-```
-cleaning cytokine frequency data
-```{r}
-Cyt<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/Cytokine.csv")
-
-Cyt<-filter(Cyt,!(Sample=="Mean")) %>%
-  filter(!Sample=="StdDev") %>%
-  filter(!Sample=="Sample") %>%
-  filter(!Sample=="NK2126_HC_X_SEA_005.fcs") %>%
-  filter(!Sample=="NK2450_HC_SM_SEA_047.fcs") %>%
-  filter(!Sample=="NK2402_LTBI_SM_SWAP_024.fcs") %>%
-  filter(!Sample=="NK2342_TB_X_SWAP_006.fcs") %>%
-  filter(!Sample=="NK2136_TB_X_WCL_003.fcs") %>%
-  filter(!Sample=="NK2168_HC_SM_WCL_040.fcs") %>%
-  filter(!Sample=="NK2421_LTBI_SM_PEP_010.fcs") %>%
-  filter(!Sample=="NK2063_TB_X_SEA_005.fcs") %>%
-  filter(!Sample=="NK2115_LTBI_X_PEP_010.fcs") %>%
-  filter(!Sample=="NK2115_LTBI_X_WCL_009.fcs") %>%
-  filter(!Sample=="NK2171_HC_X_WCL_009.fcs") %>%
-  filter(!Sample=="NK2162_HC_X_SEA_023.fcs")
-
-Split<-strsplit(as.character(Cyt$Sample),"_");
-Cyt$Donor = as.factor(sapply(Split,"[",1));
-Cyt$TB = as.factor(sapply(Split,"[",2));
-Cyt$SM = as.factor(sapply(Split,"[",3))
-Cyt$Antigen = as.factor(sapply(Split,"[",4));
-
-Cyt<-as.data.table(Cyt)
-Cyt<-Cyt[Cyt$SM!="HEL"]
-
-Cyt$SM<-factor(Cyt$SM, levels=c("N","X","SM"))
-Cyt$TB<-factor(Cyt$TB, levels=c("N","HC", "LTBI", "TB"))
-Cyt$Stim = factor(Cyt$Antigen,levels=c("UN", "WCL","PEP","SEA", "SWAP", "PMA"))
-
-Cyt<-Cyt[!(is.na(Cyt$Sample))]
-remove(Split)
-write.csv(Cyt,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/Cytokine.csv")
-
+DonorInfo<-merge(PBMC, QFT, by="StudyID")
 ```
 
-Determining antigen specific threshold vlaues
-```{r}
-UN<-filter(Cyt,Stim=="UN")
-ag<-filter(Cyt,Stim!="UN")
-ag[ag<0.03]<-NA
-Dat<-merge(ag,UN,by="Donor")
-Dat<-as.data.table(Dat)
-
-Dat$Sample<-Dat$Sample.x
-Dat$Stim<-Dat$Stim.x
-Dat$TB<-Dat$TB.x
-Dat$SM<-Dat$SM.x
-Dat$G_4_13_T<-(Dat$G_4_13_T.x-3*Dat$G_4_13_T.y)
-Dat$G_4_13_x<-(Dat$G_4_13_x.x-3*Dat$G_4_13_x.y)
-Dat$G_4_x_T<-(Dat$G_4_x_T.x-3*Dat$G_4_x_T.y)
-Dat$G_4_x_x<-(Dat$G_4_x_x.x-3*Dat$G_4_x_x.y)
-Dat$G_x_13_T<-(Dat$G_x_13_T.x-3*Dat$G_x_13_T.y)
-Dat$G_x_13_x<-(Dat$G_x_13_x.x-3*Dat$G_x_13_x.y)
-Dat$G_x_x_T<-(Dat$G_x_x_T.x-3*Dat$G_x_x_T.y)
-Dat$G_x_x_x<-(Dat$G_x_x_x.x-3*Dat$G_x_x_x.y)
-Dat$x_4_13_T<-(Dat$x_4_13_T.x-3*Dat$x_4_13_T.y)
-Dat$x_4_13_x<-(Dat$x_4_13_x.x-3*Dat$x_4_13_x.y)
-Dat$x_4_x_T<-(Dat$x_4_x_T.x-3*Dat$x_4_x_T.y)
-Dat$x_4_x_x<-(Dat$x_4_x_x.x-3*Dat$x_4_x_x.y)
-Dat$x_x_13_T<-(Dat$x_x_13_T.x-3*Dat$x_x_13_T.y)
-Dat$x_x_13_x<-(Dat$x_x_13_x.x-3*Dat$x_x_13_x.y)
-Dat$x_x_x_T<-(Dat$x_x_x_T.x-3*Dat$x_x_x_T.y)
-Dat$x_x_x_x<-Dat$x_x_x_x.x
-Dat$IFNg<-(Dat$IFNg.x-3*Dat$IFNg.y)
-Dat$IL13<-(Dat$IL13.x-3*Dat$IL13.y)
-Dat$IL4<-(Dat$IL4.x-3*Dat$IL4.y)
-Dat$TNFa<-(Dat$TNFa.x-3*Dat$TNFa.y)
-Dat$Cyt.Count<-(Dat$Cyt.Count.x-3*Dat$Cyt.Count.y)
-Dat$Cyt.Freq<-(Dat$Cyt.Freq.x-3*Dat$Cyt.Freq.y)
-Dat$TH1.Count<-(Dat$TH1.Count.x-3*Dat$TH1.Count.y)
-Dat$TH1.Freq<-(Dat$Th1.Freq.x-3*Dat$Th1.Freq.y)
-Dat$TH1.2.Count<-(Dat$TH1.2.Count.x-3*Dat$TH1.2.Count.y)
-Dat$TH1.2.Freq<-(Dat$TH1.2.Freq.x-3*Dat$TH1.2.Freq.y)
-Dat$TH2.Count<-(Dat$TH2.Count.x-3*Dat$TH2.Count.y)
-Dat$TH2.Freq<-(Dat$TH2.Freq.x-3*Dat$TH2.Freq.y)
-
-AG<-Dat[,c(1,70:88,90:93,95,97,99,101)]
-
-AG[AG <0 ]<-as.numeric(0)
-AG[AG >0 ]<-as.numeric(1)
-
-AG$Sum<-rowSums(AG[,6:28], na.rm=TRUE)
-AG$Cyt.Sum<-rowSums(AG[,21:24], na.rm = TRUE)
-AG$TH1.Sum<-rowSums(AG[,c(12,13,20,21,24,26)], na.rm = TRUE)
-AG$TH2.Sum<-rowSums(AG[,c(15,17,19,22,23,28)], na.rm = TRUE)
-AG$TH1.2.Sum<-rowSums(AG[,c(6:11,14,16,18,27)], na.rm = TRUE)
-
-x <- split(AG, AG$Stim)
-
-c<-lapply(x, function(g) colSums(g[,6:28], na.rm = TRUE))
-c2<-as.data.table(c)
-c3<-cbind(names(AG))
-cyt.tally<-cbind(c3[6:28],c2)
-
-r<-lapply(x, function(g) rowSums(g[,6:28], na.rm = TRUE))
-r2<-as.data.table(r)
-r3<-unique(cbind(as.character(AG$Donor)))
-donor.tally<-cbind(r3,r2)
-
-remove(UN, ag, Dat, c, c2, c3, r, r2, r3, x)
-write.csv(cyt.tally,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/cyt-tally.csv")
-write.csv(donor.tally,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/donor-tally.csv")
-write.csv(AG,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/threshold-data.csv")
-
+```{r read in all flowjo data tables}
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/exported/")
+filenames =list.files(pattern="*.csv")
+for (i in 1:length(filenames)) 
+    assign(filenames[i], read.csv(filenames[i]))
 ```
 
-Background subtracting cytokine frequencies
-```{r}
-UN<-filter(Cyt,Stim=="UN")
-ag<-filter(Cyt,Stim!="UN")
-Dat<-merge(ag,UN,by="Donor")
-Dat<-as.data.table(Dat)
-
-Dat$Sample<-Dat$Sample.x
-Dat$Stim<-Dat$Stim.x
-Dat$TB<-Dat$TB.x
-Dat$SM<-Dat$SM.x
-Dat$G_4_13_T<-(Dat$G_4_13_T.x-Dat$G_4_13_T.y)
-Dat$G_4_13_x<-(Dat$G_4_13_x.x-Dat$G_4_13_x.y)
-Dat$G_4_x_T<-(Dat$G_4_x_T.x-Dat$G_4_x_T.y)
-Dat$G_4_x_x<-(Dat$G_4_x_x.x-Dat$G_4_x_x.y)
-Dat$G_x_13_T<-(Dat$G_x_13_T.x-Dat$G_x_13_T.y)
-Dat$G_x_13_x<-(Dat$G_x_13_x.x-Dat$G_x_13_x.y)
-Dat$G_x_x_T<-(Dat$G_x_x_T.x-Dat$G_x_x_T.y)
-Dat$G_x_x_x<-(Dat$G_x_x_x.x-Dat$G_x_x_x.y)
-Dat$x_4_13_T<-(Dat$x_4_13_T.x-Dat$x_4_13_T.y)
-Dat$x_4_13_x<-(Dat$x_4_13_x.x-Dat$x_4_13_x.y)
-Dat$x_4_x_T<-(Dat$x_4_x_T.x-Dat$x_4_x_T.y)
-Dat$x_4_x_x<-(Dat$x_4_x_x.x-Dat$x_4_x_x.y)
-Dat$x_x_13_T<-(Dat$x_x_13_T.x-Dat$x_x_13_T.y)
-Dat$x_x_13_x<-(Dat$x_x_13_x.x-Dat$x_x_13_x.y)
-Dat$x_x_x_T<-(Dat$x_x_x_T.x-Dat$x_x_x_T.y)
-Dat$x_x_x_x<-Dat$x_x_x_x.x
-Dat$IFNg<-(Dat$IFNg.x-Dat$IFNg.y)
-Dat$IL13<-(Dat$IL13.x-Dat$IL13.y)
-Dat$IL4<-(Dat$IL4.x-Dat$IL4.y)
-Dat$TNFa<-(Dat$TNFa.x-Dat$TNFa.y)
-Dat$Cyt.Count<-(Dat$Cyt.Count.x-Dat$Cyt.Count.y)
-Dat$Cyt.Freq<-(Dat$Cyt.Freq.x-Dat$Cyt.Freq.y)
-Dat$TH1.Count<-(Dat$TH1.Count.x-Dat$TH1.Count.y)
-Dat$TH1.Freq<-(Dat$Th1.Freq.x-Dat$Th1.Freq.y)
-Dat$TH1.2.Count<-(Dat$TH1.2.Count.x-Dat$TH1.2.Count.y)
-Dat$TH1.2.Freq<-(Dat$TH1.2.Freq.x-Dat$TH1.2.Freq.y)
-Dat$TH2.Count<-(Dat$TH2.Count.x-Dat$TH2.Count.y)
-Dat$TH2.Freq<-(Dat$TH2.Freq.x-Dat$TH2.Freq.y)
-
-CYT<-Dat[,c(1,70:101)]
-CYT[CYT<0]<-0
-remove(UN, ag, Dat, Cyt)
-
-write.csv(CYT,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/Background Subtracted Cytokine.csv")
+```{r, reformat all data tables}
+l<-list(CD4_Counts.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_Positive.csv, CD4.csv, 
+        CD8_Counts.csv, CD8_Cytokine.csv, GD_Counts.csv, GD_Cytokine.csv)
+l<-lapply(l, function(x) clean_flowjo_output(x))
+l<-lapply(l, function(x) remove_samples(x))
+l<-lapply(l, function(x) splitter(x, "Sample", "_", one="Donor", two="TB", three="SM", four="Stim"))
+l<-lapply(l, function(x) class_change(x))
+for (i in 1:length(filenames)) {
+    assign(filenames[i], data.frame(l[i]))}
 ```
 
-#Antigen specific responses using threshold values
-```{r}
-Cyt<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/Cyt.csv")
-TH1<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH1.csv")
-TH2<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH2.csv")
-TH1.2<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH1-2.csv")
-
-TH1<-merge(AG,TH1,by="Sample", all=TRUE)
-TH2<-merge(AG,TH2,by="Sample", all=TRUE)
-TH1.2<-merge(AG,TH1.2,by="Sample", all=TRUE)
-Cyt<-merge(AG,Cyt,by="Sample", all=TRUE)
-
-Cyt<-Cyt[,c(1,3:5,21:25,29,30,35:51)]
-TH1<-TH1[,c(1:5,21,24,26,31,35:51)]
-TH2<-TH2[,c(1:5,22:23,28,32,35:51)]
-TH1.2<-TH1.2[,c(1:5,21:24,27,33,35:51)]
-
-write.csv(Cyt,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/Cyt_pos.csv")
-write.csv(TH1,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH1_pos.csv")
-write.csv(TH2,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH2_pos.csv")
-write.csv(TH1.2,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/TH1-2_pos.csv")
+```{r change colnames}
+newnames<-names(CD8_Cytokine.csv)
+names(CD4_Counts.csv)<-newnames[1:21]
+names(CD8_Counts.csv)<-newnames
+names(GD_Counts.csv)<-newnames
+names(GD_Cytokine.csv)<-newnames
 ```
 
-Load and rough clean all counts sets
-```{r}
-setwd("/Applications/Old Computer/Day Lab/Data_Flow")
-CD4_count<-read.csv("CD4_counts.csv")
-CD4_count<-filter(CD4_count,!(Sample=="Mean"))%>%
-    filter(!Sample=="StdDev")%>%
-    filter(!Sample=="Sample")%>%
-    filter(!Sample=="NK2126_HC_X_SEA_005.fcs") %>%
-    filter(!Sample=="NK2450_HC_SM_SEA_047.fcs") %>%
-  filter(!Sample=="NK2402_LTBI_SM_SWAP_024.fcs") %>%
-  filter(!Sample=="NK2342_TB_X_SWAP_006.fcs") %>%
-  filter(!Sample=="NK2136_TB_X_WCL_003.fcs") %>%
-  filter(!Sample=="NK2168_HC_SM_WCL_040.fcs") %>%
-  filter(!Sample=="NK2421_LTBI_SM_PEP_010.fcs") %>%
-  filter(!Sample=="NK2063_TB_X_SEA_005.fcs") %>%
-  filter(!Sample=="NK2115_LTBI_X_PEP_010.fcs") %>%
-  filter(!Sample=="NK2115_LTBI_X_WCL_009.fcs") %>%
-  filter(!Sample=="NK2171_HC_X_WCL_009.fcs") %>%
-  filter(!Sample=="NK2162_HC_X_SEA_023.fcs")
-CD4_count[CD4_count=="¥"]<-NA
+```{r subtract unstimulated from cyt freq tables}
+filenames = c("CD4_Cytokine.csv", "CD8_Cytokine.csv", "GD_Cytokine.csv")
+l<-list(CD4_Cytokine.csv, CD8_Cytokine.csv, GD_Cytokine.csv)
+l<-lapply(l, function(x) arrange(x, Donor))
+l<-lapply(l, function(x) background.subtract(x, "Donor", "Stim", "UN"))
+for (i in 1:length(filenames)) {
+    assign(filenames[i], 
+           data.frame(l[i]))}
+CD4_Cytokine.csv[CD4_Cytokine.csv<0]<-0
+CD8_Cytokine.csv[CD8_Cytokine.csv<0]<-0
+GD_Cytokine.csv[GD_Cytokine.csv<0]<-0
+```
 
-Split<-strsplit(as.character(CD4_count$Sample),"_");
-CD4_count$ID = as.factor(sapply(Split,"[",1));
-CD4_count$TB = as.factor(sapply(Split,"[",2));
-CD4_count$SM = as.factor(sapply(Split,"[",3));
-CD4_count$Antigen = as.factor(sapply(Split,"[",4));
-CD4_count$cell<-"CD4"
-
-CD4_count<-as.data.table(CD4_count)
-CD4_count<-CD4_count[CD4_count$SM!="HEL"]
-
-CD4_count$SM<-factor(CD4_count$SM, levels=c("N","X","SM"))
-CD4_count$TB<-factor(CD4_count$TB, levels=c("N","HC", "LTBI", "TB"))
-CD4_count$Stim = factor(CD4_count$Antigen,levels=c("UN", "WCL","PEP","SEA", "SWAP", "PMA"))
-CD4_count<-CD4_count[!(is.na(CD4_count$Sample))]
-
-write.csv(CD4_count,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/CD4_count.csv")
-
-CD8<-read.csv("CD8_counts.csv")
-CD8<-filter(CD8,!(Sample=="Mean"))%>%
-filter(!Sample=="StdDev")%>%
-filter(!Sample=="Sample")%>%
-    filter(!Sample=="NK2126_HC_X_SEA_005.fcs") %>%
-  filter(!Sample=="NK2450_HC_SM_SEA_047.fcs") %>%
-  filter(!Sample=="NK2402_LTBI_SM_SWAP_024.fcs") %>%
-  filter(!Sample=="NK2342_TB_X_SWAP_006.fcs") %>%
-  filter(!Sample=="NK2136_TB_X_WCL_003.fcs") %>%
-  filter(!Sample=="NK2168_HC_SM_WCL_040.fcs") %>%
-  filter(!Sample=="NK2421_LTBI_SM_PEP_010.fcs") %>%
-  filter(!Sample=="NK2063_TB_X_SEA_005.fcs") %>%
-  filter(!Sample=="NK2115_LTBI_X_PEP_010.fcs") %>%
-  filter(!Sample=="NK2115_LTBI_X_WCL_009.fcs") %>%
-  filter(!Sample=="NK2171_HC_X_WCL_009.fcs") %>%
-  filter(!Sample=="NK2162_HC_X_SEA_023.fcs")
-CD8[CD8=="¥"]<-NA
-
-Split<-strsplit(as.character(CD8$Sample),"_");
-CD8$ID = as.factor(sapply(Split,"[",1));
-CD8$TB = as.factor(sapply(Split,"[",2));
-CD8$SM = as.factor(sapply(Split,"[",3));
-CD8$Antigen = as.factor(sapply(Split,"[",4));
-CD8$cell<-"CD8"
-
-CD8<-as.data.table(CD8)
-CD8<-CD8[CD8$SM!="HEL"]
-
-CD8$SM<-factor(CD8$SM, levels=c("N","X","SM"))
-CD8$TB<-factor(CD8$TB, levels=c("N","HC", "LTBI", "TB"))
-CD8$Stim = factor(CD8$Antigen,levels=c("UN", "WCL","PEP","SEA", "SWAP", "PMA"))
-CD8<-CD8[!(is.na(CD8$Sample))]
-
-write.csv(CD8,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/CD8_count.csv")
-
-GD<-read.csv("GD_counts.csv")
-GD<-filter(GD,!(Sample=="Mean"))%>%
-filter(!Sample=="StdDev")%>%
-filter(!Sample=="Sample")%>%
-    filter(!Sample=="NK2126_HC_X_SEA_005.fcs") %>%
-  filter(!Sample=="NK2450_HC_SM_SEA_047.fcs") %>%
-  filter(!Sample=="NK2402_LTBI_SM_SWAP_024.fcs") %>%
-  filter(!Sample=="NK2342_TB_X_SWAP_006.fcs") %>%
-  filter(!Sample=="NK2136_TB_X_WCL_003.fcs") %>%
-  filter(!Sample=="NK2168_HC_SM_WCL_040.fcs") %>%
-  filter(!Sample=="NK2421_LTBI_SM_PEP_010.fcs") %>%
-  filter(!Sample=="NK2063_TB_X_SEA_005.fcs") %>%
-  filter(!Sample=="NK2115_LTBI_X_PEP_010.fcs") %>%
-  filter(!Sample=="NK2115_LTBI_X_WCL_009.fcs") %>%
-  filter(!Sample=="NK2171_HC_X_WCL_009.fcs") %>%
-  filter(!Sample=="NK2162_HC_X_SEA_023.fcs")
-GD[GD=="¥"]<-NA
-
-Split<-strsplit(as.character(GD$Sample),"_");
-GD$ID = as.factor(sapply(Split,"[",1));
-GD$TB = as.factor(sapply(Split,"[",2));
-GD$SM = as.factor(sapply(Split,"[",3));
-GD$Antigen = as.factor(sapply(Split,"[",4));
-GD$cell<-"GD"
-
-GD<-as.data.table(GD)
-GD<-GD[GD$SM!="HEL"]
-
-GD$SM<-factor(GD$SM, levels=c("N","X","SM"))
-GD$TB<-factor(GD$TB, levels=c("N","HC", "LTBI", "TB"))
-GD$Stim = factor(GD$Antigen,levels=c("UN", "WCL","PEP","SEA", "SWAP", "PMA"))
-GD<-GD[!(is.na(GD$Sample))]
-
-write.csv(GD,"/Applications/Old Computer/Day Lab/Data_Flow/Clean/GD_count.csv")
-
+```{r write files to "clean" directory}
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
+df.list<-list(CD4_Counts.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_Positive.csv,
+              CD4.csv, CD8_Counts.csv, CD8_Cytokine.csv, GD_Counts.csv,
+              GD_Cytokine.csv)
+dfnames<-list("CD4_Counts", "CD4_Cytokine", "CD4_MFI", "CD4_Positive",
+              "CD4", "CD8_Counts", "CD8_Cytokine", "GD_Counts",
+              "GD_Cytokine")
+newfilenames<-paste(dfnames, "clean", date(), ".csv", sep="_")
+#write.csv(df.list[1], file= newfilenames[1])
+write.csv(CD4_Counts.csv, file = "CD4_Counts.csv")
+write.csv(CD4_Cytokine.csv, file = "CD4_Cytokine.csv")
+write.csv(CD4_MFI.csv, file = "CD4_MFI.csv")
+write.csv(CD4_Positive.csv, file = "CD4_Positive.csv")
+write.csv(CD4.csv, file = "CD4.csv")
+write.csv(CD8_Counts.csv, file = "CD8_Counts.csv")
+write.csv(CD8_Cytokine.csv, file = "CD8_Cytokine.csv")
+write.csv(GD_Counts.csv, file = "GD_Counts.csv")
+write.csv(GD_Cytokine.csv, file = "GD_Cytokine.csv")
+write.csv(DonorInfo, file = "Donor_Info.csv")
 ```

### 2019-03-27 Cytokine_Analysis.Rmd
most recent version

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
 knitr::opts_chunk$set(warning = FALSE)
@@
 
+source("/Users/tarynam/Desktop/tryna function.R")
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
+data<-read.csv("CD4_Cytokine_clean.csv")
+    data$TB<-factor(data$TB, levels=c("N","HC","LTBI","TB"))
+    data$SM<-factor(data$SM, levels=c("N","X","SM"))
+CD4<-data
+
+data<-read.csv("CD8_Cytokine_clean.csv")
+    data$TB<-factor(data$TB, levels=c("N","HC","LTBI","TB"))
+    data$SM<-factor(data$SM, levels=c("N","X","SM"))
+CD8<-data
+
+data<-read.csv("CD4_MFI_clean.csv")
+    data$TB<-factor(data$TB, levels=c("N","HC","LTBI","TB"))
+    data$SM<-factor(data$SM, levels=c("N","X","SM"))
+MFI<-data
+
+data<-read.csv("ICS_Compass_scores_complete.csv")
+    data$TB<-factor(data$TB, levels=c("N","HC","LTBI","TB"))
+    data$SM<-factor(data$SM, levels=c("N","X","SM"))
+Compass<-data
+
+data<-read.csv("GD_Cytokine_clean.csv")
+    data$TB<-factor(data$TB, levels=c("N","HC","LTBI","TB"))
+    data$SM<-factor(data$SM, levels=c("N","X","SM"))
+GD<-data
+
+data<-read.csv("ICS_Donor_Info_clean.csv")
+data$TB<-gsub("BCG|Unknown", "Naïve", data$TB)
+data$TB<-factor(data$TB, levels=c("Naïve","Healthy Control","LTBI","Active TB"))
+data$SM<-gsub("US|Not US|Unknown", "Naïve", data$SM)
+data$SM<-factor(data$SM, levels=c("Naïve","Negative","Schisto+"))
+Donor<-data
+```
+
+```{r narrow down Yerkes Donors}
 library(dplyr)
-library(reshape2)
-Cytokine<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/Background Subtracted Cytokine.csv")
-Cytokine <- mutate(Cytokine, G_plus = G_4_13_x + G_4_x_x + G_x_13_x + G_x_x_x) %>%
-    mutate(T_plus = x_4_13_T + x_4_x_T + x_x_13_T + x_x_x_T) %>%
-    mutate(G_T_plus = G_4_x_T + G_x_13_T + G_x_x_T)%>%
-    mutate(IL4_plus = G_4_x_T + G_4_x_x + x_4_x_T + x_4_x_x) %>%
-    mutate(IL13_plus = G_x_13_T + G_x_13_x + x_x_13_T + x_x_13_x) %>%
-    mutate(IL4_IL13_plus = G_4_13_x + x_4_13_T + x_4_13_x)
-Cytokine$TB<-factor(Cytokine$TB, levels=c("N","HC","LTBI","TB"))
-Cytokine$SM<-factor(Cytokine$SM, levels=c("N","X","SM"))
+mimosa<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/CD4_MIMOSA_Complete.csv")
+HD<-filter(mimosa, grepl("HD", mimosa$Sample))%>%
+    filter(Stim!="PMA")
+HDscores<-aggregate(score~Donor, HD, FUN=sum)
+HDkeep<-head(arrange(HDscores, score), 10)
+#382 479 487 are BCG
+HDremove<-HD$Donor[!(HD$Donor %in% HDkeep$Donor)]
+CD4<-dplyr::filter(CD4, !(Donor %in% HDremove))
+MFI<-dplyr::filter(MFI, !(Donor %in% HDremove))
+Compass<-dplyr::filter(Compass, !(Donor %in% HDremove))
+Donor<-dplyr::filter(Donor, !(Donor %in% HDremove))
+```
+
+```{r viability and cell counts}
+Donor$TB[grepl("HD", Donor$StudyID)==TRUE]<-"Yerkes"
+palette.four<-c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b")
+ggplot(Donor, aes(x=SM, y=Viability, fill=TB, alpha=SM)) +
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
+    geom_jitter(width=.1,height=0, shape=16,size=3)+
+    scale_alpha_manual(values=c(0.5,0.5,1))+
+    scale_fill_manual(values = c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
+    theme_bw()+ theme(legend.position="none")+
+    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
+    facet_grid(~TB, scales="free")+
+    labs(title="Viability", 
+         subtitle="",
+         caption="",
+         x="TB and S. mansoni Status",
+         y="Live Cells (%)")
+
+ggplot(Donor, aes(x=SM, y=Cells.Vial, fill=TB, alpha=SM)) +
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
+    geom_jitter(width=.1,height=0, shape=16,size=3)+
+    scale_alpha_manual(values=c(0.5,0.5,1))+
+    scale_fill_manual(values = c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
+    theme_bw()+ theme(legend.position="none")+
+    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
+    facet_grid(~TB, scales="free")+
+    labs(title="Cell Counts", 
+         subtitle="",
+         caption="",
+         x="TB and S. mansoni Status",
+         y="Cells/mL)")
 ```
@@
 #Statistical Analysis
-I wrote a function called pvals which evaluates the relevant two sample two-tailed  p-values using the non-parametric Mann Whitney test (the command is wilcox.test in R).
-The function then corrects for multiple comparisons in two ways.
-The first uses  a Bonferonni correction to control the Family Wise Error Rate.
-The second controls the False Discovery Rate, the expected proportion of false discoveries amongst the rejected hypotheses.
-The false discovery rate is a less stringent condition than the family-wise error rate, so these methods are more powerful than the others.
-The function is below if you would like to see the details.
+I wrote a function called pvals_helminth which evaluates the relevant two sample two-tailed  p-values using the non-parametric Mann Whitney test (the command is wilcox.test in R).
+The function then corrects for multiple comparisons by controlling the False Discovery Rate, the expected proportion of false discoveries amongst the rejected hypotheses.
 
 ```{r p-value function, echo=TRUE}
-pvals<-function(data, column){
+# a function to evaluate differences in the expression of a marker (designated by the "column" argument) between disease groups
+pvals_helminth<-function(data, column){
 
-    data<-filter(data, SM!="N")
-    data$SM<-factor(data$SM, levels=c("X","SM"))
-    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
+    x<-split(data,data$TB) #creates individual dataframes for each level of TB
+    #does a wilcox test for the column of interest comparing SM- to SM+ within each level of TB
+    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"], exact=FALSE))
 
-    x<-split(data,data$TB) 
-    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"]))
+    y<-split(data,data$SM) #creates two dataframes for SM- and SM+
+    #each lapply does a wilcox test on subsets these dataframes to compare 2 levels of TB
+    #stores the output in a variable. So B will contain the comparison of HC to LTBI for both SM- and SM+
+    B<-lapply(y, function(g) 
+        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "LTBI"), exact=FALSE))
+    C<- lapply(y, function(g) 
+        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("LTBI", "TB"), exact=FALSE))
+    D<- lapply(y, function(g) 
+        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "TB"), exact=FALSE))
 
-    y<-split(data,data$SM) 
-    B<-lapply(y, function(g) 
-        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "LTBI")))
-    C<- lapply(y, function(g) 
-        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("LTBI", "TB")))
-    D<- lapply(y, function(g) 
-        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "TB")))
-
+    #To keep track of all the p-values... these should match the order below
     labels<-c("HC SM+ to HC SM-", "LTBI SM+ to LTBI SM-", "TB SM+ to TB SM-",
@@
     )
+
+    #extract the p-values from all the stored wilcox tests
     pvals<-c(A$HC$p.value, A$LTBI$p.value, A$TB$p.value,
@@
 
-    bonf_adj_pvals<-round(p.adjust(pvals, method="bonferroni"), 4)
-
+    #adjusts the p-values using the false discovery rate strategy
     fdr_adj_pvals<-round(p.adjust(pvals, method="fdr"), 4)
 
-    table<-(cbind(labels, round(pvals, 4), bonf_adj_pvals, fdr_adj_pvals))
+    #create a table with the output
+    table<-(cbind(labels, round(pvals, 4), fdr_adj_pvals))
     colnames(table)<-c("Comparison", paste(column,"p-values"," "), 
-                       "Bonferonni Corrected p-values", "FDR Corrected p-vales")
-    table
+                       "FDR Corrected")
+    table #return the table
 }
+```
+
+#Compass Graphs
+```{r}
+library(dplyr)
+DF<-dplyr::filter(Compass, TB != "N" & Stim == "WCL" & celltype =="CD4")
+ggplot(DF, aes(x=SM, y=FS, fill=TB, alpha=SM)) +
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
+    geom_jitter(width=.1,height=0, shape=16,size=4)+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_fill_manual(values = c("#1a9850" , "#2166ac", "#b2182b"))+
+    theme_bw()+ theme(legend.position="none")+
+    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
+      theme(plot.title = element_text(hjust = 0.5)) +
+    scale_x_discrete(labels=c("SM-", "SM+"))+
+    facet_grid(~TB, scales="free")+
+    labs(title="WCL-Specific CD4+ T cell Response", 
+         subtitle="",
+         caption="",
+         x="TB and S. mansoni Status",
+         y="Functionality Score")
+ggplot(DF, aes(x=SM, y=PFS, fill=TB, alpha=SM)) +
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
+    geom_jitter(width=.1,height=0, shape=16,size=3)+
+    scale_alpha_manual(values=c(0.5,0.5,1))+
+    scale_fill_manual(values = c("#1a9850" , "#2166ac", "#b2182b"))+
+    theme_bw()+ theme(legend.position="none")+
+    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
+    facet_grid(~TB, scales="free")+
+    labs(title="WCL-Specific CD4+ T cell Response", 
+         subtitle="",
+         caption="",
+         x="TB and S. mansoni Status",
+         y="Polyfunctionality Score")
+
+DF2<-merge(DF, dplyr::select(Donor, Donor, QFT))
+ggplot(DF2, aes(x=QFT, y=FS, col=SM))+geom_point()+facet_grid(~TB, scales = "free")+geom_smooth(method="lm")
 ```
@@
 I did not include all the individual data points on any of the graphs because it made it incredibly cluttered looking, however I did allow outliers to display.
-
 ```{r Boolean Graphs, cache=TRUE, echo=FALSE, warning=FALSE}
 library(dplyr)
+library(tidyr)
 library(reshape2)
 library(ggplot2)
-Boolean<-select(Cytokine, Donor, TB, SM, Stim, G_4_13_T, G_4_x_T, G_x_13_T, G_4_13_x, x_4_13_T, G_4_x_x, G_x_13_x,
-                x_4_x_T, x_x_13_T, G_x_x_T , G_x_x_x, x_x_x_T,  x_4_13_x, x_4_x_x, x_x_13_x)
 
-melt<-melt(Boolean,id.vars=c("TB","SM", "Stim"))
-melt$value<-as.numeric(melt$value)
-melt<-filter(melt, !(is.na(value)))
-melt<-filter(melt, !(value<0))
-melt$label = factor(melt$variable,levels=c("G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", 
-                                           "G_x_13_x","x_4_x_T","x_x_13_T", "G_x_x_T" , "G_x_x_x","x_x_x_T",  
-                                           "x_4_13_x","x_4_x_x", "x_x_13_x"))
-melt<-filter(melt, !is.na(label))
+Boolean<-dplyr::select(CD4, Donor, TB, SM, Stim, G_4_13_T, G_4_x_T, G_x_13_T, G_4_13_x, x_4_13_T, G_4_x_x, G_x_13_x, x_4_x_T, x_x_13_T, G_x_x_T , G_x_x_x, x_x_x_T,  x_4_13_x, x_4_x_x, x_x_13_x)%>%
+    dplyr::filter(TB!="N")
+
+melt<-melt(Boolean,id.vars=c("Donor","TB","SM", "Stim"))
 
 #all the boolean plots
-g<-ggplot(melt, aes(x=label, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
+ggplot(melt, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot(size=1, position=position_dodge(width = 1)) + 
     scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
@@
 #subset by stim
-data<-filter(melt, Stim=="PMA")
-g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
+data<-filter(melt, Stim=="SWAP")
+ggplot(data, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot(size=1, position=position_dodge(width = 1)) + 
+    scale_fill_manual(values =c("#e0e0e0", "#009E73" , "#2166ac", "#b2182b"))+
     scale_alpha_manual(values=c(0.5,0.5,1))+
     theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    facet_grid(TB~., scale="free")+
-    labs(title="PMA Cytokine Frequencies", 
+    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
+    facet_grid(TB~., scale="fixed")+
+    labs(title="SWAP-Specific Cytokine Frequencies", 
          subtitle="",
@@
 
-data<-filter(melt, Stim=="WCL")
-g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
+#3 separate booleans for TH1/2 TH1 and TH2
+#TH1-2
+data<-filter(melt, Stim=="SWAP")%>%
+    filter(variable %in% c("G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T"))
+ggplot(data, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot(size=1, position=position_dodge(width = 1))+
+    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
+    scale_alpha_manual(values=c(0.5,1))+
+    theme_bw()+ theme(legend.position="none")+
+    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
+    theme(plot.title = element_text(hjust = 0.5)) +
+    facet_grid(TB~., scale="fixed")+
+    coord_cartesian(ylim = c(0,.005))
+
+data<-filter(melt, Stim=="SWAP")%>%
+    filter(variable %in% c("G_x_x_T" , "G_x_x_x","x_x_x_T"))
+g<-ggplot(data, aes(x=variable, y=value, fill=TB, alpha=SM))
 g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
+    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
+    scale_alpha_manual(values=c(0.5,1))+
     theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    facet_grid(TB~., scale="free")+
-    labs(title="WCL Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="Cytokine Subsets",
-         y="Frequency of Cytokine+ CD4+ T cells (%)");data<-filter(melt, Stim=="PEP")
+    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
+    theme(plot.title = element_text(hjust = 0.5)) +
+    facet_grid(TB~., scale="fixed")+
+    coord_cartesian(ylim = c(0,.2))
 
-data<-filter(melt, Stim=="PEP")
-g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
+data<-filter(melt, Stim=="SWAP")%>%
+    filter(variable %in% c("x_4_13_x","x_4_x_x", "x_x_13_x"))
+g<-ggplot(data, aes(x=variable, y=value, fill=TB, alpha=SM))
 g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
+    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
+    scale_alpha_manual(values=c(0.5,1))+
     theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    facet_grid(TB~., scale="free")+
-    labs(title="PEP Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="Cytokine Subsets",
-         y="Frequency of Cytokine+ CD4+ T cells (%)");
-
-data<-filter(melt, Stim=="SEA")
-g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    facet_grid(TB~., scale="free")+
-    labs(title="SEA Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="Cytokine Subsets",
-         y="Frequency of Cytokine+ CD4+ T cells (%)");
-
-data<-filter(melt, Stim=="SWAP")
-g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    facet_grid(TB~., scale="free")+
-    labs(title="SWAP Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="Cytokine Subsets",
-         y="Frequency of Cytokine+ CD4+ T cells (%)");
+    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
+    theme(plot.title = element_text(hjust = 0.5)) +
+    facet_grid(TB~., scale="fixed")+
+    coord_cartesian(ylim = c(0,.05))
 ```
@@
 library(knitr)
-TH<-select(Cytokine, Donor, TB, SM, Stim, Cyt.Freq, TH1.Freq, TH1.2.Freq, TH2.Freq)
+TH<-dplyr::select(CD4, Donor, TB, SM, Stim, Cyt.Freq, Th1.Freq, TH1.2.Freq, TH2.Freq)
 
-melt<-melt(TH,id.vars=c("TB","SM", "Stim"))
-melt$value<-as.numeric(melt$value)
-melt<-filter(melt, !(is.na(value)))
-melt<-filter(melt, !(value<0))
-melt$label = factor(melt$variable,levels=c("Cyt.Freq", "TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
-melt<-filter(melt, !is.na(label))
+melt<-melt(TH,id.vars=c("TB","SM", "Stim"), measure.vars=c("Cyt.Freq", "Th1.Freq", "TH1.2.Freq", "TH2.Freq"))
+melt$variable = factor(melt$variable,levels=c("Cyt.Freq", "Th1.Freq", "TH1.2.Freq", "TH2.Freq"))
 
@@
     theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
+    facet_grid(variable~TB, scale="free")+
     labs(title="Non Specific CD4+ T cell Responses", 
@@
 data<-filter(TH, Stim=="PMA")
-kable(pvals(data, "TH1.Freq"), caption = "Non Specific CD4+ T cell Responses")
-kable(pvals(data, "TH2.Freq"))
-kable(pvals(data, "TH1.2.Freq"))
+kable(pvals_helminth(data, "Th1.Freq"), caption = "Non Specific CD4+ T cell Responses")
+kable(pvals_helminth(data, "TH2.Freq"))
+kable(pvals_helminth(data, "TH1.2.Freq"))
 
@@
     theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
+    facet_grid(variable~TB, scale="free")+
     labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
@@
 data<-filter(TH, Stim=="WCL")
-kable(pvals(data, "TH1.Freq"), caption = "Mtb Whole Cell Lysate Specific CD4 T cell Responses")
-kable(pvals(data, "TH2.Freq"))
-kable(pvals(data, "TH1.2.Freq"))
+kable(pvals_helminth(data, "Th1.Freq"), caption = "Mtb Whole Cell Lysate Specific CD4 T cell Responses")
+kable(pvals_helminth(data, "TH2.Freq"))
+kable(pvals_helminth(data, "TH1.2.Freq"))
 
@@
     theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
+    facet_grid(variable~TB, scale="free")+
     labs(title="Mtb Peptide Specific CD4+ T cell Responses", 
@@
 data<-filter(TH, Stim=="PEP")
-kable(pvals(data, "TH1.Freq"), caption = "Mtb Peptide Specific CD4 T cell Responses")
-kable(pvals(data, "TH2.Freq"))
-kable(pvals(data, "TH1.2.Freq"))
+kable(pvals_helminth(data, "Th1.Freq"), caption = "Mtb Peptide Specific CD4 T cell Responses")
+kable(pvals_helminth(data, "TH2.Freq"))
+kable(pvals_helminth(data, "TH1.2.Freq"))
 
@@
 
-melt<-melt(cyt,id.vars=c("TB","SM", "Stim"))
-melt$value<-as.numeric(melt$value)
-melt<-filter(melt, !(is.na(value)))
-melt<-filter(melt, !(value<0))
-melt$label = factor(melt$variable,levels=c("IFNg", "TNFa", "IL4", "IL13"))
-melt<-filter(melt, !is.na(label))
+melt<-melt(cyt,id.vars=c("TB","SM", "Stim"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
+melt$variable = factor(melt$variable,levels=c("IFNg", "TNFa", "IL4", "IL13"))
 
@@
 data<-filter(cyt, Stim=="PMA")
-kable(pvals(data, "IFNg"), caption = "Non Specific CD4 T cell Responses")
-kable(pvals(data, "TNFa"))
-kable(pvals(data, "IL4"))
-kable(pvals(data, "IL13"))
+kable(pvals_helminth(data, "IFNg"), caption = "Non Specific CD4 T cell Responses")
+kable(pvals_helminth(data, "TNFa"))
+kable(pvals_helminth(data, "IL4"))
+kable(pvals_helminth(data, "IL13"))
 
@@
 data<-filter(cyt, Stim=="WCL")
-kable(pvals(data, "IFNg"), caption = "Mtb Whole Cell Lysate Specific CD4 T cell Responses")
-kable(pvals(data, "TNFa"))
-kable(pvals(data, "IL4"))
-kable(pvals(data, "IL13"))
+kable(pvals_helminth(data, "IFNg"), caption = "Mtb Whole Cell Lysate Specific CD4 T cell Responses")
+kable(pvals_helminth(data, "TNFa"))
+kable(pvals_helminth(data, "IL4"))
+kable(pvals_helminth(data, "IL13"))
 
@@
 data<-filter(cyt, Stim=="PEP")
-kable(pvals(data, "IFNg"), caption = "Mtb Peptide Specific CD4 T cell Responses")
-kable(pvals(data, "TNFa"))
-kable(pvals(data, "IL4"))
-kable(pvals(data, "IL13"))
+kable(pvals_helminth(data, "IFNg"), caption = "Mtb Peptide Specific CD4 T cell Responses")
+kable(pvals_helminth(data, "TNFa"))
+kable(pvals_helminth(data, "IL4"))
+kable(pvals_helminth(data, "IL13"))
 
@@
 library(knitr)
-SP_lin<-rename(Cytokine, G_single =  G_x_x_x, T_single = x_x_x_T, IL4_single = x_4_x_x, IL13_single = x_x_13_x, 
-               G_T = G_x_x_T, IL4_IL13 = x_4_13_x) %>%
+SP_lin<-rename(Cytokine, G_single =  G_x_x_x, T_single = x_x_x_T, IL4_single = x_4_x_x, IL13_single = x_x_13_x, G_T = G_x_x_T, IL4_IL13 = x_4_13_x) %>%
     select(Donor, TB, SM, Stim, G_single, T_single, IL4_single, IL13_single, G_T, IL4_IL13)
 
-melt<-melt(SP_lin,id.vars=c("TB","SM", "Stim"))
-melt$value<-as.numeric(melt$value)
-melt<-filter(melt, !(is.na(value)))
-melt<-filter(melt, !(value<0))
+melt<-melt(SP_lin,id.vars=c("TB","SM", "Stim"), measure.vars=c("G_single", "T_single", "G_T", "IL4_single", "IL13_single", "IL4_IL13"))
 melt$label = factor(melt$variable,levels=c("G_single", "T_single", "G_T", "IL4_single", "IL13_single", "IL4_IL13"))
-melt<-filter(melt, !is.na(label))
 
@@
 data<-filter(SP_lin, Stim=="PMA")
-kable(pvals(data, "G_single"), caption = "Non-Specific CD4+ T cell Responses")
-kable(pvals(data, "T_single")) 
-kable(pvals(data,"G_T"))
-kable(pvals(data, "IL4_single"))
-kable(pvals(data, "IL13_single")) 
-kable(pvals(data, "IL4_IL13"))
+kable(pvals_helminth(data, "G_single"), caption = "Non-Specific CD4+ T cell Responses")
+kable(pvals_helminth(data, "T_single")) 
+kable(pvals_helminth(data,"G_T"))
+kable(pvals_helminth(data, "IL4_single"))
+kable(pvals_helminth(data, "IL13_single")) 
+kable(pvals_helminth(data, "IL4_IL13"))
 
@@
 data<-filter(SP_lin, Stim=="WCL")
-kable(pvals(data, "G_single"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
-kable(pvals(data, "T_single")) 
-kable(pvals(data,"G_T"))
-kable(pvals(data, "IL4_single"))
-kable(pvals(data, "IL13_single")) 
-kable(pvals(data, "IL4_IL13"))
+kable(pvals_helminth(data, "G_single"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
+kable(pvals_helminth(data, "T_single")) 
+kable(pvals_helminth(data,"G_T"))
+kable(pvals_helminth(data, "IL4_single"))
+kable(pvals_helminth(data, "IL13_single")) 
+kable(pvals_helminth(data, "IL4_IL13"))
 
@@
 data<-filter(SP_lin, Stim=="PEP")
-kable(pvals(data, "G_single"), caption = "Mtb Peptide Specific CD4 T cell Responses")
-kable(pvals(data, "T_single")) 
-kable(pvals(data,"G_T"))
-kable(pvals(data, "IL4_single"))
-kable(pvals(data, "IL13_single")) 
-kable(pvals(data, "IL4_IL13"))
+kable(pvals_helminth(data, "G_single"), caption = "Mtb Peptide Specific CD4 T cell Responses")
+kable(pvals_helminth(data, "T_single")) 
+kable(pvals_helminth(data,"G_T"))
+kable(pvals_helminth(data, "IL4_single"))
+kable(pvals_helminth(data, "IL13_single")) 
+kable(pvals_helminth(data, "IL4_IL13"))
 
@@
 data<-filter(SP_Ag, Stim=="PMA")
-kable(pvals(data, "G_plus"), caption = "Non Specific CD4+ T cell Responses")
-kable(pvals(data, "T_plus"))
-kable(pvals(data, "G_T_plus"))
-kable(pvals(data, "IL4_plus"))
-kable(pvals(data, "IL13_plus"))
-kable(pvals(data, "IL4_IL13_plus"))
+kable(pvals_helminth(data, "G_plus"), caption = "Non Specific CD4+ T cell Responses")
+kable(pvals_helminth(data, "T_plus"))
+kable(pvals_helminth(data, "G_T_plus"))
+kable(pvals_helminth(data, "IL4_plus"))
+kable(pvals_helminth(data, "IL13_plus"))
+kable(pvals_helminth(data, "IL4_IL13_plus"))
 
@@
 data<-filter(SP_Ag, Stim=="WCL")
-kable(pvals(data, "G_plus"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
-kable(pvals(data, "T_plus"))
-kable(pvals(data, "G_T_plus"))
-kable(pvals(data, "IL4_plus"))
-kable(pvals(data, "IL13_plus"))
-kable(pvals(data, "IL4_IL13_plus"))
+kable(pvals_helminth(data, "G_plus"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
+kable(pvals_helminth(data, "T_plus"))
+kable(pvals_helminth(data, "G_T_plus"))
+kable(pvals_helminth(data, "IL4_plus"))
+kable(pvals_helminth(data, "IL13_plus"))
+kable(pvals_helminth(data, "IL4_IL13_plus"))
 
@@
 data<-filter(SP_Ag, Stim=="PEP")
-kable(pvals(data, "G_plus"), caption = "Mtb Peptide Specific CD4+ T cell Responses")
-kable(pvals(data, "T_plus"))
-kable(pvals(data, "G_T_plus"))
-kable(pvals(data, "IL4_plus"))
-kable(pvals(data, "IL13_plus"))
-kable(pvals(data, "IL4_IL13_plus"))
+kable(pvals_helminth(data, "G_plus"), caption = "Mtb Peptide Specific CD4+ T cell Responses")
+kable(pvals_helminth(data, "T_plus"))
+kable(pvals_helminth(data, "G_T_plus"))
+kable(pvals_helminth(data, "IL4_plus"))
+kable(pvals_helminth(data, "IL13_plus"))
+kable(pvals_helminth(data, "IL4_IL13_plus"))
 ```

### 2019-03-27 MIMOSA.R

@@ -1,115 +1,149 @@
-biocLite("devtools")
+install.packages("devtools") #install devtools package
 library(devtools) #load it
 install_github("MIMOSA","RGLab",branch="master") #Need R 3.0.0, and a bunch of dependencies. The install will fail with various error messages until you install those dependencies.
-biocLite("MIMOSA")
 library(MIMOSA)
 
-DF<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/All Counts.csv")
-DF<-as.data.table(DF) 
-DF$Neg<-DF[,21]
-DF<-DF[DF$Stim!="PMA"]
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
+DF<-read.csv("CD4_Counts_Clean.csv")
 
 #All cytokine
-data<-DF[,c(2,5,26,22)]
+data<-dplyr::select(DF, Sample, Donor, TB, SM, Stim, Cyt, Total) %>%
+    dplyr::mutate(Neg = (Total - Cyt)) %>%
+    dplyr::select(-Total)
+
 E<-ConstructMIMOSAExpressionSet(data,
-                                reference=Stim%in%'UN',
-                                measure.columns=c('Neg','Cyt.Count'),
-                                other.annotations=c('Stim','ID'),
-                                default.cast.formula=component~ID+Stim,
-                                .variables=.(ID))
-test <- MIMOSA(Neg+Cyt.Count~ID|Stim,
-               data=E)
+                                reference=Stim %in% 'UN',
+                                measure.columns=c('Neg','Cyt'),
+                                other.annotations=c('Stim','Sample', 'Donor', 'TB', 'SM'),
+                                default.cast.formula=component~Donor+Stim,
+                                .variables=.(Donor))
+test <- MIMOSA(Neg+Cyt~Donor|Stim,
+               data=E,
+               method="mcmc")
+
 probabilities<-MIMOSA::getZ(test)
 counts<-MIMOSA::countsTable(test)
 fdr<-MIMOSA::fdr(test)
-MIMOSA.Results<-cbind(counts,probabilities,fdr)
+MIMOSA.Results<-data.frame(cbind(counts,probabilities,fdr))
[…]

### 2019-04-06 SUBUN.R
deleted file

@@ -1,60 +1,0 @@
-# All of this generates a fake data table that contains a "key" variable (i.e. donor)
-# a condition variable (i.e. "stim") and a set of columns of output data
-# the goal is for every donor, in each column of output data subtract out the value in that column
-# that corresponds to a specific value in the condition column
-donor<-c(rep("john", 6), rep("mary", 6), rep("jack", 6), rep("suzy", 6))
-stim<-rep(c("un","w","p","s","sw","sb"),4)
-ifng<-rnorm(24, 2)
-tnfa<-rnorm(24,10)
-IL4<-rnorm(24,5)
-IL5<-rnorm(24,7)
-IL10<-rnorm(24,13)
-IL13<-rnorm(24,19)
-IL17<-rnorm(24,17)
-IL21<-rnorm(24,11)
-IL22<-rnorm(24,3)
-
-fake<-data.frame(cbind(donor,stim,ifng,tnfa,IL4, IL5, IL10, IL13, IL17, IL21, IL22))
-remove(donor,stim,ifng,tnfa,IL4, IL5, IL10, IL13, IL17, IL21, IL22)
-#can only use data frames to use bracketed string column indexing so make sure this is a data frame not a data table
-
-#when binding data together the output data becomes facgtors and we need them to be numeric
-class_change<-function(datatable){
-    y<-dplyr::select(datatable, donor, stim)
-    x<-dplyr::select(datatable, -donor, -stim)
-    for(i in 1:length(x)){x[[i]]<-as.numeric(as.character(x[[i]]))}
-    df<-cbind(y, x)
-}
-fake<-class_change(fake)
-#we want to copy the original data table to reference back to and check our math
-fakecopy<-fake
-
-#This is going to generate the list of columns names that we want to iterate over
-numeric.only <- function(X,...){
-    returnCols <- names(X)
-    a<-sapply(X, is.numeric)
-    print(returnCols[a == "TRUE"])
-}
-colnms<-numeric.only(fake)
-
-dt<-split(fake, fake[,"donor"]) #generate a list of donor specific mini data frames 
-
-#split --> lists in alphabetical order not by original key order
-#To solve this I arranged the original data frame in alphabetical order
-
-fake<-arrange(fake, donor)
-#can we instead arrange the split to match the original order
-
-#for each column name in the list of column names
-for(col in colnms){
-    #this is a check to make sure it goes through all the columns
-    print(col) 
-    #dt.r will make a list of length of dt where each item in the list is the new column for a particular donor
[…]

### 2019-04-06 ICS Cleaning.Rmd
including age and sex

---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE )
```

```{r Donor Info}
library(dplyr)
setwd("/Applications/Old Computer/Day Lab/Flow-Data/exported")
PBMC<-read.csv("ICS_PBMC.csv")
PBMC<-dplyr::filter(PBMC, Helminth.Status=="Schisto+" | Helminth.Status== "Negative" | Helminth.Status== "Not US" | Helminth.Status=="US" | Helminth.Status== "Unknown")
PBMC$TB.Status = factor(PBMC$TB.Status,levels=c("Naïve","BCG","Healthy Control", "LTBI", "Active TB", "Unknown"))
PBMC$Viability <- gsub("%/%" , "", PBMC$Viability)
PBMC$Exp.Date <- gsub("TM", "", PBMC$Exp.Date)
PBMC<- dplyr::filter(PBMC, !(Excluded=="EX"))%>%
    dplyr::rename(TB = TB.Status, SM = Helminth.Status, StudyID = Study.ID)%>%
    dplyr::select(StudyID, TB, SM, Viability, Cells.Vial, Excluded)

QFT<- read.csv("/Applications/Old Computer/Epi Project/Data_clean/complete-kenya.csv")
QFT<- dplyr::filter(QFT, StudyID %in% PBMC$StudyID)%>%
    dplyr::select(StudyID, age, sex, QFT, SchistosomaIntensity)
QFTStudyID<-tstrsplit(QFT$StudyID, "_")

DonorInfo<-merge(PBMC, QFT, by="StudyID", all = TRUE)
DonorInfo$StudyID<-gsub("- plasma", "", DonorInfo$StudyID)
DonorInfo$StudyID<-gsub("-1", "", DonorInfo$StudyID)
DonorInfo$StudyID<-gsub("-0", "", DonorInfo$StudyID)
DonorInfo$Viability<-gsub("%" , "" , DonorInfo$Viability)
DonorInfo$Cells.Vial<-as.numeric(as.character(DonorInfo$Cells.Vial))
DonorInfo$QFT<-as.numeric(as.character(DonorInfo$QFT))
DonorInfo<-dplyr::rename(DonorInfo, Donor = StudyID)
```

```{r read in all flowjo data tables}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/exported/")
filenames =list.files(pattern="*.csv")
for (i in 1:length(filenames)) 
    assign(filenames[i], read.csv(filenames[i]))
```

```{r reformatting functions}
clean_flowjo_output<-function(datatable){
    library(dplyr)
    datatable$Sample<-gsub("HD474", "HD454", datatable$Sample)
    datatable$Sample<-gsub("HD448", "HD488", datatable$Sample)
    datatable$Sample<-gsub("NK2341_TB_SM", "NK2341_TB_X", datatable$Sample)
    datatable$Sample<-gsub("001_", "", datatable$Sample)
    datatable$Sample<-gsub("1: |2: |3: |4: |5: |6: ", "", datatable$Sample)
    datatable$Sample<-gsub("XUN", "X_UN", datatable$Sample)
    datatable$Sample<-gsub("XPMA", "X_PMA", datatable$Sample)
    datatable$Sample<-gsub("XWCL", "X_WCL", datatable$Sample)
    datatable$Sample<-gsub("XPEP", "X_PEP", datatable$Sample)
    datatable$Sample<-gsub("XSEA", "X_SEA", datatable$Sample)
    datatable$Sample<-gsub("XSWAP", "X_SWAP", datatable$Sample)
    datatable$Sample<-gsub("_ ", "_", datatable$Sample)
    datatable$Sample<-gsub(" _", "_", datatable$Sample)
    datatable<-filter(datatable,!(Sample=="Mean"))%>%
        filter(!Sample=="StdDev")%>%
        filter(!Sample=="Sample")%>%
        filter(!Sample=="Sample")
    datatable[datatable=="¥"]<-NA
    datatable[datatable=="•"]<-NA
    datatable

}

splitter<-function(data,column,splitter,one=NULL,two=NULL,three=NULL,four=NULL){
    library(data.table)

    labels<-tstrsplit(data[,column], splitter)
    data[,one]<-labels[[1]]
    data[,two]<-labels[[2]]
    data[,three]<-labels[[3]]
    data[,four]<-labels[[4]]
    return(data)
}
#demo: HEL<-splitter(HEL, "StudyID", "-", one="StudyID", two="Visit", three="Iteration")

remove_samples<-function(datatable){
    library(dplyr)
    excluded_samples<-c("NK2126_HC_X_SEA_005.fcs", 
                        "NK2450_HC_SM_SEA_047.fcs" , 
                        "NK2402_LTBI_SM_SWAP_024.fcs",
                        "NK2342_TB_X_SWAP_006.fcs",
                        "NK2136_TB_X_WCL_003.fcs",
                        "NK2168_HC_SM_WCL_040.fcs",
                        "NK2421_LTBI_SM_PEP_010.fcs",
                        "NK2063_TB_X_SEA_005.fcs",
                        "NK2115_LTBI_X_PEP_010.fcs",
                        "NK2115_LTBI_X_WCL_009.fcs",
                        "NK2171_HC_X_WCL_009.fcs",
                        "NK2162_HC_X_SEA_023.fcs")
    datatable<-filter(datatable, !(Sample %in% excluded_samples))%>%
        filter(!(Donor == "NK2111"))
    datatable
}

class_change<-function(datatable){
    y<-dplyr::select(datatable, Sample, Donor, TB, SM, Stim)
    x<-dplyr::select(datatable, -Sample, -Donor, -TB, -SM, -Stim)
    for(i in 1:length(x)){x[[i]]<-as.numeric(as.character(x[[i]]))}
    df<-cbind(y, x)
}

```

```{r, reformat all data tables}
l<-list(CD4_Counts.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_Positive.csv, CD4.csv, 
        CD8_Counts.csv, CD8_Cytokine.csv, GD_Counts.csv, GD_Cytokine.csv)
l<-lapply(l, function(x) clean_flowjo_output(x))
l<-lapply(l, function(x) splitter(x, "Sample", "_", one="Donor", two="TB", three="SM", four="Stim"))
l<-lapply(l, function(x) remove_samples(x))
l<-lapply(l, function(x) class_change(x))
for (i in 1:length(filenames)) {
    assign(filenames[i], data.frame(l[i]))}
```

```{r change colnames}
newnames<-names(CD8_Cytokine.csv)
names(CD4_Counts.csv)<-newnames[1:21]
names(CD8_Counts.csv)<-newnames
names(GD_Counts.csv)<-newnames
names(GD_Cytokine.csv)<-newnames
```

```{r update and re-organize count data}
library(dplyr)
CD4_Counts.csv<-mutate(CD4_Counts.csv,
    IFNg=  G_4_13_T + G_4_13_x + G_4_x_T + G_4_x_x + G_x_13_T + G_x_13_x + G_x_x_T + G_x_x_x,
    TNFa=  G_4_13_T + G_4_x_T + G_x_13_T + G_x_x_T + x_4_13_T + x_4_x_T + x_x_13_T + x_x_x_T,
    IL13=  G_4_13_T + G_4_13_x + G_x_13_T + G_x_13_x + x_4_13_T + x_4_13_x + x_x_13_T + x_x_13_x,
    IL4=   G_4_13_T + G_4_13_x + G_4_x_T + G_4_x_x + x_4_13_T + x_4_13_x + x_4_x_T + x_4_x_x,
    Total= G_4_13_T + G_4_x_T + G_x_13_T + G_4_13_x + x_4_13_T + G_4_x_x + G_x_13_x + x_4_x_T + x_x_13_T + G_x_x_T  + G_x_x_x + x_x_x_T +  x_4_13_x + x_4_x_x + x_x_13_x + x_x_x_x
)
xtracounts<-dplyr::select(CD4_Cytokine.csv, Sample, Cyt.Count, TH1.Count, TH1.2.Count, TH2.Count)%>%
    dplyr::rename(Cyt = Cyt.Count, TH1 = TH1.Count, TH2 = TH2.Count, TH1.2 = TH1.2.Count)

CD4_Counts.csv<-merge(CD4_Counts.csv, xtracounts, by = "Sample")
```

```{r create new combo cytokine variables}
library(dplyr)
filenames = c("CD4_Cytokine.csv", "CD8_Cytokine.csv", "GD_Cytokine.csv")
l<-list(CD4_Cytokine.csv, CD8_Cytokine.csv, GD_Cytokine.csv)
l<-lapply(l, function(x) dplyr::mutate(x, G_plus = G_4_13_x + G_4_x_x + G_x_13_x + G_x_x_x,
                   T_plus = x_4_13_T + x_4_x_T + x_x_13_T + x_x_x_T,
                   G_T_plus = G_4_x_T + G_x_13_T + G_x_x_T,
                   IL4_plus = G_4_x_T + G_4_x_x + x_4_x_T + x_4_x_x,
                   IL13_plus = G_x_13_T + G_x_13_x + x_x_13_T + x_x_13_x,
                   IL4_IL13_plus = G_4_13_x + x_4_13_T + x_4_13_x)
)

for (i in 1:length(filenames)) {
    assign(filenames[i], 
           data.frame(l[i]))}
```

```{r Cytokine MFI inclusion}
counts.ref<-CD4_Counts.csv
DF<-arrange(CD4_MFI.csv, Sample)
CD4_Counts.csv<-arrange(CD4_Counts.csv, Sample)
CD4_MFI.csv<-arrange(CD4_MFI.csv, Sample)

keynames<-colnames(counts.ref)[6:29]
for(key in keynames){
    remove<-subset(counts.ref, counts.ref[,key]<20)$Sample
    print(key)
    print(length(remove))
    DF[(DF$Sample %in% remove), grep(paste(key,"_",sep=""), names(DF))]<-NA
}

#for me
#for(key in keynames){
#    print(key)
#    test<-data.frame(cbind(CD4_Counts.csv[grep(key, names(CD4_Counts.csv))], 
#                           CD4_MFI.csv[grep(paste(key,"_",sep=""), names(CD4_MFI.csv))],
#                           DF[grep(paste(key,"_",sep=""), names(DF))]))
#    print(length(which(test[,1]<20 &! is.na(test[,2:5]))))
#    print(length(which(test[,1]<20 &! is.na(test[,6:9]))))
#}

CD4_MFI.csv<-DF
```

```{r subtract unstimulated from cyt freq tables}
#used to generate a list of character strings that are the names of the columns containing numeric data
numeric.only <- function(X,...){
      returnCols <- names(X)
      a<-sapply(X, is.numeric)
      print(returnCols[a == "TRUE"])
}

#takes a dataframe (v. important its a dataframe), splitvar is a character string of a column name
#colnms is a list of strings indicating which columns to subtract background from
#condition is a character string with the key to indicate what is being subtracted
background.subtract<-function(df, splitvar, keyvar, condition){
    df<-dplyr::arrange(df, df[,splitvar])
    dt<-split(df, df[, splitvar]) #generate a list of mini data frames for each level of splitvar
    #split --> lists in alphabetical order not by original key order
    #To solve this I arranged the original data frame in alphabetical order
    #can we instead arrange the split to match the original order

    colnms<-numeric.only(df)

    #for each column name in the list of column names
    for(col in colnms){
        #this is a check to make sure it goes through all the columns
        print(col) 
        #dt.r will make a list of length of dt where each item in the list is the new column for a particular donor
        dt.r<-lapply(dt, function(g)
        #take the values in each column and subtract out the value indicated by an "un" the stim column 
        (g[,col] - subset(g, g[,keyvar]==condition)[,col]))
        #unlist the list so you get a single column and insert it back into the original data frame
        df[,col]<-unlist(dt.r)
    }
    df
}

filenames = c("CD4_Cytokine.csv", "CD8_Cytokine.csv", "GD_Cytokine.csv")
l<-list(CD4_Cytokine.csv, CD8_Cytokine.csv, GD_Cytokine.csv)
l<-lapply(l, function(x) arrange(x, Donor))
l<-lapply(l, function(x) background.subtract(x, "Donor", "Stim", "UN"))
for (i in 1:length(filenames)) {
    assign(filenames[i], 
           data.frame(l[i]))}
CD4_Cytokine.csv[CD4_Cytokine.csv<0]<-0
CD8_Cytokine.csv[CD8_Cytokine.csv<0]<-0
GD_Cytokine.csv[GD_Cytokine.csv<0]<-0
```

```{r write files to "clean" directory}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
df.list<-list(CD4_Counts.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_Positive.csv,
              CD4.csv, CD8_Counts.csv, CD8_Cytokine.csv, GD_Counts.csv,
              GD_Cytokine.csv)
dfnames<-list("CD4_Counts", "CD4_Cytokine", "CD4_MFI", "CD4_Positive",
              "CD4", "CD8_Counts", "CD8_Cytokine", "GD_Counts",
              "GD_Cytokine")
newfilenames<-paste(dfnames, "clean", date(), ".csv", sep="_")
#write.csv(df.list[1], file= newfilenames[1])
write.csv(CD4_Counts.csv, file = "CD4_Counts_clean.csv")
write.csv(CD4_Cytokine.csv, file = "CD4_Cytokine_clean.csv")
write.csv(CD4_MFI.csv, file = "CD4_MFI_clean.csv")
write.csv(CD4_Positive.csv, file = "CD4_Positive_clean.csv")
write.csv(CD4.csv, file = "CD4_clean.csv")
write.csv(CD8_Counts.csv, file = "CD8_Counts_clean.csv")
write.csv(CD8_Cytokine.csv, file = "CD8_Cytokine_clean.csv")
write.csv(GD_Counts.csv, file = "GD_Counts_clean.csv")
write.csv(GD_Cytokine.csv, file = "GD_Cytokine_clean.csv")
write.csv(DonorInfo, file = "ICS_Donor_Info_clean.csv")
```

### 2019-04-17 ICS_Cleaning.Rmd
annotated script

---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE )
```

```{r Donor Info}
library(dplyr)
setwd("/Applications/Old Computer/Day Lab/Flow-Data/exported")
#Import information about the samples from my lab notes
    PBMC<-read.csv("ICS_PBMC.csv")
#remove samples that are other helminths since we dropped that group
#Not US, US and Unknown are from the Yerkes donors
    PBMC<-dplyr::filter(PBMC, Helminth.Status=="Schisto+" | Helminth.Status== "Negative" | Helminth.Status== "Not US" |  Helminth.Status=="US" | Helminth.Status== "Unknown")
    PBMC$Helminth.Status<-factor(PBMC$Helminth.Status, levels=c("US", "Not US", "Negative", "Schisto+", "Unknown"))
#order the levels of TB status to be from least to most exposure/disease
    PBMC$TB.Status = factor(PBMC$TB.Status,levels=c("Naïve","BCG","Healthy Control", "LTBI", "Active TB", "Unknown"))
#take out all the stuff that isn't consistent like the "TM" in experiment name and the % in viability
    PBMC$Viability <- gsub("%/%" , "", PBMC$Viability)
    PBMC$Viability<-gsub("%" , "" , PBMC$Viability)
    PBMC$Exp.Date <- gsub("TM", "", PBMC$Exp.Date)
#Some of the samples had weird suffixes that have to be removed to match the epi data  
    PBMC$Study.ID<-gsub("- plasma", "", PBMC$Study.ID)
    PBMC$Study.ID<-gsub("-1", "", PBMC$Study.ID)
    PBMC$Study.ID<-gsub("-0", "", PBMC$Study.ID)
#Remove samples that are being excluded from the ICS analysis
PBMC<- dplyr::filter(PBMC, !(Excluded=="EX"))%>%
    dplyr::rename(TB = TB.Status, SM = Helminth.Status, StudyID = Study.ID)%>% #match the variable names to ICS variable names
    dplyr::select(StudyID, TB, SM, Viability, Cells.Vial, Excluded) #take only the columns you need

#Import the clinical enrollment data from the cleaned merged epi file
    QFT<- read.csv("/Applications/Old Computer/Epi Project/Data_clean/complete-kenya.csv")
#The studyIDs of coenrolled people are NK#_TRU# and we want just the NK# to match the lab notes
    QFT$StudyID<-tstrsplit(QFT$StudyID, "_")[[1]]
#trim the data set to only those study IDs in the lab notes and select the columns you want
    QFT<- dplyr::filter(QFT, StudyID %in% PBMC$StudyID)%>%
        dplyr::select(StudyID, age, sex, QFT, SchistosomaIntensity)

#Merge the two tables together and keep all observations since the yerkes donors won't be in the epi data
    DonorInfo<-merge(PBMC, QFT, by="StudyID", all = TRUE)
#The class of variables gets messed up in the merge so re-class them
    DonorInfo$Cells.Vial<-as.numeric(as.character(DonorInfo$Cells.Vial))
    DonorInfo$QFT<-as.numeric(as.character(DonorInfo$QFT))
#Match the variable names of the ICS data
    DonorInfo<-dplyr::rename(DonorInfo, Donor = StudyID)

#Manual Edits that may get deleted after Kenya script is finalized
    DonorInfo$TB[DonorInfo$Donor=="NK2183"]<-as.character("Healthy Control")
    DonorInfo$QFT[DonorInfo$Donor=="NK2323"]<-0.04
    DonorInfo$SM[DonorInfo$Donor=="NK2325"]<-as.character("Unknown")
```

```{r read in all flowjo data tables}
#change working directory to wherever the exported files are
setwd("/Applications/Old Computer/Day Lab/Flow-Data/exported/")
#find all the csv filenames in that directory and put them in a list
filenames =list.files(pattern="*.csv")
#for each name in that list, assign it to the output of read.csv of that name
#aka read in each file by it's name
for (i in 1:length(filenames)) 
    assign(filenames[i], read.csv(filenames[i]))
```
```{r source functions used in this script}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/ICS-cleaning-functions.R")
```

```{r, reformat all data tables}
#this puts all the dataframes into a list 
#then applies each function over the dataframes in the list
#and assigns the output of that to the same named list
#so at every step the list of dataframes gets updated based on the function output

l<-list(CD4_Counts.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_Positive.csv, CD4.csv, 
        CD8_Counts.csv, CD8_Cytokine.csv, GD_Counts.csv, GD_Cytokine.csv)
l<-lapply(l, function(x) clean_flowjo_output(x))
l<-lapply(l, function(x) splitter(x, "Sample", "_", one="Donor", two="TB", three="SM", four="Stim"))
l<-lapply(l, function(x) remove_samples(x))
l<-lapply(l, function(x) fix.names(x))
l<-lapply(l, function(x) class_change(x))

#then we use the assign function to extract each data frame from the list and assign it to the filename
#associated with it from the beginning
#this only works because the list of dataframes is arranged in the same order as the list filenames 
for (i in 1:length(filenames)) {
   assign(filenames[i], data.frame(l[i]))}
```

```{r change colnames}
#I only bothered to change the names in flowjo for CD4/8 Cytokine exports so the rest are ugly
newnames<-names(CD8_Cytokine.csv)
#luckily the columns are in the same order so I can just take the names from one of those 
names(CD4_Counts.csv)<-newnames[1:21]
#and give it to all the counts tables
names(CD8_Counts.csv)<-newnames
names(GD_Counts.csv)<-newnames
names(GD_Cytokine.csv)<-newnames
```

```{r update and re-organize count data}
library(dplyr)
#again an oversight but I didn't export total counts for CD4 so I have to make them to match 
#CD8 and GD counts tables
#mutate just means make a new variable 
CD4_Counts.csv<-mutate(CD4_Counts.csv,
    #These all follow the same format
    #IFNg is the new variable. it gets assign the row sum of all the columns that have a G in the name
    IFNg=  rowSums(CD4_Counts.csv[,grep("G", names(CD4_Counts.csv))]),
    TNFa=  rowSums(CD4_Counts.csv[,grep("_T", names(CD4_Counts.csv))]),
    IL13=  rowSums(CD4_Counts.csv[,grep("13", names(CD4_Counts.csv))]),
    IL4=   rowSums(CD4_Counts.csv[,grep("4", names(CD4_Counts.csv))])
)
#pulls the columns with counts from the CD4_Cytokine export
xtracounts<-dplyr::select(CD4_Cytokine.csv, Sample, Cyt.Count, TH1.Count, TH1.2.Count, TH2.Count)%>%
    dplyr::rename(Cyt = Cyt.Count, TH1 = TH1.Count, TH2 = TH2.Count, TH1.2 = TH1.2.Count)
#and sticks them in this table
CD4_Counts.csv<-merge(CD4_Counts.csv, xtracounts, by = "Sample")

#then remove them from the other table
CD4_Cytokine.csv<-dplyr::select(CD4_Cytokine.csv, -Cyt.Count, -TH1.Count, -TH1.2.Count, -TH2.Count)
```

```{r create new combo cytokine variables}
#If you want to know the double producers without considering the opposite TH cytokines
library(dplyr)
filenames = c("CD4_Cytokine.csv", "CD8_Cytokine.csv", "GD_Cytokine.csv")
l<-list(CD4_Cytokine.csv, CD8_Cytokine.csv, GD_Cytokine.csv)
l<-lapply(l, function(x) dplyr::mutate(x, G_T = G_4_x_T + G_x_13_T + G_x_x_T,
                   IL4_IL13 = G_4_13_x + x_4_13_T + x_4_13_x)
)
for (i in 1:length(filenames)) {
    assign(filenames[i], 
           data.frame(l[i]))}
```

```{r Cytokine MFI inclusion}
#duplicate and rename the counts table so you don't fuck it up but can still use it
counts.ref<-CD4_Counts.csv
#arrange tables by Sample so the rows match up
counts.ref<-arrange(counts.ref, Sample)
CD4_MFI.csv<-arrange(CD4_MFI.csv, Sample)
#duplicate the mfi table again so you can change it but preserve the original just in case
DF<-arrange(CD4_MFI.csv, Sample)

#This is explained in detail in a separate script named "remove-samples-using-key.R"
keynames<-colnames(counts.ref)[6:29]
for(key in keynames){
    remove<-subset(counts.ref, counts.ref[,key]<20)$Sample
    print(key)
    print(length(remove))
    DF[(DF$Sample %in% remove), grep(paste(key,"_",sep=""), names(DF))]<-NA
}

#for me to test but doesn't work if you are actively working with the code. Only works if you just run the whole script. also documented in the same script as above
for(key in keynames){
    print(key)
    test<-data.frame(cbind(counts.ref[grep(key, names(counts.ref))], 
                           CD4_MFI.csv[grep(paste(key,"_",sep=""), names(CD4_MFI.csv))],
                           DF[grep(paste(key,"_",sep=""), names(DF))]))
    print(length(which(test[,1]<19 &! is.na(test[,2:5]))))
    print(length(which(test[,1]<19 &! is.na(test[,6:9]))))
}

#now that we have checked, we can assign it back to the original name
CD4_MFI.csv<-DF
```

```{r subtract unstimulated from cyt freq tables}
#the documentation for the background.subtract function is in a script in the miscellaneous folder
#just subtracts the value of the unstim from the value of the stim by donor
filenames = c("CD4_Cytokine.csv", "CD8_Cytokine.csv", "GD_Cytokine.csv")
l<-list(CD4_Cytokine.csv, CD8_Cytokine.csv, GD_Cytokine.csv)
l<-lapply(l, function(x) arrange(x, Donor))
l<-lapply(l, function(x) background.subtract(x, "Donor", "Stim", "UN"))
for (i in 1:length(filenames)) {
    assign(filenames[i], 
           data.frame(l[i]))}
#sets all negative values to zero
CD4_Cytokine.csv[CD4_Cytokine.csv<0]<-0
CD8_Cytokine.csv[CD8_Cytokine.csv<0]<-0
GD_Cytokine.csv[GD_Cytokine.csv<0]<-0
```

```{r write files to "clean" directory}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
df.list<-list(CD4_Counts.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_Positive.csv,
              CD4.csv, CD8_Counts.csv, CD8_Cytokine.csv, GD_Counts.csv,
              GD_Cytokine.csv)
dfnames<-list("CD4_Counts", "CD4_Cytokine", "CD4_MFI", "CD4_Positive",
              "CD4", "CD8_Counts", "CD8_Cytokine", "GD_Counts",
              "GD_Cytokine")
newfilenames<-paste(dfnames, "clean", date(), ".csv", sep="_")
#write.csv(df.list[1], file= newfilenames[1])
write.csv(CD4_Counts.csv, file = "CD4_Counts_clean.csv")
write.csv(CD4_Cytokine.csv, file = "CD4_Cytokine_clean.csv")
write.csv(CD4_MFI.csv, file = "CD4_MFI_clean.csv")
write.csv(CD4_Positive.csv, file = "CD4_Positive_clean.csv")
write.csv(CD4.csv, file = "CD4_clean.csv")
write.csv(CD8_Counts.csv, file = "CD8_Counts_clean.csv")
write.csv(CD8_Cytokine.csv, file = "CD8_Cytokine_clean.csv")
write.csv(GD_Counts.csv, file = "GD_Counts_clean.csv")
write.csv(GD_Cytokine.csv, file = "GD_Cytokine_clean.csv")
write.csv(DonorInfo, file = "ICS_Donor_Info_clean.csv")
```

### 2019-05-22 Cohort-Characteristics.Rmd
includes functional script for processing OG data

---
title: "Cohort_Characteristics"
author: "Taryn McLaughlin"
date: "4/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error=FALSE)
library(dplyr)
library(ggplot2)
library(knitr)
library(ggpubr)
```

```{r load data}
DonorInfo<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/ICS_Donor_Info_clean.csv")
DonorInfo<-filter(DonorInfo, SM %in% c("X", "SM"))
DonorInfo$SM<-factor(DonorInfo$SM, levels=c("X", "SM"))
DonorInfo$TB<-factor(DonorInfo$TB, levels=c("HC", "LTBI", "TB"))
DonorInfo$disease<-paste(DonorInfo$TB, DonorInfo$SM, sep="_")
DonorInfo$disease<-factor(DonorInfo$disease, levels=c("HC_X", "HC_SM",
                                                      "LTBI_X", "LTBI_SM",
                                                      "TB_X", "TB_SM"))
my_comparisons <- list(c("HC_SM", "HC_X"), c("LTBI_SM", "LTBI_X"), c("TB_SM", "TB_X"), 
                       c("HC_SM", "LTBI_SM"), c("TB_SM", "LTBI_SM"), c("HC_SM", "TB_SM"),
                       c("HC_X", "LTBI_X"), c("TB_X", "LTBI_X"), c("HC_X", "TB_X"))
DonorInfo$SchistosomaIntensity[DonorInfo$SM=="X"]<-NA
```

Generating stuff by hand
```{r by hand}
IQR_vals <- function(datatable, column) {
    library(data.table)
    y<-split(datatable, datatable$disease)
    x<-lapply(y, function(y) quantile(y[,column], probs=c(0.5, 0.25, 0.75), na.rm=TRUE))
    #quantile is a stats function that takes a vector "x" and returns the value at a specified quantile
    data.frame(t(x))
}
IQR_pvals<-function(data, column){
    x<-split(data,data$TB)
    #does a wilcox test between SM- and SM+ for whatever column you call in the column argument
    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"], exact=FALSE, na.action = na.pass))
    pvals<-c(A$HC$p.value, A$LTBI$p.value, A$TB$p.value)
    t(pvals)         
}

###N
n<-t(DonorInfo %>% dplyr::count(disease))
names(n)<-n[1,]

## AGE
fit<-lm(age~disease, data=DonorInfo)
total.p<-anova(fit)$`Pr(>F)`
#Age is different
ggplot(DonorInfo, aes(x=disease, y=age)) +
    geom_boxplot(aes(fill=TB, alpha=SM), size=1, position=position_dodge(width = 1), outlier.shape=NA) +
    geom_jitter(width=.1,height=0, shape=16,size=4)+
    scale_alpha_manual(values=c(0.5,1))+
    scale_fill_manual(values = c("#1a9850" , "#2166ac", "#b2182b"))+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons = my_comparisons,  p.adjust="bonferroni", aes(label=..p.adj..))+
    labs(subtitle = "These aren't corrected for whatever reason will figure out later")
AGE<-cbind(IQR_vals(DonorInfo, "age"),
           cbind(IQR_pvals(DonorInfo, "age"), total.p))
AGE<-dplyr::rename(AGE, HC.p = V1, LTBI.p = V2, TB.p = V3)
AGE$parameter<-"Age"

## SEX
female<-(DonorInfo %>% dplyr::group_by(sex) %>% dplyr::count(disease))[1:6,2:3]
male<-(DonorInfo %>% dplyr::group_by(sex) %>% dplyr::count(disease))[7:12,2:3]
df.sex<-merge(female, male, by="disease")
df.sex<-dplyr::rename(df.sex, female=n.x, male=n.y)
df.sex$femfreq = paste("(",
    round(df.sex$female/(df.sex$female+df.sex$male),2)*100, 
    "%", ")", sep="")
df.sex$malefreq = paste("(",
    round(df.sex$male/(df.sex$female+df.sex$male),2)*100,
    "%", ")", sep="")

total.p<-round(chisq.test(df.sex[,2:3])$p.value, 3)
HC.p<-round(chisq.test(df.sex[1:2,2:3])$p.value, 3)
LTBI.p<-round(chisq.test(df.sex[3:4,2:3])$p.value, 3)
TB.p<-round(chisq.test(df.sex[5:6,2:3])$p.value, 3)
pvals<-data.frame(cbind(HC.p, LTBI.p, TB.p, total.p))
female<-paste(df.sex$female, df.sex$femfreq, sep=" ")
male<-paste(df.sex$male, df.sex$malefreq, sep=" ")
d1<-data.frame(rbind(female, male))
names(d1)<-df.sex$disease
SEX<-cbind(d1, pvals)
SEX$parameter<-row.names(SEX)

##RACE
RACE<-paste(n[2,], "(100%)", sep=" ")

## QFT
qft<-filter(DonorInfo, TB!="TB")
qft$TB<-factor(qft$TB, levels=c("HC", "LTBI"))
qft$disease<-factor(qft$disease, levels=c("HC_X","HC_SM",
                                          "LTBI_X", "LTBI_SM"))
fit<-lm(QFT~disease, qft)
total.p<-anova(fit)$`Pr(>F)`
QFT<-IQR_vals(qft, "QFT")
x<-split(qft,qft$TB)
#does a wilcox test between SM- and SM+ for whatever column you call in the column argument
A<-lapply(x, function(g) wilcox.test(g$QFT~g$SM, exact=FALSE, na.action = na.pass))
pvals<-c(A$HC$p.value, A$LTBI$p.value, total.p)

## Schisto
eggs<-filter(DonorInfo, SM=="SM")
fit<-lm(SchistosomaIntensity~TB, eggs)
eggs.p<-anova(fit)$`Pr(>F)`[1]
EGGS<-cbind(IQR_vals(eggs, "SchistosomaIntensity"),eggs.p)

kable(AGE)
kable(SEX)
kable(QFT)
kable(EGGS)
```

https://cran.r-project.org/web/packages/tableone/vignettes/introduction.html

```{r tableone, echo = FALSE}
DonorInfo$SM<-gsub("SM", "SM+", DonorInfo$SM)
DonorInfo$SM<-gsub("X", "SM-", DonorInfo$SM)

library(tableone)
#Create a variable list which we want in Table 1
listVars <- c("age", "sex", "QFT", "SchistosomaIntensity")

#Define categorical variables
catVars <- c("TB","SM")

mediqr<- c("age", "QFT", "SchistosomaIntensity")

#Total Population
table1 <- CreateTableOne(vars = listVars, data = DonorInfo, factorVars = catVars,
                         strata = c("TB", "SM"))
test<- print(table1, showAllLevels = TRUE, nonnormal = mediqr, quote = FALSE, noSpaces = TRUE,
             printToggle = FALSE)

tableHC <- CreateTableOne(vars = listVars, data = filter(DonorInfo, TB=="HC"), factorVars = catVars, strata = "SM")
testHC<- print(tableHC, showAllLevels = TRUE, nonnormal = mediqr, quote = FALSE, noSpaces = TRUE,
             printToggle = FALSE)
tableLTBI <- CreateTableOne(vars = listVars, data = filter(DonorInfo, TB=="LTBI"), factorVars = catVars, strata = "SM")
testLTBI<- print(tableLTBI, showAllLevels = TRUE, nonnormal = mediqr, quote = FALSE, noSpaces =
                     TRUE,printToggle = FALSE)
tableTB <- CreateTableOne(vars = listVars, data = filter(DonorInfo, TB=="TB"), factorVars = catVars, strata = "SM")
testTB<- print(tableTB, showAllLevels = TRUE, nonnormal = mediqr, quote = FALSE, noSpaces = TRUE,
             printToggle = FALSE)

write.csv(test, file = "myTable.csv")
kable(test, caption = "FULL")
kable(testHC, caption = "Healthy Controls")
kable(testLTBI, caption = "LTBI")
kable(testTB, caption = "TB")
```

http://api.rpubs.com/isaac_subirana/compareGroups-demo

```{r compare groups, echo = FALSE}

library(compareGroups)
res.total<-descrTable(~ age + sex + QFT + SchistosomaIntensity, data = DonorInfo)
res <- compareGroups(SM ~ age + sex + QFT + SchistosomaIntensity, data = DonorInfo, 
                     method = c(age = 2, QFT = 2, SchistosomaIntensity = 2, sex = 3))

restab <- createTable(res, type = 1, hide.no = "no", show.n = FALSE)
test<-strataTable(restab, "TB")
export2md(test, header.labels = c(p.overall = "p-value"))
export2md(res.total)

```

Plotting stuff against age and sex
```{r age and sex vs, cache = TRUE}
compass<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/ICS_Compass_scores_complete.csv")
CD4<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/CD4_clean.csv")
CD4_cyt<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/CD4_Cytokine_clean.csv")

test<-merge(CD4, DonorInfo, by="Donor")
test<-filter(test, Stim=="UN")%>%
    select(-Donor, -X.x, -Sample, -TB.x, -TB.y, -SM.x, -SM.y, -Stim, -X.y, -Excluded)
test.age.plot<-function(datatable, column){
    g<-ggplot(datatable, aes(x=age, y=datatable[,column], color=sex))
    g<- g + geom_point()+theme_bw()+geom_smooth(method="lm", show.legend=TRUE)
    g<- g + labs(title=column)
    g
}
listnames<-names(test)
for(l in listnames){
    print(l)
    p<-test.age.plot(test, l)
    print(p)
}
### By Cytokine
test<-merge(CD4_cyt, DonorInfo, by="Donor")
test<-filter(test, Stim=="WCL")%>%
    select(-disease, -Excluded, -Cells.Vial, -Viability, -SM.y,- TB.y, -X.y, -SM.x,- TB.x, -X.x,-Sample)
test.age.plot<-function(datatable, column){
    g<-ggplot(datatable, aes(x=age, y=datatable[,column], color=sex))
    g<- g + geom_point()+theme_bw()+geom_smooth(method="lm", show.legend=TRUE)
    g<- g + labs(title=column, subtitle="WCL")
    g
}
listnames<-names(test)
for(l in listnames){
    print(l)
    p<-test.age.plot(test, l)
    print(p)
}

test<-merge(CD4_cyt, DonorInfo, by="Donor")
test<-filter(test, Stim=="PEP")%>%
    select(-disease, -Excluded, -Cells.Vial, -Viability, -SM.y,- TB.y, -X.y, -SM.x,- TB.x, -X.x,-Sample)
test.age.plot<-function(datatable, column){
    g<-ggplot(datatable, aes(x=age, y=datatable[,column], color=sex))
    g<- g + geom_point()+theme_bw()+geom_smooth(method="lm", show.legend=TRUE)
    g<- g + labs(title=column, subtitle="PEP")
    g
}
listnames<-names(test)
for(l in listnames){
    print(l)
    p<-test.age.plot(test, l)
    print(p)
}

## COMPASS
test<-merge(compass, DonorInfo, by="Donor")
test<-filter(test, Stim=="WCL", celltype=="CD4")
test.age.plot<-function(datatable, column){
    g<-ggplot(datatable, aes(x=age, y=datatable[,column], color=sex))
    g<- g + geom_point()+theme_bw()+geom_smooth(method="lm", show.legend=TRUE)
    g<- g + labs(title=column, subtitle="WCL")
    g
}
test.age.plot(test, "FS")
test.age.plot(test, "PFS")

test<-merge(CD4_cyt, DonorInfo, by="Donor")
test<-filter(test, Stim=="PEP")
test.age.plot<-function(datatable, column){
    g<-ggplot(datatable, aes(x=age, y=datatable[,column], color=sex))
    g<- g + geom_point()+theme_bw()+geom_smooth(method="lm", show.legend=TRUE)
    g<- g + labs(title=column, subtitle="PEP")
    g
}
listnames<-names(test)
for(l in listnames){
    print(l)
    p<-test.age.plot(test, l)
    print(p)
}
```

### 2019-05-22 ICS_Cleaning.Rmd
includes functional script for processing OG data

@@
 #Import the clinical enrollment data from the cleaned merged epi file
-    QFT<- read.csv("/Applications/Old Computer/Epi Project/Data_clean/complete-kenya.csv")
+    QFT<- read.csv("/Applications/Old Computer/Epi Project/Data_clean/complete-kenya-filledin.csv")
 #The studyIDs of coenrolled people are NK#_TRU# and we want just the NK# to match the lab notes
@@
 #trim the data set to only those study IDs in the lab notes and select the columns you want
-    QFT<- dplyr::filter(QFT, StudyID %in% PBMC$StudyID)%>%
-        dplyr::select(StudyID, age, sex, QFT, SchistosomaIntensity)
+    QFT2<-dplyr::select(QFT, StudyID, age, sex, QFT, SchistosomaIntensity)%>%
+        dplyr::filter(StudyID %in% PBMC$StudyID)
 
 #Merge the two tables together and keep all observations since the yerkes donors won't be in the epi data
-    DonorInfo<-merge(PBMC, QFT, by="StudyID", all = TRUE)
+    DonorInfo<-merge(PBMC, QFT2, by="StudyID", all = TRUE)
 #The class of variables gets messed up in the merge so re-class them
@@
     DonorInfo<-dplyr::rename(DonorInfo, Donor = StudyID)
+    DonorInfo$TB<-gsub("Active TB", "TB", DonorInfo$TB)
+    DonorInfo$TB<-gsub("Healthy Control", "HC", DonorInfo$TB)
+    DonorInfo$SM<-gsub("Negative", "X", DonorInfo$SM)
+    DonorInfo$SM<-gsub("Schisto+", "SM", DonorInfo$SM)
+    DonorInfo$SM<-gsub("\\+", "", DonorInfo$SM)
 
 #Manual Edits that may get deleted after Kenya script is finalized
-    DonorInfo$TB[DonorInfo$Donor=="NK2183"]<-as.character("Healthy Control")
+    DonorInfo$TB[DonorInfo$Donor=="NK2183"]<-as.character("HC")
     DonorInfo$QFT[DonorInfo$Donor=="NK2323"]<-0.04
@@
 ```
+
 ```{r source functions used in this script}
@@
 l<-lapply(l, function(x) class_change(x))
+l<-lapply(l, function(x) changeTB(x))
 

### 2019-05-22 ICS_Phenotyping.Rmd
includes functional script for processing OG data

---
title: "ICS Phenotyping"
author: "Taryn McLaughlin"
date: "12/2/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```

#Total CD4
```{r total CD4}
library(dplyr)
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
CD4<-read.csv("CD4_clean.csv")
    #since we don't care about the yerkes donors in our analysis and they mess everything up
    #we will filter them out
    data<-filter(CD4, SM!="N")
    #and then re-level both the SM and TB variable to not confuse our functions
    data$SM<-factor(data$SM, levels=c("X","SM"))
    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
CD4<-data

#creates a list of character strings for each of the columns of interest
colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){#this will loop through each character string in the list
    x<-subset(CD4, CD4$Stim=="UN") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    y<-pvals_helminth(x, col)
    print(y) #will return the output for each character string in the loop
}
```

#Removing Non-Specific Responses
```{r AG specific}
#mimosa is the program used to determine whether a response is "real" or not
#I have previously run this analysis, processed the data, and stored it locally
mimosa<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/CD4_MIMOSA_Complete.csv")
mimosa<-dplyr::rename(mimosa, id = Sample)
#select the columns that will be our sample indicator "id", as well as our "key" columns
#indicated by the "include" suffix. IF a sample has a 0 in the include column it's related data will be
#removed from the analysis
antigen.ref<-dplyr::select(mimosa, id, Cyt_include, TH1_include, TH2_include, TH1.2_include)

#read in the data set containing the actual phenotype measurements
phenotype<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/CD4_Positive_clean.csv")
#we don't have unstimulated data in the mimosa analysis by default so we remove it
phenotype<-dplyr::filter(phenotype, !(Stim=="UN"))
#we need matching ids so we create a new column that pastes the donor and the stim together to match the
#id variable in the antigen.ref data frame
phenotype$id<-paste(phenotype$Donor, phenotype$Stim, sep="_")
#again to match the other dataframe's nomenclature
names(phenotype)<-gsub("Cytokine", "Cyt", names(phenotype))

library(dplyr)
#arranging both data sets by the id variable that we will match on ensures that proper matching occurs
mimosa<-arrange(mimosa, id)
phenotype<-arrange(phenotype, id)
DF<-phenotype #so we don't alter the original phenotype dataframe just in case
keynames<-c("Cyt", "TH1", "TH1.2", "TH2") #list of character strings that contain our "key" values
#in the antigen ref we have columns named "key"_include and in the phenotype data we have columns named
#"key"_"parameter"

for(key in keynames){#loops through all the "keys" in the list and does the following for each "key"
    #creates a list of id names where the key_include column in antigen.ref is 0 -aka not a real response
    remove<-subset(antigen.ref, antigen.ref[,paste(key,"include",sep="_")]==0)$id
    print(key) #checks and balances so I can see it loop through the different keys
    print(length(remove)) #see how many sample ids will be excluded
    #this is a row column notation to index the DF data frame
    #rows are first and this is indexed to include the rows where the id variable is in the remove list
    #columns are second and this searches column names to include columns that include the "key" string
    #for each row column combo that this is true the cell is set to NA
    DF[(DF$id %in% remove), grep(paste(key,"_",sep=""), names(DF))]<-NA
}

#for me
for(key in keynames){
    print(key)
    #will bind together into a new dataframe the columns that contain the key in the name from
    #1. the original entries in the mimosa data frame, 2. the untouched phenotype data and 3. the new phenotype data with NA values
    test<-data.frame(cbind(mimosa[grep(key, names(mimosa))], 
                           phenotype[grep(paste(key,"_",sep=""), names(phenotype))],
                           DF[grep(paste(key,"_",sep=""), names(DF))]))
    #how many rows in the data have a 0 indicating they aren't a real response and have data in the corresponding columns of the phenotype data --> should be substantial
    print(length(which(test[,8]==0 &! is.na(test[,9:24]))))
        #how many rows in the data have a 0 indicating they aren't a real response and have data in the corresponding columns of the modified phenotype data --> should be 0
    print(length(which(test[,8]==0 &! is.na(test[,25:40]))))
}

phenotype<-dplyr::select(DF, -id)
write.csv(DF, "CD4_Phenotyping_clean.csv") #store for later
```

#P-values by disease status
```{r}
library(dplyr)
library(knitr)
#to see how many values are missing for each column
kable(colSums(!is.na(phenotype)))
data<-filter(phenotype, SM!="N")
data$SM <-factor(data$SM, levels=c("X","SM"))
data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
```

For these, we have to subset the data by stim since the extent to which a column's data is missing depends on the stim. NA values will really mess up the p value functions. Basically the function won't work at all if one of the 6 disease groups doesn't have any observations. It just aborts. I could fix it but I haven't yet. I've only annotated the PMA chunk but the other stims are all the same mechanism.

#PMA Phenotyping
```{r PMA}
library(dplyr)
library(knitr)
#since the extent of missingness varies by stim we have to do this each time
DF<-subset(data, data$Stim=="PMA")
#split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
x<-split(DF, list(DF$TB, DF$SM))
#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0. It's a weird double negative. But basically makes a list of the columns that have sufficient data.
y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
#each entry in the list has to be extracted and named
a<-unlist(y[1])
b<-unlist(y[2])
c<-unlist(y[3])
d<-unlist(y[5])
e<-unlist(y[6])
#then find the intersection of all the lists. Aka which column names have sufficient data in all 6 TB-SM groups
colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
colnames<-colnames[7:length(colnames)]
for(col in colnames){
    y<-pvals_helminth(DF, col)
    print(y)
}
```
#Peptide-Specific Phenotyping
```{r PEP}
library(dplyr)
library(knitr)
DF<-subset(data, data$Stim=="PEP")
x<-split(DF, list(DF$TB, DF$SM))
y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
a<-unlist(y[1])
b<-unlist(y[2])
c<-unlist(y[3])
d<-unlist(y[5])
e<-unlist(y[6])
colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
colnames<-colnames[7:length(colnames)]
for(col in colnames){
    y<-pvals_helminth(DF, col)
    print(y)
}
```
#WCL-Specific Phenotyping
```{r WCL}
library(dplyr)
library(knitr)
DF<-subset(data, data$Stim=="WCL")
x<-split(DF, list(DF$TB, DF$SM))
y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
a<-unlist(y[1])
b<-unlist(y[2])
c<-unlist(y[3])
d<-unlist(y[5])
e<-unlist(y[6])
colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
colnames<-colnames[7:length(colnames)]
for(col in colnames){
    y<-pvals_helminth(DF, col)
    print(y)
}
```
#SEA-Specific Phenotyping
```{r SEA}
library(dplyr)
library(knitr)
DF<-subset(data, data$Stim=="SEA")
x<-split(DF, list(DF$TB, DF$SM))
y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
a<-unlist(y[1])
b<-unlist(y[2])
c<-unlist(y[3])
d<-unlist(y[5])
e<-unlist(y[6])
colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
colnames<-colnames[7:length(colnames)]
for(col in colnames){
    y<-pvals_helminth(DF, col)
    print(y)
}
```
#SWAP-Specific Phenotyping
```{r SWAP}
library(dplyr)
library(knitr)
DF<-subset(data, data$Stim=="SWAP")
x<-split(DF, list(DF$TB, DF$SM))
y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
a<-unlist(y[1])
b<-unlist(y[2])
c<-unlist(y[3])
d<-unlist(y[5])
e<-unlist(y[6])
colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
colnames<-colnames[7:length(colnames)]
for(col in colnames){
    y<-pvals_helminth(DF, col)
    print(y)
}
```

#P-Values by Stim
In this case, the function is written such that I have to subset to look within each of the TB-SM subsets one by one. I'm working on writing the function to be able to handle it all. I've also included one example graph to see what I'm comparing.

```{r example ggplot}
library(ggplot2)
data<-filter(phenotype, SM!="N")
    data$SM<-factor(data$SM, levels=c("X","SM"))
    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
ggplot(data, aes(y=TH1_GATA3, x=Stim, col=TB))+
    geom_point(size=3, alpha=0.5)+geom_line(aes(group=Donor))+
    scale_color_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    theme_bw()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(SM~TB)+coord_cartesian(ylim = c(0,20))+
    labs(y= "Frequency of GATA3+ Mtb-Specific TH1 Cells")
```

#Healthy Controls SM-
```{r}
library(dplyr)
colnames<-names(dplyr::select(phenotype, -X, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(phenotype, TB=="HC" & SM=="X")
for(col in colnames){
    y<-pvals_stim(df, col)
    print(y)
}
```

#Healthy Controls SM+
```{r}
library(dplyr)
colnames<-names(dplyr::select(phenotype, -X, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(phenotype, TB=="HC" & SM=="SM")
for(col in colnames){
    y<-pvals_stim(df, col)
    print(y)
}
```

#LTBI SM-
```{r}
library(dplyr)
colnames<-names(dplyr::select(phenotype, -X, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(phenotype, TB=="LTBI" & SM=="X")
for(col in colnames){
    y<-pvals_stim(df, col)
    print(y)
}
```

#LTBI SM+
```{r}
library(dplyr)
colnames<-names(dplyr::select(phenotype, -X, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(phenotype, TB=="LTBI" & SM=="SM")
for(col in colnames){
    y<-pvals_stim(df, col)
    print(y)
}
```

#TB SM-
```{r}
library(dplyr)
colnames<-names(dplyr::select(phenotype, -X, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(phenotype, TB=="TB" & SM=="X")
for(col in colnames){
    y<-pvals_stim(df, col)
    print(y)
}
```

#TB SM+
```{r}
library(dplyr)
colnames<-names(dplyr::select(phenotype, -X, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(phenotype, TB=="TB" & SM=="SM")
for(col in colnames){
    y<-pvals_stim(df, col)
    print(y)
}
```

#P values by Stim and Helminth
I haven't written this up nicely yet but none of it is significant

### 2019-05-22 age-sex-working.R

@@ -1,0 +1,188 @@
+library(ggplot2)
+library(ggpubr)
+library(dplyr)
+
+DonorInfo<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/ICS_Donor_Info_clean.csv")
+DonorInfo<-filter(DonorInfo, SM %in% c("X", "SM"))
+DonorInfo$SM<-factor(DonorInfo$SM, levels=c("X", "SM"))
+DonorInfo$TB<-factor(DonorInfo$TB, levels=c("HC", "LTBI", "TB"))
+DonorInfo$disease<-paste(DonorInfo$TB, DonorInfo$SM, sep="_")
+DonorInfo$disease<-factor(DonorInfo$disease, levels=c("HC_X", "HC_SM",
+                                                      "LTBI_X", "LTBI_SM",
+                                                      "TB_X", "TB_SM"))
+my_comparisons <- list(c("HC_SM", "HC_X"), c("LTBI_SM", "LTBI_X"), c("TB_SM", "TB_X"), 
+                       c("HC_SM", "LTBI_SM"), c("TB_SM", "LTBI_SM"), c("HC_SM", "TB_SM"),
+                       c("HC_X", "LTBI_X"), c("TB_X", "LTBI_X"), c("HC_X", "TB_X"))
+
+IQR_vals <- function(datatable, column) {
+    library(data.table)
+    y<-split(datatable, datatable$disease)
+    x<-lapply(y, function(y) quantile(y[,column], probs=c(0.5, 0.25, 0.75), na.rm=TRUE))
+    #quantile is a stats function that takes a vector "x" and returns the value at a specified quantile
+    data.frame(t(x))
+}
+IQR_pvals<-function(data, column){
+    x<-split(data,data$TB)
+    #does a wilcox test between SM- and SM+ for whatever column you call in the column argument
+    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"], exact=FALSE, na.action = na.pass))
+    pvals<-c(A$HC$p.value, A$LTBI$p.value, A$TB$p.value)
+    fdr_adj_pvals<-round(p.adjust(pvals, method="fdr"), 4)
+    t(fdr_adj_pvals)         
+}
+
+
+###N
+n<-t(DonorInfo %>% dplyr::count(disease))
+names(n)<-n[1,]
+
+## AGE
[…]

### 2019-05-22 OG-Analysis.Rmd
basic Mtb CD4 analysis

---
title: "OG-Analysis"
author: "Taryn McLaughlin"
date: "5/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read in data files}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
CD4_summary<-read.csv("OG_CD4-bulk_summary.csv")
CD4_summary$SM<-factor(CD4_summary$SM, levels=c("X", "SM"))
CD4_OG_summary<-read.csv("OG_CD4-OG_summary.csv")
CD4_OG_summary$SM<-factor(CD4_OG_summary$SM, levels=c("X", "SM"))
PBMC<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/exported/OG_PBMC.csv")
```

```{r functions and stuff}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/Plotting_1902.R")
```

```{r PBMC}
PBMC<-filter(PBMC, !is.na(Study.ID))
PBMC<-filter(PBMC, Helminth.Status %in% c("Negative", "Schisto+"))
PBMC$Helminth.Status<-factor(PBMC$Helminth.Status, levels=c("Negative", "Schisto+"))
PBMC$TB.Status<-factor(PBMC$TB.Status, levels=c("Healthy Control", "LTBI", "Active TB"))
PBMC<-filter(PBMC, !(Study.ID=="NK2208" & Exp.Name=="TM181216"))
PBMC<-filter(PBMC, !(Study.ID=="NK2306" & Exp.Name=="TM181206"))
PBMC<-filter(PBMC, !(Study.ID=="NK2353" & Exp.Name=="TM190411"))
PBMC<-filter(PBMC, !(Study.ID=="NK2384" & Exp.Name=="TM190411"))
PBMC$TB<-PBMC$TB.Status
PBMC$SM<-PBMC$Helminth.Status

grid.arrange(plot_6(PBMC, "SM", "Viability")+labs(y="Viability"),
    plot_6(PBMC, "SM", "Cells.Vial")+labs(y="Cell Count"),
    ncol=2
)

plot_pairs<-function(data, xval, yval){
    ggplot(data, aes(x=data[,xval], y=data[,yval]))+
        geom_point(size=3, aes(color=TB, alpha=SM)) +
        geom_line(aes(group=Study.ID))+
        scale_color_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
        scale_alpha_manual(values=c(0.5,1))+
        facet_grid(~SM)+
        theme_bw()+ theme(legend.position="bottom")+
        theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
        theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
        facet_grid(SM~TB, scale="free")+
        stat_compare_means(label = "p.format", p.adjust.method = "fdr", hide.ns = TRUE, size = 4, 
                           method="wilcox.test", paired = TRUE, label.y= (max(data[,yval])*1.1))
}

melt<-melt(PBMC, id.vars=c("Study.ID", "TB", "SM"), measure.vars = c("Viability", "OG.Viability"))
plot_pairs(melt, "variable", "value")+labs(y="Viability", x="Time")

melt<-melt(PBMC, id.vars=c("Study.ID", "TB", "SM"), measure.vars = c("Cells.Vial", "OG.Cell.."))
plot_pairs(melt, "variable", "value")+labs(y="Count", x="Time")
```

```{r}
library(dplyr)
antigen.ref<-MIMOSA.Results
antigen.ref$id<-rownames(antigen.ref)
antigen.ref$id<-gsub("_Treatment", "", antigen.ref$id)
antigen.ref<-arrange(antigen.ref, id)

phenotype<-CD4_OG_summary
phenotype$id<-paste(phenotype$Donor, phenotype$Stim, sep="_")
phenotype<-arrange(phenotype, id)

remove<-subset(antigen.ref, antigen.ref[,"Pr.response"]<.7)$id
print(length(remove)) #see how many sample ids will be excluded
numeric.only <- function(X,...){
      returnCols <- names(X)
      a<-sapply(X, is.numeric)
      print(returnCols[a == "TRUE"])
}

phenotype[(phenotype$id %in% remove), numeric.only(phenotype)]<-NA
```

#Pep
```{r}
df<-filter(CD4_summary, Stim=="PEP")
top<-ggplot(df, aes(x=SM, y=OG_freq, fill=TB, alpha=SM))+
    geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    facet_grid(~TB)+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(axis.title.x=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank())+
    theme(text = element_text(size=20)) +   
    facet_grid(~TB)+
    labs(y="", x="")+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    stat_compare_means(label = "p.format")+
    coord_cartesian(ylim=c(2,30))

bottom<-ggplot(df, aes(x=SM, y=OG_freq, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB)) +
    geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    facet_grid(~TB)+
    theme_classic()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    theme(strip.background = element_blank(),strip.text.x = element_blank())+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
    facet_grid(~TB)+
    labs(y="", x="")+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    coord_cartesian(ylim=c(0,2))

grid.arrange(top, bottom, ncol=1, heights=c(4,8))

df<-filter(phenotype, Stim=="PEP", TB!="HC")
test<-filter(phenotype, Stim=="PEP", TB!="HC", !is.na(IFNg_freq))
kable(xtabs(~TB+SM, test))
df$SM<-factor(df$SM, levels=c("X", "SM"))
grid.arrange(plot_2(df, "SM", "IFNg_freq")+labs(y="IFNg+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "TNFa_freq")+labs(y="TNFa+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "IL4_freq")+labs(y="IL-4+ Proliferating CD4+ T cells (%)", x=""),
             ncol=3

)

grid.arrange(plot_2(df, "SM", "CXCR3_freq")+labs(y="CXCR3+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "CCR4_freq")+labs(y="CCR4+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "Tbet_freq")+labs(y="Tbet+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "GATA3_freq")+labs(y="GATA3+ Proliferating CD4+ T cells (%)", x=""),
             ncol=4

)
```

#WCL
```{r}
df<-filter(CD4_summary, Stim=="WCL")
top<-ggplot(df, aes(x=SM, y=OG_freq, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), 
        outlier.shape = NA, aes(fill=TB)) +geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    facet_grid(~TB)+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(axis.title.x=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank())+
    theme(text = element_text(size=20)) +   
    facet_grid(~TB)+
    labs(y="", x="")+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    stat_compare_means(label = "p.format")+
    coord_cartesian(ylim=c(5,50))

bottom<-ggplot(df, aes(x=SM, y=OG_freq, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), 
                 outlier.shape = NA, aes(fill=TB)) +
    geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    facet_grid(~TB)+
    theme_classic()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    theme(strip.background = element_blank(),strip.text.x = element_blank())+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
    facet_grid(~TB)+
    labs(y="", x="")+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    coord_cartesian(ylim=c(0,5))

grid.arrange(top, bottom, ncol=1, heights=c(4,6))

df<-filter(phenotype, Stim=="WCL", TB!="HC")
test<-filter(phenotype, Stim=="WCL", TB!="HC", !is.na(IFNg_freq))
kable(xtabs(~TB+SM, test))

grid.arrange(plot_2(df, "SM", "IFNg_freq")+labs(y="IFNg+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "TNFa_freq")+labs(y="TNFa+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "IL4_freq")+labs(y="IL-4+ Proliferating CD4+ T cells (%)", x=""),
             ncol=3

)

grid.arrange(plot_2(df, "SM", "CXCR3_freq")+labs(y="CXCR3+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "CCR4_freq")+labs(y="CCR4+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "Tbet_freq")+labs(y="Tbet+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "GATA3_freq")+labs(y="GATA3+ Proliferating CD4+ T cells (%)", x=""),
             ncol=4

)
```

### 2019-05-22 Broken Axis.R
Delete Broken Axis.R

@@ -1,25 +1,0 @@
-data<-Cytokine[Cytokine$Stim=="PMA"]
-data$mask = 0
-data$mask[data$TH2.Freq > 1.5] = 1
-data$mask = factor(data$mask, levels = c('1', '0'))
-
-library(ggplot2)
-base.plot <- function(DF, group1, group2, value) {
-        p <- ggplot(DF, aes(x=group1, y=value, col=group1))
-        p <- p + theme_bw()
-        p <- p + theme(legend.position="bottom")
-        p <- p + geom_jitter(width=.1,height=0, shape=1,size=2)
-        p <- p + scale_color_brewer(palette="Set1", guide=guide_legend(ncol=6, title=NULL))
-        p <- p + xlab("") + ylab("")
-        return(p)
-}
-
-p <- base.plot(DF=data, SM, TB, Value)
-
-p <- p + facet_grid(mask~TB, scales="free_y")
-p <- p + theme(strip.text.y = element_blank())
-#The faceting in the Y is backwards and I don't know why or how to change it
-#I have tried releveling the mask variable and that didn't work
-#And why is the bottom chunk huge compared to the top?
-#I tried using the shrink and the space options in facet_grid 
-#to make the bottom box smaller but it didn't work

### 2019-05-29 PBMC.Rmd
markdown for donor info and PBMC info

---
title: "PBMC_Donor"
author: "Taryn McLaughlin"
date: "5/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE, comment=NA)
knitr::opts_chunk$set(fig.height = 10.5, fig.width = 7.5)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
```

```{r}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
```

```{r viability and cell counts}
library(ggplot2)
Donor$TB[grepl("HD", Donor$StudyID)==TRUE]<-"Yerkes"
palette.four<-c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b")
grid.arrange(ggplot(Donor, aes(x=SM, y=Viability, fill=TB, alpha=SM)) +
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
    geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_fill_manual(values = c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    theme_bw()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
    facet_grid(~TB, scales="free")+
    labs(title="Viability", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Live Cells (%)"),

ggplot(Donor, aes(x=SM, y=Cells.Vial, fill=TB, alpha=SM)) +
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
    geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_fill_manual(values = c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    theme_bw()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
    facet_grid(~TB, scales="free")+
    labs(title="Cell Counts", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Cells/mL)")
)
```

```{r}
#This was to pick the most robust responders for something else so Idk
library(dplyr)
mimosa<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/CD4_MIMOSA_Complete.csv")
HD<-filter(mimosa, grepl("HD", mimosa$Sample))%>%
    filter(Stim!="PMA")
HDscores<-aggregate(score~Donor, HD, FUN=sum)
HDkeep<-head(arrange(HDscores, score), 10)
#382 479 487 are BCG
HDremove<-HD$Donor[!(HD$Donor %in% HDkeep$Donor)]
CD4<-dplyr::filter(CD4, !(Donor %in% HDremove))
MFI<-dplyr::filter(MFI, !(Donor %in% HDremove))
Compass<-dplyr::filter(Compass, !(Donor %in% HDremove))
Donor<-dplyr::filter(Donor, !(Donor %in% HDremove))
```

### 2019-05-29 Cytokine_Analysis.Rmd
cleaning up to source functions

--- 2018-09-26 Cytokine_Analysis.Rmd
@@
 ---
-title: "Cytokine_Analysis"
+title: "ICS_Cytokine_Analysis"
 author: "Taryn McLaughlin"
 date: "9/21/2018"
-output: pdf_document
+output: 
+  html_document: default
+  html_notebook: default
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-knitr::opts_chunk$set(warning = FALSE)
-knitr::opts_chunk$set(message = FALSE)
-knitr::opts_chunk$set(fig.height = 10.5, fig.width = 7.5)
-
+knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE, comment=NA)
 library(dplyr)
-library(reshape2)
-Cytokine<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/Background Subtracted Cytokine.csv")
-Cytokine <- mutate(Cytokine, G_plus = G_4_13_x + G_4_x_x + G_x_13_x + G_x_x_x) %>%
-    mutate(T_plus = x_4_13_T + x_4_x_T + x_x_13_T + x_x_x_T) %>%
-    mutate(G_T_plus = G_4_x_T + G_x_13_T + G_x_x_T)%>%
-    mutate(IL4_plus = G_4_x_T + G_4_x_x + x_4_x_T + x_4_x_x) %>%
-    mutate(IL13_plus = G_x_13_T + G_x_13_x + x_x_13_T + x_x_13_x) %>%
-    mutate(IL4_IL13_plus = G_4_13_x + x_4_13_T + x_4_13_x)
-Cytokine$TB<-factor(Cytokine$TB, levels=c("N","HC","LTBI","TB"))
-Cytokine$SM<-factor(Cytokine$SM, levels=c("N","X","SM"))
+library(knitr)
+library(ggplot2)
+library(gridExtra)
+library(GGally)
+library(ggpubr)
+library(cowplot)
 ```
 
-#Statistical Analysis
-I wrote a function called pvals which evaluates the relevant two sample two-tailed  p-values using the non-parametric Mann Whitney test (the command is wilcox.test in R).
-The function then corrects for multiple comparisons in two ways.
-The first uses  a Bonferonni correction to control the Family Wise Error Rate.
-The second controls the False Discovery Rate, the expected proportion of false discoveries amongst the rejected hypotheses.
-The false discovery rate is a less stringent condition than the family-wise error rate, so these methods are more powerful than the others.
-The function is below if you would like to see the details.
+```{r source}
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
+```
 
-```{r p-value function, echo=TRUE}
-pvals<-function(data, column){
+```{r load data, echo=FALSE}
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
 
-    data<-filter(data, SM!="N")
-    data$SM<-factor(data$SM, levels=c("X","SM"))
-    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
+CD4<-plot_filter(read.csv("ICS_CD4_Cytokine_clean.csv"))
 
-    x<-split(data,data$TB) 
-    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"]))
+CD8<-plot_filter(read.csv("ICS_CD8_Cytokine_clean.csv"))
 
-    y<-split(data,data$SM) 
-    B<-lapply(y, function(g) 
-        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "LTBI")))
-    C<- lapply(y, function(g) 
-        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("LTBI", "TB")))
-    D<- lapply(y, function(g) 
-        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "TB")))
+MFI<-plot_filter(read.csv("ICS_CD4_MFI_clean.csv"))
 
-    labels<-c("HC SM+ to HC SM-", "LTBI SM+ to LTBI SM-", "TB SM+ to TB SM-",
-              "HC SM+ to LTBI SM+", "LTBI SM+ to TB SM+", "HC SM+ to TB SM+",
-              "HC SM- to LTBI SM-", "LTBI SM- to TB SM-", "HC SM- to TB SM-"
-    )
-    pvals<-c(A$HC$p.value, A$LTBI$p.value, A$TB$p.value,
-             B$SM$p.value, C$SM$p.value,  D$SM$p.value,
-             B$X$p.value,  C$X$p.value,   D$X$p.value)
+Compass<-plot_filter(read.csv("ICS_Compass_scores_complete.csv"))
 
-    bonf_adj_pvals<-round(p.adjust(pvals, method="bonferroni"), 4)
+GD<-plot_filter(read.csv("ICS_GD_Cytokine_clean.csv"))
 
-    fdr_adj_pvals<-round(p.adjust(pvals, method="fdr"), 4)
+data<-read.csv("ICS_Donor_Info_clean.csv")
+data$TB<-gsub("BCG|Unknown", "Naïve", data$TB)
+data$TB<-factor(data$TB, levels=c("Naïve","Healthy Control","LTBI","Active TB"))
+data$SM<-gsub("US|Not US|Unknown", "Naïve", data$SM)
+data$SM<-factor(data$SM, levels=c("Naïve","Negative","Schisto+"))
+Donor<-data
+```
+#P-values for Cytokine Comparisons
 
-    table<-(cbind(labels, round(pvals, 4), bonf_adj_pvals, fdr_adj_pvals))
-    colnames(table)<-c("Comparison", paste(column,"p-values"," "), 
-                       "Bonferonni Corrected p-values", "FDR Corrected p-vales")
-    table
+##PMA & Ionomycin
+```{r}
+library(knitr)
+#creates a list of character strings for each of the columns of interest
+colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
+for(col in colnames){#this will loop through each character string in the list
+    a<-subset(CD4, CD4$Stim=="PMA") #subsets the data to only contain the unstimulated samples
+    #inserts each character string into the column argument of the pvals function written above
+    b<-pvals_6(a, col)
+    if(dim(b)[1]!=0)
+    print(kable(b, caption=col)) #will return the output for each character string in the loop
 }
+```
+
+##Whole Cell Lysate
+```{r}
+library(knitr)
+#creates a list of character strings for each of the columns of interest
+colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
+for(col in colnames){#this will loop through each character string in the list
+    a<-subset(CD4, CD4$Stim=="WCL") #subsets the data to only contain the unstimulated samples
+    #inserts each character string into the column argument of the pvals function written above
+    b<-pvals_6(a, col)
+    if(dim(b)[1]!=0)
+    print(kable(b, caption=col)) #will return the output for each character string in the loop
+}
+```
+
+##CFP-10 & ESAT-6
+```{r}
+library(knitr)
+#creates a list of character strings for each of the columns of interest
+colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
+for(col in colnames){#this will loop through each character string in the list
+    a<-subset(CD4, CD4$Stim=="PEP") #subsets the data to only contain the unstimulated samples
+    #inserts each character string into the column argument of the pvals function written above
+    b<-pvals_6(a, col)
+    if(dim(b)[1]!=0)
+    print(kable(b, caption=col)) #will return the output for each character string in the loop
+}
+```
+
+#Compass Graphs
+```{r}
+#CD4 Mtb Responses as measured by Compass
+DF<-dplyr::filter(Compass, TB != "N" & Stim == "WCL" & celltype == "CD4")
+
+p1<-plot_compass(DF, "FS")+
+    labs(y="Functionality Score")
+
+p2<-plot_compass(DF, "PFS")+
+    labs(y="Polyfunctionality Score")
+
+grid.arrange(p1, p2, ncol=2, top="Whole Cell Lysate Scores")
+
+DF<-dplyr::filter(Compass, TB != "N" & Stim == "PEP" & celltype == "CD4")
+
+p1<-plot_compass(DF, "FS")+
+    labs(y="Functionality Score")
+
+p2<-plot_compass(DF,"PFS")+
+    labs(y="Polyfunctionality Score")
+
+grid.arrange(p1, p2, ncol=2, top="Peptide Pool Scores")
+```
+
+```{r, fig.width=16, fig.height=12}
+library(dplyr)
+DF<-merge(Compass, dplyr::select(Donor, -X, -TB, -SM), by="Donor")
+ggpairs(dplyr::filter(DF, celltype=="CD4", Stim=="WCL", TB!="N")%>%
+            mutate(disease = paste(TB, SM, sep=" "))%>%
+            dplyr::select(age, sex, FS, PFS, disease),
+        aes(col=disease))
+
 ```
@@
 The first graph is just meant to give a snapshot of all the data at once to see if there are weird anomalous data points that we will have to examine.
-I did not include all the individual data points on any of the graphs because it made it incredibly cluttered looking, however I did allow outliers to display.
-
+I did not include all the individual data points on any of the graphs because it made it look incredibly cluttered, however I did allow outliers to display.
 ```{r Boolean Graphs, cache=TRUE, echo=FALSE, warning=FALSE}
 library(dplyr)
+library(tidyr)
 library(reshape2)
 library(ggplot2)
-Boolean<-select(Cytokine, Donor, TB, SM, Stim, G_4_13_T, G_4_x_T, G_x_13_T, G_4_13_x, x_4_13_T, G_4_x_x, G_x_13_x,
-                x_4_x_T, x_x_13_T, G_x_x_T , G_x_x_x, x_x_x_T,  x_4_13_x, x_4_x_x, x_x_13_x)
-
-melt<-melt(Boolean,id.vars=c("TB","SM", "Stim"))
-melt$value<-as.numeric(melt$value)
-melt<-filter(melt, !(is.na(value)))
-melt<-filter(melt, !(value<0))
-melt$label = factor(melt$variable,levels=c("G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", 
-                                           "G_x_13_x","x_4_x_T","x_x_13_T", "G_x_x_T" , "G_x_x_x","x_x_x_T",  
-                                           "x_4_13_x","x_4_x_x", "x_x_13_x"))
-melt<-filter(melt, !is.na(label))
-
-#all the boolean plots
-g<-ggplot(melt, aes(x=label, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    facet_grid(Stim~TB, scale="free")+
-    labs(title="Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="Cytokine Subsets",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-#subset by stim
-data<-filter(melt, Stim=="PMA")
-g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    facet_grid(TB~., scale="free")+
-    labs(title="PMA Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="Cytokine Subsets",
-         y="Frequency of Cytokine+ CD4+ T cells (%)");
-
-data<-filter(melt, Stim=="WCL")
-g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    facet_grid(TB~., scale="free")+
-    labs(title="WCL Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="Cytokine Subsets",
-         y="Frequency of Cytokine+ CD4+ T cells (%)");data<-filter(melt, Stim=="PEP")
-
-data<-filter(melt, Stim=="PEP")
-g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    facet_grid(TB~., scale="free")+
-    labs(title="PEP Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="Cytokine Subsets",
-         y="Frequency of Cytokine+ CD4+ T cells (%)");
-
-data<-filter(melt, Stim=="SEA")
-g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    facet_grid(TB~., scale="free")+
-    labs(title="SEA Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="Cytokine Subsets",
-         y="Frequency of Cytokine+ CD4+ T cells (%)");
-
-data<-filter(melt, Stim=="SWAP")
-g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    facet_grid(TB~., scale="free")+
-    labs(title="SWAP Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="Cytokine Subsets",
-         y="Frequency of Cytokine+ CD4+ T cells (%)");
+triple_boolean_plot(CD4, "PMA")
+triple_boolean_plot(CD4, "WCL")
+triple_boolean_plot(CD4, "PEP")
 ```
@@
 These are my standard lineage specific groups. 
-
-```{r TH, echo=FALSE, fig.cap="Breakdown of Cytokine Definitions"}
-knitr::include_graphics("/Applications/Old Computer/Day Lab/Data_Flow/TH_Group_Explanation.png")
-```
-
 ```{r Grouped TH Graphs, echo=FALSE}
@@
 library(knitr)
-TH<-select(Cytokine, Donor, TB, SM, Stim, Cyt.Freq, TH1.Freq, TH1.2.Freq, TH2.Freq)
-
-melt<-melt(TH,id.vars=c("TB","SM", "Stim"))
-melt$value<-as.numeric(melt$value)
-melt<-filter(melt, !(is.na(value)))
-melt<-filter(melt, !(value<0))
-melt$label = factor(melt$variable,levels=c("Cyt.Freq", "TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
-melt<-filter(melt, !is.na(label))
-
-##### PMA
-data<-filter(melt, Stim=="PMA")
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Non Specific CD4+ T cell Responses", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-data<-filter(TH, Stim=="PMA")
-kable(pvals(data, "TH1.Freq"), caption = "Non Specific CD4+ T cell Responses")
-kable(pvals(data, "TH2.Freq"))
-kable(pvals(data, "TH1.2.Freq"))
-
+colnames<-c("Cyt.Freq", "Th1.Freq", "TH1.2.Freq", "TH2.Freq")
+### PMA
+for(col in colnames){#this will loop through each character string in the list
+    #inserts each character string into the column argument of the pvals function written above
+    name<-gsub("\\.", "+", col)
+    name<-gsub("Freq","", name)
+    y<-plot_with_all_stats(CD4, "PMA", col)+
+        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
+             subtitle="PMA & Ionomycin",
+             x="",
+             y=paste(name, " CD4+ T cells (%)"))
+    print(y) #will return the output for each character string in the loop
+}
 ##### WCL
-data<-filter(melt, Stim=="WCL")
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-data<-filter(TH, Stim=="WCL")
-kable(pvals(data, "TH1.Freq"), caption = "Mtb Whole Cell Lysate Specific CD4 T cell Responses")
-kable(pvals(data, "TH2.Freq"))
-kable(pvals(data, "TH1.2.Freq"))
-
+for(col in colnames){#this will loop through each character string in the list
+    #inserts each character string into the column argument of the pvals function written above
+    name<-gsub("\\.", "+", col)
+    name<-gsub("Freq","", name)
+    y<-plot_with_all_stats(CD4, "WCL", col)+
+        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
+             subtitle="WCL",
+             x="",
+             y=paste(name, " CD4+ T cells (%)"))
+    print(y) #will return the output for each character string in the loop
+}
 ##### Peptide
-data<-filter(melt, Stim=="PEP")
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Mtb Peptide Specific CD4+ T cell Responses", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-data<-filter(TH, Stim=="PEP")
-kable(pvals(data, "TH1.Freq"), caption = "Mtb Peptide Specific CD4 T cell Responses")
-kable(pvals(data, "TH2.Freq"))
-kable(pvals(data, "TH1.2.Freq"))
-
+for(col in colnames){#this will loop through each character string in the list
+    #inserts each character string into the column argument of the pvals function written above
+    name<-gsub("\\.", "+", col)
+    name<-gsub("Freq","", name)
+    y<-plot_with_all_stats(CD4, "PEP", col)+
+        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
+             subtitle="PEP",
+             x="",
+             y=paste(name, " CD4+ T cells (%)"))
+    print(y) #will return the output for each character string in the loop
+}
 ```
@@
 library(knitr)
-cyt<-select(Cytokine, Donor, TB, SM, Stim, IFNg, TNFa, IL4, IL13)
+colnames<-c("IFNg", "TNFa", "IL4", "IL13")
 
-melt<-melt(cyt,id.vars=c("TB","SM", "Stim"))
-melt$value<-as.numeric(melt$value)
-melt<-filter(melt, !(is.na(value)))
-melt<-filter(melt, !(value<0))
-melt$label = factor(melt$variable,levels=c("IFNg", "TNFa", "IL4", "IL13"))
-melt<-filter(melt, !is.na(label))
+### PMA
+for(col in colnames){#this will loop through each character string in the list
+    #inserts each character string into the column argument of the pvals function written above
+    y<-plot_with_all_stats(CD4, "PMA", col)+
+        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
+             subtitle="PMA & Ionomycin",
+             x="",
+             y=paste(name, "+ CD4+ T cells (%)"))
+    print(y) #will return the output for each character string in the loop
+}
+####WCL
+for(col in colnames){#this will loop through each character string in the list
+    #inserts each character string into the column argument of the pvals function written above
+    name<-gsub("\\.", "+", col)
+    name<-gsub("Freq","", name)
+    y<-plot_with_all_stats(CD4, "PMA", col)+
+        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
+             subtitle="PMA & Ionomycin",
+             x="",
+             y=paste(name, "+ CD4+ T cells (%)"))
+    print(y) #will return the output for each character string in the loop
+}
 
-#### PMA
-data<-filter(melt, Stim=="PMA")
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Non-Specific CD4+ T cell Responses", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-data<-filter(cyt, Stim=="PMA")
-kable(pvals(data, "IFNg"), caption = "Non Specific CD4 T cell Responses")
-kable(pvals(data, "TNFa"))
-kable(pvals(data, "IL4"))
-kable(pvals(data, "IL13"))
-
-#### WCL
-data<-filter(melt, Stim=="WCL")
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-data<-filter(cyt, Stim=="WCL")
-kable(pvals(data, "IFNg"), caption = "Mtb Whole Cell Lysate Specific CD4 T cell Responses")
-kable(pvals(data, "TNFa"))
-kable(pvals(data, "IL4"))
-kable(pvals(data, "IL13"))
-
-#### Peptide
-data<-filter(melt, Stim=="PEP")
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Mtb Peptide Specific CD4+ T cell Responses", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-data<-filter(cyt, Stim=="PEP")
-kable(pvals(data, "IFNg"), caption = "Mtb Peptide Specific CD4 T cell Responses")
-kable(pvals(data, "TNFa"))
-kable(pvals(data, "IL4"))
-kable(pvals(data, "IL13"))
-
+### PEP
+for(col in colnames){#this will loop through each character string in the list
+    #inserts each character string into the column argument of the pvals function written above
+    name<-gsub("\\.", "+", col)
+    name<-gsub("Freq","", name)
+    y<-plot_with_all_stats(CD4, "PMA", col)+
+        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
+             subtitle="PMA & Ionomycin",
+             x="",
+             y=paste(name, "+ CD4+ T cells (%)"))
+    print(y) #will return the output for each character string in the loop
+}
 ```
-
-#True Single and Double Cytokine Positive CD4+ T Cells
-These are the frequencies of cells truly ONLY producing:  
-1. IFNg or TNFa or the combination of both AND NOT IL4 or IL13  
-2. IL-4 or IL-13 or the combination of both AND NOT IFNg or TNFa  
-These are the subgroups within the grouped TH1 vs TH2 lineage analysis. They are essentially the last 6 groups of the boolean data.
 
@@
 library(knitr)
-SP_lin<-rename(Cytokine, G_single =  G_x_x_x, T_single = x_x_x_T, IL4_single = x_4_x_x, IL13_single = x_x_13_x, 
-               G_T = G_x_x_T, IL4_IL13 = x_4_13_x) %>%
-    select(Donor, TB, SM, Stim, G_single, T_single, IL4_single, IL13_single, G_T, IL4_IL13)
 
-melt<-melt(SP_lin,id.vars=c("TB","SM", "Stim"))
-melt$value<-as.numeric(melt$value)
-melt<-filter(melt, !(is.na(value)))
-melt<-filter(melt, !(value<0))
-melt$label = factor(melt$variable,levels=c("G_single", "T_single", "G_T", "IL4_single", "IL13_single", "IL4_IL13"))
-melt<-filter(melt, !is.na(label))
-
-#### PMA
-# Th1 plots
-data<-filter(melt, Stim=="PMA" & label %in% c("G_single", "T_single", "G_T"))
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Non-Specific CD4+ Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-#TH2 plots
-data<-filter(melt, Stim=="PMA" & label %in% c("IL4_single", "IL13_single", "IL4_IL13"))
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Non-Specific CD4+ Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-data<-filter(SP_lin, Stim=="PMA")
-kable(pvals(data, "G_single"), caption = "Non-Specific CD4+ T cell Responses")
-kable(pvals(data, "T_single")) 
-kable(pvals(data,"G_T"))
-kable(pvals(data, "IL4_single"))
-kable(pvals(data, "IL13_single")) 
-kable(pvals(data, "IL4_IL13"))
-
-#### WCL
-# Th1 plots
-data<-filter(melt, Stim=="WCL" & label %in% c("G_single", "T_single", "G_T"))
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Mtb Whole Cell Lysate Specific CD4+ Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-#TH2 plots
-data<-filter(melt, Stim=="WCL" & label %in% c("IL4_single", "IL13_single", "IL4_IL13"))
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Mtb Whole Cell Lysate Specific CD4+ Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-data<-filter(SP_lin, Stim=="WCL")
-kable(pvals(data, "G_single"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
-kable(pvals(data, "T_single")) 
-kable(pvals(data,"G_T"))
-kable(pvals(data, "IL4_single"))
-kable(pvals(data, "IL13_single")) 
-kable(pvals(data, "IL4_IL13"))
-
-#### PEP
-# Th1 plots
-data<-filter(melt, Stim=="PEP" & label %in% c("G_single", "T_single", "G_T"))
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Mtb Peptide Specific CD4+ Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-#TH2 plots
-data<-filter(melt, Stim=="PEP" & label %in% c("IL4_single", "IL13_single", "IL4_IL13"))
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Mtb Peptide Specific CD4+ Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-data<-filter(SP_lin, Stim=="PEP")
-kable(pvals(data, "G_single"), caption = "Mtb Peptide Specific CD4 T cell Responses")
-kable(pvals(data, "T_single")) 
-kable(pvals(data,"G_T"))
-kable(pvals(data, "IL4_single"))
-kable(pvals(data, "IL13_single")) 
-kable(pvals(data, "IL4_IL13"))
-
-```
-
-\newpage
-
-#Lineage "Agnostic" Single and Double Cytokine Positive CD4+ T Cells
-I don't really believe that this is an appropriate way to look at the data because the cells are technically shared between some of these groupings, but I'm almost certain someone would ask the question "What if you didn't consider Il-4 and IL-13 in defining Gamma and TNF single and double positive cells" so I did it for them.
-
-```{r TH1, echo=FALSE, fig.cap="Breakdown of TH1 Cytokine Definitions"}
-knitr::include_graphics("/Applications/Old Computer/Day Lab/Data_Flow/TH1 plus explanations.png")
-```
-
-```{r TH2, echo=FALSE, fig.cap="Breakdown of TH2 Cytokine Definitions"}
-knitr::include_graphics("/Applications/Old Computer/Day Lab/Data_Flow/TH2 plus explanations.png")
-```
-
-```{r Single and Double Producers Lin Ag graphs, echo=FALSE}
-library(dplyr)
-library(reshape2)
-library(ggplot2)
-library(knitr)
-
-SP_Ag<-select(Cytokine, Donor, TB, SM, Stim, G_plus, T_plus, IL4_plus, IL13_plus, G_T_plus, IL4_IL13_plus)
-melt<-melt(SP_Ag,id.vars=c("TB","SM", "Stim"))
-melt$value<-as.numeric(melt$value)
-melt<-filter(melt, !(is.na(value)))
-melt<-filter(melt, !(value<0))
-melt$label = factor(melt$variable,levels=c("G_plus", "T_plus", "G_T_plus", "IL4_plus", "IL13_plus", "IL4_IL13_plus"))
-melt<-filter(melt, !is.na(label))
+colnames<-c("G_T" , "IL4_IL13" )
 
 ### PMA
-#TH1 plots
-data<-filter(melt, Stim=="PMA" & label %in% c("G_plus", "T_plus", "G_T_plus"))
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Non-Specific CD4+ T cell Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-#TH2 plots
-data<-filter(melt, Stim=="PMA" & label %in% c("IL4_plus", "IL13_plus", "IL4_IL13_plus"))
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Non-Specific CD4+ T cell Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
+for(col in colnames){#this will loop through each character string in the list
+    #inserts each character string into the column argument of the pvals function written above
+    name<-gsub("_", "+", col)
+    y<-plot_with_all_stats(CD4, "PMA", col)+
+        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
+             title="PMA & Ionomycin",
+             x="",
+             y=paste(name, "+ CD4+ T cells (%)"))
+    print(y) #will return the output for each character string in the loop
+}
+####WCL
+for(col in colnames){#this will loop through each character string in the list
+    #inserts each character string into the column argument of the pvals function written above
+    name<-gsub("_", "+", col)
+    y<-plot_with_all_stats(CD4, "PMA", col)+
+        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
+             subtitle="PMA & Ionomycin",
+             x="",
+             y=paste(name, "+ CD4+ T cells (%)"))
+    print(y) #will return the output for each character string in the loop
+}
 
-data<-filter(SP_Ag, Stim=="PMA")
-kable(pvals(data, "G_plus"), caption = "Non Specific CD4+ T cell Responses")
-kable(pvals(data, "T_plus"))
-kable(pvals(data, "G_T_plus"))
-kable(pvals(data, "IL4_plus"))
-kable(pvals(data, "IL13_plus"))
-kable(pvals(data, "IL4_IL13_plus"))
-
-### WCL
-#TH1 plots
-data<-filter(melt, Stim=="WCL" & label %in% c("G_plus", "T_plus", "G_T_plus"))
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-#TH2 plots
-data<-filter(melt, Stim=="WCL" & label %in% c("IL4_plus", "IL13_plus", "IL4_IL13_plus"))
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-data<-filter(SP_Ag, Stim=="WCL")
-kable(pvals(data, "G_plus"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
-kable(pvals(data, "T_plus"))
-kable(pvals(data, "G_T_plus"))
-kable(pvals(data, "IL4_plus"))
-kable(pvals(data, "IL13_plus"))
-kable(pvals(data, "IL4_IL13_plus"))
-
-### Peptide
-#TH1 plots
-data<-filter(melt, Stim=="PEP" & label %in% c("G_plus", "T_plus", "G_T_plus"))
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Mtb Peptide Specific CD4+ T cell Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-#TH2 plots
-data<-filter(melt, Stim=="PEP" & label %in% c("IL4_plus", "IL13_plus", "IL4_IL13_plus"))
-g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_x_discrete(breaks=c("N","X","SM"),
-                     labels=c("Naive", "SM(-)", "SM(+)"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(label~TB, scale="free")+
-    labs(title="Mtb Peptide Specific CD4+ T cell Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-data<-filter(SP_Ag, Stim=="PEP")
-kable(pvals(data, "G_plus"), caption = "Mtb Peptide Specific CD4+ T cell Responses")
-kable(pvals(data, "T_plus"))
-kable(pvals(data, "G_T_plus"))
-kable(pvals(data, "IL4_plus"))
-kable(pvals(data, "IL13_plus"))
-kable(pvals(data, "IL4_IL13_plus"))
+### PEP
+for(col in colnames){#this will loop through each character string in the list
+    #inserts each character string into the column argument of the pvals function written above
+    name<-gsub("_", "+", col)
+    y<-plot_with_all_stats(CD4, "PMA", col)+
+        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
+             subtitle="PMA & Ionomycin",
+             x="",
+             y=paste(name, "+ CD4+ T cells (%)"))
+    print(y) #will return the output for each character string in the loop
+}
 ```

### 2019-05-29 ICS_CD4.Rmd
pvals and graphs for total CD4

--- 2018-10-21 ICS_CD4.Rmd
@@
 ---
-title: "R Notebook"
+title: "Total CD4 Data"
 output:
@@
 ---
+This will be the markdown file looking at total CD4 phenotypes from the unstimulated condition
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE, comment=NA)
 library(dplyr)
-library(reshape2)
-Boolean<-select(Cytokine, Donor, TB, SM, Stim, G_4_13_T, G_4_x_T, G_x_13_T, G_4_13_x, x_4_13_T, G_4_x_x, G_x_13_x,
-                x_4_x_T, x_x_13_T, G_x_x_T , G_x_x_x, x_x_x_T,  x_4_13_x, x_4_x_x, x_x_13_x)
-
-melt<-melt(Boolean,id.vars=c("TB","SM", "Stim"))
-melt$value<-as.numeric(melt$value)
-melt<-filter(melt, !(is.na(value)))
-melt<-filter(melt, !(value<0))
-melt$label = factor(melt$variable,levels=c("G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T", "G_x_x_T" , "G_x_x_x","x_x_x_T",  "x_4_13_x","x_4_x_x", "x_x_13_x"))
-melt$TB<-factor(melt$TB, levels=c("N","HC","LTBI","TB"))
-melt$SM<-factor(melt$SM, levels=c("N","X","SM"))
-melt<-filter(melt, !is.na(label))
-
-data<-filter(melt, TB %in% c("HC", "N") & Stim=="PMA")
-
-g<-ggplot(melt, aes(x=label, y=value, fill=TB, alpha=SM))
-g + geom_boxplot(outlier.shape = NA, size=1) + 
-    geom_jitter(width=.1,height=0, shape=1,size=2) +
-    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#67001f"))+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    theme_bw()+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_grid(Stim~TB, scale="free")
-
-    scale_color_manual(labels = c("SM(-)", "SM(+)"), values = Palette) +
-    scale_x_discrete(breaks=c("X","SM"),
-                     labels=c("SM(-)", "SM(+)"))+
-    theme_bw()+
-    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
-    facet_wrap(~TB,nrow = 1, scale="free_x")+
-    labs(title="Cytokine Frequencies", 
-         subtitle="",
-         caption="",
-         x="TB and Helminth Status",
-         y="Frequency of Cytokine+ CD4+ T cells")
-
-```{r}
-#Antigen specific responses using total cytokine
-Cyt<-read.csv("Cyt.csv")
-TH1<-read.csv("TH1.csv")
-TH2<-read.csv("TH2.csv")
-TH1.2<-read.csv("TH1-2.csv")
-
-Split<- strsplit(as.character(Cyt$Sample)," ")
-Cyt$Sample = as.factor(sapply(Split,"[",2))
-TH1$Sample = as.factor(sapply(Split,"[",2))
-TH2$Sample = as.factor(sapply(Split,"[",2))
-TH1.2$Sample = as.factor(sapply(Split,"[",2))
-
-AG<-AG[!(AG$Sample=="NK2037_TB_HEL_T_PMA_020.fcs")]
-AG<-AG[!(AG$Sample=="NK2126_HC_X_SEA_005.fcs")]
-AG<-AG[!(AG$Sample=="NK2450_HC_SM_SEA_047.fcs")]
-AG<-AG[!(AG$Sample=="NK2402_LTBI_SM_SWAP_024.fcs")]
-AG<-AG[!(AG$Sample=="NK2342_TB_X_SWAP_006.fcs")]
-AG<-AG[!(AG$Sample=="NK2136_TB_X_WCL_003.fcs")]
-AG<-AG[!(AG$Sample=="NK2168_HC_SM_WCL_040.fcs")]
-AG<-AG[!(AG$Sample=="NK2088_LTBI_HEL_H_PEP_010.fcs")]
-
-AG<-as.data.table(AG)
-
-TH1<-merge(AG,TH1,by="Sample")
-TH2<-merge(AG,TH2,by="Sample")
-TH1.2<-merge(AG,TH1.2,by="Sample")
-Cyt<-merge(AG,Cyt,by="Sample")
-
-Cyt<-Cyt[,c(1:5,21:25,29,30,35:51)]
-TH1<-TH1[,c(1:5,21,24,26,31,35:51)]
-TH2<-TH2[,c(1:5,22:23,28,32,35:51)]
-TH1.2<-TH1.2[,c(1:5,21:24,27,33,35:51)]
-
+library(knitr)
+library(ggplot2)
+library(gridExtra)
+library(GGally)
+library(ggpubr)
+library(cowplot)
 ```
@@
 ```{r}
-
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
 ```
 
-Prep for Stacked Bars PMA as example
 ```{r}
-df<-WCL[,c(4,5,27:34)]
-df$TH1.Portion<-(df$TH1.Count/df$Cyt.Count)
-df$TH2.Portion<-(df$TH2.Count/df$Cyt.Count)
-df$TH1.2.Portion<-(df$TH1.2.Count/df$Cyt.Count)
-df$TH1.Port<-(df$TH1.Freq/df$Cyt.Freq)
-df$TH2.Port<-(df$TH2.Freq/df$Cyt.Freq)
-df$TH1.2.Port<-(df$TH1.2.Freq/df$Cyt.Freq)
-
-#unscaled frequencies
-df1<-df[,c(1,2,6,8,10)]
-df2<-melt(df1,id.vars=c("TB","SM"))
-df3<-as.data.frame(
-  ddply(df2, c("TB", "SM", "variable"), summarise,
-        median = median(value), sd = sd(value),
-        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
-df3$upper<-(df3$median+df3$error)
-df3$lower<-(df3$median-df3$error)
-
-#you have to do these to get the bars to be right
-df3$variable<-factor(df3$variable, levels=c('TH2.Freq','TH1.2.Freq','TH1.Freq',ordered=TRUE))
-df3$Label <- paste(df3$TB, df3$SM, sep='_')
-
-#adjusting error HC_X
-df3$upper[df3$variable=="TH1.2.Freq" & df3$Label=="HC_X"]<-
-  df3$median[df3$variable=="TH1.2.Freq" & df3$Label=="HC_X"]+
-  df3$median[df3$variable=="TH1.Freq" & df3$Label=="HC_X"]+
-  df3$error[df3$variable=="TH1.2.Freq" & df3$Label=="HC_X"]
-
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
+cd4<-read.csv("ICS_CD4_clean.csv")
+    #since we don't care about the yerkes donors in our analysis and they mess everything up
+    #we will filter them out
+    data<-filter(cd4, SM!="N")
+    #and then re-level both the SM and TB variable to not confuse our functions
+    data$SM<-factor(data$SM, levels=c("X","SM"))
+    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
+cd4<-data
 ```
 
-Boolean Cytokines with PMA as an example
+##Test and Return Significant Differences
+This will run through each column, and test the relevant comparisons for statistical significance.
+It also corrects when multiple comparisons are made.
+Since Healthy Controls are only compared internally, they do not get included in the multiple testing comparison.
+The function only returns the comparisons that have an initial p-value less than 0.05, which reduces clutter.
+Everything else is insignificant.
 ```{r}
-# to make the cytokine frequency graphs example with PMA
-library(dplyr)
-library(reshape2)
-Boolean<-select(data, Donor, TB, SM, G_4_13_T, G_4_x_T, G_x_13_T, G_4_13_x, x_4_13_T,G_4_x_x, G_x_13_x,x_4_x_T,x_x_13_T, G_x_x_T , G_x_x_x,x_x_x_T,  x_4_13_x,x_4_x_x, x_x_13_x)
-
-melt<-melt(Boolean,id.vars=c("TB","SM", "Stim"))
-melt$value<-as.numeric(melt$value)
-melt<-filter(melt, !(is.na(value)))
-melt<-filter(melt, !(value<0))
-melt$label = factor(melt$variable,levels=c("G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T", "G_x_x_T" , "G_x_x_x","x_x_x_T",  "x_4_13_x","x_4_x_x", "x_x_13_x"))
-
-HC<-subset(melt, TB %in% "HC")
-LTBI<-subset(melt, TB %in% "LTBI")
-ATB<-subset(melt, TB %in% "TB")
-
+library(knitr)
+#creates a list of character strings for each of the columns of interest
+colnames<-names(dplyr::select(cd4, -X, -Sample, -Donor, -Stim, -TB, -SM))
+for(col in colnames){#this will loop through each character string in the list
+    a<-subset(cd4, cd4$Stim=="UN") #subsets the data to only contain the unstimulated samples
+    #inserts each character string into the column argument of the pvals function written above
+    b<-pvals_6(a, col)
+    if(dim(b)[1]!=0)
+    print(kable(b, caption=col)) #will return the output for each character string in the loop
+}
 ```
 
-Surface Markers on CD4
-```{r}
-df1<-CD4[,c(7:9,19,20)]
-df2<-melt(df1,id.vars=c("TB","SM"))
-df3<-as.data.frame(
-  ddply(df2, c("TB", "SM", "variable"), summarise,
-        median = median(value), sd = sd(value),
-        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
-df3$upper<-(df3$median+df3$error)
-df3$lower<-(df3$median-df3$error)
+##Graph things that are significant
+Since SM- and SM+ are literally only different (p<0.05 after corrections) in Healthy Controls, I've graphed only those.
+It makes sense to me to graph it this way but am open to suggestions.
+We can start with a claim like "S mansoni on it's own has an influence on ____."
 
-#you have to do these to get the bars to be right
-df3$Label <- paste(df3$TB, df3$SM, sep='_')
-
-df3$variable<-factor(df3$variable, levels=c('CCR4.P_CXCR3.P','CCR4.N_CXCR3.P','CCR4.P_CXCR3.N', ordered=TRUE))
+####Option 1
+```{r, fig.width=8, fig.height=8}
+data<-filter(cd4, cd4$Stim=="UN", cd4$TB=="HC")
+a<-plot_hc(data, "SM", "CXCR3")+labs(title="Surface Marker Expression", subtitle="TH1")
+b<-plot_hc(data, "SM", "CCR4")+labs(title=" ", subtitle="TH2")
+c<-plot_hc(data, "SM", "Tbet")+labs(title="Transcription Factor Expression", subtitle="TH1")
+d<-plot_hc(data, "SM", "GATA3")+labs(title=" ",subtitle="TH2")
+plot_grid(a,b,c,d, ncol=2)
 ```
 
-Transcription Factors on CD4
-```{r}
-df1<-data[,c(22:24,4,5)]
-df2<-melt(df1,id.vars=c("TB","SM"))
-df3<-as.data.frame(
-  ddply(df2, c("TB", "SM", "variable"), summarise,
-        median = median(value), sd = sd(value),
-        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
-df3$upper<-(df3$median+df3$error)
-df3$lower<-(df3$median-df3$error)
-
-#you have to do these to get the bars to be right
-# (grey, neither, top), (cornflower, mix), (orchid, type 2), (navy, type 1, bottom)
-df3$Label <- paste(df3$TB, df3$SM, sep='_')
-df3$variable<-factor(df3$variable, levels=c("TH1_GATA3.P_Tbet.P",'TH1_GATA3.P_Tbet.N','TH1_GATA3.N_Tbet.P', ordered=TRUE))
-``` 
-Surface Markers on AG specific
-```{r}
-df<-TH1[TH1$Stim=="WCL"]
-df1<-df[,c(9:11,20,21)]
-df2<-melt(df1,id.vars=c("TB","SM"))
-df3<-as.data.frame(
-  ddply(df2, c("TB", "SM", "variable"), summarise,
-        median = median(value), sd = sd(value),
-        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
-df3$upper<-(df3$median+df3$error)
-df3$lower<-(df3$median-df3$error)
-
-#you have to do these to get the bars to be right
-# (grey, neither, top), (cornflower, mix), (orchid, type 2), (navy, type 1, bottom)
-df3$Label <- paste(df3$TB, df3$SM, sep='_')
-df3$variable<-factor(df3$variable, levels=c('CCR4.P_CXCR3.P','CCR4.N_CXCR3.P','CCR4.P_CXCR3.N', ordered=TRUE))
+####Option 2
+```{r, fig.width=8, fig.height=8}
+a<-plot_hc(data, "SM", "CXCR3")+labs(title="TH1", subtitle="Surface Marker")
+b<-plot_hc(data, "SM", "Tbet")+labs(title=" ", subtitle="Transcription Factor")
+c<-plot_hc(data, "SM", "CCR4")+labs(title="TH2", subtitle="Surface Marker")
+d<-plot_hc(data, "SM", "GATA3")+labs(title=" ", subtitle="Transcription Factor")
+plot_grid(a,b,c,d, ncol=2 )
 ```
 
-Transcription Factors on AG specific
+##All Possible Graphs
+For posterity... But I doubt we will include any of these
+
+###MFI
 ```{r}
-df<-TH1[TH1$Stim=="WCL"]
-df1<-df[,c(15:17,20,21)]
-df2<-melt(df1,id.vars=c("TB","SM"))
-df3<-as.data.frame(
-  ddply(df2, c("TB", "SM", "variable"), summarise,
-        median = median(value), sd = sd(value),
-        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
-df3$upper<-(df3$median+df3$error)
-df3$lower<-(df3$median-df3$error)
-
-#you have to do these to get the bars to be right
-# (grey, neither, top), (cornflower, mix), (orchid, type 2), (navy, type 1, bottom)
-df3$Label <- paste(df3$TB, df3$SM, sep='_')
-df3$variable<-factor(df3$variable, levels=c("GATA3.P_Tbet.P",'GATA3.P_Tbet.N','GATA3.N_Tbet.P', ordered=TRUE))
+for(col in colnames[1:4]){#this will loop through each character string in the list
+    x<-subset(cd4, cd4$Stim=="UN") #subsets the data to only contain the unstimulated samples
+    #inserts each character string into the column argument of the pvals function written above
+    y<-plot_6(x, "SM", col)+
+        labs(subtitle=paste(gsub("_", " ", col), "on CD4+ T cells"), 
+             caption="",
+             x="",
+             y=gsub("_", " ", col))
+    print(y) #will return the output for each character string in the loop
+}
 ```
 
-Marker Co-expression
+###Surface Markers
 ```{r}
-CD4_melt<-melt(data,id.vars=c("TB","SM"))
-CD4_melt$value<-as.numeric(CD4_melt$value)
-CD4_melt<-CD4_melt[!(is.na(CD4_melt$value))]
-CD4_stats<-as.data.frame(
-  ddply(CD4_melt, c("TB", "SM", "variable"), summarise,
-        mean = mean(value), sd = sd(value),
-        sem = sd(value)/sqrt(length(value))))
-
-TF<-subset(CD4_melt, variable %in% c("GATA3.N_Tbet.P", "GATA3.P_Tbet.N", "GATA3.P_Tbet.P"))
-tf<-subset(CD4_stats, variable %in% c("GATA3.N_Tbet.P", "GATA3.P_Tbet.N", "GATA3.P_Tbet.P"))
-CC<-subset(CD4_melt, variable %in% c("CCR4.N_CXCR3.P", "CCR4.P_CXCR3.N", "CCR4.P_CXCR3.P"))
-cc<-subset(CD4_stats, variable %in% c("CCR4.N_CXCR3.P", "CCR4.P_CXCR3.N", "CCR4.P_CXCR3.P","CCR4.N_CXCR3.N"))
+for(col in colnames[5:10]){#this will loop through each character string in the list
+    x<-subset(cd4, cd4$Stim=="UN") #subsets the data to only contain the unstimulated samples
+    #inserts each character string into the column argument of the pvals function written above
+    name<-gsub(".P", "+", col)
+    name<-gsub("_","", name)
+    name<-gsub(".N", "-", name)
+    y<-plot_6(x, "SM", col)+
+        labs(subtitle=paste(name, "(% CD4+ T cells)"), 
+             caption="",
+             x="",
+             y=paste(name, "(% CD4+ T cells)"))
+    print(y) #will return the output for each character string in the loop
+}
 ```
 
-stats? sort of read this http://rcompanion.org/rcompanion/d_08.html
+###Transcription Factors
+
+As you can see there is one crazy looking "outlier" in GATA3 from NK2586. 
 ```{r}
-data<-subset(CYT,CYT$Stim=="WCL")
-data<-filter(TH1,(Stim=="WCL" & TH1.Sum>0))
+for(col in colnames[11:16]){#this will loop through each character string in the list
+    x<-subset(cd4, cd4$Stim=="UN") #subsets the data to only contain the unstimulated samples
+    #inserts each character string into the column argument of the pvals function written above
+    name<-gsub(".P", "+", col)
+    name<-gsub("_","", name)
+    name<-gsub(".N", "-", name)
+    y<-plot_6(x, "SM", col)+
+        labs(subtitle=paste(name, "(% CD4+ T cells)"), 
+             caption="",
+             x="",
+             y=paste(name, "(% CD4+ T cells)"))
+    print(y) #will return the output for each character string in the loop
+}
+```
 
-x <- split(data, data$TB)
-lapply(x, function(g) wilcox.test(g$Cyt.Freq~g$SM))
-lapply(x, function(g) wilcox.test(g$TH1.Freq~g$SM))
-lapply(x, function(g) wilcox.test(g$TH1.2.Freq~g$SM))
-lapply(x, function(g) wilcox.test(g$TH2.Freq~g$SM))
-
-y <- split(data, data$SM)
-lapply(y, function(g) kruskal.test(g$Cyt.Freq~g$TB))
-lapply(y, function(g) kruskal.test(g$TH1.Freq~g$TB))
-lapply(y, function(g) kruskal.test(g$TH1.2.Freq~g$TB))
-lapply(y, function(g) kruskal.test(g$TH2.Freq~g$TB))
-
-data<-filter(TH1,(Stim=="WCL" & TH1.Sum>0))
-x <- split(data, data$TB)
-lapply(x, function(g) wilcox.test(g$TH1_Tbet~g$SM))
-lapply(x, function(g) wilcox.test(g$TH1_GATA3~g$SM))
-
-y <- split(data, data$SM)
-lapply(y, function(g) kruskal.test(g$TH1_Tbet~g$TB))
-lapply(y, function(g) kruskal.test(g$TH1_GATA3~g$TB))
+#Side Note
+This is the statistical evaluation of the data set without NK2586 the outlier GATA3 donor to see if it makes a difference.
+```{r}
+colnames2<-colnames[grep("GATA", colnames)]
+for(col in colnames2){#this will loop through each character string in the list
+    a<-filter(cd4, Stim=="UN", Donor!="NK2586") #subsets the data to only contain the unstimulated samples
+    #inserts each character string into the column argument of the pvals function written above
+    b<-pvals_6(a, col)
+    if(dim(b)[1]!=0)
+    print(kable(b, caption=col)) #will return the output for each character string in the loop
+}
 ```

### 2019-05-29 ICS_Cleaning.Rmd
includes edits from mismatched donors and new excluded sample Nk2586

@@
 #write.csv(df.list[1], file= newfilenames[1])
-write.csv(CD4_Counts.csv, file = "CD4_Counts_clean.csv")
-write.csv(CD4_Cytokine.csv, file = "CD4_Cytokine_clean.csv")
-write.csv(CD4_MFI.csv, file = "CD4_MFI_clean.csv")
-write.csv(CD4_Positive.csv, file = "CD4_Positive_clean.csv")
-write.csv(CD4.csv, file = "CD4_clean.csv")
-write.csv(CD8_Counts.csv, file = "CD8_Counts_clean.csv")
-write.csv(CD8_Cytokine.csv, file = "CD8_Cytokine_clean.csv")
-write.csv(GD_Counts.csv, file = "GD_Counts_clean.csv")
-write.csv(GD_Cytokine.csv, file = "GD_Cytokine_clean.csv")
+write.csv(CD4_Counts.csv, file = "ICS_CD4_Counts_clean.csv")
+write.csv(CD4_Cytokine.csv, file = "ICS_CD4_Cytokine_clean.csv")
+write.csv(CD4_MFI.csv, file = "ICS_CD4_MFI_clean.csv")
+write.csv(CD4_Positive.csv, file = "ICS_CD4_Positive_clean.csv")
+write.csv(CD4.csv, file = "ICS_CD4_clean.csv")
+write.csv(CD8_Counts.csv, file = "ICS_CD8_Counts_clean.csv")
+write.csv(CD8_Cytokine.csv, file = "ICS_CD8_Cytokine_clean.csv")
+write.csv(GD_Counts.csv, file = "ICS_GD_Counts_clean.csv")
+write.csv(GD_Cytokine.csv, file = "ICS_GD_Cytokine_clean.csv")
 write.csv(DonorInfo, file = "ICS_Donor_Info_clean.csv")

### 2019-05-30 Cytokine_Analysis.Rmd
Includes test about combining HC and LTBI

@@
 ```
-#P-values for Cytokine Comparisons
+
+#Can we use HC and LTBI together
+```{r}
+library(knitr)
+#creates a list of character strings for each of the columns of interest
+colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
+for(col in colnames){#this will loop through each character string in the list
+    a<-subset(CD4, CD4$Stim=="PMA") #subsets the data to only contain the unstimulated samples
+    #inserts each character string into the column argument of the pvals function written above
+    b<-pvals_hc(a, col)
+    print(kable(b, caption=col))
+
+    c<-filter(a, TB=="HC")
+    #will return the output for each character string in the loop
+    y<-plot_hc(c, "SM", col)
+    d<-filter(a, TB %in% c("HC", "LTBI"))
+    d$TB<-as.factor("combined")
+    x<-plot_hc(d, "SM", col)
+    grid.arrange(y,x, ncol=2)
+    }
+```
 

### 2019-05-30 ICS_CD4.Rmd
Includes test about combining HC and LTBI

@@
 cd4<-data
+```
+
+#Are HC and LTBI different or can we combine?
+
+```{r}
+library(knitr)
+#creates a list of character strings for each of the columns of interest
+colnames<-names(dplyr::select(cd4, -X, -Sample, -Donor, -Stim, -TB, -SM))
+for(col in colnames){#this will loop through each character string in the list
+    a<-subset(cd4, cd4$Stim=="UN") #subsets the data to only contain the unstimulated samples
+    #inserts each character string into the column argument of the pvals function written above
+    b<-pvals_hc(a, col)
+    print(kable(b, caption=col))
+
+    c<-filter(a, TB=="HC")
+    #will return the output for each character string in the loop
+    y<-plot_hc(c, "SM", col)
+    d<-filter(a, TB %in% c("HC", "LTBI"))
+    d$TB<-as.factor("combined")
+    x<-plot_hc(d, "SM", col)
+    grid.arrange(y,x, ncol=2)
+    }
 ```

### 2019-06-06 ICS_HC.Rmd
determining if HC and LTBI can be combined

---
title: "ICS_HC"
author: "Taryn McLaughlin"
date: "6/2/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE, comment=NA)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
```

```{r source}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")
```

```{r}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
cd4a<-read.csv("ICS_CD4_clean.csv")
cd4b<-read.csv("ICS_CD4_new.csv")

CD4<-plot_filter(read.csv("ICS_CD4_Cytokine_clean.csv"))
CD4b<-plot_filter(read.csv("ICS_CD4_Cytokine_new.csv"))

phen<-read.csv("ICS_CD4_Phenotyping_clean.csv")
phenb<-read.csv("ICS_CD4_phenotype_new_clean.csv")
```

#Are HC and LTBI different or can we combine?

##Lineage Markers on UN

```{r, fig.height=3, fig.width=6}
library(knitr)
colnames<-names(dplyr::select(cd4a, -X, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(cd4a, Stim=="UN")
df<-mutate_hc(df)
for(col in colnames){
    print(col)
    print(pvals_hc(df, col))
    grid.arrange(
       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
       ncol=2
    )
}
```

```{r, fig.height=3, fig.width=6}
colnames<-names(dplyr::select(cd4b, -X, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(cd4b, Stim=="UN")
df<-mutate_hc(df)
for(col in colnames){
    print(col)
    print(pvals_hc(df, col))
    grid.arrange(
       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
       ncol=2
    )
}
```

##Cytokine Analysis from PMA

```{r, fig.height=3, fig.width=6}
library(knitr)
colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(CD4, Stim=="PMA")
df<-mutate_hc(df)
for(col in colnames){
    print(col)
    print(pvals_hc(df, col))
    grid.arrange(
       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
       ncol=2
    )
}
```

```{r, fig.height=3, fig.width=6}
library(knitr)
colnames<-names(dplyr::select(CD4b, -X, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(CD4b, Stim=="PMA")
df<-mutate_hc(df)
for(col in colnames){
    print(col)
    print(pvals_hc(df, col))
    grid.arrange(
       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
       ncol=2
    )
}
```

#Phenotype from PMA
```{r, fig.height=3, fig.width=6}
library(knitr)
colnames<-names(dplyr::select(phen, -X.1, -X, -id, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(phen, Stim=="PMA")
df<-mutate_hc(df)
for(col in colnames){
    print(col)
    print(pvals_hc(df, col))
    grid.arrange(
       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
       ncol=2
    )
}
```

#Phenotype from PMA
```{r, fig.height=3, fig.width=6}
library(knitr)
colnames<-names(dplyr::select(phenb, -X.1, -X, -id, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(phenb, Stim=="PMA")
df<-mutate_hc(df)
for(col in colnames){
    print(col)
    print(pvals_hc(df, col))
    grid.arrange(
       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
       ncol=2
    )
}
```

```{r}
colnames<-names(dplyr::select(phen, -X, -X.1, -id, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){#this will loop through each character string in the list
    a<-subset(phen, phen$Stim=="PEP") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_6(a, col) 
    print(b)
    c<-plot_with_all_stats(a, "PEP", col)+labs(y=col)
    print(c)
     #will return the output for each character string in the loop
}

colnames<-names(dplyr::select(phenb, -X, -X.1, -id, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){#this will loop through each character string in the list
    a<-subset(phenb, phenb$Stim=="PEP") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_6(a, col) 
    print(b)
    c<-plot_with_all_stats(a, "PEP", col)+labs(y=col)
    print(c)
    #will return the output for each character string in the loop
}
```

### 2019-06-06 ICS-processing.R
Creating master ICS file from new exported flow data

@@ -1,0 +1,30 @@
+library(dplyr)
+
+setwd("/Users/tarynam/Desktop/ICS-CD4")
+filenames = list.files(pattern="csv")
+for(i in 1:length(filenames)){
+    print(filenames[i])
+    assign(filenames[i], read.table(filenames[i], header=TRUE, sep=";"))
+}
+
+l.df<-lapply(ls(pattern=".csv"), function(x) get(x))
+combined<-dplyr::bind_rows(l.df)
+
+#this has to be edited still 
+combined$Sample<-paste(combined$donor, combined$Stim, sep="_")
+remove<-c("NK2126_HC_X_SEA", 
+          "NK2450_HC_SM_SEA" , 
+          "NK2402_LTBI_SM_SWAP",
+          "NK2342_TB_X_SWAP",
+          "NK2136_TB_X_WCL",
+          "NK2168_HC_SM_WCL",
+          "NK2421_LTBI_SM_PEP",
+          "NK2063_TB_X_SEA",
+          "NK2115_LTBI_X_PEP",
+          "NK2115_LTBI_X_WCL",
+          "NK2171_HC_X_WCL",
+          "NK2162_HC_X_SEA")
+combined<-dplyr::filter(combined, !(Sample %in% remove))
+write.csv(combined, "/Users/tarynam/Desktop/ICS-CD4/clean-data/Master-ICS.csv")
+
+

### 2019-06-06 PBMC.Rmd
Including Oregon Green PBMC data

@@
 knitr::opts_chunk$set(fig.height = 10.5, fig.width = 7.5)
-library(dplyr)
+library(data.table)
 library(knitr)
@@
 library(cowplot)
+library(dplyr)
 ```
 
-```{r}
+```{r source scripts}
 source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
@@
 
-```{r viability and cell counts}
-library(ggplot2)
-Donor$TB[grepl("HD", Donor$StudyID)==TRUE]<-"Yerkes"
-palette.four<-c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b")
-grid.arrange(ggplot(Donor, aes(x=SM, y=Viability, fill=TB, alpha=SM)) +
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
-    geom_jitter(width=.1,height=0, shape=16,size=3)+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_fill_manual(values = c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
-    facet_grid(~TB, scales="free")+
-    labs(title="Viability", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Live Cells (%)"),
+```{r ICS prep}
+library(dplyr)
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/exported")
+#Import information about the samples from my lab notes
+ics<-read.csv("ICS_PBMC.csv")
+#remove samples that are other helminths since we dropped that group
+#Not US, US and Unknown are from the Yerkes donors
+    ics<-dplyr::filter(ics, Helminth.Status=="Schisto+" | Helminth.Status== "Negative" | Helminth.Status== "Not US" |  Helminth.Status=="US" | Helminth.Status== "Unknown")
+#take out all the stuff that isn't consistent like the "TM" in experiment name and the % in viability
 
-ggplot(Donor, aes(x=SM, y=Cells.Vial, fill=TB, alpha=SM)) +
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
-    geom_jitter(width=.1,height=0, shape=16,size=3)+
-    scale_alpha_manual(values=c(0.5,0.5,1))+
-    scale_fill_manual(values = c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
-    facet_grid(~TB, scales="free")+
-    labs(title="Cell Counts", 
-         subtitle="",
-         caption="",
-         x="TB and S. mansoni Status",
-         y="Cells/mL)")
+    ics$Viability <- gsub("%/%" , "", ics$Viability)
+    ics$Viability<-as.numeric(gsub("%" , "" , ics$Viability))
+    ics$Exp.Date <- gsub("TM", "", ics$Exp.Date)
+
+#Remove samples that are being excluded from the ICS analysis
+ics<- dplyr::filter(ics, !(Excluded=="EX"))%>%
+    dplyr::rename(TB = TB.Status, SM = Helminth.Status, Donor = Study.ID, Exp.Name=Exp.Date)%>% 
+    dplyr::select(Donor, TB, SM, Exp.Name, Viability, Cells.Vial, Excluded) #take only the columns you need
+
+ics$Donor<-tstrsplit(ics$Donor, "-")[[1]]
+
+ics$TB<-as.character(ics$TB)
+for(i in 1:length(ics$TB)){
+    if (ics$TB[i]=="Active TB") ics$TB[i]<-"TB"
+    else if (ics$TB[i]=="Healthy Control") ics$TB[i]<-"HC"
+    else if (ics$TB[i]=="LTBI") ics$TB[i]<-"LTBI"
+    else ics$TB[i]<-"N"
+}
+
+ics$SM<-as.character(ics$SM)
+for(i in 1:length(ics$SM)){
+    if (ics$SM[i]=="Negative") ics$SM[i]<-"X"
+    else if (ics$SM[i]=="Schisto+") ics$SM[i]<-"SM"
+    else ics$SM[i]<-"N"
+}
+
+write.csv(ics, "/Applications/Old Computer/Day Lab/Flow-Data/clean/ICS_DonorInfo_clean.csv")
+
+```
+
+```{r OG prep}
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/exported")
+og<-read.csv("OG_PBMC.csv")
+og<-rename(og, Donor=Study.ID, TB=TB.Status, SM=Helminth.Status)%>%
+    filter(Donor!="", Excluded!="EX")%>%
+    select(Donor, TB, SM, Exp.Name, Viability, Cells.Vial, OG.Viability, OG.Cell.., Excluded, Exp.Name, Cheryl)
+
+og$Donor<-tstrsplit(og$Donor, "-")[[1]]
+og$Donor<-gsub("TRU", "", og$Donor)
+
+og$TB<-as.character(og$TB)
+for(i in 1:length(og$TB)){
+    if (og$TB[i]=="Active TB") og$TB[i]<-"TB"
+    else if (og$TB[i]=="Healthy Control") og$TB[i]<-"HC"
+    else if (og$TB[i]=="LTBI") og$TB[i]<-"LTBI"
+    else og$TB[i]<-"N"
+}
+
+og$SM<-as.character(og$SM)
+for(i in 1:length(og$SM)){
+    if (og$SM[i]=="Negative") og$SM[i]<-"X"
+    else if (og$SM[i]=="Schisto+") og$SM[i]<-"SM"
+    else og$SM[i]<-"N"
+}
+
+write.csv(og, "/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_DonorInfo_clean.csv")
+```
+
+```{r combine}
+pbmc<-merge(ics, og, by=c("Donor", "TB", "SM"), all=TRUE)
+names(pbmc)<-c("Donor", "TB" , "SM" , 
+             "ICS_Exp.Name" , "ICS_Viability", "ICS_Cells.Vial", "ICS_Exluded",
+             "OG_Exp.Name" , "PreOG_Viability" , "PreOG_Cells.Vial" , "PostOG_Viability" , "PostOG_Cells.Vial" , "OG_Excluded" , "Cheryl") 
+```
+
+```{r Import Kenya data}
+#Import the clinical enrollment data from the cleaned merged epi file
+    QFT<- read.csv("/Applications/Old Computer/Epi Project/Data_clean/complete-kenya-filledin.csv")
+#The studyIDs of coenrolled people are NK#_TRU# and we want just the NK# to match the lab notes
+    QFT$Donor<-tstrsplit(QFT$StudyID, "_")[[1]]
+#trim the data set to only those study IDs in the lab notes and select the columns you want
+    QFT<-dplyr::select(QFT, Donor, age, sex, QFT, SchistosomaIntensity)%>%
+        dplyr::filter(Donor %in% pbmc$Donor)
+```
+
+```{r merge all}
+
+#Merge the two tables together and keep all observations since the yerkes donors won't be in the epi data
+    DonorInfo<-merge(pbmc, QFT, by="Donor", all = TRUE)
+#Manual Edits that may get deleted after Kenya script is finalized
+    DonorInfo$TB[DonorInfo$Donor=="NK2183"]<-as.character("HC")
+    DonorInfo$QFT[DonorInfo$Donor=="NK2323"]<-0.04
+    DonorInfo$SM[DonorInfo$Donor=="NK2325"]<-as.character("Unknown")
+
+#Used in
+    DonorInfo$used<-NA
+    for(i in 1:length(DonorInfo$used)){
+        if(!is.na(DonorInfo$ICS_Exp.Name[i]) & !is.na(DonorInfo$OG_Exp.Name[i]))
+            DonorInfo$used[i]<-"ICS/OG"
+        else if (!is.na(DonorInfo$ICS_Exp.Name[i]) & is.na(DonorInfo$OG_Exp.Name[i]))
+            DonorInfo$used[i]<-"ICS only"
+        else if (is.na(DonorInfo$ICS_Exp.Name[i]) & !is.na(DonorInfo$OG_Exp.Name[i]))
+            DonorInfo$used[i]<-"OG only"
+        else DonorInfo$used[i]<-NA
+    }
+write.csv(DonorInfo, "/Applications/Old Computer/Day Lab/Flow-Data/clean/DonorInfo_clean.csv")
+
+```
+
+```{r viability and cell counts, fig.height=5, fig.width=10}
+pbmc$SM<-gsub("SM", "SM+", pbmc$SM)
+pbmc$SM<-gsub("X", "SM-", pbmc$SM)
+pbmc$SM<-factor(pbmc$SM, levels=c("N", "SM-", "SM+"))
+pbmc$TB<-factor(pbmc$TB, levels=c("N", "HC", "LTBI", "TB"))
+
+grid.arrange(
+    plot_pbmc(pbmc, "ICS_Viability")+labs(y="Viability"),
+    plot_pbmc(pbmc, "ICS_Cells.Vial")+labs(y="Cells/Vial"),
+    top=text_grob("ICS PBMCs", size=20), ncol=2
+)
+
+grid.arrange(
+    plot_pbmc(pbmc, "PreOG_Viability")+labs(y="Viability"),
+    plot_pbmc(pbmc, "PreOG_Cells.Vial")+labs(y="Cells/Vial"),
+    top=text_grob("Pre OG Staining", size=20), ncol=2
+)
+
+grid.arrange(
+    plot_pbmc(pbmc, "PreOG_Viability")+labs(y="Viability"),
+    plot_pbmc(pbmc, "PreOG_Cells.Vial")+labs(y="Cells/Vial"),
+    top=text_grob("Post OG Staining", size=20), ncol=2
 )
@@
 ```{r}
+melted<-melt(pbmc, id.vars = c("Donor", "TB", "SM"), 
+             measure.vars = c("PreOG_Viability" , "PostOG_Viability"))
+plot_pbmc.pairs(melted)
+melted<-melt(pbmc, id.vars = c("Donor", "TB", "SM"), 
+             measure.vars = c("PreOG_Cells.Vial", "PostOG_Cells.Vial"))
+plot_pbmc.pairs(melted)
+```
+
+```{r, eval=FALSE}
 #This was to pick the most robust responders for something else so Idk

### 2019-06-06 ICS_Cleaning.Rmd
includes and merges new export

--- 2019-05-22 ICS_Cleaning.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE, message = FALSE )
-```
-
-```{r Donor Info}
-library(dplyr)
-setwd("/Applications/Old Computer/Day Lab/Flow-Data/exported")
-#Import information about the samples from my lab notes
-    PBMC<-read.csv("ICS_PBMC.csv")
-#remove samples that are other helminths since we dropped that group
-#Not US, US and Unknown are from the Yerkes donors
-    PBMC<-dplyr::filter(PBMC, Helminth.Status=="Schisto+" | Helminth.Status== "Negative" | Helminth.Status== "Not US" |  Helminth.Status=="US" | Helminth.Status== "Unknown")
-    PBMC$Helminth.Status<-factor(PBMC$Helminth.Status, levels=c("US", "Not US", "Negative", "Schisto+", "Unknown"))
-#order the levels of TB status to be from least to most exposure/disease
-    PBMC$TB.Status = factor(PBMC$TB.Status,levels=c("Naïve","BCG","Healthy Control", "LTBI", "Active TB", "Unknown"))
-#take out all the stuff that isn't consistent like the "TM" in experiment name and the % in viability
-    PBMC$Viability <- gsub("%/%" , "", PBMC$Viability)
-    PBMC$Viability<-gsub("%" , "" , PBMC$Viability)
-    PBMC$Exp.Date <- gsub("TM", "", PBMC$Exp.Date)
-#Some of the samples had weird suffixes that have to be removed to match the epi data  
-    PBMC$Study.ID<-gsub("- plasma", "", PBMC$Study.ID)
-    PBMC$Study.ID<-gsub("-1", "", PBMC$Study.ID)
-    PBMC$Study.ID<-gsub("-0", "", PBMC$Study.ID)
-#Remove samples that are being excluded from the ICS analysis
-PBMC<- dplyr::filter(PBMC, !(Excluded=="EX"))%>%
-    dplyr::rename(TB = TB.Status, SM = Helminth.Status, StudyID = Study.ID)%>% #match the variable names to ICS variable names
-    dplyr::select(StudyID, TB, SM, Viability, Cells.Vial, Excluded) #take only the columns you need
-
-#Import the clinical enrollment data from the cleaned merged epi file
-    QFT<- read.csv("/Applications/Old Computer/Epi Project/Data_clean/complete-kenya-filledin.csv")
-#The studyIDs of coenrolled people are NK#_TRU# and we want just the NK# to match the lab notes
-    QFT$StudyID<-tstrsplit(QFT$StudyID, "_")[[1]]
-#trim the data set to only those study IDs in the lab notes and select the columns you want
-    QFT2<-dplyr::select(QFT, StudyID, age, sex, QFT, SchistosomaIntensity)%>%
-        dplyr::filter(StudyID %in% PBMC$StudyID)
-
-#Merge the two tables together and keep all observations since the yerkes donors won't be in the epi data
-    DonorInfo<-merge(PBMC, QFT2, by="StudyID", all = TRUE)
-#The class of variables gets messed up in the merge so re-class them
-    DonorInfo$Cells.Vial<-as.numeric(as.character(DonorInfo$Cells.Vial))
-    DonorInfo$QFT<-as.numeric(as.character(DonorInfo$QFT))
-#Match the variable names of the ICS data
-    DonorInfo<-dplyr::rename(DonorInfo, Donor = StudyID)
-    DonorInfo$TB<-gsub("Active TB", "TB", DonorInfo$TB)
-    DonorInfo$TB<-gsub("Healthy Control", "HC", DonorInfo$TB)
-    DonorInfo$SM<-gsub("Negative", "X", DonorInfo$SM)
-    DonorInfo$SM<-gsub("Schisto+", "SM", DonorInfo$SM)
-    DonorInfo$SM<-gsub("\\+", "", DonorInfo$SM)
-
-#Manual Edits that may get deleted after Kenya script is finalized
-    DonorInfo$TB[DonorInfo$Donor=="NK2183"]<-as.character("HC")
-    DonorInfo$QFT[DonorInfo$Donor=="NK2323"]<-0.04
-    DonorInfo$SM[DonorInfo$Donor=="NK2325"]<-as.character("Unknown")
 ```
@@
 
-l<-list(CD4_Counts.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_Positive.csv, CD4.csv, 
-        CD8_Counts.csv, CD8_Cytokine.csv, GD_Counts.csv, GD_Cytokine.csv)
+l<-list(CD4_Counts.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_new.csv,
+        CD4_phenotype_new.csv, CD4_phenotype.csv, CD4.csv, 
+        CD8_Counts.csv, CD8_Cytokine.csv, 
+        GD_Counts.csv, GD_Cytokine.csv
+        )
 l<-lapply(l, function(x) clean_flowjo_output(x))
 l<-lapply(l, function(x) splitter(x, "Sample", "_", one="Donor", two="TB", three="SM", four="Stim"))
+l<-lapply(l, function(x) fix.names(x))
 l<-lapply(l, function(x) remove_samples(x))
-l<-lapply(l, function(x) fix.names(x))
+l<-lapply(l, function(x) remove_donors_ICS(x))
 l<-lapply(l, function(x) class_change(x))
@@
 names(GD_Cytokine.csv)<-newnames
+
+CD4_Cytokine.csv <- dplyr::rename(CD4_Cytokine.csv, TH1.Freq = Th1.Freq)
 ```
@@
 
+#creating variables for TH1 and TH2 that include co-prouducing cells
+CD4_Counts.csv$TH1.Total<-(CD4_Counts.csv$TH1 + CD4_Counts.csv$TH1.2)
+CD4_Counts.csv$TH2.Total<-(CD4_Counts.csv$TH2 + CD4_Counts.csv$TH1.2)
+
+#Creating a total counts column
+CD4_Counts.csv$Total<-CD4_Counts.csv$x_x_x_x + CD4_Counts.csv$Cyt
+
 #then remove them from the other table
@@
 
-```{r create new combo cytokine variables}
+```{r update and organize cytokine variables}
 #If you want to know the double producers without considering the opposite TH cytokines
 library(dplyr)
-filenames = c("CD4_Cytokine.csv", "CD8_Cytokine.csv", "GD_Cytokine.csv")
-l<-list(CD4_Cytokine.csv, CD8_Cytokine.csv, GD_Cytokine.csv)
-l<-lapply(l, function(x) dplyr::mutate(x, G_T = G_4_x_T + G_x_13_T + G_x_x_T,
-                   IL4_IL13 = G_4_13_x + x_4_13_T + x_4_13_x)
+CD4_Cytokine.csv<-dplyr::mutate(CD4_Cytokine.csv, G_T = G_4_x_T + G_x_13_T + G_x_x_T,
+                   IL4_IL13 = G_4_13_x + x_4_13_T + x_4_13_x,
+                   TH1.Total = TH1.Freq + TH1.2.Freq,
+                   TH2.Total = TH2.Freq + TH1.2.Freq
                    )
-for (i in 1:length(filenames)) {
-    assign(filenames[i], 
-           data.frame(l[i]))}
+```
+
+```{r update and merge phenotype data}
+CD4<-merge(CD4.csv, CD4_new.csv, by=c("Sample","Donor","TB","SM","Stim" ), all=TRUE)
+CD4_phenotype<-merge(CD4_phenotype.csv, CD4_phenotype_new.csv, by=c("Sample","Donor","TB","SM","Stim" ), all=TRUE)
 ```
@@
 setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
-df.list<-list(CD4_Counts.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_Positive.csv,
-              CD4.csv, CD8_Counts.csv, CD8_Cytokine.csv, GD_Counts.csv,
-              GD_Cytokine.csv)
-dfnames<-list("CD4_Counts", "CD4_Cytokine", "CD4_MFI", "CD4_Positive",
-              "CD4", "CD8_Counts", "CD8_Cytokine", "GD_Counts",
-              "GD_Cytokine")
-newfilenames<-paste(dfnames, "clean", date(), ".csv", sep="_")
-#write.csv(df.list[1], file= newfilenames[1])
-write.csv(CD4_Counts.csv, file = "CD4_Counts_clean.csv")
-write.csv(CD4_Cytokine.csv, file = "CD4_Cytokine_clean.csv")
-write.csv(CD4_MFI.csv, file = "CD4_MFI_clean.csv")
-write.csv(CD4_Positive.csv, file = "CD4_Positive_clean.csv")
-write.csv(CD4.csv, file = "CD4_clean.csv")
-write.csv(CD8_Counts.csv, file = "CD8_Counts_clean.csv")
-write.csv(CD8_Cytokine.csv, file = "CD8_Cytokine_clean.csv")
-write.csv(GD_Counts.csv, file = "GD_Counts_clean.csv")
-write.csv(GD_Cytokine.csv, file = "GD_Cytokine_clean.csv")
-write.csv(DonorInfo, file = "ICS_Donor_Info_clean.csv")
+write.csv(CD4_Counts.csv, file = "ICS_CD4_Counts_clean.csv")
+write.csv(CD4_Cytokine.csv, file = "ICS_CD4_Cytokine_clean.csv")
+write.csv(CD4_MFI.csv, file = "ICS_CD4_MFI_clean.csv")
+write.csv(CD4_phenotype, file = "ICS_CD4_phenotype_clean.csv")
+write.csv(CD4, file = "ICS_CD4_clean.csv")
+write.csv(CD8_Counts.csv, file = "ICS_CD8_Counts_clean.csv")
+write.csv(CD8_Cytokine.csv, file = "ICS_CD8_Cytokine_clean.csv")
+write.csv(GD_Counts.csv, file = "ICS_GD_Counts_clean.csv")
+write.csv(GD_Cytokine.csv, file = "ICS_GD_Cytokine_clean.csv")
 ```

### 2019-06-10 cross_check_donorinfo.R
to check that flow data labels match donor info

@@ -1,0 +1,23 @@
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
+filenames =list.files(pattern="OG")
+
+pbmc<-read.csv("DonorInfo_clean.csv")
+pbmc<-dplyr::select(pbmc, Donor, TB, SM, used, OG_Excluded)%>%
+    dplyr::filter(grepl("OG", pbmc$used))
+
+test<-read.csv("OG_GDonly_OG_summary_clean.csv")
+#Selecting and renaming variables
+test2<-dplyr::select(test, Donor, Stim, TB, SM)
+
+inter<-intersect(pbmc$Donor, test$Donor)
+pbmc$Donor[!(pbmc$Donor %in% inter)]
+test$Donor[!(test$Donor %in% inter)]
+
+compare<-merge(test2, pbmc, by="Donor", all=TRUE)
+
+for(i in 1:length(compare)){compare[[i]]<-as.character(compare[[i]])}
+length(which(compare$TB.x != compare$TB.y))
+length(which(compare$SM.x != compare$SM.y))
+
+
+

### 2019-06-10 ICS-processing.R
boolean data gates added for markers and cytokines

@@ -9,22 +9,6 @@ for(i in 1:length(filenames)){
 
 l.df<-lapply(ls(pattern=".csv"), function(x) get(x))
 combined<-dplyr::bind_rows(l.df)
-
-#this has to be edited still 
-combined$Sample<-paste(combined$donor, combined$Stim, sep="_")
-remove<-c("NK2126_HC_X_SEA", 
-          "NK2450_HC_SM_SEA" , 
-          "NK2402_LTBI_SM_SWAP",
-          "NK2342_TB_X_SWAP",
-          "NK2136_TB_X_WCL",
-          "NK2168_HC_SM_WCL",
-          "NK2421_LTBI_SM_PEP",
-          "NK2063_TB_X_SEA",
-          "NK2115_LTBI_X_PEP",
-          "NK2115_LTBI_X_WCL",
-          "NK2171_HC_X_WCL",
-          "NK2162_HC_X_SEA")
-combined<-dplyr::filter(combined, !(Sample %in% remove))
 write.csv(combined, "/Users/tarynam/Desktop/ICS-CD4/clean-data/Master-ICS.csv")
 
 

### 2019-06-10 Cytokine_Analysis.Rmd
Took PMA out to put in HC script

--- 2019-05-29 Cytokine_Analysis.Rmd
@@
 source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")
 ```
@@
 CD4<-plot_filter(read.csv("ICS_CD4_Cytokine_clean.csv"))
+CD4b<-plot_filter(read.csv("ICS_Cytokine_new.csv"))
 
@@
 ```
-#P-values for Cytokine Comparisons
 
@@
     #inserts each character string into the column argument of the pvals function written above
-    b<-pvals_6(a, col)
+    b<-pvals_4(a, col)
     if(dim(b)[1]!=0)
@@
     #inserts each character string into the column argument of the pvals function written above
-    b<-pvals_6(a, col)
+    b<-pvals_4(a, col)
+    if(dim(b)[1]!=0)
+    print(kable(b, caption=col)) #will return the output for each character string in the loop
+}
+colnames<-c("TH1_total", "TH2_total")
+for(col in colnames){#this will loop through each character string in the list
+    a<-subset(CD4b, CD4b$Stim=="WCL") #subsets the data to only contain the unstimulated samples
+    #inserts each character string into the column argument of the pvals function written above
+    b<-pvals_4(a, col)
     if(dim(b)[1]!=0)
@@
     #inserts each character string into the column argument of the pvals function written above
-    b<-pvals_6(a, col)
+    b<-pvals_4(a, col)
+    if(dim(b)[1]!=0)
+    print(kable(b, caption=col)) #will return the output for each character string in the loop
+}
+colnames<-c("TH1_total", "TH2_total")
+for(col in colnames){#this will loop through each character string in the list
+    a<-subset(CD4b, CD4b$Stim=="PEP") #subsets the data to only contain the unstimulated samples
+    #inserts each character string into the column argument of the pvals function written above
+    b<-pvals_4(a, col)
     if(dim(b)[1]!=0)
@@
              x="",
-             y=paste(name, "+ CD4+ T cells (%)"))
+             y=paste(col, "+ CD4+ T cells (%)"))
     print(y) #will return the output for each character string in the loop
@@
     name<-gsub("Freq","", name)
-    y<-plot_with_all_stats(CD4, "PMA", col)+
+    y<-plot_with_all_stats(CD4, "WCL", col)+
         labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
-             subtitle="PMA & Ionomycin",
+             subtitle="WCL",
              x="",
-             y=paste(name, "+ CD4+ T cells (%)"))
+             y=paste(col, "+ CD4+ T cells (%)"))
     print(y) #will return the output for each character string in the loop
@@
     name<-gsub("Freq","", name)
-    y<-plot_with_all_stats(CD4, "PMA", col)+
+    y<-plot_with_all_stats(CD4, "PEP", col)+
         labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
-             subtitle="PMA & Ionomycin",
+             subtitle="PEP",
              x="",
-             y=paste(name, "+ CD4+ T cells (%)"))
+             y=paste(col, "+ CD4+ T cells (%)"))
     print(y) #will return the output for each character string in the loop
@@
     name<-gsub("_", "+", col)
-    y<-plot_with_all_stats(CD4, "PMA", col)+
+    y<-plot_with_all_stats(CD4, "WCL", col)+
         labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
-             subtitle="PMA & Ionomycin",
+             subtitle="WCL",
              x="",
-             y=paste(name, "+ CD4+ T cells (%)"))
+             y=paste(col, "+ CD4+ T cells (%)"))
     print(y) #will return the output for each character string in the loop
@@
     name<-gsub("_", "+", col)
-    y<-plot_with_all_stats(CD4, "PMA", col)+
+    y<-plot_with_all_stats(CD4, "WCL", col)+
         labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
-             subtitle="PMA & Ionomycin",
+             subtitle="PEP",
              x="",
-             y=paste(name, "+ CD4+ T cells (%)"))
+             y=paste(col, "+ CD4+ T cells (%)"))
     print(y) #will return the output for each character string in the loop
@@
 ```
+
+```{r}
+library(dplyr)
+library(reshape2)
+library(ggplot2)
+library(knitr)
+
+colnames<-c("TH1_Total" , "TH2_Total" )
+
+### PMA
+for(col in colnames){#this will loop through each character string in the list
+    #inserts each character string into the column argument of the pvals function written above
+    name<-gsub("_", "+", col)
+    y<-plot_with_all_stats(CD4b, "PMA", col)+
+        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
+             title="PMA & Ionomycin",
+             x="",
+             y=paste(col, "+ CD4+ T cells (%)"))
+    print(y) #will return the output for each character string in the loop
+}
+```

### 2019-06-10 ICS_CD4.Rmd
took out HC stuff and put it in separate script

--- 2019-05-29 ICS_CD4.Rmd
@@
 setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
-cd4<-read.csv("ICS_CD4_clean.csv")
-    #since we don't care about the yerkes donors in our analysis and they mess everything up
-    #we will filter them out
-    data<-filter(cd4, SM!="N")
-    #and then re-level both the SM and TB variable to not confuse our functions
-    data$SM<-factor(data$SM, levels=c("X","SM"))
-    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
-cd4<-data
+cd4a<-read.csv("ICS_CD4_clean.csv")
+cd4b<-read.csv("ICS_CD4_new.csv")
+```
+
+#Are HC and LTBI different or can we combine?
+```{r, fig.height=3, fig.width=6}
+library(knitr)
+colnames<-names(dplyr::select(cd4a, -X, -Sample, -Donor, -Stim, -TB, -SM))
+df<-filter(cd4a, Stim=="UN")
+df<-mutate_hc(df)
+for(col in colnames){
+    print(col)
+    print(pvals_hc(df, col))
+    grid.arrange(
+       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
+       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
+       ncol=2
+    )
+}
+```
+
+```{r}
+colnames<-names(dplyr::select(cd4b, -X, -Sample, -Donor, -Stim, -TB, -SM))
+df<-filter(cd4b, Stim=="UN")
+df<-mutate_hc(df)
+for(col in colnames){
+    print(col)
+    print(pvals_hc(df, col))
+    grid.arrange(
+       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
+       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
+       ncol=2
+    )
+}
 ```

### 2019-06-10 ICS_Cleaning.Rmd
matched up TB/SM/donor nomenclature

@@
 
+```{r ICS compass}
+ICS_Compass_scores_complete.csv<-changeTB(ICS_Compass_scores_complete.csv)
+ICS_Compass_scores_complete.csv<-changeHD(ICS_Compass_scores_complete.csv)
+ICS_Compass_scores_complete.csv<-fix.names(ICS_Compass_scores_complete.csv)
+ICS_Compass_scores_complete.csv<-remove_samples(ICS_Compass_scores_complete.csv)
+ICS_Compass_scores_complete.csv<-remove_donors(ICS_Compass_scores_complete.csv)
+write.csv(ICS_Compass_scores_complete.csv,
+          "/Applications/Old Computer/Day Lab/Flow-Data/clean/ICS_compass_complete_clean.csv"
+)
+```
+
 ```{r, reformat all data tables}
@@
 l<-lapply(l, function(x) remove_samples(x))
-l<-lapply(l, function(x) remove_donors_ICS(x))
+l<-lapply(l, function(x) remove_donors(x))
 l<-lapply(l, function(x) class_change(x))
 l<-lapply(l, function(x) changeTB(x))
+l<-lapply(l, function(x) changeHD(x))
 
@@
 
+```{r remove specific CD8/GD tables}
+pbmc<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/ICS_DonorInfo_clean.csv")
+cd8.remove<-filter(pbmc, grepl("CD8", Excluded))$Donor
+CD8_Counts.csv<-filter(CD8_Counts.csv, !(Donor %in% cd8.remove))
+CD8_Cytokine.csv<-filter(CD8_Cytokine.csv, !(Donor %in% cd8.remove))
+
+gd.remove<-filter(pbmc, grepl("CD4-", Excluded))$Donor
+GD_Counts.csv<-filter(GD_Counts.csv, !(Donor %in% cd8.remove))
+GD_Cytokine.csv<-filter(GD_Cytokine.csv, !(Donor %in% cd8.remove))
+
+```
+
 ```{r write files to "clean" directory}

### 2019-06-10 PBMC.Rmd
double checked with primary files so this is all correct

@@
 og<-read.csv("OG_PBMC.csv")
-og<-rename(og, Donor=Study.ID, TB=TB.Status, SM=Helminth.Status)%>%
-    filter(Donor!="", Excluded!="EX")%>%
-    select(Donor, TB, SM, Exp.Name, Viability, Cells.Vial, OG.Viability, OG.Cell.., Excluded, Exp.Name, Cheryl)
+og<-dplyr::rename(og, Donor=Study.ID, TB=TB.Status, SM=Helminth.Status)%>%
+    dplyr::filter(Donor!="", Excluded!="EX")%>%
+    dplyr::select(Donor, TB, SM, Exp.Name, Viability, Cells.Vial, OG.Viability, OG.Cell..,
+                  Excluded, Exp.Name, Cheryl)
 
@@
 names(pbmc)<-c("Donor", "TB" , "SM" , 
-             "ICS_Exp.Name" , "ICS_Viability", "ICS_Cells.Vial", "ICS_Exluded",
+             "ICS_Exp.Name" , "ICS_Viability", "ICS_Cells.Vial", "ICS_Excluded",
              "OG_Exp.Name" , "PreOG_Viability" , "PreOG_Cells.Vial" , "PostOG_Viability" , "PostOG_Cells.Vial" , "OG_Excluded" , "Cheryl") 
@@
 #Merge the two tables together and keep all observations since the yerkes donors won't be in the epi data
-    DonorInfo<-merge(pbmc, QFT, by="Donor", all = TRUE)
-#Manual Edits that may get deleted after Kenya script is finalized
+    DonorInfo <- merge(pbmc, QFT, by="Donor", all = TRUE)
     DonorInfo$TB[DonorInfo$Donor=="NK2183"]<-as.character("HC")
-    DonorInfo$QFT[DonorInfo$Donor=="NK2323"]<-0.04
-    DonorInfo$SM[DonorInfo$Donor=="NK2325"]<-as.character("Unknown")
+    DonorInfo <- filter(DonorInfo, Donor!="NK2325")
+    DonorInfo$SM[DonorInfo$Donor=="NK2384"]<-as.character("X")
 
@@
 HDremove<-HD$Donor[!(HD$Donor %in% HDkeep$Donor)]
-CD4<-dplyr::filter(CD4, !(Donor %in% HDremove))
-MFI<-dplyr::filter(MFI, !(Donor %in% HDremove))
-Compass<-dplyr::filter(Compass, !(Donor %in% HDremove))
-Donor<-dplyr::filter(Donor, !(Donor %in% HDremove))
 ```

### 2019-06-12 Exploratory Pairs Scatterplots.R
cleaning up

@@ -1,32 +1,0 @@
-# another option
-makePairs <- function(data) 
-{
-        grid <- expand.grid(x = 1:ncol(data), y = 1:ncol(data))
-        grid <- subset(grid, x != y)
-        all <- do.call("rbind", lapply(1:nrow(grid), function(i) {
-                xcol <- grid[i, "x"]
-                ycol <- grid[i, "y"]
-                data.frame(xvar = names(data)[ycol], yvar = names(data)[xcol], 
-                           x = data[, xcol], y = data[, ycol], data)
-        }))
-        all$xvar <- factor(all$xvar, levels = names(data))
-        all$yvar <- factor(all$yvar, levels = names(data))
-        densities <- do.call("rbind", lapply(1:ncol(data), function(i) {
-                data.frame(xvar = names(data)[i], yvar = names(data)[i], x = data[, i])
-        }))
-        list(all=all, densities=densities)
-}
-
-# expand iris data frame for pairs plot
-gg1 = makePairs(CD4[,3:18])
-
-# new data frame mega iris
-mega_CD4 = data.frame(gg1$all, Species=rep(CD4$TB, length=nrow(gg1$all)))
-
-# pairs plot
-ggplot(mega_CD4, aes_string(x = "x", y = "y")) + 
-        facet_grid(xvar ~ yvar, scales = "free") + 
-        geom_point(aes(colour=Species), na.rm = TRUE, alpha=0.8) + 
-        stat_density(aes(x = x, y = ..scaled.. * diff(range(x)) + min(x)), 
-                     data = gg1$densities, position = "identity", 
-                     colour = "grey20", geom = "line")

### 2019-06-12 EVC copy.R
cleaning up

@@ -1,154 +1,0 @@
-setwd("/Users/oldmac/Desktop/Day Lab/Data")
-NK_ICS<-read.csv("NK_ICS.csv")
-OG<-read.csv("EVC_OG.csv")
-IFN<-read.csv("ELISA_IFN.csv")
-IL13<-read.csv("ELISA_IL13.csv")
-
-##ELISA###
-IFN$Cytokine<-as.character("IFNg")
-IL13$Cytokine<-as.character("13")
-setnames(IFN,old="Dilution.Factor",new="Dilution")
-ELISA<-rbind(IFN,IL13)
-splStudyID = strsplit(as.character(ELISA$Sample)," ");
-ELISA$Donor = as.factor(sapply(splStudyID,"[",1));
-ELISA$Antigen <- as.factor(sapply(splStudyID, "[", 2));
-ELISA$Result[ELISA$Result>500]<-500
-ELISA$Adjusted<-(ELISA$Result*ELISA$Dilution)
-
-## combine and melt
-melted<-melt(ELISA, id.vars=c("Donor", "Antigen", "Cytokine"))
-#for some reason it stores all the values as character so coerce them to numeric to run stats
-melted$value<-as.numeric(as.character(melted$value))
-stats<-as.data.frame(
-  ddply(melted, c("Antigen", "Donor", "Cytokine", "variable"), summarise,
-        mean = mean(value), sd = sd(value),
-        sem = sd(value)/sqrt(length(value))))
-
-# re-separate data sets
-ELISA.Mean<-subset(stats,stats$variable=="Adjusted")
-ELISA.Mean<-ELISA.Mean[!("Antigen"=="Pep")]
-
-
-ELISA.Mean[,"Group"]<-NA
-ELISA.Mean$Group[ELISA.Mean$Donor=="HD235"]<-as.character("TH2")
-ELISA.Mean$Group[ELISA.Mean$Donor=="HD355"]<-as.character("HD")
-ELISA.Mean$Group[ELISA.Mean$Donor=="HD203"]<-as.character("HD")
-ELISA.Mean$Group[ELISA.Mean$Donor=="PS16-1001"]<-as.character("SA")
-ELISA.Mean$Group[ELISA.Mean$Donor=="PS16-1002"]<-as.character("SA")
-ELISA.Mean$Group[is.na(ELISA.Mean$Group)]<-as.character("SM")
[…]

### 2019-06-12 Merging 3 Columns in R For Loop.R

@@ -1,43 +1,0 @@
-TB.1<-c("LTBI", "LTBI", "LTBI", "LTBI", "LTBI", 
-        "ACTIVE", "ACTIVE", "ACTIVE", "HEALTHY CONTROL", "HEALTHY CONTROL",
-        NA,NA,NA,NA,NA)
-TB.2<-c(NA, NA, "LTBI", "LTBI", "HEALTHY CONTROL", 
-        "ACTIVE", "ACTIVE", NA, "HEALTHY CONTROL", "HEALTHY CONTROL",
-        "HEALTHY CONTROL","ACTIVE","LTBI",NA,NA)
-TB.3<-c(NA, "ACTIVE", NA, NA, "HEALTHY CONTROL", 
-        NA, NA, NA,"HEALTHY CONTROL", NA,
-        "HEALTHY CONTROL","ACTIVE","LTBI","LTBI", "HEALTHY CONTROL")
-library(data.table)
-TB<-as.data.table(cbind(TB.1,TB.2,TB.3))
-TB[,"TB.Final"]<-NA
-
-###I'm trying to write a for-loop that combines the information in columns 1-3
-#such that if TB.1 is not NA then TB.Final takes that value
-#if TB.1 is NA then TB.Final takes the value in TB.2
-#if TB.2 is NA then TB.Final takes the value in TB.3
-
-#This one doesn't incorporate the information in TB.3 without an error message
-for (i in 1:length(TB$TB.Final)){
-        if(!is.na(TB$TB.1[i])){
-                TB$TB.Final[i] <- TB$TB.1[i]}
-        
-        else if(!is.na(TB$TB.2[i])){
-                TB$TB.Final[i] <- TB$TB.2[i]} 
-        
-        else{
-                TB$TB.Final[i] <- TB$TB.3[i]} 
-}
-
-#This one also doesn't incorporate the information in TB.3 without an error message
-for (i in 1:length(TB$TB.Final)){
-        if(!is.na(TB$TB.1[i])){
-                TB$TB.Final[i] <- TB$TB.1[i]}
-        
-        else if(is.na(TB$TB.1[i])){
-                TB$TB.Final[i] <- TB$TB.2[i]} 
-        
-        else if (is.na(TB$TB.1[i] & is.na(TB$TB.2[i]))){
-                TB$TB.Final[i] <- TB$TB.3[i]} 
-}
-
-

### 2019-06-12 Trying to make NA values in specific cells.R

@@ -1,42 +1,0 @@
-names<-c("ID1","ID2","ID3", "ID4", "ID5", "ID6", "ID7", "ID8", "ID9", "ID10")
-data<-matrix(data=rep(1:8, 5), nrow=10)
-data<-as.data.table(cbind(names,data))
-names(data)<-c("ID","CXCR3_MFI","blank","blank2","CXCR3_Freq")
-
-#I'm super lost on how to combine various tools to achieve the following
-#I want to make certain values of columns that contain the string "CXCR3" NA
-#based on the ID in the ID column
-
-#I know I can select columns containing "CXCR3" using contains in dplyr
-library(dplyr)
-df1<-data %>% select(contains("CXCR3"))
-#but I can't combine that with "ID" these both give errors
-df1<-data %>% select(contains("CXCR3|ID"))
-df1<-data %>% select(contains("CXCR3") | contains("ID"))
-
-#So then I tried a million variations of grep commands
-#This one gives me a 3 rows all columns
-df2 <- data [grep("CXCR3|ID", names(data))]
-#This one has an error
-df3 <- data [grepl("CXCR3|ID", names(data))]
-#This one give me a logical vector of all FALSE
-df4 <- grepl("ID | CXCR3", names(data))
-
-#this gives me the correct column names but I can never seem to subset the data based on this
-colnames<-names(data)[grep("ID|CXCR3", names(data))]
-df<-select(data, (names(data) %in% colnames))
-df<-data[(names(data) %in% colnames)]
-df<-subset(data, names(data) %in% colnames)
-
-#So then I gave up but this is what I want to do 
-#but without having to type out every column name <-NA
-important_cols = grep('CXCR3', names(data))
-for (i in 1:length(data)){
-        if(data$ID[i] %in% c("ID1", "ID4", "ID5")){
-                for(j in important_cols){
-                        data[i,j] <- NA;
-                }        
-        }
-}
-
-

### 2019-06-12 multiple correlation analysis.R

@@ -1,17 +1,0 @@
-multiple correlation analysis
-
-mosthighlycorrelated <- function(mydataframe,numtoreport)
-{
-  # find the correlations
-  cormatrix <- cor(mydataframe)
-  # set the correlations on the diagonal or lower triangle to zero,
-  # so they will not be reported as the highest ones:
-  diag(cormatrix) <- 0
-  cormatrix[lower.tri(cormatrix)] <- 0
-  # flatten the matrix into a dataframe for easy sorting
-  fm <- as.data.frame(as.table(cormatrix))
-  # assign human-friendly names
-  names(fm) <- c("First.Variable", "Second.Variable","Correlation")
-  # sort and print the top n correlations
-  head(fm[order(abs(fm$Correlation),decreasing=T),],n=numtoreport)
-}

### 2019-06-12 remove-samples-from-key-table.R

@@ -1,77 +1,0 @@
-# All of this generates a fake data table that contains a "key" variable (i.e. donor)
-# a condition variable (i.e. "stim") and a set of columns of output data
-# the goal is for every donor, in each column of output data subtract out the value in that column
-# that corresponds to a specific value in the condition column
-donor<-c(rep("john", 6), rep("mary", 6), rep("jack", 6), rep("suzy", 6))
-stim<-rep(c("un","w","p","s","sw","sb"),4)
-Cyt<-sample(20:200, 24)
-TH1<-sample(20:200, 24)
-TH2<-sample(20:200, 24)
-TH1.2<-sample(20:200, 24)
-
-Cyt_ifng<-sample(100:1000, 24)
-Cyt_tnfa<-sample(100:1000, 24)
-Cyt_IL4<-sample(100:1000, 24)
-Cyt_IL13<-sample(100:1000, 24)
-TH1_ifng<-sample(100:1000, 24)
-TH1_tnfa<-sample(100:1000, 24)
-TH1_IL4<-sample(100:1000, 24)
-TH1_IL13<-sample(100:1000, 24)
-TH2_ifng<-sample(100:1000, 24)
-TH2_tnfa<-sample(100:1000, 24)
-TH2_IL4<-sample(100:1000, 24)
-TH2_IL13<-sample(100:1000, 24)
-TH1.2_ifng<-sample(100:1000, 24)
-TH1.2_tnfa<-sample(100:1000, 24)
-TH1.2_IL4<-sample(100:1000, 24)
-TH1.2_IL13<-sample(100:1000, 24)
-
-fake_count<-data.frame(cbind(donor,stim, Cyt, TH1, TH2, TH1.2))
-fake_mfi<-data.frame(cbind(donor,stim,Cyt_ifng,Cyt_tnfa,Cyt_IL4,Cyt_IL13,TH1_ifng,TH1_tnfa,
-                           TH1_IL4,TH1_IL13,TH2_ifng,TH2_tnfa,TH2_IL4,TH2_IL13,TH1.2_ifng,TH1.2_tnfa,
-                           TH1.2_IL4,TH1.2_IL13))
-remove(donor, stim, Cyt, TH1, TH2, TH1.2,Cyt_ifng,Cyt_tnfa,Cyt_IL4,Cyt_IL13,TH1_ifng,TH1_tnfa,
[…]

### 2019-06-12 subtract-background.R

@@ -1,42 +1,0 @@
-#used to generate a list of character strings that are the names of the columns containing numeric data
-numeric.only <- function(X,...){
-    returnCols <- names(X)
-    a<-sapply(X, is.numeric)
-    print(returnCols[a == "TRUE"])
-}
-
-#takes a dataframe (v. important its a dataframe), splitvar is a character string of a column name
-#colnms is a list of strings indicating which columns to subtract background from
-#condition is a character string with the key to indicate what is being subtracted
-background.subtract<-function(df, splitvar, keyvar, condition){
-    dt<-split(df, df[, splitvar]) #generate a list of mini data frames for each level of splitvar
-    #split --> lists in alphabetical order not by original key order
-    #To solve this I arranged the original data frame in alphabetical order
-    #can we instead arrange the split to match the original order
-    
-    colnms<-numeric.only(df)
-    
-    #for each column name in the list of column names
-    for(col in colnms){
-        #this is a check to make sure it goes through all the columns
-        print(col) 
-        #dt.r will make a list of length of dt where each item in the list is the new column for a particular donor
-        dt.r<-lapply(dt, function(g)
-            #take the values in each column and subtract out the value indicated by an "un" the stim column 
-            (g[,col] - subset(g, g[,keyvar]==condition)[,col]))
-        #unlist the list so you get a single column and insert it back into the original data frame
-        df[,col]<-unlist(dt.r)
-    }
-    df
-}
-
-filenames = c("CD4_Cytokine.csv", "CD8_Cytokine.csv", "GD_Cytokine.csv")
-l<-list(CD4_Cytokine.csv, CD8_Cytokine.csv, GD_Cytokine.csv)
-l<-lapply(l, function(x) arrange(x, Donor))
-l<-lapply(l, function(x) background.subtract(x, "Donor", "Stim", "UN"))
-for (i in 1:length(filenames)) {
-    assign(filenames[i], 
-           data.frame(l[i]))}
-CD4_Cytokine.csv[CD4_Cytokine.csv<0]<-0
-CD8_Cytokine.csv[CD8_Cytokine.csv<0]<-0
-GD_Cytokine.csv[GD_Cytokine.csv<0]<-0

### 2019-06-12 tryna function.R

@@ -1,110 +1,0 @@
-splitter<-function(data,column,splitter,one=NULL,two=NULL,three=NULL,four=NULL){
-        library(data.table)
-        
-        labels<-tstrsplit(data[,column], splitter)
-        data[,one]<-labels[[1]]
-        data[,two]<-labels[[2]]
-        data[,three]<-labels[[3]]
-        data[,four]<-labels[[4]]
-        return(data)
-}
-
-base.plot <- function(DF, group1, group2, value) {
-        p <- ggplot(DF, aes(x=group1, y=value, col=group1))
-        p <- p + theme_bw()
-        p <- p + theme(legend.position="bottom")
-        p <- p + geom_jitter(width=.1,height=0, shape=1,size=2)
-        p <- p + scale_color_brewer(palette="Set1", guide=guide_legend(ncol=6, title=NULL))
-        p <- p + xlab("") + ylab("")
-        return(p)
-}
-
-replace.btwn.tables2<-function(datatable1, datatable2, id.var, ref.var, new.var, new.vals){
-    library(dplyr)
-    inter <- intersect(datatable1[,id.var], datatable2[,ref.var])
-    tbl1a <- subset(datatable1, datatable1[,id.var] %in% inter)
-    tbl1b <- subset(datatable1, !(datatable1[,id.var] %in% inter))
-    tbl2  <- subset(datatable2, datatable2[,ref.var] %in% inter)
-    tbl1a[,new.var] <- replace(x = tbl1a[,new.var],
-                           list = (tbl1a[,id.var] %in% tbl2[,ref.var]),
-                           values = as.character(tbl2[,new.vals]))
-    new.table<-rbind(tbl1a, tbl1b)
-}
-
-pvals<-function(data, column){
-    
-    data<-filter(data, SM!="N")
-    data$SM<-factor(data$SM, levels=c("X","SM"))
-    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
-    
-    x<-split(data,data$TB) 
-    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"]))
-    
-    y<-split(data,data$SM) 
-    B<-lapply(y, function(g) wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "LTBI")))
[…]

### 2019-06-13 ICS_Phenotyping.Rmd
just LTBI and TB. Includes total to mtb. includes by stim

@@
 date: "12/2/2018"
-output: pdf_document
+output: html_document
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, error=TRUE)
+knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, comment=NA, warning=FALSE)
 ```
 
-#Total CD4
-```{r total CD4}
-library(dplyr)
-setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
-CD4<-read.csv("CD4_clean.csv")
-    #since we don't care about the yerkes donors in our analysis and they mess everything up
-    #we will filter them out
-    data<-filter(CD4, SM!="N")
-    #and then re-level both the SM and TB variable to not confuse our functions
-    data$SM<-factor(data$SM, levels=c("X","SM"))
-    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
-CD4<-data
-
-#creates a list of character strings for each of the columns of interest
-colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
-for(col in colnames){#this will loop through each character string in the list
-    x<-subset(CD4, CD4$Stim=="UN") #subsets the data to only contain the unstimulated samples
-    #inserts each character string into the column argument of the pvals function written above
-    y<-pvals_helminth(x, col)
-    print(y) #will return the output for each character string in the loop
-}
+```{r}
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
 ```
@@
 #Removing Non-Specific Responses
-```{r AG specific}
+```{r original export}
 #mimosa is the program used to determine whether a response is "real" or not
 #I have previously run this analysis, processed the data, and stored it locally
-mimosa<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/CD4_MIMOSA_Complete.csv")
+mimosa<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/ICS_CD4_MIMOSA_Complete.csv")
 mimosa<-dplyr::rename(mimosa, id = Sample)
@@
 #read in the data set containing the actual phenotype measurements
-phenotype<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/CD4_Positive_clean.csv")
+phenotype<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/ICS_CD4_phenotype_clean.csv")
 #we don't have unstimulated data in the mimosa analysis by default so we remove it
@@
 phenotype<-dplyr::select(DF, -id)
-write.csv(DF, "CD4_Phenotyping_clean.csv") #store for later
+phenotype<-filter(phenotype, Stim!="UN", SM!="N", TB!="HC")
+phenotype$TB<-factor(phenotype$TB, levels=c("LTBI", "TB"))
+phenotype$SM<-factor(phenotype$SM, levels=c("X", "SM"))
+write.csv(DF, "/Applications/Old Computer/Day Lab/Flow-Data/Clean/ICS_CD4_phenotyping_clean_NAs.csv") #store for later
 ```
 
-#P-values by disease status
 ```{r}
@@
 #to see how many values are missing for each column
-kable(colSums(!is.na(phenotype)))
-data<-filter(phenotype, SM!="N")
-data$SM <-factor(data$SM, levels=c("X","SM"))
-data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
+l<-split(phenotype, phenotype$Stim)
+y<-lapply(l, function (g) (colSums(!is.na(g[,c(8, 25, 42, 59)]))))
+numbers<-data.frame(rbind(y[[1]], y[[2]], y[[3]], y[[4]], y[[6]]))
+names(numbers)<-gsub("_MFI_CCR4", "", names(numbers))
+row.names(numbers)<-c("PEP", "PMA", "SEA", "SWAP", "WCL")
+kable(numbers, caption = "Number of samples remaining after exclusion by MIMOSA")
 ```
+
+#P-values by disease status
 
@@
 
-#PMA Phenotyping
+##PMA
 ```{r PMA}
@@
 #since the extent of missingness varies by stim we have to do this each time
-DF<-subset(data, data$Stim=="PMA")
+DF<-subset(phenotype, phenotype$Stim=="PMA")
 #split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
@@
 c<-unlist(y[3])
-d<-unlist(y[5])
-e<-unlist(y[6])
-#then find the intersection of all the lists. Aka which column names have sufficient data in all 6 TB-SM groups
-colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
+d<-unlist(y[4])
+#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
+test<-Reduce(intersect, list(a,b,c, d))
 #this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
-colnames<-colnames[7:length(colnames)]
+colnames<-test[grep("_", test)]
 for(col in colnames){
-    y<-pvals_helminth(DF, col)
-    print(y)
+    y<-pvals_4(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
 }
 ```
-#Peptide-Specific Phenotyping
+
+##PEP
 ```{r PEP}
@@
 library(knitr)
-DF<-subset(data, data$Stim=="PEP")
+DF<-subset(phenotype, phenotype$Stim=="PEP")
+#split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
 x<-split(DF, list(DF$TB, DF$SM))
+#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0.
+It's a weird double negative.
+But basically makes a list of the columns that have sufficient data.
 y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
+#each entry in the list has to be extracted and named
 a<-unlist(y[1])
@@
 c<-unlist(y[3])
-d<-unlist(y[5])
-e<-unlist(y[6])
-colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
-colnames<-colnames[7:length(colnames)]
+d<-unlist(y[4])
+#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
+test<-Reduce(intersect, list(a,b,c, d))
+#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
+colnames<-test[grep("_", test)]
 for(col in colnames){
-    y<-pvals_helminth(DF, col)
-    print(y)
+    y<-pvals_4(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
 }
 ```
-#WCL-Specific Phenotyping
+
+#WCL
 ```{r WCL}
@@
 library(knitr)
-DF<-subset(data, data$Stim=="WCL")
+DF<-subset(phenotype, phenotype$Stim=="WCL")
+#split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
 x<-split(DF, list(DF$TB, DF$SM))
+#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0.
+It's a weird double negative.
+But basically makes a list of the columns that have sufficient data.
 y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
+#each entry in the list has to be extracted and named
 a<-unlist(y[1])
@@
 c<-unlist(y[3])
-d<-unlist(y[5])
-e<-unlist(y[6])
-colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
-colnames<-colnames[7:length(colnames)]
+d<-unlist(y[4])
+#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
+test<-Reduce(intersect, list(a,b,c, d))
+#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
+colnames<-test[grep("_", test)]
 for(col in colnames){
-    y<-pvals_helminth(DF, col)
-    print(y)
+    y<-pvals_4(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
 }
 ```
-#SEA-Specific Phenotyping
-```{r SEA}
-library(dplyr)
-library(knitr)
-DF<-subset(data, data$Stim=="SEA")
-x<-split(DF, list(DF$TB, DF$SM))
+
+#Comparing Mtb-Specific to Total
+```{r}
+total<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/ICS_CD4_clean.csv")
+total$cell<-"Total"
+names<-c("Donor", "TB", "SM", "Stim", names(phenotype)[grep("TH1_", names(phenotype))])
+mtb_th1<-select(phenotype, names)
+names(mtb_th1)<-gsub("TH1_", "", names(mtb_th1))
+mtb_th1$cell<-"Mtb-Specific"
+test<-rbindlist(list(total, mtb_th1), fill=TRUE)
+    test<-filter(test, SM!="N", TB!="HC") 
+    test$SM<-factor(test$SM, levels=c("X","SM"))
+    test$TB<-factor(test$TB, levels=c("LTBI","TB"))
+test2<-merge(total, mtb_th1, by=c("Donor","TB","SM","Stim"))
+    test2<-filter(test2, SM!="N", TB!="HC") 
+    test2$SM<-factor(test2$SM, levels=c("X","SM"))
+    test2$TB<-factor(test2$TB, levels=c("LTBI","TB"))
+```
+
+##PMA
+```{r}
+df1<-filter(test, Stim=="PMA")
+df2<-filter(test2, Stim=="PMA")
+
+x<-split(subset(mtb_th1, Stim=="PMA"), list(mtb_th1$TB, mtb_th1$SM))
+#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0.
+It's a weird double negative.
+But basically makes a list of the columns that have sufficient data.
 y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
+#each entry in the list has to be extracted and named
 a<-unlist(y[1])
@@
 c<-unlist(y[3])
-d<-unlist(y[5])
-e<-unlist(y[6])
-colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
-colnames<-colnames[7:length(colnames)]
+d<-unlist(y[4])
+#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
+n<-Reduce(intersect, list(a,b,c, d))
+#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
+colnames<-n[5:36]
 for(col in colnames){
-    y<-pvals_helminth(DF, col)
-    print(y)
-}
+    a<-pvals_total.mtb(df1, col)
+    df2$new<-df2[,paste(col,"x",sep=".")]-df2[,paste(col,"y",sep=".")]
+    b<-pvals_4(df2, "new")
+    if(dim(a)[1]>1 | dim(b)[1]>1){
+        print(col)
+        print(a)
+        print(b)}}
 ```
-#SWAP-Specific Phenotyping
-```{r SWAP}
-library(dplyr)
-library(knitr)
-DF<-subset(data, data$Stim=="SWAP")
-x<-split(DF, list(DF$TB, DF$SM))
+
+##PEP
+```{r}
+df1<-filter(test, Stim=="PEP")
+df2<-filter(test2, Stim=="PEP")
+
+x<-split(subset(mtb_th1, Stim=="PEP"), list(mtb_th1$TB, mtb_th1$SM))
+#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0.
+It's a weird double negative.
+But basically makes a list of the columns that have sufficient data.
 y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
+#each entry in the list has to be extracted and named
 a<-unlist(y[1])
@@
 c<-unlist(y[3])
-d<-unlist(y[5])
-e<-unlist(y[6])
-colnames<-intersect(intersect((intersect(intersect(a,b),c)),d),e)
-colnames<-colnames[7:length(colnames)]
+d<-unlist(y[4])
+#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
+n<-Reduce(intersect, list(a,b,c, d))
+#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
+colnames<-n[5:36]
 for(col in colnames){
-    y<-pvals_helminth(DF, col)
-    print(y)
-}
+    a<-pvals_total.mtb(df1, col)
+    df2$new<-df2[,paste(col,"x",sep=".")]-df2[,paste(col,"y",sep=".")]
+    b<-pvals_4(df2, "new")
+    if(dim(a)[1]>1 | dim(b)[1]>1){
+        print(col)
+        print(a)
+        print(b)}}
 ```
 
-#P-Values by Stim
+##WCL
+```{r}
+df1<-filter(test, Stim=="WCL")
+df2<-filter(test2, Stim=="WCL")
+
+x<-split(subset(mtb_th1, Stim=="WCL"), list(mtb_th1$TB, mtb_th1$SM))
+#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0.
+It's a weird double negative.
+But basically makes a list of the columns that have sufficient data.
+y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
+#each entry in the list has to be extracted and named
+a<-unlist(y[1])
+b<-unlist(y[2])
+c<-unlist(y[3])
+d<-unlist(y[4])
+#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
+n<-Reduce(intersect, list(a,b,c, d))
+#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
+colnames<-n[5:36]
+for(col in colnames){
+    a<-pvals_total.mtb(df1, col)
+    df2$new<-df2[,paste(col,"x",sep=".")]-df2[,paste(col,"y",sep=".")]
+    b<-pvals_4(df2, "new")
+    if(dim(a)[1]>1 | dim(b)[1]>1){
+        print(col)
+        print(a)
+        print(b)}}
+```
+
+#Comparing by Stim
 In this case, the function is written such that I have to subset to look within each of the TB-SM subsets one by one.
@@
 library(ggplot2)
-data<-filter(phenotype, SM!="N")
-    data$SM<-factor(data$SM, levels=c("X","SM"))
-    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))
-ggplot(data, aes(y=TH1_GATA3, x=Stim, col=TB))+
-    geom_point(size=3, alpha=0.5)+geom_line(aes(group=Donor))+
-    scale_color_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
-    theme_bw()+ theme(legend.position="none")+
-    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    facet_grid(SM~TB)+coord_cartesian(ylim = c(0,20))+
-    labs(y= "Frequency of GATA3+ Mtb-Specific TH1 Cells")
-```
-
-#Healthy Controls SM-
-```{r}
-library(dplyr)
-colnames<-names(dplyr::select(phenotype, -X, -Sample, -Donor, -Stim, -TB, -SM))
-df<-filter(phenotype, TB=="HC" & SM=="X")
+data<-filter(phenotype, Stim %in% c("PMA", "PEP", "WCL"))
+data$Stim<-factor(data$Stim, levels=c("PMA","PEP","WCL"))
+colnames<-names(data)[grep("TH1_", names(data))]
 for(col in colnames){
-    y<-pvals_stim(df, col)
+    y<-plot_stim(data, col)
     print(y)
@@
 ```
-
-#Healthy Controls SM+
-```{r}
-library(dplyr)
-colnames<-names(dplyr::select(phenotype, -X, -Sample, -Donor, -Stim, -TB, -SM))
-df<-filter(phenotype, TB=="HC" & SM=="SM")
-for(col in colnames){
-    y<-pvals_stim(df, col)
-    print(y)
-}
-```
-
-#LTBI SM-
-```{r}
-library(dplyr)
-colnames<-names(dplyr::select(phenotype, -X, -Sample, -Donor, -Stim, -TB, -SM))
-df<-filter(phenotype, TB=="LTBI" & SM=="X")
-for(col in colnames){
-    y<-pvals_stim(df, col)
-    print(y)
-}
-```
-
-#LTBI SM+
-```{r}
-library(dplyr)
-colnames<-names(dplyr::select(phenotype, -X, -Sample, -Donor, -Stim, -TB, -SM))
-df<-filter(phenotype, TB=="LTBI" & SM=="SM")
-for(col in colnames){
-    y<-pvals_stim(df, col)
-    print(y)
-}
-```
-
-#TB SM-
-```{r}
-library(dplyr)
-colnames<-names(dplyr::select(phenotype, -X, -Sample, -Donor, -Stim, -TB, -SM))
-df<-filter(phenotype, TB=="TB" & SM=="X")
-for(col in colnames){
-    y<-pvals_stim(df, col)
-    print(y)
-}
-```
-
-#TB SM+
-```{r}
-library(dplyr)
-colnames<-names(dplyr::select(phenotype, -X, -Sample, -Donor, -Stim, -TB, -SM))
-df<-filter(phenotype, TB=="TB" & SM=="SM")
-for(col in colnames){
-    y<-pvals_stim(df, col)
-    print(y)
-}
-```
-
-#P values by Stim and Helminth
-I haven't written this up nicely yet but none of it is significant

### 2019-06-13 OG_cleaning.Rmd
just LTBI and TB. Includes total to mtb. includes by stim

---
title: "OG_cleaning"
author: "Taryn McLaughlin"
date: "6/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/exported/")
#find all the csv filenames in that directory and put them in a list
filenames =list.files(pattern="OG")
#for each name in that list, assign it to the output of read.csv of that name
#aka read in each file by it's name
for (i in 1:length(filenames)) 
    assign(filenames[i], read.csv(filenames[i]))
```

```{r source functions used in this script}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/OG-cleaning-functions.R")
```

```{r}
l<-lapply(ls(pattern="summary.csv"), function(x) get(x))

l<-lapply(l, function(x) split_og(x))
l<-lapply(l, function(x) clean_ognames(x))
l<-lapply(l, function(x) remove_samples(x))
l<-lapply(l, function(x) remove_donors(x))
l<-lapply(l, function(x) changeTB(x))

filenames<-ls(pattern="summary.csv")
for(i in 1:15){
assign(filenames[i], data.frame(l[i]))}
```

```{r}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")

write.csv(OG_CD3_bulk_summary.csv, "OG_CD3_bulk_summary_clean.csv")
write.csv(OG_CD3_OG_summary.csv, "OG_CD3_OG_summary_clean.csv")

write.csv(OG_CD4_bulk_summary.csv, "OG_CD4_bulk_summary_clean.csv")
write.csv(OG_CD4_OG_summary.csv, "OG_CD4_OG_summary_clean.csv")
write.csv(OG_CD4_OG.GT_summary.csv, "OG_CD4_OG.GT_summary_clean.csv")
write.csv(OG_CD4_OG.IL4_summary.csv, "OG_CD4_OG.IL4_summary_clean.csv")
write.csv(OG_CD4_OG.TH1_summary.csv, "OG_CD4_OG.TH1_summary_clean.csv")
write.csv(OG_CD4_OG.TNFa_summary.csv, "OG_CD4_OG.TNFa_summary_clean.csv")

write.csv(OG_CD8_bulk_summary.csv, "OG_CD8_bulk_summary_clean.csv")
write.csv(OG_CD8_OG_summary.csv, "OG_CD8_OG_summary_clean.csv")

write.csv(OG_gd_bulk_summary.csv, "OG_gd_bulk_summary_clean.csv")
write.csv(OG_GDonly_bulk_summary.csv, "OG_GDonly_bulk_summary_clean.csv")
write.csv(OG_gd_OG_summary.csv, "OG_gd_OG_summary_clean.csv")
write.csv(OG_GDonly_OG_summary.csv, "OG_GDonly_OG_summary_clean.csv")

write.csv(OG_counts_summary.csv, "OG_counts_summary_clean.csv")
```

### 2019-06-17 Cohort-Characteristics.Rmd
Includes both OG and ICS data

@@
 ```{r load data}
-DonorInfo<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/ICS_Donor_Info_clean.csv")
+DonorInfo<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/DonorInfo_clean.csv")
 DonorInfo<-filter(DonorInfo, SM %in% c("X", "SM"))
@@
 DonorInfo$SchistosomaIntensity[DonorInfo$SM=="X"]<-NA
+
 ```
@@
                          strata = c("TB", "SM"))
-test<- print(table1, showAllLevels = TRUE, nonnormal = mediqr, quote = FALSE, noSpaces = TRUE,
-             printToggle = FALSE)
+test<- print(table1, showAllLevels = TRUE, nonnormal = mediqr, quote = FALSE, noSpaces = TRUE, printToggle = FALSE, na.print=FALSE)
 
@@
 library(compareGroups)
-res.total<-descrTable(~ age + sex + QFT + SchistosomaIntensity, data = DonorInfo)
+res.total<-descrTable(~ age + sex + QFT + SchistosomaIntensity, data = DonorInfo, na.action = na.exclude)
+
 res <- compareGroups(SM ~ age + sex + QFT + SchistosomaIntensity, data = DonorInfo,
@@
 
-restab <- createTable(res, type = 1, hide.no = "no", show.n = FALSE)
+restab <- createTable(res, type = 1, hide.no = "no", show.n = FALSE,
+                      show.p.overall = TRUE, show.p.mul = TRUE,
+                      q.type=c(2,3))
 test<-strataTable(restab, "TB")
@@
 Plotting stuff against age and sex
-```{r age and sex vs, cache = TRUE}
+```{r age and sex vs, cache = TRUE, eval=FALSE}
 compass<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/ICS_Compass_scores_complete.csv")

### 2019-06-17 Cytokine_Analysis.Rmd
narrowed down to LTBI and TB

@@
 library(cowplot)
+library(tidyr)
+library(reshape2)
 ```
@@
 CD4<-plot_filter(read.csv("ICS_CD4_Cytokine_clean.csv"))
-CD4b<-plot_filter(read.csv("ICS_Cytokine_new.csv"))
 
@@
 
-Compass<-plot_filter(read.csv("ICS_Compass_scores_complete.csv"))
+Compass<-plot_filter(read.csv("ICS_Compass_scores_complete_clean.csv"))
 
@@
 
-data<-read.csv("ICS_Donor_Info_clean.csv")
+data<-read.csv("ICS_DonorInfo_clean.csv")
 data$TB<-gsub("BCG|Unknown", "Naïve", data$TB)
@@
 }
-colnames<-c("TH1_total", "TH2_total")
-for(col in colnames){#this will loop through each character string in the list
-    a<-subset(CD4b, CD4b$Stim=="WCL") #subsets the data to only contain the unstimulated samples
-    #inserts each character string into the column argument of the pvals function written above
-    b<-pvals_4(a, col)
-    if(dim(b)[1]!=0)
-    print(kable(b, caption=col)) #will return the output for each character string in the loop
-}
 ```
@@
 }
-colnames<-c("TH1_total", "TH2_total")
-for(col in colnames){#this will loop through each character string in the list
-    a<-subset(CD4b, CD4b$Stim=="PEP") #subsets the data to only contain the unstimulated samples
-    #inserts each character string into the column argument of the pvals function written above
-    b<-pvals_4(a, col)
-    if(dim(b)[1]!=0)
-    print(kable(b, caption=col)) #will return the output for each character string in the loop
-}
 ```
 
-#Compass Graphs
+```{r}
+grid.arrange(
+    plot_2(subset(df, df$Stim=="PMA"), "disease", "IFNg")+
+        labs(y="IFNg+ CD4+ T cells (%)", title="PMA & Ionomycin"),
+    plot_2(subset(df, df$Stim=="PEP"), "disease", "IFNg")+
+        labs(y="", title="CFP-10 & ESAT-6"),
+    ncol=2)
+
+grid.arrange(
+    plot_2(subset(df, df$Stim=="PMA"), "disease", "G_T")+
+        labs(y="IFNg+TNFa+ CD4+ T cells (%)", title="PMA & Ionomycin"),
+    plot_2(subset(df, df$Stim=="PEP"), "disease", "G_T")+
+        labs(y="", title="CFP-10 & ESAT-6"),
+    ncol=2)
+
+grid.arrange(
+    plot_2(subset(df, df$Stim=="PMA"), "disease", "x_4_13_T")+
+        labs(y="IL4+TNFa+ CD4+ T cells (%)", title="PMA & Ionomycin"),
+    plot_2(subset(df, df$Stim=="PEP"), "disease", "x_4_13_T")+
+        labs(y="", title="CFP-10 & ESAT-6"),
+    ncol=2) 
+```
+
+#Extra Graphs
+
+##Compass Graphs
 ```{r}
@@
 
-```{r, fig.width=16, fig.height=12}
-library(dplyr)
-DF<-merge(Compass, dplyr::select(Donor, -X, -TB, -SM), by="Donor")
-ggpairs(dplyr::filter(DF, celltype=="CD4", Stim=="WCL", TB!="N")%>%
-            mutate(disease = paste(TB, SM, sep=" "))%>%
-            dplyr::select(age, sex, FS, PFS, disease),
-        aes(col=disease))
+##Boolean Graphs
+The first graph is just meant to give a snapshot of all the data at once to see if there are weird anomalous data points that we will have to examine.
+I did not include all the individual data points on any of the graphs because it made it look incredibly cluttered, however I did allow outliers to display.
 
+```{r Boolean Graphs, cache=TRUE, echo=FALSE, warning=FALSE, fig.width=20}
+pma<-filter(CD4, Stim=="PMA")
+wcl<-filter(CD4, Stim=="WCL")
+pep<-filter(CD4, Stim=="PEP")
+triple_boolean_plot(pma)
+triple_boolean_plot(wcl)
+triple_boolean_plot(pep)
 ```
 
-#Boolean Graphs
-The first graph is just meant to give a snapshot of all the data at once to see if there are weird anomalous data points that we will have to examine.
-I did not include all the individual data points on any of the graphs because it made it look incredibly cluttered, however I did allow outliers to display.
-```{r Boolean Graphs, cache=TRUE, echo=FALSE, warning=FALSE}
-library(dplyr)
-library(tidyr)
-library(reshape2)
-library(ggplot2)
-triple_boolean_plot(CD4, "PMA")
-triple_boolean_plot(CD4, "WCL")
-triple_boolean_plot(CD4, "PEP")
-```
-
-#T cell Lineage Specific Grouped Analysis
+##T cell Lineage Specific Grouped Analysis
 These are my standard lineage specific groups. 
@@
 
-#Individual Cytokines
+##Individual Cytokines
 These are the frequencies of cells producing individual cytokines without considering whether or not they are single positive, double positive etc.
@@
 ```
-
-```{r}
-library(dplyr)
-library(reshape2)
-library(ggplot2)
-library(knitr)
-
-colnames<-c("TH1_Total" , "TH2_Total" )
-
-### PMA
-for(col in colnames){#this will loop through each character string in the list
-    #inserts each character string into the column argument of the pvals function written above
-    name<-gsub("_", "+", col)
-    y<-plot_with_all_stats(CD4b, "PMA", col)+
-        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
-             title="PMA & Ionomycin",
-             x="",
-             y=paste(col, "+ CD4+ T cells (%)"))
-    print(y) #will return the output for each character string in the loop
-}
-```

### 2019-06-17 ICS_Phenotyping.Rmd
narrowed down to LTBI and TB. Includes mtb to total comps

@@
 
+```{r}
+test<-mutate_TB(filter(DF, !is.na(TH1_GATA3)))
+kable(xtabs(~TB+SM, test))
+plot_2(test, "disease", "TH1_GATA3")
+plot_2(test, "disease", "TH1_t_x_x_4")
+plot_2(test, "disease", "TH1_t_x_x_x")
+plot_2(test, "disease", "TH1_x_g_3_4")
+plot_2(test, "disease", "TH1_x_g_3_x")
+plot_2(test, "disease", "TH1_x_g_x_4")
+plot_2(test, "disease", "TH1_x_x_3_x")
+
+grid.arrange(plot_2(test, "disease", "TH1_t_x_x_x")+labs(y="Tbet+ CD4+ T cells"),
+plot_2(test, "disease", "TH1_t_x_x_4")+labs(y="CCR4+Tbet+ CD4+ T Cells (%)"),
+    plot_2(test, "disease", "TH1_GATA3")+labs(y="GATA3+ CD4+ T cells (%)"),
+ncol=3
+)
+```
+
 #WCL
@@
 #Comparing Mtb-Specific to Total
+This is only within TH1 cells
 ```{r}
@@
 
+```{r}
+t<-mutate_TB(filter(df1, !is.na(GATA3)))
+t$cell<-factor(t$cell, levels=c("Total", "Mtb-Specific"))
+plot_compare(t, "cell", "GATA3")
+plot_compare(t, "cell", "t_x_x_4")
+plot_compare(t, "cell", "t_x_x_x")
+plot_compare(t, "cell", "x_g_3_4")
+plot_compare(t, "cell", "x_g_3_x")
+plot_compare(t, "cell", "x_g_x_4")
+plot_compare(t, "cell", "x_x_3_x")
+```
+
 ##WCL

### 2019-06-17 ICS_CD4.Rmd
LTBI/TB only

@@
 setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
-cd4a<-read.csv("ICS_CD4_clean.csv")
-cd4b<-read.csv("ICS_CD4_new.csv")
-```
-
-#Are HC and LTBI different or can we combine?
-```{r, fig.height=3, fig.width=6}
-library(knitr)
-colnames<-names(dplyr::select(cd4a, -X, -Sample, -Donor, -Stim, -TB, -SM))
-df<-filter(cd4a, Stim=="UN")
-df<-mutate_hc(df)
-for(col in colnames){
-    print(col)
-    print(pvals_hc(df, col))
-    grid.arrange(
-       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
-       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
-       ncol=2
-    )
-}
-```
-
-```{r}
-colnames<-names(dplyr::select(cd4b, -X, -Sample, -Donor, -Stim, -TB, -SM))
-df<-filter(cd4b, Stim=="UN")
-df<-mutate_hc(df)
-for(col in colnames){
-    print(col)
-    print(pvals_hc(df, col))
-    grid.arrange(
-       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
-       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
-       ncol=2
-    )
-}
+cd4<-read.csv("ICS_CD4_clean.csv")
 ```
@@
 
-####Option 1
-```{r, fig.width=8, fig.height=8}
-data<-filter(cd4, cd4$Stim=="UN", cd4$TB=="HC")
-a<-plot_hc(data, "SM", "CXCR3")+labs(title="Surface Marker Expression", subtitle="TH1")
-b<-plot_hc(data, "SM", "CCR4")+labs(title=" ", subtitle="TH2")
-c<-plot_hc(data, "SM", "Tbet")+labs(title="Transcription Factor Expression", subtitle="TH1")
-d<-plot_hc(data, "SM", "GATA3")+labs(title=" ",subtitle="TH2")
+##Significant on Total CD4+ T cells
+```{r, fig.width=12, fig.height=8}
+data<-mutate_TB(filter(cd4, cd4$Stim=="UN"))
+a<-plot_2(data, "disease", "CCR4")
+b<-plot_2(data, "disease", "t_g_x_4")
+c<-plot_2(data, "disease", "GATA3")
+d<-plot_2(data, "disease", "GATA3.P_Tbet.N")
 plot_grid(a,b,c,d, ncol=2)
@@
 
-####Option 2
-```{r, fig.width=8, fig.height=8}
-a<-plot_hc(data, "SM", "CXCR3")+labs(title="TH1", subtitle="Surface Marker")
-b<-plot_hc(data, "SM", "Tbet")+labs(title=" ", subtitle="Transcription Factor")
-c<-plot_hc(data, "SM", "CCR4")+labs(title="TH2", subtitle="Surface Marker")
-d<-plot_hc(data, "SM", "GATA3")+labs(title=" ", subtitle="Transcription Factor")
-plot_grid(a,b,c,d, ncol=2 )
+##These are significant amongst Peptide-Specific CD4+ T cells
+```{r}
+b<-plot_2(data, "disease", "Tbet")+labs(y="Tbet+ Total CD4+ T cells (%)")
+c<-plot_2(data, "disease", "t_x_x_4")+labs(y="CCR4+Tbet+ Total CD4+ T cells (%)")
+d<-plot_2(data, "disease", "GATA3")+labs(y="GATA3+ Total CD4+ T cells (%)")
+plot_grid(b,c,d, ncol=3)
 ```

### 2019-06-17 ICS_Cleaning.Rmd
sums over boolean phenotype data to create double positives

@@
 
+```{r create new combo variables from exported boolean phenotype}
+CD4<-mutate(CD4,
+      #These all follow the same format
+       #IFNg is the new variable. it gets assign the row sum of all the columns that have a G in the name
+       TbetCXCR3=  rowSums(CD4[,grep("t_._3_.", names(CD4))]),
+       TbetCCR4=  rowSums(CD4[,grep("t_._._4", names(CD4))]),
+       GATA3CXCR3=  rowSums(CD4[,grep("._g_3_.", names(CD4))]),
+       GATA3CCR4=  rowSums(CD4[,grep("._g_._4", names(CD4))])
+)
+
+CD4_phenotype<-mutate(CD4_phenotype,
+      #These all follow the same format
+       #IFNg is the new variable. it gets assign the row sum of all the columns that have a G in the name
+       TH1_TbetCXCR3=  rowSums(CD4_phenotype[,grep("TH1_t_._3_.", names(CD4_phenotype))]),
+       TH1_TbetCCR4=  rowSums(CD4_phenotype[,grep("TH1_t_._._4", names(CD4_phenotype))]),
+       TH1_GATA3CXCR3=  rowSums(CD4_phenotype[,grep("TH1_._g_3_.", names(CD4_phenotype))]),
+       TH1_GATA3CCR4=  rowSums(CD4_phenotype[,grep("TH1_._g_._4", names(CD4_phenotype))]),
+      TH2_TbetCXCR3=  rowSums(CD4_phenotype[,grep("TH2_t_._3_.", names(CD4_phenotype))]),
+      TH2_TbetCCR4=  rowSums(CD4_phenotype[,grep("TH2_t_._._4", names(CD4_phenotype))]),
+      TH2_GATA3CXCR3=  rowSums(CD4_phenotype[,grep("TH2_._g_3_.", names(CD4_phenotype))]),
+      TH2_GATA3CCR4=  rowSums(CD4_phenotype[,grep("TH2_._g_._4", names(CD4_phenotype))])
+)
+
+```
+
 ```{r Cytokine MFI inclusion}

### 2019-06-17 ICS_HC.Rmd
updating to reflect that new export was combined with old in Cleaning script

@@
 setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
-cd4a<-read.csv("ICS_CD4_clean.csv")
-cd4b<-read.csv("ICS_CD4_new.csv")
+cd4<-read.csv("ICS_CD4_clean.csv")
 
 CD4<-plot_filter(read.csv("ICS_CD4_Cytokine_clean.csv"))
-CD4b<-plot_filter(read.csv("ICS_CD4_Cytokine_new.csv"))
 
-phen<-read.csv("ICS_CD4_Phenotyping_clean.csv")
-phenb<-read.csv("ICS_CD4_phenotype_new_clean.csv")
+phen<-read.csv("ICS_CD4_phenotyping_clean_NAs.csv")
 ```
@@
 library(knitr)
-colnames<-names(dplyr::select(cd4a, -X, -Sample, -Donor, -Stim, -TB, -SM))
-df<-filter(cd4a, Stim=="UN")
-df<-mutate_hc(df)
-for(col in colnames){
-    print(col)
-    print(pvals_hc(df, col))
-    grid.arrange(
-       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
-       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
-       ncol=2
-    )
-}
-```
-
-```{r, fig.height=3, fig.width=6}
-colnames<-names(dplyr::select(cd4b, -X, -Sample, -Donor, -Stim, -TB, -SM))
-df<-filter(cd4b, Stim=="UN")
+colnames<-names(dplyr::select(cd4, -X, -Sample, -Donor, -Stim, -TB, -SM))
+df<-filter(cd4, Stim=="UN")
 df<-mutate_hc(df)
@@
 
-```{r, fig.height=3, fig.width=6}
-library(knitr)
-colnames<-names(dplyr::select(CD4b, -X, -Sample, -Donor, -Stim, -TB, -SM))
-df<-filter(CD4b, Stim=="PMA")
-df<-mutate_hc(df)
-for(col in colnames){
-    print(col)
-    print(pvals_hc(df, col))
-    grid.arrange(
-       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
-       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
-       ncol=2
-    )
-}
-```
-
 #Phenotype from PMA
@@
 
-#Phenotype from PMA
-```{r, fig.height=3, fig.width=6}
-library(knitr)
-colnames<-names(dplyr::select(phenb, -X.1, -X, -id, -Sample, -Donor, -Stim, -TB, -SM))
-df<-filter(phenb, Stim=="PMA")
-df<-mutate_hc(df)
-for(col in colnames){
-    print(col)
-    print(pvals_hc(df, col))
-    grid.arrange(
-       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
-       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
-       ncol=2
-    )
-}
-```
-
-```{r}
-colnames<-names(dplyr::select(phen, -X, -X.1, -id, -Sample, -Donor, -Stim, -TB, -SM))
-for(col in colnames){#this will loop through each character string in the list
-    a<-subset(phen, phen$Stim=="PEP") #subsets the data to only contain the unstimulated samples
-    #inserts each character string into the column argument of the pvals function written above
-    b<-pvals_6(a, col) 
-    print(b)
-    c<-plot_with_all_stats(a, "PEP", col)+labs(y=col)
-    print(c)
-     #will return the output for each character string in the loop
-}
-
-colnames<-names(dplyr::select(phenb, -X, -X.1, -id, -Sample, -Donor, -Stim, -TB, -SM))
-for(col in colnames){#this will loop through each character string in the list
-    a<-subset(phenb, phenb$Stim=="PEP") #subsets the data to only contain the unstimulated samples
-    #inserts each character string into the column argument of the pvals function written above
-    b<-pvals_6(a, col) 
-    print(b)
-    c<-plot_with_all_stats(a, "PEP", col)+labs(y=col)
-    print(c)
-    #will return the output for each character string in the loop
-}
-```
+##Actual analysis

### 2019-07-02 ICS_CD4.Rmd
Includes summed boolean and proportion figures

@@
 cd4<-read.csv("ICS_CD4_clean.csv")
-```
+grid.arrange(
+    grid.arrange(
+        grid.arrange(tableGrob(xtabs(~TB+SM, test)), HC_T, ncol=2),
+        two, ncol=1), 
+            b, ncol=2)```
 
@@
 #creates a list of character strings for each of the columns of interest
-colnames<-names(dplyr::select(cd4, -X, -Sample, -Donor, -Stim, -TB, -SM))
+colnames<-names(dplyr::select(cd4, -X, -X.1,-Sample, -Donor, -Stim, -TB, -SM))
 for(col in colnames){#this will loop through each character string in the list
@@
 ##Significant on Total CD4+ T cells
-```{r, fig.width=12, fig.height=8}
+```{r, fig.width=16, fig.height=6}
 data<-mutate_TB(filter(cd4, cd4$Stim=="UN"))
@@
 d<-plot_2(data, "disease", "GATA3.P_Tbet.N")
-plot_grid(a,b,c,d, ncol=2)
+e<-plot_2(data, "disease", "GATA3CCR4")
+plot_grid(a,b,ncol=2)
+plot_grid(c,d,e,ncol=3)
 ```
@@
 ##These are significant amongst Peptide-Specific CD4+ T cells
-```{r}
+```{r, fig.width=12, fig.height=6}
 b<-plot_2(data, "disease", "Tbet")+labs(y="Tbet+ Total CD4+ T cells (%)")
@@
 
-###MFI
-```{r}
-for(col in colnames[1:4]){#this will loop through each character string in the list
-    x<-subset(cd4, cd4$Stim=="UN") #subsets the data to only contain the unstimulated samples
-    #inserts each character string into the column argument of the pvals function written above
-    y<-plot_6(x, "SM", col)+
-        labs(subtitle=paste(gsub("_", " ", col), "on CD4+ T cells"), 
-             caption="",
-             x="",
-             y=gsub("_", " ", col))
-    print(y) #will return the output for each character string in the loop
-}
+```{r, fig.width=16, fig.height=6}
+test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = grep("._._._.", names(cd4)))
+test2<-filter(test2, variable!="x_x_x_x")
+ggplot(test2, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot()+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_fill_manual(values =c("#2166ac", "#b2182b"))+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
+    theme(legend.position="none")+
+    labs(title="Lineage Marker Expression on Total CD4+ T cells",
+         y="Frequency of Marker+ CD4+ T cells (%)")
 ```
@@
 for(col in colnames[5:10]){#this will loop through each character string in the list
-    x<-subset(cd4, cd4$Stim=="UN") #subsets the data to only contain the unstimulated samples
+    x<-plot_filter(subset(cd4, cd4$Stim=="UN")) #subsets the data to only contain the unstimulated samples
     #inserts each character string into the column argument of the pvals function written above
@@
     name<-gsub(".N", "-", name)
-    y<-plot_6(x, "SM", col)+
+    y<-plot_with_all_stats(x, col)+
         labs(subtitle=paste(name, "(% CD4+ T cells)"), 
@@
 ###Transcription Factors
-
 As you can see there is one crazy looking "outlier" in GATA3 from NK2586. 
@@
 for(col in colnames[11:16]){#this will loop through each character string in the list
-    x<-subset(cd4, cd4$Stim=="UN") #subsets the data to only contain the unstimulated samples
+    x<-plot_filter(subset(cd4, cd4$Stim=="UN")) #subsets the data to only contain the unstimulated samples
     #inserts each character string into the column argument of the pvals function written above
@@
 
-#Side Note
-This is the statistical evaluation of the data set without NK2586 the outlier GATA3 donor to see if it makes a difference.
+##Combination Gates
 ```{r}
-colnames2<-colnames[grep("GATA", colnames)]
-for(col in colnames2){#this will loop through each character string in the list
-    a<-filter(cd4, Stim=="UN", Donor!="NK2586") #subsets the data to only contain the unstimulated samples
+for(col in colnames[17:48]){#this will loop through each character string in the list
+    x<-plot_filter(subset(cd4, cd4$Stim=="UN")) #subsets the data to only contain the unstimulated samples
     #inserts each character string into the column argument of the pvals function written above
-    b<-pvals_6(a, col)
-    if(dim(b)[1]!=0)
-    print(kable(b, caption=col)) #will return the output for each character string in the loop
+    y<-plot_with_all_stats(x, col)+
+        labs(subtitle=col, 
+           x="",
+           y="")
+    print(y) #will return the output for each character string in the loop
 }
 ```
+
+###MFI
+```{r}
+for(col in colnames[1:4]){#this will loop through each character string in the list
+    x<-plot_filter(subset(cd4, cd4$Stim=="UN")) #subsets the data to only contain the unstimulated samples
+    #inserts each character string into the column argument of the pvals function written above
+    y<-plot_with_all_stats(x, col)+
+        labs(subtitle=paste(gsub("_", " ", col), "on CD4+ T cells"), 
+             caption="",
+             x="",
+             y=gsub("_", " ", col))
+    print(y) #will return the output for each character string in the loop
+}
+```

### 2019-07-02 ICS_Cleaning.Rmd
proportions calculations

--- 2019-06-10 ICS_Cleaning.Rmd
@@
 
+```{r create new combo variables from exported boolean phenotype}
+CD4<-mutate(CD4,
+      #These all follow the same format
+       #IFNg is the new variable. it gets assign the row sum of all the columns that have a G in the name
+       TbetCXCR3=  rowSums(CD4[,grep("t_._3_.", names(CD4))]),
+       TbetCCR4=  rowSums(CD4[,grep("t_._._4", names(CD4))]),
+       GATA3CXCR3=  rowSums(CD4[,grep("._g_3_.", names(CD4))]),
+       GATA3CCR4=  rowSums(CD4[,grep("._g_._4", names(CD4))]),
+       CXCR3.Tbetprop = TbetCXCR3/Tbet,
+       CCR4.Tbetprop = TbetCCR4/Tbet,
+       CXCR3.GATA3prop= GATA3CXCR3/GATA3,
+       CCR4.GATA3prop= GATA3CCR4/GATA3
+)
+
+CD4_phenotype<-mutate(CD4_phenotype,
+      #These all follow the same format
+       #IFNg is the new variable. it gets assign the row sum of all the columns that have a G in the name
+       TH1_TbetCXCR3=  rowSums(CD4_phenotype[,grep("TH1_t_._3_.", names(CD4_phenotype))]),
+       TH1_TbetCCR4=  rowSums(CD4_phenotype[,grep("TH1_t_._._4", names(CD4_phenotype))]),
+       TH1_GATA3CXCR3=  rowSums(CD4_phenotype[,grep("TH1_._g_3_.", names(CD4_phenotype))]),
+       TH1_GATA3CCR4=  rowSums(CD4_phenotype[,grep("TH1_._g_._4", names(CD4_phenotype))]),
+      TH2_TbetCXCR3=  rowSums(CD4_phenotype[,grep("TH2_t_._3_.", names(CD4_phenotype))]),
+      TH2_TbetCCR4=  rowSums(CD4_phenotype[,grep("TH2_t_._._4", names(CD4_phenotype))]),
+      TH2_GATA3CXCR3=  rowSums(CD4_phenotype[,grep("TH2_._g_3_.", names(CD4_phenotype))]),
+      TH2_GATA3CCR4=  rowSums(CD4_phenotype[,grep("TH2_._g_._4", names(CD4_phenotype))]),
+
+       TH1_CXCR3.Tbetprop = TH1_TbetCXCR3/TH1_Tbet,
+       TH1_CCR4.Tbetprop = TH1_TbetCCR4/TH1_Tbet,
+       TH1_CXCR3.GATA3prop= TH1_GATA3CXCR3/TH1_GATA3,
+       TH1_CCR4.GATA3prop= TH1_GATA3CCR4/TH1_GATA3,
+
+       TH2_CXCR3.Tbetprop = TH2_TbetCXCR3/TH2_Tbet,
+       TH2_CCR4.Tbetprop = TH2_TbetCCR4/TH2_Tbet,
+       TH2_CXCR3.GATA3prop= TH2_GATA3CXCR3/TH2_GATA3,
+       TH2_CCR4.GATA3prop= TH2_GATA3CCR4/TH2_GATA3
+)
+
+```
+
 ```{r Cytokine MFI inclusion}

### 2019-07-02 ICS_HC.Rmd
Includes boolean graphs for phenotype

@@
 setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
-cd4<-read.csv("ICS_CD4_clean.csv")
+cd4<-plot_filter(read.csv("ICS_CD4_clean.csv"))
 
@@
 
-phen<-read.csv("ICS_CD4_phenotyping_clean_NAs.csv")
+phen<-plot_filter(read.csv("ICS_CD4_phenotyping_clean_NAs.csv"))
 ```
 
-#Are HC and LTBI different or can we combine?
+##Actual analysis
 
+```{r fig.width=16, fig.height=8}
+HC<-mutate_hc(subset(cd4, cd4$Stim=="UN"))
+test<-melt(HC, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = grep("._._._.", names(cd4)))
+test2<-filter(test, variable!="x_x_x_x")
+ggplot(test2, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot()+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_fill_manual(values =c("#1a9850"))+
+    stat_compare_means(label="p.signif", hide.ns = TRUE)+
+    theme(legend.position="none")+
+    labs(title="Lineage Marker Expression on Total CD4+ T cells in Healthy Adults")
+```
+
+```{r Figure 1a, }
+HC<-mutate_hc(subset(cd4, cd4$Stim=="UN"))
+HC_T<-ggplot(HC, aes(x=SM, y=Tbet))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
+    geom_jitter(width=.1,height=0, shape=16,size=2)+
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_x_discrete(labels=c("SM-", "SM+"))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(axis.text.x = element_text(vjust=0.5)) + 
+    theme(text = element_text(size=14)) +
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE, label.y= (max(HC$Tbet)*1.1),
+                       label.x.npc = "center")+
+    labs(title="",
+         x="",
+         y=paste("Total T-bet+ CD4+ T cells (%)"))
+
+HC_T3<-ggplot(HC, aes(x=SM, y=TbetCXCR3))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
+    geom_jitter(width=.1,height=0, shape=16,size=2)+
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_x_discrete(labels=c("SM-", "SM+"))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(axis.text.x = element_text(vjust=0.5)) + 
+    theme(text = element_text(size=14)) +   
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TbetCXCR3)*1.1),
+                       label.x.npc = "center")+
+    labs(title="",
+         x="",
+         y=paste("T-bet+CXCR3+ CD4+ T cells (%)"))
+
+HC_T4<-ggplot(HC, aes(x=SM, y=TbetCCR4))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
+    geom_jitter(width=.1,height=0, shape=16,size=2)+
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_x_discrete(labels=c("SM-", "SM+"))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(axis.text.x = element_text(vjust=0.5)) + 
+    theme(text = element_text(size=14)) +   
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TbetCXCR3)*1.1),
+                       label.x.npc = "center")+
+    labs(title="",
+         x="",
+         y=paste("T-bet+CXCR3+ CD4+ T cells (%)"))
+
+HC_lin<-grid.arrange(HC_T, HC_T3, HC_T4, ncol=3, top=text_grob("Expression of TH1 Lineage Markers", size=20))
+HC_lin
+```
+
+###Difference aesthetics Same Cytokine Graph
+```{r}
+HC<-mutate_hc(filter(CD4, Stim=="PMA"))
+test<-melt(HC, id.vars=c("Donor", "TB", "SM"), measure.vars = c("TH1.Freq", "TH2.Freq", "TH1.2.Freq"))
+
+one<-ggplot(test, aes(x=SM, y=value, fill=TB, alpha=SM))+
+    geom_boxplot(size=1, position=position_dodge(width = 0.9), outlier.shape = NA) +
+      geom_point(position=position_jitterdodge(dodge.width=0.9)) +
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(text = element_text(size=14)) +
+    theme(strip.text.x = element_text(size = 12))+
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE,
+                       label.x.npc = "center")+
+    facet_wrap(~variable, scales="free")+
+    labs(title="Cytokine Frequencies",
+         x="",
+         y=paste("Freq of Cytokine+ CD4+ T cells (%)"))
+
+two<-ggplot(test, aes(x=SM, y=value, fill=TB, alpha=SM))+
+    geom_boxplot(size=1, position=position_dodge(width = 0.9), outlier.shape = NA) +
+      geom_point(position=position_jitterdodge(dodge.width=0.9)) +
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(text = element_text(size=14)) +
+    theme(strip.text.x = element_text(size = 12))+
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE,
+                       label.x.npc = "center")+
+    facet_wrap(~variable)+
+    labs(title="Cytokine Frequencies",
+         x="",
+         y=paste("Freq of Cytokine+ CD4+ T cells (%)"))
+
+three<-ggplot(test, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot(size=1, position=position_dodge(width = 0.9), outlier.shape = NA) +
+      geom_point(position=position_jitterdodge(dodge.width=0.9)) +
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(text = element_text(size=14)) +
+    theme(strip.text.x = element_text(size = 12))+
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE,
+                       label.x.npc = "center")+
+    labs(title="Cytokine Frequencies",
+         x="",
+         y=paste("Freq of Cytokine+ CD4+ T cells (%)"))
+
+one
+two
+three
+```
+
+##Different Ways to Visualize TH1 and TH2 cell marker expression
+```{r Figure 1b, hide=TRUE}
+HC<-mutate_hc(filter(phen, Stim=="PMA"));
+HC_1T<-ggplot(HC, aes(x=SM, y=TH1_Tbet))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
+    geom_jitter(width=.1,height=0, shape=16,size=2)+
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_x_discrete(labels=c("SM-", "SM+"))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(axis.text.x = element_text(vjust=0.5)) + 
+    theme(text = element_text(size=14)) +
+    theme(strip.text.x = element_text(size = 12))+
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH1_Tbet)*1.1),
+                       label.x.npc = "center")+
+    labs(title="TH1 Cells",
+         x="",
+         y=paste("Total T-bet+ CD4+ T cells (%)"))
+HC_2T<-ggplot(HC, aes(x=SM, y=TH2_Tbet))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
+    geom_jitter(width=.1,height=0, shape=16,size=2)+
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_x_discrete(labels=c("SM-", "SM+"))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(axis.text.x = element_text(vjust=0.5)) + 
+    theme(text = element_text(size=14)) +
+    theme(strip.text.x = element_text(size = 12))+
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH1_Tbet)*1.1),
+                       label.x.npc = "center")+
+    labs(title="TH2 Cells",
+         x="",
+         y=paste(""))
+
+HC_1G<-ggplot(HC, aes(x=SM, y=TH1_GATA3))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
+    geom_jitter(width=.1,height=0, shape=16,size=2)+
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_x_discrete(labels=c("SM-", "SM+"))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(axis.text.x = element_text(vjust=0.5)) + 
+    theme(text = element_text(size=14)) +
+    theme(strip.text.x = element_text(size = 12))+
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH2_GATA3)*1.1),
+                       label.x.npc = "center")+
+    labs(title="TH1 Cells",
+         x="",
+         y=paste("Total GATA3+ CD4+ T cells (%)"))
+HC_2G<-ggplot(HC, aes(x=SM, y=TH2_GATA3))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
+    geom_jitter(width=.1,height=0, shape=16,size=2)+
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_x_discrete(labels=c("SM-", "SM+"))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(axis.text.x = element_text(vjust=0.5)) + 
+    theme(text = element_text(size=14)) +
+    theme(strip.text.x = element_text(size = 12))+
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH2_GATA3)*1.1),
+                       label.x.npc = "center")+
+    labs(title="TH2 Cells",
+         x="",
+         y=paste(""))
+
+HC_13<-ggplot(HC, aes(x=SM, y=TH1_CXCR3))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
+    geom_jitter(width=.1,height=0, shape=16,size=2)+
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_x_discrete(labels=c("SM-", "SM+"))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(axis.text.x = element_text(vjust=0.5)) + 
+    theme(text = element_text(size=14)) +
+    theme(strip.text.x = element_text(size = 12))+
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH1_CXCR3)*1.1),
+                       label.x.npc = "center")+
+    labs(title="TH1 Cells",
+         x="",
+         y=paste("Total CXCR3+ CD4+ T cells (%)"))
+HC_23<-ggplot(HC, aes(x=SM, y=TH2_CXCR3))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
+    geom_jitter(width=.1,height=0, shape=16,size=2)+
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_x_discrete(labels=c("SM-", "SM+"))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(axis.text.x = element_text(vjust=0.5)) + 
+    theme(text = element_text(size=14)) +
+    theme(strip.text.x = element_text(size = 12))+
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH1_CXCR3)*1.1),
+                       label.x.npc = "center")+
+    labs(title="TH2 Cells",
+         x="",
+         y=paste(""))
+
+HC_14<-ggplot(HC, aes(x=SM, y=TH1_CCR4))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
+    geom_jitter(width=.1,height=0, shape=16,size=2)+
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_x_discrete(labels=c("SM-", "SM+"))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(axis.text.x = element_text(vjust=0.5)) + 
+    theme(text = element_text(size=14)) +
+    theme(strip.text.x = element_text(size = 12))+
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH2_CCR4)*1.1),
+                       label.x.npc = "center")+
+    labs(title="TH1 Cells",
+         x="",
+         y=paste("Total CXCR3+ CD4+ T cells (%)"))
+HC_24<-ggplot(HC, aes(x=SM, y=TH2_CCR4))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
+    geom_jitter(width=.1,height=0, shape=16,size=2)+
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_x_discrete(labels=c("SM-", "SM+"))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(axis.text.x = element_text(vjust=0.5)) + 
+    theme(text = element_text(size=14)) +
+    theme(strip.text.x = element_text(size = 12))+
+    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH2_CCR4)*1.1),
+                       label.x.npc = "center")+
+    labs(title="TH2 Cells",
+         x="",
+         y=paste(""))
+```
+
+```{r, fig.width=16, fig.height=8}
+one.b<-grid.arrange(
+    HC_T<-grid.arrange(HC_1T, HC_2T, ncol=2, top=text_grob("T-bet", size=20)),
+    HC_3<-grid.arrange(HC_13, HC_23, ncol=2, top=text_grob("CXCR3", size=20)),
+    HC_G<-grid.arrange(HC_1G, HC_2G, ncol=2, top=text_grob("GATA3", size=20)),
+    HC_4<-grid.arrange(HC_14, HC_24, ncol=2, top=text_grob("CCR4", size=20)),
+    ncol=4)
+```
+
+```{r}
+HC<-mutate_hc(filter(phen, Stim=="PMA"));
+vars <- names(HC)[grep("TH1_|TH2_", names(HC))]
+vars1 <- c("TH1_Tbet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4",
+           "TH2_Tbet", "TH2_GATA3", "TH2_CXCR3", "TH2_CCR4")
+vars2 <- vars[grep(".P|.N", vars)]
+vars3 <- vars[grep("._._._.", vars)]
+
+```
+
+```{r}
+test<-melt(HC, id.vars=c("Donor", "TB", "SM"), vars1)
+test$subset<-tstrsplit(test$variable, "_")[[1]]
+test$subset<-paste(test$subset, "Cytokine + CD4+ T Cells", sep=" ")
+test$variable<-gsub("TH1_","", test$variable)
+test$variable<-gsub("TH2_","", test$variable)
+
+plot_markers<-function(df, marker){
+    ggplot(df[df$variable==marker,], aes(x=SM, y=value, fill=TB, alpha=SM))+
+    geom_boxplot(size=1, position=position_dodge(width = 0.9), outlier.shape = NA) +
+    geom_point(position=position_jitterdodge(dodge.width=0.9)) +
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(text = element_text(size=14)) +
+    theme(strip.text = element_text(size = 14))+
+    facet_grid(variable~subset, scales="free")+
+    labs(y="", x="")+
+    stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE,
+                       label.x.npc = "center")}
+
+grid.arrange(
+    plot_markers(test, "Tbet")+theme(axis.text.x = element_blank()),
+    plot_markers(test, "GATA3")
+        +theme(axis.text.x = element_blank())+theme(strip.text.x = element_blank()),
+    plot_markers(test, "CXCR3")
+        +theme(axis.text.x = element_blank())+theme(strip.text.x = element_blank()),
+    plot_markers(test, "CCR4")
+        +theme(strip.text.x = element_blank())+        scale_x_discrete(labels=c("SM-", "SM+")),
+    ncol=1
+)
+```
+
+```{r fig.width=16, fig.height=10}
+test<-melt(HC, id.vars=c("Donor", "TB", "SM"), vars2)
+test$subset<-tstrsplit(test$variable, "_")[[1]]
+test$subset<-paste(test$subset, "Cytokine Producing CD4+ T Cells", sep=" ")
+test$variable<-gsub("TH1_","", test$variable)
+test$variable<-gsub("TH2_","", test$variable)
+test<-filter(test, variable!="CCR4.N_CXCR3.N" , variable!="GATA3.N_Tbet.N")
+
+ggplot(test, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot(size=1, position=position_dodge(width = 0.9), outlier.shape = NA) +
+    geom_point(position=position_jitterdodge(dodge.width=0.9)) +
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(text = element_text(size=14)) +
+    theme(strip.text.x = element_text(size = 20))+
+    facet_wrap(~subset, ncol=1)+
+    stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE,
+                       label.x.npc = "center")
+
+test<-melt(HC, id.vars=c("Donor", "TB", "SM"), vars3)
+test$subset<-tstrsplit(test$variable, "_")[[1]]
+test$subset<-paste(test$subset, "Cytokine Producing CD4+ T Cells", sep=" ")
+test$variable<-gsub("TH1_","", test$variable)
+test$variable<-gsub("TH2_","", test$variable)
+test<-filter(test, -x_x_x_x)
+
+ggplot(test, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot(size=1, position=position_dodge(width = 0.9), outlier.shape = NA) +
+    geom_point(position=position_jitterdodge(dodge.width=0.9)) +
+    scale_fill_manual(values="#1a9850")+
+    scale_alpha_manual(values=c(0.5,1))+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5))+
+    theme(text = element_text(size=14)) +
+    theme(strip.text.x = element_text(size = 20))+
+    facet_wrap(~subset, ncol=1)+
+    stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 4, 
+                       method="wilcox.test", paired = FALSE,
+                       label.x.npc = "center")
+```
+
+#Proof that we can combine hc and ltbi
 ##Lineage Markers on UN
@@
 ```
-
-##Actual analysis

### 2019-07-02 ICS_Phenotyping.Rmd
boolean graphs for phenotype

@@
 source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
 ```
@@
     #will bind together into a new dataframe the columns that contain the key in the name from
-    #1. the original entries in the mimosa data frame, 2. the untouched phenotype data and 3. the new phenotype data with NA values
+    #1. the original entries in the mimosa data frame, 2. the untouched phenotype data and 3. the new DF<-subset(CD4, CD4$Stim=="PEP")
     test<-data.frame(cbind(mimosa[grep(key, names(mimosa))], 
@@
 
+```{r, fig.width=16, fig.height=8}
+test<-mutate_TB(filter(DF, !is.na(TH1_GATA3)))
+test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = grep("TH1_._._._.", names(DF)))
+test2<-filter(test2, variable!="TH1_x_x_x_x")
+test2$variable<-gsub("TH1_", "", test2$variable)
+ggplot(test2, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot()+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_fill_manual(values =c("#2166ac", "#b2182b"))+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    labs(title="Lineage Marker Expression on Mtb-Specific TH1 CD4+ T cells",
+         y="Frequency of Marker+ Mtb-Specific T cells")
+kable(xtabs(~TB+SM, test))
+```
+
+```{r, fig.width=16, fig.height=8}
+test<-mutate_TB(filter(DF, !is.na(TH2_GATA3)))
+test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = grep("TH2_._._._.", names(DF)))
+test2<-filter(test2, variable!="TH2_x_x_x_x")
+test2$variable<-gsub("TH2_", "",test2$variable)
+ggplot(test2, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot()+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_fill_manual(values =c("#2166ac", "#b2182b"))+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    labs(title="Lineage Marker Expression on Mtb-Specific TH2 CD4+ T cells",
+         y="Frequency of Marker+ Mtb-Specific T cells")
+kable(xtabs(~TB+SM, test))
+```
+
 ```{r}
@@
 plot_2(test, "disease", "TH1_x_x_3_x")
+plot_2(test, "disease", "TH1_GATA3CXCR3")
+plot_2(test, "disease", "TH1_GATA3CCR4")
+```
 
+```{r, fig.width=12, fig.height=6}
 grid.arrange(plot_2(test, "disease", "TH1_t_x_x_x")+labs(y="Tbet+ CD4+ T cells"),
@@
 library(ggplot2)
-data<-filter(phenotype, Stim %in% c("PMA", "PEP", "WCL"))
+data<-mutate_TB(filter(phenotype, Stim %in% c("PMA", "PEP", "WCL")))
 data$Stim<-factor(data$Stim, levels=c("PMA","PEP","WCL"))

### 2019-07-02 OG_cleaning.Rmd
add OG cleaning functions

@@
 
+```{r create new combo variables from exported boolean phenotype}
+l<-lapply(ls(pattern="bulk_summary"), function(x) get(x))
+l<-lapply(l, function(x) sum_boolean(x))
+filenames<-ls(pattern="bulk_summary")
+for(i in 1:5){
+assign(filenames[i], data.frame(l[i]))}
+
+l<-lapply(ls(pattern="_OG"), function(x) get(x))
+l<-lapply(l, function(x) sum_boolean(x))
+filenames<-ls(pattern="_OG")
+for(i in 1:9){
+assign(filenames[i], data.frame(l[i]))}
+```
+
+```{r}
+l<-lapply(ls(pattern="bulk_summary"), function(x) get(x))
+l<-lapply(l, function(x) OG_background.subtract(x, "Donor", "Stim", "UN", "OG_freq"))
+filenames<-ls(pattern="bulk_summary")
+for(i in 1:5){
+assign(filenames[i], data.frame(l[i]))}
+```
+
 ```{r}

### 2019-07-02 OG-Analysis.Rmd
completely overhauled. based on ICS

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment=NA)
+```
+
+```{r}
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
+source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")
 ```
@@
 setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
-CD4_summary<-read.csv("OG_CD4-bulk_summary.csv")
-CD4_summary$SM<-factor(CD4_summary$SM, levels=c("X", "SM"))
-CD4_OG_summary<-read.csv("OG_CD4-OG_summary.csv")
-CD4_OG_summary$SM<-factor(CD4_OG_summary$SM, levels=c("X", "SM"))
-PBMC<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/exported/OG_PBMC.csv")
+CD4<-dplyr::select(read.csv("OG_CD4_bulk_summary_clean.csv"), -X, -X.1)
+CD4_OG<-dplyr::select(read.csv("OG_CD4_OG_summary_clean.csv"), -X, -X.1)
+CD4_OG_TH1<-dplyr::select(read.csv("OG_CD4_OG.TH1_summary_clean.csv"), -X, -X.1)
+CD4_OG_GT<-dplyr::select(read.csv("OG_CD4_OG.TH1_summary_clean.csv"), -X, -X.1)
+CD4_OG_TNF<-dplyr::select(read.csv("OG_CD4_OG.TNFa_summary_clean.csv"), -X, -X.1)
+PBMC<-dplyr::select(read.csv("DonorInfo_clean.csv"), -X)
+    PBMC<-PBMC[grep("OG", PBMC$used),]
+mimosa<-dplyr::select(read.csv("OG_MIMOSA_clean.csv"), -X)
+    mimosa<-filter(mimosa, celltype=="CD4" & Exprate==0.01)
 ```
 
-```{r functions and stuff}
-source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
-source("/Users/tarynam/Desktop/Miscellaneous-Scripts/Plotting_1902.R")
-```
-
-```{r PBMC}
-PBMC<-filter(PBMC, !is.na(Study.ID))
-PBMC<-filter(PBMC, Helminth.Status %in% c("Negative", "Schisto+"))
-PBMC$Helminth.Status<-factor(PBMC$Helminth.Status, levels=c("Negative", "Schisto+"))
-PBMC$TB.Status<-factor(PBMC$TB.Status, levels=c("Healthy Control", "LTBI", "Active TB"))
-PBMC<-filter(PBMC, !(Study.ID=="NK2208" & Exp.Name=="TM181216"))
-PBMC<-filter(PBMC, !(Study.ID=="NK2306" & Exp.Name=="TM181206"))
-PBMC<-filter(PBMC, !(Study.ID=="NK2353" & Exp.Name=="TM190411"))
-PBMC<-filter(PBMC, !(Study.ID=="NK2384" & Exp.Name=="TM190411"))
-PBMC$TB<-PBMC$TB.Status
-PBMC$SM<-PBMC$Helminth.Status
-
-grid.arrange(plot_6(PBMC, "SM", "Viability")+labs(y="Viability"),
-    plot_6(PBMC, "SM", "Cells.Vial")+labs(y="Cell Count"),
-    ncol=2
-)
-
-plot_pairs<-function(data, xval, yval){
-    ggplot(data, aes(x=data[,xval], y=data[,yval]))+
-        geom_point(size=3, aes(color=TB, alpha=SM)) +
-        geom_line(aes(group=Study.ID))+
-        scale_color_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
-        scale_alpha_manual(values=c(0.5,1))+
-        facet_grid(~SM)+
-        theme_bw()+ theme(legend.position="bottom")+
-        theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-        theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
-        facet_grid(SM~TB, scale="free")+
-        stat_compare_means(label = "p.format", p.adjust.method = "fdr", hide.ns = TRUE, size = 4, 
-                           method="wilcox.test", paired = TRUE, label.y= (max(data[,yval])*1.1))
+#pep
+```{r pep total}
+library(dplyr)
+library(knitr)
+DF<-subset(CD4, CD4$Stim=="WCL")
+colnames<-names(numeric_only(DF))
+for(col in colnames){
+    y<-pvals_4_loose(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
 }
-
-melt<-melt(PBMC, id.vars=c("Study.ID", "TB", "SM"), measure.vars = c("Viability", "OG.Viability"))
-plot_pairs(melt, "variable", "value")+labs(y="Viability", x="Time")
-
-melt<-melt(PBMC, id.vars=c("Study.ID", "TB", "SM"), measure.vars = c("Cells.Vial", "OG.Cell.."))
-plot_pairs(melt, "variable", "value")+labs(y="Count", x="Time")
 ```
@@
 ```{r}
+test<-mutate_TB(filter(DF, !is.na(OG_freq)))
+kable(xtabs(~TB+SM, test))
+plot_2(test, "disease", "OG_freq")
+```
+
+#Removing Low Responders
+```{r}
 library(dplyr)
-antigen.ref<-MIMOSA.Results
-antigen.ref$id<-rownames(antigen.ref)
-antigen.ref$id<-gsub("_Treatment", "", antigen.ref$id)
+antigen.ref<-mimosa
+antigen.ref$id<-gsub("_Treatment", "", antigen.ref$Sample)
 antigen.ref<-arrange(antigen.ref, id)
 
-phenotype<-CD4_OG_summary
+#Total OG
+phenotype<-CD4_OG
 phenotype$id<-paste(phenotype$Donor, phenotype$Stim, sep="_")
@@
 
-remove<-subset(antigen.ref, antigen.ref[,"Pr.response"]<.7)$id
+remove<-subset(antigen.ref, antigen.ref[,"Pr.response"]<.7 & antigen.ref[,"fdr"]>0.03)$id
 print(length(remove)) #see how many sample ids will be excluded
@@
 phenotype[(phenotype$id %in% remove), numeric.only(phenotype)]<-NA
+OG<-phenotype
+
+#OG GT
+phenotype<-CD4_OG_GT
+phenotype$id<-paste(phenotype$Donor, phenotype$Stim, sep="_")
+phenotype<-arrange(phenotype, id)
+
+remove<-subset(antigen.ref, antigen.ref[,"Pr.response"]<.7 & antigen.ref[,"fdr"]>0.03)$id
+print(length(remove)) #see how many sample ids will be excluded
+numeric.only <- function(X,...){
+      returnCols <- names(X)
+      a<-sapply(X, is.numeric)
+      print(returnCols[a == "TRUE"])
+}
+
+phenotype[(phenotype$id %in% remove), numeric.only(phenotype)]<-NA
+GT<-phenotype
+
+#OG TH1
+phenotype<-CD4_OG_TH1
+phenotype$id<-paste(phenotype$Donor, phenotype$Stim, sep="_")
+phenotype<-arrange(phenotype, id)
+
+remove<-subset(antigen.ref, antigen.ref[,"Pr.response"]<.7 & antigen.ref[,"fdr"]>0.03)$id
+print(length(remove)) #see how many sample ids will be excluded
+numeric.only <- function(X,...){
+      returnCols <- names(X)
+      a<-sapply(X, is.numeric)
+      print(returnCols[a == "TRUE"])
+}
+
+phenotype[(phenotype$id %in% remove), numeric.only(phenotype)]<-NA
+TH1<-phenotype
+
+#OG TNF
+phenotype<-CD4_OG_TNF
+phenotype$id<-paste(phenotype$Donor, phenotype$Stim, sep="_")
+phenotype<-arrange(phenotype, id)
+
+remove<-subset(antigen.ref, antigen.ref[,"Pr.response"]<.7 & antigen.ref[,"fdr"]>0.03)$id
+print(length(remove)) #see how many sample ids will be excluded
+numeric.only <- function(X,...){
+      returnCols <- names(X)
+      a<-sapply(X, is.numeric)
+      print(returnCols[a == "TRUE"])
+}
+
+phenotype[(phenotype$id %in% remove), numeric.only(phenotype)]<-NA
+TNF<-phenotype
 ```
 
-#Pep
-```{r}
-df<-filter(CD4_summary, Stim=="PEP")
-top<-ggplot(df, aes(x=SM, y=OG_freq, fill=TB, alpha=SM))+
-    geom_jitter(width=.1,height=0, shape=16,size=3)+
-    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,1))+
-    facet_grid(~TB)+
-    theme_classic()+ 
-    theme(legend.position="none")+
-    theme(axis.title.x=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank())+
-    theme(text = element_text(size=20)) +   
-    facet_grid(~TB)+
-    labs(y="", x="")+
-    scale_x_discrete(labels=c("SM-", "SM+"))+
-    stat_compare_means(label = "p.format")+
-    coord_cartesian(ylim=c(2,30))
-
-bottom<-ggplot(df, aes(x=SM, y=OG_freq, fill=TB, alpha=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB)) +
-    geom_jitter(width=.1,height=0, shape=16,size=3)+
-    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
-    scale_alpha_manual(values=c(0.5,1))+
-    facet_grid(~TB)+
-    theme_classic()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    theme(strip.background = element_blank(),strip.text.x = element_blank())+
-    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
-    facet_grid(~TB)+
-    labs(y="", x="")+
-    scale_x_discrete(labels=c("SM-", "SM+"))+
-    coord_cartesian(ylim=c(0,2))
-
-grid.arrange(top, bottom, ncol=1, heights=c(4,8))
-
-df<-filter(phenotype, Stim=="PEP", TB!="HC")
-test<-filter(phenotype, Stim=="PEP", TB!="HC", !is.na(IFNg_freq))
-kable(xtabs(~TB+SM, test))
-df$SM<-factor(df$SM, levels=c("X", "SM"))
-grid.arrange(plot_2(df, "SM", "IFNg_freq")+labs(y="IFNg+ Proliferating CD4+ T cells (%)", x=""),
-             plot_2(df, "SM", "TNFa_freq")+labs(y="TNFa+ Proliferating CD4+ T cells (%)", x=""),
-             plot_2(df, "SM", "IL4_freq")+labs(y="IL-4+ Proliferating CD4+ T cells (%)", x=""),
-             ncol=3
-
-)
-
-grid.arrange(plot_2(df, "SM", "CXCR3_freq")+labs(y="CXCR3+ Proliferating CD4+ T cells (%)", x=""),
-             plot_2(df, "SM", "CCR4_freq")+labs(y="CCR4+ Proliferating CD4+ T cells (%)", x=""),
-             plot_2(df, "SM", "Tbet_freq")+labs(y="Tbet+ Proliferating CD4+ T cells (%)", x=""),
-             plot_2(df, "SM", "GATA3_freq")+labs(y="GATA3+ Proliferating CD4+ T cells (%)", x=""),
-             ncol=4
-
-)
+#SEB
+```{r SEB}
+library(dplyr)
+library(knitr)
+DF<-subset(OG, OG$Stim=="SEB")
+#split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
+x<-split(DF, list(DF$TB, DF$SM))
+#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0.
+It's a weird double negative.
+But basically makes a list of the columns that have sufficient data.
+y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
+#each entry in the list has to be extracted and named
+a<-unlist(y[1])
+b<-unlist(y[2])
+c<-unlist(y[3])
+d<-unlist(y[4])
+#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
+test<-Reduce(intersect, list(a,b,c,d))
+#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
+colnames<-c(test[1:31], test[38:49])
+for(col in colnames){
+    y<-pvals_4_loose(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
+}
 ```
 
-#WCL
 ```{r}
-df<-filter(CD4_summary, Stim=="WCL")
-top<-ggplot(df, aes(x=SM, y=OG_freq, fill=TB, alpha=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), 
-        outlier.shape = NA, aes(fill=TB)) +geom_jitter(width=.1,height=0, shape=16,size=3)+
-    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
+test<-mutate_TB(filter(DF, !is.na(OG_freq)))
+kable(xtabs(~TB+SM, test))
+```
+
+```{r}
+vars<-names(test)[grep("._._.", names(test))]
+vars<-vars[-grep("._._._.", vars)]
+test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
+test2<-filter(test2, variable!="x_x_x")
+ggplot(test2, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot()+
     scale_alpha_manual(values=c(0.5,1))+
-    facet_grid(~TB)+
-    theme_classic()+ 
+    scale_fill_manual(values =c("#2166ac", "#b2182b"))+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
     theme(legend.position="none")+
-    theme(axis.title.x=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank())+
-    theme(text = element_text(size=20)) +   
-    facet_grid(~TB)+
-    labs(y="", x="")+
-    scale_x_discrete(labels=c("SM-", "SM+"))+
-    stat_compare_means(label = "p.format")+
-    coord_cartesian(ylim=c(5,50))
+    theme(plot.title = element_text(hjust = 0.5)) +
+    labs(title="Cytokine Production by Proliferating CD4+ T cells",
+         y="Frequency of Cytokine+ CD4+ T cells")
+```
 
-bottom<-ggplot(df, aes(x=SM, y=OG_freq, fill=TB, alpha=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), 
-                 outlier.shape = NA, aes(fill=TB)) +
-    geom_jitter(width=.1,height=0, shape=16,size=3)+
-    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
+```{r, fig.width=16, fig.height=8}
+vars<-names(test)[grep("._._._.", names(test))]
+test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
+test2<-filter(test2, variable!="x_x_x_x")
+ggplot(test2, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot()+
     scale_alpha_manual(values=c(0.5,1))+
-    facet_grid(~TB)+
-    theme_classic()+ theme(legend.position="none")+
-    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
-    theme(strip.background = element_blank(),strip.text.x = element_blank())+
-    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
-    facet_grid(~TB)+
-    labs(y="", x="")+
-    scale_x_discrete(labels=c("SM-", "SM+"))+
-    coord_cartesian(ylim=c(0,5))
+    scale_fill_manual(values =c("#2166ac", "#b2182b"))+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    labs(title="Lineage Marker Expression on Proliferating CD4+ T cells",
+         y="Frequency of Marker+ CD4+ T cells")
+```
 
-grid.arrange(top, bottom, ncol=1, heights=c(4,6))
+##PEP
+```{r PEP}
+library(dplyr)
+library(knitr)
+DF<-subset(OG, OG$Stim=="PEP")
+#split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
+x<-split(DF, list(DF$TB, DF$SM))
+#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0.
+It's a weird double negative.
+But basically makes a list of the columns that have sufficient data.
+y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
+#each entry in the list has to be extracted and named
+a<-unlist(y[1])
+b<-unlist(y[2])
+c<-unlist(y[3])
+d<-unlist(y[4])
+#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
+test<-Reduce(intersect, list(a,b,c,d))
+#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
+colnames<-c(test[1:31], test[38:49])
+for(col in colnames){
+    y<-pvals_4_loose(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
+}
+```
 
-df<-filter(phenotype, Stim=="WCL", TB!="HC")
-test<-filter(phenotype, Stim=="WCL", TB!="HC", !is.na(IFNg_freq))
+```{r}
+test<-mutate_TB(filter(DF, !is.na(OG_freq)))
 kable(xtabs(~TB+SM, test))
+```
 
-grid.arrange(plot_2(df, "SM", "IFNg_freq")+labs(y="IFNg+ Proliferating CD4+ T cells (%)", x=""),
-             plot_2(df, "SM", "TNFa_freq")+labs(y="TNFa+ Proliferating CD4+ T cells (%)", x=""),
-             plot_2(df, "SM", "IL4_freq")+labs(y="IL-4+ Proliferating CD4+ T cells (%)", x=""),
-             ncol=3
+```{r}
+vars<-names(test)[grep("._._.", names(test))]
+vars<-vars[-grep("._._._.", vars)]
+test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
+test2<-filter(test2, variable!="x_x_x")
+ggplot(test2, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot()+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_fill_manual(values =c("#2166ac", "#b2182b"))+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    labs(title="Cytokine Production by Mtb-Specific CD4+ T cells",
+         y="Frequency of Cytokine+ Mtb-Specific T cells")
+```
 
-)
+```{r, fig.width=16, fig.height=8}
+vars<-names(test)[grep("._._._.", names(test))]
+test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
+test2<-filter(test2, variable!="x_x_x_x")
+ggplot(test2, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot()+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_fill_manual(values =c("#2166ac", "#b2182b"))+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    labs(title="Lineage Marker Expression on Mtb-Specific CD4+ T cells",
+         y="Frequency of Marker+ Mtb-Specific T cells")
+```
 
-grid.arrange(plot_2(df, "SM", "CXCR3_freq")+labs(y="CXCR3+ Proliferating CD4+ T cells (%)", x=""),
-             plot_2(df, "SM", "CCR4_freq")+labs(y="CCR4+ Proliferating CD4+ T cells (%)", x=""),
-             plot_2(df, "SM", "Tbet_freq")+labs(y="Tbet+ Proliferating CD4+ T cells (%)", x=""),
-             plot_2(df, "SM", "GATA3_freq")+labs(y="GATA3+ Proliferating CD4+ T cells (%)", x=""),
-             ncol=4
+```{r}
+plot_2(test, "disease", "x_G_4")
+```
 
-)
+##WCL
+```{r WCL}
+library(dplyr)
+library(knitr)
+DF<-subset(OG, OG$Stim=="WCL")
+#split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
+x<-split(DF, list(DF$TB, DF$SM))
+#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0.
+It's a weird double negative.
+But basically makes a list of the columns that have sufficient data.
+y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
+#each entry in the list has to be extracted and named
+a<-unlist(y[1])
+b<-unlist(y[2])
+c<-unlist(y[3])
+d<-unlist(y[4])
+#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
+test<-Reduce(intersect, list(a,b,c,d))
+#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
+colnames<-c(test[1:31], test[38:49])
+for(col in colnames){
+    y<-pvals_4_loose(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
+}
 ```
+
+```{r}
+test<-mutate_TB(filter(DF, !is.na(OG_freq)))
+kable(xtabs(~TB+SM, test))
+```
+
+```{r}
+vars<-names(test)[grep("._._.", names(test))]
+vars<-vars[-grep("._._._.", vars)]
+test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
+test2<-filter(test2, variable!="x_x_x")
+ggplot(test2, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot()+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_fill_manual(values =c("#2166ac", "#b2182b"))+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    labs(title="Cytokine Production by Mtb-Specific CD4+ T cells",
+         y="Frequency of Cytokine+ Mtb-Specific T cells")
+```
+
+```{r, fig.width=16, fig.height=8}
+vars<-names(test)[grep("._._._.", names(test))]
+test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
+test2<-filter(test2, variable!="x_x_x_x")
+ggplot(test2, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot()+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_fill_manual(values =c("#2166ac", "#b2182b"))+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    labs(title="Lineage Marker Expression on Mtb-Specific CD4+ T cells",
+         y="Frequency of Marker+ Mtb-Specific T cells")
+```
+
+```{r}
+plot_2(test, "disease", "IL4_freq")
+plot_2(test, "disease", "T_x_4")
+plot_2(test, "disease", "T_x_x")
+plot_2(test, "disease", "x_G_x")
+plot_2(test, "disease", "x_x_x_4")
+```
+
+##PEP
+```{r PEP th1}
+library(dplyr)
+library(knitr)
+DF<-subset(TH1, TH1$Stim=="PEP")
+#split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
+x<-split(DF, list(DF$TB, DF$SM))
+#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0.
+It's a weird double negative.
+But basically makes a list of the columns that have sufficient data.
+y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
+#each entry in the list has to be extracted and named
+a<-unlist(y[1])
+b<-unlist(y[2])
+c<-unlist(y[3])
+d<-unlist(y[4])
+#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
+test<-Reduce(intersect, list(a,b,c,d))
+#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
+colnames<-c(test[1:31], test[38:49])
+for(col in colnames){
+    y<-pvals_4_loose(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
+}
+```
+
+##PEP
+```{r PEP tnfa}
+library(dplyr)
+library(knitr)
+DF<-subset(TNF, TNF$Stim=="PEP")
+#split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
+x<-split(DF, list(DF$TB, DF$SM))
+#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0.
+It's a weird double negative.
+But basically makes a list of the columns that have sufficient data.
+y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
+#each entry in the list has to be extracted and named
+a<-unlist(y[1])
+b<-unlist(y[2])
+c<-unlist(y[3])
+d<-unlist(y[4])
+#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
+test<-Reduce(intersect, list(a,b,c,d))
+#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
+colnames<-c(test[1:31], test[38:49])
+for(col in colnames){
+    y<-pvals_4_loose(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
+}
+```
+
+##PEP
+```{r PEP gt}
+library(dplyr)
+library(knitr)
+DF<-subset(GT, GT$Stim=="PEP")
+#split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
+x<-split(DF, list(DF$TB, DF$SM))
+#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0.
+It's a weird double negative.
+But basically makes a list of the columns that have sufficient data.
+y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
+#each entry in the list has to be extracted and named
+a<-unlist(y[1])
+b<-unlist(y[2])
+c<-unlist(y[3])
+d<-unlist(y[4])
+#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
+test<-Reduce(intersect, list(a,b,c,d))
+#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
+colnames<-c(test[1:31], test[38:49])
+for(col in colnames){
+    y<-pvals_4_loose(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
+}
+```

### 2019-07-02 MIMOSA.R
mimosa moved

@@ -1,149 +1,0 @@
-install.packages("devtools") #install devtools package
-library(devtools) #load it
-install_github("MIMOSA","RGLab",branch="master") #Need R 3.0.0, and a bunch of dependencies. The install will fail with various error messages until you install those dependencies.
-library(MIMOSA)
-
-setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
-DF<-read.csv("CD4_Counts_Clean.csv")
-
-#All cytokine
-data<-dplyr::select(DF, Sample, Donor, TB, SM, Stim, Cyt, Total) %>%
-    dplyr::mutate(Neg = (Total - Cyt)) %>%
-    dplyr::select(-Total)
-
-E<-ConstructMIMOSAExpressionSet(data,
-                                reference=Stim %in% 'UN',
-                                measure.columns=c('Neg','Cyt'),
-                                other.annotations=c('Stim','Sample', 'Donor', 'TB', 'SM'),
-                                default.cast.formula=component~Donor+Stim,
-                                .variables=.(Donor))
-test <- MIMOSA(Neg+Cyt~Donor|Stim,
-               data=E,
-               method="mcmc")
-
-probabilities<-MIMOSA::getZ(test)
-counts<-MIMOSA::countsTable(test)
-fdr<-MIMOSA::fdr(test)
-MIMOSA.Results<-data.frame(cbind(counts,probabilities,fdr))
-MIMOSA.summary<-MIMOSA::getW(test)
-
-#TH1 
-data<-dplyr::select(DF, Sample, Donor, TB, SM, Stim, TH1, Total) %>%
-    dplyr::mutate(Neg = (Total - TH1)) %>%
-    dplyr::select(-Total)
-
-E1<-ConstructMIMOSAExpressionSet(data,
-                                reference=Stim %in% 'UN',
-                                measure.columns=c('Neg','TH1'),
-                                other.annotations=c('Stim','Sample', 'Donor', 'TB', 'SM'),
-                                default.cast.formula=component~Donor+Stim,
-                                .variables=.(Donor))
-
[…]

### 2019-07-02 MIMOSA_Morgan.R

@@ -1,102 +1,0 @@
-biocLite("devtools")
-library(devtools) #load it
-install_github("MIMOSA","RGLab",branch="master") #Need R 3.0.0, and a bunch of dependencies. The install will fail with various error messages until you install those dependencies.
-biocLite("MIMOSA")
-library(MIMOSA)
-
-DF<-read.csv("/Users/tarynam/Desktop/QFT-_HIV+_donor_file/CD8_TB_HIVexport_180824.csv")
-DF<-as.data.table(DF) 
-
-
-#GT
-DF$Neg<-DF$COUNT-DF$GT_Count
-DF<-DF[DF$CONDITION!="SEB"]
-data<-dplyr::select(DF, DONOR, CONDITION, DISEASE, Neg, GT_Count)
-E<-ConstructMIMOSAExpressionSet(data,
-                                reference=CONDITION%in%'unstim',
-                                measure.columns=c('Neg','GT_Count'),
-                                other.annotations=c('CONDITION','DONOR'),
-                                default.cast.formula=component~DONOR+CONDITION,
-                                .variables=.(DONOR))
-test <- MIMOSA(Neg+GT_Count~DONOR|CONDITION,
-               data=E, method="mcmc")
-probabilities<-MIMOSA::getZ(test)
-counts<-MIMOSA::countsTable(test)
-fdr<-MIMOSA::fdr(test)
-MIMOSA.Results1<-cbind(counts,probabilities,fdr)
-MIMOSA.summary<-MIMOSA::getW(test)
-
-#TNF-any
-DF$Neg<-DF$COUNT-DF$TNFa_Count
-DF<-DF[DF$CONDITION!="SEB"]
-data<-dplyr::select(DF, DONOR, CONDITION, DISEASE, Neg, TNFa_Count)
-E<-ConstructMIMOSAExpressionSet(data,
-                                reference=CONDITION%in%'unstim',
-                                measure.columns=c('Neg','TNFa_Count'),
-                                other.annotations=c('CONDITION','DONOR'),
-                                default.cast.formula=component~DONOR+CONDITION,
-                                .variables=.(DONOR))
-test <- MIMOSA(Neg+TNFa_Count~DONOR|CONDITION,
-               data=E, method="mcmc")
-probabilities<-MIMOSA::getZ(test)
-counts<-MIMOSA::countsTable(test)
-fdr<-MIMOSA::fdr(test)
[…]

### 2019-07-28 OG-Analysis.Rmd
Tested HC bulk and SEB

@@
 mimosa<-dplyr::select(read.csv("OG_MIMOSA_clean.csv"), -X)
-    mimosa<-filter(mimosa, celltype=="CD4" & Exprate==0.01)
+    mimosa<-dplyr::filter(mimosa, celltype=="CD4" & Exprate==0.01)
+```
+
+#HC
+went through total cd4 un and seb --> no dif  
+went through og cd4 seb --> no dif  
+want the graph just in case
+
+maybe this can go in a supp figure or something...  
+went through og_GT seb and there are more tbet+gata3+ccr4+ cells  
+went through og_TH1 seb and there are more tbet+gata3+ccr4+ cells  
+went through og_TNF seb and there are more tbet+gata3+ccr4+ cells
+
+```{r HC}
+library(dplyr)
+library(knitr)
+DF<-mutate_hc(subset(CD4, CD4$Stim=="SEB"))
+plot_hc(DF, "SM","OG_freq")
 ```
@@
 library(knitr)
-DF<-subset(CD4, CD4$Stim=="WCL")
-colnames<-names(numeric_only(DF))
+DF<-subset(CD4, CD4$Stim=="PEP")
+colnames<-numeric_only(DF)
 for(col in colnames){

### 2019-07-28 Cytokine_Analysis.Rmd
allows for p-values that are <0.05 before correction
Fixed Graphs since changed graphing function

@@
 ```{r}
+df<-mutate_TB(CD4)
 grid.arrange(
@@
 DF<-dplyr::filter(Compass, TB != "N" & Stim == "WCL" & celltype == "CD4")
+my_comparisons <- list(c("LTBI SM", "LTBI X"), c("TB SM", "TB X"),
+                       c("TB SM", "LTBI SM"), c("TB X", "LTBI X"))
 
@@
 library(knitr)
-colnames<-c("Cyt.Freq", "Th1.Freq", "TH1.2.Freq", "TH2.Freq")
+colnames<-c("Cyt.Freq", "TH1.Freq", "TH1.2.Freq", "TH2.Freq")
 ### PMA
@@
     name<-gsub("Freq","", name)
-    y<-plot_with_all_stats(CD4, "PMA", col)+
+    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="PMA"), col)+
         labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
@@
     name<-gsub("Freq","", name)
-    y<-plot_with_all_stats(CD4, "WCL", col)+
+    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="WCL"), col)+
         labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
@@
     name<-gsub("Freq","", name)
-    y<-plot_with_all_stats(CD4, "PEP", col)+
+    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="PEP"), col)+
         labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
@@
     #inserts each character string into the column argument of the pvals function written above
-    y<-plot_with_all_stats(CD4, "PMA", col)+
+    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="PMA"), col)+
         labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
@@
     name<-gsub("Freq","", name)
-    y<-plot_with_all_stats(CD4, "WCL", col)+
+    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="WCL"), col)+
         labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
@@
     name<-gsub("Freq","", name)
-    y<-plot_with_all_stats(CD4, "PEP", col)+
+    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="PEP"), col)+
         labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
@@
     name<-gsub("_", "+", col)
-    y<-plot_with_all_stats(CD4, "PMA", col)+
+    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="PMA"), col)+
         labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
@@
     name<-gsub("_", "+", col)
-    y<-plot_with_all_stats(CD4, "WCL", col)+
+    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="WCL"), col)+
         labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
@@
     name<-gsub("_", "+", col)
-    y<-plot_with_all_stats(CD4, "WCL", col)+
+    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="PEP"), col)+
         labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 

### 2019-07-29 OG-Analysis.Rmd
added in a IFNg subset

@@
 CD4_OG_TNF<-dplyr::select(read.csv("OG_CD4_OG.TNFa_summary_clean.csv"), -X, -X.1)
+CD4_OG_IFNg<-dplyr::select(read.csv("OG_CD4_OG.IFNg_summary_clean.csv"), -X, -X.1)
 PBMC<-dplyr::select(read.csv("DonorInfo_clean.csv"), -X)
@@
 
-#OG TNF
+#OG IFNg
+phenotype<-CD4_OG_IFNg
+phenotype$id<-paste(phenotype$Donor, phenotype$Stim, sep="_")
+phenotype<-arrange(phenotype, id)
+
+remove<-subset(antigen.ref, antigen.ref[,"Pr.response"]<.7 & antigen.ref[,"fdr"]>0.03)$id
+print(length(remove)) #see how many sample ids will be excluded
+numeric.only <- function(X,...){
+      returnCols <- names(X)
+      a<-sapply(X, is.numeric)
+      print(returnCols[a == "TRUE"])
+}
+
+phenotype[(phenotype$id %in% remove), numeric.only(phenotype)]<-NA
+TNF<-phenotype
+
 phenotype<-CD4_OG_TNF
@@
 phenotype[(phenotype$id %in% remove), numeric.only(phenotype)]<-NA
-TNF<-phenotype
+IFNg<-phenotype
 ```
@@
 ```
+
+##PEP
+```{r PEP IFNg}
+library(dplyr)
+library(knitr)
+DF<-subset(IFNg, IFNg$Stim=="PEP")
+#split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
+x<-split(DF, list(DF$TB, DF$SM))
+#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0.
+It's a weird double negative.
+But basically makes a list of the columns that have sufficient data.
+y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
+#each entry in the list has to be extracted and named
+a<-unlist(y[1])
+b<-unlist(y[2])
+c<-unlist(y[3])
+d<-unlist(y[4])
+#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
+test<-Reduce(intersect, list(a,b,c,d))
+#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
+colnames<-c(test[1:31], test[38:49])
+for(col in colnames){
+    y<-pvals_4_loose(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
+}
+```

### 2019-07-29 OG_cleaning.Rmd
added in a IFNg subset

@@
 write.csv(OG_CD4_OG.TNFa_summary.csv, "OG_CD4_OG.TNFa_summary_clean.csv")
+write.csv(OG_CD4_OG.IFNg_summary.csv, "OG_CD4_OG.IFNg_summary_clean.csv")
 

### 2019-07-29 OG-Analysis.Rmd
write files with low responders removed

@@
 OG<-phenotype
+write.csv(OG, "/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_CD4_OG_summary_clean_NAs.csv")
 
@@
 GT<-phenotype
+write.csv(GT, "/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_CD4_OG.GT_summary_clean_NAs.csv")
 
@@
 TH1<-phenotype
+write.csv(TH1, "/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_CD4_OG.TH1_summary_clean_NAs.csv")
 
@@
 phenotype[(phenotype$id %in% remove), numeric.only(phenotype)]<-NA
-TNF<-phenotype
+IFNg<-phenotype
+write.csv(IFNg, "/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_CD4_OG.IFNg_summary_clean_NAs.csv")
 
+#OG TNFa
 phenotype<-CD4_OG_TNF
@@
 phenotype[(phenotype$id %in% remove), numeric.only(phenotype)]<-NA
-IFNg<-phenotype
+TNF<-phenotype
+write.csv(TNF, "/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_CD4_OG.TNFa_summary_clean_NAs.csv")
 ```
@@
 
-##PEP
+##PEP TH1
 ```{r PEP th1}
@@
 
-##PEP
+##PEP TNF
 ```{r PEP tnfa}
@@
 
-##PEP
+##PEP GT
 ```{r PEP gt}
@@
 
-##PEP
+##PEP IFNg
 ```{r PEP IFNg}

### 2019-07-31 OG-Analysis_GD.Rmd
Start of OG gamma delta analysis

---
title: "OG-Analysis"
author: "Taryn McLaughlin"
date: "5/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment=NA)
library(dplyr)
library(knitr)
library(ggplot2)
```

```{r}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")
```

```{r read in data files}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
gd<-dplyr::select(read.csv("OG_gd_bulk_summary_clean.csv"), -X, -X.1)
gd_OG<-dplyr::select(read.csv("OG_gd_OG_summary_clean.csv"), -X, -X.1)
GD<-dplyr::select(read.csv("OG_GDonly_bulk_summary_clean.csv"), -X, -X.1)
GD_OG<-dplyr::select(read.csv("OG_GDonly_OG_summary_clean.csv"), -X, -X.1)
CD3<-dplyr::select(read.csv("OG_CD3_bulk_summary_clean.csv"), -X, -X.1)
PBMC<-dplyr::select(read.csv("DonorInfo_clean.csv"), -X)
    PBMC<-PBMC[grep("OG", PBMC$used),]
mimosa<-dplyr::select(read.csv("OG_MIMOSA_clean.csv"), -X)
    mimosa<-dplyr::filter(mimosa, celltype=="GD" & Exprate==0.01)
```

##Frequency of Gamma Deltas to start with
```{r}
pvals_gd(subset(CD3, CD3$Stim=="UN"), "GD_freq")
plot_with_all_stats(subset(CD3, CD3$Stim=="UN"), "GD_freq")
plot_3(subset(CD3, CD3$Stim=="UN"), "GD_freq")
pvals_gd(subset(CD3, CD3$Stim=="SWAP"), "GD_freq")
```

##Proliferation of Gamma Deltas
###Including cd4/cd8
```{r gd}
DF<-subset(gd, gd$Stim=="SWAP")
plot_with_all_stats(DF, "OG_freq")
plot_3(DF,  "OG_freq")
ggplot(DF, aes(x=SM, y=OG_freq, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
colnames<-numeric_only(DF)
for(col in colnames){
    y<-pvals_collapse(DF, col)
    if(dim(y)[1]>0){
        print(col)
        print(y)}
}
```

###Excluding CD4/CD8
```{r GD}
DF<-subset(GD, GD$Stim=="SWAP")
plot_with_all_stats(DF, "OG_freq")
plot_3(DF,  "OG_freq")
ggplot(DF, aes(x=SM, y=OG_freq, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
colnames<-numeric_only(DF)
for(col in colnames){
    y<-pvals_collapse(DF, col)
    if(dim(y)[1]>0){
        print(col)
        print(y)}
}
```

##Are starting Frequencies related to expansion
```{r}
start<-select(CD3, Sample, Stim, Donor, GD_freq)
middle1<-filter(start, Stim=="UN")
middle2<-filter(start, Stim=="SWAP")
end<-select(gd, Sample, Stim, TB, SM, Donor, OG_freq)%>%filter(Stim=="SWAP")

test<-merge(middle1, middle2, by=c("Donor"))
test2<-merge(test, end, by="Donor")
melt<-melt(test2, id.vars=c("Donor", "TB", "SM"), measure.vars=c("GD_freq.x", "GD_freq.y"))
ggplot(melt, aes(x=variable, y=value, color=TB))+geom_point()+geom_line(aes(group=Donor))+facet_grid(~TB)+scale_x_discrete(labels=c("Starting freq", "Swap Freq", "Proliferating Freq"))

test2$ratio<-test2$GD_freq.y/test2$GD_freq.x
plot_with_all_stats(test2, "ratio")+labs(title="Expansion of gd")
plot_3(test2, "ratio")+labs(title="Expansion of gd")

test2$dif<-test2$GD_freq.y-test2$GD_freq.x
plot_with_all_stats(test2, "dif")+labs(title="Expansion of gd")
plot_3(test2, "dif")+labs(title="Expansion of gd")
```

#Phenotyping
##Removing Low Responders
```{r}
antigen.ref<-mimosa
antigen.ref$id<-gsub("_Treatment", "", antigen.ref$Sample)
antigen.ref<-arrange(antigen.ref, id)

#Total gd
phenotype<-gd_OG
phenotype$id<-paste(phenotype$Donor, phenotype$Stim, sep="_")
phenotype<-arrange(phenotype, id)

remove<-subset(antigen.ref, antigen.ref[,"Pr.response"]<.7 & antigen.ref[,"fdr"]>0.03)$id
print(length(remove)) #see how many sample ids will be excluded
numeric.only <- function(X,...){
      returnCols <- names(X)
      a<-sapply(X, is.numeric)
      print(returnCols[a == "TRUE"])
}

phenotype[(phenotype$id %in% remove), numeric.only(phenotype)]<-NA
OG_gd<-phenotype
write.csv(OG_gd, "/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_gd_OG_summary_clean_NAs.csv")

antigen.ref<-mimosa
antigen.ref$id<-gsub("_Treatment", "", antigen.ref$Sample)
antigen.ref<-arrange(antigen.ref, id)

#gd only
phenotype<-gd_OG
phenotype$id<-paste(phenotype$Donor, phenotype$Stim, sep="_")
phenotype<-arrange(phenotype, id)

remove<-subset(antigen.ref, antigen.ref[,"Pr.response"]<.7 & antigen.ref[,"fdr"]>0.03)$id
print(length(remove)) #see how many sample ids will be excluded
numeric.only <- function(X,...){
      returnCols <- names(X)
      a<-sapply(X, is.numeric)
      print(returnCols[a == "TRUE"])
}

phenotype[(phenotype$id %in% remove), numeric.only(phenotype)]<-NA
OG_GD<-phenotype
write.csv(OG_GD, "/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_GDonly_OG_summary_clean_NAs.csv")
```

##gd
```{r gd og SWAP}
library(dplyr)
library(knitr)
DF<-subset(OG_gd, OG_gd$Stim=="SWAP")
#split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
x<-split(DF, list(DF$TB, DF$SM))
#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0. It's a weird double negative. But basically makes a list of the columns that have sufficient data.
y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
#each entry in the list has to be extracted and named
a<-unlist(y[1])
b<-unlist(y[2])
c<-unlist(y[3])
d<-unlist(y[4])
#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
test<-Reduce(intersect, list(a,b,c,d))
#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
colnames<-c(test[1:39], test[45:56])
for(col in colnames){
    y<-pvals_collapse(DF, col)
    if(dim(y)[1]>0){
        print(col)
        print(y)}
}
```

```{r}
test<-filter(DF, !is.na(OG_freq))
kable(xtabs(~TB+SM, test))
```

```{r}
vars<-names(test)[grep("._._.", names(test))]
vars<-vars[-grep("._._._.", vars)]
test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
test2<-filter(test2, variable!="x_x_x")
ggplot(test2, aes(x=variable, y=value, fill=TB))+
    geom_boxplot()+

    scale_fill_manual(values =c("#1a9850", "#2166ac", "#b2182b"))+
    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(title="Cytokine Production by Proliferating CD4+ T cells",
         y="Frequency of Cytokine+ CD4+ T cells")

plot_with_all_stats(DF, "IL4_freq")
ggplot(DF, aes(x=SM, y=IL4_freq, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
```

```{r, fig.width=16, fig.height=8}
vars<-names(test)[grep("._._._.", names(test))]
test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
test2<-filter(test2, variable!="x_x_x_x")
ggplot(test2, aes(x=variable, y=value, fill=TB))+
    geom_boxplot()+

    scale_fill_manual(values =c("#1a9850", "#2166ac", "#b2182b"))+
    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(title="Lineage Marker Expression on Proliferating GD T cells",
         y="Frequency of Marker+ GD T cells")

```

##GD
```{r GD og swap}
library(dplyr)
library(knitr)
DF<-subset(OG_GD, OG_GD$Stim=="SWAP")
#split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
x<-split(DF, list(DF$TB, DF$SM))
#for each dataframe in the list, return the column names of the columns where the sum of the number of rows that do not have missing data is more than 0. It's a weird double negative. But basically makes a list of the columns that have sufficient data.
y<-lapply(x, function(g) names(g[, colSums(!is.na(g)) > 0]))
#each entry in the list has to be extracted and named
a<-unlist(y[1])
b<-unlist(y[2])
c<-unlist(y[3])
d<-unlist(y[4])
#then find the intersection of all the lists. Aka which column names have sufficient data in all 4 TB-SM groups
test<-Reduce(intersect, list(a,b,c,d))
#this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
colnames<-c(test[1:39], test[45:56])
for(col in colnames){
    y<-pvals_collapse(DF, col)
    if(dim(y)[1]>0){
        print(col)
        print(y)}
}
```

```{r}
test<-filter(DF, !is.na(OG_freq))
kable(xtabs(~TB+SM, test))
```

```{r}
vars<-names(test)[grep("._._.", names(test))]
vars<-vars[-grep("._._._.", vars)]
test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
test2<-filter(test2, variable!="x_x_x")
ggplot(test2, aes(x=variable, y=value, fill=TB))+
    geom_boxplot()+

    scale_fill_manual(values =c("#1a9850", "#2166ac", "#b2182b"))+
    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(title="Cytokine Production by SWAP-specific CD4+ T cells",
         y="Frequency of Cytokine+ SWAP-specific T cells")
```

```{r, fig.width=16, fig.height=8}
vars<-names(test)[grep("._._._.", names(test))]
test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
test2<-filter(test2, variable!="x_x_x_x")
ggplot(test2, aes(x=variable, y=value, fill=TB))+
    geom_boxplot()+
    scale_fill_manual(values =c("#1a9850", "#2166ac", "#b2182b"))+
    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(title="Lineage Marker Expression on SWAP-specific CD4+ T cells",
         y="Frequency of Marker+ SWAP-specific T cells")
```

### 2019-07-31 ICS-processing.R
had to re-export someone

@@ -1,6 +1,6 @@
 library(dplyr)
 
-setwd("/Users/tarynam/Desktop/ICS-CD4")
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/ICS-CD4")
 filenames = list.files(pattern="csv")
 for(i in 1:length(filenames)){
     print(filenames[i])
@@ -9,6 +9,6 @@ for(i in 1:length(filenames)){
 
 l.df<-lapply(ls(pattern=".csv"), function(x) get(x))
 combined<-dplyr::bind_rows(l.df)
-write.csv(combined, "/Users/tarynam/Desktop/ICS-CD4/clean-data/Master-ICS.csv")
+write.csv(combined, "/Applications/Old Computer/Day Lab/Flow-Data/ICS-CD4/Master-ICS.csv")
 
 

### 2019-08-12 ICS_Phenotyping.Rmd
added in PMA phenotype graph

@@
 }
+```
+
+```{r}
+test<-mutate_TB(filter(DF, !is.na(TH1_GATA3)))
+test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = grep("TH1_._._._.", names(DF)))
+test2<-filter(test2, variable!="TH1_x_x_x_x")
+test2$variable<-gsub("TH1_", "", test2$variable)
+ggplot(test2, aes(x=variable, y=value, fill=TB, alpha=SM))+
+    geom_boxplot()+
+    scale_alpha_manual(values=c(0.5,1))+
+    scale_fill_manual(values =c("#2166ac", "#b2182b"))+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
+    theme(legend.position="none")+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    labs(title="Lineage Marker Expression on Mtb-Specific TH1 CD4+ T cells",
+         y="Frequency of Marker+ Mtb-Specific T cells")+
+    coord_cartesian(ylim=c(0,10))
+kable(xtabs(~TB+SM, test))
 ```

### 2019-08-12 OG-Analysis_GD.Rmd
includes starting frequencies and ratios for expansion

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment=NA)
+knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment=NA, warning=FALSE)
 library(dplyr)
@@
 library(ggplot2)
+library(ggpmisc)
 ```
@@
 CD3<-dplyr::select(read.csv("OG_CD3_bulk_summary_clean.csv"), -X, -X.1)
+CD3_OG<-dplyr::select(read.csv("OG_CD3_OG_summary_clean.csv"), -X, -X.1)
 PBMC<-dplyr::select(read.csv("DonorInfo_clean.csv"), -X)
@@
 
+phenotype$include<-1
+for(i in 1:length(phenotype$include)){
+    if(is.na(phenotype$OG_freq[i])){
+        phenotype$include[i]<-0
+    }
+}
+kable(xtabs(~include+SM+TB, phenotype[phenotype$Stim=="SWAP",]))
+x<-split(phenotype[phenotype$Stim=="SWAP",], list(phenotype$SM, phenotype$TB))
+lapply(x, function(g) table(g$include))
+
+#gd only
 antigen.ref<-mimosa
@@
 
-#gd only
-phenotype<-gd_OG
+phenotype<-GD_OG
 phenotype$id<-paste(phenotype$Donor, phenotype$Stim, sep="_")
@@
 write.csv(OG_GD, "/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_GDonly_OG_summary_clean_NAs.csv")
+
+phenotype$include<-1
+for(i in 1:length(phenotype$include)){
+    if(is.na(phenotype$OG_freq[i])){
+        phenotype$include[i]<-0
+    }
+}
+kable(xtabs(~include+SM+TB, phenotype[phenotype$Stim=="SWAP",]))
+x<-split(phenotype[phenotype$Stim=="SWAP",], list(phenotype$SM, phenotype$TB))
+kable(sapply(x, function(g) table(g$include)))
 ```
@@
 test2<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
-test2<-filter(test2, variable!="x_x_x")
 ggplot(test2, aes(x=variable, y=value, fill=TB))+
@@
 
-    scale_fill_manual(values =c("#1a9850", "#2166ac", "#b2182b"))+
+scale_fill_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))+
     stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
-    theme(legend.position="none")+
+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    labs(title="Cytokine Production by Proliferating CD4+ T cells",
+         y="Frequency of Cytokine+ CD4+ T cells")
+
+ggplot(test2, aes(x=variable, y=value, fill=SM))+
+    geom_boxplot()+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
+
     theme(plot.title = element_text(hjust = 0.5)) +
@@
 plot_with_all_stats(DF, "IL4_freq")
+
+plot_3(DF, "IL4_freq")
+plot_3(DF, "TNFa_freq")
+plot_3(DF, "IFNg_freq")
+
 ggplot(DF, aes(x=SM, y=IL4_freq, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
+
+ggplot(DF, aes(x=SM, y=TNFa_freq, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
+
+ggplot(DF, aes(x=SM, y=IFNg_freq, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
 ```
@@
 
-    scale_fill_manual(values =c("#1a9850", "#2166ac", "#b2182b"))+
+scale_fill_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))+
     stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
-    theme(legend.position="none")+
+
     theme(plot.title = element_text(hjust = 0.5)) +
@@
          y="Frequency of Marker+ GD T cells")
+
+```
+
+```{r}
+plot_3(PBMC[PBMC$SM=="SM",], "SchistosomaIntensity")
+
+formula<-y~x
+DF<-merge(select(PBMC, Donor, SchistosomaIntensity), CD3[CD3$Stim=="SWAP",], by="Donor")
+ggplot(DF[DF$SM=="SM",], aes(x=OG_freq, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="CD3")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
+
+DF<-merge(select(PBMC, Donor, SchistosomaIntensity), CD3_OG[CD3_OG$Stim=="SWAP",], by="Donor")
+ggplot(DF[DF$SM=="SM",], aes(x=GD_freq, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="CD3")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
+
+DF<-merge(select(PBMC, Donor, SchistosomaIntensity), GD[GD$Stim=="SWAP",], by="Donor")
+ggplot(DF[DF$SM=="SM",], aes(x=OG_freq, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="GD")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
+
+DF<-merge(select(PBMC, Donor, SchistosomaIntensity), GD_OG[GD_OG$Stim=="SWAP",], by="Donor")
+ggplot(DF[DF$SM=="SM",], aes(x=IL4_freq, y=SchistosomaIntensity, color=TB))+
+    geom_point()+
+    scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))+
+    stat_smooth(method="lm", formula = formula)+
+    stat_poly_eq(aes(label = paste(..rr.label..)), 
+               label.x.npc = "right", label.y.npc = 0.15,
+               formula = formula, parse = TRUE, size = 3)
+
+DF<-merge(select(PBMC, Donor, SchistosomaIntensity), GD_OG[GD_OG$Stim=="SWAP",], by="Donor")
+ggplot(DF[DF$SM=="SM",], aes(x=TNFa_freq, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="GD")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
+
+DF<-merge(select(PBMC, Donor, SchistosomaIntensity), GD_OG[GD_OG$Stim=="SWAP",], by="Donor")
+ggplot(DF[DF$SM=="SM",], aes(x=IFNg_freq, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="GD")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
 
@@
 #this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
-colnames<-c(test[1:39], test[45:56])
+colnames<-c(test[1:32], test[38:49])
 for(col in colnames){
@@
     geom_boxplot()+
-    scale_fill_manual(values =c("#1a9850", "#2166ac", "#b2182b"))+
+scale_fill_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))+
     stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
-    theme(legend.position="none")+
+
     theme(plot.title = element_text(hjust = 0.5)) +

### 2019-08-12 PBMC.Rmd
separated out pre and post og viabilities

@@
 grid.arrange(
-    plot_pbmc(pbmc, "PreOG_Viability")+labs(y="Viability"),
-    plot_pbmc(pbmc, "PreOG_Cells.Vial")+labs(y="Cells/Vial"),
+    plot_pbmc(pbmc, "PostOG_Viability")+labs(y="Viability"),
+    plot_pbmc(pbmc, "PostOG_Cells.Vial")+labs(y="Cells/Vial"),
     top=text_grob("Post OG Staining", size=20), ncol=2

### 2019-08-12 PaperFigures.Rmd
from EVC script. with TH cats and boolean phenotype

---
title: "Paper Figures"
author: "Taryn McLaughlin"
date: "7/29/2019"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment=NA, warning=FALSE, echo=FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
library(tidyr)
library(reshape2)
library(cowplot)
library(corrplot)
library(sjPlot)
```

```{r}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")

baseplot<-function(data, xval, yval, colors){
    ggplot(data, aes(data[,xval], data[,yval], fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    geom_point(shape=16,size=2, position = position_jitterdodge(dodge.width=1))+
    scale_fill_manual(values = colors)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    stat_compare_means(aes(label = sprintf("p = %5.3f", as.numeric(..p.format..))),
                       label.y= (max(data[,yval], na.rm = TRUE)*1.1), label.x=1.3,
                       size=6)+
    labs(x="")
}

booleanplot<-function(data, xval, yval, colors){
    ggplot(data, aes(data[,xval], data[,yval], fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    geom_point(shape=16,size=2, position = position_jitterdodge(dodge.width=1))+
    scale_fill_manual(values = colors)+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(text = element_text(size=20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
    labs(x="")
}

healthy.cols<- c("SM+" = "#1a9850", "SM-" ="#91cf60")
ltbi.cols<- c("SM+" = "#2166ac", "SM-" = "#85ABD1")
tb.cols<- c("SM+" = "#b2182b", "SM-"="#E59EA6")
```

```{r}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
CD4<-read.csv("ICS_CD4_clean.csv")
CYT<-read.csv("ICS_CD4_Cytokine_clean.csv")
PHEN<-read.csv("ICS_CD4_phenotyping_clean_NAs.csv")
OG<-read.csv("OG_CD4_bulk_summary_clean.csv")
OGPHEN<-read.csv("OG_CD4_OG_summary_clean_NAs.csv")
OG.IFNg<-read.csv("OG_CD4_OG.IFNg_summary_clean_NAs.csv")
OG.GT<-read.csv("OG_CD4_OG.GT_summary_clean_NAs.csv")
luminex<-read.csv("/Applications/Old Computer/Day Lab/Luminex/Clean Data/Luminex_combined_background.txt")
```

#Healthy Adults
####The frequency of T-bet+ CD4+ T cells is lower amongst SM+ Healthy Adults  
```{r, fig.height=6, fig.width=3}
cd4<-filter_hc(subset(CD4, CD4$Stim=="UN"))
cyt<-filter_hc(subset(CYT, CYT$Stim=="PMA"))
phen<-filter_hc(subset(PHEN, PHEN$Stim=="PMA"))
og<-filter_hc(subset(OG, OG$Stim=="SEB"))
ogphen<-filter_hc(subset(OGPHEN, OGPHEN$Stim=="SEB"))
baseplot(cd4, "SM", "T-bet", healthy.cols)+labs(y="Frequency T-bet+ CD4+ T cells (%)")
```

####Generic cytokine production from CD4+ T cells stimulated with PMA & Ionomycin does not differ between SM- and SM+ Healthy Adults
```{r, fig.height=6, fig.width=10}
melt<-melt(cyt, id.vars=c("TB","SM"), measure.vars=c("Cyt.Freq", "TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
cyt_groups <- as_labeller(c("Cyt.Freq" = "Total", "TH1.Freq" = "TH1", 
                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))
baseplot(melt, "SM", "value", healthy.cols)+
    facet_wrap(~variable, ncol=4, labeller=cyt_groups)+
    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
```

####Generic TH1 and TH2 cytokine producing cells are stratified by lineage-specific transcription factors, and CCR4
Amongst TH1 cytokine producing CD4+ T cells, moderate frequencies express the transcription factor T-bet while very low frequencies express GATA3. In contrast, amongst TH2 cytokine producing CD4+ T cells, low frequencies express T-bet while high frequencies express GATA3. Similarly high frequencies of TH2 cytokine producing cells express CCR4 compared to TH1 cytokine producing cells. Both TH1 and TH2 cells have equivalent moderate frequencies of CXCR3 expressing cells. There are no differences by S. mansoni status.

*Which of these orientations makes the point better?*

```{r, fig.height=10, fig.width=8}
test<-melt(phen, id.vars=c("TB", "SM"), 
           measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4",
           "TH2_T-bet", "TH2_GATA3", "TH2_CXCR3", "TH2_CCR4"))

test$subset<-tstrsplit(test$variable, "_")[[1]]
test$subset<-paste(test$subset, "Cytokine+ CD4+ T Cells", sep=" ")
test$variable<-gsub("TH1_","", test$variable)
test$variable<-gsub("TH2_","", test$variable)
test$variable<-factor(test$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

booleanplot(test, "SM","value", healthy.cols)+
    facet_grid(variable~subset, scales="free")+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(y="Frequency of Marker+ CD4+ T cells (%)")
```

```{r, fig.height=8, fig.width=10}
booleanplot(test, "SM","value", healthy.cols)+
    facet_grid(subset~variable, scales="free")+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(y="Frequency of Marker+ CD4+ T cells (%)")
```

#QFT+

####QFT+ individuals coinfected with S. mansoni have higher Mtb peptide-specific total cytokine production, specifically due to higher co-production of IFNγ and TNFa
```{r, fig.height=6, fig.width=10}
cyt<-filter_ltbi(subset(CYT, CYT$Stim=="PEP"))
phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
og<-filter_ltbi(subset(OG, OG$Stim=="PEP"))
ogphen<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
oggt<-filter_ltbi(subset(OG.GT, OG.GT$Stim=="PEP"))

#generate cytokine table for graphing
melt<-melt(cyt, 
    id.vars=c("TB","SM"), 
    measure.vars=c("Cyt.Freq", "TH1.Freq", "TH1.2.Freq", "TH2.Freq", "G_x_x_T"))
cyt_groups <- as_labeller(c("Cyt.Freq" = "Total", "TH1.Freq" = "TH1", 
                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2", "G_x_x_T"="IFNγ\n&\nTNFa"))
#graph with facets
a<-baseplot(melt[melt$variable!="G_x_x_T",], "SM", "value", ltbi.cols)+
    facet_wrap(~variable, ncol=4, labeller=cyt_groups)+
    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
b<-baseplot(melt[melt$variable=="G_x_x_T",], "SM", "value", ltbi.cols)+
    facet_wrap(~variable, ncol=4, labeller=cyt_groups)+
    labs(y="")
grid.arrange(a,b, ncol=2, widths=c(3,1))
```

####Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers, however this does not differ by S. mansoni status
*which of these formats do you like? It may be better to look at the active TB data first*
```{r, fig.height=6, fig.width=10}
#graph phenotype
melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
test<-filter(melt, variable!="TH1_x_x_x_x")
test$variable<-gsub("TH1_", "", test$variable)
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" ,       
"t_x_3_x" ,         
"t_x_x_4" ,         
"t_x_x_x" ,
"t_g_3_4" ,         
"t_g_3_x" ,         
"t_g_x_4" ,         
"t_g_x_x" ,
"x_g_3_4" ,         
"x_g_3_x" ,         
"x_g_x_4" ,         
"x_g_x_x" ,
"x_x_3_4" ,          
"x_x_3_x" ,         
"x_x_x_4"           
))

T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
breaks <- levels(test$variable)

booleanplot(test, "variable", "value", ltbi.cols)+
    scale_x_discrete(labels = labels)+
    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free", ncol=4)+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")

```

####There are no differences in Mtb-specific proliferation betweeen SM+ and SM- QF+ individuals. The cytokine production from expanded Mtb-specific CD4+ T cells only differs with regards to IFNγ IL-4 co-production, however the frequency of these cells is incredibly low. The dominant cytokine subgroup is still cells that co-produce IFNγ and TNFa. 
```{r, fig.width=10, fig.height=6}
a<-baseplot(og, "SM", "OG_freq", ltbi.cols)+
    labs(y="Proliferating CD4+ T cells (%)")+coord_cartesian(ylim=c(0,5))
melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._.", names(ogphen)))
test<-melt[-grep("._._._.", melt$variable),]
test<-filter(test, variable!="x_x_x")
test$variable<-factor(test$variable, levels=c(
"T_G_x" ,       
"T_x_x" ,         
"x_G_x" ,         
"T_G_4" ,
"T_x_4" ,         
"x_G_4" ,         
"x_x_4"          
))

T <- c("+", "+", "-", "+", "+", "-", "-")
G <- c("+", "-", "+", "+", "-", "+", "-")
il4 <- c("-", "-", "-", "+", "+", "+", "+")

labels <- paste0(T, "\n", G, "\n", il4)
breaks <- levels(test$variable)

b<-booleanplot(test, "variable", "value", ltbi.cols)+
    scale_x_discrete(labels = labels)+
    labs(y="Cytokine+ Proliferating\nCD4+ T cells (%)")

cowplot::plot_grid(a, b, ncol=2, rel_widths=c(1.5,3.5), align="h")
```

####Furthermore, IFNγ and TNFa are also the two dominant cytokines produced when analyzed by luminex. The concentration of all cytokines measured by luminex are equivalent between SM- and SM+ groups.
```{r, fig.height=10, fig.width=8}
lum<-filter_ltbi(subset(luminex, luminex$Stim=="PEP"))
names(lum)<-gsub("IL", "IL-", names(lum))
names(lum)<-gsub("IFNg", "IFNγ", names(lum))
names(lum)<-gsub("TNFa", "TNFα", names(lum))

melt<-melt(lum, id.vars=c("TB", "SM"), measure.vars=grep("pg", names(lum)))
melt$variable<-gsub("_pg.ml", "", melt$variable)
melt$variable<- factor(melt$variable, levels=c("TNFα", "IFNγ", "IL-21", "IL-4", "IL-5", "IL-13", "IL-17A", "IL-22", "IL-10"))
booleanplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+labs(y="pg/mL")
```

####However the correlation between cytokines differs between groups. Of note, the correlation between IFNγ and TNFa is stronger amongst SM+ individuals than SM- individuals.
```{r}
df<-lum[,3:11]
names(lum)<-gsub("_pg.ml", "", names(lum))
pca1 <- prcomp(df, scale = TRUE)
pcas<-data.frame(pca1$x)
rotations<-data.frame(pca1$rotation)
rotations$cytokine<-row.names(rotations)
melt<-melt(rotations, id.vars="cytokine")

stims<-factor(lum$Stim)
worms<-factor(lum$SM)
a<-ggplot(pcas, aes(x=pcas[,1], y=pcas[,2]))+geom_point(aes(col=worms))+labs(y="PCA2", x="PCA1")

neg <- cor(lum[lum$SM=="SM-",3:11])
order <- corrMatOrder(neg, order = "hclust", hclust.method="centroid")
order.neg <- neg[order, order]
res1 <- cor.mtest(lum[lum$SM=="SM-",3:11], conf.level = .95)
pvals <- res1$p[order, order]
corrplot(order.neg, method = "circle", insig = "label_sig", tl.col = "black", 
         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = .9, pch.col = "white",
         type="upper")

pos <- cor(lum[lum$SM=="SM+",3:11])
order.pos <- pos[order, order]
res1 <- cor.mtest(lum[lum$SM=="SM+",3:11], conf.level = .95)
pvals <- res1$p[order, order]
corrplot(order.pos, method = "circle", insig = "label_sig", tl.col = "black", 
         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = .9, pch.col = "white",
         type="upper")

lum_x<-lum[lum$SM=="SM-",]
df_x<-as.matrix(lum_x[,3:11])
row.names(df_x)<-lum_x$Sample
hc_x <- hclust(as.dist(1-cor(df_x)))
x<-plot(hc_x)

lum_sm<-lum[lum$SM=="SM+",]
df_sm<-as.matrix(lum_sm[,3:11])
row.names(df_sm)<-lum_sm$Sample
hc_sm <- hclust(as.dist(1-cor(df_sm)))
sm<-plot(hc_sm)

df<-as.matrix(lum[,3:11])
#heatmap(t(df_x), Rowv=as.dendrogram(hc_x))
#heatmap(t(df_sm), Rowv=as.dendrogram(hc_sm))
#heatmap(t(df), Rowv=as.dendrogram(hc_sm))
```

####Similar to the overnight assay, expanded Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers, however this does not differ by S. mansoni status.  
*I maybe want to do some analysis about the fact that now higher frequencies of cells express T-bet and/or GATA3 and that there seems to be a preferential expansion of CCR4+ cells but I honestly don't know how to test that*
```{r, fig.height=6, fig.width=10}
#graph phenotype
melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._._.", names(ogphen)))
test<-filter(melt, variable!="x_x_x_x")
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" ,       
"t_x_3_x" ,         
"t_x_x_4" ,         
"t_x_x_x" ,
"t_g_3_4" ,         
"t_g_3_x" ,         
"t_g_x_4" ,         
"t_g_x_x" ,
"x_g_3_4" ,         
"x_g_3_x" ,         
"x_g_x_4" ,         
"x_g_x_x" ,
"x_x_3_4" ,          
"x_x_3_x" ,         
"x_x_x_4"           
))

T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
breaks <- levels(test$variable)

booleanplot(test, "variable", "value", ltbi.cols)+
    scale_x_discrete(labels = labels)+
    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
    labs(y="Frequency of Transcription Factor \n Mtb-Specific CD4+ T cells (%)")
```

#TB  
####Similar to QFT+, individuals with active TB, co-infected with S. mansoni maintain function Mtb-specific CD4+ T cell cytokine production. There are higher frequencies of TH1/2 cytokine co-producing CD4+ T cells amongst SM+ individuals, however the frequency of these cells is incredibly low. The dominant cytokine subgroup is still TH1 cytokine producing cells.
*Again this is where I need an aesthetics opinion*
```{r tb cytokine ics, fig.width=10, fig.height=6}
cyt<-filter_tb(subset(CYT, CYT$Stim=="PEP"))
phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
og<-filter_tb(subset(OG, OG$Stim=="PEP"))
ogphen<-filter_tb(subset(OGPHEN, OGPHEN$Stim=="PEP"))

#generate cytokine table for graphing
melt<-melt(cyt, 
    id.vars=c("TB","SM"), 
    measure.vars=c("Cyt.Freq", "TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
cyt_groups <- as_labeller(c("Cyt.Freq" = "Total", "TH1.Freq" = "TH1", 
                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))
#graph cytokine with facets
baseplot(melt, "SM", "value", tb.cols)+
    facet_wrap(~variable, ncol=4, labeller=cyt_groups)+
    labs(y="Frequency of Cytokine+ \nCD4+ T cells (%)")
```

####Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers. There are numerous phenotypic differences amongst Mtb-specific TH1 cytokine producing cells betweeen SM- and SM+ groups. While T-bet is still the dominant transcription factor expressed in both groups, most differences are due to higher frequencies of GATA3 and CCR4 co-expressing cells.
```{r, fig.width=10, fig.height=6}
#graph phenotype
melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
test<-filter(melt, variable!="TH1_x_x_x_x")
test$variable<-gsub("TH1_", "", test$variable)
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" ,       
"t_x_3_x" ,         
"t_x_x_4" ,         
"t_x_x_x" ,
"t_g_3_4" ,         
"t_g_3_x" ,         
"t_g_x_4" ,         
"t_g_x_x" ,
"x_g_3_4" ,         
"x_g_3_x" ,         
"x_g_x_4" ,         
"x_g_x_x" ,
"x_x_3_4" ,          
"x_x_3_x" ,         
"x_x_x_4"           
))

T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
breaks <- levels(test$variable)

booleanplot(test, "variable", "value", tb.cols)+
    scale_x_discrete(labels = labels)+
    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")

booleanplot(test[grep("g", test$variable),], "variable", "value", tb.cols)+
    scale_x_discrete(labels = labels[5:12])+
    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

booleanplot(test, "variable", "value", tb.cols)+facet_wrap(~tf, scales="free", ncol=4)+
    scale_x_discrete(labels = labels, breaks=breaks)+
    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")

```

### 2019-08-12 PaperFigures.Rmd
with plots for all 6 groups but TB got messed up

@@
 date: "7/29/2019"
-output: pdf_document
+output: html_document
 ---
@@
     geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    geom_point(shape=16,size=2, position = position_jitterdodge(dodge.width=1))+
     scale_fill_manual(values = colors)+
@@
     theme(plot.subtitle = element_text(hjust = 0.5)) +
-    stat_compare_means(aes(label = sprintf("p = %5.3f", as.numeric(..p.format..))),
-                       label.y= (max(data[,yval], na.rm = TRUE)*1.1), label.x=1.3,
-                       size=6)+
-    labs(x="")
+    stat_compare_means(
+        label="p.signif", hide.ns = TRUE, size=8,
+        label.x=1.3,
+        label.y = boxplot.stats(data[,yval])$stats[5]*1.4)+
+    coord_cartesian(ylim=c(0, boxplot.stats(data[,yval])$stats[5]*1.5))+
+    labs(x="", y="")
 }
@@
     stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
-    labs(x="")
+    labs(x="", y="")
 }
+
+allplot<-function(data, xval, yval, colors){
+    ggplot(data, aes(data[,xval], data[,yval], fill=disease))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
+    scale_fill_manual(values = colors)+
+    theme_classic()+ 
+    scale_x_discrete(labels=c("HC\nSM-", "HC\nSM+", "LTBI\nSM-", "LTBI\nSM+", "TB\nSM-", "TB\nSM+"))+
+    theme(legend.position="none")+
+    theme(text = element_text(size=20))+
+    theme(plot.subtitle = element_text(hjust = 0.5)) +
+    stat_compare_means(
+        comparisons = list(c("HC SM+", "HC SM-"),c("LTBI SM+", "LTBI SM-"), c("TB SM+", "TB SM-")),
+        label="p.signif", hide.ns = TRUE, size=8,
+        label.y = boxplot.stats(data[,yval])$stats[5]*1.4)+
+        coord_cartesian(ylim=c(0, boxplot.stats(data[,yval])$stats[5]*1.5))+
+        labs(x="", y="")
+}
+
+all_comps<-list(c("HC SM+", "HC SM-"),c("LTBI SM+", "LTBI SM-"), c("TB SM+", "TB SM-"))
 
@@
 tb.cols<- c("SM+" = "#b2182b", "SM-"="#E59EA6")
+all.cols<- c("HC SM+" = "#1a9850", "HC SM-" ="#91cf60",
+             "LTBI SM+" = "#2166ac", "LTBI SM-" = "#85ABD1",
+             "TB SM+" = "#b2182b", "TB SM-"="#E59EA6")
 ```
@@
 
+#All groups
+```{r, fig.height=6, fig.width=3}
+cd4<-plot_filter(subset(CD4, CD4$Stim=="UN"))
+cyt<-plot_filter(subset(CYT, CYT$Stim=="PMA"))
+phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))
+og<-plot_filter(subset(OG, OG$Stim=="SEB"))
+ogphen<-plot_filter(subset(OGPHEN, OGPHEN$Stim=="SEB"))
+```
+
+```{r, fig.width=16, fig.height=8}
+melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+
+a<-allplot(melt[melt$variable=="T-bet",], "disease", "value", all.cols)+    
+    facet_wrap(~variable)+
+    labs(y="(%) Total CD4+ T cells")
+
+b<-allplot(melt[melt$variable=="GATA3",], "disease", "value", all.cols)+    
+    facet_wrap(~variable)
+
+c<-allplot(melt[melt$variable=="CXCR3",], "disease", "value", all.cols)+    
+    facet_wrap(~variable)
+
+d<-allplot(melt[melt$variable=="CCR4",], "disease", "value", all.cols)+    
+    facet_wrap(~variable)
+
+cowplot::plot_grid(a, b, c, d,ncol=4)
+```
+
+```{r, fig.width=16, fig.height=8}
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
+cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
+                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))
+
+a<-allplot(melt[melt$variable=="IFNg",], "disease", "value", all.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)+
+    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
+
+b<-allplot(melt[melt$variable=="TNFa",], "disease", "value", all.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)
+
+c<-allplot(melt[melt$variable=="IL4",], "disease", "value", all.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)
+
+d<-allplot(melt[melt$variable=="IL13",], "disease", "value", all.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)
+
+cowplot::plot_grid(a, b, c, d,ncol=4)
+```
+
 #Healthy Adults
@@
 ogphen<-filter_hc(subset(OGPHEN, OGPHEN$Stim=="SEB"))
-baseplot(cd4, "SM", "T-bet", healthy.cols)+labs(y="Frequency T-bet+ CD4+ T cells (%)")
 ```
@@
 ```{r, fig.height=6, fig.width=10}
-melt<-melt(cyt, id.vars=c("TB","SM"), measure.vars=c("Cyt.Freq", "TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
-cyt_groups <- as_labeller(c("Cyt.Freq" = "Total", "TH1.Freq" = "TH1", 
+melt<-melt(cyt, id.vars=c("TB","SM"), measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
+cyt_groups <- as_labeller(c("TH1.Freq" = "TH1", 
                     "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))
@@
 
-```{r, fig.height=10, fig.width=8}
+```{r, fig.height=10, fig.width=5}
 test<-melt(phen, id.vars=c("TB", "SM"), 
@@
 test$subset<-tstrsplit(test$variable, "_")[[1]]
-test$subset<-paste(test$subset, "Cytokine+ CD4+ T Cells", sep=" ")
+test$subset<-paste(test$subset, "Cytokine+\nCD4+ T Cells", sep=" ")
 test$variable<-gsub("TH1_","", test$variable)
@@
 
-```{r, fig.height=8, fig.width=10}
-booleanplot(test, "SM","value", healthy.cols)+
-    facet_grid(subset~variable, scales="free")+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    labs(y="Frequency of Marker+ CD4+ T cells (%)")
-```
+#LTBI
 
-#QFT+
-
-####QFT+ individuals coinfected with S. mansoni have higher Mtb peptide-specific total cytokine production, specifically due to higher co-production of IFNγ and TNFa
+####LTBI individuals coinfected with S. mansoni have higher Mtb peptide-specific total cytokine production, specifically due to higher co-production of IFNγ and TNFa
 ```{r, fig.height=6, fig.width=10}
@@
     labs(y="")
-grid.arrange(a,b, ncol=2, widths=c(3,1))
+```
+
+```{r, fig.width=16, fig.height=8}
+melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+
+a<-baseplot(melt[melt$variable=="T-bet",], "disease", "value", ltbi.cols)+    
+    facet_wrap(~variable)+
+    labs(y="(%) Total CD4+ T cells")
+
+b<-baseplot(melt[melt$variable=="GATA3",], "disease", "value", ltbi.cols)+    
+    facet_wrap(~variable)
+
+c<-baseplot(melt[melt$variable=="CXCR3",], "disease", "value", ltbi.cols)+    
+    facet_wrap(~variable)
+
+d<-baseplot(melt[melt$variable=="CCR4",], "disease", "value", ltbi.cols)+    
+    facet_wrap(~variable)
+
+cowplot::plot_grid(a, b, c, d,ncol=4)
+```
+
+```{r, fig.width=16, fig.height=8}
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
+cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
+                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))
+
+a<-baseplot(melt[melt$variable=="IFNg",], "disease", "value", ltbi.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)+
+    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
+
+b<-baseplot(melt[melt$variable=="TNFa",], "disease", "value", ltbi.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)
+
+c<-baseplot(melt[melt$variable=="IL4",], "disease", "value", ltbi.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)
+
+d<-baseplot(melt[melt$variable=="IL13",], "disease", "value", ltbi.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)
+
+cowplot::plot_grid(a, b, c, d,ncol=4)
+```
+
+```{r, fig.height=6, fig.width=10}
+#generate cytokine table for graphing
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
+cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
+                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))
+#graph with facets
+baseplot(melt, "SM", "value", ltbi.cols)+
+    facet_wrap(~variable, ncol=4, labeller=cyt_groups)+
+    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
 ```
@@
 breaks <- levels(test$variable)
-
-booleanplot(test, "variable", "value", ltbi.cols)+
-    scale_x_discrete(labels = labels)+
-    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
-    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")
 
@@
 corrplot(order.neg, method = "circle", insig = "label_sig", tl.col = "black", 
-         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = .9, pch.col = "white",
+         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = 2, pch.col = "white",
          type="upper")
@@
 corrplot(order.pos, method = "circle", insig = "label_sig", tl.col = "black", 
-         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = .9, pch.col = "white",
+         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = 2, pch.col = "white",
          type="upper")
@@
 ####Similar to the overnight assay, expanded Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers, however this does not differ by S. mansoni status.
+
 *I maybe want to do some analysis about the fact that now higher frequencies of cells express T-bet and/or GATA3 and that there seems to be a preferential expansion of CCR4+ cells but I honestly don't know how to test that*
@@
 
-booleanplot(test, "variable", "value", ltbi.cols)+
-    scale_x_discrete(labels = labels)+
+test$tf<-NA
+test$tf[grep("x_g", test$variable)]<-"GATA3"
+test$tf[grep("t_x", test$variable)]<-"T-bet"
+test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
+test$tf[grep("x_x_._.", test$variable)]<-"None"
+test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
+
+booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free", ncol=4)+
+    scale_x_discrete(labels = labels, breaks=breaks)+
     stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
-    labs(y="Frequency of Transcription Factor \n Mtb-Specific CD4+ T cells (%)")
+    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")
 ```
@@
     id.vars=c("TB","SM"), 
-    measure.vars=c("Cyt.Freq", "TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
-cyt_groups <- as_labeller(c("Cyt.Freq" = "Total", "TH1.Freq" = "TH1", 
+    measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
+cyt_groups <- as_labeller(c("TH1.Freq" = "TH1", 
                     "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))
 #graph cytokine with facets
-baseplot(melt, "SM", "value", tb.cols)+
+keep <- baseplot(melt, "SM", "value", tb.cols)+
     facet_wrap(~variable, ncol=4, labeller=cyt_groups)+
     labs(y="Frequency of Cytokine+ \nCD4+ T cells (%)")
+```
+
+```{r, fig.width=16, fig.height=8}
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
+cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
+                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))
+
+a<-baseplot(melt[melt$variable=="IFNg",], "disease", "value", tb.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)+
+    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
+
+b<-baseplot(melt[melt$variable=="TNFa",], "disease", "value", tb.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)
+
+c<-baseplot(melt[melt$variable=="IL4",], "disease", "value", tb.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)
+
+d<-baseplot(melt[melt$variable=="IL13",], "disease", "value", tb.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)
+
+cowplot::plot_grid(a, b, c, d,ncol=4)
 ```
@@
 
-booleanplot(test, "variable", "value", tb.cols)+
-    scale_x_discrete(labels = labels)+
-    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")
-
-booleanplot(test[grep("g", test$variable),], "variable", "value", tb.cols)+
-    scale_x_discrete(labels = labels[5:12])+
-    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")
-
 test$tf<-NA

### 2019-08-13 ICS-processing.R
added in line to fix 2054 export

@@ -7,8 +7,9 @@ for(i in 1:length(filenames)){
     assign(filenames[i], read.table(filenames[i], header=TRUE, sep=";"))
 }
 
+NK2054_TB_SM.csv$donor<-"NK2054_TB_SM"
+
 l.df<-lapply(ls(pattern=".csv"), function(x) get(x))
 combined<-dplyr::bind_rows(l.df)
-write.csv(combined, "/Applications/Old Computer/Day Lab/Flow-Data/ICS-CD4/Master-ICS.csv")
-
+write.csv(combined, "/Applications/Old Computer/Day Lab/Flow-Data/Master-ICS.csv")
 

### 2019-08-13 ICS_Cleaning.Rmd
adding in IFNg and TNFa gates

--- 2019-06-10 ICS_Cleaning.Rmd
@@
 
+```{r create new combo variables from exported boolean phenotype}
+CD4<-mutate(CD4,
+      #These all follow the same format
+       #IFNg is the new variable. it gets assign the row sum of all the columns that have a G in the name
+       TbetCXCR3=  rowSums(CD4[,grep("t_._3_.", names(CD4))]),
+       TbetCCR4=  rowSums(CD4[,grep("t_._._4", names(CD4))]),
+       GATA3CXCR3=  rowSums(CD4[,grep("._g_3_.", names(CD4))]),
+       GATA3CCR4=  rowSums(CD4[,grep("._g_._4", names(CD4))]),
+       CXCR3.Tbetprop = TbetCXCR3/Tbet,
+       CCR4.Tbetprop = TbetCCR4/Tbet,
+       CXCR3.GATA3prop= GATA3CXCR3/GATA3,
+       CCR4.GATA3prop= GATA3CCR4/GATA3
+)
+
+CD4_phenotype<-mutate(CD4_phenotype,
+      #These all follow the same format
+      ifng_Tbet=  rowSums(CD4_phenotype[,grep("ifng_t_._._.", names(CD4_phenotype))]),
+      ifng_GATA3=  rowSums(CD4_phenotype[,grep("ifng_._g_._.", names(CD4_phenotype))]),
+      ifng_CCR4=  rowSums(CD4_phenotype[,grep("ifng_._._._4", names(CD4_phenotype))]),
+      ifng_CXCR3=  rowSums(CD4_phenotype[,grep("ifng_._._3_.", names(CD4_phenotype))]),
+      tnfa_Tbet=  rowSums(CD4_phenotype[,grep("tnfa_t_._._.", names(CD4_phenotype))]),
+      tnfa_GATA3=  rowSums(CD4_phenotype[,grep("tnfa_._g_._.", names(CD4_phenotype))]),
+      tnfa_CCR4=  rowSums(CD4_phenotype[,grep("tnfa_._._._4", names(CD4_phenotype))]),
+      tnfa_CXCR3=  rowSums(CD4_phenotype[,grep("tnfa_._._3_.", names(CD4_phenotype))]),
+
+      TH1_TbetCXCR3=  rowSums(CD4_phenotype[,grep("TH1_t_._3_.", names(CD4_phenotype))]),
+      TH1_TbetCCR4=  rowSums(CD4_phenotype[,grep("TH1_t_._._4", names(CD4_phenotype))]),
+      TH1_GATA3CXCR3=  rowSums(CD4_phenotype[,grep("TH1_._g_3_.", names(CD4_phenotype))]),
+      TH1_GATA3CCR4=  rowSums(CD4_phenotype[,grep("TH1_._g_._4", names(CD4_phenotype))]),
+      TH2_TbetCXCR3=  rowSums(CD4_phenotype[,grep("TH2_t_._3_.", names(CD4_phenotype))]),
+      TH2_TbetCCR4=  rowSums(CD4_phenotype[,grep("TH2_t_._._4", names(CD4_phenotype))]),
+      TH2_GATA3CXCR3=  rowSums(CD4_phenotype[,grep("TH2_._g_3_.", names(CD4_phenotype))]),
+      TH2_GATA3CCR4=  rowSums(CD4_phenotype[,grep("TH2_._g_._4", names(CD4_phenotype))]),
+      ifng_TbetCXCR3=  rowSums(CD4_phenotype[,grep("ifng_t_._3_.", names(CD4_phenotype))]),
+      ifng_TbetCCR4=  rowSums(CD4_phenotype[,grep("ifng_t_._._4", names(CD4_phenotype))]),
+      ifng_GATA3CXCR3=  rowSums(CD4_phenotype[,grep("ifng_._g_3_.", names(CD4_phenotype))]),
+      ifng_GATA3CCR4=  rowSums(CD4_phenotype[,grep("ifng_._g_._4", names(CD4_phenotype))]),
+      tnfa_TbetCXCR3=  rowSums(CD4_phenotype[,grep("tnfa_t_._3_.", names(CD4_phenotype))]),
+      tnfa_TbetCCR4=  rowSums(CD4_phenotype[,grep("tnfa_t_._._4", names(CD4_phenotype))]),
+      tnfa_GATA3CXCR3=  rowSums(CD4_phenotype[,grep("tnfa_._g_3_.", names(CD4_phenotype))]),
+      tnfa_GATA3CCR4=  rowSums(CD4_phenotype[,grep("tnfa_._g_._4", names(CD4_phenotype))]),
+
+       TH1_CXCR3.Tbetprop = TH1_TbetCXCR3/TH1_Tbet,
+       TH1_CCR4.Tbetprop = TH1_TbetCCR4/TH1_Tbet,
+       TH1_CXCR3.GATA3prop= TH1_GATA3CXCR3/TH1_GATA3,
+       TH1_CCR4.GATA3prop= TH1_GATA3CCR4/TH1_GATA3,
+
+       TH2_CXCR3.Tbetprop = TH2_TbetCXCR3/TH2_Tbet,
+       TH2_CCR4.Tbetprop = TH2_TbetCCR4/TH2_Tbet,
+       TH2_CXCR3.GATA3prop= TH2_GATA3CXCR3/TH2_GATA3,
+       TH2_CCR4.GATA3prop= TH2_GATA3CCR4/TH2_GATA3,
+
+       ifng_CXCR3.Tbetprop = ifng_TbetCXCR3/ifng_Tbet,
+       ifng_CCR4.Tbetprop = ifng_TbetCCR4/ifng_Tbet,
+       ifng_CXCR3.GATA3prop= ifng_GATA3CXCR3/ifng_GATA3,
+       ifng_CCR4.GATA3prop= ifng_GATA3CCR4/ifng_GATA3,
+
+       tnfa_CXCR3.Tbetprop = tnfa_TbetCXCR3/tnfa_Tbet,
+       tnfa_CCR4.Tbetprop = tnfa_TbetCCR4/tnfa_Tbet,
+       tnfa_CXCR3.GATA3prop= tnfa_GATA3CXCR3/tnfa_GATA3,
+       tnfa_CCR4.GATA3prop= tnfa_GATA3CCR4/tnfa_GATA3
+)
+
+```
+
 ```{r Cytokine MFI inclusion}

### 2019-08-13 ICS_Phenotyping.Rmd
adding in IFNg and TNFa analysis

@@
 #removed from the analysis
-antigen.ref<-dplyr::select(mimosa, id, Cyt_include, TH1_include, TH2_include, TH1.2_include)
+antigen.ref<-dplyr::select(mimosa, id, Cyt_include, TH1_include, TH2_include, TH1.2_include, ifng_include, tnfa_include)
 
@@
 DF<-phenotype #so we don't alter the original phenotype dataframe just in case
-keynames<-c("Cyt", "TH1", "TH1.2", "TH2") #list of character strings that contain our "key" values
+keynames<-c("Cyt", "TH1", "TH1.2", "TH2", "ifng", "tnfa") #list of character strings that contain our "key" values
 #in the antigen ref we have columns named "key"_include and in the phenotype data we have columns named

### 2019-08-20 PaperFigures.Rmd
all 6 groups for total cd4 and pma

@@
 
-```{r}
+```{r plot functions}
 source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
@@
         comparisons = list(c("HC SM+", "HC SM-"),c("LTBI SM+", "LTBI SM-"), c("TB SM+", "TB SM-")),
-        label="p.signif", hide.ns = TRUE, size=8,
+        label="p.signif", size=8,
         label.y = boxplot.stats(data[,yval])$stats[5]*1.4)+
@@
 
-```{r}
+```{r load data}
 setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
@@
 #All groups
-```{r, fig.height=6, fig.width=3}
+Dont worry I can change this so the ns bars don't show up, especially for the cytokines since I have no idea why that giant NS is there. But for now I'm leaving them so we know.
+```{r}
 cd4<-plot_filter(subset(CD4, CD4$Stim=="UN"))
@@
 #Healthy Adults
-####The frequency of T-bet+ CD4+ T cells is lower amongst SM+ Healthy Adults  
-```{r, fig.height=6, fig.width=3}
-cd4<-filter_hc(subset(CD4, CD4$Stim=="UN"))
-cyt<-filter_hc(subset(CYT, CYT$Stim=="PMA"))
-phen<-filter_hc(subset(PHEN, PHEN$Stim=="PMA"))
-og<-filter_hc(subset(OG, OG$Stim=="SEB"))
-ogphen<-filter_hc(subset(OGPHEN, OGPHEN$Stim=="SEB"))
-```
-
+*Is this superfluous now?*  
 ####Generic cytokine production from CD4+ T cells stimulated with PMA & Ionomycin does not differ between SM- and SM+ Healthy Adults
-```{r, fig.height=6, fig.width=10}
+```{r, fig.width=18, fig.height=9.5}
 melt<-melt(cyt, id.vars=c("TB","SM"), measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
@@
     facet_wrap(~variable, ncol=4, labeller=cyt_groups)+
-    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
+    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")+coord_cartesian(ylim=c(0,40))
 ```
@@
 
-*Which of these orientations makes the point better?*
-
-```{r, fig.height=10, fig.width=5}
+```{r, fig.width=5, fig.height=10}
 test<-melt(phen, id.vars=c("TB", "SM"), 
@@
 
-####LTBI individuals coinfected with S. mansoni have higher Mtb peptide-specific total cytokine production, specifically due to higher co-production of IFNγ and TNFa
-```{r, fig.height=6, fig.width=10}
+####LTBI individuals coinfected with S. mansoni have higher Mtb peptide-specific cytokine production, specifically due to higher production of IFNγ
+```{r, narrow down ltbi data}
 cyt<-filter_ltbi(subset(CYT, CYT$Stim=="PEP"))
@@
 oggt<-filter_ltbi(subset(OG.GT, OG.GT$Stim=="PEP"))
-
-#generate cytokine table for graphing
-melt<-melt(cyt, 
-    id.vars=c("TB","SM"), 
-    measure.vars=c("Cyt.Freq", "TH1.Freq", "TH1.2.Freq", "TH2.Freq", "G_x_x_T"))
-cyt_groups <- as_labeller(c("Cyt.Freq" = "Total", "TH1.Freq" = "TH1", 
-                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2", "G_x_x_T"="IFNγ\n&\nTNFa"))
-#graph with facets
-a<-baseplot(melt[melt$variable!="G_x_x_T",], "SM", "value", ltbi.cols)+
-    facet_wrap(~variable, ncol=4, labeller=cyt_groups)+
-    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
-b<-baseplot(melt[melt$variable=="G_x_x_T",], "SM", "value", ltbi.cols)+
-    facet_wrap(~variable, ncol=4, labeller=cyt_groups)+
-    labs(y="")
 ```
 
-```{r, fig.width=16, fig.height=8}
-melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+```{r cytokine, fig.width=16, fig.height=8}
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
+cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
+                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))
 
-a<-baseplot(melt[melt$variable=="T-bet",], "disease", "value", ltbi.cols)+    
+a<-baseplot(melt[melt$variable=="IFNg",], "SM", "value", ltbi.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)+
+    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
+
+b<-baseplot(melt[melt$variable=="TNFa",], "SM", "value", ltbi.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)
+
+c<-baseplot(melt[melt$variable=="IL4",], "SM", "value", ltbi.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)
+
+d<-baseplot(melt[melt$variable=="IL13",], "SM", "value", ltbi.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)
+
+cowplot::plot_grid(a, b, c, d,ncol=4)
+```
+
+####Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers, however this does not differ by S. mansoni status
+```{r ics main four markers, fig.width=16, fig.height=8}
+melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("ifng_T-bet", "ifng_GATA3", "ifng_CXCR3", "ifng_CCR4"))
+melt$variable<-gsub("ifng_", "", melt$variable)
+
+a<-baseplot(melt[melt$variable=="T-bet",], "SM", "value", ltbi.cols)+    
     facet_wrap(~variable)+
@@
 
-b<-baseplot(melt[melt$variable=="GATA3",], "disease", "value", ltbi.cols)+    
+b<-baseplot(melt[melt$variable=="GATA3",], "SM", "value", ltbi.cols)+    
+    facet_wrap(~variable)
+c<-baseplot(melt[melt$variable=="CXCR3",], "SM", "value", ltbi.cols)+    
     facet_wrap(~variable)
 
-c<-baseplot(melt[melt$variable=="CXCR3",], "disease", "value", ltbi.cols)+    
-    facet_wrap(~variable)
-
-d<-baseplot(melt[melt$variable=="CCR4",], "disease", "value", ltbi.cols)+    
+d<-baseplot(melt[melt$variable=="CCR4",], "SM", "value", ltbi.cols)+    
     facet_wrap(~variable)
@@
 
-```{r, fig.width=16, fig.height=8}
-melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
-cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
-                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))
-
-a<-baseplot(melt[melt$variable=="IFNg",], "disease", "value", ltbi.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)+
-    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-b<-baseplot(melt[melt$variable=="TNFa",], "disease", "value", ltbi.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)
-
-c<-baseplot(melt[melt$variable=="IL4",], "disease", "value", ltbi.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)
-
-d<-baseplot(melt[melt$variable=="IL13",], "disease", "value", ltbi.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)
-
-cowplot::plot_grid(a, b, c, d,ncol=4)
-```
-
-```{r, fig.height=6, fig.width=10}
-#generate cytokine table for graphing
-melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
-cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
-                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))
-#graph with facets
-baseplot(melt, "SM", "value", ltbi.cols)+
-    facet_wrap(~variable, ncol=4, labeller=cyt_groups)+
-    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
-```
-
-####Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers, however this does not differ by S. mansoni status
-*which of these formats do you like? It may be better to look at the active TB data first*
-```{r, fig.height=6, fig.width=10}
-#graph phenotype
-melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
-test<-filter(melt, variable!="TH1_x_x_x_x")
-test$variable<-gsub("TH1_", "", test$variable)
-test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
-
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
-
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
-
-test$tf<-NA
-test$tf[grep("x_g", test$variable)]<-"GATA3"
-test$tf[grep("t_x", test$variable)]<-"T-bet"
-test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
-test$tf[grep("x_x_._.", test$variable)]<-"None"
-test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
-
-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free", ncol=4)+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
-    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")
-
-```
-
 ####There are no differences in Mtb-specific proliferation betweeen SM+ and SM- QF+ individuals.
 The cytokine production from expanded Mtb-specific CD4+ T cells only differs with regards to IFNγ IL-4 co-production, however the frequency of these cells is incredibly low.
-The dominant cytokine subgroup is still cells that co-produce IFNγ and TNFa.
-```{r, fig.width=10, fig.height=6}
+The dominant cytokine subgroup is still cells that co-produce IFNγ and TNFα.
+```{r og and cyt, fig.width=16, fig.height=8}
 a<-baseplot(og, "SM", "OG_freq", ltbi.cols)+
@@
 ####Furthermore, IFNγ and TNFa are also the two dominant cytokines produced when analyzed by luminex. The concentration of all cytokines measured by luminex are equivalent between SM- and SM+ groups.
-```{r, fig.height=10, fig.width=8}
+```{r luminex, fig.width=12, fig.height=16}
 lum<-filter_ltbi(subset(luminex, luminex$Stim=="PEP"))
@@
 melt$variable<-gsub("_pg.ml", "", melt$variable)
-melt$variable<- factor(melt$variable, levels=c("TNFα", "IFNγ", "IL-21", "IL-4", "IL-5", "IL-13", "IL-17A", "IL-22", "IL-10"))
+melt$variable<- factor(melt$variable, levels=c("TNFα", "IFNγ", "IL-21", "IL-17A", "IL-22", "IL-10", "IL-4", "IL-5", "IL-13"))
 booleanplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+labs(y="pg/mL")
@@
 ####However the correlation between cytokines differs between groups. Of note, the correlation between IFNγ and TNFa is stronger amongst SM+ individuals than SM- individuals.
-```{r}
+```{r cor plots}
 df<-lum[,3:11]
@@
          type="upper")
-
-lum_x<-lum[lum$SM=="SM-",]
-df_x<-as.matrix(lum_x[,3:11])
-row.names(df_x)<-lum_x$Sample
-hc_x <- hclust(as.dist(1-cor(df_x)))
-x<-plot(hc_x)
-
-lum_sm<-lum[lum$SM=="SM+",]
-df_sm<-as.matrix(lum_sm[,3:11])
-row.names(df_sm)<-lum_sm$Sample
-hc_sm <- hclust(as.dist(1-cor(df_sm)))
-sm<-plot(hc_sm)
-
-df<-as.matrix(lum[,3:11])
-#heatmap(t(df_x), Rowv=as.dendrogram(hc_x))
-#heatmap(t(df_sm), Rowv=as.dendrogram(hc_sm))
-#heatmap(t(df), Rowv=as.dendrogram(hc_sm))
 ```
@@
 *I maybe want to do some analysis about the fact that now higher frequencies of cells express T-bet and/or GATA3 and that there seems to be a preferential expansion of CCR4+ cells but I honestly don't know how to test that*
-```{r, fig.height=6, fig.width=10}
+
+```{r og main four markers, fig.width=16, fig.height=8}
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+
+a<-baseplot(melt[melt$variable=="T-bet",], "disease", "value", ltbi.cols)+    
+    facet_wrap(~variable)+
+    labs(y="(%) Total CD4+ T cells")+coord_cartesian(ylim=c(0,100))
+
+b<-baseplot(melt[melt$variable=="GATA3",], "disease", "value", ltbi.cols)+    
+    facet_wrap(~variable)+coord_cartesian(ylim=c(0,100))
+
+c<-baseplot(melt[melt$variable=="CXCR3",], "disease", "value", ltbi.cols)+    
+    facet_wrap(~variable)+coord_cartesian(ylim=c(0,100))
+
+d<-baseplot(melt[melt$variable=="CCR4",], "disease", "value", ltbi.cols)+    
+    facet_wrap(~variable)+coord_cartesian(ylim=c(0,100))
+
+cowplot::plot_grid(a, b, c, d,ncol=4)
+```
+
+```{r og phenotype, fig.width=16, fig.height=8}
 #graph phenotype
@@
 *Again this is where I need an aesthetics opinion*
-```{r tb cytokine ics, fig.width=10, fig.height=6}
+```{r tb cytokine ics, fig.width=16, fig.height=8}
 cyt<-filter_tb(subset(CYT, CYT$Stim=="PEP"))
@@
 While T-bet is still the dominant transcription factor expressed in both groups, most differences are due to higher frequencies of GATA3 and CCR4 co-expressing cells.
-```{r, fig.width=10, fig.height=6}
+
+```{r main four markers, fig.width=16, fig.height=8}
+melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4"))
+melt$variable<-gsub("TH1_", "", melt$variable)
+
+a<-baseplot(melt[melt$variable=="T-bet",], "SM", "value", tb.cols)+    
+    facet_wrap(~variable)+
+    labs(y="(%) Total CD4+ T cells")
+
+b<-baseplot(melt[melt$variable=="GATA3",], "SM", "value", tb.cols)+    
+    facet_wrap(~variable)
+
+c<-baseplot(melt[melt$variable=="CXCR3",], "SM", "value", tb.cols)+    
+    facet_wrap(~variable)
+
+d<-baseplot(melt[melt$variable=="CCR4",], "SM", "value", tb.cols)+    
+    facet_wrap(~variable)
+
+cowplot::plot_grid(a, b, c, d,ncol=4)
+```
+
+```{r, fig.width=16, fig.height=8}
 #graph phenotype
@@
 ```
+
+#Supplementary
+
+```{r ics phenotyping, fig.width=16, fig.height=8}
+phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
+
+melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
+test<-filter(melt, variable!="TH1_x_x_x_x")
+test$variable<-gsub("TH1_", "", test$variable)
+test$variable<-factor(test$variable, levels=c(
+"t_x_3_4" ,       
+"t_x_3_x" ,         
+"t_x_x_4" ,         
+"t_x_x_x" ,
+"t_g_3_4" ,         
+"t_g_3_x" ,         
+"t_g_x_4" ,         
+"t_g_x_x" ,
+"x_g_3_4" ,         
+"x_g_3_x" ,         
+"x_g_x_4" ,         
+"x_g_x_x" ,
+"x_x_3_4" ,          
+"x_x_3_x" ,         
+"x_x_x_4"           
+))
+
+T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
+G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
+c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
+c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
+
+labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
+breaks <- levels(test$variable)
+
+test$tf<-NA
+test$tf[grep("x_g", test$variable)]<-"GATA3"
+test$tf[grep("t_x", test$variable)]<-"T-bet"
+test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
+test$tf[grep("x_x_._.", test$variable)]<-"None"
+test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
+
+booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free", ncol=4)+
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
+    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")
+
+```
+
+```{r}
+lum_x<-lum[lum$SM=="SM-",]
+df_x<-as.matrix(lum_x[,3:11])
+row.names(df_x)<-lum_x$Sample
+hc_x <- hclust(as.dist(1-cor(df_x)))
+x<-plot(hc_x)
+
+lum_sm<-lum[lum$SM=="SM+",]
+df_sm<-as.matrix(lum_sm[,3:11])
+row.names(df_sm)<-lum_sm$Sample
+hc_sm <- hclust(as.dist(1-cor(df_sm)))
+sm<-plot(hc_sm)
+
+df<-as.matrix(lum[,3:11])
+#heatmap(t(df_x), Rowv=as.dendrogram(hc_x))
+#heatmap(t(df_sm), Rowv=as.dendrogram(hc_sm))
+#heatmap(t(df), Rowv=as.dendrogram(hc_sm))
+```

### 2019-08-21 PaperFigures.Rmd
added in yerkes donors for reference

@@
     theme(plot.subtitle = element_text(hjust = 0.5)) +
-    stat_compare_means(
-        label="p.signif", hide.ns = TRUE, size=8,
-        label.x=1.3,
-        label.y = boxplot.stats(data[,yval])$stats[5]*1.4)+
-    coord_cartesian(ylim=c(0, boxplot.stats(data[,yval])$stats[5]*1.5))+
     labs(x="", y="")
@@
     theme(plot.subtitle = element_text(hjust = 0.5)) +
-    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=10)+
     labs(x="", y="")
@@
     ggplot(data, aes(data[,xval], data[,yval], fill=disease))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
+    geom_boxplot(size=1, position=position_dodge2(width=0.2, padding = 0.2, preserve = "single"), outlier.shape = NA)+
     scale_fill_manual(values = colors)+
     theme_classic()+ 
-    scale_x_discrete(labels=c("HC\nSM-", "HC\nSM+", "LTBI\nSM-", "LTBI\nSM+", "TB\nSM-", "TB\nSM+"))+
-    theme(legend.position="none")+
+    facet_wrap(~variable)+
+    scale_x_discrete(labels=c("\nNaive","SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
     theme(text = element_text(size=20))+
-    theme(plot.subtitle = element_text(hjust = 0.5)) +
-    stat_compare_means(
-        comparisons = list(c("HC SM+", "HC SM-"),c("LTBI SM+", "LTBI SM-"), c("TB SM+", "TB SM-")),
-        label="p.signif", size=8,
-        label.y = boxplot.stats(data[,yval])$stats[5]*1.4)+
-        coord_cartesian(ylim=c(0, boxplot.stats(data[,yval])$stats[5]*1.5))+
+    theme(plot.subtitle = element_text(hjust = 0.5))+
+    stat_compare_means(method = 'wilcox.test', label="p.signif", hide.ns=TRUE, size=10)+
     labs(x="", y="")
 }
+
+cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
+                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))
 
@@
 tb.cols<- c("SM+" = "#b2182b", "SM-"="#E59EA6")
-all.cols<- c("HC SM+" = "#1a9850", "HC SM-" ="#91cf60",
+all.cols<- c("Naive" = "#C8D0D8","HC SM+" = "#1a9850", "HC SM-" ="#91cf60",
              "LTBI SM+" = "#2166ac", "LTBI SM-" = "#85ABD1",
@@
 ```{r}
-cd4<-plot_filter(subset(CD4, CD4$Stim=="UN"))
-cyt<-plot_filter(subset(CYT, CYT$Stim=="PMA"))
-phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))
-og<-plot_filter(subset(OG, OG$Stim=="SEB"))
-ogphen<-plot_filter(subset(OGPHEN, OGPHEN$Stim=="SEB"))
+cd4<-filter_all(subset(CD4, CD4$Stim=="UN"))
+cyt<-filter_all(subset(CYT, CYT$Stim=="PMA"))
+phen<-filter_all(subset(PHEN, PHEN$Stim=="PMA"))
+og<-filter_all(subset(OG, OG$Stim=="SEB"))
+ogphen<-filter_all(subset(OGPHEN, OGPHEN$Stim=="SEB"))
 ```
 
-```{r, fig.width=16, fig.height=8}
+```{r, fig.width=20, fig.height=8}
 melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
 
-a<-allplot(melt[melt$variable=="T-bet",], "disease", "value", all.cols)+    
-    facet_wrap(~variable)+
-    labs(y="(%) Total CD4+ T cells")
+allplot(melt, "TB", "value", all.cols)+    
+    facet_wrap(~variable, scales="free", ncol=4)+
+    labs(y="(%) Total CD4+ T cells")+
+    theme(legend.position="bottom")
 
-b<-allplot(melt[melt$variable=="GATA3",], "disease", "value", all.cols)+    
-    facet_wrap(~variable)
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
 
-c<-allplot(melt[melt$variable=="CXCR3",], "disease", "value", all.cols)+    
-    facet_wrap(~variable)
-
-d<-allplot(melt[melt$variable=="CCR4",], "disease", "value", all.cols)+    
-    facet_wrap(~variable)
-
-cowplot::plot_grid(a, b, c, d,ncol=4)
-```
-
-```{r, fig.width=16, fig.height=8}
-melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
-cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
-                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))
-
-a<-allplot(melt[melt$variable=="IFNg",], "disease", "value", all.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)+
-    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-b<-allplot(melt[melt$variable=="TNFa",], "disease", "value", all.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)
-
-c<-allplot(melt[melt$variable=="IL4",], "disease", "value", all.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)
-
-d<-allplot(melt[melt$variable=="IL13",], "disease", "value", all.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)
-
-cowplot::plot_grid(a, b, c, d,ncol=4)
+allplot(melt, "TB", "value", all.cols)+    
+    facet_wrap(~variable, scales="free", labeller = cyt_groups, ncol=4)+
+    labs(y="(%) Total CD4+ T cells")+
+    theme(legend.position="bottom") 
 ```
@@
 ```{r, fig.width=18, fig.height=9.5}
-melt<-melt(cyt, id.vars=c("TB","SM"), measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
-cyt_groups <- as_labeller(c("TH1.Freq" = "TH1", 
+melt<-melt(subset(CYT, CYT$Stim=="PMA"), 
+           id.vars=c("TB","SM"), 
+           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
+th_groups <- as_labeller(c("TH1.Freq" = "TH1", 
                     "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))
-baseplot(melt, "SM", "value", healthy.cols)+
-    facet_wrap(~variable, ncol=4, labeller=cyt_groups)+
-    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")+coord_cartesian(ylim=c(0,40))
+
+cyt<-filter_hc(melt)
+a<-baseplot(cyt, "SM", "value", healthy.cols)+
+    facet_wrap(~variable, ncol=4, labeller=th_groups)+
+    labs(y="(%) CD4+ T cells")+coord_cartesian(ylim=c(0,45))
+
+cyt<-filter_ltbi(melt)
+b<-baseplot(cyt, "SM", "value", ltbi.cols)+
+    facet_wrap(~variable, ncol=4, labeller=th_groups)+
+    coord_cartesian(ylim=c(0,45))
+
+cyt<-filter_tb(melt)
+c<-baseplot(cyt, "SM", "value", tb.cols)+
+    facet_wrap(~variable, ncol=4, labeller=th_groups)+
+    coord_cartesian(ylim=c(0,45))
+
+cowplot::plot_grid(a, b, c, ncol=3)
 ```
@@
 
-```{r, fig.width=5, fig.height=10}
+```{r, fig.width=16, fig.height=20}
+phen<-filter_hc(subset(PHEN, PHEN$Stim=="PMA"))
 test<-melt(phen, id.vars=c("TB", "SM"), 
@@
 
-booleanplot(test, "SM","value", healthy.cols)+
+a<-booleanplot(test, "SM","value", healthy.cols)+
     facet_grid(variable~subset, scales="free")+
     theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    labs(y="Frequency of Marker+ CD4+ T cells (%)")
+    labs(y="(%) Cytokine+ CD4+ T cells")
+
+phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PMA"))
+test<-melt(phen, id.vars=c("TB", "SM"), 
+           measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4",
+           "TH2_T-bet", "TH2_GATA3", "TH2_CXCR3", "TH2_CCR4"))
+
+test$subset<-tstrsplit(test$variable, "_")[[1]]
+test$subset<-paste(test$subset, "Cytokine+\nCD4+ T Cells", sep=" ")
+test$variable<-gsub("TH1_","", test$variable)
+test$variable<-gsub("TH2_","", test$variable)
+test$variable<-factor(test$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+
+b<-booleanplot(test, "SM","value", ltbi.cols)+
+    facet_grid(variable~subset, scales="free")+
+    theme(panel.border = element_rect(fill = NA, colour = "black"))
+
+phen<-filter_tb(subset(PHEN, PHEN$Stim=="PMA"))
+test<-melt(phen, id.vars=c("TB", "SM"), 
+           measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4",
+           "TH2_T-bet", "TH2_GATA3", "TH2_CXCR3", "TH2_CCR4"))
+
+test$subset<-tstrsplit(test$variable, "_")[[1]]
+test$subset<-paste(test$subset, "Cytokine+\nCD4+ T Cells", sep=" ")
+test$variable<-gsub("TH1_","", test$variable)
+test$variable<-gsub("TH2_","", test$variable)
+test$variable<-factor(test$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+
+c<-booleanplot(test, "SM","value", tb.cols)+
+    facet_grid(variable~subset, scales="free")+
+    theme(panel.border = element_rect(fill = NA, colour = "black"))
+
+cowplot::plot_grid(a, b, c, ncol=3)
 ```
@@
 melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
-cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
-                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))
 
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
+    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y = (.95*.15))+
+    coord_cartesian(ylim=c(0,.15))+
+    labs(y="(%) CD4+ T cells")
 
 b<-baseplot(melt[melt$variable=="TNFa",], "SM", "value", ltbi.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)
+    facet_wrap(~variable, labeller = cyt_groups)+
+    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y = (.95*.15))+
+    coord_cartesian(ylim=c(0,.15))
 
 c<-baseplot(melt[melt$variable=="IL4",], "SM", "value", ltbi.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)
+    facet_wrap(~variable, labeller = cyt_groups)+
+    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y = (.95*.013))+
+    coord_cartesian(ylim=c(0,.013))
 
 d<-baseplot(melt[melt$variable=="IL13",], "SM", "value", ltbi.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)
+    facet_wrap(~variable, labeller = cyt_groups)+
+    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y = (.95*.02))+
+    coord_cartesian(ylim=c(0,.02))
 
@@
 melt$variable<-gsub("ifng_", "", melt$variable)
+melt$variable <- factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))
 
-a<-baseplot(melt[melt$variable=="T-bet",], "SM", "value", ltbi.cols)+    
+a<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
     facet_wrap(~variable)+
-    labs(y="(%) Total CD4+ T cells")
+    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y=(.98*70))+
+    coord_cartesian(ylim=c(0,70))+
+    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
 
-b<-baseplot(melt[melt$variable=="GATA3",], "SM", "value", ltbi.cols)+    
-    facet_wrap(~variable)
-c<-baseplot(melt[melt$variable=="CXCR3",], "SM", "value", ltbi.cols)+    
-    facet_wrap(~variable)
+b<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
+    facet_wrap(~variable)+
+    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y=(.98*100))+
+    coord_cartesian(ylim=c(0,100))
 
-d<-baseplot(melt[melt$variable=="CCR4",], "SM", "value", ltbi.cols)+    
-    facet_wrap(~variable)
+cowplot::plot_grid(a,b,ncol=2)
+```
+
+#TB 
+```{r, narrow down tb data}
+cyt<-filter_tb(subset(CYT, CYT$Stim=="PEP"))
+phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
+```
+####Similar to QFT+, individuals with active TB, co-infected with S.
+mansoni maintain function Mtb-specific CD4+ T cell cytokine production.
+There are higher frequencies of TH1/2 cytokine co-producing CD4+ T cells amongst SM+ individuals, however the frequency of these cells is incredibly low.
+The dominant cytokine subgroup is still TH1 cytokine producing cells.
+*Again this is where I need an aesthetics opinion*
+
+```{r, fig.width=16, fig.height=8}
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
+
+a<-baseplot(melt[melt$variable=="IFNg",], "disease", "value", tb.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)+
+    labs(y="(%) CD4+ T cells")+
+    coord_cartesian(ylim=c(0,.06))
+
+b<-baseplot(melt[melt$variable=="TNFa",], "disease", "value", tb.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)+
+    coord_cartesian(ylim=c(0,.15))
+
+c<-baseplot(melt[melt$variable=="IL4",], "disease", "value", tb.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)+
+    coord_cartesian(ylim=c(0,.02))
+
+d<-baseplot(melt[melt$variable=="IL13",], "disease", "value", tb.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)+
+    coord_cartesian(ylim=c(0,.02))
 
 cowplot::plot_grid(a, b, c, d,ncol=4)
+```
+
+####Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers.
+There are numerous phenotypic differences amongst Mtb-specific TH1 cytokine producing cells betweeen SM- and SM+ groups.
+While T-bet is still the dominant transcription factor expressed in both groups, most differences are due to higher frequencies of GATA3 and CCR4 co-expressing cells.
+
+```{r main four markers, fig.width=16, fig.height=8}
+melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4"))
+melt$variable<-gsub("TH1_", "", melt$variable)
+melt$variable<-factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+
+a<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", tb.cols)+    
+    facet_wrap(~variable)+
+    stat_compare_means(label="p.format", size=6, label.y=(.98*80))+
+    coord_cartesian(ylim=c(0,80))+
+    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
+
+b<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", tb.cols)+    
+    facet_wrap(~variable)+
+    stat_compare_means(label="p.format", size=6, label.y=(.98*80))+
+    coord_cartesian(ylim=c(0,80))
+
+cowplot::plot_grid(a, b, ncol=2)
+```
+
+```{r, fig.width=16, fig.height=8}
+#graph phenotype
+melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
+test<-filter(melt, variable!="TH1_x_x_x_x")
+test$variable<-gsub("TH1_", "", test$variable)
+test$variable<-factor(test$variable, levels=c(
+"t_x_3_4" ,       
+"t_x_3_x" ,         
+"t_x_x_4" ,         
+"t_x_x_x" ,
+"t_g_3_4" ,         
+"t_g_3_x" ,         
+"t_g_x_4" ,         
+"t_g_x_x" ,
+"x_g_3_4" ,         
+"x_g_3_x" ,         
+"x_g_x_4" ,         
+"x_g_x_x" ,
+"x_x_3_4" ,          
+"x_x_3_x" ,         
+"x_x_x_4"           
+))
+
+T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
+G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
+c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
+c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
+
+labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
+breaks <- levels(test$variable)
+
+test$tf<-NA
+test$tf[grep("x_g", test$variable)]<-"GATA3"
+test$tf[grep("t_x", test$variable)]<-"T-bet"
+test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
+test$tf[grep("x_x_._.", test$variable)]<-"None"
+test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
+
+booleanplot(test, "variable", "value", tb.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")+
+    coord_cartesian(ylim=c(0,55))+
+    stat_compare_means(label="p.signif", hide.ns=TRUE, label.y=50, size=10)
+
 ```
@@
 a<-baseplot(og, "SM", "OG_freq", ltbi.cols)+
-    labs(y="Proliferating CD4+ T cells (%)")+coord_cartesian(ylim=c(0,5))
+    labs(y="(%) CD4+ T cells")+
+    facet_wrap(~TB, labeller = as_labeller(c("LTBI"="Proliferation")))+
+    coord_cartesian(ylim=c(0,3))
+
 melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._.", names(ogphen)))
@@
 ))
+test$group<-"TH1/2"
+for(i in 1:length(test$group)){
+    if(test$variable[i] %in% c("T_G_x", "T_x_x", "x_G_x")){
+        test$group[i]<- "TH1"
+    }
+    else if (test$variable[i] == "x_x_4"){
+        test$group[i]<- "TH2"
+    }
+}
 
@@
 b<-booleanplot(test, "variable", "value", ltbi.cols)+
-    scale_x_discrete(labels = labels)+
-    labs(y="Cytokine+ Proliferating\nCD4+ T cells (%)")
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    facet_grid(~group, scales="free_x", space="free_x")+
+    labs(y="(%) Proliferating CD4+ T cells")
 
-cowplot::plot_grid(a, b, ncol=2, rel_widths=c(1.5,3.5), align="h")
+cowplot::plot_grid(a, b, ncol=2, rel_widths=c(1,3.5), align="h")
+```
+
+####Similar to the overnight assay, expanded Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers, however this does not differ by S. mansoni status.
+
+*I maybe want to do some analysis about the fact that now higher frequencies of cells express T-bet and/or GATA3 and that there seems to be a preferential expansion of CCR4+ cells but I honestly don't know how to test that*
+
+```{r og main four markers, fig.width=16, fig.height=8}
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+melt$variable<-factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+
+a<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
+    facet_wrap(~variable)+
+    stat_compare_means(label="p.format", size=6, label.y=(.98*100))+
+    coord_cartesian(ylim=c(0,100))+
+    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
+
+b<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
+    facet_wrap(~variable)+
+    stat_compare_means(label="p.format", size=6, label.y=(.98*100))+
+    coord_cartesian(ylim=c(0,100))
+
+cowplot::plot_grid(a,b,ncol=2)
+```
+
+```{r og phenotype, fig.width=16, fig.height=8}
+#graph phenotype
+melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._._.", names(ogphen)))
+test<-filter(melt, variable!="x_x_x_x")
+test$variable<-factor(test$variable, levels=c(
+"t_x_3_4" ,       
+"t_x_3_x" ,         
+"t_x_x_4" ,         
+"t_x_x_x" ,
+"t_g_3_4" ,         
+"t_g_3_x" ,         
+"t_g_x_4" ,         
+"t_g_x_x" ,
+"x_g_3_4" ,         
+"x_g_3_x" ,         
+"x_g_x_4" ,         
+"x_g_x_x" ,
+"x_x_3_4" ,          
+"x_x_3_x" ,         
+"x_x_x_4"           
+))
+
+T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
+G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
+c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
+c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
+
+labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
+breaks <- levels(test$variable)
+
+test$tf<-NA
+test$tf[grep("x_g", test$variable)]<-"GATA3"
+test$tf[grep("t_x", test$variable)]<-"T-bet"
+test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
+test$tf[grep("x_x_._.", test$variable)]<-"None"
+test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
+
+booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
+    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
 ```
@@
 
-####Similar to the overnight assay, expanded Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers, however this does not differ by S. mansoni status.
-
-*I maybe want to do some analysis about the fact that now higher frequencies of cells express T-bet and/or GATA3 and that there seems to be a preferential expansion of CCR4+ cells but I honestly don't know how to test that*
-
-```{r og main four markers, fig.width=16, fig.height=8}
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
-
-a<-baseplot(melt[melt$variable=="T-bet",], "disease", "value", ltbi.cols)+    
-    facet_wrap(~variable)+
-    labs(y="(%) Total CD4+ T cells")+coord_cartesian(ylim=c(0,100))
-
-b<-baseplot(melt[melt$variable=="GATA3",], "disease", "value", ltbi.cols)+    
-    facet_wrap(~variable)+coord_cartesian(ylim=c(0,100))
-
-c<-baseplot(melt[melt$variable=="CXCR3",], "disease", "value", ltbi.cols)+    
-    facet_wrap(~variable)+coord_cartesian(ylim=c(0,100))
-
-d<-baseplot(melt[melt$variable=="CCR4",], "disease", "value", ltbi.cols)+    
-    facet_wrap(~variable)+coord_cartesian(ylim=c(0,100))
-
-cowplot::plot_grid(a, b, c, d,ncol=4)
-```
-
-```{r og phenotype, fig.width=16, fig.height=8}
-#graph phenotype
-melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._._.", names(ogphen)))
-test<-filter(melt, variable!="x_x_x_x")
-test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
-
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
-
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
-
-test$tf<-NA
-test$tf[grep("x_g", test$variable)]<-"GATA3"
-test$tf[grep("t_x", test$variable)]<-"T-bet"
-test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
-test$tf[grep("x_x_._.", test$variable)]<-"None"
-test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
-
-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free", ncol=4)+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
-    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")
-```
-
-#TB  
-####Similar to QFT+, individuals with active TB, co-infected with S.
-mansoni maintain function Mtb-specific CD4+ T cell cytokine production.
-There are higher frequencies of TH1/2 cytokine co-producing CD4+ T cells amongst SM+ individuals, however the frequency of these cells is incredibly low.
-The dominant cytokine subgroup is still TH1 cytokine producing cells.
-*Again this is where I need an aesthetics opinion*
-```{r tb cytokine ics, fig.width=16, fig.height=8}
-cyt<-filter_tb(subset(CYT, CYT$Stim=="PEP"))
-phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
-og<-filter_tb(subset(OG, OG$Stim=="PEP"))
-ogphen<-filter_tb(subset(OGPHEN, OGPHEN$Stim=="PEP"))
-
-#generate cytokine table for graphing
-melt<-melt(cyt, 
-    id.vars=c("TB","SM"), 
-    measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
-cyt_groups <- as_labeller(c("TH1.Freq" = "TH1", 
-                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))
-#graph cytokine with facets
-keep <- baseplot(melt, "SM", "value", tb.cols)+
-    facet_wrap(~variable, ncol=4, labeller=cyt_groups)+
-    labs(y="Frequency of Cytokine+ \nCD4+ T cells (%)")
-```
-
-```{r, fig.width=16, fig.height=8}
-melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
-cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
-                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))
-
-a<-baseplot(melt[melt$variable=="IFNg",], "disease", "value", tb.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)+
-    labs(y="Frequency of Cytokine+ CD4+ T cells (%)")
-
-b<-baseplot(melt[melt$variable=="TNFa",], "disease", "value", tb.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)
-
-c<-baseplot(melt[melt$variable=="IL4",], "disease", "value", tb.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)
-
-d<-baseplot(melt[melt$variable=="IL13",], "disease", "value", tb.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)
-
-cowplot::plot_grid(a, b, c, d,ncol=4)
-```
-
-####Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers.
-There are numerous phenotypic differences amongst Mtb-specific TH1 cytokine producing cells betweeen SM- and SM+ groups.
-While T-bet is still the dominant transcription factor expressed in both groups, most differences are due to higher frequencies of GATA3 and CCR4 co-expressing cells.
-
-```{r main four markers, fig.width=16, fig.height=8}
-melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4"))
-melt$variable<-gsub("TH1_", "", melt$variable)
-
-a<-baseplot(melt[melt$variable=="T-bet",], "SM", "value", tb.cols)+    
-    facet_wrap(~variable)+
-    labs(y="(%) Total CD4+ T cells")
-
-b<-baseplot(melt[melt$variable=="GATA3",], "SM", "value", tb.cols)+    
-    facet_wrap(~variable)
-
-c<-baseplot(melt[melt$variable=="CXCR3",], "SM", "value", tb.cols)+    
-    facet_wrap(~variable)
-
-d<-baseplot(melt[melt$variable=="CCR4",], "SM", "value", tb.cols)+    
-    facet_wrap(~variable)
-
-cowplot::plot_grid(a, b, c, d,ncol=4)
-```
-
-```{r, fig.width=16, fig.height=8}
-#graph phenotype
-melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
-test<-filter(melt, variable!="TH1_x_x_x_x")
-test$variable<-gsub("TH1_", "", test$variable)
-test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
-
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
-
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
-
-test$tf<-NA
-test$tf[grep("x_g", test$variable)]<-"GATA3"
-test$tf[grep("t_x", test$variable)]<-"T-bet"
-test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
-test$tf[grep("x_x_._.", test$variable)]<-"None"
-test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
-
-booleanplot(test, "variable", "value", tb.cols)+facet_wrap(~tf, scales="free", ncol=4)+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")
-
-```
-
 #Supplementary
@@
     stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
-    labs(y="Frequency of Transcription Factor \n Mtb-Specific TH1 cells (%)")
+    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
 

### 2019-08-26 Cytokine_Analysis.Rmd
proportions of tf cells making cytokine

@@
 
-CD4<-plot_filter(read.csv("ICS_CD4_Cytokine_clean.csv"))
+CD4<-read.csv("ICS_CD4_Cytokine_clean.csv")
 
-CD8<-plot_filter(read.csv("ICS_CD8_Cytokine_clean.csv"))
+CD8<-read.csv("ICS_CD8_Cytokine_clean.csv")
 
-MFI<-plot_filter(read.csv("ICS_CD4_MFI_clean.csv"))
+MFI<-read.csv("ICS_CD4_MFI_clean.csv")
 
-Compass<-plot_filter(read.csv("ICS_Compass_scores_complete_clean.csv"))
+Compass<-read.csv("ICS_Compass_scores_complete_clean.csv")
 
-GD<-plot_filter(read.csv("ICS_GD_Cytokine_clean.csv"))
+GD<-read.csv("ICS_GD_Cytokine_clean.csv")
 
@@
     #inserts each character string into the column argument of the pvals function written above
-    b<-pvals_4(a, col)
-    if(dim(b)[1]!=0)
-    print(kable(b, caption=col)) #will return the output for each character string in the loop
+    b<-pvals_6(a, col)
+    if(dim(b)[1]!=0){
+    print(kable(b, caption=col))#will return the output for each character string in the loop
+    y<-plot_with_all_stats(a, col)+labs(title=col)
+    print(y)}
     }

### 2019-08-26 ICS_Cleaning.Rmd
proportions of tf cells making cytokine

@@
 
-l<-list(CD4_Counts.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_new.csv,
+l<-list(CD4_Counts.csv, CD4_cytokine_new.csv, CD4_Cytokine.csv, CD4_MFI.csv, CD4_new.csv,
         CD4_phenotype_new.csv, CD4_phenotype.csv, CD4.csv, 
@@
 ```{r update and organize cytokine variables}
+library(dplyr)
+CD4_Cytokine<-merge(CD4_Cytokine.csv, CD4_cytokine_new.csv, by=c("Sample","Donor","TB","SM","Stim" ), all=TRUE)
 #If you want to know the double producers without considering the opposite TH cytokines
-library(dplyr)
-CD4_Cytokine.csv<-dplyr::mutate(CD4_Cytokine.csv, G_T = G_4_x_T + G_x_13_T + G_x_x_T,
+CD4_Cytokine<-dplyr::mutate(CD4_Cytokine, 
+         G_T = G_4_x_T + G_x_13_T + G_x_x_T,
          IL4_IL13 = G_4_13_x + x_4_13_T + x_4_13_x,
@@
          TH2.Total = TH2.Freq + TH1.2.Freq
+                   )
+#To collapse boolean gates in transcription factor exported data
+CD4_Cytokine<-dplyr::mutate(CD4_Cytokine, 
+        tbet_ifng=  rowSums(CD4_Cytokine[,grep("tbet_G", names(CD4_Cytokine))]),
+        tbet_tnfa=  rowSums(CD4_Cytokine
+                    [intersect(grep("tbet", names(CD4_Cytokine)) , grep("_T", names(CD4_Cytokine)))]),
+        tbet_il4=   rowSums(CD4_Cytokine[,grep("tbet_._4_", names(CD4_Cytokine))]),
+        tbet_il13=  rowSums(CD4_Cytokine[,grep("tbet_._._13", names(CD4_Cytokine))]),
+        gata3_ifng= rowSums(CD4_Cytokine[,grep("gata3_G", names(CD4_Cytokine))]),
+        gata3_tnfa= rowSums(CD4_Cytokine
+                    [intersect(grep("gata3", names(CD4_Cytokine)) , grep("_T", names(CD4_Cytokine)))]),
+        gata3_il4=  rowSums(CD4_Cytokine[,grep("gata3_._4_", names(CD4_Cytokine))]),
+        gata3_il13= rowSums(CD4_Cytokine[,grep("gata3_._._13", names(CD4_Cytokine))])
+            )
+
+CD4_Cytokine<-dplyr::mutate(CD4_Cytokine, 
+        tbet_th1=  tbet_G_x_x_T + tbet_G_x_x_x + tbet_x_x_x_T,
+        tbet_th2=  tbet_x_4_13_x + tbet_x_4_x_x + tbet_x_x_13_x,
+        tbet_th1.2=tbet_G_4_13_T + tbet_G_4_13_x + tbet_G_4_x_T + tbet_G_4_x_x + 
+            tbet_G_x_13_T + tbet_G_x_13_x + tbet_x_4_13_T + tbet_x_4_x_T + tbet_x_x_13_T,
+        gata3_th1=  gata3_G_x_x_T + gata3_G_x_x_x + gata3_x_x_x_T,
+        gata3_th2=  gata3_x_4_13_x + gata3_x_4_x_x + gata3_x_x_13_x,
+        gata3_th1.2=gata3_G_4_13_T + gata3_G_4_13_x + gata3_G_4_x_T + gata3_G_4_x_x + 
+            gata3_G_x_13_T + gata3_G_x_13_x + gata3_x_4_13_T + gata3_x_4_x_T + gata3_x_x_13_T 
             )
@@
 #just subtracts the value of the unstim from the value of the stim by donor
-filenames = c("CD4_Cytokine.csv", "CD8_Cytokine.csv", "GD_Cytokine.csv")
-l<-list(CD4_Cytokine.csv, CD8_Cytokine.csv, GD_Cytokine.csv)
+filenames = c("CD4_Cytokine", "CD8_Cytokine.csv", "GD_Cytokine.csv")
+l<-list(CD4_Cytokine, CD8_Cytokine.csv, GD_Cytokine.csv)
 l<-lapply(l, function(x) arrange(x, Donor))
@@
 #sets all negative values to zero
-CD4_Cytokine.csv[CD4_Cytokine.csv<0]<-0
+CD4_Cytokine[CD4_Cytokine<0]<-0
 CD8_Cytokine.csv[CD8_Cytokine.csv<0]<-0
@@
 write.csv(CD4_Counts.csv, file = "ICS_CD4_Counts_clean.csv")
-write.csv(CD4_Cytokine.csv, file = "ICS_CD4_Cytokine_clean.csv")
+write.csv(CD4_Cytokine, file = "ICS_CD4_Cytokine_clean.csv")
 write.csv(CD4_MFI.csv, file = "ICS_CD4_MFI_clean.csv")

### 2019-08-26 PaperFigures.Rmd
took out yerkes donors for committee meeting figures

@@
     facet_wrap(~variable)+
-    scale_x_discrete(labels=c("\nNaive","SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
     theme(text = element_text(size=20))+
@@
 ```{r}
-cd4<-filter_all(subset(CD4, CD4$Stim=="UN"))
-cyt<-filter_all(subset(CYT, CYT$Stim=="PMA"))
-phen<-filter_all(subset(PHEN, PHEN$Stim=="PMA"))
-og<-filter_all(subset(OG, OG$Stim=="SEB"))
-ogphen<-filter_all(subset(OGPHEN, OGPHEN$Stim=="SEB"))
+cd4<-plot_filter(subset(CD4, CD4$Stim=="UN"))
+cyt<-plot_filter(subset(CYT, CYT$Stim=="PMA"))
+phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))
+og<-plot_filter(subset(OG, OG$Stim=="SEB"))
+ogphen<-plot_filter(subset(OGPHEN, OGPHEN$Stim=="SEB"))
 ```
 
-```{r, fig.width=20, fig.height=8}
+```{r, fig.width=15, fig.height=8}
 melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
@@
 ####Generic cytokine production from CD4+ T cells stimulated with PMA & Ionomycin does not differ between SM- and SM+ Healthy Adults
-```{r, fig.width=18, fig.height=9.5}
+```{r, fig.width=15, fig.height=8}
 melt<-melt(subset(CYT, CYT$Stim=="PMA"), 
@@
 
-```{r, fig.width=16, fig.height=20}
+```{r, fig.width=15, fig.height=8}
 phen<-filter_hc(subset(PHEN, PHEN$Stim=="PMA"))
@@
 
-```{r cytokine, fig.width=16, fig.height=8}
+```{r cytokine, fig.width=15, fig.height=8}
 melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
@@
 ####Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers, however this does not differ by S. mansoni status
-```{r ics main four markers, fig.width=16, fig.height=8}
+```{r ics main four markers, fig.width=15, fig.height=8}
 melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("ifng_T-bet", "ifng_GATA3", "ifng_CXCR3", "ifng_CCR4"))
@@
 
-```{r, fig.width=16, fig.height=8}
+```{r, fig.width=15, fig.height=8}
 melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
@@
 
-```{r main four markers, fig.width=16, fig.height=8}
+```{r main four markers, fig.width=15, fig.height=8}
 melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4"))
@@
 
-```{r, fig.width=16, fig.height=8}
+```{r, fig.width=15, fig.height=8}
 #graph phenotype
@@
 The dominant cytokine subgroup is still cells that co-produce IFNγ and TNFα.
-```{r og and cyt, fig.width=16, fig.height=8}
+```{r og and cyt, fig.width=15, fig.height=8}
 a<-baseplot(og, "SM", "OG_freq", ltbi.cols)+
@@
 
-cowplot::plot_grid(a, b, ncol=2, rel_widths=c(1,3.5), align="h")
+cowplot::plot_grid(a, b, ncol=2, rel_widths=c(1.25,3.75), align="h")
 ```

### 2019-08-27 PaperFigures.Rmd
added in proliferation data

@@
 OG.GT<-read.csv("OG_CD4_OG.GT_summary_clean_NAs.csv")
+OG.th1<-read.csv("OG_CD4_OG.TH1_summary_clean_NAs.csv")
 luminex<-read.csv("/Applications/Old Computer/Day Lab/Luminex/Clean Data/Luminex_combined_background.txt")
@@
 #All groups
-Dont worry I can change this so the ns bars don't show up, especially for the cytokines since I have no idea why that giant NS is there. But for now I'm leaving them so we know.
+Statistics were performed using the Mann-Whitney U test for non-parametric data. P-values <0.05 were considered significant.
 ```{r}
@@
 ```{r, fig.width=15, fig.height=8}
-melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))
 
-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, scales="free", ncol=4)+
+tf<-allplot(melt, "TB", "value", all.cols)+    
+    facet_wrap(~variable, ncol=2)+
+    theme(legend.position="none") +
+    labs(y="(%) Total CD4+ T cells")
+
+melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))
+
+sm<-allplot(melt, "TB", "value", all.cols)+    
+    facet_wrap(~variable, ncol=2)+
+    theme(legend.position="none")
+
+cowplot::plot_grid(tf, sm)
+
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
+
+one<-allplot(melt, "TB", "value", all.cols)+    
+    facet_wrap(~variable, scales="free", labeller = cyt_groups, ncol=2)+
     labs(y="(%) Total CD4+ T cells")+
-    theme(legend.position="bottom")
+    theme(legend.position="none")
 
-melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))
 
-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, scales="free", labeller = cyt_groups, ncol=4)+
-    labs(y="(%) Total CD4+ T cells")+
-    theme(legend.position="bottom") 
+four<-allplot(melt, "TB", "value", all.cols)+    
+    facet_wrap(~variable, scales="free", labeller = cyt_groups, ncol=1)+
+    theme(legend.position="none")+
+    coord_cartesian(ylim=c(0,7))
+
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL13"))
+
+thirteen<-allplot(melt, "TB", "value", all.cols)+    
+    facet_wrap(~variable, scales="free", labeller = cyt_groups, ncol=1)+
+    theme(legend.position="none")+
+    coord_cartesian(ylim=c(0,3))
+
+cowplot::plot_grid(one, four, thirteen, ncol=3, rel_widths=c(7.5,3.75, 3.75))
 ```
 
-#Healthy Adults
-*Is this superfluous now?*  
+```{r pvals for ifng and tnfa}
+y<-split(cyt,cyt$SM) 
+A<-lapply(y, function(g) kruskal.test(IFNg~TB, g))
+B<-lapply(y, function(g) kruskal.test(TNFa~TB, g))
+
+pval<-c(A$`SM-`$p.value, A$`SM+`$p.value, B$`SM-`$p.value, B$`SM+`$p.value)
+comparison<-c("SM- IFNγ", "SM+ IFNγ", "SM- TNFα", "SM+ TNFα")
+kable(cbind(comparison, pval))
+```
+
 ####Generic cytokine production from CD4+ T cells stimulated with PMA & Ionomycin does not differ between SM- and SM+ Healthy Adults
@@
 phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
-og<-filter_ltbi(subset(OG, OG$Stim=="PEP"))
-ogphen<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
-oggt<-filter_ltbi(subset(OG.GT, OG.GT$Stim=="PEP"))
 ```
@@
 ```
+
 ####Similar to QFT+, individuals with active TB, co-infected with S.
@@
 
-```{r, fig.width=15, fig.height=8}
-#graph phenotype
-melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
-test<-filter(melt, variable!="TH1_x_x_x_x")
-test$variable<-gsub("TH1_", "", test$variable)
-test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
+```{r receptor on gata3, fig.width=7.5, fig.height=8}
+melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("TH1_GATA3CCR4", "TH1_GATA3CXCR3"))
+melt$variable<-gsub("TH1_", "", melt$variable)
+melt$variable<-gsub("GATA3", "GATA3+", melt$variable)
+melt$variable<-gsub("CXCR3", "CXCR3+", melt$variable)
+melt$variable<-gsub("CCR4", "CCR4+", melt$variable)
 
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
+baseplot(melt, "SM", "value", tb.cols)+    
+    facet_wrap(~variable)+
+    stat_compare_means(label="p.format", size=6, label.y=(.98*15))+
+    coord_cartesian(ylim=c(0,15))+
+    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
+```
 
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
+#Proliferation Assay
+##All Groups Responses to SEB
+```{r}
+og<-plot_filter(subset(OG, OG$Stim=="SEB"))
+ogphen<-plot_filter(subset(OGPHEN, OGPHEN$Stim=="SEB"))
+```
 
-test$tf<-NA
-test$tf[grep("x_g", test$variable)]<-"GATA3"
-test$tf[grep("t_x", test$variable)]<-"T-bet"
-test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
-test$tf[grep("x_x_._.", test$variable)]<-"None"
-test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
+```{r, proliferation and total cyt, fig.width=15, fig.height=8}
+melt<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars=c("OG_freq", "IFNg_freq", "TNFa_freq", "IL4_freq"))
 
-booleanplot(test, "variable", "value", tb.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")+
-    coord_cartesian(ylim=c(0,55))+
-    stat_compare_means(label="p.signif", hide.ns=TRUE, label.y=50, size=10)
+groups <- as_labeller(c("IFNg_freq" = "IFNγ", 
+                    "TNFa_freq" = "TNFα", "IL4_freq" = "IL-4", "OG_freq" = "Proliferation"))
 
+allplot(melt, "TB", "value", all.cols)+
+    facet_wrap(~variable, scales="free", labeller = groups, ncol=4)+
+    labs(y="(%) Total CD4+ T cells")+
+    theme(legend.position="bottom") 
+```
+
+```{r, cyt from proliferating, fig.width=12, fig.height=8}
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg_freq", "TNFa_freq", "IL4_freq"))
+
+groups <- as_labeller(c("IFNg_freq" = "IFNγ", 
+                    "TNFa_freq" = "TNFα", "IL4_freq" = "IL-4"))
+
+allplot(melt, "TB", "value", all.cols)+
+    facet_wrap(~variable, scales="free", labeller = groups, ncol=3)+
+    labs(y="(%) Proliferating CD4+ T cells")+
+    theme(legend.position="bottom") 
+```
+
+```{r, marker on proliferating, fig.width=15, fig.height=8}
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet_freq", "GATA3_freq", "CXCR3_freq", "CCR4_freq"))
+
+melt$variable<-gsub("_freq", "", melt$variable)
+melt$variable<-factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+
+allplot(melt, "TB", "value", all.cols)+
+    facet_wrap(~variable, scales="free", ncol=4)+
+    labs(y="(%) Total CD4+ T cells")+
+    theme(legend.position="bottom") 
+```
+
+##Mtb-Specific Responses in LTBI
+```{r}
+og<-filter_ltbi(subset(OG, OG$Stim=="PEP"))
+ogphen<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
+oggt<-filter_ltbi(subset(OG.GT, OG.GT$Stim=="PEP"))
+ogth1<-filter_ltbi(subset(OG.th1, OG.GT$Stim=="PEP"))
 ```
@@
 *I maybe want to do some analysis about the fact that now higher frequencies of cells express T-bet and/or GATA3 and that there seems to be a preferential expansion of CCR4+ cells but I honestly don't know how to test that*
-
+###On proliferating cells
 ```{r og main four markers, fig.width=16, fig.height=8}
@@
     coord_cartesian(ylim=c(0,100))+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
+    labs(y="(%) Proliferating Mtb-Specific cells")
 
@@
 
+###On TH1 proliferating cells
+```{r og th1 main four markers, fig.width=16, fig.height=8}
+melt<-melt(ogth1, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+melt$variable<-factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+
+a<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
+    facet_wrap(~variable)+
+    stat_compare_means(label="p.format", size=6, label.y=(.98*100))+
+    coord_cartesian(ylim=c(0,100))+
+    labs(y="(%) Proliferating Mtb-Specific cells")
+
+b<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
+    facet_wrap(~variable)+
+    stat_compare_means(label="p.format", size=6, label.y=(.98*100))+
+    coord_cartesian(ylim=c(0,100))
+
+cowplot::plot_grid(a,b,ncol=2)
+```
+
+#Luminex Data
+
+####Furthermore, IFNγ and TNFa are also the two dominant cytokines produced when analyzed by luminex. The concentration of all cytokines measured by luminex are equivalent between SM- and SM+ groups.
+
+*Honestly I will have to just remake the whole thing a couple times and stitch them together to fix the axes so just ignore them for now*
+```{r luminex, fig.width=12, fig.height=16}
+lum<-filter_ltbi(subset(luminex, luminex$Stim=="PEP"))
+names(lum)<-gsub("IL", "IL-", names(lum))
+names(lum)<-gsub("IFNg", "IFNγ", names(lum))
+names(lum)<-gsub("TNFa", "TNFα", names(lum))
+
+melt<-melt(lum, id.vars=c("TB", "SM"), measure.vars=grep("pg", names(lum)))
+melt$variable<-gsub("_pg.ml", "", melt$variable)
+melt$variable<- factor(melt$variable, levels=c("TNFα", "IFNγ", "IL-21", "IL-17A", "IL-22", "IL-10", "IL-4", "IL-5", "IL-13"))
+
+booleanplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+
+    labs(y="pg/mL")
+```
+
+####However the correlation between cytokines differs between groups. Of note, the correlation between IFNγ and TNFa is stronger amongst SM+ individuals than SM- individuals.
+```{r cor plots}
+df<-lum[,3:11]
+names(lum)<-gsub("_pg.ml", "", names(lum))
+pca1 <- prcomp(df, scale = TRUE)
+pcas<-data.frame(pca1$x)
+rotations<-data.frame(pca1$rotation)
+rotations$cytokine<-row.names(rotations)
+melt<-melt(rotations, id.vars="cytokine")
+
+stims<-factor(lum$Stim)
+worms<-factor(lum$SM)
+a<-ggplot(pcas, aes(x=pcas[,1], y=pcas[,2]))+geom_point(aes(col=worms))+labs(y="PCA2", x="PCA1")
+
+neg <- cor(lum[lum$SM=="SM-",3:11])
+order <- corrMatOrder(neg, order = "hclust", hclust.method="centroid")
+order.neg <- neg[order, order]
+res1 <- cor.mtest(lum[lum$SM=="SM-",3:11], conf.level = .95)
+pvals <- res1$p[order, order]
+corrplot(order.neg, method = "circle", insig = "label_sig", tl.col = "black", 
+         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = 2, pch.col = "white",
+         type="upper")
+
+pos <- cor(lum[lum$SM=="SM+",3:11])
+order.pos <- pos[order, order]
+res1 <- cor.mtest(lum[lum$SM=="SM+",3:11], conf.level = .95)
+pvals <- res1$p[order, order]
+corrplot(order.pos, method = "circle", insig = "label_sig", tl.col = "black", 
+         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = 2, pch.col = "white",
+         type="upper")
+```
+
+#Supplementary
+###Peptide-Specific Cytokine Production  
+```{r, fig.width=15, fig.height=8}
+melt<-melt(subset(CYT, CYT$Stim=="PEP"), 
+           id.vars=c("TB","SM"), 
+           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
+th_groups <- as_labeller(c("TH1.Freq" = "TH1", 
+                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))
+
+cyt<-filter_hc(melt)
+a<-baseplot(cyt, "SM", "value", healthy.cols)+
+    facet_wrap(~variable, ncol=4, labeller=th_groups)+
+    labs(y="(%) CD4+ T cells")+coord_cartesian(ylim=c(0,.2))
+
+cyt<-filter_ltbi(melt)
+b<-baseplot(cyt, "SM", "value", ltbi.cols)+
+    facet_wrap(~variable, ncol=4, labeller=th_groups)+
+    coord_cartesian(ylim=c(0,.2))
+
+cyt<-filter_tb(melt)
+c<-baseplot(cyt, "SM", "value", tb.cols)+
+    facet_wrap(~variable, ncol=4, labeller=th_groups)+
+    coord_cartesian(ylim=c(0,.2))
+
+cowplot::plot_grid(a, b, c, ncol=3)
+```
+
+##Some Boolean Cytokine graphs will go here for posterity
+*but I need to make them pretty and I dont' feel like doing that right now*
+```{r}
+cyt<-filter_ltbi(CYT[CYT$Stim=="PMA",])
+test<-melt(cyt, id.vars=c("TB", "SM"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
+test$variable<-factor(test$variable, levels=c(
+"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
+"G_x_x_T" , "G_x_x_x","x_x_x_T",
+"x_4_13_x","x_4_x_x", "x_x_13_x"
+))
+
+G   <-c("+", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
+il4 <-c("+", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
+il13<-c("+", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
+T   <-c("+", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")
+
+labels <- paste0(G, "\n", T, "\n", il4, "\n", il13)
+breaks <- levels(test$variable)
+
+test$th<-"TH1/2"
+test$th[grep("._x_x_.", test$variable)]<-"TH1"
+test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
+test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))
+
+booleanplot(test, "variable", "value", ltbi.cols)+facet_grid(~th, scales="free", space="free_x")+
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
+    labs(y="(%) CD4+ T cells",
+         title="PMA")
+```
+
+###Peptide Mtb-specific TH1 Cytokine+ Cell Phenotypes
+```{r ics phenotyping, fig.width=16, fig.height=8}
+phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
+
+melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
+test<-filter(melt, variable!="TH1_x_x_x_x")
+test$variable<-gsub("TH1_", "", test$variable)
+test$variable<-factor(test$variable, levels=c(
+"t_x_3_4" ,       
+"t_x_3_x" ,         
+"t_x_x_4" ,         
+"t_x_x_x" ,
+"t_g_3_4" ,         
+"t_g_3_x" ,         
+"t_g_x_4" ,         
+"t_g_x_x" ,
+"x_g_3_4" ,         
+"x_g_3_x" ,         
+"x_g_x_4" ,         
+"x_g_x_x" ,
+"x_x_3_4" ,          
+"x_x_3_x" ,         
+"x_x_x_4"           
+))
+
+T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
+G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
+c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
+c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
+
+labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
+breaks <- levels(test$variable)
+
+test$tf<-NA
+test$tf[grep("x_g", test$variable)]<-"GATA3"
+test$tf[grep("t_x", test$variable)]<-"T-bet"
+test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
+test$tf[grep("x_x_._.", test$variable)]<-"None"
+test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
+
+booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x",ncol=4)+
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
+    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells",
+         title="ICS Assay")
+
+```
+
+```{r, fig.width=15, fig.height=8}
+phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
+
+#graph phenotype
+melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
+test<-filter(melt, variable!="TH1_x_x_x_x")
+test$variable<-gsub("TH1_", "", test$variable)
+test$variable<-factor(test$variable, levels=c(
+"t_x_3_4" ,       
+"t_x_3_x" ,         
+"t_x_x_4" ,         
+"t_x_x_x" ,
+"t_g_3_4" ,         
+"t_g_3_x" ,         
+"t_g_x_4" ,         
+"t_g_x_x" ,
+"x_g_3_4" ,         
+"x_g_3_x" ,         
+"x_g_x_4" ,         
+"x_g_x_x" ,
+"x_x_3_4" ,          
+"x_x_3_x" ,         
+"x_x_x_4"           
+))
+
+T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
+G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
+c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
+c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
+
+labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
+breaks <- levels(test$variable)
+
+test$tf<-NA
+test$tf[grep("x_g", test$variable)]<-"GATA3"
+test$tf[grep("t_x", test$variable)]<-"T-bet"
+test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
+test$tf[grep("x_x_._.", test$variable)]<-"None"
+test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
+
+booleanplot(test, "variable", "value", tb.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")+
+    coord_cartesian(ylim=c(0,55))+
+    stat_compare_means(label="p.signif", hide.ns=TRUE, label.y=50, size=10)
+
+```
+
 ```{r og phenotype, fig.width=16, fig.height=8}
+ogphen<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
 #graph phenotype
@@
     stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
-```
-
-####Furthermore, IFNγ and TNFa are also the two dominant cytokines produced when analyzed by luminex. The concentration of all cytokines measured by luminex are equivalent between SM- and SM+ groups.
-```{r luminex, fig.width=12, fig.height=16}
-lum<-filter_ltbi(subset(luminex, luminex$Stim=="PEP"))
-names(lum)<-gsub("IL", "IL-", names(lum))
-names(lum)<-gsub("IFNg", "IFNγ", names(lum))
-names(lum)<-gsub("TNFa", "TNFα", names(lum))
-
-melt<-melt(lum, id.vars=c("TB", "SM"), measure.vars=grep("pg", names(lum)))
-melt$variable<-gsub("_pg.ml", "", melt$variable)
-melt$variable<- factor(melt$variable, levels=c("TNFα", "IFNγ", "IL-21", "IL-17A", "IL-22", "IL-10", "IL-4", "IL-5", "IL-13"))
-booleanplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+labs(y="pg/mL")
-```
-
-####However the correlation between cytokines differs between groups. Of note, the correlation between IFNγ and TNFa is stronger amongst SM+ individuals than SM- individuals.
-```{r cor plots}
-df<-lum[,3:11]
-names(lum)<-gsub("_pg.ml", "", names(lum))
-pca1 <- prcomp(df, scale = TRUE)
-pcas<-data.frame(pca1$x)
-rotations<-data.frame(pca1$rotation)
-rotations$cytokine<-row.names(rotations)
-melt<-melt(rotations, id.vars="cytokine")
-
-stims<-factor(lum$Stim)
-worms<-factor(lum$SM)
-a<-ggplot(pcas, aes(x=pcas[,1], y=pcas[,2]))+geom_point(aes(col=worms))+labs(y="PCA2", x="PCA1")
-
-neg <- cor(lum[lum$SM=="SM-",3:11])
-order <- corrMatOrder(neg, order = "hclust", hclust.method="centroid")
-order.neg <- neg[order, order]
-res1 <- cor.mtest(lum[lum$SM=="SM-",3:11], conf.level = .95)
-pvals <- res1$p[order, order]
-corrplot(order.neg, method = "circle", insig = "label_sig", tl.col = "black", 
-         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = 2, pch.col = "white",
-         type="upper")
-
-pos <- cor(lum[lum$SM=="SM+",3:11])
-order.pos <- pos[order, order]
-res1 <- cor.mtest(lum[lum$SM=="SM+",3:11], conf.level = .95)
-pvals <- res1$p[order, order]
-corrplot(order.pos, method = "circle", insig = "label_sig", tl.col = "black", 
-         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = 2, pch.col = "white",
-         type="upper")
-```
-
-#Supplementary
-
-```{r ics phenotyping, fig.width=16, fig.height=8}
-phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
-
-melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
-test<-filter(melt, variable!="TH1_x_x_x_x")
-test$variable<-gsub("TH1_", "", test$variable)
-test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
-
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
-
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
-
-test$tf<-NA
-test$tf[grep("x_g", test$variable)]<-"GATA3"
-test$tf[grep("t_x", test$variable)]<-"T-bet"
-test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
-test$tf[grep("x_x_._.", test$variable)]<-"None"
-test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
-
-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free", ncol=4)+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
-
+    labs(y="(%) Proliferating Mtb-Specific cells")
 ```

### 2019-08-28 PaperFigures.Rmd
added in lines between plots

@@
 author: "Taryn McLaughlin"
-date: "7/29/2019"
+date: "8/27/2019"
 output: html_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment=NA, warning=FALSE, echo=FALSE)
+knitr::opts_chunk$set(message = FALSE, comment=NA, warning=FALSE, echo=FALSE)
 library(dplyr)
@@
     theme_classic()+ theme(legend.position="none")+
+    theme(panel.border = element_rect(fill = NA, colour = "black"))+
     theme(text = element_text(size=20))+
@@
     theme_classic()+ 
+    theme(legend.position="none") +
+    theme(panel.border = element_rect(fill = NA, colour = "black"))+
     facet_wrap(~variable)+
@@
 phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))
-og<-plot_filter(subset(OG, OG$Stim=="SEB"))
-ogphen<-plot_filter(subset(OGPHEN, OGPHEN$Stim=="SEB"))
 ```
@@
 tf<-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, ncol=2)+
-    theme(legend.position="none") +
     labs(y="(%) Total CD4+ T cells")
@@
 
-sm<-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, ncol=2)+
-    theme(legend.position="none")
+sm<-allplot(melt, "TB", "value", all.cols)
 
@@
 one<-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, scales="free", labeller = cyt_groups, ncol=2)+
-    labs(y="(%) Total CD4+ T cells")+
-    theme(legend.position="none")
+    facet_wrap(~variable, labeller = cyt_groups, ncol=2)+
+    labs(y="(%) Total CD4+ T cells")
 
-melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4", "IL13"))
 
-four<-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, scales="free", labeller = cyt_groups, ncol=1)+
-    theme(legend.position="none")+
-    coord_cartesian(ylim=c(0,7))
+two<-allplot(melt, "TB", "value", all.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups, ncol=2)+
+    coord_cartesian(ylim=c(0,8))
 
-melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL13"))
-
-thirteen<-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, scales="free", labeller = cyt_groups, ncol=1)+
-    theme(legend.position="none")+
-    coord_cartesian(ylim=c(0,3))
-
-cowplot::plot_grid(one, four, thirteen, ncol=3, rel_widths=c(7.5,3.75, 3.75))
+cowplot::plot_grid(one, two)
 ```
@@
 ```{r cytokine, fig.width=15, fig.height=8}
-melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
-
-a<-baseplot(melt[melt$variable=="IFNg",], "SM", "value", ltbi.cols)+    
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
+a<-baseplot(melt, "SM", "value", ltbi.cols)+    
     facet_wrap(~variable, labeller = cyt_groups)+
@@
 
-b<-baseplot(melt[melt$variable=="TNFa",], "SM", "value", ltbi.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y = (.95*.15))+
-    coord_cartesian(ylim=c(0,.15))
-
-c<-baseplot(melt[melt$variable=="IL4",], "SM", "value", ltbi.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y = (.95*.013))+
-    coord_cartesian(ylim=c(0,.013))
-
-d<-baseplot(melt[melt$variable=="IL13",], "SM", "value", ltbi.cols)+    
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4", "IL13"))
+b<-baseplot(melt, "SM", "value", ltbi.cols)+    
     facet_wrap(~variable, labeller = cyt_groups)+
@@
 
-cowplot::plot_grid(a, b, c, d,ncol=4)
+cowplot::plot_grid(a, b)
 ```
@@
 ```{r, fig.width=15, fig.height=8}
-melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
+a<-baseplot(melt, "disease", "value", tb.cols)+    
+    facet_wrap(~variable, labeller = cyt_groups)+
+    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y=(.95*.15))+
+    coord_cartesian(ylim=c(0,.15))+
+    labs(y="(%) CD4+ T cells")
 
-a<-baseplot(melt[melt$variable=="IFNg",], "disease", "value", tb.cols)+    
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4", "IL13"))
+b<-baseplot(melt, "disease", "value", tb.cols)+ 
     facet_wrap(~variable, labeller = cyt_groups)+
-    labs(y="(%) CD4+ T cells")+
-    coord_cartesian(ylim=c(0,.06))
-
-b<-baseplot(melt[melt$variable=="TNFa",], "disease", "value", tb.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)+
-    coord_cartesian(ylim=c(0,.15))
-
-c<-baseplot(melt[melt$variable=="IL4",], "disease", "value", tb.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)+
+    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y=(.95*.02))+
     coord_cartesian(ylim=c(0,.02))
 
-d<-baseplot(melt[melt$variable=="IL13",], "disease", "value", tb.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups)+
-    coord_cartesian(ylim=c(0,.02))
-
-cowplot::plot_grid(a, b, c, d,ncol=4)
+cowplot::plot_grid(a, b)
 ```
@@
 *Honestly I will have to just remake the whole thing a couple times and stitch them together to fix the axes so just ignore them for now*
+
 ```{r luminex, fig.width=12, fig.height=16}
@@
     labs(y="pg/mL")
+
+booleanplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+
+    labs(y="pg/mL")+
+    coord_cartesian(ylim=c(0,40))
 ```

### 2019-08-29 PaperFigures.Rmd
added boolean cytokine for supp

@@
     theme(legend.position="none")+
-    theme(text = element_text(size=20))+
+    theme(text = element_text(size=20), axis.text.x = element_text(size=20))+
     theme(plot.subtitle = element_text(hjust = 0.5)) +
-    stat_compare_means(label="p.signif", hide.ns = TRUE, size=10)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
     labs(x="", y="")
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y = (.95*.15))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y = (.95*.15))+
     coord_cartesian(ylim=c(0,.15))+
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y = (.95*.02))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y = (.95*.02))+
     coord_cartesian(ylim=c(0,.02))
@@
     facet_wrap(~variable)+
-    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y=(.98*70))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.98*70))+
     coord_cartesian(ylim=c(0,70))+
@@
     facet_wrap(~variable)+
-    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y=(.98*100))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.98*100))+
     coord_cartesian(ylim=c(0,100))
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y=(.95*.15))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.95*.15))+
     coord_cartesian(ylim=c(0,.15))+
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(label="p.format", size=6, label.x=1.3, label.y=(.95*.02))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.95*.02))+
     coord_cartesian(ylim=c(0,.02))
@@
     facet_wrap(~variable)+
-    stat_compare_means(label="p.format", size=6, label.y=(.98*80))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.y=(.98*80))+
     coord_cartesian(ylim=c(0,80))+
@@
     facet_wrap(~variable)+
-    stat_compare_means(label="p.format", size=6, label.y=(.98*80))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.y=(.98*80))+
     coord_cartesian(ylim=c(0,80))
@@
 ```{r receptor on gata3, fig.width=7.5, fig.height=8}
-melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("TH1_GATA3CCR4", "TH1_GATA3CXCR3"))
-melt$variable<-gsub("TH1_", "", melt$variable)
-melt$variable<-gsub("GATA3", "GATA3+", melt$variable)
-melt$variable<-gsub("CXCR3", "CXCR3+", melt$variable)
-melt$variable<-gsub("CCR4", "CCR4+", melt$variable)
+phen$TH1_GATA4notCCR4<-rowSums(phen[,grep("TH1_._g_._x", names(phen))])
+melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("TH1_GATA3CCR4", "TH1_GATA4notCCR4"))
+
+groups <- as_labeller(c("TH1_GATA3CCR4" = "GATA3+CCR4+", 
+                    "TH1_GATA4notCCR4" = "GATA3+CCR4-"))
 
 baseplot(melt, "SM", "value", tb.cols)+    
-    facet_wrap(~variable)+
-    stat_compare_means(label="p.format", size=6, label.y=(.98*15))+
-    coord_cartesian(ylim=c(0,15))+
+    facet_wrap(~variable, labeller = groups)+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3)+
     labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
@@
 ```{r, proliferation and total cyt, fig.width=15, fig.height=8}
-melt<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars=c("OG_freq", "IFNg_freq", "TNFa_freq", "IL4_freq"))
-
-groups <- as_labeller(c("IFNg_freq" = "IFNγ", 
-                    "TNFa_freq" = "TNFα", "IL4_freq" = "IL-4", "OG_freq" = "Proliferation"))
-
-allplot(melt, "TB", "value", all.cols)+
-    facet_wrap(~variable, scales="free", labeller = groups, ncol=4)+
-    labs(y="(%) Total CD4+ T cells")+
-    theme(legend.position="bottom") 
-```
-
-```{r, cyt from proliferating, fig.width=12, fig.height=8}
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg_freq", "TNFa_freq", "IL4_freq"))
-
-groups <- as_labeller(c("IFNg_freq" = "IFNγ", 
+groups <- as_labeller(c("OG_freq" = "Proliferation","IFNg_freq" = "IFNγ", 
                     "TNFa_freq" = "TNFα", "IL4_freq" = "IL-4"))
 
-allplot(melt, "TB", "value", all.cols)+
-    facet_wrap(~variable, scales="free", labeller = groups, ncol=3)+
-    labs(y="(%) Proliferating CD4+ T cells")+
-    theme(legend.position="bottom") 
+part1<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG_freq")
+og<-allplot(part1, "TB", "value", all.cols)+
+    facet_wrap(~variable, scales="free", labeller = groups)+
+    labs(y="(%) Total CD4+ T cells")
+
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg_freq", "TNFa_freq"))
+
+one<-allplot(melt, "TB", "value", all.cols)+    
+    facet_wrap(~variable, labeller = groups, ncol=2)+
+    labs(y="(%) Proliferating CD4+ T cells")
+
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4_freq"))
+
+two<-allplot(melt, "TB", "value", all.cols)+    
+    facet_wrap(~variable, labeller = groups)
+
+cowplot::plot_grid(og, one, two, ncol=3, rel_widths = c(1.2,2,1))
 ```
@@
 ```{r, marker on proliferating, fig.width=15, fig.height=8}
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet_freq", "GATA3_freq", "CXCR3_freq", "CCR4_freq"))
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))
 
-melt$variable<-gsub("_freq", "", melt$variable)
-melt$variable<-factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))
+tf<-allplot(melt, "TB", "value", all.cols)+    
+    labs(y="(%) Proliferating CD4+ T cells")
 
-allplot(melt, "TB", "value", all.cols)+
-    facet_wrap(~variable, scales="free", ncol=4)+
-    labs(y="(%) Total CD4+ T cells")+
-    theme(legend.position="bottom") 
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))
+
+sm<-allplot(melt, "TB", "value", all.cols)
+
+cowplot::plot_grid(tf, sm)
 ```
@@
 ```{r og and cyt, fig.width=15, fig.height=8}
-a<-baseplot(og, "SM", "OG_freq", ltbi.cols)+
-    labs(y="(%) CD4+ T cells")+
-    facet_wrap(~TB, labeller = as_labeller(c("LTBI"="Proliferation")))+
-    coord_cartesian(ylim=c(0,3))
+groups <- as_labeller(c("OG_freq" = "Proliferation","IFNg_freq" = "IFNγ", 
+                    "TNFa_freq" = "TNFα", "IL4_freq" = "IL-4"))
 
-melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._.", names(ogphen)))
-test<-melt[-grep("._._._.", melt$variable),]
-test<-filter(test, variable!="x_x_x")
-test$variable<-factor(test$variable, levels=c(
-"T_G_x" ,       
-"T_x_x" ,         
-"x_G_x" ,         
-"T_G_4" ,
-"T_x_4" ,         
-"x_G_4" ,         
-"x_x_4"          
-))
-test$group<-"TH1/2"
-for(i in 1:length(test$group)){
-    if(test$variable[i] %in% c("T_G_x", "T_x_x", "x_G_x")){
-        test$group[i]<- "TH1"
-    }
-    else if (test$variable[i] == "x_x_4"){
-        test$group[i]<- "TH2"
-    }
-}
+melt<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG_freq")
+og<-baseplot(melt, "TB", "value", ltbi.cols)+
+    facet_wrap(~variable, scales="free", labeller = groups)+
+    coord_cartesian(ylim=c(0,3))+
+    labs(y="(%) Total CD4+ T cells")
 
-T <- c("+", "+", "-", "+", "+", "-", "-")
-G <- c("+", "-", "+", "+", "-", "+", "-")
-il4 <- c("-", "-", "-", "+", "+", "+", "+")
-
-labels <- paste0(T, "\n", G, "\n", il4)
-breaks <- levels(test$variable)
-
-b<-booleanplot(test, "variable", "value", ltbi.cols)+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    facet_grid(~group, scales="free_x", space="free_x")+
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg_freq", "TNFa_freq"))
+one<-baseplot(melt, "TB", "value", ltbi.cols)+    
+    facet_wrap(~variable, labeller = groups, ncol=2)+
     labs(y="(%) Proliferating CD4+ T cells")
 
-cowplot::plot_grid(a, b, ncol=2, rel_widths=c(1.25,3.75), align="h")
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4_freq"))
+two<-baseplot(melt, "TB", "value", ltbi.cols)+    
+    facet_wrap(~variable, labeller = groups)
+
+cowplot::plot_grid(og, one, two, ncol=3, rel_widths = c(1.2,2,1))
 ```
@@
     facet_wrap(~variable)+
-    stat_compare_means(label="p.format", size=6, label.y=(.98*100))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.98*100))+
     coord_cartesian(ylim=c(0,100))+
@@
     facet_wrap(~variable)+
-    stat_compare_means(label="p.format", size=6, label.y=(.98*100))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.98*100))+
     coord_cartesian(ylim=c(0,100))
@@
 
-```{r luminex, fig.width=12, fig.height=16}
+```{r luminex, fig.width=12, fig.height=32}
 lum<-filter_ltbi(subset(luminex, luminex$Stim=="PEP"))
@@
 
-booleanplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+
+a<-baseplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+
     labs(y="pg/mL")
 
-booleanplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+
+b<-baseplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+
     labs(y="pg/mL")+
     coord_cartesian(ylim=c(0,40))
+
+cowplot::plot_grid(a,b, ncol=1, align = "v")
 ```
@@
 
-#Supplementary
-###Peptide-Specific Cytokine Production  
-```{r, fig.width=15, fig.height=8}
-melt<-melt(subset(CYT, CYT$Stim=="PEP"), 
-           id.vars=c("TB","SM"), 
-           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
-th_groups <- as_labeller(c("TH1.Freq" = "TH1", 
-                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))
-
-cyt<-filter_hc(melt)
-a<-baseplot(cyt, "SM", "value", healthy.cols)+
-    facet_wrap(~variable, ncol=4, labeller=th_groups)+
-    labs(y="(%) CD4+ T cells")+coord_cartesian(ylim=c(0,.2))
-
-cyt<-filter_ltbi(melt)
-b<-baseplot(cyt, "SM", "value", ltbi.cols)+
-    facet_wrap(~variable, ncol=4, labeller=th_groups)+
-    coord_cartesian(ylim=c(0,.2))
-
-cyt<-filter_tb(melt)
-c<-baseplot(cyt, "SM", "value", tb.cols)+
-    facet_wrap(~variable, ncol=4, labeller=th_groups)+
-    coord_cartesian(ylim=c(0,.2))
-
-cowplot::plot_grid(a, b, c, ncol=3)
+##ICS Supplementary
+```{r}
+booleanplot2<-function(data, xval, yval, colors){
+    ggplot(data, aes(data[,xval], data[,yval], fill=disease))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
+    scale_fill_manual(values = colors)+
+    theme_classic()+ 
+    theme(legend.position="none")+
+    theme(text = element_text(size=20), axis.text.x = element_text(size=20))+
+    theme(plot.subtitle = element_text(hjust = 0.5)) +
+    labs(x="", y="")
+}
 ```
 
-##Some Boolean Cytokine graphs will go here for posterity
+###Boolean PMA
 *but I need to make them pretty and I dont' feel like doing that right now*
-```{r}
-cyt<-filter_ltbi(CYT[CYT$Stim=="PMA",])
-test<-melt(cyt, id.vars=c("TB", "SM"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
+```{r boolean pma cyt, fig.width=15, fig.height=12}
+cyt<-plot_filter(CYT[CYT$Stim=="PMA",])
+test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
+test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
 test$variable<-factor(test$variable, levels=c(
@@
 
-booleanplot(test, "variable", "value", ltbi.cols)+facet_grid(~th, scales="free", space="free_x")+
+th1.2<-booleanplot2(test[test$th=="TH1/2",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
     scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
+    coord_cartesian(ylim=c(0,3))+
+    theme(strip.text.y = element_blank(), strip.background.y = element_blank())+
+    labs(y="(%) CD4+ T cells")
+
+th1<-booleanplot2(test[test$th=="TH1",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
+    coord_cartesian(ylim=c(0,25))+
+    theme(strip.text.y = element_blank(), strip.background.y = element_blank())
+
+th2<-booleanplot2(test[test$th=="TH2",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
+    coord_cartesian(ylim=c(0,3))
+
+cowplot::plot_grid(th1.2, th1, th2, ncol=3, rel_widths = c(8,3.5,3.5))
+```
+
+###Boolean Phenotypes
+###TH1
+```{r pma th1 boolean, fig.width=20, fig.height=18}
+phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))
+
+melt<-melt(phen, id.vars=c("TB", "SM", "disease"), measure.vars = grep("TH1_._._._.", names(phen)))
+test<-filter(melt, variable!="TH1_x_x_x_x")
+test$variable<-gsub("TH1_", "", test$variable)
+test$variable<-factor(test$variable, levels=c(
+"t_x_3_4" ,       
+"t_x_3_x" ,         
+"t_x_x_4" ,         
+"t_x_x_x" ,
+"t_g_3_4" ,         
+"t_g_3_x" ,         
+"t_g_x_4" ,         
+"t_g_x_x" ,
+"x_g_3_4" ,         
+"x_g_3_x" ,         
+"x_g_x_4" ,         
+"x_g_x_x" ,
+"x_x_3_4" ,          
+"x_x_3_x" ,         
+"x_x_x_4"           
+))
+
+T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
+G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
+c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
+c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
+
+labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
+breaks <- levels(test$variable)
+
+test$tf<-NA
+test$tf[grep("x_g", test$variable)]<-"GATA3"
+test$tf[grep("t_x", test$variable)]<-"T-bet"
+test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
+test$tf[grep("x_x_._.", test$variable)]<-"None"
+test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
+
+booleanplot2(test, "variable", "value", all.cols)+
+    facet_grid(TB~tf, scales="free_x")+
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
+    coord_cartesian(ylim=c(0,45))
+
+```
+###TH2  
+```{r pma th2 boolean, fig.width=20, fig.height=18}
+phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))
+
+melt<-melt(phen, id.vars=c("TB", "SM", "disease"), measure.vars = grep("TH2_._._._.", names(phen)))
+test<-filter(melt, variable!="TH2_x_x_x_x")
+test$variable<-gsub("TH2_", "", test$variable)
+test$variable<-factor(test$variable, levels=c(
+"t_x_3_4" ,       
+"t_x_3_x" ,         
+"t_x_x_4" ,         
+"t_x_x_x" ,
+"t_g_3_4" ,         
+"t_g_3_x" ,         
+"t_g_x_4" ,         
+"t_g_x_x" ,
+"x_g_3_4" ,         
+"x_g_3_x" ,         
+"x_g_x_4" ,         
+"x_g_x_x" ,
+"x_x_3_4" ,          
+"x_x_3_x" ,         
+"x_x_x_4"           
+))
+
+T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
+G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
+c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
+c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
+
+labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
+breaks <- levels(test$variable)
+
+test$tf<-NA
+test$tf[grep("x_g", test$variable)]<-"GATA3"
+test$tf[grep("t_x", test$variable)]<-"T-bet"
+test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
+test$tf[grep("x_x_._.", test$variable)]<-"None"
+test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
+
+booleanplot2(test, "variable", "value", all.cols)+
+    facet_grid(TB~tf, scales="free_x")+
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
+    coord_cartesian(ylim=c(0,45))
+
+```
+
+###Peptide-Specific Cytokine Production  
+```{r, fig.width=15, fig.height=8}
+melt<-melt(subset(CYT, CYT$Stim=="PEP"), 
+           id.vars=c("TB","SM"), 
+           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
+th_groups <- as_labeller(c("TH1.Freq" = "TH1", 
+                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))
+
+cyt<-filter_ltbi(melt)
+b<-baseplot(cyt, "SM", "value", ltbi.cols)+
+    facet_wrap(~variable, ncol=4, labeller=th_groups)+
+    coord_cartesian(ylim=c(0,.2))
+
+cyt<-filter_tb(melt)
+c<-baseplot(cyt, "SM", "value", tb.cols)+
+    facet_wrap(~variable, ncol=4, labeller=th_groups)+
+    coord_cartesian(ylim=c(0,.2))
+
+cowplot::plot_grid(b, c, ncol=3)
+```
+
+```{r, fig.width=16, fig.height=8}
+cyt<-plot_filter(CYT[CYT$Stim=="PEP",])
+cyt<-filter(cyt, TB!="HC")
+test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
+test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
+test$variable<-factor(test$variable, levels=c(
+"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
+"G_x_x_T" , "G_x_x_x","x_x_x_T",
+"x_4_13_x","x_4_x_x", "x_x_13_x"
+))
+
+G   <-c("+", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
+il4 <-c("+", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
+il13<-c("+", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
+T   <-c("+", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")
+
+labels <- paste0(G, "\n", T, "\n", il4, "\n", il13)
+breaks <- levels(test$variable)
+
+test$th<-"TH1/2"
+test$th[grep("._x_x_.", test$variable)]<-"TH1"
+test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
+test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))
+
+booleanplot2(test, "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
+    coord_cartesian(ylim=c(0,.2))+
     labs(y="(%) CD4+ T cells",
@@
     scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
     labs(y="(%) Mtb-Specific TH1 Cytokine+ cells",
@@
     coord_cartesian(ylim=c(0,55))+
-    stat_compare_means(label="p.signif", hide.ns=TRUE, label.y=50, size=10)
+    stat_compare_means(method="wilcox.test",label="p.signif", hide.ns=TRUE, label.y=50, size=10)
+
+```
+
+##Proliferation Supplementary
+```{r}
+ogphen<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
+melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._.", names(ogphen)))
+test<-melt[-grep("._._._.", melt$variable),]
+test<-filter(test, variable!="x_x_x")
+test$variable<-factor(test$variable, levels=c(
+"T_G_x" ,       
+"T_x_x" ,         
+"x_G_x" ,         
+"T_G_4" ,
+"T_x_4" ,         
+"x_G_4" ,         
+"x_x_4"          
+))
+test$group<-"TH1/2"
+for(i in 1:length(test$group)){
+    if(test$variable[i] %in% c("T_G_x", "T_x_x", "x_G_x")){
+        test$group[i]<- "TH1"
+    }
+    else if (test$variable[i] == "x_x_4"){
+        test$group[i]<- "TH2"
+    }
+}
+
+T <- c("+", "+", "-", "+", "+", "-", "-")
+G <- c("+", "-", "+", "+", "-", "+", "-")
+il4 <- c("-", "-", "-", "+", "+", "+", "+")
+
+labels <- paste0(T, "\n", G, "\n", il4)
+breaks <- levels(test$variable)
+
+booleanplot(test, "variable", "value", ltbi.cols)+
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    facet_grid(~group, scales="free_x", space="free_x")+
+    labs(y="(%) Proliferating CD4+ T cells")
 
@@
 ```{r og phenotype, fig.width=16, fig.height=8}
-ogphen<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
 #graph phenotype
@@
     scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(label="p.signif", hide.ns = TRUE, size=6)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=6)+
     labs(y="(%) Proliferating Mtb-Specific cells")

### 2019-08-29 gamma-delta.Rmd
initial commit

---
title: "gamma-delta-figures"
author: "Taryn McLaughlin"
date: "8/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment=NA, warning=FALSE, echo=FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
library(tidyr)
library(reshape2)
library(cowplot)
library(corrplot)
library(sjPlot)
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")
```

```{r}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
PBMC<-dplyr::select(read.csv("DonorInfo_clean.csv"), -X)
    PBMC<-PBMC[grep("OG", PBMC$used),]
    plot_3(PBMC[PBMC$SM=="SM",], "SchistosomaIntensity")+
    labs(y="Eggs/gram")
```

#ICS Assay

```{r, fig.width=12, fig.height=8}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
compass<-read.csv("ICS_compass_complete_clean.csv")
compass$TB<-factor(compass$TB, levels=c("N", "HC", "LTBI", "TB"))

include<-subset(compass, compass$Stim %in% c("SEA", "SWAP"))
plot_cell(include, "FS")+
    facet_grid(Stim~TB)+
    scale_x_discrete(labels = c("CD4", "CD8", "CD4-CD8-"))+
    labs(y="Functionality Score",
         title="Includes SM-")

exclude<-subset(include, include$SM!="X")
plot_cell(exclude, "FS")+
    facet_grid(Stim~TB)+
    scale_x_discrete(labels = c("CD4", "CD8", "CD4-CD8-"))+
    labs(y="Functionality Score",
         title="Excludes SM-")
```

#Proliferation Assay
```{r load og data}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
gd<-dplyr::select(read.csv("OG_gd_bulk_summary_clean.csv"), -X, -X.1)
gd_OG<-dplyr::select(read.csv("OG_gd_OG_summary_clean_NAs.csv"), -X)
GD<-dplyr::select(read.csv("OG_GDonly_bulk_summary_clean.csv"), -X, -X.1)
GD_OG<-dplyr::select(read.csv("OG_GDonly_OG_summary_clean_NAs.csv"), -X)
CD3<-dplyr::select(read.csv("OG_CD3_bulk_summary_clean.csv"), -X, -X.1)
CD3_OG<-dplyr::select(read.csv("OG_CD3_OG_summary_clean.csv"), -X, -X.1)
mimosa<-dplyr::select(read.csv("OG_MIMOSA_clean.csv"), -X)
    mimosa<-dplyr::filter(mimosa, celltype=="GD" & Exprate==0.01)
```

###*I was conflicted so all the plots on the left will include the SM- donors, whereas the plots on the right will exclude them.  I still have to export the proliferation data for the yerkes donors but that's a good chunk of time/effort so I'm waiting until the other paper is further along*

###Starting Frequency of Gamma Delta Cells
```{r}
include<-subset(CD3, CD3$Stim=="UN")
a<-plot_3(include, "GD_freq")+
    coord_cartesian(ylim=c(0,35))+
    labs(y="(%) CD3+ T cells",
         title="Includes SM-")

exclude<-subset(include, include$SM=="SM")
b<-plot_3(exclude, "GD_freq")+
    coord_cartesian(ylim=c(0,35))+
    labs(y="(%) CD3+ T cells",
         title="Excludes SM-")

cowplot::plot_grid(a,b)
```

###Proliferation of Gamma Delta Cells
```{r}
include<-subset(gd, gd$Stim=="SWAP")
a<-plot_3(include, "OG_freq")+
    labs(y="(%) γδ T cells")

exclude<-subset(include, include$SM=="SM")
b<-plot_3(exclude, "OG_freq")+
    labs(y="(%) γδ T cells")

cowplot::plot_grid(a,b)
```

###Proportion of proliferating Gamma Delta Cells
```{r, fig.width=12, fig.height=6}
include<-subset(CD3_OG, CD3_OG$Stim=="SWAP")
melt<-melt(include, id.vars=c("TB", "SM"), measure.vars=c("CD4_freq", "CD8_freq", "GD_freq"))
melt$celltype<-gsub("_freq", "",melt$variable)

a<-plot_3(melt, "value")+
    facet_grid(~variable, labeller = as_labeller(c("CD4_freq"="CD4", "CD8_freq"="CD8", "GD_freq"="γδ")))+
    #coord_cartesian(ylim=c(0,65))+
    labs(y="(%) Proliferating T cells",
         title="Includes SM-")

exclude<-subset(melt, melt$SM=="SM")
b<-plot_3(melt, "value")+
    facet_grid(~variable, labeller = as_labeller(c("CD4_freq"="CD4", "CD8_freq"="CD8", "GD_freq"="γδ")))+
    labs(y="(%) Proliferating T cells",
         title="Excludes SM-")

cowplot::plot_grid(a,b)
```

```{r, fig.width=12, fig.height=6}
include<-subset(CD3_OG, CD3_OG$Stim=="SWAP")
melt<-melt(include, id.vars=c("TB", "SM"), measure.vars=c("CD4_freq", "CD8_freq", "GD_freq"))
melt$celltype<-gsub("_freq", "",melt$variable)

a<-plot_cell(melt, "value")+
    facet_grid(~TB)+
    scale_x_discrete(labels = c("CD4", "CD8", "γδ"))+
    labs(y="(%) Proliferating+ T cells",
         title="Includes SM-")

exclude<-subset(melt, melt$SM=="SM")
b<-plot_cell(melt, "value")+
    facet_grid(~TB)+
    scale_x_discrete(labels = c("CD4", "CD8", "γδ"))+
    labs(y="(%) Proliferating+ T cells",
         title="Excludes SM-")

cowplot::plot_grid(a,b)
```

###IL-4
```{r}
include<-subset(gd_OG, gd_OG$Stim=="SWAP")
a<-plot_3(include, "IL4_freq")+
    labs(y="(%) Proliferating γδ T cells")

exclude<-subset(include, include$SM=="SM")
b<-plot_3(exclude, "IL4_freq")+
    labs(y="(%) Proliferating γδ T cells")

cowplot::plot_grid(a,b)
```

###TNFa
```{r}
include<-subset(gd_OG, gd_OG$Stim=="SWAP")
a<-plot_3(include, "TNFa_freq")+
    labs(y="(%) Proliferating γδ T cells")

exclude<-subset(include, include$SM=="SM")
b<-plot_3(exclude, "TNFa_freq")+
    labs(y="(%) Proliferating γδ T cells")

cowplot::plot_grid(a,b)
```

###All Cytokine Combinations
```{r}
test<-subset(gd_OG, gd_OG$Stim=="SWAP")
vars<-names(test)[grep("._._.", names(test))]
vars<-vars[-grep("._._._.", vars)]

include<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
include<-filter(include, variable!="x_x_x")
include$variable<-factor(include$variable, levels=c(
"T_G_x" ,       
"T_x_x" ,         
"x_G_x" ,         
"T_G_4" ,
"T_x_4" ,         
"x_G_4" ,         
"x_x_4"          
))

T <- c("+", "+", "-", "+", "+", "-", "-")
G <- c("+", "-", "+", "+", "-", "+", "-")
il4 <- c("-", "-", "-", "+", "+", "+", "+")

labels <- paste0(T, "\n", G, "\n", il4)
breaks <- levels(include$variable)

ggplot(include, aes(x=variable, y=value, fill=TB))+
    geom_boxplot(outlier.shape = NA)+
    scale_fill_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))+
    scale_x_discrete(labels=labels, breaks=breaks)+
    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(title="Includes SM-",
         y="(%) Proliferating γδ T cells")

exclude<-subset(include, include$SM=="SM")
ggplot(exclude, aes(x=variable, y=value, fill=TB))+
    geom_boxplot(outlier.shape = NA)+
    scale_fill_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))+
    scale_x_discrete(labels=labels, breaks=breaks)+
    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(title="Excludes SM-",
         y="(%) Proliferating γδ T cells")
```

### 2019-09-12 ICS_Cleaning.Rmd
fix the TH2 definition because it wasn't strict enough

@@
 knitr::opts_chunk$set(echo = TRUE, message = FALSE )
+
 ```
@@
       TH1_GATA3CCR4=  rowSums(CD4_phenotype[,grep("TH1_._g_._4", names(CD4_phenotype))]),
+            TH1_GATA3notCCR4=  rowSums(CD4_phenotype[,grep("TH1_._g_._x", names(CD4_phenotype))]),
       TH2_TbetCXCR3=  rowSums(CD4_phenotype[,grep("TH2_t_._3_.", names(CD4_phenotype))]),

### 2019-09-12 ICS_Phenotyping.Rmd
fix the TH2 definition because it wasn't strict enough

@@
          y="Frequency of Marker+ Mtb-Specific T cells")+
-    coord_cartesian(ylim=c(0,10))
+    coord_cartesian(ylim=c(0,20))
 kable(xtabs(~TB+SM, test))

### 2019-09-12 OG-Analysis.Rmd
removed _freq from names

@@
 library(knitr)
-DF<-mutate_hc(subset(CD4, CD4$Stim=="SEB"))
-plot_hc(DF, "SM","OG_freq")
+DF<-mutate_hc(dplyr::filter(CD4, Stim=="SEB"))
+plot_hc(DF, "SM","OG")
 ```
@@
 library(knitr)
-DF<-subset(CD4, CD4$Stim=="PEP")
+DF<-dplyr::filter(CD4, Stim=="PEP")
 colnames<-numeric_only(DF)
@@
 ```{r}
-test<-mutate_TB(filter(DF, !is.na(OG_freq)))
+test<-mutate_TB(filter(DF, !is.na(OG)))
 kable(xtabs(~TB+SM, test))
-plot_2(test, "disease", "OG_freq")
+plot_2(test, "disease", "OG")
 ```
@@
 library(knitr)
-DF<-subset(OG, OG$Stim=="SEB")
+DF<-dplyr::filter(OG, Stim=="SEB")
 #split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
@@
 #this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
-colnames<-c(test[1:31], test[38:49])
+colnames<-c(test[1:31], test[38:45])
 for(col in colnames){
@@
 ```{r}
-test<-mutate_TB(filter(DF, !is.na(OG_freq)))
+test<-mutate_TB(filter(DF, !is.na(OG)))
 kable(xtabs(~TB+SM, test))
@@
 library(knitr)
-DF<-subset(OG, OG$Stim=="PEP")
+DF<-dplyr::filter(OG, Stim=="PEP")
 #split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
@@
 library(knitr)
-DF<-subset(OG, OG$Stim=="WCL")
+DF<-dplyr::filter(OG, Stim=="WCL")
 #split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
@@
 ```{r}
-plot_2(test, "disease", "IL4_freq")
+plot_2(test, "disease", "IL4")
 plot_2(test, "disease", "T_x_4")
@@
 library(knitr)
-DF<-subset(TH1, TH1$Stim=="PEP")
+DF<-dplyr::filter(TH1, Stim=="PEP")
 #split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
@@
 library(knitr)
-DF<-subset(TNF, TNF$Stim=="PEP")
+DF<-dplyr::filter(TNF, Stim=="PEP")
 #split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
@@
 library(knitr)
-DF<-subset(GT, GT$Stim=="PEP")
+DF<-dplyr::filter(GT, GT$Stim=="PEP")
 #split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest
@@
 library(knitr)
-DF<-subset(IFNg, IFNg$Stim=="PEP")
+DF<-dplyr::filter(IFNg, Stim=="PEP")
 #split by both TB and SM so you have a list of 6 data frames- each of which is a group of interest

### 2019-09-12 OG_cleaning.Rmd
removed _freq from names

@@
 l<-lapply(l, function(x) changeTB(x))
+l<-lapply(l, function(x) cleanvariablenames(x))
 
 filenames<-ls(pattern="summary.csv")
-for(i in 1:15){
+for(i in 1:16){
 assign(filenames[i], data.frame(l[i]))}
@@
 l<-lapply(l, function(x) sum_boolean(x))
+l<-lapply(l, function(x) sum_boolean2(x))
 filenames<-ls(pattern="bulk_summary")
@@
 l<-lapply(l, function(x) sum_boolean(x))
+l<-lapply(l, function(x) sum_boolean2(x))
 filenames<-ls(pattern="_OG")
@@
 l<-lapply(ls(pattern="bulk_summary"), function(x) get(x))
-l<-lapply(l, function(x) OG_background.subtract(x, "Donor", "Stim", "UN", "OG_freq"))
+l<-lapply(l, function(x) OG_background.subtract(x, "Donor", "Stim", "UN", "OG"))
 filenames<-ls(pattern="bulk_summary")

### 2019-09-12 PaperFigures.Rmd
added in space for flow plots and letters for labeling

@@
 author: "Taryn McLaughlin"
-date: "8/27/2019"
+date: "8/28/2019"
 output: html_document
@@
     theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=20))+
-    theme(plot.subtitle = element_text(hjust = 0.5)) +
+    theme(text = element_text(size=16))+
+    theme(plot.title = element_text(hjust = 0.5)) +
     labs(x="", y="")
@@
     theme(legend.position="none")+
-    theme(text = element_text(size=20), axis.text.x = element_text(size=20))+
-    theme(plot.subtitle = element_text(hjust = 0.5)) +
+    theme(text = element_text(size=16), axis.text.x = element_text(size=16))+
+    theme(plot.title = element_text(hjust = 0.5)) +
     stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
@@
     scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
-    theme(text = element_text(size=20))+
-    theme(plot.subtitle = element_text(hjust = 0.5))+
+    theme(text = element_text(size=16))+
+    theme(plot.title = element_text(hjust = 0.5))+
     stat_compare_means(method = 'wilcox.test', label="p.signif", hide.ns=TRUE, size=10)+
@@
 PHEN<-read.csv("ICS_CD4_phenotyping_clean_NAs.csv")
-OG<-read.csv("OG_CD4_bulk_summary_clean.csv")
+OG<-data.table(read.csv("OG_CD4_bulk_summary_clean.csv"))
 OGPHEN<-read.csv("OG_CD4_OG_summary_clean_NAs.csv")
@@
 
+#Figure 1
 ```{r, fig.width=15, fig.height=8}
@@
 
-sm<-allplot(melt, "TB", "value", all.cols)
-
-cowplot::plot_grid(tf, sm)
-
-melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
-
-one<-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups, ncol=2)+
+sm<-allplot(melt, "TB", "value", all.cols) +    
     labs(y="(%) Total CD4+ T cells")
 
-melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4", "IL13"))
-
-two<-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, labeller = cyt_groups, ncol=2)+
-    coord_cartesian(ylim=c(0,8))
-
-cowplot::plot_grid(one, two)
+first_row = cowplot::plot_grid(NULL, NULL, labels = c('A' , 'B'), nrow=1, label_size = 20, rel_widths = c(3,2))
+second_row = cowplot::plot_grid(tf, sm,labels = c('C', 'D'), nrow=1, label_size = 20)
+figure1<- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,2))
 ```
@@
 ```{r pvals for ifng and tnfa}
+kable(table(cd4$disease), caption="Total CD4 Phenotypes")
+kable(table(cyt$disease), caption="Total CD4 Cytokine")
+
 y<-split(cyt,cyt$SM) 
@@
 comparison<-c("SM- IFNγ", "SM+ IFNγ", "SM- TNFα", "SM+ TNFα")
-kable(cbind(comparison, pval))
-```
-
+kable(cbind(comparison, pval), caption="Stats across Tb groups")
+```
+
+#Figure 2
 ####Generic cytokine production from CD4+ T cells stimulated with PMA & Ionomycin does not differ between SM- and SM+ Healthy Adults
 ```{r, fig.width=15, fig.height=8}
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
+
+one<-allplot(melt, "TB", "value", all.cols)+ 
+    facet_wrap(~variable, labeller = cyt_groups)+
+    labs(y="(%) Total CD4+ T cells")
+
+melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4", "IL13"))
+
+two<-allplot(melt, "TB", "value", all.cols)+
+    facet_wrap(~variable, labeller = cyt_groups)+
+    coord_cartesian(ylim=c(0,8))+
+    labs(y="(%) Total CD4+ T cells")
+
 melt<-melt(subset(CYT, CYT$Stim=="PMA"), 
@@
     facet_wrap(~variable, ncol=4, labeller=th_groups)+
-    labs(y="(%) CD4+ T cells")+coord_cartesian(ylim=c(0,45))
+    labs(y="(%) Total CD4+ T cells")+coord_cartesian(ylim=c(0,45))+
+    labs(title="Healthy Controls")
 
@@
     facet_wrap(~variable, ncol=4, labeller=th_groups)+
-    coord_cartesian(ylim=c(0,45))
+    coord_cartesian(ylim=c(0,45))+
+    labs(title="LTBI")
 
@@
     facet_wrap(~variable, ncol=4, labeller=th_groups)+
-    coord_cartesian(ylim=c(0,45))
-
-cowplot::plot_grid(a, b, c, ncol=3)
-```
-
+    coord_cartesian(ylim=c(0,45))+
+    labs(title="TB")
+
+first_row = cowplot::plot_grid(NULL, NULL, labels = c('A', 'B'), nrow=1, label_size = 20)
+second_row = cowplot::plot_grid(one, two, labels = c('C', 'D'), nrow=1, label_size = 20)
+third_row = cowplot::plot_grid(a, b, c, nrow=1, labels=c('E'), label_size = 20)
+figure2 = cowplot::plot_grid(first_row, second_row, third_row, ncol=1, rel_heights = c(1,2,2))
+```
+
+#Figure 3
 ####Generic TH1 and TH2 cytokine producing cells are stratified by lineage-specific transcription factors, and CCR4
@@
     theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    labs(y="(%) Cytokine+ CD4+ T cells")
+    labs(y="(%) Cytokine+ CD4+ T cells",
+         title="HC")
 
@@
     facet_grid(variable~subset, scales="free")+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))
+    theme(panel.border = element_rect(fill = NA, colour = "black"))+
+    labs(y="(%) Cytokine+ CD4+ T cells",
+         title="LTBI")
 
@@
     facet_grid(variable~subset, scales="free")+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))
-
-cowplot::plot_grid(a, b, c, ncol=3)
+    theme(panel.border = element_rect(fill = NA, colour = "black"))+
+    labs(y="(%) Cytokine+ CD4+ T cells",
+         title="TB")
+
+first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
+second_row = cowplot::plot_grid(a, b, c, nrow=1, labels=c('B', 'C', 'D'), label_size = 20)
+figure3 <- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,4))
 ```
@@
 #LTBI
-
+#Figure 4
 ####LTBI individuals coinfected with S. mansoni have higher Mtb peptide-specific cytokine production, specifically due to higher production of IFNγ
@@
 phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
+
+kable(table(cyt$disease), caption="Pep cytokine")
+kable(table(phen$disease[!is.na(phen$TH1_MFI_CCR4)]), caption="Pep phenotype")
 ```
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y = (.95*.15))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y = (.95*.15))+
     coord_cartesian(ylim=c(0,.15))+
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y = (.95*.02))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y = (.95*.02))+
     coord_cartesian(ylim=c(0,.02))
 
-cowplot::plot_grid(a, b)
 ```
@@
 
-a<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
+tf<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.98*70))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*70))+
     coord_cartesian(ylim=c(0,70))+
@@
 
-b<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
+sm<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.98*100))+
-    coord_cartesian(ylim=c(0,100))
-
-cowplot::plot_grid(a,b,ncol=2)
+    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
+    coord_cartesian(ylim=c(0,100))+
+    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
+
+first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
+second_row = cowplot::plot_grid(a, b, nrow=1, labels=c('B', 'C'), label_size = 20)
+third_row = cowplot::plot_grid(tf, sm, nrow=1, labels=c('D', 'E'), label_size = 20)
+figure4 <- cowplot::plot_grid(first_row, second_row, third_row, ncol=1, rel_heights = c(1,2,2))
 ```
@@
 #TB 
+#Figure 5
 ```{r, narrow down tb data}
@@
 phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
+
+kable(table(cyt$disease), caption="Pep cytokine")
+kable(table(phen$disease[!is.na(phen$TH1_MFI_CCR4)]), caption="Pep phenotype")
 ```
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.95*.15))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.95*.15))+
     coord_cartesian(ylim=c(0,.15))+
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.95*.02))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.95*.02))+
     coord_cartesian(ylim=c(0,.02))
-
-cowplot::plot_grid(a, b)
 ```
@@
 
-a<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", tb.cols)+    
+tf<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", tb.cols)+    
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.y=(.98*80))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*80))+
     coord_cartesian(ylim=c(0,80))+
@@
 
-b<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", tb.cols)+    
+sm<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", tb.cols)+    
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.y=(.98*80))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*80))+
     coord_cartesian(ylim=c(0,80))
 
-cowplot::plot_grid(a, b, ncol=2)
-```
-
-```{r receptor on gata3, fig.width=7.5, fig.height=8}
 phen$TH1_GATA4notCCR4<-rowSums(phen[,grep("TH1_._g_._x", names(phen))])
@@
 
-baseplot(melt, "SM", "value", tb.cols)+    
+both<-baseplot(melt, "SM", "value", tb.cols)+    
     facet_wrap(~variable, labeller = groups)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3)+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3)+
     labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
+
+tbcyt<-cowplot::plot_grid(NULL, a, b, nrow=1, labels=c("A","B", "C"), label_size = 20, rel_widths = c(1.5,1,1))
+tbphen<-cowplot::plot_grid(tf, sm, both, nrow=1, labels=c("D","E","F"), label_size = 20)
+figure5<-cowplot::plot_grid(tbcyt, tbphen, nrow=2, rel_heights = c(2,3))
 ```
@@
 #Proliferation Assay
+#Figure 6
 ##All Groups Responses to SEB
 ```{r}
-og<-plot_filter(subset(OG, OG$Stim=="SEB"))
-ogphen<-plot_filter(subset(OGPHEN, OGPHEN$Stim=="SEB"))
+og<-plot_filter(dplyr::filter(OG, Stim=="SEB"))
+ogphen<-plot_filter(dplyr::filter(OGPHEN, Stim=="SEB"))
+
+kable(table(og$disease), caption="Total Proliferation")
+kable(table(ogphen$disease[!is.na(ogphen$IFNg)]), caption="SEB phenotype")
 ```
@@
 ```{r, proliferation and total cyt, fig.width=15, fig.height=8}
-groups <- as_labeller(c("OG_freq" = "Proliferation","IFNg_freq" = "IFNγ", 
-                    "TNFa_freq" = "TNFα", "IL4_freq" = "IL-4"))
-
-part1<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG_freq")
+groups <- as_labeller(c("OG" = "Proliferation","IFNg" = "IFNγ", 
+                    "TNFa" = "TNFα", "IL4" = "IL-4"))
+
+part1<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG")
 og<-allplot(part1, "TB", "value", all.cols)+
@@
 
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg_freq", "TNFa_freq"))
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
 
@@
 
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4_freq"))
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))
 
@@
     facet_wrap(~variable, labeller = groups)
-
-cowplot::plot_grid(og, one, two, ncol=3, rel_widths = c(1.2,2,1))
 ```
@@
 
-cowplot::plot_grid(tf, sm)
-```
-
+first_row = cowplot::plot_grid(NULL, labels = c('A', 'B'), nrow=1, label_size = 20)
+second_row = cowplot::plot_grid(og, one, two, nrow=1, rel_widths = c(1.2,2,1), labels=c("B", "C","D"), label_size = 20)
+third_row = cowplot::plot_grid(tf, sm, labels=c("E", "F"), nrow=1, label_size = 20)
+figure6<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3)
+```
+
+#Figure 7
 ##Mtb-Specific Responses in LTBI
 ```{r}
-og<-filter_ltbi(subset(OG, OG$Stim=="PEP"))
-ogphen<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
-oggt<-filter_ltbi(subset(OG.GT, OG.GT$Stim=="PEP"))
-ogth1<-filter_ltbi(subset(OG.th1, OG.GT$Stim=="PEP"))
+og<-filter_ltbi(dplyr::filter(OG, Stim=="PEP"))
+ogphen<-filter_ltbi(dplyr::filter(OGPHEN, Stim=="PEP"))
+oggt<-filter_ltbi(dplyr::filter(OG.GT, Stim=="PEP"))
+ogth1<-filter_ltbi(dplyr::filter(OG.th1, Stim=="PEP"))
+
+kable(table(og$disease), caption="PEP Proliferation")
+kable(table(ogphen$disease[!is.na(ogphen$IFNg)]), caption="PEP phenotype")
 ```
@@
 ```{r og and cyt, fig.width=15, fig.height=8}
-groups <- as_labeller(c("OG_freq" = "Proliferation","IFNg_freq" = "IFNγ", 
-                    "TNFa_freq" = "TNFα", "IL4_freq" = "IL-4"))
-
-melt<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG_freq")
+groups <- as_labeller(c("OG" = "Proliferation","IFNg" = "IFNγ", 
+                    "TNFa" = "TNFα", "IL4" = "IL-4"))
+
+melt<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG")
 og<-baseplot(melt, "TB", "value", ltbi.cols)+
@@
 
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg_freq", "TNFa_freq"))
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
 one<-baseplot(melt, "TB", "value", ltbi.cols)+    
@@
 
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4_freq"))
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))
 two<-baseplot(melt, "TB", "value", ltbi.cols)+    
     facet_wrap(~variable, labeller = groups)
-
-cowplot::plot_grid(og, one, two, ncol=3, rel_widths = c(1.2,2,1))
 ```
@@
 
-a<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
+tf<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.98*100))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
     coord_cartesian(ylim=c(0,100))+
@@
 
-b<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
+sm<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.98*100))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
     coord_cartesian(ylim=c(0,100))
 
-cowplot::plot_grid(a,b,ncol=2)
+first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
+second_row = cowplot::plot_grid(og, one, two, nrow=1, rel_widths = c(1.2,2,1), labels=c("B", "C","D"), label_size = 20)
+third_row = cowplot::plot_grid(tf, sm, labels=c("E", "F"), nrow=1, label_size = 20)
+figure7<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3)
 ```
@@
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.98*100))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
     coord_cartesian(ylim=c(0,100))+
@@
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.98*100))+
+    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
     coord_cartesian(ylim=c(0,100))
 
-cowplot::plot_grid(a,b,ncol=2)
+ogth1<-cowplot::plot_grid(a,b,ncol=2)
 ```
@@
 
-cowplot::plot_grid(a,b, ncol=1, align = "v")
+lumplot<-cowplot::plot_grid(a,b, NULL, ncol=1, align = "v", labels=c("A","A", "B"), label_size = 20)
 ```
@@
 
-neg <- cor(lum[lum$SM=="SM-",3:11])
-order <- corrMatOrder(neg, order = "hclust", hclust.method="centroid")
+neg <- cor(lum[lum$SM=="SM-",3:11], method="spearman")
+order <- corrMatOrder(neg, order = "hclust")
 order.neg <- neg[order, order]
@@
 
-pos <- cor(lum[lum$SM=="SM+",3:11])
+pos <- cor(lum[lum$SM=="SM+",3:11], method="spearman")
 order.pos <- pos[order, order]
@@
          type="upper")
+```
+#Exporting Data
+```{r}
+setwd("/Users/tarynam/Desktop/Manuscript1")
+cowplot::save_plot("figure1.png", figure1, base_height=8, base_width=16) #Figure 1
+cowplot::save_plot("figure2.png", figure2, base_height=14, base_width=16) #Figure 2
+cowplot::save_plot("figure3.png", figure3, base_height=14, base_width=16)
+cowplot::save_plot("figure4.png", figure4, base_height=14, base_width=16)
+cowplot::save_plot("figure5.png", figure5, base_height=14, base_width=16)
+cowplot::save_plot("figure6.png", figure6, base_height=16, base_width=16)
+cowplot::save_plot("figure7.png", figure7, base_height=16, base_width=16)
+cowplot::save_plot("figure8.png", lumplot, base_height=32, base_width=16)
 ```
@@
 
-cowplot::plot_grid(th1.2, th1, th2, ncol=3, rel_widths = c(8,3.5,3.5))
+supp1<-cowplot::plot_grid(th1.2, th1, th2, ncol=3, rel_widths = c(8,3.5,3.5))
 ```
@@
 
+df<-data.frame(cbind(levels(test$variable),T,G,c3,c4))
+dfm<-melt(df, measure.vars=c("T", "G","c3","c4"))
+dfm$variable<-factor(dfm$variable, levels=c("c4","c3","G", "T"))
+
+data_table <- ggplot(dfm, aes(x = V1, y = variable, label = format(value))) +
+    geom_text(size = 10) + 
+    theme_bw() + 
+    labs(x="",y="")+
+    theme(strip.text.x = element_blank(), strip.background.x = element_blank())+
+    scale_y_discrete(labels=c("T-bet", "GATA3", "CXCR3", "CCR4"))+
+    theme(axis.text.y = element_text(size=20))+
+    theme(panel.grid.major = element_blank(), 
+          panel.grid.minor = element_blank(),
+          panel.border = element_blank(),
+          axis.title = element_blank(),
+          axis.text.x = element_blank(),
+          axis.ticks = element_blank())
+
 test$tf<-NA
@@
 
-booleanplot2(test, "variable", "value", all.cols)+
+supp2a<-booleanplot2(test, "variable", "value", all.cols)+
     facet_grid(TB~tf, scales="free_x")+
@@
     stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
-    coord_cartesian(ylim=c(0,45))
+    coord_cartesian(ylim=c(0,65))
 
@@
 
-booleanplot2(test, "variable", "value", all.cols)+
+supp2b<-booleanplot2(test, "variable", "value", all.cols)+
     facet_grid(TB~tf, scales="free_x")+
@@
     stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
-    coord_cartesian(ylim=c(0,45))
-
+    coord_cartesian(ylim=c(0,65))
+```
+
+```{r}
+supp2<-cowplot::plot_grid(supp2a, supp2b, nrow=2, labels=c("A", "B"))
 ```
@@
 
-cowplot::plot_grid(b, c, ncol=3)
+supp3<-cowplot::plot_grid(b, c, ncol=3)
 ```
@@
 cyt<-plot_filter(CYT[CYT$Stim=="PEP",])
-cyt<-filter(cyt, TB!="HC")
 test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
@@
 
-G   <-c("+", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
-il4 <-c("+", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
-il13<-c("+", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
-T   <-c("+", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")
+G   <-c("IFNγ +", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
+il4 <-c("IL-4 +", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
+il13<-c("IL-13 +", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
+T   <-c("TNFα +", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")
 
@@
 
-booleanplot2(test, "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
+supp4a<-booleanplot2(test[test$TB=="LTBI",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
     scale_x_discrete(labels = labels, breaks=breaks)+
     stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
-    coord_cartesian(ylim=c(0,.2))+
-    labs(y="(%) CD4+ T cells",
-         title="PMA")
+    coord_cartesian(ylim=c(0,.16))+
+    labs(y="(%) CD4+ T cells")
+
+supp5a<-booleanplot2(test[test$TB=="TB",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
+    scale_x_discrete(labels = labels, breaks=breaks)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
+    coord_cartesian(ylim=c(0,.15))+
+    labs(y="(%) CD4+ T cells")
 ```
@@
 
-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x",ncol=4)+
+supp4b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x",ncol=4)+
     scale_x_discrete(labels = labels, breaks=breaks)+
     stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells",
-         title="ICS Assay")
-
+    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
 ```
@@
 
-booleanplot(test, "variable", "value", tb.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
+supp5b<-booleanplot(test, "variable", "value", tb.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
     scale_x_discrete(labels = labels, breaks=breaks)+
@@
     stat_compare_means(method="wilcox.test",label="p.signif", hide.ns=TRUE, label.y=50, size=10)
-
+```
+
+###Comparing Mtb-specific TH1 cells to total CD4 T cells in peptide stim  
+```{r, fig.width=12, fig.height=6}
+total<-filter_ltbi(subset(CD4, CD4$Stim=="UN"))
+total$cell<-"Total"
+phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
+names<-c("Donor", "TB", "SM", "Stim", "disease", names(phen)[grep("TH1_", names(phen))])
+mtb_th1<-select(phen, names)
+names(mtb_th1)<-gsub("TH1_", "", names(mtb_th1))
+mtb_th1$cell<-"Mtb-Specific"
+long<-rbindlist(list(total, mtb_th1), fill=TRUE)
+
+melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
+melt$cell<-factor(melt$cell, levels=c("Total", "Mtb-Specific"))
+melt$x<-paste(melt$SM, melt$cell, sep=" ")
+melt$x<-factor(melt$x, levels=c("SM- Total", "SM- Mtb-Specific","SM+ Total", "SM+ Mtb-Specific"))
+
+ggplot(melt, aes(x=x, y=value, fill=SM))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
+    facet_wrap(~variable, scales="free", nrow=1)+
+    theme_classic()+
+    scale_fill_manual(values = ltbi.cols)+
+    scale_x_discrete(labels=c("Total", "Mtb", "Total", "Mtb"))+
+    theme(panel.border = element_rect(fill = NA, colour = "black"))+
+    theme(text = element_text(size=16))+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    stat_compare_means(comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")) 
+                       ,method="wilcox.test", label="p.format")+
+    labs(x="", y="")
+```
+
+###Comparing Mtb-specific TH1 cells to PMA TH1 cells  
+```{r, fig.width=12, fig.height=6}
+pma<-filter_ltbi(subset(PHEN, PHEN$Stim=="PMA"))
+pma$cell<-"PMA"
+pep<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
+pep$cell<-"Mtb"
+long<-rbindlist(list(pma, pep), fill=TRUE)
+names<-c("Donor", "TB", "SM", "Stim", "disease", "cell", names(long)[grep("TH1_", names(long))])
+long<-select(long, names)
+names(long)<-gsub("TH1_", "", names(long))
+
+melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
+melt$x<-paste(melt$SM, melt$cell, sep=" ")
+melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))
+
+ggplot(melt, aes(x=x, y=value, fill=SM))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
+    facet_wrap(~variable, scales="free", nrow=1)+
+    theme_classic()+
+    scale_fill_manual(values = ltbi.cols)+
+    scale_x_discrete(labels=c("PMA", "Mtb", "PMA", "Mtb"))+
+    theme(panel.border = element_rect(fill = NA, colour = "black"))+
+    theme(text = element_text(size=16))+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    stat_compare_means(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")) 
+                       ,method="wilcox.test", label="p.format")+
+    labs(x="", y="")
+```
+
+###Comparing Mtb-specific TH1 cells to total CD4 T cells in peptide stim  
+```{r, fig.width=12, fig.height=6}
+total<-filter_tb(subset(CD4, CD4$Stim=="UN"))
+total$cell<-"Total"
+phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
+names<-c("Donor", "TB", "SM", "Stim", "disease", names(phen)[grep("TH1_", names(phen))])
+mtb_th1<-select(phen, names)
+names(mtb_th1)<-gsub("TH1_", "", names(mtb_th1))
+mtb_th1$cell<-"Mtb-Specific"
+long<-rbindlist(list(total, mtb_th1), fill=TRUE)
+
+melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
+melt$cell<-factor(melt$cell, levels=c("Total", "Mtb-Specific"))
+melt$x<-paste(melt$SM, melt$cell, sep=" ")
+melt$x<-factor(melt$x, levels=c("SM- Total", "SM- Mtb-Specific","SM+ Total", "SM+ Mtb-Specific"))
+
+ggplot(melt, aes(x=x, y=value, fill=SM))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
+    facet_wrap(~variable, scales="free", nrow=1)+
+    theme_classic()+
+    scale_fill_manual(values = tb.cols)+
+    scale_x_discrete(labels=c("Total", "Mtb", "Total", "Mtb"))+
+    theme(panel.border = element_rect(fill = NA, colour = "black"))+
+    theme(text = element_text(size=16))+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    stat_compare_means(comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")) 
+                       ,method="wilcox.test", label="p.format")+
+    labs(x="", y="")
+
+```
+
+###Comparing Mtb-specific TH1 cells to PMA TH1 cells  
+```{r, fig.width=12, fig.height=6}
+pma<-filter_tb(subset(PHEN, PHEN$Stim=="PMA"))
+pma$cell<-"PMA"
+pep<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
+pep$cell<-"Mtb"
+long<-rbindlist(list(pma, pep), fill=TRUE)
+names<-c("Donor", "TB", "SM", "Stim", "disease", "cell", names(long)[grep("TH1_", names(long))])
+long<-select(long, names)
+names(long)<-gsub("TH1_", "", names(long))
+
+melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
+melt$x<-paste(melt$SM, melt$cell, sep=" ")
+melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))
+
+ggplot(melt, aes(x=x, y=value, fill=SM))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
+    facet_wrap(~variable, scales="free", nrow=1)+
+    theme_classic()+
+    scale_fill_manual(values = tb.cols)+
+    scale_x_discrete(labels=c("PMA", "Mtb", "PMA", "Mtb"))+
+    theme(panel.border = element_rect(fill = NA, colour = "black"))+
+    theme(text = element_text(size=16))+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    stat_compare_means(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")) 
+                       ,method="wilcox.test", label="p.format")+
+    labs(x="", y="")
+```
+
+```{r}
+supp4<-cowplot::plot_grid(supp4a, supp4b, nrow=2, labels=c("A", "B"))
+supp5<-cowplot::plot_grid(supp5a, supp5b, nrow=2, labels=c("A", "B"))
 ```
@@
 
-booleanplot(test, "variable", "value", ltbi.cols)+
+supp6a<-booleanplot(test, "variable", "value", ltbi.cols)+
     scale_x_discrete(labels = labels, breaks=breaks)+
@@
 
-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
+supp6b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
     scale_x_discrete(labels = labels, breaks=breaks)+
@@
     labs(y="(%) Proliferating Mtb-Specific cells")
+```
+
+```{r}
+supp6<-cowplot::plot_grid(supp6a, supp6b, nrow=2, labels=c("A", "B"), rel_widths = c(3,4))
+```
+
+###Comparing Mtb-specific TH1 cells to PMA TH1 cells  
+```{r, fig.width=12, fig.height=6}
+un<-filter_ltbi(subset(OG, OG$Stim=="UN"))
+un$cell<-"total"
+seb<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="SEB"))
+seb$cell<-"seb"
+long<-rbindlist(list(un, seb), fill=TRUE)
+
+melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
+melt$x<-paste(melt$SM, melt$cell, sep=" ")
+melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))
+
+ggplot(long, aes(x=cell, y=long[,"T-bet"], color=disease))+
+    geom_point(size=3, alpha=0.5)+geom_line(aes(group=Donor))+
+    facet_wrap(~TB, scales="free", nrow=1)+
+    theme_classic()+
+    scale_fill_manual(values = ltbi.cols)+
+    scale_x_discrete(labels=c("PMA", "Mtb", "PMA", "Mtb"))+
+    theme(panel.border = element_rect(fill = NA, colour = "black"))+
+    theme(text = element_text(size=16))+
+    theme(plot.title = element_text(hjust = 0.5)) +
+    stat_compare_means(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")) 
+                       ,method="wilcox.test", label="p.format")+
+    labs(x="", y="")
 ```
@@
 ```
+
+#Exporting Data
+```{r}
+setwd("/Users/tarynam/Desktop/Manuscript1")
+cowplot::save_plot("supp1.png", supp1, base_height=14, base_width=16)
+cowplot::save_plot("supp2.png", supp2, base_height=28, base_width=16)
+cowplot::save_plot("supp3.png", supp3, base_height=7, base_width=16)
+cowplot::save_plot("supp4.png", supp4, base_height=14, base_width=16)
+cowplot::save_plot("supp5.png", supp5, base_height=14, base_width=16)
+cowplot::save_plot("supp6.png", supp6, base_height=14, base_width=16)
+```

### 2019-09-15 gamma-delta.Rmd
pretty figures

@@
 compass$TB<-factor(compass$TB, levels=c("N", "HC", "LTBI", "TB"))
+kable(pvals_SM(compass[compass$Stim=="SWAP",], "FS"), caption="SWAP")
+kable(pvals_SM(compass[compass$Stim=="SEA",], "FS"), caption="SEA")
 
@@
 exclude<-subset(include, include$SM!="X")
-plot_cell(exclude, "FS")+
+keep<-plot_cell(exclude, "FS")+
     facet_grid(Stim~TB)+
@@
 CD3_OG<-dplyr::select(read.csv("OG_CD3_OG_summary_clean.csv"), -X, -X.1)
+CD4<-dplyr::select(read.csv("OG_CD4_bulk_summary_clean.csv"), -X, -X.1)
+CD8<-dplyr::select(read.csv("OG_CD8_bulk_summary_clean.csv"), -X, -X.1)
+
 mimosa<-dplyr::select(read.csv("OG_MIMOSA_clean.csv"), -X)
@@
 
-###*I was conflicted so all the plots on the left will include the SM- donors, whereas the plots on the right will exclude them.
+*I was conflicted so I ran a test to see if there were statistical differences between sm- and sm+ and when there are I plot both.
+All the plots on the left will include the SM- donors, whereas the plots on the right will exclude them.
 I still have to export the proliferation data for the yerkes donors but that's a good chunk of time/effort so I'm waiting until the other paper is further along*
@@
 include<-subset(CD3, CD3$Stim=="UN")
-a<-plot_3(include, "GD_freq")+
+kable(pvals_SM(include, "GD"))
+a<-plot_3(include, "GD")+
     coord_cartesian(ylim=c(0,35))+
@@
 exclude<-subset(include, include$SM=="SM")
-b<-plot_3(exclude, "GD_freq")+
+b<-plot_3(exclude, "GD")+
     coord_cartesian(ylim=c(0,35))+
@@
 
-###Proliferation of Gamma Delta Cells
-```{r}
-include<-subset(gd, gd$Stim=="SWAP")
-a<-plot_3(include, "OG_freq")+
-    labs(y="(%) γδ T cells")
+###Proportion of proliferating Gamma Delta Cells
+```{r, fig.width=12, fig.height=8}
+include<-subset(CD3_OG, CD3_OG$Stim=="SWAP")
+colnames<-c("CD4", "CD8", "GD")
+for(col in colnames){
+    y<-pvals_SM(include, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
+}
 
-exclude<-subset(include, include$SM=="SM")
-b<-plot_3(exclude, "OG_freq")+
-    labs(y="(%) γδ T cells")
+include<-subset(CD3_OG, CD3_OG$Stim%in%c("SEA","SWAP"))
+melt<-melt(include, id.vars=c("TB", "SM", "Stim"), measure.vars=colnames)
+melt$variable<-gsub("GD", "γδ", melt$variable)
 
-cowplot::plot_grid(a,b)
-```
-
-###Proportion of proliferating Gamma Delta Cells
-```{r, fig.width=12, fig.height=6}
-include<-subset(CD3_OG, CD3_OG$Stim=="SWAP")
-melt<-melt(include, id.vars=c("TB", "SM"), measure.vars=c("CD4_freq", "CD8_freq", "GD_freq"))
-melt$celltype<-gsub("_freq", "",melt$variable)
+melt$celltype<-gsub("", "",melt$variable)
 
 a<-plot_3(melt, "value")+
-    facet_grid(~variable, labeller = as_labeller(c("CD4_freq"="CD4", "CD8_freq"="CD8", "GD_freq"="γδ")))+
+    facet_grid(Stim~variable)+
     #coord_cartesian(ylim=c(0,65))+
@@
 b<-plot_3(melt, "value")+
-    facet_grid(~variable, labeller = as_labeller(c("CD4_freq"="CD4", "CD8_freq"="CD8", "GD_freq"="γδ")))+
+    facet_grid(Stim~variable, labeller = as_labeller(c("CD4"="CD4", "CD8"="CD8", "GD"="γδ")))+
     labs(y="(%) Proliferating T cells",
@@
 
-cowplot::plot_grid(a,b)
+a
 ```
 
-```{r, fig.width=12, fig.height=6}
-include<-subset(CD3_OG, CD3_OG$Stim=="SWAP")
-melt<-melt(include, id.vars=c("TB", "SM"), measure.vars=c("CD4_freq", "CD8_freq", "GD_freq"))
-melt$celltype<-gsub("_freq", "",melt$variable)
+```{r, fig.width=12, fig.height=8}
+include<-subset(CD3_OG, CD3_OG$Stim%in%c("SEA","SWAP"))
+melt<-melt(include, id.vars=c("TB", "SM", "Stim"), measure.vars=c("CD4", "CD8", "GD"))
+melt$celltype<-gsub("", "",melt$variable)
 
 a<-plot_cell(melt, "value")+
-    facet_grid(~TB)+
+    facet_grid(Stim~TB)+
     scale_x_discrete(labels = c("CD4", "CD8", "γδ"))+
@@
 b<-plot_cell(melt, "value")+
-    facet_grid(~TB)+
+    facet_grid(Stim~TB)+
     scale_x_discrete(labels = c("CD4", "CD8", "γδ"))+
@@
          title="Excludes SM-")
+
+a
+```
+
+###Proliferation of Gamma Delta Cells
+```{r}
+include<-subset(gd, gd$Stim=="SEA")
+pvals_SM(include, "OG")
+a<-plot_3(include, "OG")+
+    facet_grid(~Stim)+
+    labs(y="(%) γδ T cells")
+
+include<-subset(gd, gd$Stim=="SWAP")
+pvals_SM(include, "OG")
+b<-plot_3(include, "OG")+
+    facet_grid(~Stim)
+
+cowplot::plot_grid(a,b)
+```
+
+###Proliferation of CD4 T Cells
+```{r}
+include<-subset(CD4, CD4$Stim=="SEA")
+pvals_SM(include, "OG")
+a<-plot_3(include, "OG")+
+    facet_grid(~Stim)+
+    labs(y="(%) CD4 T cells")
+
+include<-subset(CD4, CD4$Stim=="SWAP")
+pvals_SM(include, "OG")
+b<-plot_3(include, "OG")+
+    facet_grid(~Stim)+
+    labs(y="")
+
+cowplot::plot_grid(a,b)
+```
+
+###Proliferation of CD8 T Cells
+```{r}
+include<-subset(CD8, CD8$Stim=="SEA")
+pvals_SM(include, "OG")
+a<-plot_3(include, "OG")+
+    facet_grid(~Stim)+
+    labs(y="(%) CD8 T cells")
+
+include<-subset(CD8, CD8$Stim=="SWAP")
+pvals_SM(include, "OG")
+b<-plot_3(include, "OG")+
+    facet_grid(~Stim)+
+    labs(y="")
 
@@
 include<-subset(gd_OG, gd_OG$Stim=="SWAP")
-a<-plot_3(include, "IL4_freq")+
+kable(pvals_SM(include, "IL4"))
+a<-plot_3(include, "IL4")+
     labs(y="(%) Proliferating γδ T cells")
@@
 exclude<-subset(include, include$SM=="SM")
-b<-plot_3(exclude, "IL4_freq")+
-    labs(y="(%) Proliferating γδ T cells")
+b<-plot_3(exclude, "IL4")+
+    labs(y="(%) Proliferating γδ T cells")+
+    labs(y="")
 
@@
 include<-subset(gd_OG, gd_OG$Stim=="SWAP")
-a<-plot_3(include, "TNFa_freq")+
+kable(pvals_SM(include, "TNFa"))
+a<-plot_3(include, "TNFa")+
     labs(y="(%) Proliferating γδ T cells")
@@
 exclude<-subset(include, include$SM=="SM")
-b<-plot_3(exclude, "TNFa_freq")+
+b<-plot_3(exclude, "TNFa")+
     labs(y="(%) Proliferating γδ T cells")
 
-cowplot::plot_grid(a,b)
+cowplot::plot_grid(a,NULL)
 ```
@@
 vars<-vars[-grep("._._._.", vars)]
+for(v in vars){
+    y<-pvals_SM(test, v)
+    if(dim(y)[1]>0){
+        print(v)
+        print(y)}
+}
 

### 2019-09-15 OG-Analysis_GD.Rmd
took out _freq naming

@@
 
+##Checking if I can combine SM- and SM+ in TB groups
+###CD3 T cells
+```{r}
+DF<-subset(CD3, CD3$Stim=="SWAP")
+colnames<-c("GD","OG")
+for(col in colnames){
+    y<-pvals_SM(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
+}
+
+DF<-subset(CD3_OG, CD3_OG$Stim=="SWAP")
+colnames<-c("GD")
+for(col in colnames){
+    y<-pvals_SM(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
+}
+
+```
+###total gamma delta
+```{r}
+DF<-subset(gd, gd$Stim=="SWAP")
+colnames<-numeric_only(DF)
+for(col in colnames){
+    y<-pvals_SM(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
+}
+
+```
+
+###proliferating gamma deltas
+```{r}
+DF<-subset(gd_OG, gd_OG$Stim=="SWAP")
+colnames<-numeric_only(DF)
+for(col in colnames){
+    y<-pvals_SM(DF, col)
+    if(dim(y)[1]>0){
+        print(col)
+        print(y)}
+}
+```
+
 ##Frequency of Gamma Deltas to start with
 ```{r}
-pvals_gd(subset(CD3, CD3$Stim=="UN"), "GD_freq")
-plot_with_all_stats(subset(CD3, CD3$Stim=="UN"), "GD_freq")
-plot_3(subset(CD3, CD3$Stim=="UN"), "GD_freq")
-pvals_gd(subset(CD3, CD3$Stim=="SWAP"), "GD_freq")
+pvals_gd(subset(CD3, CD3$Stim=="UN"), "GD")
+plot_with_all_stats(subset(CD3, CD3$Stim=="UN"), "GD")
+plot_3(subset(CD3, CD3$Stim=="UN"), "GD")
+pvals_gd(subset(CD3, CD3$Stim=="SWAP"), "GD")
 ```
@@
 DF<-subset(gd, gd$Stim=="SWAP")
-plot_with_all_stats(DF, "OG_freq")
-plot_3(DF,  "OG_freq")
-ggplot(DF, aes(x=SM, y=OG_freq, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
+DF$SM<-factor(DF$SM, levels=c("X", "SM"))
+plot_with_all_stats(DF, "OG")
+plot_3(DF,  "OG")
+ggplot(DF, aes(x=SM, y=OG, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means(label.y=85, label.x=1.3)+scale_x_discrete(labels=c("SM-", "SM+"))
 colnames<-numeric_only(DF)
@@
 DF<-subset(GD, GD$Stim=="SWAP")
-plot_with_all_stats(DF, "OG_freq")
-plot_3(DF,  "OG_freq")
-ggplot(DF, aes(x=SM, y=OG_freq, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
+plot_with_all_stats(DF, "OG")
+plot_3(DF,  "OG")
+ggplot(DF, aes(x=SM, y=OG, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
 colnames<-numeric_only(DF)
@@
 ```{r}
-start<-select(CD3, Sample, Stim, Donor, GD_freq)
+start<-select(CD3, Sample, Stim, Donor, GD)
 middle1<-filter(start, Stim=="UN")
 middle2<-filter(start, Stim=="SWAP")
-end<-select(gd, Sample, Stim, TB, SM, Donor, OG_freq)%>%filter(Stim=="SWAP")
+end<-select(gd, Sample, Stim, TB, SM, Donor, OG)%>%filter(Stim=="SWAP")
 
@@
 test2<-merge(test, end, by="Donor")
-melt<-melt(test2, id.vars=c("Donor", "TB", "SM"), measure.vars=c("GD_freq.x", "GD_freq.y"))
+melt<-melt(test2, id.vars=c("Donor", "TB", "SM"), measure.vars=c("GD.x", "GD.y"))
 ggplot(melt, aes(x=variable, y=value, color=TB))+geom_point()+geom_line(aes(group=Donor))+facet_grid(~TB)+scale_x_discrete(labels=c("Starting freq", "Swap Freq", "Proliferating Freq"))
 
-test2$ratio<-test2$GD_freq.y/test2$GD_freq.x
+test2$ratio<-test2$GD.y/test2$GD.x
 plot_with_all_stats(test2, "ratio")+labs(title="Expansion of gd")
@@
 
-test2$dif<-test2$GD_freq.y-test2$GD_freq.x
+test2$dif<-test2$GD.y-test2$GD.x
 plot_with_all_stats(test2, "dif")+labs(title="Expansion of gd")
@@
 for(i in 1:length(phenotype$include)){
-    if(is.na(phenotype$OG_freq[i])){
+    if(is.na(phenotype$OG[i])){
         phenotype$include[i]<-0
@@
 #this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
-colnames<-c(test[1:39], test[45:56])
+colnames<-c(test[1:39], test[45:52])
 for(col in colnames){
@@
 ```{r}
-test<-filter(DF, !is.na(OG_freq))
+test<-filter(DF, !is.na(OG))
 kable(xtabs(~TB+SM, test))
@@
 
-plot_with_all_stats(DF, "IL4_freq")
+plot_with_all_stats(DF, "IL4")
 
-plot_3(DF, "IL4_freq")
-plot_3(DF, "TNFa_freq")
-plot_3(DF, "IFNg_freq")
+plot_3(DF, "IL4")
+plot_3(DF, "TNFa")
+plot_3(DF, "IFNg")
 
-ggplot(DF, aes(x=SM, y=IL4_freq, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
+ggplot(DF, aes(x=SM, y=IL4, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
 
-ggplot(DF, aes(x=SM, y=TNFa_freq, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
+ggplot(DF, aes(x=SM, y=TNFa, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
 
-ggplot(DF, aes(x=SM, y=IFNg_freq, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
+ggplot(DF, aes(x=SM, y=IFNg, fill=SM))+geom_boxplot(outlier.shape = NA)+geom_jitter(width=.1)+stat_compare_means()
 ```
@@
 DF<-merge(select(PBMC, Donor, SchistosomaIntensity), CD3[CD3$Stim=="SWAP",], by="Donor")
-ggplot(DF[DF$SM=="SM",], aes(x=OG_freq, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="CD3")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
+ggplot(DF[DF$SM=="SM",], aes(x=OG, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="CD3")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
 
 DF<-merge(select(PBMC, Donor, SchistosomaIntensity), CD3_OG[CD3_OG$Stim=="SWAP",], by="Donor")
-ggplot(DF[DF$SM=="SM",], aes(x=GD_freq, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="CD3")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
+ggplot(DF[DF$SM=="SM",], aes(x=GD, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="CD3")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
 
 DF<-merge(select(PBMC, Donor, SchistosomaIntensity), GD[GD$Stim=="SWAP",], by="Donor")
-ggplot(DF[DF$SM=="SM",], aes(x=OG_freq, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="GD")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
+ggplot(DF[DF$SM=="SM",], aes(x=OG, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="GD")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
 
 DF<-merge(select(PBMC, Donor, SchistosomaIntensity), GD_OG[GD_OG$Stim=="SWAP",], by="Donor")
-ggplot(DF[DF$SM=="SM",], aes(x=IL4_freq, y=SchistosomaIntensity, color=TB))+
+ggplot(DF[DF$SM=="SM",], aes(x=IL4, y=SchistosomaIntensity, color=TB))+
     geom_point()+
@@
 DF<-merge(select(PBMC, Donor, SchistosomaIntensity), GD_OG[GD_OG$Stim=="SWAP",], by="Donor")
-ggplot(DF[DF$SM=="SM",], aes(x=TNFa_freq, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="GD")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
+ggplot(DF[DF$SM=="SM",], aes(x=TNFa, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="GD")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
 
 DF<-merge(select(PBMC, Donor, SchistosomaIntensity), GD_OG[GD_OG$Stim=="SWAP",], by="Donor")
-ggplot(DF[DF$SM=="SM",], aes(x=IFNg_freq, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="GD")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
+ggplot(DF[DF$SM=="SM",], aes(x=IFNg, y=SchistosomaIntensity, color=TB))+geom_point()+stat_smooth(method="lm")+labs(title="GD")+scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))
 
@@
 #this will also include metadata so we remove those column names and can phenotype using the standard pvals function from the beginning
-colnames<-c(test[1:32], test[38:49])
+colnames<-c(test[1:32])
 for(col in colnames){

### 2019-09-17 PaperFigures.Rmd
change the way p-values display

@@
     theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
+    theme(text = element_text(size=20),
+          strip.text = element_text(size = 20))+
     theme(plot.title = element_text(hjust = 0.5)) +
@@
     theme(legend.position="none")+
-    theme(text = element_text(size=16), axis.text.x = element_text(size=16))+
+    theme(text = element_text(size=20), 
+          axis.text.x = element_text(size=16), 
+          strip.text = element_text(size = 20))+
     theme(plot.title = element_text(hjust = 0.5)) +
@@
     facet_wrap(~variable)+
-    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
-    theme(text = element_text(size=16))+
+    theme(text = element_text(size=20),
+         strip.text = element_text(size = 20))+
     theme(plot.title = element_text(hjust = 0.5))+
-    stat_compare_means(method = 'wilcox.test', label="p.signif", hide.ns=TRUE, size=10)+
     labs(x="", y="")
@@
 tf<-allplot(melt, "TB", "value", all.cols)+    
-    labs(y="(%) Total CD4+ T cells")
+    labs(y="(%) Total CD4+ T cells")+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    stat_compare_means(method = 'wilcox.test', label="p.signif", hide.ns=TRUE, size=10, label.y=25)+
+    coord_cartesian(ylim=c(0,30))
 
@@
 sm<-allplot(melt, "TB", "value", all.cols) +    
-    labs(y="(%) Total CD4+ T cells")
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    labs(y="(%) Total CD4+ T cells")+
+    stat_compare_means(method = 'wilcox.test', label="p.signif", hide.ns=TRUE, size=10, label.y=60)
 
@@
 figure1<- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,2))
+```
+
+#Figure 1 alt B
+
+```{r, fig.width=15, fig.height=8}
+melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))
+
+tf<-allplot(melt, "TB", "value", all.cols)+    
+    labs(y="(%) Total CD4+ T cells")+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    stat_compare_means(method = 'wilcox.test', label="p.format", size=6, label.y=.9*30)+
+    coord_cartesian(ylim=c(0,30))
+
+melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))
+
+sm<-allplot(melt, "TB", "value", all.cols) +    
+    labs(y="(%) Total CD4+ T cells")+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    stat_compare_means(method = 'wilcox.test', label="p.format", size=6, label.y=.9*65)+
+    coord_cartesian(ylim=c(0,65))
+
+first_row = cowplot::plot_grid(NULL, NULL, labels = c('A' , 'B'), nrow=1, label_size = 20, rel_widths = c(3,2))
+second_row = cowplot::plot_grid(tf, sm,labels = c('C', 'D'), nrow=1, label_size = 20)
+figure1B<- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,2))
+```
+
+#Figure 1 alt C
+
+```{r, fig.width=15, fig.height=8}
+melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))
+
+tf<-allplot(melt, "disease", "value", all.cols)+    
+    labs(y="(%) Total CD4+ T cells")+
+    scale_x_discrete(labels=c("HC\nSM-", "HC\nSM+", 
+                              "LTBI\nSM-", "LTBI\nSM+", 
+                              "TB\nSM-", "TB\nSM+"))+
+    stat_compare_means(comparisons=list(c("HC SM-", "HC SM+")), method = 'wilcox.test', label="p.format", size=6, label.y=.9*30)+
+    coord_cartesian(ylim=c(0,30))
+
+melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))
+
+sm<-allplot(melt, "disease", "value", all.cols) +    
+    labs(y="(%) Total CD4+ T cells")+    
+    scale_x_discrete(labels=c("HC\nSM-", "HC\nSM+", 
+                              "LTBI\nSM-", "LTBI\nSM+", 
+                              "TB\nSM-", "TB\nSM+"))+
+    stat_compare_means(comparisons=list(c("TB SM-", "TB SM+")),method = 'wilcox.test', label="p.format", size=6, label.y=.9*65)+
+    coord_cartesian(ylim=c(0,65))
+
+first_row = cowplot::plot_grid(NULL, NULL, labels = c('A' , 'B'), nrow=1, label_size = 20, rel_widths = c(3,2))
+second_row = cowplot::plot_grid(tf, sm,labels = c('C', 'D'), nrow=1, label_size = 20)
+figure1C<- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,2))
 ```
@@
     facet_wrap(~variable, labeller = cyt_groups)+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
     labs(y="(%) Total CD4+ T cells")
@@
     facet_wrap(~variable, labeller = cyt_groups)+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
     coord_cartesian(ylim=c(0,8))+
@@
     labs(y="(%) Total CD4+ T cells")+coord_cartesian(ylim=c(0,45))+
-    labs(title="Healthy Controls")
+    labs(title="HC")
 
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y = (.95*.15))+
+    stat_compare_means(comparisons=list(c("SM-","SM+")), method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y = (.90*.15))+
     coord_cartesian(ylim=c(0,.15))+
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y = (.95*.02))+
     coord_cartesian(ylim=c(0,.02))
@@
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*70))+
     coord_cartesian(ylim=c(0,70))+
@@
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
     coord_cartesian(ylim=c(0,100))+
@@
 melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
-a<-baseplot(melt, "disease", "value", tb.cols)+    
+a<-baseplot(melt, "SM", "value", tb.cols)+    
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.95*.15))+
     coord_cartesian(ylim=c(0,.15))+
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.95*.02))+
     coord_cartesian(ylim=c(0,.02))
@@
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*80))+
+    stat_compare_means(comparisons=list(c("SM-","SM+")), method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.98*80))+
     coord_cartesian(ylim=c(0,80))+
@@
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*80))+
     coord_cartesian(ylim=c(0,80))
@@
     facet_wrap(~variable, labeller = groups)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3)+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
+    stat_compare_means(comparisons=list(c("SM-","SM+")), method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=.98*13)+
+    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")+
+    coord_cartesian(ylim=c(0,13))
 
@@
 tbphen<-cowplot::plot_grid(tf, sm, both, nrow=1, labels=c("D","E","F"), label_size = 20)
-figure5<-cowplot::plot_grid(tbcyt, tbphen, nrow=2, rel_heights = c(2,3))
+figure5<-cowplot::plot_grid(tbcyt, tbphen, nrow=2, rel_heights = c(2,2))
 ```
@@
     facet_wrap(~variable, scales="free", labeller = groups)+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
     labs(y="(%) Total CD4+ T cells")
@@
 two<-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, labeller = groups)
+    facet_wrap(~variable, labeller = groups)+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    coord_cartesian(ylim=c(0,5))
 ```
@@
 
-first_row = cowplot::plot_grid(NULL, labels = c('A', 'B'), nrow=1, label_size = 20)
-second_row = cowplot::plot_grid(og, one, two, nrow=1, rel_widths = c(1.2,2,1), labels=c("B", "C","D"), label_size = 20)
-third_row = cowplot::plot_grid(tf, sm, labels=c("E", "F"), nrow=1, label_size = 20)
-figure6<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3)
+first_row = cowplot::plot_grid(NULL, NULL, labels = c('A', 'B'), nrow=1, label_size = 20, rel_widths = c(3,1))
+second_row = cowplot::plot_grid(og, one, two, nrow=1, rel_widths = c(1,2,1), labels=c("C","D","E"), label_size = 20)
+third_row= cowplot::plot_grid(tf, sm, labels=c('F', 'G'), nrow=1, label_size = 20)
+
+figure6<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3, rel_heights = c(1.5,2,2))
 ```
@@
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
     coord_cartesian(ylim=c(0,100))
@@
 second_row = cowplot::plot_grid(og, one, two, nrow=1, rel_widths = c(1.2,2,1), labels=c("B", "C","D"), label_size = 20)
-third_row = cowplot::plot_grid(tf, sm, labels=c("E", "F"), nrow=1, label_size = 20)
-figure7<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3)
+third_row = cowplot::plot_grid(tf, sm, labels=c('E', 'F'), nrow=1, label_size = 20)
+figure7<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3, rel_heights = c(2,2,2))
 ```
@@
 
-lumplot<-cowplot::plot_grid(a,b, NULL, ncol=1, align = "v", labels=c("A","A", "B"), label_size = 20)
+c<-baseplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+
+    labs(y="pg/mL")+
+    coord_cartesian(ylim=c(0,20))
+
+lumplot<-cowplot::plot_grid(a,b,c, NULL, ncol=1, align = "v", labels=c("A","A", "A", "B"), label_size = 20)
 ```
@@
 setwd("/Users/tarynam/Desktop/Manuscript1")
+
 cowplot::save_plot("figure1.png", figure1, base_height=8, base_width=16) #Figure 1
+cowplot::save_plot("figure1alt1.png", figure1B, base_height=8, base_width=16) #Figure 1
+cowplot::save_plot("figure1alt2.png", figure1C, base_height=8, base_width=16) #Figure 1
 cowplot::save_plot("figure2.png", figure2, base_height=14, base_width=16) #Figure 2
 cowplot::save_plot("figure3.png", figure3, base_height=14, base_width=16)
-cowplot::save_plot("figure4.png", figure4, base_height=14, base_width=16)
+cowplot::save_plot("figure4.png", figure4, base_height=14, base_width=12)
 cowplot::save_plot("figure5.png", figure5, base_height=14, base_width=16)
@@
 cowplot::save_plot("figure7.png", figure7, base_height=16, base_width=16)
-cowplot::save_plot("figure8.png", lumplot, base_height=32, base_width=16)
+cowplot::save_plot("figure8.png", lumplot, base_height=48, base_width=16)
 ```

### 2019-09-18 PaperFigures.Rmd
took SEB phenotype out of main figs

@@
 
-first_row = cowplot::plot_grid(NULL, NULL, labels = c('A' , 'B'), nrow=1, label_size = 20, rel_widths = c(3,2))
-second_row = cowplot::plot_grid(tf, sm,labels = c('C', 'D'), nrow=1, label_size = 20)
+first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20, rel_widths = c(3,2))
+second_row = cowplot::plot_grid(tf, sm,labels = c('B', 'C'), nrow=1, label_size = 20)
 figure1<- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,2))
@@
 #Figure 1 alt B
-
 ```{r, fig.width=15, fig.height=8}
@@
 #Figure 1 alt C
-
 ```{r, fig.width=15, fig.height=8}
@@
 
-first_row = cowplot::plot_grid(NULL, NULL, labels = c('A' , 'B'), nrow=1, label_size = 20, rel_widths = c(3,2))
-second_row = cowplot::plot_grid(tf, sm,labels = c('C', 'D'), nrow=1, label_size = 20)
+first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20, rel_widths = c(3,2))
+second_row = cowplot::plot_grid(tf, sm,labels = c('B', 'C'), nrow=1, label_size = 20)
 figure1C<- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,2))
@@
 
-first_row = cowplot::plot_grid(NULL, NULL, labels = c('A', 'B'), nrow=1, label_size = 20)
-second_row = cowplot::plot_grid(one, two, labels = c('C', 'D'), nrow=1, label_size = 20)
-third_row = cowplot::plot_grid(a, b, c, nrow=1, labels=c('E'), label_size = 20)
+first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
+second_row = cowplot::plot_grid(one, two, labels = c('B', 'C'), nrow=1, label_size = 20)
+third_row = cowplot::plot_grid(a, b, c, nrow=1, labels=c('D'), label_size = 20)
 figure2 = cowplot::plot_grid(first_row, second_row, third_row, ncol=1, rel_heights = c(1,2,2))
@@
     coord_cartesian(ylim=c(0,100))+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
+    labs(y="")
+
+first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
+second_row = cowplot::plot_grid(a, b, tf, sm, nrow=1, labels=c('B', 'C', 'D', 'E'), label_size = 20)
+figure4opt1 <- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(3,4))
 
@@
 third_row = cowplot::plot_grid(tf, sm, nrow=1, labels=c('D', 'E'), label_size = 20)
-figure4 <- cowplot::plot_grid(first_row, second_row, third_row, ncol=1, rel_heights = c(1,2,2))
+figure4opt2 <- cowplot::plot_grid(first_row, second_row, third_row, ncol=1, rel_heights = c(2.5,4,4))
+
 ```
@@
 ```{r}
+groups <- as_labeller(c("OG" = "Proliferation","IFNg" = "IFNγ", 
+                    "TNFa" = "TNFα", "IL4" = "IL-4"))
+
 og<-plot_filter(dplyr::filter(OG, Stim=="SEB"))
-ogphen<-plot_filter(dplyr::filter(OGPHEN, Stim=="SEB"))
+part1<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG")
+seb<-allplot(part1, "TB", "value", all.cols)+
+    facet_wrap(~variable, scales="free", labeller = groups)+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    labs(y="(%) Total CD4+ T cells",
+         title="SEB")
+
+og<-plot_filter(dplyr::filter(OG, Stim=="PEP", TB %in% c("LTBI", "TB")))
+part2<-melt(og, id.vars=c("TB", "SM", "disease"), measure.vars="OG")
+pep<-allplot(part2, "TB", "value", all.cols)+
+    facet_wrap(~variable, scales="free", labeller = groups)+
+    scale_x_discrete(labels=c("SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    coord_cartesian(ylim=c(0,3))+
+    labs(title="CFP-10 & ESAT-6")
+
+first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
+second_row = cowplot::plot_grid(seb, pep, nrow=1, rel_widths = c(3,2), labels=c("B", "C"), label_size = 20)
+figure6<-cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1.5,3))
 
@@
 kable(table(ogphen$disease[!is.na(ogphen$IFNg)]), caption="SEB phenotype")
-```
-
-```{r, proliferation and total cyt, fig.width=15, fig.height=8}
-groups <- as_labeller(c("OG" = "Proliferation","IFNg" = "IFNγ", 
-                    "TNFa" = "TNFα", "IL4" = "IL-4"))
-
-part1<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG")
-og<-allplot(part1, "TB", "value", all.cols)+
-    facet_wrap(~variable, scales="free", labeller = groups)+
-    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
-    labs(y="(%) Total CD4+ T cells")
-
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
-
-one<-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, labeller = groups, ncol=2)+
-    labs(y="(%) Proliferating CD4+ T cells")
-
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))
-
-two<-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, labeller = groups)+
-    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
-    coord_cartesian(ylim=c(0,5))
-```
-
-```{r, marker on proliferating, fig.width=15, fig.height=8}
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))
-
-tf<-allplot(melt, "TB", "value", all.cols)+    
-    labs(y="(%) Proliferating CD4+ T cells")
-
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))
-
-sm<-allplot(melt, "TB", "value", all.cols)
-
-first_row = cowplot::plot_grid(NULL, NULL, labels = c('A', 'B'), nrow=1, label_size = 20, rel_widths = c(3,1))
-second_row = cowplot::plot_grid(og, one, two, nrow=1, rel_widths = c(1,2,1), labels=c("C","D","E"), label_size = 20)
-third_row= cowplot::plot_grid(tf, sm, labels=c('F', 'G'), nrow=1, label_size = 20)
-
-figure6<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3, rel_heights = c(1.5,2,2))
 ```
@@
 ```{r og and cyt, fig.width=15, fig.height=8}
-groups <- as_labeller(c("OG" = "Proliferation","IFNg" = "IFNγ", 
+groups <- as_labeller(c("IFNg" = "IFNγ", 
                     "TNFa" = "TNFα", "IL4" = "IL-4"))
-
-melt<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG")
-og<-baseplot(melt, "TB", "value", ltbi.cols)+
-    facet_wrap(~variable, scales="free", labeller = groups)+
-    coord_cartesian(ylim=c(0,3))+
-    labs(y="(%) Total CD4+ T cells")
 
@@
 first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
-second_row = cowplot::plot_grid(og, one, two, nrow=1, rel_widths = c(1.2,2,1), labels=c("B", "C","D"), label_size = 20)
+second_row = cowplot::plot_grid(one, two, nrow=1, rel_widths = c(2,1), labels=c("B", "C","D"), label_size = 20)
 third_row = cowplot::plot_grid(tf, sm, labels=c('E', 'F'), nrow=1, label_size = 20)
 figure7<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3, rel_heights = c(2,2,2))
-```
-
-###On TH1 proliferating cells
-```{r og th1 main four markers, fig.width=16, fig.height=8}
-melt<-melt(ogth1, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
-melt$variable<-factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))
-
-a<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
-    facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
-    coord_cartesian(ylim=c(0,100))+
-    labs(y="(%) Proliferating Mtb-Specific cells")
-
-b<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
-    facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
-    coord_cartesian(ylim=c(0,100))
-
-ogth1<-cowplot::plot_grid(a,b,ncol=2)
 ```
@@
 cowplot::save_plot("figure1.png", figure1, base_height=8, base_width=16) #Figure 1
-cowplot::save_plot("figure1alt1.png", figure1B, base_height=8, base_width=16) #Figure 1
+#cowplot::save_plot("figure1alt1.png", figure1B, base_height=8, base_width=16) #Figure 1
 cowplot::save_plot("figure1alt2.png", figure1C, base_height=8, base_width=16) #Figure 1
@@
 cowplot::save_plot("figure3.png", figure3, base_height=14, base_width=16)
-cowplot::save_plot("figure4.png", figure4, base_height=14, base_width=12)
-cowplot::save_plot("figure5.png", figure5, base_height=14, base_width=16)
-cowplot::save_plot("figure6.png", figure6, base_height=16, base_width=16)
-cowplot::save_plot("figure7.png", figure7, base_height=16, base_width=16)
+
+cowplot::save_plot("figure4-1.png", figure4opt1, base_height=10, base_width=16)
+cowplot::save_plot("figure4-2.png", figure4opt2, base_height=15, base_width=14)
+
+cowplot::save_plot("figure5.png", figure5, base_height=11, base_width=16)
+cowplot::save_plot("figure6.png", figure6, base_height=10, base_width=12)
+cowplot::save_plot("figure7.png", figure7, base_height=16, base_width=12)
 cowplot::save_plot("figure8.png", lumplot, base_height=48, base_width=16)
@@
 ##Proliferation Supplementary
+```{r}
+ogphen<-plot_filter(subset(OGPHEN, OGPHEN$Stim=="SEB"))
+
+#cytokines
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
+one<-allplot(melt, "TB", "value", all.cols)+    
+    facet_wrap(~variable, labeller = groups, ncol=2)+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    labs(y="(%) Proliferating CD4+ T cells")
+
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))
+two<-allplot(melt, "TB", "value", all.cols)+    
+    facet_wrap(~variable, labeller = groups)+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    coord_cartesian(ylim=c(0,5))
+
+#phenotypes
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))
+tf<-allplot(melt, "TB", "value", all.cols)+    
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    labs(y="(%) Proliferating CD4+ T cells")
+
+melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))
+sm<-allplot(melt, "TB", "value", all.cols)+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))
+
+first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20, rel_widths = c(3,1))
+second_row = cowplot::plot_grid(og, one, two, nrow=1, rel_widths = c(1,2,1), labels=c("B","C","D"), label_size = 20)
+third_row= cowplot::plot_grid(tf, sm, labels=c('F', 'G'), nrow=1, label_size = 20)
+
+supp6<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3)
+```
+
 ```{r}
@@
 
-supp6a<-booleanplot(test, "variable", "value", ltbi.cols)+
+supp7a<-booleanplot(test, "variable", "value", ltbi.cols)+
     scale_x_discrete(labels = labels, breaks=breaks)+
@@
 
-supp6b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
+supp7b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
     scale_x_discrete(labels = labels, breaks=breaks)+
@@
 ```{r}
-supp6<-cowplot::plot_grid(supp6a, supp6b, nrow=2, labels=c("A", "B"), rel_widths = c(3,4))
+supp7<-cowplot::plot_grid(supp7a, supp7b, nrow=2, labels=c("A", "B"), rel_widths = c(3,4))
 ```
@@
 cowplot::save_plot("supp6.png", supp6, base_height=14, base_width=16)
-```
+cowplot::save_plot("supp7.png", supp6, base_height=14, base_width=16)
+```

### 2019-09-28 OG_cleaning.Rmd
added bit to get OG neg cells

@@
 filenames<-ls(pattern="summary.csv")
-for(i in 1:16){
+for(i in 1:17){
 assign(filenames[i], data.frame(l[i]))}
@@
 filenames<-ls(pattern="_OG")
-for(i in 1:9){
+for(i in 1:10){
 assign(filenames[i], data.frame(l[i]))}
+
+OG_CD4_noOG_summary.csv<-sum_boolean(OG_CD4_noOG_summary.csv)
+OG_CD4_noOG_summary.csv<-sum_boolean2(OG_CD4_noOG_summary.csv)
 ```
@@
 write.csv(OG_CD4_OG.IFNg_summary.csv, "OG_CD4_OG.IFNg_summary_clean.csv")
+write.csv(OG_CD4_noOG_summary.csv, "OG_CD4_noOG_summary_clean.csv")
 

### 2019-09-28 PaperFigures.Rmd
separated main figures from supplementary figures

--- 2019-09-12 PaperFigures.Rmd
@@
     theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
+    theme(text = element_text(size=20),
+          axis.text.x = element_text(size = 16),
+          strip.text = element_text(size = 20))+
+    theme(plot.title = element_text(hjust = 0.5, size=20)) +
     labs(x="", y="")
@@
     theme(legend.position="none")+
-    theme(text = element_text(size=16), axis.text.x = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
+    theme(text = element_text(size=20), 
+          axis.text.x = element_text(size=16), 
+          strip.text = element_text(size = 20))+
+    theme(plot.title = element_text(hjust = 0.5, size=20)) +
     stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
@@
     facet_wrap(~variable)+
-    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5))+
-    stat_compare_means(method = 'wilcox.test', label="p.signif", hide.ns=TRUE, size=10)+
+    theme(text = element_text(size=20),
+            axis.text.x = element_text(size = 16),
+            strip.text = element_text(size = 20))+
+    theme(plot.title = element_text(hjust = 0.5, size=20))+
     labs(x="", y="")
@@
 OGPHEN<-read.csv("OG_CD4_OG_summary_clean_NAs.csv")
+OGnope<-read.csv("OG_CD4_noOG_summary_clean.csv")
 OG.IFNg<-read.csv("OG_CD4_OG.IFNg_summary_clean_NAs.csv")
@@
 tf<-allplot(melt, "TB", "value", all.cols)+    
-    labs(y="(%) Total CD4+ T cells")
+    labs(y="(%) Total CD4 T cells")+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    stat_compare_means(method = 'wilcox.test', label="p.signif", hide.ns=TRUE, size=10, label.y=25)+
+    coord_cartesian(ylim=c(0,30))
 
@@
 sm<-allplot(melt, "TB", "value", all.cols) +    
-    labs(y="(%) Total CD4+ T cells")
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    labs(y="")+
+    stat_compare_means(method = 'wilcox.test', label="p.signif", hide.ns=TRUE, size=10, label.y=60)
 
-first_row = cowplot::plot_grid(NULL, NULL, labels = c('A' , 'B'), nrow=1, label_size = 20, rel_widths = c(3,2))
-second_row = cowplot::plot_grid(tf, sm,labels = c('C', 'D'), nrow=1, label_size = 20)
+first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20, rel_widths = c(3,2))
+second_row = cowplot::plot_grid(tf, sm,labels = c('B', ''), nrow=1, label_size = 20)
 figure1<- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,2))
+figure1
+```
+
+#Figure 1 alt
+```{r, fig.width=15, fig.height=8}
+melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))
+
+tf<-allplot(melt, "disease", "value", all.cols)+    
+    labs(y="(%) Total CD4 T cells")+
+    scale_x_discrete(labels=c("HC\nSM-", "HC\nSM+", 
+                              "LTBI\nSM-", "LTBI\nSM+", 
+                              "TB\nSM-", "TB\nSM+"))+
+    stat_compare_means(comparisons=list(c("HC SM-", "HC SM+")), method = 'wilcox.test', label="p.format", size=6, label.y=.9*30)+
+    coord_cartesian(ylim=c(0,30))
+
+melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))
+
+sm<-allplot(melt, "disease", "value", all.cols) +    
+    labs(y="")+    
+    scale_x_discrete(labels=c("HC\nSM-", "HC\nSM+", 
+                              "LTBI\nSM-", "LTBI\nSM+", 
+                              "TB\nSM-", "TB\nSM+"))+
+    stat_compare_means(comparisons=list(c("TB SM-", "TB SM+")), 
+                       method = 'wilcox.test', label="p.format", size=6, label.y=.9*65)+
+    coord_cartesian(ylim=c(0,65))
+
+first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20, rel_widths = c(3,2))
+second_row = cowplot::plot_grid(tf, sm,labels = c('B', ''), nrow=1, label_size = 20)
+figure1alt<- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,2))
+figure1alt
 ```
@@
 #Figure 2
-####Generic cytokine production from CD4+ T cells stimulated with PMA & Ionomycin does not differ between SM- and SM+ Healthy Adults
-```{r, fig.width=15, fig.height=8}
+####Generic cytokine production from CD4 T cells stimulated with PMA & Ionomycin does not differ between SM- and SM+ Healthy Adults
+```{r, fig.width=15, fig.height=15}
 melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    labs(y="(%) Total CD4+ T cells")
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    labs(y="(%) Total CD4 T cells")
 
@@
     facet_wrap(~variable, labeller = cyt_groups)+
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
     coord_cartesian(ylim=c(0,8))+
-    labs(y="(%) Total CD4+ T cells")
+    labs(y="")
 
@@
     facet_wrap(~variable, ncol=4, labeller=th_groups)+
-    labs(y="(%) Total CD4+ T cells")+coord_cartesian(ylim=c(0,45))+
-    labs(title="Healthy Controls")
+    labs(y="(%) Total CD4 T cells")+coord_cartesian(ylim=c(0,45))+
+    labs(title="HC")
 
@@
 
-first_row = cowplot::plot_grid(NULL, NULL, labels = c('A', 'B'), nrow=1, label_size = 20)
-second_row = cowplot::plot_grid(one, two, labels = c('C', 'D'), nrow=1, label_size = 20)
-third_row = cowplot::plot_grid(a, b, c, nrow=1, labels=c('E'), label_size = 20)
+first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
+second_row = cowplot::plot_grid(one, two, labels = c('B', ''), nrow=1, label_size = 20)
+third_row = cowplot::plot_grid(a, b, c, nrow=1, labels=c('C'), label_size = 20)
 figure2 = cowplot::plot_grid(first_row, second_row, third_row, ncol=1, rel_heights = c(1,2,2))
+figure2
+```
+
+some stats comparing th1 to th2 for reference. 
+```{r}
+x<-filter(plot_filter(melt), variable!="TH1.2.Freq")
+y<-split(x,list(x$SM, x$TB)) 
+A<-lapply(y, function(g) wilcox.test(value~variable, g))
+
+pval<-c(A$`SM-.HC`$p.value, A$`SM+.HC`$p.value,
+        A$`SM-.LTBI`$p.value, A$`SM+.LTBI`$p.value, 
+        A$`SM-.TB`$p.value, A$`SM+.TB`$p.value)
+group<-c("HC SM-", "HC SM+", "LTBI SM-", "LTBI SM+", "TB SM-", "TB SM+")
+table<-data.frame(cbind(group, pval))
+table$pval<-as.numeric(as.character(table$pval))
+table$pval[table$pval<.0001]<-"<0.0001"
+kable(table, caption="TH1 to TH2")
+```
+
+some stats comparing th1 to th1/2 for reference. 
+```{r}
+x<-filter(plot_filter(melt), variable!="TH2.Freq")
+y<-split(x,list(x$SM, x$TB)) 
+A<-lapply(y, function(g) wilcox.test(value~variable, g))
+
+pval<-c(A$`SM-.HC`$p.value, A$`SM+.HC`$p.value,
+        A$`SM-.LTBI`$p.value, A$`SM+.LTBI`$p.value, 
+        A$`SM-.TB`$p.value, A$`SM+.TB`$p.value)
+group<-c("HC SM-", "HC SM+", "LTBI SM-", "LTBI SM+", "TB SM-", "TB SM+")
+table<-data.frame(cbind(group, pval))
+table$pval<-as.numeric(as.character(table$pval))
+table$pval[table$pval<.0001]<-"<0.0001"
+kable(table, caption="TH1 to TH1/2")
 ```
@@
 ####Generic TH1 and TH2 cytokine producing cells are stratified by lineage-specific transcription factors, and CCR4
-Amongst TH1 cytokine producing CD4+ T cells, moderate frequencies express the transcription factor T-bet while very low frequencies express GATA3.
-In contrast, amongst TH2 cytokine producing CD4+ T cells, low frequencies express T-bet while high frequencies express GATA3.
+Amongst TH1 cytokine producing CD4 T cells, moderate frequencies express the transcription factor T-bet while very low frequencies express GATA3.
+In contrast, amongst TH2 cytokine producing CD4 T cells, low frequencies express T-bet while high frequencies express GATA3.
 Similarly high frequencies of TH2 cytokine producing cells express CCR4 compared to TH1 cytokine producing cells.
@@
 test$subset<-tstrsplit(test$variable, "_")[[1]]
-test$subset<-paste(test$subset, "Cytokine+\nCD4+ T Cells", sep=" ")
+test$subset<-paste(test$subset, "Cytokine+\nCD4 T Cells", sep=" ")
 test$variable<-gsub("TH1_","", test$variable)
@@
     theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    labs(y="(%) Cytokine+ CD4+ T cells",
+
+    labs(y="(%) Cytokine+ CD4 T cells",
          title="HC")
@@
     theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    labs(y="(%) Cytokine+ CD4+ T cells",
+    theme(plot.margin=margin(l=-1,unit="cm"))+
+    labs(y="",
          title="LTBI")
@@
     theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    labs(y="(%) Cytokine+ CD4+ T cells",
+    theme(plot.margin=margin(l=-1,unit="cm"))+
+    labs(y="",
          title="TB")
 
-first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
-second_row = cowplot::plot_grid(a, b, c, nrow=1, labels=c('B', 'C', 'D'), label_size = 20)
-figure3 <- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,4))
+second_row <- cowplot::plot_grid(a,b,c, nrow=1)
+figure3 <- cowplot::plot_grid(NULL, second_row, labels = c('A', 'B'), label_size = 20, ncol=1, rel_heights = c(1.5,4.5))
+figure3
 ```
@@
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y = (.95*.15))+
+    stat_compare_means(comparisons=list(c("SM-","SM+")), method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y = (.90*.15))+
     coord_cartesian(ylim=c(0,.15))+
-    labs(y="(%) CD4+ T cells")
+    labs(y="(%) CD4 T cells")
 
@@
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*70))+
     coord_cartesian(ylim=c(0,70))+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
+    labs(y="(%) Mtb-Specific\nTH1 Cytokine+ cells")
 
@@
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
     coord_cartesian(ylim=c(0,100))+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
+    labs(y="")
 
 first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
-second_row = cowplot::plot_grid(a, b, nrow=1, labels=c('B', 'C'), label_size = 20)
-third_row = cowplot::plot_grid(tf, sm, nrow=1, labels=c('D', 'E'), label_size = 20)
-figure4 <- cowplot::plot_grid(first_row, second_row, third_row, ncol=1, rel_heights = c(1,2,2))
+second_row = cowplot::plot_grid(a, b, tf, sm, nrow=1, labels=c('B', '', 'C', ''), label_size = 20)
+figure4opt1 <- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(3,4))
+figure4opt1
+
+#first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
+#second_row = cowplot::plot_grid(a, b, nrow=1, labels=c('B', 'C'), label_size = 20)
+#third_row = cowplot::plot_grid(tf, sm, nrow=1, labels=c('D', 'E'), label_size = 20)
+#figure4opt2 <- cowplot::plot_grid(first_row, second_row, third_row, ncol=1, rel_heights = c(2.5,4,4))
+
 ```
@@
 ####Similar to QFT+, individuals with active TB, co-infected with S.
-mansoni maintain function Mtb-specific CD4+ T cell cytokine production.
-There are higher frequencies of TH1/2 cytokine co-producing CD4+ T cells amongst SM+ individuals, however the frequency of these cells is incredibly low.
+mansoni maintain function Mtb-specific CD4 T cell cytokine production.
+There are higher frequencies of TH1/2 cytokine co-producing CD4 T cells amongst SM+ individuals, however the frequency of these cells is incredibly low.
 The dominant cytokine subgroup is still TH1 cytokine producing cells.
@@
 melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
-a<-baseplot(melt, "disease", "value", tb.cols)+    
+a<-baseplot(melt, "SM", "value", tb.cols)+    
     facet_wrap(~variable, labeller = cyt_groups)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.95*.15))+
     coord_cartesian(ylim=c(0,.15))+
-    labs(y="(%) CD4+ T cells")
+    labs(y="(%) CD4 T cells")
 
@@
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*80))+
+    stat_compare_means(comparisons=list(c("SM-","SM+")), method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.98*80))+
     coord_cartesian(ylim=c(0,80))+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
+    labs(y="(%) Mtb-Specific\nTH1 Cytokine+ cells")
 
@@
     facet_wrap(~variable, labeller = groups)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3)+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
+    stat_compare_means(comparisons=list(c("SM-","SM+")), method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=.98*13)+
+    labs(y="(%) Mtb-Specific\nTH1 Cytokine+ cells")+
+    coord_cartesian(ylim=c(0,13))
 
-tbcyt<-cowplot::plot_grid(NULL, a, b, nrow=1, labels=c("A","B", "C"), label_size = 20, rel_widths = c(1.5,1,1))
-tbphen<-cowplot::plot_grid(tf, sm, both, nrow=1, labels=c("D","E","F"), label_size = 20)
-figure5<-cowplot::plot_grid(tbcyt, tbphen, nrow=2, rel_heights = c(2,3))
+tbcyt<-cowplot::plot_grid(NULL, a, b, nrow=1, labels=c("A","B", ""), label_size = 20, rel_widths = c(1.5,1,1))
+tbphen<-cowplot::plot_grid(tf, sm, both, nrow=1, labels=c("C","","D"), label_size = 20)
+figure5<-cowplot::plot_grid(tbcyt, tbphen, nrow=2, rel_heights = c(2,2))
+figure5
 ```
@@
 ```{r}
-og<-plot_filter(dplyr::filter(OG, Stim=="SEB"))
-ogphen<-plot_filter(dplyr::filter(OGPHEN, Stim=="SEB"))
-
-kable(table(og$disease), caption="Total Proliferation")
-kable(table(ogphen$disease[!is.na(ogphen$IFNg)]), caption="SEB phenotype")
-```
-
-```{r, proliferation and total cyt, fig.width=15, fig.height=8}
 groups <- as_labeller(c("OG" = "Proliferation","IFNg" = "IFNγ", 
@@
 
+og<-plot_filter(dplyr::filter(OG, Stim=="SEB"))
 part1<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG")
-og<-allplot(part1, "TB", "value", all.cols)+
+seb<-allplot(part1, "TB", "value", all.cols)+
     facet_wrap(~variable, scales="free", labeller = groups)+
-    labs(y="(%) Total CD4+ T cells")
+    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    labs(y="(%) Total CD4 T cells",
+         title="SEB")
 
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
+og<-plot_filter(dplyr::filter(OG, Stim=="PEP", TB %in% c("LTBI", "TB")))
+part2<-melt(og, id.vars=c("TB", "SM", "disease"), measure.vars="OG")
+pep<-allplot(part2, "TB", "value", all.cols)+
+    facet_wrap(~variable, scales="free", labeller = groups)+
+    scale_x_discrete(labels=c("SM- SM+\nLTBI", "SM- SM+\nTB"))+
+    coord_cartesian(ylim=c(0,4))+
+    labs(y="(%) Total CD4 T cells",
+        title="CFP-10 & ESAT-6")
 
-one<-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, labeller = groups, ncol=2)+
-    labs(y="(%) Proliferating CD4+ T cells")
+pep2<-allplot(part2, "disease", "value", all.cols)+
+    facet_wrap(~variable, scales="free", labeller = groups)+
+    stat_compare_means(comparisons=list(c("TB SM-", "LTBI SM-"), c("TB SM+", "LTBI SM+")),
+                       method="wilcox.test", size=6)+
+    coord_cartesian(ylim=c(0,35))+
+    labs(y="(%) Total CD4 T cells",
+        title="CFP-10 & ESAT-6")
 
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))
+first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
+second_row = cowplot::plot_grid(seb, pep, nrow=1, rel_widths = c(3,2), labels=c("B", "C"), label_size = 20)
+figure6a<-cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1.5,3))
 
-two<-allplot(melt, "TB", "value", all.cols)+    
-    facet_wrap(~variable, labeller = groups)
-```
+second_row = cowplot::plot_grid(seb, pep2, nrow=1, rel_widths = c(3,2), labels=c("B", "C"), label_size = 20)
+figure6b<-cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1.5,3))
 
-```{r, marker on proliferating, fig.width=15, fig.height=8}
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))
-
-tf<-allplot(melt, "TB", "value", all.cols)+    
-    labs(y="(%) Proliferating CD4+ T cells")
-
-melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))
-
-sm<-allplot(melt, "TB", "value", all.cols)
-
-first_row = cowplot::plot_grid(NULL, labels = c('A', 'B'), nrow=1, label_size = 20)
-second_row = cowplot::plot_grid(og, one, two, nrow=1, rel_widths = c(1.2,2,1), labels=c("B", "C","D"), label_size = 20)
-third_row = cowplot::plot_grid(tf, sm, labels=c("E", "F"), nrow=1, label_size = 20)
-figure6<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3)
+kable(table(og$disease), caption="Total Proliferation")
+figure6a
+figure6b
 ```
@@
 ####There are no differences in Mtb-specific proliferation betweeen SM+ and SM- QF+ individuals.
-The cytokine production from expanded Mtb-specific CD4+ T cells only differs with regards to IFNγ IL-4 co-production, however the frequency of these cells is incredibly low.
+The cytokine production from expanded Mtb-specific CD4 T cells only differs with regards to IFNγ IL-4 co-production, however the frequency of these cells is incredibly low.
 The dominant cytokine subgroup is still cells that co-produce IFNγ and TNFα.
 ```{r og and cyt, fig.width=15, fig.height=8}
-groups <- as_labeller(c("OG" = "Proliferation","IFNg" = "IFNγ", 
+groups <- as_labeller(c("IFNg" = "IFNγ", 
                     "TNFa" = "TNFα", "IL4" = "IL-4"))
 
-melt<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG")
-og<-baseplot(melt, "TB", "value", ltbi.cols)+
-    facet_wrap(~variable, scales="free", labeller = groups)+
-    coord_cartesian(ylim=c(0,3))+
-    labs(y="(%) Total CD4+ T cells")
-
 melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
-one<-baseplot(melt, "TB", "value", ltbi.cols)+    
+one<-baseplot(melt, "disease", "value", ltbi.cols)+    
     facet_wrap(~variable, labeller = groups, ncol=2)+
-    labs(y="(%) Proliferating CD4+ T cells")
+    labs(y="(%) Proliferating\nMtb-Specific CD4 T cells")
 
 melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))
-two<-baseplot(melt, "TB", "value", ltbi.cols)+    
+two<-baseplot(melt, "disease", "value", ltbi.cols)+    
     facet_wrap(~variable, labeller = groups)
@@
 ###On proliferating cells
-```{r og main four markers, fig.width=16, fig.height=8}
+```{r og main four markers, fig.width=12, fig.height=12}
 melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
@@
     facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
     coord_cartesian(ylim=c(0,100))+
-    labs(y="(%) Proliferating Mtb-Specific cells")
+    labs(y="(%) Proliferating\nMtb-Specific CD4 T cells")
 
@@
 first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
-second_row = cowplot::plot_grid(og, one, two, nrow=1, rel_widths = c(1.2,2,1), labels=c("B", "C","D"), label_size = 20)
-third_row = cowplot::plot_grid(tf, sm, labels=c("E", "F"), nrow=1, label_size = 20)
-figure7<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3)
-```
-
-###On TH1 proliferating cells
-```{r og th1 main four markers, fig.width=16, fig.height=8}
-melt<-melt(ogth1, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
-melt$variable<-factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))
-
-a<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
-    facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
-    coord_cartesian(ylim=c(0,100))+
-    labs(y="(%) Proliferating Mtb-Specific cells")
-
-b<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
-    facet_wrap(~variable)+
-    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
-    coord_cartesian(ylim=c(0,100))
-
-ogth1<-cowplot::plot_grid(a,b,ncol=2)
+second_row = cowplot::plot_grid(one, two, nrow=1, rel_widths = c(2,1), labels=c("B"), label_size = 20)
+third_row = cowplot::plot_grid(tf, sm, labels=c('C'), nrow=1, label_size = 20)
+figure7<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3, rel_heights = c(2,2,2))
+figure7
 ```
@@
 
-```{r luminex, fig.width=12, fig.height=32}
+```{r luminex, fig.width=12, fig.height=4}
 lum<-filter_ltbi(subset(luminex, luminex$Stim=="PEP"))
@@
 
-a<-baseplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+
-    labs(y="pg/mL")
+a<-baseplot(melt[melt$variable=="TNFα",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
+    coord_cartesian(ylim=c(0,1500))+
+    theme(plot.margin=margin(l=-0.5,unit="cm"))
+b<-baseplot(melt[melt$variable=="IFNγ",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
+    coord_cartesian(ylim=c(0,18000))+
+    theme(plot.margin=margin(l=-0.5,unit="cm"))
+c<-baseplot(melt[melt$variable=="IL-21",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
+    coord_cartesian(ylim=c(0,60))+
+    theme(plot.margin=margin(l=-0.5,unit="cm"))
+d<-baseplot(melt[melt$variable=="IL-17A",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
+    coord_cartesian(ylim=c(0,20))+
+    theme(plot.margin=margin(l=-0.5,unit="cm"))
+e<-baseplot(melt[melt$variable=="IL-22",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
+    coord_cartesian(ylim=c(0,40))+
+    theme(plot.margin=margin(l=-0.5,unit="cm"))
+f<-baseplot(melt[melt$variable=="IL-10",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
+    coord_cartesian(ylim=c(0,20))+
+    theme(plot.margin=margin(l=-0.5,unit="cm"))
+g<-baseplot(melt[melt$variable=="IL-4",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
+    coord_cartesian(ylim=c(0,40))+
+    theme(plot.margin=margin(l=-0.5,unit="cm"))
+h<-baseplot(melt[melt$variable=="IL-5",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
+    coord_cartesian(ylim=c(0,15))+
+    theme(plot.margin=margin(l=-0.5,unit="cm"))
+i<-baseplot(melt[melt$variable=="IL-13",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
+    coord_cartesian(ylim=c(0,400))+
+    theme(plot.margin=margin(l=-0.5,unit="cm"))
 
-b<-baseplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+
-    labs(y="pg/mL")+
-    coord_cartesian(ylim=c(0,40))
-
-lumplot<-cowplot::plot_grid(a,b, NULL, ncol=1, align = "v", labels=c("A","A", "B"), label_size = 20)
+grid<-cowplot::plot_grid(a,b,c,d,e,f,g,h,i, nrow=1, align="hv")
+lumplot<-cowplot::plot_grid(grid, NULL, ncol=1, align = "v", labels=c("A", "B"), rel_heights=c(3,1), label_size = 20)
+lumplot
 ```
@@
 ```
-#Exporting Data
-```{r}
-setwd("/Users/tarynam/Desktop/Manuscript1")
-cowplot::save_plot("figure1.png", figure1, base_height=8, base_width=16) #Figure 1
-cowplot::save_plot("figure2.png", figure2, base_height=14, base_width=16) #Figure 2
-cowplot::save_plot("figure3.png", figure3, base_height=14, base_width=16)
-cowplot::save_plot("figure4.png", figure4, base_height=14, base_width=16)
-cowplot::save_plot("figure5.png", figure5, base_height=14, base_width=16)
-cowplot::save_plot("figure6.png", figure6, base_height=16, base_width=16)
-cowplot::save_plot("figure7.png", figure7, base_height=16, base_width=16)
-cowplot::save_plot("figure8.png", lumplot, base_height=32, base_width=16)
-```
-
-##ICS Supplementary
-```{r}
-booleanplot2<-function(data, xval, yval, colors){
-    ggplot(data, aes(data[,xval], data[,yval], fill=disease))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    scale_fill_manual(values = colors)+
-    theme_classic()+ 
-    theme(legend.position="none")+
-    theme(text = element_text(size=20), axis.text.x = element_text(size=20))+
-    theme(plot.subtitle = element_text(hjust = 0.5)) +
-    labs(x="", y="")
-}
-```
-
-###Boolean PMA
-*but I need to make them pretty and I dont' feel like doing that right now*
-```{r boolean pma cyt, fig.width=15, fig.height=12}
-cyt<-plot_filter(CYT[CYT$Stim=="PMA",])
-test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
-test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
-test$variable<-factor(test$variable, levels=c(
-"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
-"G_x_x_T" , "G_x_x_x","x_x_x_T",
-"x_4_13_x","x_4_x_x", "x_x_13_x"
-))
-
-G   <-c("+", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
-il4 <-c("+", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
-il13<-c("+", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
-T   <-c("+", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")
-
-labels <- paste0(G, "\n", T, "\n", il4, "\n", il13)
-breaks <- levels(test$variable)
-
-test$th<-"TH1/2"
-test$th[grep("._x_x_.", test$variable)]<-"TH1"
-test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
-test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))
-
-th1.2<-booleanplot2(test[test$th=="TH1/2",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
-    coord_cartesian(ylim=c(0,3))+
-    theme(strip.text.y = element_blank(), strip.background.y = element_blank())+
-    labs(y="(%) CD4+ T cells")
-
-th1<-booleanplot2(test[test$th=="TH1",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
-    coord_cartesian(ylim=c(0,25))+
-    theme(strip.text.y = element_blank(), strip.background.y = element_blank())
-
-th2<-booleanplot2(test[test$th=="TH2",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
-    coord_cartesian(ylim=c(0,3))
-
-supp1<-cowplot::plot_grid(th1.2, th1, th2, ncol=3, rel_widths = c(8,3.5,3.5))
-```
-
-###Boolean Phenotypes
-###TH1
-```{r pma th1 boolean, fig.width=20, fig.height=18}
-phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))
-
-melt<-melt(phen, id.vars=c("TB", "SM", "disease"), measure.vars = grep("TH1_._._._.", names(phen)))
-test<-filter(melt, variable!="TH1_x_x_x_x")
-test$variable<-gsub("TH1_", "", test$variable)
-test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
-
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
-
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
-
-df<-data.frame(cbind(levels(test$variable),T,G,c3,c4))
-dfm<-melt(df, measure.vars=c("T", "G","c3","c4"))
-dfm$variable<-factor(dfm$variable, levels=c("c4","c3","G", "T"))
-
-data_table <- ggplot(dfm, aes(x = V1, y = variable, label = format(value))) +
-    geom_text(size = 10) + 
-    theme_bw() + 
-    labs(x="",y="")+
-    theme(strip.text.x = element_blank(), strip.background.x = element_blank())+
-    scale_y_discrete(labels=c("T-bet", "GATA3", "CXCR3", "CCR4"))+
-    theme(axis.text.y = element_text(size=20))+
-    theme(panel.grid.major = element_blank(), 
-          panel.grid.minor = element_blank(),
-          panel.border = element_blank(),
-          axis.title = element_blank(),
-          axis.text.x = element_blank(),
-          axis.ticks = element_blank())
-
-test$tf<-NA
-test$tf[grep("x_g", test$variable)]<-"GATA3"
-test$tf[grep("t_x", test$variable)]<-"T-bet"
-test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
-test$tf[grep("x_x_._.", test$variable)]<-"None"
-test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
-
-supp2a<-booleanplot2(test, "variable", "value", all.cols)+
-    facet_grid(TB~tf, scales="free_x")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
-    coord_cartesian(ylim=c(0,65))
-
-```
-###TH2  
-```{r pma th2 boolean, fig.width=20, fig.height=18}
-phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))
-
-melt<-melt(phen, id.vars=c("TB", "SM", "disease"), measure.vars = grep("TH2_._._._.", names(phen)))
-test<-filter(melt, variable!="TH2_x_x_x_x")
-test$variable<-gsub("TH2_", "", test$variable)
-test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
-
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
-
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
-
-test$tf<-NA
-test$tf[grep("x_g", test$variable)]<-"GATA3"
-test$tf[grep("t_x", test$variable)]<-"T-bet"
-test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
-test$tf[grep("x_x_._.", test$variable)]<-"None"
-test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
-
-supp2b<-booleanplot2(test, "variable", "value", all.cols)+
-    facet_grid(TB~tf, scales="free_x")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
-    coord_cartesian(ylim=c(0,65))
-```
-
-```{r}
-supp2<-cowplot::plot_grid(supp2a, supp2b, nrow=2, labels=c("A", "B"))
-```
-
-###Peptide-Specific Cytokine Production  
-```{r, fig.width=15, fig.height=8}
-melt<-melt(subset(CYT, CYT$Stim=="PEP"), 
-           id.vars=c("TB","SM"), 
-           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
-th_groups <- as_labeller(c("TH1.Freq" = "TH1", 
-                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))
-
-cyt<-filter_ltbi(melt)
-b<-baseplot(cyt, "SM", "value", ltbi.cols)+
-    facet_wrap(~variable, ncol=4, labeller=th_groups)+
-    coord_cartesian(ylim=c(0,.2))
-
-cyt<-filter_tb(melt)
-c<-baseplot(cyt, "SM", "value", tb.cols)+
-    facet_wrap(~variable, ncol=4, labeller=th_groups)+
-    coord_cartesian(ylim=c(0,.2))
-
-supp3<-cowplot::plot_grid(b, c, ncol=3)
-```
-
-```{r, fig.width=16, fig.height=8}
-cyt<-plot_filter(CYT[CYT$Stim=="PEP",])
-test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
-test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
-test$variable<-factor(test$variable, levels=c(
-"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
-"G_x_x_T" , "G_x_x_x","x_x_x_T",
-"x_4_13_x","x_4_x_x", "x_x_13_x"
-))
-
-G   <-c("IFNγ +", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
-il4 <-c("IL-4 +", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
-il13<-c("IL-13 +", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
-T   <-c("TNFα +", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")
-
-labels <- paste0(G, "\n", T, "\n", il4, "\n", il13)
-breaks <- levels(test$variable)
-
-test$th<-"TH1/2"
-test$th[grep("._x_x_.", test$variable)]<-"TH1"
-test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
-test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))
-
-supp4a<-booleanplot2(test[test$TB=="LTBI",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
-    coord_cartesian(ylim=c(0,.16))+
-    labs(y="(%) CD4+ T cells")
-
-supp5a<-booleanplot2(test[test$TB=="TB",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
-    coord_cartesian(ylim=c(0,.15))+
-    labs(y="(%) CD4+ T cells")
-```
-
-###Peptide Mtb-specific TH1 Cytokine+ Cell Phenotypes
-```{r ics phenotyping, fig.width=16, fig.height=8}
-phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
-
-melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
-test<-filter(melt, variable!="TH1_x_x_x_x")
-test$variable<-gsub("TH1_", "", test$variable)
-test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
-
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
-
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
-
-test$tf<-NA
-test$tf[grep("x_g", test$variable)]<-"GATA3"
-test$tf[grep("t_x", test$variable)]<-"T-bet"
-test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
-test$tf[grep("x_x_._.", test$variable)]<-"None"
-test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
-
-supp4b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x",ncol=4)+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
-```
-
-```{r, fig.width=15, fig.height=8}
-phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
-
-#graph phenotype
-melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
-test<-filter(melt, variable!="TH1_x_x_x_x")
-test$variable<-gsub("TH1_", "", test$variable)
-test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
-
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
-
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
-
-test$tf<-NA
-test$tf[grep("x_g", test$variable)]<-"GATA3"
-test$tf[grep("t_x", test$variable)]<-"T-bet"
-test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
-test$tf[grep("x_x_._.", test$variable)]<-"None"
-test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
-
-supp5b<-booleanplot(test, "variable", "value", tb.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")+
-    coord_cartesian(ylim=c(0,55))+
-    stat_compare_means(method="wilcox.test",label="p.signif", hide.ns=TRUE, label.y=50, size=10)
-```
-
-###Comparing Mtb-specific TH1 cells to total CD4 T cells in peptide stim  
-```{r, fig.width=12, fig.height=6}
-total<-filter_ltbi(subset(CD4, CD4$Stim=="UN"))
-total$cell<-"Total"
-phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
-names<-c("Donor", "TB", "SM", "Stim", "disease", names(phen)[grep("TH1_", names(phen))])
-mtb_th1<-select(phen, names)
-names(mtb_th1)<-gsub("TH1_", "", names(mtb_th1))
-mtb_th1$cell<-"Mtb-Specific"
-long<-rbindlist(list(total, mtb_th1), fill=TRUE)
-
-melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
-melt$cell<-factor(melt$cell, levels=c("Total", "Mtb-Specific"))
-melt$x<-paste(melt$SM, melt$cell, sep=" ")
-melt$x<-factor(melt$x, levels=c("SM- Total", "SM- Mtb-Specific","SM+ Total", "SM+ Mtb-Specific"))
-
-ggplot(melt, aes(x=x, y=value, fill=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    facet_wrap(~variable, scales="free", nrow=1)+
-    theme_classic()+
-    scale_fill_manual(values = ltbi.cols)+
-    scale_x_discrete(labels=c("Total", "Mtb", "Total", "Mtb"))+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
-    stat_compare_means(comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")) 
-                       ,method="wilcox.test", label="p.format")+
-    labs(x="", y="")
-```
-
-###Comparing Mtb-specific TH1 cells to PMA TH1 cells  
-```{r, fig.width=12, fig.height=6}
-pma<-filter_ltbi(subset(PHEN, PHEN$Stim=="PMA"))
-pma$cell<-"PMA"
-pep<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
-pep$cell<-"Mtb"
-long<-rbindlist(list(pma, pep), fill=TRUE)
-names<-c("Donor", "TB", "SM", "Stim", "disease", "cell", names(long)[grep("TH1_", names(long))])
-long<-select(long, names)
-names(long)<-gsub("TH1_", "", names(long))
-
-melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
-melt$x<-paste(melt$SM, melt$cell, sep=" ")
-melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))
-
-ggplot(melt, aes(x=x, y=value, fill=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    facet_wrap(~variable, scales="free", nrow=1)+
-    theme_classic()+
-    scale_fill_manual(values = ltbi.cols)+
-    scale_x_discrete(labels=c("PMA", "Mtb", "PMA", "Mtb"))+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
-    stat_compare_means(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")) 
-                       ,method="wilcox.test", label="p.format")+
-    labs(x="", y="")
-```
-
-###Comparing Mtb-specific TH1 cells to total CD4 T cells in peptide stim  
-```{r, fig.width=12, fig.height=6}
-total<-filter_tb(subset(CD4, CD4$Stim=="UN"))
-total$cell<-"Total"
-phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
-names<-c("Donor", "TB", "SM", "Stim", "disease", names(phen)[grep("TH1_", names(phen))])
-mtb_th1<-select(phen, names)
-names(mtb_th1)<-gsub("TH1_", "", names(mtb_th1))
-mtb_th1$cell<-"Mtb-Specific"
-long<-rbindlist(list(total, mtb_th1), fill=TRUE)
-
-melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
-melt$cell<-factor(melt$cell, levels=c("Total", "Mtb-Specific"))
-melt$x<-paste(melt$SM, melt$cell, sep=" ")
-melt$x<-factor(melt$x, levels=c("SM- Total", "SM- Mtb-Specific","SM+ Total", "SM+ Mtb-Specific"))
-
-ggplot(melt, aes(x=x, y=value, fill=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    facet_wrap(~variable, scales="free", nrow=1)+
-    theme_classic()+
-    scale_fill_manual(values = tb.cols)+
-    scale_x_discrete(labels=c("Total", "Mtb", "Total", "Mtb"))+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
-    stat_compare_means(comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")) 
-                       ,method="wilcox.test", label="p.format")+
-    labs(x="", y="")
-
-```
-
-###Comparing Mtb-specific TH1 cells to PMA TH1 cells  
-```{r, fig.width=12, fig.height=6}
-pma<-filter_tb(subset(PHEN, PHEN$Stim=="PMA"))
-pma$cell<-"PMA"
-pep<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
-pep$cell<-"Mtb"
-long<-rbindlist(list(pma, pep), fill=TRUE)
-names<-c("Donor", "TB", "SM", "Stim", "disease", "cell", names(long)[grep("TH1_", names(long))])
-long<-select(long, names)
-names(long)<-gsub("TH1_", "", names(long))
-
-melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
-melt$x<-paste(melt$SM, melt$cell, sep=" ")
-melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))
-
-ggplot(melt, aes(x=x, y=value, fill=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    facet_wrap(~variable, scales="free", nrow=1)+
-    theme_classic()+
-    scale_fill_manual(values = tb.cols)+
-    scale_x_discrete(labels=c("PMA", "Mtb", "PMA", "Mtb"))+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
-    stat_compare_means(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")) 
-                       ,method="wilcox.test", label="p.format")+
-    labs(x="", y="")
-```
-
-```{r}
-supp4<-cowplot::plot_grid(supp4a, supp4b, nrow=2, labels=c("A", "B"))
-supp5<-cowplot::plot_grid(supp5a, supp5b, nrow=2, labels=c("A", "B"))
-```
-
-##Proliferation Supplementary
-```{r}
-ogphen<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
-melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._.", names(ogphen)))
-test<-melt[-grep("._._._.", melt$variable),]
-test<-filter(test, variable!="x_x_x")
-test$variable<-factor(test$variable, levels=c(
-"T_G_x" ,       
-"T_x_x" ,         
-"x_G_x" ,         
-"T_G_4" ,
-"T_x_4" ,         
-"x_G_4" ,         
-"x_x_4"          
-))
-test$group<-"TH1/2"
-for(i in 1:length(test$group)){
-    if(test$variable[i] %in% c("T_G_x", "T_x_x", "x_G_x")){
-        test$group[i]<- "TH1"
-    }
-    else if (test$variable[i] == "x_x_4"){
-        test$group[i]<- "TH2"
-    }
-}
-
-T <- c("+", "+", "-", "+", "+", "-", "-")
-G <- c("+", "-", "+", "+", "-", "+", "-")
-il4 <- c("-", "-", "-", "+", "+", "+", "+")
-
-labels <- paste0(T, "\n", G, "\n", il4)
-breaks <- levels(test$variable)
-
-supp6a<-booleanplot(test, "variable", "value", ltbi.cols)+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    facet_grid(~group, scales="free_x", space="free_x")+
-    labs(y="(%) Proliferating CD4+ T cells")
-
-```
-
-```{r og phenotype, fig.width=16, fig.height=8}
-#graph phenotype
-melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._._.", names(ogphen)))
-test<-filter(melt, variable!="x_x_x_x")
-test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
-
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
-
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
-
-test$tf<-NA
-test$tf[grep("x_g", test$variable)]<-"GATA3"
-test$tf[grep("t_x", test$variable)]<-"T-bet"
-test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
-test$tf[grep("x_x_._.", test$variable)]<-"None"
-test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
-
-supp6b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=6)+
-    labs(y="(%) Proliferating Mtb-Specific cells")
-```
-
-```{r}
-supp6<-cowplot::plot_grid(supp6a, supp6b, nrow=2, labels=c("A", "B"), rel_widths = c(3,4))
-```
-
-###Comparing Mtb-specific TH1 cells to PMA TH1 cells  
-```{r, fig.width=12, fig.height=6}
-un<-filter_ltbi(subset(OG, OG$Stim=="UN"))
-un$cell<-"total"
-seb<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="SEB"))
-seb$cell<-"seb"
-long<-rbindlist(list(un, seb), fill=TRUE)
-
-melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
-melt$x<-paste(melt$SM, melt$cell, sep=" ")
-melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))
-
-ggplot(long, aes(x=cell, y=long[,"T-bet"], color=disease))+
-    geom_point(size=3, alpha=0.5)+geom_line(aes(group=Donor))+
-    facet_wrap(~TB, scales="free", nrow=1)+
-    theme_classic()+
-    scale_fill_manual(values = ltbi.cols)+
-    scale_x_discrete(labels=c("PMA", "Mtb", "PMA", "Mtb"))+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
-    stat_compare_means(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")) 
-                       ,method="wilcox.test", label="p.format")+
-    labs(x="", y="")
-```
-
-```{r}
-lum_x<-lum[lum$SM=="SM-",]
-df_x<-as.matrix(lum_x[,3:11])
-row.names(df_x)<-lum_x$Sample
-hc_x <- hclust(as.dist(1-cor(df_x)))
-x<-plot(hc_x)
-
-lum_sm<-lum[lum$SM=="SM+",]
-df_sm<-as.matrix(lum_sm[,3:11])
-row.names(df_sm)<-lum_sm$Sample
-hc_sm <- hclust(as.dist(1-cor(df_sm)))
-sm<-plot(hc_sm)
-
-df<-as.matrix(lum[,3:11])
-#heatmap(t(df_x), Rowv=as.dendrogram(hc_x))
-#heatmap(t(df_sm), Rowv=as.dendrogram(hc_sm))
-#heatmap(t(df), Rowv=as.dendrogram(hc_sm))
-```
 
@@
 setwd("/Users/tarynam/Desktop/Manuscript1")
-cowplot::save_plot("supp1.png", supp1, base_height=14, base_width=16)
-cowplot::save_plot("supp2.png", supp2, base_height=28, base_width=16)
-cowplot::save_plot("supp3.png", supp3, base_height=7, base_width=16)
-cowplot::save_plot("supp4.png", supp4, base_height=14, base_width=16)
-cowplot::save_plot("supp5.png", supp5, base_height=14, base_width=16)
-cowplot::save_plot("supp6.png", supp6, base_height=14, base_width=16)
+
+cowplot::save_plot("figure1.png", figure1, base_height=8, base_width=16) #Figure 1
+#cowplot::save_plot("figure1alt1.png", figure1B, base_height=8, base_width=16) #Figure 1
+cowplot::save_plot("figure1alt2.png", figure1alt, base_height=8, base_width=16) #Figure 1
+cowplot::save_plot("figure2.png", figure2, base_height=14, base_width=16) #Figure 2
+cowplot::save_plot("figure3.png", figure3, base_height=16, base_width=16)
+
+cowplot::save_plot("figure4.png", figure4opt1, base_height=10, base_width=16)
+
+cowplot::save_plot("figure5.png", figure5, base_height=11, base_width=16)
+cowplot::save_plot("figure6a.png", figure6a, base_height=9, base_width=9)
+cowplot::save_plot("figure6b.png", figure6b, base_height=9, base_width=9)
+cowplot::save_plot("figure7.png", figure7, base_height=14, base_width=12)
+
+cowplot::save_plot("figure8.png", lumplot, base_height=6, base_width=18)
 ```

### 2019-09-28 SuppFigures.Rmd
separated main figures from supplementary figures

---
title: "supplementary figures"
author: "Taryn McLaughlin"
date: "9/26/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, comment=NA, warning=FALSE, echo=FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
library(tidyr)
library(reshape2)
library(cowplot)
library(corrplot)
library(sjPlot)
```

```{r plot functions}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")

baseplot<-function(data, xval, yval, colors){
    ggplot(data, aes(data[,xval], data[,yval], fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    scale_fill_manual(values = colors)+
    theme_classic()+ theme(legend.position="none")+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=20),
          axis.text.x = element_text(size = 16),
          strip.text = element_text(size = 20))+
    theme(plot.title = element_text(hjust = 0.5, size=20)) +
    labs(x="", y="")
}

booleanplot<-function(data, xval, yval, colors){
    ggplot(data, aes(data[,xval], data[,yval], fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    scale_fill_manual(values = colors)+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(text = element_text(size=20), 
          axis.text.x = element_text(size=16), 
          strip.text = element_text(size = 20))+
    theme(plot.title = element_text(hjust = 0.5, size=20)) +
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
    labs(x="", y="")
}

allplot<-function(data, xval, yval, colors){
    ggplot(data, aes(data[,xval], data[,yval], fill=disease))+
    geom_boxplot(size=1, position=position_dodge2(width=0.2, padding = 0.2, preserve = "single"), outlier.shape = NA)+
    scale_fill_manual(values = colors)+
    theme_classic()+ 
    theme(legend.position="none") +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    facet_wrap(~variable)+
    theme(text = element_text(size=20),
            axis.text.x = element_text(size = 16),
            strip.text = element_text(size = 20))+
    theme(plot.title = element_text(hjust = 0.5, size=20))+
    labs(x="", y="")
}

cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))

all_comps<-list(c("HC SM+", "HC SM-"),c("LTBI SM+", "LTBI SM-"), c("TB SM+", "TB SM-"))

healthy.cols<- c("SM+" = "#1a9850", "SM-" ="#91cf60")
ltbi.cols<- c("SM+" = "#2166ac", "SM-" = "#85ABD1")
tb.cols<- c("SM+" = "#b2182b", "SM-"="#E59EA6")
all.cols<- c("Naive" = "#C8D0D8","HC SM+" = "#1a9850", "HC SM-" ="#91cf60",
             "LTBI SM+" = "#2166ac", "LTBI SM-" = "#85ABD1",
             "TB SM+" = "#b2182b", "TB SM-"="#E59EA6")
```

```{r load data}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
CD4<-read.csv("ICS_CD4_clean.csv")
CYT<-read.csv("ICS_CD4_Cytokine_clean.csv")
PHEN<-read.csv("ICS_CD4_phenotyping_clean_NAs.csv")
OG<-data.table(read.csv("OG_CD4_bulk_summary_clean.csv"))
OGPHEN<-read.csv("OG_CD4_OG_summary_clean_NAs.csv")
OGnope<-read.csv("OG_CD4_noOG_summary_clean.csv")
OG.IFNg<-read.csv("OG_CD4_OG.IFNg_summary_clean_NAs.csv")
OG.GT<-read.csv("OG_CD4_OG.GT_summary_clean_NAs.csv")
OG.th1<-read.csv("OG_CD4_OG.TH1_summary_clean_NAs.csv")
luminex<-read.csv("/Applications/Old Computer/Day Lab/Luminex/Clean Data/Luminex_combined_background.txt")
```

##ICS Supplementary
```{r}
booleanplot2<-function(data, xval, yval, colors){
    ggplot(data, aes(data[,xval], data[,yval], fill=disease))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    scale_fill_manual(values = colors)+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(size=20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    labs(x="", y="")
}
```

###Boolean PMA
*but I need to make them pretty and I dont' feel like doing that right now*
```{r boolean pma cyt, fig.width=15, fig.height=12}
cyt<-plot_filter(CYT[CYT$Stim=="PMA",])
test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
test$variable<-factor(test$variable, levels=c(
"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
"G_x_x_T" , "G_x_x_x","x_x_x_T",
"x_4_13_x","x_4_x_x", "x_x_13_x"
))

G   <-c("+", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
il4 <-c("+", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
il13<-c("+", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
T   <-c("+", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")

labels <- paste0(G, "\n", T, "\n", il4, "\n", il13)
breaks <- levels(test$variable)

test$th<-"TH1/2"
test$th[grep("._x_x_.", test$variable)]<-"TH1"
test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))

th1.2<-booleanplot2(test[test$th=="TH1/2",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
    coord_cartesian(ylim=c(0,3))+
    theme(strip.text.y = element_blank(), strip.background.y = element_blank())+
    labs(y="(%) CD4 T cells")

th1<-booleanplot2(test[test$th=="TH1",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
    coord_cartesian(ylim=c(0,25))+
    theme(strip.text.y = element_blank(), strip.background.y = element_blank())

th2<-booleanplot2(test[test$th=="TH2",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
    coord_cartesian(ylim=c(0,3))

supp1<-cowplot::plot_grid(th1.2, th1, th2, ncol=3, rel_widths = c(8,3.5,3.5))
```

###Boolean Phenotypes
###TH1
```{r pma th1 boolean, fig.width=20, fig.height=18}
phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))

melt<-melt(phen, id.vars=c("TB", "SM", "disease"), measure.vars = grep("TH1_._._._.", names(phen)))
test<-filter(melt, variable!="TH1_x_x_x_x")
test$variable<-gsub("TH1_", "", test$variable)
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" ,       
"t_x_3_x" ,         
"t_x_x_4" ,         
"t_x_x_x" ,
"t_g_3_4" ,         
"t_g_3_x" ,         
"t_g_x_4" ,         
"t_g_x_x" ,
"x_g_3_4" ,         
"x_g_3_x" ,         
"x_g_x_4" ,         
"x_g_x_x" ,
"x_x_3_4" ,          
"x_x_3_x" ,         
"x_x_x_4"           
))

T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
breaks <- levels(test$variable)

df<-data.frame(cbind(levels(test$variable),T,G,c3,c4))
dfm<-melt(df, measure.vars=c("T", "G","c3","c4"))
dfm$variable<-factor(dfm$variable, levels=c("c4","c3","G", "T"))

data_table <- ggplot(dfm, aes(x = V1, y = variable, label = format(value))) +
    geom_text(size = 10) + 
    theme_bw() + 
    labs(x="",y="")+
    theme(strip.text.x = element_blank(), strip.background.x = element_blank())+
    scale_y_discrete(labels=c("T-bet", "GATA3", "CXCR3", "CCR4"))+
    theme(axis.text.y = element_text(size=20))+
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks = element_blank())

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

supp2a<-booleanplot2(test, "variable", "value", all.cols)+
    facet_grid(TB~tf, scales="free_x")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
    coord_cartesian(ylim=c(0,65))

```
###TH2  
```{r pma th2 boolean, fig.width=20, fig.height=18}
phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))

melt<-melt(phen, id.vars=c("TB", "SM", "disease"), measure.vars = grep("TH2_._._._.", names(phen)))
test<-filter(melt, variable!="TH2_x_x_x_x")
test$variable<-gsub("TH2_", "", test$variable)
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" ,       
"t_x_3_x" ,         
"t_x_x_4" ,         
"t_x_x_x" ,
"t_g_3_4" ,         
"t_g_3_x" ,         
"t_g_x_4" ,         
"t_g_x_x" ,
"x_g_3_4" ,         
"x_g_3_x" ,         
"x_g_x_4" ,         
"x_g_x_x" ,
"x_x_3_4" ,          
"x_x_3_x" ,         
"x_x_x_4"           
))

T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
breaks <- levels(test$variable)

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

supp2b<-booleanplot2(test, "variable", "value", all.cols)+
    facet_grid(TB~tf, scales="free_x")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
    coord_cartesian(ylim=c(0,65))
```

```{r}
supp2<-cowplot::plot_grid(supp2a, supp2b, nrow=2, labels=c("A", "B"))
```

###Peptide-Specific Cytokine Production  
```{r, fig.width=15, fig.height=8}
melt<-melt(subset(CYT, CYT$Stim=="PEP"), 
           id.vars=c("TB","SM"), 
           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
th_groups <- as_labeller(c("TH1.Freq" = "TH1", 
                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))

cyt<-filter_ltbi(melt)
b<-baseplot(cyt, "SM", "value", ltbi.cols)+
    facet_wrap(~variable, ncol=4, labeller=th_groups)+
    coord_cartesian(ylim=c(0,.2))

cyt<-filter_tb(melt)
c<-baseplot(cyt, "SM", "value", tb.cols)+
    facet_wrap(~variable, ncol=4, labeller=th_groups)+
    coord_cartesian(ylim=c(0,.2))

supp3<-cowplot::plot_grid(b, c, ncol=3)
```

```{r, fig.width=16, fig.height=8}
cyt<-plot_filter(CYT[CYT$Stim=="PEP",])
test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
test$variable<-factor(test$variable, levels=c(
"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
"G_x_x_T" , "G_x_x_x","x_x_x_T",
"x_4_13_x","x_4_x_x", "x_x_13_x"
))

G   <-c("IFNγ +", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
il4 <-c("IL-4 +", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
il13<-c("IL-13 +", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
T   <-c("TNFα +", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")

labels <- paste0(G, "\n", T, "\n", il4, "\n", il13)
breaks <- levels(test$variable)

test$th<-"TH1/2"
test$th[grep("._x_x_.", test$variable)]<-"TH1"
test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))

supp4a<-booleanplot2(test[test$TB=="LTBI",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
    coord_cartesian(ylim=c(0,.16))+
    labs(y="(%) CD4 T cells")

supp5a<-booleanplot2(test[test$TB=="TB",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
    coord_cartesian(ylim=c(0,.15))+
    labs(y="(%) CD4 T cells")
```

###Peptide Mtb-specific TH1 Cytokine+ Cell Phenotypes
```{r ics phenotyping, fig.width=16, fig.height=24}
phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))

melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
test<-filter(melt, variable!="TH1_x_x_x_x")
test$variable<-gsub("TH1_", "", test$variable)
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" ,       
"t_x_3_x" ,         
"t_x_x_4" ,         
"t_x_x_x" ,
"t_g_3_4" ,         
"t_g_3_x" ,         
"t_g_x_4" ,         
"t_g_x_x" ,
"x_g_3_4" ,         
"x_g_3_x" ,         
"x_g_x_4" ,         
"x_g_x_x" ,
"x_x_3_4" ,          
"x_x_3_x" ,         
"x_x_x_4"           
))

T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
breaks <- levels(test$variable)

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

supp4b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x",ncol=4)+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
```

```{r, fig.width=15, fig.height=24}
phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))

#graph phenotype
melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
test<-filter(melt, variable!="TH1_x_x_x_x")
test$variable<-gsub("TH1_", "", test$variable)
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" ,       
"t_x_3_x" ,         
"t_x_x_4" ,         
"t_x_x_x" ,
"t_g_3_4" ,         
"t_g_3_x" ,         
"t_g_x_4" ,         
"t_g_x_x" ,
"x_g_3_4" ,         
"x_g_3_x" ,         
"x_g_x_4" ,         
"x_g_x_x" ,
"x_x_3_4" ,          
"x_x_3_x" ,         
"x_x_x_4"           
))

T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
breaks <- levels(test$variable)

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

supp5b<-booleanplot(test, "variable", "value", tb.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
    scale_x_discrete(labels = labels, breaks=breaks)+
    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")+
    coord_cartesian(ylim=c(0,55))+
    stat_compare_means(method="wilcox.test",label="p.signif", hide.ns=TRUE, label.y=50, size=10)
```

###Comparing Mtb-specific TH1 cells to total CD4 T cells in peptide stim  
```{r, fig.width=12, fig.height=6}
total<-filter_ltbi(subset(CD4, CD4$Stim=="UN"))
total$cell<-"Total"
phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
names<-c("Donor", "TB", "SM", "Stim", "disease", names(phen)[grep("TH1_", names(phen))])
mtb_th1<-select(phen, names)
names(mtb_th1)<-gsub("TH1_", "", names(mtb_th1))
mtb_th1$cell<-"Mtb-Specific"
long<-rbindlist(list(total, mtb_th1), fill=TRUE)

melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
melt$cell<-factor(melt$cell, levels=c("Total", "Mtb-Specific"))
melt$x<-paste(melt$SM, melt$cell, sep=" ")
melt$x<-factor(melt$x, levels=c("SM- Total", "SM- Mtb-Specific","SM+ Total", "SM+ Mtb-Specific"))

ggplot(melt, aes(x=x, y=value, fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    facet_wrap(~variable, scales="free", nrow=1)+
    theme_classic()+
    scale_fill_manual(values = ltbi.cols)+
    scale_x_discrete(labels=c("Total", "Mtb", "Total", "Mtb"))+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")) 
                       ,method="wilcox.test", label="p.format")+
    labs(x="", y="")
```

###Comparing Mtb-specific TH1 cells to PMA TH1 cells  
```{r, fig.width=12, fig.height=6}
pma<-filter_ltbi(subset(PHEN, PHEN$Stim=="PMA"))
pma$cell<-"PMA"
pep<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
pep$cell<-"Mtb"
long<-rbindlist(list(pma, pep), fill=TRUE)
names<-c("Donor", "TB", "SM", "Stim", "disease", "cell", names(long)[grep("TH1_", names(long))])
long<-select(long, names)
names(long)<-gsub("TH1_", "", names(long))

melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
melt$x<-paste(melt$SM, melt$cell, sep=" ")
melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))

ggplot(melt, aes(x=x, y=value, fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    facet_wrap(~variable, scales="free", nrow=1)+
    theme_classic()+
    scale_fill_manual(values = ltbi.cols)+
    scale_x_discrete(labels=c("PMA", "Mtb", "PMA", "Mtb"))+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")) 
                       ,method="wilcox.test", label="p.format")+
    labs(x="", y="")
```

###Comparing Mtb-specific TH1 cells to total CD4 T cells in peptide stim  
```{r, fig.width=12, fig.height=6}
total<-filter_tb(subset(CD4, CD4$Stim=="UN"))
total$cell<-"Total"
phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
names<-c("Donor", "TB", "SM", "Stim", "disease", names(phen)[grep("TH1_", names(phen))])
mtb_th1<-select(phen, names)
names(mtb_th1)<-gsub("TH1_", "", names(mtb_th1))
mtb_th1$cell<-"Mtb-Specific"
long<-rbindlist(list(total, mtb_th1), fill=TRUE)

melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
melt$cell<-factor(melt$cell, levels=c("Total", "Mtb-Specific"))
melt$x<-paste(melt$SM, melt$cell, sep=" ")
melt$x<-factor(melt$x, levels=c("SM- Total", "SM- Mtb-Specific","SM+ Total", "SM+ Mtb-Specific"))

ggplot(melt, aes(x=x, y=value, fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    facet_wrap(~variable, scales="free", nrow=1)+
    theme_classic()+
    scale_fill_manual(values = tb.cols)+
    scale_x_discrete(labels=c("Total", "Mtb", "Total", "Mtb"))+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")) 
                       ,method="wilcox.test", label="p.format")+
    labs(x="", y="")

```

###Comparing Mtb-specific TH1 cells to PMA TH1 cells  
```{r, fig.width=12, fig.height=6}
pma<-filter_tb(subset(PHEN, PHEN$Stim=="PMA"))
pma$cell<-"PMA"
pep<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
pep$cell<-"Mtb"
long<-rbindlist(list(pma, pep), fill=TRUE)
names<-c("Donor", "TB", "SM", "Stim", "disease", "cell", names(long)[grep("TH1_", names(long))])
long<-select(long, names)
names(long)<-gsub("TH1_", "", names(long))

melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
melt$x<-paste(melt$SM, melt$cell, sep=" ")
melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))

ggplot(melt, aes(x=x, y=value, fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    facet_wrap(~variable, scales="free", nrow=1)+
    theme_classic()+
    scale_fill_manual(values = tb.cols)+
    scale_x_discrete(labels=c("PMA", "Mtb", "PMA", "Mtb"))+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")) 
                       ,method="wilcox.test", label="p.format")+
    labs(x="", y="")
```

```{r, fig.width=12, fig.height=6}
supp4<-cowplot::plot_grid(supp4a, supp4b, nrow=2, labels=c("A", "B"))
supp5<-cowplot::plot_grid(supp5a, supp5b, nrow=2, labels=c("A", "B"))
supp1
supp2
supp3
supp4
supp5
```

##Proliferation Supplementary
```{r}
ogphen<-plot_filter(subset(OGPHEN, OGPHEN$Stim=="SEB"))
groups <- as_labeller(c("OG" = "Proliferation","IFNg" = "IFNγ","TNFa" = "TNFα", "IL4" = "IL-4"))
#cytokines
melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
one<-allplot(melt, "TB", "value", all.cols)+    
    facet_wrap(~variable, labeller = groups, ncol=2)+
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
    labs(y="(%) Proliferating CD4 T cells")

melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))
two<-allplot(melt, "TB", "value", all.cols)+    
    facet_wrap(~variable, labeller = groups)+
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
    coord_cartesian(ylim=c(0,5))

#phenotypes
melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))
tf<-allplot(melt, "TB", "value", all.cols)+    
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
    labs(y="(%) Proliferating CD4 T cells")

melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))
sm<-allplot(melt, "TB", "value", all.cols)+
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))

first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20, rel_widths = c(3,1))
second_row = cowplot::plot_grid(one, two, nrow=1, rel_widths = c(2,1), labels=c("B","C"), label_size = 20)
third_row= cowplot::plot_grid(tf, sm, labels=c('F', 'G'), nrow=1, label_size = 20)

supp6<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3)
```

```{r}
ogphen<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._.", names(ogphen)))
test<-melt[-grep("._._._.", melt$variable),]
test<-filter(test, variable!="x_x_x")
test$variable<-factor(test$variable, levels=c(
"T_G_x" ,       
"T_x_x" ,         
"x_G_x" ,         
"T_G_4" ,
"T_x_4" ,         
"x_G_4" ,         
"x_x_4"          
))
test$group<-"TH1/2"
for(i in 1:length(test$group)){
    if(test$variable[i] %in% c("T_G_x", "T_x_x", "x_G_x")){
        test$group[i]<- "TH1"
    }
    else if (test$variable[i] == "x_x_4"){
        test$group[i]<- "TH2"
    }
}

T <- c("+", "+", "-", "+", "+", "-", "-")
G <- c("+", "-", "+", "+", "-", "+", "-")
il4 <- c("-", "-", "-", "+", "+", "+", "+")

labels <- paste0(T, "\n", G, "\n", il4)
breaks <- levels(test$variable)

supp7a<-booleanplot(test, "variable", "value", ltbi.cols)+
    scale_x_discrete(labels = labels, breaks=breaks)+
    facet_grid(~group, scales="free_x", space="free_x")+
    labs(y="(%) Proliferating CD4 T cells")

```

```{r og phenotype, fig.width=16, fig.height=24}
#graph phenotype
melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._._.", names(ogphen)))
test<-filter(melt, variable!="x_x_x_x")
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" ,       
"t_x_3_x" ,         
"t_x_x_4" ,         
"t_x_x_x" ,
"t_g_3_4" ,         
"t_g_3_x" ,         
"t_g_x_4" ,         
"t_g_x_x" ,
"x_g_3_4" ,         
"x_g_3_x" ,         
"x_g_x_4" ,         
"x_g_x_x" ,
"x_x_3_4" ,          
"x_x_3_x" ,         
"x_x_x_4"           
))

T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
breaks <- levels(test$variable)

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

supp7b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=6)+
    labs(y="(%) Proliferating Mtb-Specific cells")
```

```{r, fig.width=12, fig.height=12}
supp7<-cowplot::plot_grid(supp7a, supp7b, nrow=2, labels=c("A", "B"), rel_widths = c(3,4))
supp7
```

###Comparing SEB-proliferating cells to non-proliferating cells in UN 
```{r, fig.width=12, fig.height=6}
un<-filter_ltbi(dplyr::filter(OG, Stim=="UN"))
un$cell<-"Total"
seb<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="SEB"))
seb$cell<-"SEB"
long<-rbindlist(list(un, seb), fill=TRUE)
long$cell<-factor(long$cell, levels=c("Total", "SEB"))
long<-data.frame(long)

melt<-melt(long, id.vars=c("TB", "SM", "disease", "Donor", "cell"), 
           measure.vars=c("T.bet", "GATA3", "CXCR3","CCR4"))
melt$x<-paste(melt$SM, melt$cell)
melt$x<-factor(melt$x, levels=c("SM- Total", "SM- SEB","SM+ Total", "SM+ SEB"))

ggplot(melt, aes(x=x, y=value, fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    facet_wrap(~variable, scales="free", nrow=1)+
    theme_classic()+
    scale_fill_manual(values = ltbi.cols)+
    scale_x_discrete(labels=c("Total", "SEB", "Total", "SEB"))+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons=list(c("SM- Total", "SM- SEB"),c("SM+ Total", "SM+ SEB")) 
                       ,method="wilcox.test", label="p.format")+
    labs(x="", y="")
```

###Comparing CFP-10 and ESAT-6-proliferating cells to non-proliferating cells in UN
```{r, fig.width=12, fig.height=6}
un<-filter_ltbi(dplyr::filter(OG, Stim=="UN"))
un$cell<-"Total"
ce<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
ce$cell<-"PEP"
long<-rbindlist(list(un, ce), fill=TRUE)
long$cell<-factor(long$cell, levels=c("Total", "PEP"))
long<-data.frame(long)

melt<-melt(long, id.vars=c("TB", "SM", "disease", "Donor", "cell"), 
           measure.vars=c("T.bet", "GATA3", "CXCR3","CCR4"))
melt$x<-paste(melt$SM, melt$cell)
melt$x<-factor(melt$x, levels=c("SM- Total", "SM- PEP", 
                                "SM+ Total", "SM+ PEP"))

ggplot(melt, aes(x=x, y=value, fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    facet_wrap(~variable, scales="free", nrow=1)+
    theme_classic()+
    scale_fill_manual(values = ltbi.cols)+
    scale_x_discrete(labels=c("Total", "CFP-10 &\nESAT-6", "Total", "CFP-10 &\nESAT-6"))+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons=list(c("SM- Total", "SM- PEP"),c("SM+ Total", "SM+ PEP")), 
        method="wilcox.test", label="p.format")+
    labs(x="", y="")

```

###Comparing CFP-10 and ESAT-6-proliferating to non-proliferating cells
```{r, fig.width=12, fig.height=6}
un<-filter_ltbi(dplyr::filter(OGnope, Stim=="PEP"))
un$cell<-"Non-Proliferating"
ce<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
ce$cell<-"Proliferating"
long<-rbindlist(list(un, ce), fill=TRUE)
long$cell<-factor(long$cell, levels=c("Non-Proliferating", "Proliferating"))
long<-data.frame(long)

melt<-melt(long, id.vars=c("TB", "SM", "disease", "Donor", "cell"), 
           measure.vars=c("T.bet", "GATA3", "CXCR3","CCR4"))
melt$x<-paste(melt$SM, melt$cell)
melt$x<-factor(melt$x, levels=c("SM- Non-Proliferating", "SM- Proliferating", 
                                "SM+ Non-Proliferating", "SM+ Proliferating"))

ggplot(melt, aes(x=x, y=value, fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    facet_wrap(~variable, scales="free", nrow=1)+
    theme_classic()+
    scale_fill_manual(values = ltbi.cols)+
    scale_x_discrete(labels=c("Non-\nProliferating", "Proliferating", 
                              "Non-\nProliferating", "Proliferating"))+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons=list(c("SM- Non-Proliferating", "SM- Proliferating"),
                                        c("SM+ Non-Proliferating", "SM+ Proliferating")), 
        method="wilcox.test", label="p.format")+
    labs(x="", y="")

```

```{r}
lum_x<-lum[lum$SM=="SM-",]
df_x<-as.matrix(lum_x[,3:11])
row.names(df_x)<-lum_x$Sample
hc_x <- hclust(as.dist(1-cor(df_x)))
x<-plot(hc_x)

lum_sm<-lum[lum$SM=="SM+",]
df_sm<-as.matrix(lum_sm[,3:11])
row.names(df_sm)<-lum_sm$Sample
hc_sm <- hclust(as.dist(1-cor(df_sm)))
sm<-plot(hc_sm)

df<-as.matrix(lum[,3:11])
#heatmap(t(df_x), Rowv=as.dendrogram(hc_x))
#heatmap(t(df_sm), Rowv=as.dendrogram(hc_sm))
#heatmap(t(df), Rowv=as.dendrogram(hc_sm))
```

#Exporting Data
```{r}
setwd("/Users/tarynam/Desktop/Manuscript1")
cowplot::save_plot("supp1.png", supp1, base_height=14, base_width=16)
cowplot::save_plot("supp2.png", supp2, base_height=28, base_width=16)
cowplot::save_plot("supp3.png", supp3, base_height=7, base_width=16)
cowplot::save_plot("supp4.png", supp4, base_height=14, base_width=16)
cowplot::save_plot("supp5.png", supp5, base_height=14, base_width=16)
cowplot::save_plot("supp6.png", supp6, base_height=14, base_width=16)
cowplot::save_plot("supp7.png", supp6, base_height=14, base_width=16)
```

#Extra plots

To see if the variation in T-bet and GATA3 in the SEB proliferation was due to original variation or proliferation. It's not.
```{r}
test<-merge(OG, OGPHEN, by=c("Sample", "Donor", "Stim", "TB", "SM"))
test<-filter(test, Stim=="SEB")
ggplot(test, aes(x=OG.x, y=Tbet.y, color=TB))+geom_point()+stat_smooth(method="lm")
ggplot(test, aes(x=OG.x, y=GATA3.y, color=TB))+geom_point()+stat_smooth(method="lm")
ggplot(test, aes(x=Tbet.y, y=GATA3.y, color=TB))+geom_point()+stat_smooth(method="lm")

part1<-filter(CD4, Stim=="UN")
part2<-filter(OGPHEN, Stim=="SEB")
test<-merge(part1, part2, by=c("Donor", "TB", "SM"))
ggplot(test, aes(x=Tbet.x, y=Tbet.y))+geom_point()+stat_smooth(method="lm")
ggplot(test, aes(x=GATA3.x, y=GATA3.y))+geom_point()+stat_smooth(method="lm")
```

### 2019-10-03 PaperFigures.Rmd
change figure sizes/labeling

@@
 source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")
-
-baseplot<-function(data, xval, yval, colors){
-    ggplot(data, aes(data[,xval], data[,yval], fill=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    scale_fill_manual(values = colors)+
-    theme_classic()+ theme(legend.position="none")+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=20),
-          axis.text.x = element_text(size = 16),
-          strip.text = element_text(size = 20))+
-    theme(plot.title = element_text(hjust = 0.5, size=20)) +
-    labs(x="", y="")
-}
-
-booleanplot<-function(data, xval, yval, colors){
-    ggplot(data, aes(data[,xval], data[,yval], fill=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    scale_fill_manual(values = colors)+
-    theme_classic()+ 
-    theme(legend.position="none")+
-    theme(text = element_text(size=20), 
-          axis.text.x = element_text(size=16), 
-          strip.text = element_text(size = 20))+
-    theme(plot.title = element_text(hjust = 0.5, size=20)) +
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
-    labs(x="", y="")
-}
-
-allplot<-function(data, xval, yval, colors){
-    ggplot(data, aes(data[,xval], data[,yval], fill=disease))+
-    geom_boxplot(size=1, position=position_dodge2(width=0.2, padding = 0.2, preserve = "single"), outlier.shape = NA)+
-    scale_fill_manual(values = colors)+
-    theme_classic()+ 
-    theme(legend.position="none") +
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    facet_wrap(~variable)+
-    theme(text = element_text(size=20),
-            axis.text.x = element_text(size = 16),
-            strip.text = element_text(size = 20))+
-    theme(plot.title = element_text(hjust = 0.5, size=20))+
-    labs(x="", y="")
-}
-
-cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
-                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))
-
-all_comps<-list(c("HC SM+", "HC SM-"),c("LTBI SM+", "LTBI SM-"), c("TB SM+", "TB SM-"))
-
-healthy.cols<- c("SM+" = "#1a9850", "SM-" ="#91cf60")
-ltbi.cols<- c("SM+" = "#2166ac", "SM-" = "#85ABD1")
-tb.cols<- c("SM+" = "#b2182b", "SM-"="#E59EA6")
-all.cols<- c("Naive" = "#C8D0D8","HC SM+" = "#1a9850", "HC SM-" ="#91cf60",
-             "LTBI SM+" = "#2166ac", "LTBI SM-" = "#85ABD1",
-             "TB SM+" = "#b2182b", "TB SM-"="#E59EA6")
 ```
@@
 
-pval<-c(A$`SM-.HC`$p.value, A$`SM+.HC`$p.value,
-        A$`SM-.LTBI`$p.value, A$`SM+.LTBI`$p.value, 
-        A$`SM-.TB`$p.value, A$`SM+.TB`$p.value)
-group<-c("HC SM-", "HC SM+", "LTBI SM-", "LTBI SM+", "TB SM-", "TB SM+")
-table<-data.frame(cbind(group, pval))
-table$pval<-as.numeric(as.character(table$pval))
-table$pval[table$pval<.0001]<-"<0.0001"
-kable(table, caption="TH1 to TH2")
-```
-
-some stats comparing th1 to th1/2 for reference. 
-```{r}
 x<-filter(plot_filter(melt), variable!="TH2.Freq")
 y<-split(x,list(x$SM, x$TB)) 
-A<-lapply(y, function(g) wilcox.test(value~variable, g))
+B<-lapply(y, function(g) wilcox.test(value~variable, g))
 
-pval<-c(A$`SM-.HC`$p.value, A$`SM+.HC`$p.value,
-        A$`SM-.LTBI`$p.value, A$`SM+.LTBI`$p.value, 
-        A$`SM-.TB`$p.value, A$`SM+.TB`$p.value)
-group<-c("HC SM-", "HC SM+", "LTBI SM-", "LTBI SM+", "TB SM-", "TB SM+")
-table<-data.frame(cbind(group, pval))
-table$pval<-as.numeric(as.character(table$pval))
-table$pval[table$pval<.0001]<-"<0.0001"
-kable(table, caption="TH1 to TH1/2")
+hc<-c(A$`SM-.HC`$p.value, A$`SM+.HC`$p.value,
+      B$`SM-.HC`$p.value, B$`SM+.HC`$p.value)
+
+ltbi<-c(A$`SM-.LTBI`$p.value, A$`SM+.LTBI`$p.value, 
+        B$`SM-.LTBI`$p.value, B$`SM+.LTBI`$p.value)
+
+tb<-c(A$`SM-.TB`$p.value, A$`SM+.TB`$p.value,
+      B$`SM-.TB`$p.value, B$`SM+.TB`$p.value)
+
+label<-c("SM- TH1 to TH1/2", "SM- TH1 to TH2", "SM+ TH1 to TH1/2", "SM+ TH1 to TH2")
+
+hc.tab<-data.frame(cbind(label, hc, p.adjust(hc, method="bonferroni")))
+names(hc.tab)<-c("Comparison", "P-value", "Bonferroni Adjust P-val")
+kable(hc.tab, caption="HC")
+
+ltbi.tab<-data.frame(cbind(label, ltbi, p.adjust(ltbi, method="bonferroni")))
+names(hc.tab)<-c("Comparison", "P-value", "Bonferroni Adjust P-val")
+kable(ltbi.tab, caption="LTBI")
+
+tb.tab<-data.frame(cbind(label, tb, p.adjust(tb, method="bonferroni")))
+names(hc.tab)<-c("Comparison", "P-value", "Bonferroni Adjust P-val")
+kable(tb.tab, caption="TB")
 ```
@@
 melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4", "IL13"))
-b<-baseplot(melt, "disease", "value", tb.cols)+ 
+b<-baseplot(melt, "SM", "value", tb.cols)+ 
     facet_wrap(~variable, labeller = cyt_groups)+
@@
 melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
-one<-baseplot(melt, "disease", "value", ltbi.cols)+    
+one<-baseplot(melt, "SM", "value", ltbi.cols)+    
     facet_wrap(~variable, labeller = groups, ncol=2)+
@@
 melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))
-two<-baseplot(melt, "disease", "value", ltbi.cols)+    
+two<-baseplot(melt, "SM", "value", ltbi.cols)+    
     facet_wrap(~variable, labeller = groups)
@@
     coord_cartesian(ylim=c(0,1500))+
-    theme(plot.margin=margin(l=-0.5,unit="cm"))
+    labs(y="pg/mL")
 b<-baseplot(melt[melt$variable=="IFNγ",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
@@
 ```{r}
-setwd("/Users/tarynam/Desktop/Manuscript1")
+setwd("/Users/tarynam/Desktop/Manuscript1/figures")
+cowplot::save_plot("figure8.png", lumplot, base_height=6, base_width=18)
 

### 2019-10-03 SuppFigures.Rmd
change figure sizes/labeling

@@
 library(sjPlot)
+library(ggsignif)
 ```
@@
 PHEN<-read.csv("ICS_CD4_phenotyping_clean_NAs.csv")
-OG<-data.table(read.csv("OG_CD4_bulk_summary_clean.csv"))
+OG<-read.csv("OG_CD4_bulk_summary_clean.csv")
 OGPHEN<-read.csv("OG_CD4_OG_summary_clean_NAs.csv")
@@
 
-##ICS Supplementary
-```{r}
-booleanplot2<-function(data, xval, yval, colors){
-    ggplot(data, aes(data[,xval], data[,yval], fill=disease))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    scale_fill_manual(values = colors)+
-    theme_classic()+ 
-    theme(legend.position="none")+
-    theme(text = element_text(size=20), axis.text.x = element_text(size=20))+
-    theme(plot.subtitle = element_text(hjust = 0.5)) +
-    labs(x="", y="")
-}
+```{r, all the labels I use multiple times}
+#cytokines
+G   <-c("+", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
+il4 <-c("+", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
+il13<-c("+", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
+T   <-c("+", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")
+
+cyt_labels <- paste0(G, "\n", T, "\n", il4, "\n", il13)
+cyt_breaks <- c(
+"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
+"G_x_x_T" , "G_x_x_x","x_x_x_T",
+"x_4_13_x","x_4_x_x", "x_x_13_x")
+
+#TH groups
+th_groups <- as_labeller(c("TH1.Freq" = "TH1", 
+                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))
+
+#Phenotype boolean labels
+T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
+G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
+c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
+c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
+
+phen_labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
+phen_breaks <- c(
+"t_x_3_4" , "t_x_3_x" , "t_x_x_4" , "t_x_x_x" ,
+"t_g_3_4" , "t_g_3_x" , "t_g_x_4" , "t_g_x_x" ,
+"x_g_3_4" , "x_g_3_x" , "x_g_x_4" , "x_g_x_x" ,
+"x_x_3_4" , "x_x_3_x" , "x_x_x_4")
 ```
 
-###Boolean PMA
-*but I need to make them pretty and I dont' feel like doing that right now*
-```{r boolean pma cyt, fig.width=15, fig.height=12}
+#ICS Supplementary
+
+##2 Boolean PMA
+```{r}
 cyt<-plot_filter(CYT[CYT$Stim=="PMA",])
@@
 
-G   <-c("+", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
-il4 <-c("+", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
-il13<-c("+", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
-T   <-c("+", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")
-
-labels <- paste0(G, "\n", T, "\n", il4, "\n", il13)
-breaks <- levels(test$variable)
-
 test$th<-"TH1/2"
@@
 th1.2<-booleanplot2(test[test$th=="TH1/2",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
+    scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
     stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
@@
 th1<-booleanplot2(test[test$th=="TH1",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
+    scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
     stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
@@
 th2<-booleanplot2(test[test$th=="TH2",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
+    scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
     stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
     coord_cartesian(ylim=c(0,3))
-
-supp1<-cowplot::plot_grid(th1.2, th1, th2, ncol=3, rel_widths = c(8,3.5,3.5))
 ```
 
-###Boolean Phenotypes
-###TH1
-```{r pma th1 boolean, fig.width=20, fig.height=18}
+```{r, fig.height=10, fig.width=10}
+supp2<-cowplot::plot_grid(th1.2, th1, th2, ncol=3, rel_widths = c(8,3.5,3.5))
+supp2
+```
+
+##3 PMA Phenotypes
+```{r}
 phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))
 
-melt<-melt(phen, id.vars=c("TB", "SM", "disease"), measure.vars = grep("TH1_._._._.", names(phen)))
-test<-filter(melt, variable!="TH1_x_x_x_x")
+test<-melt(phen, id.vars=c("TB", "SM", "disease"), measure.vars = grep("TH._._._._.", names(phen)))
+test$thgroup<-"TH1"
+test$thgroup[grep("TH2", test$variable)]<-"TH2"
+
+test<-test[-(grep("x_x_x_x", test$variable)),]
 test$variable<-gsub("TH1_", "", test$variable)
+test$variable<-gsub("TH2_", "", test$variable)
 test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
-
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
-
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
-
-df<-data.frame(cbind(levels(test$variable),T,G,c3,c4))
-dfm<-melt(df, measure.vars=c("T", "G","c3","c4"))
-dfm$variable<-factor(dfm$variable, levels=c("c4","c3","G", "T"))
-
-data_table <- ggplot(dfm, aes(x = V1, y = variable, label = format(value))) +
-    geom_text(size = 10) + 
-    theme_bw() + 
-    labs(x="",y="")+
-    theme(strip.text.x = element_blank(), strip.background.x = element_blank())+
-    scale_y_discrete(labels=c("T-bet", "GATA3", "CXCR3", "CCR4"))+
-    theme(axis.text.y = element_text(size=20))+
-    theme(panel.grid.major = element_blank(), 
-          panel.grid.minor = element_blank(),
-          panel.border = element_blank(),
-          axis.title = element_blank(),
-          axis.text.x = element_blank(),
-          axis.ticks = element_blank())
+"t_x_3_4" , "t_x_3_x" , "t_x_x_4" , "t_x_x_x" ,
+"t_g_3_4" , "t_g_3_x" , "t_g_x_4" , "t_g_x_x" ,
+"x_g_3_4" , "x_g_3_x" , "x_g_x_4" , "x_g_x_x" ,
+"x_x_3_4" , "x_x_3_x" , "x_x_x_4"))
 
@@
 
-supp2a<-booleanplot2(test, "variable", "value", all.cols)+
+supp3a<-booleanplot2(test[test$thgroup=="TH1",], "variable", "value", all.cols)+
     facet_grid(TB~tf, scales="free_x")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
+    scale_x_discrete(labels = phen_labels, breaks=phen_breaks)+
     stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
-    coord_cartesian(ylim=c(0,65))
+    coord_cartesian(ylim=c(0,65))+
+    theme(plot.margin=margin(l=1,unit="cm"))+
+    labs(y="(%) TH1 Cytokine+ CD4 T cells")
+
+supp3b<-booleanplot2(test[test$thgroup=="TH2",], "variable", "value", all.cols)+
+    facet_grid(TB~tf, scales="free_x")+
+    scale_x_discrete(labels = phen_labels, breaks=phen_breaks)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
+    coord_cartesian(ylim=c(0,65))+
+    theme(plot.margin=margin(l=1,unit="cm"))+
+    labs(y="(%) TH2 Cytokine+ CD4 T cells")
 
 ```
-###TH2  
-```{r pma th2 boolean, fig.width=20, fig.height=18}
-phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))
 
-melt<-melt(phen, id.vars=c("TB", "SM", "disease"), measure.vars = grep("TH2_._._._.", names(phen)))
-test<-filter(melt, variable!="TH2_x_x_x_x")
-test$variable<-gsub("TH2_", "", test$variable)
+```{r, fig.height=12, fig.width=10}
+supp3<-cowplot::plot_grid(supp3a, supp3b, nrow=2, labels=c("A", "B"), label_size = 20)
+supp3
+```
+
+##4 LTBI ICS extra 
+```{r}
+cyt<-filter_ltbi(CYT[CYT$Stim=="PEP",])
+
+#boolean cytokine
+test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
+test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
 test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
+"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
+"G_x_x_T" , "G_x_x_x","x_x_x_T",
+"x_4_13_x","x_4_x_x", "x_x_13_x"
 ))
 
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
+test$th<-"TH1/2"
+test$th[grep("._x_x_.", test$variable)]<-"TH1"
+test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
+test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))
 
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
+a<-booleanplot2(test, "variable", "value", all.cols)+facet_grid(~th, scales="free", space="free")+
+    scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
+    coord_cartesian(ylim=c(0,.16))+
+    theme(plot.margin=margin(l=1,unit="cm"))+
+    labs(y="(%) CD4 T cells")
+
+#condensed TH cytokine
+test<-melt(cyt,
+           id.vars=c("TB","SM"), 
+           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
+
+b<-baseplot(test, "SM", "value", ltbi.cols)+
+    facet_wrap(~variable, ncol=4, labeller=th_groups)+
+    coord_cartesian(ylim=c(0,.2))+
+    labs(y="(%) CD4 T cells")
+
+#phenotypes
+phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
+
+melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
+test<-filter(melt, variable!="TH1_x_x_x_x")
+test$variable<-gsub("TH1_", "", test$variable)
+test$variable<-factor(test$variable, levels=c(
+"t_x_3_4" , "t_x_3_x" , "t_x_x_4" , "t_x_x_x" ,
+"t_g_3_4" , "t_g_3_x" , "t_g_x_4" , "t_g_x_x" ,
+"x_g_3_4" , "x_g_3_x" , "x_g_x_4" , "x_g_x_x" ,
+"x_x_3_4" , "x_x_3_x" , "x_x_x_4"))
 
@@
 
-supp2b<-booleanplot2(test, "variable", "value", all.cols)+
-    facet_grid(TB~tf, scales="free_x")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
-    coord_cartesian(ylim=c(0,65))
+c<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x",ncol=4)+
+    scale_x_discrete(labels = phen_labels, breaks=phen_breaks)+
+    theme(legend.position = "right", legend.key.size = unit(4, "cm"))+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
+    theme(plot.margin=margin(l=0.5,unit="cm"))+
+    labs(y="(%) Mtb-Specific\nTH1 Cytokine+ cells")
 ```
 
-```{r}
-supp2<-cowplot::plot_grid(supp2a, supp2b, nrow=2, labels=c("A", "B"))
+```{r, fig.height=10, fig.width=12}
+#compiling together
+suppab<-cowplot::plot_grid(a, b, nrow=1, labels=c("A", "B"), label_size = 20,
+                           align="hv", rel_widths = c(6,3))
+suppc<-cowplot::plot_grid(c, nrow=1, labels=c("C"), label_size = 20)
+supp4<-cowplot::plot_grid(suppab, suppc, align="hv", nrow=2)
+supp4
 ```
 
-###Peptide-Specific Cytokine Production  
-```{r, fig.width=15, fig.height=8}
-melt<-melt(subset(CYT, CYT$Stim=="PEP"), 
-           id.vars=c("TB","SM"), 
-           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
-th_groups <- as_labeller(c("TH1.Freq" = "TH1", 
-                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))
-
-cyt<-filter_ltbi(melt)
-b<-baseplot(cyt, "SM", "value", ltbi.cols)+
-    facet_wrap(~variable, ncol=4, labeller=th_groups)+
-    coord_cartesian(ylim=c(0,.2))
-
-cyt<-filter_tb(melt)
-c<-baseplot(cyt, "SM", "value", tb.cols)+
-    facet_wrap(~variable, ncol=4, labeller=th_groups)+
-    coord_cartesian(ylim=c(0,.2))
-
-supp3<-cowplot::plot_grid(b, c, ncol=3)
-```
-
-```{r, fig.width=16, fig.height=8}
-cyt<-plot_filter(CYT[CYT$Stim=="PEP",])
-test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
-test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
-test$variable<-factor(test$variable, levels=c(
-"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
-"G_x_x_T" , "G_x_x_x","x_x_x_T",
-"x_4_13_x","x_4_x_x", "x_x_13_x"
-))
-
-G   <-c("IFNγ +", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
-il4 <-c("IL-4 +", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
-il13<-c("IL-13 +", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
-T   <-c("TNFα +", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")
-
-labels <- paste0(G, "\n", T, "\n", il4, "\n", il13)
-breaks <- levels(test$variable)
-
-test$th<-"TH1/2"
-test$th[grep("._x_x_.", test$variable)]<-"TH1"
-test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
-test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))
-
-supp4a<-booleanplot2(test[test$TB=="LTBI",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
-    coord_cartesian(ylim=c(0,.16))+
-    labs(y="(%) CD4 T cells")
-
-supp5a<-booleanplot2(test[test$TB=="TB",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
-    coord_cartesian(ylim=c(0,.15))+
-    labs(y="(%) CD4 T cells")
-```
-
-###Peptide Mtb-specific TH1 Cytokine+ Cell Phenotypes
-```{r ics phenotyping, fig.width=16, fig.height=24}
-phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
-
-melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
-test<-filter(melt, variable!="TH1_x_x_x_x")
-test$variable<-gsub("TH1_", "", test$variable)
-test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
-
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
-
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
-
-test$tf<-NA
-test$tf[grep("x_g", test$variable)]<-"GATA3"
-test$tf[grep("t_x", test$variable)]<-"T-bet"
-test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
-test$tf[grep("x_x_._.", test$variable)]<-"None"
-test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
-
-supp4b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x",ncol=4)+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
-```
-
-```{r, fig.width=15, fig.height=24}
-phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
-
-#graph phenotype
-melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
-test<-filter(melt, variable!="TH1_x_x_x_x")
-test$variable<-gsub("TH1_", "", test$variable)
-test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
-
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
-
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
-
-test$tf<-NA
-test$tf[grep("x_g", test$variable)]<-"GATA3"
-test$tf[grep("t_x", test$variable)]<-"T-bet"
-test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
-test$tf[grep("x_x_._.", test$variable)]<-"None"
-test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
-
-supp5b<-booleanplot(test, "variable", "value", tb.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")+
-    coord_cartesian(ylim=c(0,55))+
-    stat_compare_means(method="wilcox.test",label="p.signif", hide.ns=TRUE, label.y=50, size=10)
-```
-
-###Comparing Mtb-specific TH1 cells to total CD4 T cells in peptide stim  
-```{r, fig.width=12, fig.height=6}
+##5 Comparing Mtb-specific CD4 T cells to other CD4 T cells in LTBI
+```{r}
 total<-filter_ltbi(subset(CD4, CD4$Stim=="UN"))
@@
 melt$x<-factor(melt$x, levels=c("SM- Total", "SM- Mtb-Specific","SM+ Total", "SM+ Mtb-Specific"))
+melt<-data.frame(melt)
 
-ggplot(melt, aes(x=x, y=value, fill=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    facet_wrap(~variable, scales="free", nrow=1)+
-    theme_classic()+
-    scale_fill_manual(values = ltbi.cols)+
-    scale_x_discrete(labels=c("Total", "Mtb", "Total", "Mtb"))+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
-    stat_compare_means(comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")) 
-                       ,method="wilcox.test", label="p.format")+
-    labs(x="", y="")
+a1<-baseplot(melt[(melt$variable %in% c("T-bet", "GATA3")),], "x", "value", ltbi.cols)+
+    facet_wrap(~variable, nrow=1)+
+    scale_x_discrete(labels=c("Total", "CFP-10 &\nESAT-6", "Total", "CFP-10 &\nESAT-6"))+
+    labs(y="(%) CD4 T cells")+
+    geom_signif(
+        comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")),
+                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = .98*32)+    
+    coord_cartesian(ylim=c(0,32))
+
+a2<-baseplot(melt[(melt$variable %in% c("CXCR3", "CCR4")),], "x", "value", ltbi.cols)+
+    facet_wrap(~variable, nrow=1)+
+    scale_x_discrete(labels=c("Total", "CFP-10 &\nESAT-6", "Total", "CFP-10 &\nESAT-6"))+
+    geom_signif(
+        comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")),
+                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = 98)+    
+    coord_cartesian(ylim=c(0,100))
 ```
 
-###Comparing Mtb-specific TH1 cells to PMA TH1 cells  
-```{r, fig.width=12, fig.height=6}
+```{r}
 pma<-filter_ltbi(subset(PHEN, PHEN$Stim=="PMA"))
@@
 melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))
+melt<-data.frame(melt)
 
-ggplot(melt, aes(x=x, y=value, fill=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    facet_wrap(~variable, scales="free", nrow=1)+
-    theme_classic()+
-    scale_fill_manual(values = ltbi.cols)+
-    scale_x_discrete(labels=c("PMA", "Mtb", "PMA", "Mtb"))+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
-    stat_compare_means(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")) 
-                       ,method="wilcox.test", label="p.format")+
-    labs(x="", y="")
+b1<-baseplot(melt[(melt$variable %in% c("T-bet", "GATA3")),], "x", "value", ltbi.cols)+
+    facet_wrap(~variable, nrow=1)+
+    scale_x_discrete(labels=c("PMA", "CFP-10 &\nESAT-6", "PMA", "CFP-10 &\nESAT-6"))+
+    labs(y="(%) CD4 T cells")+
+    geom_signif(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")), 
+                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = .98*40)+    
+    coord_cartesian(ylim=c(0,40))
+
+b2<-baseplot(melt[(melt$variable %in% c("CXCR3", "CCR4")),], "x", "value", ltbi.cols)+
+    facet_wrap(~variable, nrow=1)+
+    scale_x_discrete(labels=c("PMA", "CFP-10 &\nESAT-6", "PMA", "CFP-10 &\nESAT-6"))+
+    geom_signif(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")), 
+                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = 98)+    
+    coord_cartesian(ylim=c(0,100))
 ```
 
-###Comparing Mtb-specific TH1 cells to total CD4 T cells in peptide stim  
-```{r, fig.width=12, fig.height=6}
+```{r, fig.height=10, fig.width=12}
+supp5<-cowplot::plot_grid(a1,a2,b1,b2, ncol=2, labels=c("A","","B",""), label_size = 20)
+supp5
+```
+
+##6 TB ICS extra
+```{r}
+cyt<-filter_tb(CYT[CYT$Stim=="PEP",])
+
+#boolean cytokine
+test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
+test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
+test$variable<-factor(test$variable, levels=c(
+"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
+"G_x_x_T" , "G_x_x_x","x_x_x_T",
+"x_4_13_x","x_4_x_x", "x_x_13_x"
+))
+
+test$th<-"TH1/2"
+test$th[grep("._x_x_.", test$variable)]<-"TH1"
+test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
+test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))
+
+a<-booleanplot2(test, "variable", "value", all.cols)+facet_grid(~th, scales="free", space="free")+
+    scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
+    coord_cartesian(ylim=c(0,.16))+
+    theme(plot.margin=margin(l=1,unit="cm"))+
+    labs(y="(%) CD4 T cells")
+
+#condensed TH cytokine
+test<-melt(cyt,
+           id.vars=c("TB","SM"), 
+           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
+
+b<-baseplot(test, "SM", "value", tb.cols)+
+    facet_wrap(~variable, ncol=4, labeller=th_groups)+
+    coord_cartesian(ylim=c(0,.2))+
+    labs(y="(%) CD4 T cells")
+
+#phenotypes
+phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
+
+melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
+test<-filter(melt, variable!="TH1_x_x_x_x")
+test$variable<-gsub("TH1_", "", test$variable)
+test$variable<-factor(test$variable, levels=c(
+"t_x_3_4" , "t_x_3_x" , "t_x_x_4" , "t_x_x_x" ,
+"t_g_3_4" , "t_g_3_x" , "t_g_x_4" , "t_g_x_x" ,
+"x_g_3_4" , "x_g_3_x" , "x_g_x_4" , "x_g_x_x" ,
+"x_x_3_4" , "x_x_3_x" , "x_x_x_4"))
+
+test$tf<-NA
+test$tf[grep("x_g", test$variable)]<-"GATA3"
+test$tf[grep("t_x", test$variable)]<-"T-bet"
+test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
+test$tf[grep("x_x_._.", test$variable)]<-"None"
+test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))
+
+c<-booleanplot(test, "variable", "value", tb.cols)+facet_wrap(~tf, scales="free_x",ncol=4)+
+    scale_x_discrete(labels = phen_labels, breaks=phen_breaks)+
+    theme(legend.position = "right", legend.key.size = unit(4, "cm"))+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
+    coord_cartesian(ylim=c(0,50))+
+    labs(y="(%) Mtb-Specific\nTH1 Cytokine+ cells")
+```
+
+```{r, fig.height=10, fig.width=12}
+#compiling together
+suppab<-cowplot::plot_grid(a, b, nrow=1, labels=c("A", "B"), label_size = 20, 
+                           align="hv", rel_widths = c(6,3))
+suppc<-cowplot::plot_grid(c, nrow=1, labels=c("C"), label_size = 20)
+supp6<-cowplot::plot_grid(suppab, suppc, align="v", nrow=2)
+supp6
+```
+
+##7 Comparing Mtb-specific CD4 T cells to others in TB
+```{r}
 total<-filter_tb(subset(CD4, CD4$Stim=="UN"))
@@
 melt$x<-factor(melt$x, levels=c("SM- Total", "SM- Mtb-Specific","SM+ Total", "SM+ Mtb-Specific"))
+melt<-data.frame(melt)
 
-ggplot(melt, aes(x=x, y=value, fill=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    facet_wrap(~variable, scales="free", nrow=1)+
-    theme_classic()+
-    scale_fill_manual(values = tb.cols)+
-    scale_x_discrete(labels=c("Total", "Mtb", "Total", "Mtb"))+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
-    stat_compare_means(comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")) 
-                       ,method="wilcox.test", label="p.format")+
-    labs(x="", y="")
+a1<-baseplot(melt[(melt$variable %in% c("T-bet", "GATA3")),], "x", "value", tb.cols)+
+    facet_wrap(~variable, nrow=1)+
+    scale_x_discrete(labels=c("Total", "CFP-10 &\nESAT-6", "Total", "CFP-10 &\nESAT-6"))+
+    labs(y="(%) CD4 T cells")+
+    geom_signif(
+        comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")),
+                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = .98*80)+    
+    coord_cartesian(ylim=c(0,80))
+
+a2<-baseplot(melt[(melt$variable %in% c("CXCR3", "CCR4")),], "x", "value", tb.cols)+
+    facet_wrap(~variable, nrow=1)+
+    scale_x_discrete(labels=c("Total", "CFP-10 &\nESAT-6", "Total", "CFP-10 &\nESAT-6"))+
+    geom_signif(
+        comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")),
+                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = .98*80)+    
+    coord_cartesian(ylim=c(0,80))
 
@@
 
-###Comparing Mtb-specific TH1 cells to PMA TH1 cells  
-```{r, fig.width=12, fig.height=6}
+```{r}
 pma<-filter_tb(subset(PHEN, PHEN$Stim=="PMA"))
@@
 melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))
+melt<-data.frame(melt)
 
-ggplot(melt, aes(x=x, y=value, fill=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    facet_wrap(~variable, scales="free", nrow=1)+
-    theme_classic()+
-    scale_fill_manual(values = tb.cols)+
-    scale_x_discrete(labels=c("PMA", "Mtb", "PMA", "Mtb"))+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
-    stat_compare_means(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")) 
-                       ,method="wilcox.test", label="p.format")+
-    labs(x="", y="")
+b1<-baseplot(melt[(melt$variable %in% c("T-bet", "GATA3")),], "x", "value", tb.cols)+
+    facet_wrap(~variable, nrow=1)+
+    scale_x_discrete(labels=c("PMA", "CFP-10 &\nESAT-6", "PMA", "CFP-10 &\nESAT-6"))+
+    labs(y="(%) CD4 T cells")+
+    geom_signif(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")), 
+                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = .98*80)+    
+    coord_cartesian(ylim=c(0,80))
+
+b2<-baseplot(melt[(melt$variable %in% c("CXCR3", "CCR4")),], "x", "value", tb.cols)+
+    facet_wrap(~variable, nrow=1)+
+    scale_x_discrete(labels=c("PMA", "CFP-10 &\nESAT-6", "PMA", "CFP-10 &\nESAT-6"))+
+    geom_signif(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")), 
+                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = .98*80)+    
+    coord_cartesian(ylim=c(0,80))
 ```
 
-```{r, fig.width=12, fig.height=6}
-supp4<-cowplot::plot_grid(supp4a, supp4b, nrow=2, labels=c("A", "B"))
-supp5<-cowplot::plot_grid(supp5a, supp5b, nrow=2, labels=c("A", "B"))
-supp1
-supp2
-supp3
-supp4
-supp5
+```{r, fig.height=10, fig.width=12}
+supp7<-cowplot::plot_grid(a1,a2,b1,b2, ncol=2, labels=c("A","","B",""), label_size = 20)
+supp7
 ```
 
-##Proliferation Supplementary
+#Proliferation Supplementary
+
+##8 SEB
 ```{r}
@@
     scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
-    labs(y="(%) Proliferating CD4 T cells")
+    labs(y="(%) Proliferating\nCD4 T cells")
 
@@
 first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20, rel_widths = c(3,1))
-second_row = cowplot::plot_grid(one, two, nrow=1, rel_widths = c(2,1), labels=c("B","C"), label_size = 20)
-third_row= cowplot::plot_grid(tf, sm, labels=c('F', 'G'), nrow=1, label_size = 20)
-
-supp6<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3)
+second_row = cowplot::plot_grid(one, two, nrow=1, rel_widths = c(2,1), labels=c("B"), label_size = 20)
+third_row= cowplot::plot_grid(tf, sm, labels=c('C'), nrow=1, label_size = 20)
 ```
 
+```{r, fig.height=12, fig.width=10}
+supp8<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3)
+supp8
+```
+
+##9 LTBI Peptide
 ```{r}
 ogphen<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
+
+#Cytokine
 melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._.", names(ogphen)))
@@
 test$variable<-factor(test$variable, levels=c(
-"T_G_x" ,       
-"T_x_x" ,         
-"x_G_x" ,         
-"T_G_4" ,
-"T_x_4" ,         
-"x_G_4" ,         
-"x_x_4"          
-))
+"T_G_x" , "T_x_x" , "x_G_x" ,         
+"T_G_4" , "T_x_4" , "x_G_4" , 
+"x_x_4"))
+
 test$group<-"TH1/2"
@@
 
-supp7a<-booleanplot(test, "variable", "value", ltbi.cols)+
+a<-booleanplot(test, "variable", "value", ltbi.cols)+
     scale_x_discrete(labels = labels, breaks=breaks)+
+    theme(legend.position = "right", legend.key.size = unit(6, "cm"))+
     facet_grid(~group, scales="free_x", space="free_x")+
-    labs(y="(%) Proliferating CD4 T cells")
+    labs(y="(%) Proliferating\nMtb-Specific CD4 T cells")
 
-```
-
-```{r og phenotype, fig.width=16, fig.height=24}
-#graph phenotype
+#phenotype
 melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._._.", names(ogphen)))
@@
 test$variable<-factor(test$variable, levels=c(
-"t_x_3_4" ,       
-"t_x_3_x" ,         
-"t_x_x_4" ,         
-"t_x_x_x" ,
-"t_g_3_4" ,         
-"t_g_3_x" ,         
-"t_g_x_4" ,         
-"t_g_x_x" ,
-"x_g_3_4" ,         
-"x_g_3_x" ,         
-"x_g_x_4" ,         
-"x_g_x_x" ,
-"x_x_3_4" ,          
-"x_x_3_x" ,         
-"x_x_x_4"           
-))
-
-T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
-G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
-c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
-c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")
-
-labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
-breaks <- levels(test$variable)
+"t_x_3_4" , "t_x_3_x" , "t_x_x_4" , "t_x_x_x" ,
+"t_g_3_4" , "t_g_3_x" , "t_g_x_4" , "t_g_x_x" ,
+"x_g_3_4" , "x_g_3_x" , "x_g_x_4" , "x_g_x_x" ,
+"x_x_3_4" , "x_x_3_x" , "x_x_x_4"))
 
@@
 
-supp7b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
-    scale_x_discrete(labels = labels, breaks=breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=6)+
-    labs(y="(%) Proliferating Mtb-Specific cells")
+b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
+    scale_x_discrete(labels = phen_labels, breaks=phen_breaks)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=12)+
+    labs(y="(%) Proliferating\nMtb-Specific CD4 T cells")
 ```
 
-```{r, fig.width=12, fig.height=12}
-supp7<-cowplot::plot_grid(supp7a, supp7b, nrow=2, labels=c("A", "B"), rel_widths = c(3,4))
-supp7
+```{r, fig.height=10, fig.width=10}
+supp9<-cowplot::plot_grid(a, b, nrow=2, labels=c("A", "B"), label_size = 20, rel_widths = c(9,10))
+supp9
 ```
 
-###Comparing SEB-proliferating cells to non-proliferating cells in UN 
-```{r, fig.width=12, fig.height=6}
-un<-filter_ltbi(dplyr::filter(OG, Stim=="UN"))
-un$cell<-"Total"
-seb<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="SEB"))
-seb$cell<-"SEB"
-long<-rbindlist(list(un, seb), fill=TRUE)
-long$cell<-factor(long$cell, levels=c("Total", "SEB"))
-long<-data.frame(long)
-
-melt<-melt(long, id.vars=c("TB", "SM", "disease", "Donor", "cell"), 
-           measure.vars=c("T.bet", "GATA3", "CXCR3","CCR4"))
-melt$x<-paste(melt$SM, melt$cell)
-melt$x<-factor(melt$x, levels=c("SM- Total", "SM- SEB","SM+ Total", "SM+ SEB"))
-
-ggplot(melt, aes(x=x, y=value, fill=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    facet_wrap(~variable, scales="free", nrow=1)+
-    theme_classic()+
-    scale_fill_manual(values = ltbi.cols)+
-    scale_x_discrete(labels=c("Total", "SEB", "Total", "SEB"))+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
-    stat_compare_means(comparisons=list(c("SM- Total", "SM- SEB"),c("SM+ Total", "SM+ SEB")) 
-                       ,method="wilcox.test", label="p.format")+
-    labs(x="", y="")
-```
-
-###Comparing CFP-10 and ESAT-6-proliferating cells to non-proliferating cells in UN
-```{r, fig.width=12, fig.height=6}
-un<-filter_ltbi(dplyr::filter(OG, Stim=="UN"))
-un$cell<-"Total"
-ce<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
-ce$cell<-"PEP"
-long<-rbindlist(list(un, ce), fill=TRUE)
-long$cell<-factor(long$cell, levels=c("Total", "PEP"))
-long<-data.frame(long)
-
-melt<-melt(long, id.vars=c("TB", "SM", "disease", "Donor", "cell"), 
-           measure.vars=c("T.bet", "GATA3", "CXCR3","CCR4"))
-melt$x<-paste(melt$SM, melt$cell)
-melt$x<-factor(melt$x, levels=c("SM- Total", "SM- PEP", 
-                                "SM+ Total", "SM+ PEP"))
-
-ggplot(melt, aes(x=x, y=value, fill=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    facet_wrap(~variable, scales="free", nrow=1)+
-    theme_classic()+
-    scale_fill_manual(values = ltbi.cols)+
-    scale_x_discrete(labels=c("Total", "CFP-10 &\nESAT-6", "Total", "CFP-10 &\nESAT-6"))+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
-    stat_compare_means(comparisons=list(c("SM- Total", "SM- PEP"),c("SM+ Total", "SM+ PEP")), 
-        method="wilcox.test", label="p.format")+
-    labs(x="", y="")
-
-```
-
-###Comparing CFP-10 and ESAT-6-proliferating to non-proliferating cells
-```{r, fig.width=12, fig.height=6}
+##10 CFP-10 and ESAT-6-proliferating to other cells
+```{r}
 un<-filter_ltbi(dplyr::filter(OGnope, Stim=="PEP"))
@@
 
-ggplot(melt, aes(x=x, y=value, fill=SM))+
-    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
-    facet_wrap(~variable, scales="free", nrow=1)+
-    theme_classic()+
-    scale_fill_manual(values = ltbi.cols)+
-    scale_x_discrete(labels=c("Non-\nProliferating", "Proliferating", 
-                              "Non-\nProliferating", "Proliferating"))+
-    theme(panel.border = element_rect(fill = NA, colour = "black"))+
-    theme(text = element_text(size=16))+
-    theme(plot.title = element_text(hjust = 0.5)) +
-    stat_compare_means(comparisons=list(c("SM- Non-Proliferating", "SM- Proliferating"),
+a1<-baseplot(melt[(melt$variable %in% c("T.bet", "GATA3")),], "x", "value", ltbi.cols)+
+    facet_wrap(~variable, nrow=1)+
+    scale_x_discrete(labels=c("NP", "P", 
+                              "NP", "P"))+
+    geom_signif(comparisons=list(c("SM- Non-Proliferating", "SM- Proliferating"),
                                  c("SM+ Non-Proliferating", "SM+ Proliferating")),
-        method="wilcox.test", label="p.format")+
-    labs(x="", y="")
+                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = 100)+
+    labs(y="(%) CD4 T cells")+
+    coord_cartesian(ylim=c(0,105))
+
+a2<-baseplot(melt[(melt$variable %in% c("CXCR3", "CCR4")),], "x", "value", ltbi.cols)+
+    facet_wrap(~variable, nrow=1)+
+    scale_x_discrete(labels=c("NP", "P", 
+                              "NP", "P"))+
+    geom_signif(comparisons=list(c("SM- Non-Proliferating", "SM- Proliferating"),
+                                 c("SM+ Non-Proliferating", "SM+ Proliferating")),
+                test = "wilcox.test", map_signif_level=TRUE, textsize = 8)+
+    coord_cartesian(ylim=c(0,105))
 
@@
 ```{r}
+un<-filter_ltbi(dplyr::filter(OGPHEN, Stim=="SEB"))
+un$cell<-"SEB"
+ce<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
+ce$cell<-"PEP"
+long<-rbindlist(list(un, ce), fill=TRUE)
+long$cell<-factor(long$cell, levels=c("SEB", "PEP"))
+long<-data.frame(long)
+
+melt<-melt(long, id.vars=c("TB", "SM", "disease", "Donor", "cell"), 
+           measure.vars=c("T.bet", "GATA3", "CXCR3","CCR4"))
+melt$x<-paste(melt$SM, melt$cell)
+melt$x<-factor(melt$x, levels=c("SM- SEB", "SM- PEP", 
+                                "SM+ SEB", "SM+ PEP"))
+
+b1<-baseplot(melt[(melt$variable %in% c("T.bet", "GATA3")),], "x", "value", ltbi.cols)+
+    facet_wrap(~variable, nrow=1)+
+    scale_x_discrete(labels=c("SEB", "CFP-10 &\nESAT-6", "SEB", "CFP-10 &\nESAT-6"))+
+    labs(y="(%) CD4 T cells")+
+    geom_signif(comparisons=list(c("SM- SEB", "SM- PEP"),c("SM+ SEB", "SM+ PEP")),
+                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = 100)+
+    coord_cartesian(ylim=c(0,105))
+
+#not significant so didn't include
+b2<-baseplot(melt[(melt$variable %in% c("CXCR3", "CCR4")),], "x", "value", ltbi.cols)+
+    facet_wrap(~variable, nrow=1)+
+    scale_x_discrete(labels=c("SEB", "CFP-10 &\nESAT-6", "SEB", "CFP-10 &\nESAT-6"))+
+    coord_cartesian(ylim=c(0,100))
+```
+
+```{r, fig.height=10, fig.width=12}
+supp10<-cowplot::plot_grid(a1,a2, b1,b2, ncol=2, labels=c("A", "","B", ""), label_size = 20)
+supp10
+```
+
+##11 Dendrograms
+```{r}
+lum<-filter_ltbi(subset(luminex, luminex$Stim=="PEP"))
+names(lum)<-gsub("IL", "IL-", names(lum))
+names(lum)<-gsub("IFNg", "IFNγ", names(lum))
+names(lum)<-gsub("TNFa", "TNFα", names(lum))
+
 lum_x<-lum[lum$SM=="SM-",]
@@
 setwd("/Users/tarynam/Desktop/Manuscript1")
-cowplot::save_plot("supp1.png", supp1, base_height=14, base_width=16)
-cowplot::save_plot("supp2.png", supp2, base_height=28, base_width=16)
-cowplot::save_plot("supp3.png", supp3, base_height=7, base_width=16)
-cowplot::save_plot("supp4.png", supp4, base_height=14, base_width=16)
-cowplot::save_plot("supp5.png", supp5, base_height=14, base_width=16)
-cowplot::save_plot("supp6.png", supp6, base_height=14, base_width=16)
-cowplot::save_plot("supp7.png", supp6, base_height=14, base_width=16)
+cowplot::save_plot("supp2.png", supp2, base_height=8, base_width=10)
+cowplot::save_plot("supp3.png", supp3, base_height=14, base_width=10)
+cowplot::save_plot("supp4.png", supp4, base_height=12, base_width=12)
+cowplot::save_plot("supp5.png", supp5, base_height=15, base_width=16)
+cowplot::save_plot("supp6.png", supp6, base_height=12, base_width=12)
+cowplot::save_plot("supp7.png", supp7, base_height=15, base_width=16)
+cowplot::save_plot("supp8.png", supp8, base_height=14, base_width=16)
+cowplot::save_plot("supp9.png", supp9, base_height=14, base_width=16)
+cowplot::save_plot("supp10.png", supp10, base_height=15, base_width=16)
 ```
@@
 test<-filter(test, Stim=="SEB")
-ggplot(test, aes(x=OG.x, y=Tbet.y, color=TB))+geom_point()+stat_smooth(method="lm")
-ggplot(test, aes(x=OG.x, y=GATA3.y, color=TB))+geom_point()+stat_smooth(method="lm")
-ggplot(test, aes(x=Tbet.y, y=GATA3.y, color=TB))+geom_point()+stat_smooth(method="lm")
+ggplot(test, aes(x=OG.x, y=Tbet.y, color=TB))+geom_point()+stat_smooth(method="lm")+
+    labs(x="Proliferation (%) CD4 T cells",
+         y="T-bet (%) Proliferating CD4 T cells")
+ggplot(test, aes(x=OG.x, y=GATA3.y, color=TB))+geom_point()+stat_smooth(method="lm")+
+        labs(x="Proliferation (%) CD4 T cells",
+         y="GATA3 (%) Proliferating CD4 T cells")
+ggplot(test, aes(x=Tbet.y, y=GATA3.y, color=TB))+geom_point()+stat_smooth(method="lm")+
+    labs(x="T-bet (%) Proliferating CD4 T cells",
+         y="GATA3 (%) Proliferating CD4 T cells")
+
+part1<-filter(OGPHEN, Stim=="UN")
+part2<-filter(OGPHEN, Stim=="SEB")
+test<-merge(part1, part2, by=c("Donor", "TB", "SM"))
+ggplot(test, aes(x=Tbet.x, y=Tbet.y, color=TB))+geom_point()+stat_smooth(method="lm")+
+        labs(x="(%) CD4 T cells",
+         y="(%) Proliferating CD4 T cells",
+         title="T-bet")
+ggplot(test, aes(x=GATA3.x, y=GATA3.y, color=TB))+geom_point()+stat_smooth(method="lm")+
+        labs(x="(%) CD4 T cells",
+         y="(%) Proliferating CD4 T cells",
+         title="GATA3")
 
@@
 test<-merge(part1, part2, by=c("Donor", "TB", "SM"))
-ggplot(test, aes(x=Tbet.x, y=Tbet.y))+geom_point()+stat_smooth(method="lm")
-ggplot(test, aes(x=GATA3.x, y=GATA3.y))+geom_point()+stat_smooth(method="lm")
+ggplot(test, aes(x=Tbet.x, y=Tbet.y, color=TB))+geom_point()+stat_smooth(method="lm")+
+        labs(x="(%) ICS CD4 T cells",
+         y="(%) Proliferating CD4 T cells",
+         title="T-bet")
+ggplot(test, aes(x=GATA3.x, y=GATA3.y, color=TB))+geom_point()+stat_smooth(method="lm")+
+        labs(x="(%) ICS CD4 T cells",
+         y="(%) Proliferating CD4 T cells",
+         title="GATA3")
 ```

### 2019-11-11 PaperFigures.Rmd
re-ordered TH groups

@@
     coord_cartesian(ylim=c(0,1500))+
+    theme(plot.margin = unit(c(0,0,0,0), "cm"))+
     labs(y="pg/mL")
 b<-baseplot(melt[melt$variable=="IFNγ",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
-    coord_cartesian(ylim=c(0,18000))+
-    theme(plot.margin=margin(l=-0.5,unit="cm"))
+    theme(plot.margin = unit(c(0,0,0,0), "cm"))+
+    coord_cartesian(ylim=c(0,18000))
 c<-baseplot(melt[melt$variable=="IL-21",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
-    coord_cartesian(ylim=c(0,60))+
-    theme(plot.margin=margin(l=-0.5,unit="cm"))
+    theme(plot.margin = unit(c(0,0,0,0), "cm"))+
+    coord_cartesian(ylim=c(0,60))
 d<-baseplot(melt[melt$variable=="IL-17A",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
-    coord_cartesian(ylim=c(0,20))+
-    theme(plot.margin=margin(l=-0.5,unit="cm"))
+    theme(plot.margin = unit(c(0,0,0,0), "cm"))+
+    coord_cartesian(ylim=c(0,20))
 e<-baseplot(melt[melt$variable=="IL-22",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
-    coord_cartesian(ylim=c(0,40))+
-    theme(plot.margin=margin(l=-0.5,unit="cm"))
+    theme(plot.margin = unit(c(0,0,0,0), "cm"))+
+    coord_cartesian(ylim=c(0,40))
 f<-baseplot(melt[melt$variable=="IL-10",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
-    coord_cartesian(ylim=c(0,20))+
-    theme(plot.margin=margin(l=-0.5,unit="cm"))
+    theme(plot.margin = unit(c(0,0,0,0), "cm"))+
+    coord_cartesian(ylim=c(0,20))
 g<-baseplot(melt[melt$variable=="IL-4",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
-    coord_cartesian(ylim=c(0,40))+
-    theme(plot.margin=margin(l=-0.5,unit="cm"))
+    theme(plot.margin = unit(c(0,0,0,0), "cm"))+
+    coord_cartesian(ylim=c(0,40))
 h<-baseplot(melt[melt$variable=="IL-5",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
-    coord_cartesian(ylim=c(0,15))+
-    theme(plot.margin=margin(l=-0.5,unit="cm"))
+    theme(plot.margin = unit(c(0,0,0,0), "cm"))+
+    coord_cartesian(ylim=c(0,15))
 i<-baseplot(melt[melt$variable=="IL-13",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
-    coord_cartesian(ylim=c(0,400))+
-    theme(plot.margin=margin(l=-0.5,unit="cm"))
+    theme(plot.margin = unit(c(0,0,0,0), "cm"))+
+    coord_cartesian(ylim=c(0,400))
 
@@
 ```{r}
-setwd("/Users/tarynam/Desktop/Manuscript1/figures")
-cowplot::save_plot("figure8.png", lumplot, base_height=6, base_width=18)
+setwd("/Users/tarynam/Desktop/Mtb Manuscript/figures")
+cowplot::save_plot("figure8.png", lumplot, base_height=5, base_width=18)
 

### 2019-11-11 SuppFigures.Rmd
re-ordered TH groups

@@
 
+th1<-booleanplot2(test[test$th=="TH1",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
+    scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
+    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
+    coord_cartesian(ylim=c(0,25))+
+    theme(strip.text.y = element_blank(), strip.background.y = element_blank())+
+    labs(y="(%) CD4 T cells")
+
 th1.2<-booleanplot2(test[test$th=="TH1/2",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
@@
     coord_cartesian(ylim=c(0,3))+
-    theme(strip.text.y = element_blank(), strip.background.y = element_blank())+
-    labs(y="(%) CD4 T cells")
-
-th1<-booleanplot2(test[test$th=="TH1",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
-    scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
-    coord_cartesian(ylim=c(0,25))+
     theme(strip.text.y = element_blank(), strip.background.y = element_blank())
@@
     stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
-    coord_cartesian(ylim=c(0,3))
+    coord_cartesian(ylim=c(0,1.5))
 ```
@@
 ```{r, fig.height=10, fig.width=10}
-supp2<-cowplot::plot_grid(th1.2, th1, th2, ncol=3, rel_widths = c(8,3.5,3.5))
+supp2<-cowplot::plot_grid(th1, th1.2, th2, ncol=3, rel_widths = c(3,8,3.5))
 supp2
@@
 test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
-test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))
+test$th<-factor(test$th, levels=c("TH1", "TH1/2", "TH2"))
 
@@
     scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
-    coord_cartesian(ylim=c(0,.16))+
+    stat_compare_means(method="wilcox.test", label="p.signif", 
+                       hide.ns = TRUE, exact=FALSE, size=10, label.y=(.98*.12))+
+    coord_cartesian(ylim=c(0,.12))+
     theme(plot.margin=margin(l=1,unit="cm"))+
@@
     scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
-    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
-    coord_cartesian(ylim=c(0,.16))+
+    stat_compare_means(method="wilcox.test", label="p.signif", 
+                       hide.ns = TRUE, exact=FALSE, size=10, label.y=(.98*0.11))+
+    coord_cartesian(ylim=c(0,.11))+
     theme(plot.margin=margin(l=1,unit="cm"))+
@@
 ```{r}
-setwd("/Users/tarynam/Desktop/Manuscript1")
+setwd("/Users/tarynam/Desktop/Mtb Manuscript/")
 cowplot::save_plot("supp2.png", supp2, base_height=8, base_width=10)

### 2019-11-23 OG-Analysis_GD.Rmd
Added in Yerkes Donors

@@
     PBMC<-PBMC[grep("OG", PBMC$used),]
-mimosa<-dplyr::select(read.csv("OG_MIMOSA_clean.csv"), -X)
-    mimosa<-dplyr::filter(mimosa, celltype=="GD" & Exprate==0.01)
+mimosa<-dplyr::select(read.csv("OG_MIMOSA_complete.csv"), -X)
+    mimosa<-dplyr::filter(mimosa, celltype=="GD")
 ```
@@
 antigen.ref<-mimosa
-antigen.ref$id<-gsub("_Treatment", "", antigen.ref$Sample)
 antigen.ref<-arrange(antigen.ref, id)

### 2019-11-23 OG_cleaning.Rmd
Added in Yerkes Donors

@@
 l<-lapply(l, function(x) sum_boolean(x))
-l<-lapply(l, function(x) sum_boolean2(x))
 filenames<-ls(pattern="_OG")

### 2019-11-23 gamma-delta.Rmd
Added in Yerkes Donors

@@
 library(sjPlot)
+library(ggpmisc)
 source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
@@
 source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")
-```
-
-```{r}
-setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
-PBMC<-dplyr::select(read.csv("DonorInfo_clean.csv"), -X)
-    PBMC<-PBMC[grep("OG", PBMC$used),]
-    plot_3(PBMC[PBMC$SM=="SM",], "SchistosomaIntensity")+
-    labs(y="Eggs/gram")
 ```
@@
 exclude<-subset(include, include$SM!="X")
-keep<-plot_cell(exclude, "FS")+
+plot_cell(exclude, "FS")+
+    facet_grid(Stim~TB)+
+    scale_x_discrete(labels = c("CD4", "CD8", "CD4-CD8-"))+
+    labs(y="Functionality Score",
+         title="Excludes SM-")
+```
+
+```{r, fig.width=12, fig.height=8}
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
+compass<-read.csv("ICS_compass_complete_clean.csv")
+compass$TB<-factor(compass$TB, levels=c("N", "HC", "LTBI", "TB"))
+kable(pvals_SM(compass[compass$Stim=="SWAP",], "FS"), caption="SWAP")
+kable(pvals_SM(compass[compass$Stim=="SEA",], "FS"), caption="SEA")
+
+include<-subset(compass, compass$Stim %in% c("SEA", "SWAP"))
+plot_cell(include, "FS")+
+    facet_grid(Stim~TB)+
+    scale_x_discrete(labels = c("CD4", "CD8", "CD4-CD8-"))+
+    labs(y="Functionality Score",
+         title="Includes SM-")
+
+exclude<-subset(include, include$SM!="X")
+plot_cell(exclude, "FS")+
     facet_grid(Stim~TB)+
@@
 
-mimosa<-dplyr::select(read.csv("OG_MIMOSA_clean.csv"), -X)
-    mimosa<-dplyr::filter(mimosa, celltype=="GD" & Exprate==0.01)
+mimosa<-dplyr::select(read.csv("OG_MIMOSA_complete.csv"), -X)
+```
+
+###Proliferation 
+```{r}
+include<-subset(CD3, CD3$Stim%in%c("SEA","SWAP"))
+include$TB<-factor(include$TB, levels=c("N", "HC", "LTBI", "TB"))
+pvals_SM(include, "OG")
+a<-plot_3(include, "OG", "TB")+
+    facet_grid(~Stim)+
+    labs(y="(%) CD3 T cells")
+exclude<-subset(include, include$SM!="X")
+b<-plot_3(exclude, "OG", "TB")+
+    facet_grid(~Stim)+
+    labs(y="(%) CD3 T cells")
+cowplot::plot_grid(a,b, ncol=2)
+
+a<-ggplot(include, aes(x=Stim, y=OG))+
+    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
+    geom_jitter(width=.1,height=0, shape=16,size=2)+
+    scale_fill_manual(values =c("N" = "#da70d6","HC" = "#00AFBB", "LTBI" = "#E7B800","TB"="#FC4E07"))+
+    facet_grid(~TB)+
+    theme_classic()+ theme(legend.position="none")+
+    theme(text = element_text(size=16)) + 
+    theme(plot.subtitle = element_text(hjust = 0.5)) +
+    stat_compare_means(label="p.signif")+
+    labs(y="(%) CD3 T cells")
+
+exclude<-subset(include, include$SM!="X")
+b<-ggplot(include, aes(x=Stim, y=OG))+
+    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
+    geom_jitter(width=.1,height=0, shape=16,size=2)+
+    scale_fill_manual(values =c("N" = "#da70d6","HC" = "#00AFBB", "LTBI" = "#E7B800","TB"="#FC4E07"))+
+    facet_grid(~TB)+
+    theme_classic()+ theme(legend.position="none")+
+    theme(text = element_text(size=16)) + 
+    theme(plot.subtitle = element_text(hjust = 0.5)) +
+    stat_compare_means(label="p.signif")+
+    labs(y="(%) CD3 T cells")
+
+cowplot::plot_grid(a,b)
+
 ```
@@
 
-###Starting Frequency of Gamma Delta Cells
-```{r}
-include<-subset(CD3, CD3$Stim=="UN")
-kable(pvals_SM(include, "GD"))
-a<-plot_3(include, "GD")+
-    coord_cartesian(ylim=c(0,35))+
-    labs(y="(%) CD3+ T cells",
-         title="Includes SM-")
-
-exclude<-subset(include, include$SM=="SM")
-b<-plot_3(exclude, "GD")+
-    coord_cartesian(ylim=c(0,35))+
-    labs(y="(%) CD3+ T cells",
-         title="Excludes SM-")
-
-cowplot::plot_grid(a,b)
-```
-
-###Proportion of proliferating Gamma Delta Cells
+###Proportion of Proliferating Cells
 ```{r, fig.width=12, fig.height=8}
@@
 include<-subset(CD3_OG, CD3_OG$Stim%in%c("SEA","SWAP"))
+
 melt<-melt(include, id.vars=c("TB", "SM", "Stim"), measure.vars=colnames)
 melt$variable<-gsub("GD", "γδ", melt$variable)
+melt$celltype<-gsub("", "",melt$variable)
+melt$TB<-factor(melt$TB, levels=c("N", "HC", "LTBI", "TB"))
 
-melt$celltype<-gsub("", "",melt$variable)
-
-a<-plot_3(melt, "value")+
+a<-plot_3(melt, "value", "TB")+
     facet_grid(Stim~variable)+
@@
 
-exclude<-subset(melt, melt$SM=="SM")
-b<-plot_3(melt, "value")+
-    facet_grid(Stim~variable, labeller = as_labeller(c("CD4"="CD4", "CD8"="CD8", "GD"="γδ")))+
+exclude<-subset(melt, melt$SM!="X")
+b<-plot_3(exclude, "value", "TB")+
+    facet_grid(Stim~variable)+
     labs(y="(%) Proliferating T cells",
@@
 a
+b
 ```
 
+###compare cell types
 ```{r, fig.width=12, fig.height=8}
@@
 melt$celltype<-gsub("", "",melt$variable)
+melt$TB<-factor(melt$TB, levels=c("N", "HC", "LTBI", "TB"))
 
-a<-plot_cell(melt, "value")+
+plot_cell(melt, "value")+
+    facet_grid(Stim~.)+
+    scale_x_discrete(labels = c("CD4", "CD8", "γδ"))+
+    labs(y="(%) Proliferating+ T cells",
+         title="total")
+
+plot_cell(melt, "value")+
     facet_grid(Stim~TB)+
@@
 
-exclude<-subset(melt, melt$SM=="SM")
-b<-plot_cell(melt, "value")+
+exclude<-subset(melt, melt$SM!="X")
+plot_cell(exclude, "value")+
     facet_grid(Stim~TB)+
@@
          title="Excludes SM-")
+```
 
-a
+###Starting Frequency of Gamma Delta Cells
+```{r}
+include<-subset(CD3, CD3$Stim=="UN")
+include$TB<-factor(include$TB, levels=c("N", "HC", "LTBI", "TB"))
+kable(pvals_SM(include, "GD"))
+a<-plot_3(include, "GD", "TB")+
+    coord_cartesian(ylim=c(0,35))+
+    labs(y="(%) CD3+ T cells",
+         title="Includes SM-")
+
+exclude<-subset(include, include$SM!="X")
+b<-plot_3(exclude, "GD", "TB")+
+    coord_cartesian(ylim=c(0,35))+
+    labs(y="(%) CD3+ T cells",
+         title="Excludes SM-")
+
+cowplot::plot_grid(a,b)
 ```
@@
 ###Proliferation of Gamma Delta Cells
-```{r}
+```{r, fig.height=8}
 include<-subset(gd, gd$Stim=="SEA")
+include$TB<-factor(include$TB, levels=c("N", "HC", "LTBI", "TB"))
 pvals_SM(include, "OG")
-a<-plot_3(include, "OG")+
+a<-plot_3(include, "OG", "TB")+
+    facet_grid(~Stim)+
+    labs(y="(%) γδ T cells")
+exclude<-subset(include, include$SM!="X")
+c<-plot_3(exclude, "OG", "TB")+
     facet_grid(~Stim)+
@@
 include<-subset(gd, gd$Stim=="SWAP")
+include$TB<-factor(include$TB, levels=c("N", "HC", "LTBI", "TB"))
 pvals_SM(include, "OG")
-b<-plot_3(include, "OG")+
-    facet_grid(~Stim)
+b<-plot_3(include, "OG", "TB")+
+    facet_grid(~Stim)+
+    labs(y="(%) γδ T cells")
+exclude<-subset(include, include$SM!="X")
+d<-plot_3(exclude, "OG", "TB")+
+    facet_grid(~Stim)+
+    labs(y="(%) γδ T cells")
 
-cowplot::plot_grid(a,b)
+cowplot::plot_grid(a,b, c,d, ncol=2)
 ```
@@
 ###Proliferation of CD4 T Cells
-```{r}
+```{r, fig.height=8}
 include<-subset(CD4, CD4$Stim=="SEA")
+include$TB<-factor(include$TB, levels=c("N", "HC", "LTBI", "TB"))
 pvals_SM(include, "OG")
-a<-plot_3(include, "OG")+
+a<-plot_3(include, "OG", "TB")+
+    facet_grid(~Stim)+
+    labs(y="(%) CD4 T cells")
+exclude<-subset(include, include$SM!="X")
+c<-plot_3(exclude, "OG", "TB")+
     facet_grid(~Stim)+
@@
 include<-subset(CD4, CD4$Stim=="SWAP")
+include$TB<-factor(include$TB, levels=c("N", "HC", "LTBI", "TB"))
 pvals_SM(include, "OG")
-b<-plot_3(include, "OG")+
+b<-plot_3(include, "OG", "TB")+
     facet_grid(~Stim)+
-    labs(y="")
+    labs(y="(%) CD4 T cells")
+exclude<-subset(include, include$SM!="X")
+d<-plot_3(exclude, "OG", "TB")+
+    facet_grid(~Stim)+
+    labs(y="(%) CD4 T cells")
 
-cowplot::plot_grid(a,b)
+cowplot::plot_grid(a,b, c,d, ncol=2)
 ```
@@
 ###Proliferation of CD8 T Cells
-```{r}
+```{r, fig.height=8}
 include<-subset(CD8, CD8$Stim=="SEA")
 pvals_SM(include, "OG")
-a<-plot_3(include, "OG")+
+include$TB<-factor(include$TB, levels=c("N", "HC", "LTBI", "TB"))
+a<-plot_3(include, "OG", "TB")+
+    facet_grid(~Stim)+
+    labs(y="(%) CD8 T cells")
+exclude<-subset(include, include$SM!="X")
+c<-plot_3(exclude, "OG", "TB")+
     facet_grid(~Stim)+
@@
 include<-subset(CD8, CD8$Stim=="SWAP")
+include$TB<-factor(include$TB, levels=c("N", "HC", "LTBI", "TB"))
 pvals_SM(include, "OG")
-b<-plot_3(include, "OG")+
+c<-plot_3(include, "OG", "TB")+
     facet_grid(~Stim)+
+    labs(y="(%) CD8 T cells")
+exclude<-subset(include, include$SM!="X")
+d<-plot_3(include, "OG", "TB")+
+    facet_grid(~Stim)+
+    labs(y="(%) CD8 T cells")
+
+cowplot::plot_grid(a,b, c,d, ncol=2)
+```
+
+```{r}
+table1<-xtabs(~SM+TB, gd_OG)
+table2<-xtabs(~SM+TB, gd_OG[!is.na(gd_OG$IFNg),])
+```
+
+###Total Cytokine
+```{r}
+include<-subset(gd_OG, gd_OG$Stim=="SWAP")
+include$TB<-factor(include$TB, levels=c("N", "HC", "LTBI", "TB"))
+kable(pvals_SM(include, "cyt"))
+a<-plot_3(include, "cyt", "TB")+
+    labs(y="(%) Proliferating γδ T cells")
+
+exclude<-subset(include, include$SM!="X")
+b<-plot_3(exclude, "cyt", "TB")+
+    labs(y="(%) Proliferating γδ T cells")+
     labs(y="")
@@
 include<-subset(gd_OG, gd_OG$Stim=="SWAP")
+include$TB<-factor(include$TB, levels=c("N", "HC", "LTBI", "TB"))
 kable(pvals_SM(include, "IL4"))
-a<-plot_3(include, "IL4")+
+a<-plot_3(include, "IL4", "TB")+
     labs(y="(%) Proliferating γδ T cells")
 
-exclude<-subset(include, include$SM=="SM")
-b<-plot_3(exclude, "IL4")+
+exclude<-subset(include, include$SM!="X")
+b<-plot_3(exclude, "IL4", "TB")+
+    labs(y="(%) Proliferating γδ T cells")+
+    labs(y="")
+
+cowplot::plot_grid(a,b)
+```
+
+###IFNg
+```{r}
+include<-subset(gd_OG, gd_OG$Stim=="SWAP")
+include$TB<-factor(include$TB, levels=c("N", "HC", "LTBI", "TB"))
+kable(pvals_SM(include, "IFNg"))
+a<-plot_3(include, "IFNg", "TB")+
+    labs(y="(%) Proliferating γδ T cells")
+
+exclude<-subset(include, include$SM!="X")
+b<-plot_3(exclude, "IFNg", "TB")+
     labs(y="(%) Proliferating γδ T cells")+
@@
 include<-subset(gd_OG, gd_OG$Stim=="SWAP")
+include$TB<-factor(include$TB, levels=c("N", "HC", "LTBI", "TB"))
 kable(pvals_SM(include, "TNFa"))
-a<-plot_3(include, "TNFa")+
+a<-plot_3(include, "TNFa", "TB")+
     labs(y="(%) Proliferating γδ T cells")
 
-exclude<-subset(include, include$SM=="SM")
-b<-plot_3(exclude, "TNFa")+
+exclude<-subset(include, include$SM!="X")
+b<-plot_3(exclude, "TNFa", "TB")+
     labs(y="(%) Proliferating γδ T cells")
 
-cowplot::plot_grid(a,NULL)
+cowplot::plot_grid(a,b)
 ```
@@
 include<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
+include$TB<-factor(include$TB, levels=c("N", "HC", "LTBI", "TB"))
 include<-filter(include, variable!="x_x_x")
@@
     geom_boxplot(outlier.shape = NA)+
-    scale_fill_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))+
+        scale_fill_manual(values =c("N" = "#da70d6","HC" = "#00AFBB", "LTBI" = "#E7B800","TB"="#FC4E07"))+
     scale_x_discrete(labels=labels, breaks=breaks)+
@@
 
-exclude<-subset(include, include$SM=="SM")
+exclude<-subset(include, include$SM!="X")
 ggplot(exclude, aes(x=variable, y=value, fill=TB))+
     geom_boxplot(outlier.shape = NA)+
-    scale_fill_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))+
+        scale_fill_manual(values =c("N" = "#da70d6","HC" = "#00AFBB", "LTBI" = "#E7B800","TB"="#FC4E07"))+
     scale_x_discrete(labels=labels, breaks=breaks)+
@@
 ```
+
+###Egg data
+```{r}
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
+PBMC<-dplyr::select(read.csv("DonorInfo_clean.csv"), -X)
+PBMC<-PBMC[grep("OG", PBMC$used),]
+plot_3(PBMC[PBMC$SM=="SM",], "SchistosomaIntensity", "TB")+
+    labs(y="Eggs/gram")+
+    stat_compare_means(comparisons=list(c("HC", "LTBI"), c("LTBI", "TB"), c("HC", "TB")))
+
+setwd("/Applications/Old Computer/Epi Project/Data_clean")
+egg<-dplyr::select(read.csv("complete-kenya-filledin.csv"), -X)%>%
+    dplyr::filter(SchistosomaPositive==1, !is.na(TB))
+plot_3(egg, "SchistosomaIntensity", "TB")+
+    labs(y="Eggs/gram")+
+    stat_compare_means(comparisons=list(c("HEALTHY CONTROL", "LTBI"), 
+                                        c("LTBI", "ACTIVE"), 
+                                        c("HEALTHY CONTROL", "ACTIVE")))
+
+DF<-merge(dplyr::select(PBMC, Donor, SchistosomaIntensity), gd_OG[gd_OG$Stim=="SWAP",], by="Donor")
+formula<-y~x
+
+ggplot(DF[DF$SM=="SM",], aes(x=IL4, y=SchistosomaIntensity, color=TB))+
+    geom_point()+
+    scale_color_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))+
+    stat_smooth(method="lm", formula=formula)+
+    stat_poly_eq(aes(label = paste(..rr.label..)), 
+               label.x.npc = "right", label.y.npc = 0.15,
+               formula = formula, parse = TRUE, size = 3)
+```
+
+```{r}
+p1<-dplyr::select(PBMC, Donor, SchistosomaIntensity)
+p2<-dplyr::filter(gd, Stim=="SWAP")%>%dplyr::select(Donor, OG)
+p3<-dplyr::select(gd_OG, -OG, -Sample, -id)%>%dplyr::filter(Stim=="SWAP" & SM=="SM")
+df<-merge(p1, p2, by="Donor")
+DF<-merge(df, p3, by="Donor")
+Mat<-as.matrix(dplyr::select(DF, -Donor, -TB, -SM, -Stim)%>%dplyr::filter(!is.na(IFNg)))
+M<-cor(Mat)
+corrplot(M)
+```
+
+#Comparing N to SM- to SM+
+
+```{r}
+include<-subset(CD3, CD3$Stim%in%c("SEA","SWAP"))
+include$SM<-factor(include$SM, levels=c("N", "X", "SM"))
+plot_3(include, "OG", "SM")+
+    facet_grid(~Stim)+
+    stat_compare_means(comparisons = list(c("N", "X"),c("N", "SM"),c("X", "SM")),
+        label="p.signif")+
+    labs(y="(%) Proliferating T cells",
+         x="SM",
+         title="CD3")
+
+include<-subset(CD4, CD4$Stim%in%c("SEA","SWAP"))
+include$SM<-factor(include$SM, levels=c("N", "X", "SM"))
+plot_3(include, "OG", "SM")+
+    facet_grid(~Stim)+
+    stat_compare_means(comparisons = list(c("N", "X"),c("N", "SM"),c("X", "SM")),
+        label="p.signif")+
+    labs(y="(%) Proliferating T cells",
+         x="SM",
+         title="CD4")
+
+include<-subset(CD8, CD8$Stim%in%c("SEA","SWAP"))
+include$SM<-factor(include$SM, levels=c("N", "X", "SM"))
+plot_3(include, "OG", "SM")+
+    facet_grid(~Stim)+
+    stat_compare_means(comparisons = list(c("N", "X"),c("N", "SM"),c("X", "SM")),
+        label="p.signif")+
+    labs(y="(%) Proliferating T cells",
+         x="SM",
+         title="CD8")
+
+include<-subset(gd, gd$Stim%in%c("SEA","SWAP"))
+include$SM<-factor(include$SM, levels=c("N", "X", "SM"))
+plot_3(include, "OG", "SM")+
+    facet_grid(~Stim)+
+    stat_compare_means(comparisons = list(c("N", "X"),c("N", "SM"),c("X", "SM")),
+        label="p.signif")+
+    labs(y="(%) Proliferating T cells",
+         x="SM",
+         title="GD")
+```
+
+```{r}
+include<-subset(CD3_OG, CD3_OG$Stim%in%c("SEA","SWAP"))
+colnames<-c("CD4", "CD8", "GD")
+
+melt<-melt(include, id.vars=c("TB", "SM", "Stim"), measure.vars=colnames)
+melt$variable<-gsub("GD", "γδ", melt$variable)
+melt$celltype<-gsub("", "",melt$variable)
+melt$TB<-factor(melt$TB, levels=c("N", "X", "SM"))
+
+plot_3(melt, "value", "SM")+
+    facet_grid(Stim~variable)+
+    stat_compare_means(comparisons = list(c("N", "X"),c("N", "SM"),c("X", "SM")),
+        label="p.signif")+
+    labs(y="(%) Proliferating T cells",
+         x="SM",
+         title="Includes SM-")
+```

### 2020-05-01 PBMC.Rmd
after review

@@
 #trim the data set to only those study IDs in the lab notes and select the columns you want
-    QFT<-dplyr::select(QFT, Donor, age, sex, QFT, SchistosomaIntensity)%>%
+    QFT<-dplyr::select(QFT, Donor, age, sex, QFT, HB, SchistosomaIntensity)%>%
         dplyr::filter(Donor %in% pbmc$Donor)

### 2020-05-01 PaperFigures.Rmd
after review

@@
----
+phen---
 title: "Paper Figures"

### 2020-05-01 SuppFigures.Rmd
after review

@@
 #ICS Supplementary
+```{r}
+pep<-plot_filter(CYT[(CYT$Stim=="PEP" & CYT$TB!="HC"),])
+my_comparisons <- list(c("HC","LTBI"), 
+                           c("TB", "LTBI"), 
+                           c("HC", "TB"))
+
+ggplot(pep, aes(x=TB, y=IFNg, fill=TB))+
+    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA)+
+    scale_fill_manual(values =c("#2166ac", "#b2182b"))+
+    theme_classic()+ theme(legend.position="none")+
+    theme(text = element_text(size=16)) + 
+    theme(plot.subtitle = element_text(hjust = 0.5)) +
+    stat_compare_means(method="wilcox.test", label.x=1.3, size=10, label="p.format")
+
+```
 

### 2020-05-01 gd_paperfigures.Rmd
first commit

---
title: "Untitled"
author: "Taryn McLaughlin"
date: "1/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, comment=NA, warning=FALSE, echo=FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(tidyr)
library(reshape2)
library(cowplot)
library(PMCMR)
library(COMPASS)
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")
remove<-c("HD10" , "HD30" , "HD211" , "HD341" , "HD382" , "HD396", "HD398" ,"HD421", "HD446" ,"HD469", "HD478" ,"HD479", "HD487" ,"HD495" ,"HD211" ,"HD384", "HD454")
gd_read<-function(filename){
    library(dplyr) #load it in the function to overcome namespace issues
    datatable<-dplyr::select(read.csv(filename), -X)
    datatable<-filter(datatable, Stim %in% c("SEA", "SWAP")) #No SM data in primary files
    datatable<-filter(datatable, SM!="X") #No SM data in primary files
    datatable<-datatable[!(datatable$Donor %in% remove),]
    datatable$SM<-factor(datatable$SM, levels=c("N", "SM"))
    datatable$TB<-factor(datatable$TB, levels=c("N", "HC", "LTBI", "TB"))
    datatable
    datatable
}
#old
#scale_fill_manual(values =c("N" = "#da70d6","HC" = "#00AFBB", "LTBI" = "#E7B800","TB"="#FC4E07"))+
#new
#scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
```

```{r}
library(tableone)
nk<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/DonorInfo_clean.csv")
sm<-nk[nk$SM=="SM",]
yerkes<-read.csv("/Applications/Old Computer/Day Lab/Flow/Yerkes Donor Info.csv")
yerkes<-yerkes[yerkes$BCG.status=="N",]
yerkes<-yerkes[!(is.na(yerkes$IgG)),]
keep<-gsub("HD","",nk$Donor[nk$TB=="N"])
yerkes<-yerkes[yerkes$Donor %in% keep,]
gd<-rbindlist(list(sm, yerkes), fill=TRUE)
gd$TB[is.na(gd$TB)]<-"N"
gd$SM[is.na(gd$SM)]<-"SM"

#Create a variable list which we want in Table 1
listVars <- c("age", "sex", "SchistosomaIntensity", "QFT")
#Define categorical variables
catVars <- c( "sex")
mediqr<- c("age", "SchistosomaIntensity","QFT")
#Total Population
table1 <- CreateTableOne(vars = listVars, data = gd, strata = "TB",factorVars = catVars)
tab<-print(table1, showAllLevels = TRUE, nonnormal = mediqr, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
## Save to a CSV file
write.csv(tab, file = "/Users/tarynam/Desktop/manuscript-sm/cohort.csv")
```

```{r}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
compass<-rbind(gd_read("ICS_compass_sea.csv"), gd_read("ICS_compass_swap.csv"))
cd4<-CYT<-gd_read("ICS_CD4_Cytokine_clean.csv")
cd8<-CYT<-gd_read("ICS_CD8_Cytokine_clean.csv")
cd3<-CYT<-gd_read("ICS_CD3_Cytokine_clean.csv")
neg<-CYT<-gd_read("ICS_GD_Cytokine_clean.csv")
gd<-gd_read("OG_gd_bulk_summary_clean.csv")
gd_OG<-gd_read("OG_gd_OG_summary_clean_NAs.csv")
CD3<-gd_read("OG_CD3_bulk_summary_clean.csv")
CD3_OG<-gd_read("OG_CD3_OG_summary_clean.csv")
CD4<-gd_read("OG_CD4_bulk_summary_clean.csv")
CD8<-gd_read("OG_CD8_bulk_summary_clean.csv")
mimosa<-read.csv("OG_MIMOSA_complete.csv")
load("/Users/tarynam/Desktop/Flow-scripts/Computational Packages/COMPASS/all_sea2.RData")
load("/Users/tarynam/Desktop/Flow-scripts/Computational Packages/COMPASS/all_SWAP2.RData")
load("/Users/tarynam/Desktop/Flow-scripts/Computational Packages/COMPASS/all_pma2.RData")
```

# ICS  
For whatever very annoying reason I can't seem to put the compass plots into the grids I make. I couldnt' with the corrplot either so not too surprised.  
```{r compass heatmaps sea}
fit_sea$data$meta$celltype<-gsub("GD", "CD4-CD8-", fit_sea$data$meta$celltype)
fit_sea$data$meta$celltype<-factor(fit_sea$data$meta$celltype, levels=c("CD4", "CD8", "CD4-CD8-"))

Var2 = c("#c3c3c3", "#878787" , "#000000")
names(Var2) = c("CD4", "CD8", "CD4-CD8-")
ann_colors_rows = list(celltype = Var2)

plot(fit_sea, palette = colorRampPalette(c("white", "#3f0086"))(50),
     row_annotation = "celltype", row_annotation_colors = ann_colors_rows, border_color=NA,
     markers=c("IL13", "IL4", "TNFa", "IFNg"), order_by_max_functionality = FALSE, threshold = -1)

```

```{r compass heatmaps swap}
fit_swap$data$meta$celltype<-gsub("GD", "CD4-CD8-", fit_swap$data$meta$celltype)
fit_swap$data$meta$celltype<-factor(fit_swap$data$meta$celltype, levels=c("CD4", "CD8", "CD4-CD8-"))

Var2 = c("#c3c3c3", "#878787" , "#000000")
names(Var2) = c("CD4", "CD8", "CD4-CD8-")
ann_colors_rows = list(celltype = Var2)

plot(fit_swap, palette = colorRampPalette(c("white", "#3CB371"))(50),
     row_annotation = "celltype", row_annotation_colors = ann_colors_rows, border_color=NA,
     markers=c("IL13", "IL4", "TNFa", "IFNg"), order_by_max_functionality = FALSE, threshold = -1)
```

```{r literally just to get the grey bars}
fit_swap$data$meta$celltype<-gsub("GD", "CD4-CD8-", fit_swap$data$meta$celltype)
M<-fit_swap$fit$mean_gamma
cytlabels<-data.frame(fit_swap$fit$categories[,1:4])
for(n in 1:ncol(cytlabels)){
    cytlabels[,n]<-as.factor(as.character(cytlabels[,n]))
}
cytlabels<-cytlabels[1:15,]
cytlabels2<-select(cytlabels, IL13, IL4, TNFa, IFNg)
names(cytlabels2)<-c("IL13", "IL4", "TNFα", "IFNγ")
row_ann <- fit_swap$data$meta
M<-M[which(row_ann$SM!="X"),]
M<-M[,1:15]
rownames(row_ann)<-row_ann$unique
row_ann<-row_ann[which(row_ann$SM!="X"),]
row_ann<-select(row_ann, celltype, TB)

Var1 = c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b")
names(Var1) = c("N", "HC", "LTBI", "TB")

Var2 = c("#A9A9A9", "#696969" , "#000000")
names(Var2) = c("CD4", "CD8", "CD4-CD8-")

ann_colors_rows = list(celltype = Var2)

Var3 = c("white", "#A9A9A9")
names(Var3) = c("0", "1")
ann_colors_cols = list(IFNγ = Var3, TNFα=Var3, IL4=Var3, IL13=Var3)

pheatmap(M, cluster_rows=FALSE, cluster_cols=FALSE,
         color = colorRampPalette(c("white", "navy"))(50),
         annotation = cytlabels2, annotation_colors = ann_colors_cols,
         show_rownames=FALSE,
         row_annotation = row_ann, row_annotation_colors = ann_colors_rows)
```

```{r fig one, fig.width=6, fig.height=3}
melt<-melt(compass, id.vars=c("group","celltype", "Stim"), measure.vars=c("FS", "PFS"))

a<-ggplot(melt[melt$celltype=="CD4",], aes(x=Stim, y=value))+
    geom_boxplot(aes(fill=Stim),size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values =c("SEA" = "#5700b8","SWAP" = "#3CB371"))+
    facet_grid(~variable)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(subtitle="CD4",y="Score",
         x="")+
    coord_cartesian(ylim=c(0,.4))

b<-ggplot(melt[melt$celltype=="CD8",], aes(x=Stim, y=value))+
    geom_boxplot(aes(fill=Stim),size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values =c("SEA" = "#5700b8","SWAP" = "#3CB371"))+
    facet_grid(~variable)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(label="p.signif", method="wilcox.test", size=8, label.x=1.4, label.y=.35)+
    labs(subtitle="CD8",y="Score",
         x="")+
    coord_cartesian(ylim=c(0,.4))

c<-ggplot(melt[melt$celltype=="GD",], aes(x=Stim, y=value))+
    geom_boxplot(aes(fill=Stim),size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values =c("SEA" = "#5700b8","SWAP" = "#3CB371"))+
    facet_grid(~variable)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(label="p.signif", method="wilcox.test", size=8, label.x=1.4, label.y=.35)+
    labs(subtitle="CD4-CD8-",y="Score",
         x="")+
    coord_cartesian(ylim=c(0,.4))

bottom<-cowplot::plot_grid(a, b, c, nrow=1, labels=c("", "", ""), label_size = 24)
fig1<-cowplot::plot_grid(NULL, NULL, bottom, labels=c("A", "B", "C"), nrow=3, label_size=24, rel_heights = c(1.5,6,2.5))
bottom
```
```{r, supp fig 2}
cd4$celltype<-"CD4"
cd4$Cyt.Freq<-rowSums(cd4[grep("_", names(cd4))[1:15]])
cd8$celltype<-"CD8"
cd8$Cyt.Freq<-rowSums(cd8[grep("_", names(cd8))[1:15]])
neg$celltype<-"CD4-CD8-"
neg$Cyt.Freq<-rowSums(neg[grep("_", names(neg))[1:15]])
ics<-rbindlist(list(cd4, cd8, neg), fill=TRUE)
ics$celltype<-factor(ics$celltype, levels=c("CD4", "CD8", "CD4-CD8-"))

supp2<-ggplot(ics, aes(x=Stim, y=Cyt.Freq))+
    geom_boxplot(aes(fill=Stim),size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values =c("SEA" = "#5700b8","SWAP" = "#3CB371"))+
    facet_wrap(~celltype, scales="free")+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(y="Cytokine+ (%)",
         x="")

melt<-melt(ics, id.vars=c("celltype", "Stim"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
melt$variable<-factor(melt$variable, levels=c("Cyt.Freq", "IFNg", "TNFa", "IL4", "IL13"))
cyt.groups<-as_labeller(c("IL4" = "IL-4", "IFNg" = "IFNγ", "TNFa" = "TNFα", "IL13"="IL-13", "Cyt.Freq"="Total"))

a<-ggplot(melt[melt$celltype=="CD4",], aes(x=Stim, y=value))+
    geom_boxplot(aes(fill=Stim),size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values =c("SEA" = "#5700b8","SWAP" = "#3CB371"))+
    facet_grid(~variable, scales="free", labeller=cyt.groups)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(label="p.signif", method="wilcox.test", size=8, label.x=1.4, label.y=.2)+
    labs(subtitle="CD4",y="Cytokine Frequency (%)",
         x="")

b<-ggplot(melt[melt$celltype=="CD8",], aes(x=Stim, y=value))+
    geom_boxplot(aes(fill=Stim),size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values =c("SEA" = "#5700b8","SWAP" = "#3CB371"))+
    facet_grid(~variable, scales="free", labeller=cyt.groups)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(label="p.signif", method="wilcox.test", size=8, label.x=1.4, label.y=.25)+
    labs(subtitle="CD8",y="Cytokine Frequency (%)",
         x="")

c<-ggplot(melt[melt$celltype=="CD4-CD8-",], aes(x=Stim, y=value))+
    geom_boxplot(aes(fill=Stim),size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values =c("SEA" = "#5700b8","SWAP" = "#3CB371"))+
    facet_grid(~variable, scales="free", labeller=cyt.groups)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(label="p.signif", method="wilcox.test", size=8, label.x=1.4, label.y = 15)+
    labs(subtitle="CD4-CD8-",y="Cytokine Frequency (%)",
         x="")+
    coord_cartesian(ylim=c(0,16))

ics_celltype<-cowplot::plot_grid(a, b, c, ncol=1, labels=c("A", "B", "C"), label_size = 26)
```

```{r supp fig 2 pma compass}
fit_pma$data$meta$celltype<-gsub("GD", "CD4-CD8-", fit_pma$data$meta$celltype)
fit_pma$data$meta$celltype<-factor(fit_pma$data$meta$celltype, levels=c("CD4", "CD8", "CD4-CD8-"))

Var2 = c("#c3c3c3", "#878787" , "#000000")
names(Var2) = c("CD4", "CD8", "CD4-CD8-")
ann_colors_rows = list(celltype = Var2)

plot(fit_pma, palette = colorRampPalette(c("white", "#000080"))(50),
     row_annotation = "celltype", row_annotation_colors = ann_colors_rows, border_color=NA,
     markers=c("IL13", "IL4", "TNFa", "IFNg"), order_by_max_functionality = FALSE, threshold = -1)

datatable<-dplyr::select(read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/ICS_compass_pma.csv"), -X)
datatable<-filter(datatable, Stim =="PMA") #No SM data in primary files
datatable<-filter(datatable, SM!="X") #No SM data in primary files
datatable<-datatable[!(datatable$Donor %in% remove),]
datatable$SM<-factor(datatable$SM, levels=c("N", "SM"))
datatable$TB<-factor(datatable$TB, levels=c("N", "HC", "LTBI", "TB"))
melt<-melt(datatable, id.vars=c("celltype", "Stim"), measure.vars=c("FS", "PFS"))

a<-ggplot(melt, aes(x=celltype, y=value))+
    geom_boxplot(aes(fill=Stim),size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_x_discrete(labels = c("CD4", "CD8", "CD4-CD8-"))+
    scale_fill_manual(values =c("PMA" = "#000094"))+
    facet_grid(~variable)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(comparisons=list(c("CD4", "CD8"), c("CD4", "GD")),
                       label="p.signif", method="wilcox.test", size=8, hide.ns=TRUE, step.increase = .2)+
    labs(subtitle="",y="Score",
         x="")+
    coord_cartesian(ylim=c(0,1.3))

supp3<-cowplot::plot_grid(NULL, a, ncol=1, rel_heights = c(6,2.5),
                          labels=c("A", "B"), label_size = 26)
```

## Split by celltype, compare Stims within TB  
don't need to run stats because ggplot handles... lots of differences but doesn't tell us much more than the previous figure. Might be good first supplement.  
```{r supp fig 3, fig.width=10, fig.height=10}
##CD4
cd4fs<-ggplot(compass[compass$celltype=="CD4",], aes(x=Stim, y=FS))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_grid(~TB)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(label="p.signif", method="wilcox.test", size=8, label.x=1.4, label.y=.35, hide.ns = TRUE)+
    labs(subtitle="Functionality Scores", y="CD4 Functionality Score",
         x="")+
    coord_cartesian(ylim=c(0,.4))

cd4pfs<-ggplot(compass[compass$celltype=="CD4",], aes(x=Stim, y=PFS))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_grid(~TB)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(label="p.signif", method="wilcox.test", size=8, label.x=1.4, label.y=.16, hide.ns = TRUE)+
    labs(subtitle="Polyfunctionality Scores", y="CD4 Polyfunctionality Score",
         x="")+
    coord_cartesian(ylim=c(0,.18))

cd4<-cowplot::plot_grid(cd4fs, cd4pfs, nrow=1)

##CD8
cd8fs<-ggplot(compass[compass$celltype=="CD8",], aes(x=Stim, y=FS))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_grid(~TB)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(label="p.signif", method="wilcox.test", size=8, label.x=1.4, label.y=.35, hide.ns = TRUE)+
    labs(y="CD8 Functionality Score",
         x="")+
        coord_cartesian(ylim=c(0,.4))

cd8pfs<-ggplot(compass[compass$celltype=="CD8",], aes(x=Stim, y=PFS))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_grid(~TB)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(label="p.signif", method="wilcox.test", size=8, label.x=1.4, label.y=.16, hide.ns = TRUE)+
    labs(y="CD8 Polyfunctionality Score",
         x="")+
    coord_cartesian(ylim=c(0,.18))

cd8<-cowplot::plot_grid(cd8fs, cd8pfs, nrow=1)

##GD
gdfs<-ggplot(compass[compass$celltype=="GD",], aes(x=Stim, y=FS))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_grid(~TB)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(label="p.signif", method="wilcox.test", size=8, label.x=1.4, label.y=.35, hide.ns = TRUE)+
    labs(y="CD4-CD8- Functionality Score",
         x="")+
    coord_cartesian(ylim=c(0,.4))

gdpfs<-ggplot(compass[compass$celltype=="GD",], aes(x=Stim, y=PFS))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_grid(~TB)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(label="p.signif", method="wilcox.test", size=8, label.x=1.4, label.y=.16, hide.ns = TRUE)+
    labs(y="CD4-CD8- Polyfunctionality Score",
         x="")+
    coord_cartesian(ylim=c(0,.18))

gds<-cowplot::plot_grid(gdfs, gdpfs, nrow=1)

supp4<-cowplot::plot_grid(cd4, cd8, gds, ncol=1, labels=c("A", "B", "C"), label_size=26)
```

## Split by Stim, compare celltypes within TB
Within the SWAP stim in the LTBI and TB groups, celltype responses are different. Might be figure 2.  
### SEA  
```{r}
sea<-compass[compass$Stim=="SEA",]
sea$celltype<-factor(sea$celltype)
x<-split(sea, sea$TB)
lapply(x, function(g) posthoc.kruskal.nemenyi.test(FS~celltype, g)$p.value)

sea<-compass[compass$Stim=="SEA",]
sea$celltype<-factor(sea$celltype)
x<-split(sea, sea$TB)
lapply(x, function(g) posthoc.kruskal.nemenyi.test(PFS~celltype, g)$p.value)
```
### SWAP  
```{r}
#FS
swap<-compass[compass$Stim=="SWAP",]
swap$celltype<-factor(swap$celltype)
x<-split(swap, swap$TB)
lapply(x, function(g) posthoc.kruskal.nemenyi.test(FS~celltype, g)$p.value)

#PFS
swap<-compass[compass$Stim=="SWAP",]
swap$celltype<-factor(swap$celltype)
x<-split(swap, swap$TB)
lapply(x, function(g) posthoc.kruskal.nemenyi.test(PFS~celltype, g)$p.value)
```

```{r fig 2}
fs<-ggplot(compass, aes(x=celltype, y=FS))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=26)) + 
    facet_grid(Stim~TB)+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    scale_x_discrete(labels = c("CD4", "CD8", "CD4-CD8-"))+
    stat_compare_means(comparisons=list(c("GD", "CD8"), c("GD", "CD4")), 
            label="p.signif" ,label_y=0.4, hide.ns = TRUE, size=8, step.increase=0.15)+
    labs(y="Functionality Score",
         x="")+
    coord_cartesian(ylim=c(0,.56))

supp6<-ggplot(compass, aes(x=celltype, y=PFS))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=26)) + 
    facet_grid(Stim~TB)+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    scale_x_discrete(labels = c("CD4", "CD8", "CD4-CD8-"))+
    stat_compare_means(comparisons=list(c("GD", "CD8"), c("GD", "CD4")), 
            label="p.signif" ,label_y=0.2, hide.ns = TRUE, size=8, step.increase=0.15)+
    labs(y="Polyfunctionality Score",
         x="")+
    coord_cartesian(ylim=c(0,0.26))

fig2<-cowplot::plot_grid(NULL, fs, ncol=1, labels=c("A", "B"), label_size=26, rel_heights = c(3,4))
two.hor<-cowplot::plot_grid(NULL, fs, ncol=2, labels=c("A", "B"), label_size=26, rel_widths = c(5,4))
two.v3<-cowplot::plot_grid(NULL, fs, ncol=2, labels=c("A", "B"), label_size=26, rel_widths = c(4,6))
```

## Split by Stim compare TB within celltypes
Only SEA difference was CD8 comparing N to LTBI and no SWAP differences so supplementary if anything.
```{r}
sea<-compass[compass$Stim=="SEA",]
sea$TB<-factor(sea$TB)
x<-split(sea, sea$celltype)
lapply(x, function(g) posthoc.kruskal.nemenyi.test(FS~TB, g)$p.value)
lapply(x, function(g) posthoc.kruskal.nemenyi.test(PFS~TB, g)$p.value)

swap<-compass[compass$Stim=="SWAP",]
swap$celltype<-factor(swap$celltype)
x<-split(swap, swap$celltype)
lapply(x, function(g) posthoc.kruskal.nemenyi.test(FS~TB, g)$p.value)
lapply(x, function(g) posthoc.kruskal.nemenyi.test(PFS~TB, g)$p.value)
```

```{r supp fig 4, fig.width=10, fig.height=12}
melt<-melt(compass[compass$Stim=="SEA",], id.vars=c("group","celltype", "TB","Stim"), measure.vars=c("FS", "PFS"))

a<-ggplot(melt[melt$celltype=="CD4",], aes(x=TB, y=value))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5), plot.title = element_text(hjust = 0.5)) +
    facet_grid(~variable)+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(title="SEA", subtitle="CD4",
         y="Score",
         x="")+
    coord_cartesian(ylim=c(0,.35))

b<-ggplot(melt[melt$celltype=="CD8",], aes(x=TB, y=value))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    facet_grid(~variable)+
    stat_compare_means(comparisons=list(c("N", "HC")), 
            label="p.signif", size=8, hid.ns=TRUE, step.increase=0.15, label.y = .3)+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(subtitle="CD8",
         y="Score",
         x="")+
    coord_cartesian(ylim=c(0,.35))

c<-ggplot(melt[melt$celltype=="GD",], aes(x=TB, y=value))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    facet_grid(~variable)+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(subtitle="CD4-CD8-",
         y="Score",
         x="")+
    coord_cartesian(ylim=c(0,.4))

supp5sea<-cowplot::plot_grid(a, b, c, ncol=1, labels=c("A", "B", "C"), label_size=26)
```

```{r supp fig 4, fig.width=10, fig.height=12}
melt<-melt(compass[compass$Stim=="SWAP",], id.vars=c("group","celltype", "TB","Stim"), measure.vars=c("FS", "PFS"))

a<-ggplot(melt[melt$celltype=="CD4",], aes(x=TB, y=value))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5), plot.title = element_text(hjust = 0.5)) +
    facet_grid(~variable)+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(title="SWAP",subtitle="CD4",
         y="Score",
         x="")+
    coord_cartesian(ylim=c(0,.35))

b<-ggplot(melt[melt$celltype=="CD8",], aes(x=TB, y=value))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons=list(c("N", "TB")), 
            label="p.signif", size=8, hid.ns=TRUE, step.increase=0.15, label.y = .3)+
    facet_grid(~variable)+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(subtitle="CD8",
         y="Score",
         x="")+
    coord_cartesian(ylim=c(0,.35))

c<-ggplot(melt[melt$celltype=="GD",], aes(x=TB, y=value))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20),
              axis.text.x = element_text(size = 16),
              strip.text = element_text(size = 20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    facet_grid(~variable)+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(subtitle="CD4-CD8-",
         y="Score",
         x="")+
    coord_cartesian(ylim=c(0,.4))

supp5swap<-cowplot::plot_grid(a, b, c, ncol=1, labels=c("D", "E", "F"), label_size=26)
supp5<-cowplot::plot_grid(supp5sea, supp5swap, nrow=1)
```

# OG

## Proliferation 
```{r fig 3}
include<-subset(CD3, CD3$Stim%in%c("SEA","SWAP"))
b<-ggplot(include, aes(x=Stim, y=OG))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_grid(~TB)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=26)) + 
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(label="p.signif", method="wilcox.test", size=8, label.x=1.4, label.y=45)+
    labs(y="Proliferating\nCD3+ T cells (%)",
         x="")+
    coord_cartesian(ylim=c(0,50))
fig3<-cowplot::plot_grid(NULL, b, ncol=1, rel_heights = c(3,4), labels = c("A", "B"), label_size=26)

```

```{r}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
datatable<-read.csv("OG_CD3_bulk_summary_clean.csv")
    datatable<-filter(datatable, Stim %in% c("SEA", "SWAP", "SEB")) #No SM data in primary files
    datatable<-filter(datatable, SM!="X") #No SM data in primary files
    datatable<-datatable[!(datatable$Donor %in% remove),]
    datatable$SM<-factor(datatable$SM, levels=c("N", "SM"))
    datatable$TB<-factor(datatable$TB, levels=c("N", "HC", "LTBI", "TB"))
datatable$Stim<-factor(datatable$Stim, levels=c("SEB", "SEA", "SWAP"))

x<-split(datatable, datatable$Stim)
lapply(x, function(g) posthoc.kruskal.nemenyi.test(OG~TB, g)$p.value)

seaplot<-ggplot(datatable[datatable$Stim=="SEA",], aes(x=TB, y=OG))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_wrap(~Stim, scales="free_y")+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=26)) + 
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(y="Proliferating CD3+ T cells (%)",
         x="")+
    coord_cartesian(ylim=c(0,1))

swapplot<-ggplot(datatable[datatable$Stim=="SWAP",], aes(x=TB, y=OG))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_wrap(~Stim, scales="free_y")+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=26)) + 
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(y="Proliferating CD3+ T cells (%)",
         x="")+
    coord_cartesian(ylim=c(0,40))

sebplot<-ggplot(datatable[datatable$Stim=="SEB",], aes(x=TB, y=OG))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_wrap(~Stim, scales="free_y")+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=26)) + 
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(y="Proliferating CD3+ T cells (%)",
         x="")+
    coord_cartesian(ylim=c(0,80))

supp7<-cowplot::plot_grid(sebplot, seaplot, swapplot, nrow=1, labels=c("A", "B", "C"), label_size = 26)
```

## Proportion of Proliferating Cells
P-values
```{r}
colnames<-c("CD4", "CD8", "GD")
include<-subset(CD3_OG, CD3_OG$Stim%in%c("SEA","SWAP"))
include$id<-paste(include$Donor, include$Stim, sep="_")
include$celltype<-"CD3"
test<-merge(include, mimosa, by=c("id", "celltype"))
test2<-filter(test, test$include==1)

kable(table(test2$Stim, test2$TB))

melt<-melt(test2, id.vars=c("TB", "SM", "Stim"), measure.vars=colnames)
melt$variable<-gsub("GD", "γδ", melt$variable)
melt$celltype<-gsub("", "",melt$variable)
melt$TB<-factor(melt$TB, levels=c("N", "HC", "LTBI", "TB"))

#SEA---> just don't include
#sea<-melt[melt$Stim=="SEA",]
#sea$celltype<-factor(sea$celltype)
#x<-split(sea, sea$TB)
#lapply(x, function(g) posthoc.kruskal.nemenyi.test(value~celltype, g)$p.value)

#SWAP
swap<-melt[melt$Stim=="SWAP",]
swap$celltype<-factor(swap$celltype)
x<-split(swap, swap$TB)
lapply(x, function(g) posthoc.kruskal.nemenyi.test(value~celltype, g)$p.value)
```

## Compare cell types
```{r, fig 4, fig.width=8, fig.height=6}
b1<-ggplot(swap, aes(x=celltype, y=value))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=26)) + 
    facet_grid(Stim~TB)+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    scale_x_discrete(labels = c("CD4", "CD8", "γδ"))+
    stat_compare_means(comparisons=list(c("CD8","γδ"), c("CD4", "γδ")),  
                       hide.ns=TRUE, label="p.signif", size=8, step.increase = .11)+
    labs(y="Subset of Proliferating T cells (%)",
         x="")+
    coord_cartesian(ylim=c(0,130))
b<-ggplot(swap, aes(x=celltype, y=value))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=26)) + 
    facet_grid(~TB)+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    scale_x_discrete(labels = c("CD4", "CD8", "γδ"))+
    labs(y="Subset of Proliferating T cells (%)",
         x="")
fig4<-cowplot::plot_grid(NULL, b1, nrow=1, rel_widths = c(4,6), labels = c("A", "B"), label_size=26)
four<-cowplot::plot_grid(NULL, b, nrow=1, rel_heights = c(4,6), labels = c("A", "B"), label_size=26)
```

## GD specific stuff  
```{r fig 5}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
datatable<-read.csv("OG_CD3_bulk_summary_clean.csv")
    datatable<-filter(datatable, Stim %in% c("UN", "SWAP")) #No SM data in primary files
    datatable<-filter(datatable, SM!="X") #No SM data in primary files
    datatable<-datatable[!(datatable$Donor %in% remove),]
    datatable$SM<-factor(datatable$SM, levels=c("N", "SM"))
    datatable$TB<-factor(datatable$TB, levels=c("N", "HC", "LTBI", "TB"))

unplot<-ggplot(datatable[datatable$Stim=="UN",], aes(x=TB, y=GD))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    theme_classic()+ theme(legend.position="none")+
    facet_grid(~Stim)+
    theme(text = element_text(size=26)) + 
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(y="γδ T cells (% CD3 T cells)",
         x="")

setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
datatable<-read.csv("OG_GD_bulk_summary_clean.csv")
    datatable<-filter(datatable, Stim %in% c("SEA", "SWAP", "SEB")) #No SM data in primary files
    datatable<-filter(datatable, SM!="X") #No SM data in primary files
    datatable<-datatable[!(datatable$Donor %in% remove),]
    datatable$SM<-factor(datatable$SM, levels=c("N", "SM"))
    datatable$TB<-factor(datatable$TB, levels=c("N", "HC", "LTBI", "TB"))
datatable$Stim<-factor(datatable$Stim, levels=c("SEB", "SEA", "SWAP"))

x<-split(datatable, datatable$Stim)
lapply(x, function(g) posthoc.kruskal.nemenyi.test(OG~TB, g)$p.value)

seaplot<-ggplot(datatable[datatable$Stim=="SEA",], aes(x=TB, y=OG))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_wrap(~Stim, scales="free_y")+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=26)) + 
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(y="Proliferating γδ T cells (%)",
         x="")+
    coord_cartesian(ylim=c(0,5))

swapplot<-ggplot(datatable[datatable$Stim=="SWAP",], aes(x=TB, y=OG))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_wrap(~Stim, scales="free_y")+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=26)) + 
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(comparisons=list(c("N", "LTBI"), c("N","TB")), label="p.signif", size=8, step.increase = 0.07)+
    labs(y="Proliferating γδ T cells (%)",
         x="")+
    coord_cartesian(ylim=c(0,106))

sebplot<-ggplot(datatable[datatable$Stim=="SEB",], aes(x=TB, y=OG))+
    geom_boxplot(aes(fill=TB), size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_wrap(~Stim, scales="free_y")+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=26)) + 
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(y="Proliferating γδ T cells (%)",
         x="")+
    coord_cartesian(ylim=c(0,106))

fig5<-cowplot::plot_grid(unplot ,sebplot, seaplot, swapplot, nrow=1, 
                         labels=c("A", "B", "C", "D"), label_size = 26)
```

```{r fig 6}
include<-subset(gd_OG, gd_OG$Stim=="SWAP")
posthoc.kruskal.nemenyi.test(TNFa~TB, include)$p.value
posthoc.kruskal.nemenyi.test(IFNg~TB, include)$p.value
posthoc.kruskal.nemenyi.test(IL4~TB, include)$p.value

melt<-melt(include, id.vars=c("TB"), measure.vars=c("TNFa", "IFNg", "IL4"))
cyt.groups<-as_labeller(c("IL4" = "IL-4", "IFNg" = "IFNγ", "TNFa" = "TNFα"))
cytokine<-ggplot(melt, aes(x=TB, y=value, fill=TB))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_grid(~variable, scales="free", labeller = cyt.groups)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=26)) + 
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    stat_compare_means(comparisons=list(c("N","TB")), hide.ns=TRUE,
        label="p.signif", size=8, step.increase = .11)+
    labs(y="Cytokine+ Proliferating γδ T cells (%)",
         x="")+
    coord_cartesian(ylim=c(0,110))

fig6<-cowplot::plot_grid(NULL, cytokine, labels=c("A", "B"), nrow=2, label_size=26)
```

```{r}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
datatable<-read.csv("OG_gd_OG_summary_clean_NAs.csv")
datatable<-filter(datatable, Stim %in% c("SEA", "SWAP", "SEB")) #No SM data in primary files
datatable<-filter(datatable, SM!="X") #No SM data in primary files
datatable<-datatable[!(datatable$Donor %in% remove),]
datatable$SM<-factor(datatable$SM, levels=c("N", "SM"))
datatable$TB<-factor(datatable$TB, levels=c("N", "HC", "LTBI", "TB"))
datatable$Stim<-factor(datatable$Stim, levels=c("SEB", "SEA", "SWAP"))

include<-subset(datatable, datatable$Stim=="SEB")
posthoc.kruskal.nemenyi.test(TNFa~TB, include)$p.value
posthoc.kruskal.nemenyi.test(IFNg~TB, include)$p.value
posthoc.kruskal.nemenyi.test(IL4~TB, include)$p.value

melt<-melt(include, id.vars=c("TB"), measure.vars=c("TNFa", "IFNg", "IL4"))
cyt.groups<-as_labeller(c("IL4" = "IL-4", "IFNg" = "IFNγ", "TNFa" = "TNFα"))
ggplot(melt, aes(x=TB, y=value, fill=TB))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA)+
    scale_fill_manual(values = c("N" ="#e0e0e0","HC"= "#1a9850" ,"LTBI"= "#2166ac","TB"= "#b2182b"))+
    facet_grid(~variable, scales="free", labeller = cyt.groups)+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=26)) + 
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(y="(%) Proliferating γδ T cells",
         x="")
```

```{r}
prolif<-dplyr::select(gd, Donor, Stim, OG)
prolif$celltype<-"GD"
test<-merge(prolif, compass, by=c("Donor", "Stim", "celltype"))
cyt<-dplyr::select(neg, Donor, Stim, celltype, Cyt.Freq)
cyt$celltype<-gsub("CD4-CD8-", "GD", cyt$celltype)
test2<-merge(test, cyt, by=c("Donor", "Stim", "celltype"))

ggplot(test2, aes(x=OG, y=FS))+
  geom_point()+
  facet_grid(~Stim)+
  geom_smooth(method="lm")

ggplot(test2, aes(x=OG, y=Cyt.Freq))+
  geom_point()+
  facet_grid(~Stim)+
  geom_smooth(method="lm")

cor.test(~OG+Cyt.Freq, data=test2, method="spearman")
cor.test(~OG+Cyt.Freq, data=test2, method="kendall")
cor.test(~OG+FS, data=test2, method="spearman")
cor.test(~OG+FS, data=test2, method="kendall")
```

```{r}
setwd("/Users/tarynam/Desktop/manuscript-sm/")
save_plot("fig1.png", fig1, base_height = 18, base_width = 15)
save_plot("fig2.png", fig2, base_height = 16, base_width = 16)
save_plot("fig3.png", fig3, base_height = 8, base_width = 10)
save_plot("fig4.png", fig4, base_height = 7, base_width = 18)
save_plot("fig5.png", fig5, base_height=8, base_width = 24)
save_plot("fig6.png", fig6, base_height=15, base_width = 12)
save_plot("supp2.png", supp2, base_height = 6, base_width = 10)
save_plot("supp3.png", supp3, base_height = 18, base_width = 12)
save_plot("supp4.png", supp4, base_height = 16, base_width = 15)
save_plot("supp5.png", supp5, base_height = 16, base_width = 20)
save_plot("supp6.png", supp6, base_height = 9, base_width = 16)
save_plot("supp7.png", supp7, base_height = 8, base_width = 16)
```

### 2020-05-01 Computational Packages/COMPASS/Compass_analysis.Rmd
first commit

--- 2018-10-21 Compass_analysis.Rmd
@@
 library(COMPASS)
-DF<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/CD4_count.csv")
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
+DF<-read.csv("ICS_CD4_Counts_clean.csv")
 compass_data<-DF
 compass_data<-rename(compass_data, 
-    "IFNg&IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count",
-    "IFNg&IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.1",
-    "IFNg&IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.2",
-    "IFNg&IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.3",
-    "IFNg&!IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.4",
-    "IFNg&!IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.5",
-    "IFNg&!IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.6",
-    "IFNg&!IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.7",
-    "!IFNg&IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.8",
-    "!IFNg&IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.9",
-    "!IFNg&IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.10",
-    "!IFNg&IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.11",
-    "!IFNg&!IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.12",
-    "!IFNg&!IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.13",
-    "!IFNg&!IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.14",
-    "!IFNg&!IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.15")
+    "IFNg&IL4&IL13&TNFa" = "G_4_13_T",
+    "IFNg&IL4&IL13&!TNFa" = "G_4_13_x",
+    "IFNg&IL4&!IL13&TNFa" = "G_4_x_T",
+    "IFNg&IL4&!IL13&!TNFa" =  "G_4_x_x" ,
+    "IFNg&!IL4&IL13&TNFa" = "G_x_13_T" ,
+    "IFNg&!IL4&IL13&!TNFa" = "G_x_13_x" ,
+    "IFNg&!IL4&!IL13&TNFa" = "G_x_x_T",
+    "IFNg&!IL4&!IL13&!TNFa" = "G_x_x_x",
+    "!IFNg&IL4&IL13&TNFa" = "x_4_13_T",
+    "!IFNg&IL4&IL13&!TNFa" = "x_4_13_x",
+    "!IFNg&IL4&!IL13&TNFa" = "x_4_x_T",
+    "!IFNg&IL4&!IL13&!TNFa" = "x_4_x_x",
+    "!IFNg&!IL4&IL13&TNFa" = "x_x_13_T",
+    "!IFNg&!IL4&IL13&!TNFa" = "x_x_13_x",
+    "!IFNg&!IL4&!IL13&TNFa" = "x_x_x_T",
+    "!IFNg&!IL4&!IL13&!TNFa" = "x_x_x_x")
 
@@
 compass_data$Stim<-factor(compass_data$Stim, levels=c("UN", "PMA", "WCL", "PEP", "SEA", "SWAP"))
+compass_data$ID<-compass_data$Donor
 metadata<-dplyr::select(compass_data, c("ID","TB","SM","Stim"))

### 2020-05-01 Computational Packages/MIMOSA/MIMOSA-priors.Rmd
first commit

---
title: "MIMOSA Priors"
author: "Taryn McLaughlin"
date: "10/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, message = FALSE)
library(MIMOSA)
```

**Background:** When we spoke in April, I asked if MIMOSA could be used in flow experiments with dye dilution assays since the expected positive counts are much higher than in an intracellular cytokine stain. You said that it should be possible by changing the priors, but would need to be optimized. The priors are specified by the EXPRATE argument in the .fitMCMC function which is called by the MIMOSA function. I'm having trouble getting my edits to EXPRATE to stick.

This is just the example in the ReadMe on your Github to show that the package is in fact working correctly. Though I don't know why I get a warning error since it's literally copied from the help file.

```{r, cache = TRUE}
library(MIMOSA)
data(ICS)
E<-ConstructMIMOSAExpressionSet(ICS,
                    reference=ANTIGEN%in%'negctrl',measure.columns=c('CYTNUM','NSUB'),
                    other.annotations=c('CYTOKINE','TCELLSUBSET','ANTIGEN','UID'),
                    default.cast.formula=component~UID+ANTIGEN+CYTOKINE+TCELLSUBSET,
                    .variables=.(TCELLSUBSET,CYTOKINE,UID),
                    featureCols=1,ref.append.replace='_REF')

result<-MIMOSA(NSUB+CYTNUM~UID+TCELLSUBSET+CYTOKINE|ANTIGEN,
               data=E, method='mcmc',
               subset=RefTreat%in%'Treatment'&ANTIGEN%in%'ENV',
               ref=ANTIGEN%in%'ENV'&RefTreat%in%'Reference')

head(getZ(result))
```

I pulled this documentation from .fitMCMC by calling fixInNamespace which pulls up the function in a new window for you to approve. I didn't actually click any buttons, I just copied the function.

```{r, eval = FALSE}
fixInNamespace(".fitMCMC", .fitMCMC, ns="MIMOSA")
function (data, inits = NULL, iter = 250000, burn = 50000, thin = 1, 
    tune = 100, outfile = basename(tempfile(tmpdir = ".", fileext = ".dat")), 
    alternative = "greater", UPPER = 0.5, LOWER = 0.15, FAST = TRUE, 
    EXPRATE = 1e-04, pXi = c(1, 1), seed = 10) 
{
    set.seed(seed)
    alternative <- match.arg(alternative, c("greater", "not equal"))
    data <- icsdata2mvicsdata(data)
    if (is.null(inits)) {
        r <- MDMix(data)
        inits <- list(alpha.s = r@par.stim, alpha.u = r@par.unstim, 
            q = r@w[1], z = round(r@z))
    }
    if (alternative == "greater") {
        ps <- t(do.call(cbind, apply(data$n.stim, 1, function(x) (data.frame(prop.table(x))[-1L, 
            , drop = FALSE]))))
        pu <- t(do.call(cbind, apply(data$n.unstim, 1, function(x) (data.frame(prop.table(x))[-1L, 
            , drop = FALSE]))))
        filter <- sapply(1:nrow(ps), function(i) all(ps[i, ] <= 
            pu[i, ]))
        FILTER = TRUE
    }
    else {
        filter <- rep(FALSE, nrow(data$n.stim))
        FILTER <- FALSE
    }
    result <- .Call("C_fitMCMC", as.matrix(data$n.stim), as.matrix(data$n.unstim), 
        as.vector(inits$alpha.s), as.vector(inits$alpha.u), as.vector(inits$q), 
        as.matrix(inits$z), as.vector(iter), as.vector(burn), 
        as.vector(thin), as.numeric(tune), as.character(outfile), 
        as.vector(filter), as.numeric(UPPER), as.numeric(LOWER), 
        FILTER, FAST, as.numeric(EXPRATE), as.numeric(pXi))
    if (inherits(result, "character")) {
        return(result)
    }
    result$z <- cbind(result$z, 1 - result$z)
    result$getmcmc <- function(x = outfile) {
        mcmc(read.table(x, sep = "\t", header = TRUE))
    }
    result$getP <- function(x = paste(outfile, "P", sep = ""), 
        thin = 1) {
        if (thin > 1) {
            nc <- length(strsplit(readLines(x, 1), "\t")[[1]])
            thins <- paste("p", paste(rep(";n", thin - 1), collapse = ""), 
                sep = "")
            s <- sprintf("sed -n '%s' %s|cut -f %s-%s", thins, 
                x, (nc/3 + 1), nc)
            con <- pipe(s)
            d <- do.call(rbind, lapply(strsplit(readLines(con), 
                "\t")[-1L], as.numeric))
            colnames(d) <- strsplit(readLines(x, 1), "\t")[[1]][(nc/3 + 
                1):nc]
            d <- split(as.list(data.frame(d)), gl(nc/3, 2))
            close(con)
        }
        else {
            d <- mcmc(read.table(x, sep = "\t", header = TRUE))
            nc <- ncol(d)
            d <- split(as.list(data.frame(d[, (nc/3 + 1):nc])), 
                gl(nc/3, 2))
        }
        d
    }
    attr(result, "class") <- c(attr(result, "class"), "MDMixResult")
    attr(result, "pData") <- attr(data, "pData")
    result$n.stim <- data$n.stim
    result$n.unstim <- data$n.unstim
    result
}
```

So now here is my new fitMCMC function

```{r}
library(MIMOSA)
my.fitMCMC<-function (data, inits = NULL, iter = 250000, burn = 50000, thin = 1, 
          tune = 100, outfile = basename(tempfile(tmpdir = ".", fileext = ".dat")), 
          alternative = "greater", UPPER = 0.5, LOWER = 0.15, FAST = TRUE, 
          EXPRATE = .01, pXi = c(1, 1), seed = 10)
{
    set.seed(seed)
    alternative <- match.arg(alternative, c("greater", "not equal"))
    data <- MIMOSA:::icsdata2mvicsdata(data)
    if (is.null(inits)) {
        r <- MDMix(data)
        inits <- list(alpha.s = r@par.stim, alpha.u = r@par.unstim, 
                      q = r@w[1], z = round(r@z))
    }
    if (alternative == "greater") {
        ps <- t(do.call(cbind, apply(data$n.stim, 1, function(x) (data.frame(prop.table(x))[-1L, 
                                                                                            , drop = FALSE]))))
        pu <- t(do.call(cbind, apply(data$n.unstim, 1, function(x) (data.frame(prop.table(x))[-1L, 
                                                                                              , drop = FALSE]))))
        filter <- sapply(1:nrow(ps), function(i) all(ps[i, ] <= 
                                                         pu[i, ]))
        FILTER = TRUE
    }
    else {
        filter <- rep(FALSE, nrow(data$n.stim))
        FILTER <- FALSE
    }
    result <- .Call("C_fitMCMC", as.matrix(data$n.stim), as.matrix(data$n.unstim), 
                    as.vector(inits$alpha.s), as.vector(inits$alpha.u), as.vector(inits$q), 
                    as.matrix(inits$z), as.vector(iter), as.vector(burn), 
                    as.vector(thin), as.numeric(tune), as.character(outfile), 
                    as.vector(filter), as.numeric(UPPER), as.numeric(LOWER), 
                    FILTER, FAST, as.numeric(EXPRATE), as.numeric(pXi))
    if (inherits(result, "character")) {
        return(result)
    }
    result$z <- cbind(result$z, 1 - result$z)
    result$getmcmc <- function(x = outfile) {
        coda:::mcmc(read.table(x, sep = "\t", header = TRUE))
    }
    result$getP <- function(x = paste(outfile, "P", sep = ""), 
                            thin = 1) {
        if (thin > 1) {
            nc <- length(strsplit(readLines(x, 1), "\t")[[1]])
            thins <- paste("p", paste(rep(";n", thin - 1), collapse = ""), 
                           sep = "")
            s <- sprintf("sed -n '%s' %s|cut -f %s-%s", thins, 
                         x, (nc/3 + 1), nc)
            con <- pipe(s)
            d <- do.call(rbind, lapply(strsplit(readLines(con), 
                                                "\t")[-1L], as.numeric))
            colnames(d) <- strsplit(readLines(x, 1), "\t")[[1]][(nc/3 + 
                                                                     1):nc]
            d <- split(as.list(data.frame(d)), gl(nc/3, 2))
            close(con)
        }
        else {
            d <- coda:::mcmc(read.table(x, sep = "\t", header = TRUE))
            nc <- ncol(d)
            d <- split(as.list(data.frame(d[, (nc/3 + 1):nc])), 
                       gl(nc/3, 2))
        }
        d
    }
    attr(result, "class") <- c(attr(result, "class"), "MDMixResult")
    attr(result, "pData") <- attr(data, "pData")
    result$n.stim <- data$n.stim
    result$n.unstim <- data$n.unstim
    result
}
```

I then assign it to the MIMOSA namespace and call the MIMOSA function and it runs just fine.

```{r, cache = TRUE}
library(MIMOSA)
assignInNamespace(".fitMCMC", my.fitMCMC, ns="MIMOSA")
result_2<-MIMOSA(NSUB+CYTNUM~UID+TCELLSUBSET+CYTOKINE|ANTIGEN,
               data=E, method='mcmc',
               subset=RefTreat%in%'Treatment'&ANTIGEN%in%'ENV',
               ref=ANTIGEN%in%'ENV'&RefTreat%in%'Reference')
```

Now it runs just fine but I get nearly the same results with both the original .fitMCMC and my.fitMCMC which I feel like shouldn't happen when the priors are different.

```{r}
library(knitr)
z1<-cbind(countsTable(result), fdr(result), getZ(result))
z2<-cbind(countsTable(result_2), fdr(result_2), getZ(result_2))

kable(head(z1, 10), caption = "Output for original fitMCMC")
kable(head(z2, 10), caption = "Output for new fitMCMC")

kable(colMeans(z1), caption = "Means for original fitMCMC")
kable(colMeans(z2), caption = "Means for new fitMCMC")

#testing for equality between the probability of fdr
all.equal(z1[,5], z2[,5])
#testing for equality between the probability of no response columns
all.equal(z1[,6], z2[,6])
#testing for equality between the probability of response columns
all.equal(z1[,7], z2[,7])
```

### 2020-05-01 Computational Packages/MIMOSA/MIMOSA_priors_compare.Rmd
first commit

---
title: "MIMOSA_OG-test"
author: "Taryn McLaughlin"
date: "5/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, error=FALSE)
library(dplyr)
library(ggplot2)
```

```{r}
counts<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_counts_summary_clean.csv")

pkgs <- c("MIMOSA","Biobase","ggplot2","MASS", "plyr","reshape")
    # see what packages are currently installed
installed_pacakges <- row.names(installed.packages()) # loop over the needed packages
for(p in pkgs){
        # check if package is installed
        already_installed <- p %in% installed_pacakges
        # if not already installed, install it
        if(!already_installed){ install.packages(p)
        }
        # and load package
        library(p, character.only = TRUE)}

source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")    
```

```{r standard}
## CD4
data<-dplyr::select(counts, Donor, TB, SM, Stim, CD4_count, CD4_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('CD4_neg','CD4_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(CD4_neg+CD4_count~Donor|Stim,
               data=E,
               method="mcmc")

CD4.probabilities<-MIMOSA::getZ(test)
CD4.counts<-MIMOSA::countsTable(test)
CD4.fdr<-MIMOSA::fdr(test)
CD4.Results<-data.frame(cbind(CD4.counts,CD4.probabilities,CD4.fdr))
CD4.Results$celltype<-"CD4"
CD4.summary<-MIMOSA::getW(test)

## CD8
data<-dplyr::select(counts, Donor, TB, SM, Stim, CD8_count, CD8_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('CD8_neg','CD8_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(CD8_neg+CD8_count~Donor|Stim,
               data=E,
               method="mcmc")

CD8.probabilities<-MIMOSA::getZ(test)
CD8.counts<-MIMOSA::countsTable(test)
CD8.fdr<-MIMOSA::fdr(test)
CD8.Results<-data.frame(cbind(CD8.counts,CD8.probabilities,CD8.fdr))
CD8.Results$celltype<-"CD8"
CD8.summary<-MIMOSA::getW(test)

## GD
data<-dplyr::select(counts, Donor, TB, SM, Stim, GD_count, GD_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('GD_neg','GD_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(GD_neg+GD_count~Donor|Stim,
               data=E,
               method="mcmc")

GD.probabilities<-MIMOSA::getZ(test)
GD.counts<-MIMOSA::countsTable(test)
GD.fdr<-MIMOSA::fdr(test)
GD.Results<-data.frame(cbind(GD.counts,GD.probabilities,GD.fdr))
GD.Results$celltype<-"GD"
GD.summary<-MIMOSA::getW(test)

## CD3
data<-dplyr::select(counts, Donor, TB, SM, Stim, CD3_count, CD3_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('CD3_neg','CD3_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(CD3_neg+CD3_count~Donor|Stim,
               data=E,
               method="mcmc")

CD3.probabilities<-MIMOSA::getZ(test)
CD3.counts<-MIMOSA::countsTable(test)
CD3.fdr<-MIMOSA::fdr(test)
CD3.Results<-data.frame(cbind(CD3.counts,CD3.probabilities,CD3.fdr))
CD3.Results$celltype<-"CD3"
CD3.summary<-MIMOSA::getW(test)

CD3.Results$Sample<-rownames(CD3.Results)
CD4.Results$Sample<-rownames(CD4.Results)
CD8.Results$Sample<-rownames(CD8.Results)
GD.Results$Sample<-rownames(GD.Results)

l<-list(CD3.Results, CD4.Results, CD8.Results, GD.Results)
l<-lapply(l, function(x) splitter(x, "Sample", "_", one="Donor", two="Stim"))
MIMOSA.Results<-rbindlist(l)
MIMOSA.Results$Exprate<-0.0001
```

For each of these test runs, I have changed the prior expected rate. The prior expected rate reflects the 
```{r exprate 0.01}
my.fitMCMC<-function (data, inits = NULL, iter = 250000, burn = 50000, thin = 1, 
          tune = 100, outfile = basename(tempfile(tmpdir = ".", fileext = ".dat")), 
          alternative = "greater", UPPER = 0.5, LOWER = 0.15, FAST = TRUE, 
          EXPRATE = .01, pXi = c(1, 1), seed = 10)
{
    set.seed(seed)
    alternative <- match.arg(alternative, c("greater", "not equal"))
    data <- MIMOSA:::icsdata2mvicsdata(data)
    if (is.null(inits)) {
        r <- MDMix(data)
        inits <- list(alpha.s = r@par.stim, alpha.u = r@par.unstim, 
                      q = r@w[1], z = round(r@z))
    }
    if (alternative == "greater") {
        ps <- t(do.call(cbind, apply(data$n.stim, 1, function(x) (data.frame(prop.table(x))[-1L, 
                                                                                            , drop = FALSE]))))
        pu <- t(do.call(cbind, apply(data$n.unstim, 1, function(x) (data.frame(prop.table(x))[-1L, 
                                                                                              , drop = FALSE]))))
        filter <- sapply(1:nrow(ps), function(i) all(ps[i, ] <= 
                                                         pu[i, ]))
        FILTER = TRUE
    }
    else {
        filter <- rep(FALSE, nrow(data$n.stim))
        FILTER <- FALSE
    }
    result <- .Call("C_fitMCMC", as.matrix(data$n.stim), as.matrix(data$n.unstim), 
                    as.vector(inits$alpha.s), as.vector(inits$alpha.u), as.vector(inits$q), 
                    as.matrix(inits$z), as.vector(iter), as.vector(burn), 
                    as.vector(thin), as.numeric(tune), as.character(outfile), 
                    as.vector(filter), as.numeric(UPPER), as.numeric(LOWER), 
                    FILTER, FAST, as.numeric(EXPRATE), as.numeric(pXi))
    if (inherits(result, "character")) {
        return(result)
    }
    result$z <- cbind(result$z, 1 - result$z)
    result$getmcmc <- function(x = outfile) {
        coda:::mcmc(read.table(x, sep = "\t", header = TRUE))
    }
    result$getP <- function(x = paste(outfile, "P", sep = ""), 
                            thin = 1) {
        if (thin > 1) {
            nc <- length(strsplit(readLines(x, 1), "\t")[[1]])
            thins <- paste("p", paste(rep(";n", thin - 1), collapse = ""), 
                           sep = "")
            s <- sprintf("sed -n '%s' %s|cut -f %s-%s", thins, 
                         x, (nc/3 + 1), nc)
            con <- pipe(s)
            d <- do.call(rbind, lapply(strsplit(readLines(con), 
                                                "\t")[-1L], as.numeric))
            colnames(d) <- strsplit(readLines(x, 1), "\t")[[1]][(nc/3 + 
                                                                     1):nc]
            d <- split(as.list(data.frame(d)), gl(nc/3, 2))
            close(con)
        }
        else {
            d <- coda:::mcmc(read.table(x, sep = "\t", header = TRUE))
            nc <- ncol(d)
            d <- split(as.list(data.frame(d[, (nc/3 + 1):nc])), 
                       gl(nc/3, 2))
        }
        d
    }
    attr(result, "class") <- c(attr(result, "class"), "MDMixResult")
    attr(result, "pData") <- attr(data, "pData")
    result$n.stim <- data$n.stim
    result$n.unstim <- data$n.unstim
    result
}

assignInNamespace(".fitMCMC", my.fitMCMC, ns="MIMOSA")
```

```{r run 0.01, cache=TRUE}
## CD4
data<-dplyr::select(counts, Donor, TB, SM, Stim, CD4_count, CD4_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('CD4_neg','CD4_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(CD4_neg+CD4_count~Donor|Stim,
               data=E,
               method="mcmc")

CD4.probabilities<-MIMOSA::getZ(test)
CD4.counts<-MIMOSA::countsTable(test)
CD4.fdr<-MIMOSA::fdr(test)
CD4.Results<-data.frame(cbind(CD4.counts,CD4.probabilities,CD4.fdr))
CD4.Results$celltype<-"CD4"
CD4.summary<-MIMOSA::getW(test)

## CD8
data<-dplyr::select(counts, Donor, TB, SM, Stim, CD8_count, CD8_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('CD8_neg','CD8_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(CD8_neg+CD8_count~Donor|Stim,
               data=E,
               method="mcmc")

CD8.probabilities<-MIMOSA::getZ(test)
CD8.counts<-MIMOSA::countsTable(test)
CD8.fdr<-MIMOSA::fdr(test)
CD8.Results<-data.frame(cbind(CD8.counts,CD8.probabilities,CD8.fdr))
CD8.Results$celltype<-"CD8"
CD8.summary<-MIMOSA::getW(test)

## GD
data<-dplyr::select(counts, Donor, TB, SM, Stim, GD_count, GD_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('GD_neg','GD_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(GD_neg+GD_count~Donor|Stim,
               data=E,
               method="mcmc")

GD.probabilities<-MIMOSA::getZ(test)
GD.counts<-MIMOSA::countsTable(test)
GD.fdr<-MIMOSA::fdr(test)
GD.Results<-data.frame(cbind(GD.counts,GD.probabilities,GD.fdr))
GD.Results$celltype<-"GD"
GD.summary<-MIMOSA::getW(test)

## CD3
data<-dplyr::select(counts, Donor, TB, SM, Stim, CD3_count, CD3_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('CD3_neg','CD3_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(CD3_neg+CD3_count~Donor|Stim,
               data=E,
               method="mcmc")

CD3.probabilities<-MIMOSA::getZ(test)
CD3.counts<-MIMOSA::countsTable(test)
CD3.fdr<-MIMOSA::fdr(test)
CD3.Results<-data.frame(cbind(CD3.counts,CD3.probabilities,CD3.fdr))
CD3.Results$celltype<-"CD3"
CD3.summary<-MIMOSA::getW(test)

CD3.Results$Sample<-rownames(CD3.Results)
CD4.Results$Sample<-rownames(CD4.Results)
CD8.Results$Sample<-rownames(CD8.Results)
GD.Results$Sample<-rownames(GD.Results)

l<-list(CD3.Results, CD4.Results, CD8.Results, GD.Results)
l<-lapply(l, function(x) splitter(x, "Sample", "_", one="Donor", two="Stim"))
MIMOSA.Results_01<-rbindlist(l)
MIMOSA.Results_01$Exprate<-0.01
```

```{r exprate 0.05}
my.fitMCMC<-function (data, inits = NULL, iter = 250000, burn = 50000, thin = 1, 
          tune = 100, outfile = basename(tempfile(tmpdir = ".", fileext = ".dat")), 
          alternative = "greater", UPPER = 0.5, LOWER = 0.15, FAST = TRUE, 
          EXPRATE = .05, pXi = c(1, 1), seed = 10)
{
    set.seed(seed)
    alternative <- match.arg(alternative, c("greater", "not equal"))
    data <- MIMOSA:::icsdata2mvicsdata(data)
    if (is.null(inits)) {
        r <- MDMix(data)
        inits <- list(alpha.s = r@par.stim, alpha.u = r@par.unstim, 
                      q = r@w[1], z = round(r@z))
    }
    if (alternative == "greater") {
        ps <- t(do.call(cbind, apply(data$n.stim, 1, function(x) (data.frame(prop.table(x))[-1L, 
                                                                                            , drop = FALSE]))))
        pu <- t(do.call(cbind, apply(data$n.unstim, 1, function(x) (data.frame(prop.table(x))[-1L, 
                                                                                              , drop = FALSE]))))
        filter <- sapply(1:nrow(ps), function(i) all(ps[i, ] <= 
                                                         pu[i, ]))
        FILTER = TRUE
    }
    else {
        filter <- rep(FALSE, nrow(data$n.stim))
        FILTER <- FALSE
    }
    result <- .Call("C_fitMCMC", as.matrix(data$n.stim), as.matrix(data$n.unstim), 
                    as.vector(inits$alpha.s), as.vector(inits$alpha.u), as.vector(inits$q), 
                    as.matrix(inits$z), as.vector(iter), as.vector(burn), 
                    as.vector(thin), as.numeric(tune), as.character(outfile), 
                    as.vector(filter), as.numeric(UPPER), as.numeric(LOWER), 
                    FILTER, FAST, as.numeric(EXPRATE), as.numeric(pXi))
    if (inherits(result, "character")) {
        return(result)
    }
    result$z <- cbind(result$z, 1 - result$z)
    result$getmcmc <- function(x = outfile) {
        coda:::mcmc(read.table(x, sep = "\t", header = TRUE))
    }
    result$getP <- function(x = paste(outfile, "P", sep = ""), 
                            thin = 1) {
        if (thin > 1) {
            nc <- length(strsplit(readLines(x, 1), "\t")[[1]])
            thins <- paste("p", paste(rep(";n", thin - 1), collapse = ""), 
                           sep = "")
            s <- sprintf("sed -n '%s' %s|cut -f %s-%s", thins, 
                         x, (nc/3 + 1), nc)
            con <- pipe(s)
            d <- do.call(rbind, lapply(strsplit(readLines(con), 
                                                "\t")[-1L], as.numeric))
            colnames(d) <- strsplit(readLines(x, 1), "\t")[[1]][(nc/3 + 
                                                                     1):nc]
            d <- split(as.list(data.frame(d)), gl(nc/3, 2))
            close(con)
        }
        else {
            d <- coda:::mcmc(read.table(x, sep = "\t", header = TRUE))
            nc <- ncol(d)
            d <- split(as.list(data.frame(d[, (nc/3 + 1):nc])), 
                       gl(nc/3, 2))
        }
        d
    }
    attr(result, "class") <- c(attr(result, "class"), "MDMixResult")
    attr(result, "pData") <- attr(data, "pData")
    result$n.stim <- data$n.stim
    result$n.unstim <- data$n.unstim
    result
}

assignInNamespace(".fitMCMC", my.fitMCMC, ns="MIMOSA")
```

```{r run 0.05, cache=TRUE}
## CD4
data<-dplyr::select(counts, Donor, TB, SM, Stim, CD4_count, CD4_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('CD4_neg','CD4_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(CD4_neg+CD4_count~Donor|Stim,
               data=E,
               method="mcmc")

CD4.probabilities<-MIMOSA::getZ(test)
CD4.counts<-MIMOSA::countsTable(test)
CD4.fdr<-MIMOSA::fdr(test)
CD4.Results<-data.frame(cbind(CD4.counts,CD4.probabilities,CD4.fdr))
CD4.Results$celltype<-"CD4"
CD4.summary<-MIMOSA::getW(test)

## CD8
data<-dplyr::select(counts, Donor, TB, SM, Stim, CD8_count, CD8_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('CD8_neg','CD8_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(CD8_neg+CD8_count~Donor|Stim,
               data=E,
               method="mcmc")

CD8.probabilities<-MIMOSA::getZ(test)
CD8.counts<-MIMOSA::countsTable(test)
CD8.fdr<-MIMOSA::fdr(test)
CD8.Results<-data.frame(cbind(CD8.counts,CD8.probabilities,CD8.fdr))
CD8.Results$celltype<-"CD8"
CD8.summary<-MIMOSA::getW(test)

## GD
data<-dplyr::select(counts, Donor, TB, SM, Stim, GD_count, GD_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('GD_neg','GD_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(GD_neg+GD_count~Donor|Stim,
               data=E,
               method="mcmc")

GD.probabilities<-MIMOSA::getZ(test)
GD.counts<-MIMOSA::countsTable(test)
GD.fdr<-MIMOSA::fdr(test)
GD.Results<-data.frame(cbind(GD.counts,GD.probabilities,GD.fdr))
GD.Results$celltype<-"GD"
GD.summary<-MIMOSA::getW(test)

## CD3
data<-dplyr::select(counts, Donor, TB, SM, Stim, CD3_count, CD3_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('CD3_neg','CD3_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(CD3_neg+CD3_count~Donor|Stim,
               data=E,
               method="mcmc")

CD3.probabilities<-MIMOSA::getZ(test)
CD3.counts<-MIMOSA::countsTable(test)
CD3.fdr<-MIMOSA::fdr(test)
CD3.Results<-data.frame(cbind(CD3.counts,CD3.probabilities,CD3.fdr))
CD3.Results$celltype<-"CD3"
CD3.summary<-MIMOSA::getW(test)

CD3.Results$Sample<-rownames(CD3.Results)
CD4.Results$Sample<-rownames(CD4.Results)
CD8.Results$Sample<-rownames(CD8.Results)
GD.Results$Sample<-rownames(GD.Results)

l<-list(CD3.Results, CD4.Results, CD8.Results, GD.Results)
l<-lapply(l, function(x) splitter(x, "Sample", "_", one="Donor", two="Stim"))
MIMOSA.Results_05<-rbindlist(l)
MIMOSA.Results_05$Exprate<-0.05
```

```{r exprate 0.1}
my.fitMCMC<-function (data, inits = NULL, iter = 250000, burn = 50000, thin = 1, 
          tune = 100, outfile = basename(tempfile(tmpdir = ".", fileext = ".dat")), 
          alternative = "greater", UPPER = 0.5, LOWER = 0.15, FAST = TRUE, 
          EXPRATE = .1, pXi = c(1, 1), seed = 10)
{
    set.seed(seed)
    alternative <- match.arg(alternative, c("greater", "not equal"))
    data <- MIMOSA:::icsdata2mvicsdata(data)
    if (is.null(inits)) {
        r <- MDMix(data)
        inits <- list(alpha.s = r@par.stim, alpha.u = r@par.unstim, 
                      q = r@w[1], z = round(r@z))
    }
    if (alternative == "greater") {
        ps <- t(do.call(cbind, apply(data$n.stim, 1, function(x) (data.frame(prop.table(x))[-1L, 
                                                                                            , drop = FALSE]))))
        pu <- t(do.call(cbind, apply(data$n.unstim, 1, function(x) (data.frame(prop.table(x))[-1L, 
                                                                                              , drop = FALSE]))))
        filter <- sapply(1:nrow(ps), function(i) all(ps[i, ] <= 
                                                         pu[i, ]))
        FILTER = TRUE
    }
    else {
        filter <- rep(FALSE, nrow(data$n.stim))
        FILTER <- FALSE
    }
    result <- .Call("C_fitMCMC", as.matrix(data$n.stim), as.matrix(data$n.unstim), 
                    as.vector(inits$alpha.s), as.vector(inits$alpha.u), as.vector(inits$q), 
                    as.matrix(inits$z), as.vector(iter), as.vector(burn), 
                    as.vector(thin), as.numeric(tune), as.character(outfile), 
                    as.vector(filter), as.numeric(UPPER), as.numeric(LOWER), 
                    FILTER, FAST, as.numeric(EXPRATE), as.numeric(pXi))
    if (inherits(result, "character")) {
        return(result)
    }
    result$z <- cbind(result$z, 1 - result$z)
    result$getmcmc <- function(x = outfile) {
        coda:::mcmc(read.table(x, sep = "\t", header = TRUE))
    }
    result$getP <- function(x = paste(outfile, "P", sep = ""), 
                            thin = 1) {
        if (thin > 1) {
            nc <- length(strsplit(readLines(x, 1), "\t")[[1]])
            thins <- paste("p", paste(rep(";n", thin - 1), collapse = ""), 
                           sep = "")
            s <- sprintf("sed -n '%s' %s|cut -f %s-%s", thins, 
                         x, (nc/3 + 1), nc)
            con <- pipe(s)
            d <- do.call(rbind, lapply(strsplit(readLines(con), 
                                                "\t")[-1L], as.numeric))
            colnames(d) <- strsplit(readLines(x, 1), "\t")[[1]][(nc/3 + 
                                                                     1):nc]
            d <- split(as.list(data.frame(d)), gl(nc/3, 2))
            close(con)
        }
        else {
            d <- coda:::mcmc(read.table(x, sep = "\t", header = TRUE))
            nc <- ncol(d)
            d <- split(as.list(data.frame(d[, (nc/3 + 1):nc])), 
                       gl(nc/3, 2))
        }
        d
    }
    attr(result, "class") <- c(attr(result, "class"), "MDMixResult")
    attr(result, "pData") <- attr(data, "pData")
    result$n.stim <- data$n.stim
    result$n.unstim <- data$n.unstim
    result
}

assignInNamespace(".fitMCMC", my.fitMCMC, ns="MIMOSA")
```

```{r run 0.1, cache=TRUE}
## CD4
data<-dplyr::select(counts, Donor, TB, SM, Stim, CD4_count, CD4_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('CD4_neg','CD4_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(CD4_neg+CD4_count~Donor|Stim,
               data=E,
               method="mcmc")

CD4.probabilities<-MIMOSA::getZ(test)
CD4.counts<-MIMOSA::countsTable(test)
CD4.fdr<-MIMOSA::fdr(test)
CD4.Results<-data.frame(cbind(CD4.counts,CD4.probabilities,CD4.fdr))
CD4.Results$celltype<-"CD4"
CD4.summary<-MIMOSA::getW(test)

## CD8
data<-dplyr::select(counts, Donor, TB, SM, Stim, CD8_count, CD8_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('CD8_neg','CD8_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(CD8_neg+CD8_count~Donor|Stim,
               data=E,
               method="mcmc")

CD8.probabilities<-MIMOSA::getZ(test)
CD8.counts<-MIMOSA::countsTable(test)
CD8.fdr<-MIMOSA::fdr(test)
CD8.Results<-data.frame(cbind(CD8.counts,CD8.probabilities,CD8.fdr))
CD8.Results$celltype<-"CD8"
CD8.summary<-MIMOSA::getW(test)

## GD
data<-dplyr::select(counts, Donor, TB, SM, Stim, GD_count, GD_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('GD_neg','GD_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(GD_neg+GD_count~Donor|Stim,
               data=E,
               method="mcmc")

GD.probabilities<-MIMOSA::getZ(test)
GD.counts<-MIMOSA::countsTable(test)
GD.fdr<-MIMOSA::fdr(test)
GD.Results<-data.frame(cbind(GD.counts,GD.probabilities,GD.fdr))
GD.Results$celltype<-"GD"
GD.summary<-MIMOSA::getW(test)

## CD3
data<-dplyr::select(counts, Donor, TB, SM, Stim, CD3_count, CD3_neg)
E<-ConstructMIMOSAExpressionSet(data,
                                reference=Stim %in% 'UN',
                                measure.columns=c('CD3_neg','CD3_count'),
                                other.annotations=c('Stim', 'Donor', 'TB', 'SM'),
                                default.cast.formula=component~Donor+Stim,
                                .variables=.(Donor))

test <- MIMOSA(CD3_neg+CD3_count~Donor|Stim,
               data=E,
               method="mcmc")

CD3.probabilities<-MIMOSA::getZ(test)
CD3.counts<-MIMOSA::countsTable(test)
CD3.fdr<-MIMOSA::fdr(test)
CD3.Results<-data.frame(cbind(CD3.counts,CD3.probabilities,CD3.fdr))
CD3.Results$celltype<-"CD3"
CD3.summary<-MIMOSA::getW(test)

CD3.Results$Sample<-rownames(CD3.Results)
CD4.Results$Sample<-rownames(CD4.Results)
CD8.Results$Sample<-rownames(CD8.Results)
GD.Results$Sample<-rownames(GD.Results)

l<-list(CD3.Results, CD4.Results, CD8.Results, GD.Results)
l<-lapply(l, function(x) splitter(x, "Sample", "_", one="Donor", two="Stim"))
MIMOSA.Results_10<-rbindlist(l)
MIMOSA.Results_10$Exprate<-0.1
```

```{r}
Results<-list(MIMOSA.Results, MIMOSA.Results_01, MIMOSA.Results_05, MIMOSA.Results_10)
Final<-rbindlist(Results)
write.csv(Final, "/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_MIMOSA_clean.csv")
```

```{r fig.height=25, fig.width=25}
Final<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_MIMOSA_clean.csv")
PBMC<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_DonorInfo_clean.csv")
PBMC<-dplyr::filter(PBMC, SM %in% c("X", "SM+"))
PBMC$SM<-factor(PBMC$SM, levels=c("X", "SM"))
PBMC$TB<-factor(PBMC$TB, levels=c("HC", "LTBI", "TB"))

Final<-merge(PBMC, Final, by="Donor")
ggplot(Final, aes(x=Exprate, y=Pr.response, col=TB))+
    geom_line(aes(group=Donor))+
    geom_point(shape=21, size=1, stroke = 1)+
    geom_hline(yintercept = 0.7, size=1, linetype=2)+
    theme(text = element_text(size=40))+
    facet_grid(celltype~Stim)+
    theme_classic()+theme(panel.border = element_rect(colour = "black", fill=NA))

```

```{r fig.height=25, fig.width=25}
ggplot(Final, aes(x=Exprate, y=fdr, col=TB))+
    geom_line(aes(group=Donor))+
    geom_point(shape=21, size=1, stroke = 1)+
    geom_hline(yintercept = 0.03, size=1, linetype=2)+
    theme(text = element_text(size=40))+
    facet_grid(celltype~Stim)+
    theme_classic()+theme(panel.border = element_rect(colour = "black", fill=NA))

```

### 2020-05-01 Computational Packages/COMPASS/COMPASS_ICS.R

@@ -1,0 +1,510 @@
+#source("http://www.bioconductor.org/biocLite.R")
+#biocLite("COMPASS")
+library(COMPASS)
+library(dplyr)
+library(data.table)
+CD4_compass<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/ICS_CD4_Counts_clean.csv")
+CD4_compass$celltype<-"CD4"
+CD8_compass<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/ICS_CD8_Counts_clean.csv")
+CD8_compass$celltype<-"CD8"
+GD_compass<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/ICS_GD_Counts_clean.csv")
+GD_compass$celltype<-"GD"
+
+
+test<-merge(CD4_compass, CD8_compass, 
+            by=c("Sample", "Donor", "TB", "SM", "Stim"), all=TRUE)
+
+test2<-merge(test, GD_compass, 
+             by=c("Sample", "Donor", "TB", "SM", "Stim"), all=TRUE)
+
+CD3<-dplyr::select(test2, Sample, Donor, TB, SM, Stim)
+
+CD3$G_4_13_T<-test2$G_4_13_T + test2$G_4_13_T.x + test2$G_4_13_T.y
+CD3$G_4_13_x<-test2$G_4_13_x + test2$G_4_13_x.x + test2$G_4_13_x.y
+CD3$G_4_x_T<-test2$G_4_x_T   + test2$G_4_x_T.x  + test2$G_4_x_T.y
+CD3$G_4_x_x<-test2$G_4_x_x   + test2$G_4_x_x.x  + test2$G_4_x_x.y
+
+CD3$G_x_13_T<-test2$G_x_13_T + test2$G_x_13_T.x + test2$G_x_13_T.y
+CD3$G_x_13_x<-test2$G_x_13_x + test2$G_x_13_x.x + test2$G_x_13_x.y
+CD3$G_x_x_T<-test2$G_x_x_T   + test2$G_x_x_T.x  + test2$G_x_x_T.y
+CD3$G_x_x_x<-test2$G_x_x_x   + test2$G_x_x_x.x  + test2$G_x_x_x.y
+
+CD3$x_4_13_T<-test2$x_4_13_T + test2$x_4_13_T.x + test2$x_4_13_T.y
[…]

### 2020-05-01 Computational Packages/COMPASS/COMPASS_tutorial.R

@@ -1,0 +1,104 @@
+library(dplyr)
+library(data.table)
+
+setwd("/Users/tarynam/Desktop/Lisa/")
+compass<-read.csv("MIMI_compass.csv")
+tiger<-dplyr::select(compass, Sample, unique, ID, Stim) #this should just me the meta data
+kitten<- compass[,2:17] #This should just be the count columns
+
+### TWO WAYS TO RELABEL CYTOKINES 
+
+# IF YOU HAVE USE THE .P .N NOMENCLATURE
+# do this for each cytokine make sure the capitalization matches whatever you have
+names(kitten) <- gsub("IL2.p", "&IL2", names(kitten))
+names(kitten) <- gsub("IL2.n", "&!IL2", names(kitten))
+# Whatever the first cytokine is gets just a "!" instead of a "&!"
+names(kitten) <- gsub("&!IFNG", "!IFNG", names(kitten))
+
+# IF TAKING THE NAMES STRAIGHT FROM FLOWJO AND NOT MANUALLY EDITING THE FILE
+# These are the original column names in order. 
+# The nice thing is your can manually reorder the cytokines for the future graphs
+# Ex: I flipped the TNF and IL17 so that the cytokine ordering looks better in the plots
+#i+2+17+t+  --->  IFNγ&IL2&TNFα&IL17
+#i+2+17+t-  --->  IFNγ&IL2&!TNFα&IL17
+#i+2+17-t+  --->  IFNγ&IL2&TNFα&!IL17
+#i+2+17-t-  --->  IFNγ&IL2&!TNFα&!IL17
+#i+2-17+t+  --->  IFNγ&!IL2&TNFα&IL17
+#i+2-17+t-  --->  IFNγ&!IL2&!TNFα&IL17
+#i+2-17-t+  --->  IFNγ&!IL2&TNFα&!IL17
+#i+2-17-t-  --->  IFNγ&!IL2&!TNFα&!IL17
+#i-2+17+t+  --->  !IFNγ&IL2&TNFα&IL17
+#i-2+17+t-  --->  !IFNγ&IL2&!TNFα&IL17
+#i-2+17-t+  --->  !IFNγ&IL2&TNFα&!IL17
+#i-2+17-t-  --->  !IFNγ&IL2&!TNFα&!IL17
[…]

### 2020-05-01 Computational Packages/MIMOSA/MIMOSA_ICS.R

@@ -1,0 +1,300 @@
+install.packages("devtools") #install devtools package
+library(devtools) #load it
+install_github("MIMOSA","RGLab",branch="master") 
+#Need R 3.0.0, and a bunch of dependencies.
+#The install will fail with various error messages until you install those dependencies.
+library(MIMOSA)
+library(dplyr)
+library(data.table)
+
+setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
+DF<-read.csv("ICS_CD4_Counts_Clean.csv")
+DF<-mutate(DF, Total =(x_x_x_x + Cyt))
+
+#All cytokine
+data<-dplyr::select(DF, Sample, Donor, TB, SM, Stim, Cyt, Total) %>%
+    dplyr::mutate(Neg = (Total - Cyt)) %>%
+    dplyr::select(-Total)
+
+E<-ConstructMIMOSAExpressionSet(data,
+    reference=Stim %in% 'UN',
+    measure.columns=c('Neg','Cyt'),
+    other.annotations=c('Stim','Sample', 'Donor', 'TB', 'SM'),
+    default.cast.formula=component~Donor+Stim,
+    .variables=.(Donor))
+test <- MIMOSA(Neg+Cyt~Donor|Stim,
+    data=E,
+    method="mcmc")
+
+probabilities<-MIMOSA::getZ(test)
+counts<-MIMOSA::countsTable(test)
+fdr<-MIMOSA::fdr(test)
+MIMOSA.Results<-data.frame(cbind(counts,probabilities,fdr))
+MIMOSA.summary<-MIMOSA::getW(test)
+
+#TH1 
+data<-dplyr::select(DF, Sample, Donor, TB, SM, Stim, TH1, Total) %>%
+    dplyr::mutate(Neg = (Total - TH1)) %>%
+    dplyr::select(-Total)
+
+E1<-ConstructMIMOSAExpressionSet(data,
+                                reference=Stim %in% 'UN',
+                                measure.columns=c('Neg','TH1'),
+                                other.annotations=c('Stim','Sample', 'Donor', 'TB', 'SM'),
+                                default.cast.formula=component~Donor+Stim,
+                                .variables=.(Donor))
+
+test1 <- MIMOSA(Neg+TH1~Donor|Stim,
[…]

### 2020-05-01 Computational Packages/MIMOSA/MIMOSA_Jeremiah.R

@@ -1,0 +1,107 @@
+library(MIMOSA)
+library(dplyr)
+library(data.table)
+
+###Here is the updated script
+setwd("/Users/tarynam/Desktop/Jeremiah/")
+#I manually edited these column names in the csv file because they were horrible to format
+DF<-read.csv("JKK Flow Data Absolute Counts 2019_AUG_19.csv")
+names(DF)<-gsub("CD4_CD4.", "CD4_", names(DF))
+names(DF)<-gsub("CD8_CD8.", "CD8_", names(DF))
+a<-tstrsplit(DF$Sample, " ")[[5]]
+b<-tstrsplit(a, "_")[[1]]
+b<-gsub(",2f,", ".", b)
+DF$condition<-b
+
+
+#CD8 IFNg
+bubblegum<-dplyr::select(DF, PATIENT.ID, condition, CD8_IFNgP, CD8_Total) %>%
+    dplyr::mutate(Neg = (CD8_Total - CD8_IFNgP)) %>%
+    dplyr::select(-CD8_Total)
+
+doozie<-ConstructMIMOSAExpressionSet(bubblegum,
+                                     reference=condition %in% 'uns',
+                                     measure.columns=c('Neg','CD8_IFNgP'),
+                                     other.annotations=c('condition','PATIENT.ID'),
+                                     default.cast.formula=component~PATIENT.ID+condition,
+                                     .variables=.(PATIENT.ID))
+
+banana <- MIMOSA(Neg+CD8_IFNgP~PATIENT.ID|condition,
+                 data=doozie,
+                 method="mcmc")
+
+probabilities<-MIMOSA::getZ(banana)
+counts<-MIMOSA::countsTable(banana)
+fdr<-MIMOSA::fdr(banana)
+MIMOSA.Results_ifng<-data.frame(cbind(counts,probabilities,fdr))
+MIMOSA.summary_ifng<-MIMOSA::getW(banana)
+MIMOSA.Results_ifng$measure<-"ifng"
+MIMOSA.Results_ifng$Sample<-rownames(MIMOSA.Results_ifng)
+names(MIMOSA.Results_ifng)<-gsub("CD8_IFNgP","Pos", names(MIMOSA.Results_ifng))
+MIMOSA.Results_ifng$Sample<-gsub(".ESAT6", "", MIMOSA.Results_ifng$Sample)
+#comapre to cheryl
[…]

### 2020-05-01 Computational Packages/MIMOSA/MIMOSA_Morgan.R

@@ -1,0 +1,102 @@
+biocLite("devtools")
+library(devtools) #load it
+install_github("MIMOSA","RGLab",branch="master") #Need R 3.0.0, and a bunch of dependencies. The install will fail with various error messages until you install those dependencies.
+biocLite("MIMOSA")
+library(MIMOSA)
+
+DF<-read.csv("/Users/tarynam/Desktop/QFT-_HIV+_donor_file/CD8_TB_HIVexport_180824.csv")
+DF<-as.data.table(DF) 
+
+
+#GT
+DF$Neg<-DF$COUNT-DF$GT_Count
+DF<-DF[DF$CONDITION!="SEB"]
+data<-dplyr::select(DF, DONOR, CONDITION, DISEASE, Neg, GT_Count)
+E<-ConstructMIMOSAExpressionSet(data,
+                                reference=CONDITION%in%'unstim',
+                                measure.columns=c('Neg','GT_Count'),
+                                other.annotations=c('CONDITION','DONOR'),
+                                default.cast.formula=component~DONOR+CONDITION,
+                                .variables=.(DONOR))
+test <- MIMOSA(Neg+GT_Count~DONOR|CONDITION,
+               data=E, method="mcmc")
+probabilities<-MIMOSA::getZ(test)
+counts<-MIMOSA::countsTable(test)
+fdr<-MIMOSA::fdr(test)
+MIMOSA.Results1<-cbind(counts,probabilities,fdr)
+MIMOSA.summary<-MIMOSA::getW(test)
+
+#TNF-any
+DF$Neg<-DF$COUNT-DF$TNFa_Count
+DF<-DF[DF$CONDITION!="SEB"]
+data<-dplyr::select(DF, DONOR, CONDITION, DISEASE, Neg, TNFa_Count)
+E<-ConstructMIMOSAExpressionSet(data,
+                                reference=CONDITION%in%'unstim',
+                                measure.columns=c('Neg','TNFa_Count'),
+                                other.annotations=c('CONDITION','DONOR'),
+                                default.cast.formula=component~DONOR+CONDITION,
+                                .variables=.(DONOR))
+test <- MIMOSA(Neg+TNFa_Count~DONOR|CONDITION,
+               data=E, method="mcmc")
+probabilities<-MIMOSA::getZ(test)
+counts<-MIMOSA::countsTable(test)
+fdr<-MIMOSA::fdr(test)
[…]

### 2020-05-01 Computational Packages/MIMOSA/MIMOSA_OG.R

@@ -1,0 +1,203 @@
+counts<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_counts_summary_clean.csv")
+
+pkgs <- c("MIMOSA","Biobase","ggplot2","MASS", "plyr","reshape")
+    # see what packages are currently installed
+installed_pacakges <- row.names(installed.packages()) # loop over the needed packages
+for(p in pkgs){
+        # check if package is installed
+        already_installed <- p %in% installed_pacakges
+        # if not already installed, install it
+        if(!already_installed){ install.packages(p)
+        }
+        # and load package
+        library(p, character.only = TRUE)}    
+    
+
+#When I first tried to sub in a new value for EXPRATE, I couldn't find the actual function
+#That .fitMCMC executes so I left the brackets empty
+my.fitMCMC<-function (data, inits = NULL, iter = 250000, burn = 50000, thin = 1, 
+          tune = 100, outfile = basename(tempfile(tmpdir = ".", fileext = ".dat")), 
+          alternative = "greater", UPPER = 0.5, LOWER = 0.15, FAST = TRUE, 
+          EXPRATE = .01, pXi = c(1, 1), seed = 10)
+{
+    set.seed(seed)
+    alternative <- match.arg(alternative, c("greater", "not equal"))
+    data <- MIMOSA:::icsdata2mvicsdata(data)
+    if (is.null(inits)) {
+        r <- MDMix(data)
+        inits <- list(alpha.s = r@par.stim, alpha.u = r@par.unstim, 
+                      q = r@w[1], z = round(r@z))
+    }
+    if (alternative == "greater") {
+        ps <- t(do.call(cbind, apply(data$n.stim, 1, function(x) (data.frame(prop.table(x))[-1L, 
+                                                                                            , drop = FALSE]))))
+        pu <- t(do.call(cbind, apply(data$n.unstim, 1, function(x) (data.frame(prop.table(x))[-1L, 
+                                                                                              , drop = FALSE]))))
+        filter <- sapply(1:nrow(ps), function(i) all(ps[i, ] <= 
+                                                         pu[i, ]))
+        FILTER = TRUE
[…]

### 2020-05-01 Computational Packages/MIMOSA/MIMOSA_TBRU.R

@@ -1,0 +1,40 @@
+install.packages("remotes") #install devtools package
+library(remotes) #load it
+install_github("RGLab/MIMOSA",ref="trunk") 
+#Need R 3.0.0, and a bunch of dependencies.
+#The install will fail with various error messages until you install those dependencies.
+library(MIMOSA)
+library(dplyr)
+library(data.table)
+
+###Here is the updated script
+setwd("/Users/tarynam/Desktop/MIMOSA_docs_for_troubleshooting")
+DF<-read.csv("TBRU Project 1_Counts_FOR COMPASS (1).csv")
+#CD4.p is the total CD4 count
+#CD4.p_IFNg.p is the IFNg postive count
+#Sample_ID is the ID v
+
+#IFNg
+bubblegum<-dplyr::select(DF, Sample_ID, TimePoint, Stim, CD4.p, CD4.p_IFNg.p) %>%
+    dplyr::mutate(Neg = (CD4.p - CD4.p_IFNg.p)) %>%
+    dplyr::select(-CD4.p)
+
+E<-ConstructMIMOSAExpressionSet(bubblegum,
+                                reference=Stim %in% 'US',
+                                measure.columns=c('Neg','CD4.p_IFNg.p'),
+                                other.annotations=c('Stim','Sample_ID', 'TimePoint'),
+                                default.cast.formula=component~Sample_ID+Stim+TimePoint,
+                                .variables=.(Sample_ID, TimePoint))
+
+results <- MIMOSA(Neg+CD4.p_IFNg.p~Sample_ID+TimePoint|Stim,
+               data=E,
+               method="mcmc",
+               ref= RefTreat%in%'Reference')
+
+probabilities<-MIMOSA::getZ(results)
+counts<-MIMOSA::countsTable(results)
+fdr<-MIMOSA::fdr(results)
+MIMOSA.Results_ifng<-data.frame(cbind(counts,probabilities,fdr))
+MIMOSA.summary_ifng<-MIMOSA::getW(results)
+
+

### 2020-05-01 Computational Packages/MIMOSA/MIMOSA_demo.R

@@ -1,0 +1,112 @@
+install.packages("remotes") #install devtools package
+library(remotes) #load it
+install_github("RGLab/MIMOSA",ref="trunk") 
+#Need R 3.0.0, and a bunch of dependencies.
+#The install will fail with various error messages until you install those dependencies.
+library(MIMOSA)
+library(dplyr)
+library(data.table)
+
+###Here is the updated script
+setwd("/Users/tarynam/Desktop/")
+DF<-read.csv("JKK Flow Data Absolute Counts 2019_AUG_19.csv")
+#show data table --> column names
+DF$cd40 <- DF$CD4.40.p_107.p_G.p + DF$CD4.40.p_107.p_G.n + 
+    DF$CD4.40.p_107.n_G.p + DF$CD4.40.p_107.n_G.n
+DF$total<- DF$CD4.40.p_107.p_G.p + DF$CD4.40.p_107.p_G.n + 
+    DF$CD4.40.p_107.n_G.p + DF$CD4.40.p_107.n_G.n + 
+    DF$CD4.40.n_107.p_G.p + DF$CD4.40.n_107.p_G.n + 
+    DF$CD4.40.n_107.n_G.p + DF$CD4.40.n_107.n_G.n
+##easier for me this way
+CD4<-mutate(DF,
+            #These all follow the same format
+            #IFNg is the new variable. it gets assign the row sum of all the columns that have a G in the name
+            cd40=  rowSums(DF[,grep("CD4.40.p_107.._G..", names(DF))]),
+            cd107=  rowSums(DF[,grep("CD4.40.._107.p_G..", names(DF))]),
+            ifng=  rowSums(DF[,grep("CD4.40.._107.._G.p", names(DF))]),
+            total=   rowSums(DF[,grep("CD4.40.._107.._G..", names(DF))])
+)
+CD4$condition <- tstrsplit(CD4$TUBE.NAME, " ")[[3]]
+CD4$condition<-gsub("CFP10_ESAT6", "CFP10/ESAT6", CD4$condition)
+names(CD4)
+
+#IFNg
+bubblegum<-dplyr::select(CD4, PATIENT.ID, condition, ifng, total) %>%
+    dplyr::mutate(Neg = (total - ifng)) %>%
+    dplyr::select(-total)
+
+doozie<-ConstructMIMOSAExpressionSet(bubblegum,
+                                reference=condition %in% 'uns',
+                                measure.columns=c('Neg','ifng'),
+                                other.annotations=c('condition','PATIENT.ID'),
[…]

### 2020-05-01 Computational Packages/MIMOSA/change-mimosa_priors.R

@@ -1,0 +1,24 @@
+data(ICS)
+E<-ConstructMIMOSAExpressionSet(ICS,
+                                reference=ANTIGEN%in%'negctrl',measure.columns=c('CYTNUM','NSUB'),
+                                other.annotations=c('CYTOKINE','TCELLSUBSET','ANTIGEN','UID'),
+                                default.cast.formula=component~UID+ANTIGEN+CYTOKINE+TCELLSUBSET,
+                                .variables=.(TCELLSUBSET,CYTOKINE,UID),
+                                featureCols=1,
+                                ref.append.replace='_REF')
+
+my.fitMCMC<-function(data, inits = NULL, iter = 250000, burn = 50000, thin = 1,
+         tune = 100, outfile = basename(tempfile(tmpdir = ".", fileext = ".dat")),
+         alternative = "greater", UPPER = 0.5, LOWER = 0.15, FAST = TRUE,
+         EXPRATE = .1, pXi = c(1, 1), seed = 10){}
+
+fixInNamespace(".fitMCMC", my.fitMCMC, ns="MIMOSA")
+
+result2<-MIMOSA(NSUB+CYTNUM~UID+TCELLSUBSET+CYTOKINE|ANTIGEN,
+               data=E, 
+               method='EM',
+               subset=RefTreat%in%'Treatment'&ANTIGEN%in%'ENV',
+               ref=ANTIGEN%in%'ENV'&RefTreat%in%'Reference', 
+               RT=FALSE
+)
+
